### 2019-02-28 paper.tex
Granger Causality paper

%% Use \input{...} to include a standard preamble

\title{VAR Models and Granger Causality}
\author{R. J. Kinnear, R. R. Mazumdar}

\begin{document}
\maketitle
\abstract{}

\section{Introduction}

\printbibliography
\end{document}

### 2019-03-01 paper.tex
Humble beginnings.
Some update to introduction.

@@
-%% Use \input{...} to include a standard preamble
+\documentclass[12pt]{article}
 
-\title{VAR Models and Granger Causality}
+%% bibliography stuff -- this needs come before the preamble inclusion
+\usepackage[backend=bibtex,sorting=none]{biblatex}
+\bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
+\usepackage{hyperref}
+
+\input{\string~/Documents/academics/global_academics/latex_preamble}
+
+\graphicspath{{./figures/}}
+
+\title{Structure Learning for $VAR(p)$ Models}
 \author{R. J. Kinnear, R. R. Mazumdar}
@@
 \maketitle
-\abstract{}
+\abstract{We study Granger Causality and propose a structure learning
+  heuristic for uncovering a parsimonious representation of large
+  $\mathsf{VAR}(p)$ models.}
 
-\section{Introduction}
+\section{Introduction and Review}
+Consider a collection of stochastic processes producing observations
+at discrete time intervals.  Are the underlying processes dependent?
+Can we quantify any of the underlying relationships?  Can the arrow of
+time help us to distinguish a directionality or flow of dependence
+among our observed series?  In this paper we contribute to the
+understanding of the notion of Granger-Causality
+\cite{granger1969investigating} as a tool for answering these questions.
+
+Though the notion of causality is a philosophically slippery concept,
+it is fundamental to the way we understand the world and to the
+progress of science in general.  Indeed, without faith in the
+consistency of causal interactions the results of experimental science
+could not be generalized or applied in any meaningful way.  In the
+case of Granger-Causality, we state that if an event '$A$' provides us
+with unique (that is, not available anywhere else) information about a
+later event $B$, then $A$ must have a causal impact on $B$.  As
+opposed to the notion of Causation promoted by Pearl
+\cite{pearl2000art}, this is an entirely model-free notion of cause,
+and instead leverages the intuition that a cause must precede it's
+effect.
+
+In practice, Granger's notion of causation it not a convincing test
+for \textit{true} causation, since our statements about causation are
+highly dependent upon the data that we are able to observe.  We prefer
+instead to interpret Granger causality as a means of uncovering a flow
+of ``information'' or ``energy'' through some underlying graph of
+interactions.  Though this graph cannot be observed directly, we will
+infer it's presence as a latent structure among our observed time
+series data.
+
+\section{Theory}
+\section{Structure Learning} 
+\section{Application}
+\section{Conclusion}
 
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2019-03-01 paper.tex
Some review and definitions in the paper.

@@
 \section{Introduction and Review}
+\label{sec:introduction}
 Consider a collection of stochastic processes producing observations
@@
 
+Finding the ``best'' graph structure consistent with observed data is
+generally an extremely challenging problem, though the comparison of
+quality between different structures, and hence the notion of
+``best'', needs to be quantified.  In applications where we are
+interested merely in minimizing the mean squared error of a linear
+one-step-ahead predictor, then we will naturally desire an entirely
+dense graph of connections, since each edge can only serve to reduce
+estimation error.  However, since the number of edges scales
+quadratically in $n$, it is imperative to infer a sparse causality
+graph for large systems, both to avoid overfitting observed data, as
+well as to aid the interpretability of the results.
+
+In \cite{bach2004learning} the authors apply a local search heuristic
+to an AIC penalized approximation of the likelihood where at each
+iteration an edge is either added, removed, or reversed.  This is a
+common approach to combinatorial optimization due to it's simplicity,
+but is liable to get stuck in shallow local minima.
+
+A second and wildly successful heuristic is the LASSO regularizer
+\cite{tibshirani1996regression}, which can be understood as a natural
+convex relaxation to penalizing the count of the non-zero edges.  The
+LASSO enjoys strong theoretical guarantees \cite{wainwright2009sharp},
+even in the case of time series data \cite{basu2015}
+\cite{wong2016lasso}.
+
+From a Bayesian perspective, spike-and-slab priors can be used to
+estimate the probability of an edge's inclusion, after which a graph
+can be chosen heuristically by thresholding or searching for the MAP
+graph structure.
+
+Our contribution is to provide an alternative heuristic based on
+minimum spanning trees, inspired by our result in section
+\ref{sec:theory} that the true causality graph can be recovered from
+pairwise information alone in the case that the underlying graph is a
+directed forest.  We compare this algorithm to a grouped LASSO
+\cite{yuan2006model}, local search, and a Bayesian posterior
+probability thresholding scheme.
+
+%% -Can we also formulate an SDP relaxation?
+%% -Fit trees jointly or via coordinate descent?
+
 \section{Theory}
+\label{sec:theory}
+Consider a wide sense stationary, vector valued and discrete time
+stochastic process $x(t) \in L^n_2(\Omega);\ t \in \Z$, having the matrix
+valued covariance sequence $R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ and spectra $S(\omega)$:
+
+\begin{equation}
+  \label{eqn:fourier_pair}
+  \begin{aligned}
+    R(\tau) &= \frac{1}{2\pi}\int_{-\pi}^\pi S(\omega) e^{j\tau\omega}\d \omega,\\
+    S(\omega) &= \sum_{\tau=-\infty}^\infty R(\tau)e^{-j\tau\omega}.
+  \end{aligned}
+\end{equation}
+
+\begin{definition}[Granger Causality]
+  For the WSS series $x(t)$ we say that component $x_j$ \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x(t)$) and write $x_j \rightarrow x_i$ if given Hilbert spaces $\mathcal{H}^{(i)}_{t - 1}$, $\mathcal{H}^{(i, j)}_{t - 1}$ generated by $\{x_i(t - \tau)\}_{\tau = 1}^\infty$ and $\{x_i(t - \tau), x_j(t - s\}_{\tau, s = 1}^\infty$ respectively we have
+
+\begin{equation}
+  \xi[x_i(t) \ |\ \mathcal{H}^{(i)}_{t - 1}] > \xi[x_i(t) \ |\ \mathcal{H}^{(i, j)}_{t - 1}],
+\end{equation}
+
+where $\xi[x \ |\ \mathcal{H}] = \E (x - \hat{\E}[x \ |\ \mathcal{H}])^2$ and $\hat{\E}[x \ |\ \mathcal{H}] = \text{proj}_{\mathcal{H}}(x)$ denotes the (unique) projection onto the Hilbert space $\mathcal{H}$.
+\end{definition}
+
+This notion captures the idea that the process $x_j$ provides
+information about $x_i$ that is not available from elsewhere.  The
+caveat ``with respect to $x(t)$'' is important in that GC relations
+can change when components are added to or removed from $x(t)$,
+e.g. new GC relations can arise if we remove the observations of a
+common cause, and existing GC relations can disappear if we observe a
+new mediating series.
+
+The notion is closely related to the information theoretic measure of
+transfer entropy, indeed, if everything in Gaussian then they are
+equivalent \cite{barnett2009granger}.
+
+We continue to review some pertinent facts.
+
+As a consequence of the Wold decomposition theorem \cite{lindquist},
+every such sequence has the ``moving average'' $MA(\infty)$ representation
+
+\begin{equation}
+\label{eqn:wold}
+  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) z(t - \tau);\ z(t) \overset{\text{i.i.d.}}{\sim} \mathcal{N}(0, I),
+\end{equation}
+
+where $c(t)$ is a perfectly predictable sequence\footnote{$c(t)$ is
+  perfectly predictable if any sample $c(t_0)$ is enough to determine
+  $c(t)$ for every $t$.  For example,
+  $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi, \pi]$}, and
+$z(t) \overset{\text{i.i.d.}}{\sim} \mathcal{N}(0, I)$.  We will assume
+that $c(t)$ can be effectively estimated and removed and take
+$c(t) = 0$.  Furthermore, we assume that the spectra is uniformly
+bounded away from $0$:
+$\exists \delta:\ \delta I \preceq S(\omega) \preceq \frac{1}{\delta} I\ \forall \omega$ so that the Wold
+representation can be inverted to yield the $VAR(\infty)$ form
+
+\begin{equation}
+  x(t) = \sum_{\tau = 1}^\infty B(\tau) x(t - \tau) + v(t);\ v(t) \overset{\text{i.i.d.}}{\sim} \mathcal{N}(0, \Sigma_v).
+\end{equation}
+
 \section{Structure Learning}
+\subsection{Edge-Wise Grouped LASSO}
+\subsection{Bayesian Posterior Thresholding}
+\subsection{Pairwise Minimum Error Spanning Trees}
 \section{Application}

### 2019-03-01 paper.tex
Some outlining and rough statements of theorems.

@@
 \usepackage{hyperref}
+
+\def\gc{\overset{\text{GC}}{\rightarrow}}
+\def\te{\overset{\mathcal{T}}{\rightarrow}}
 
@@
 \label{sec:theory}
-Consider a wide sense stationary, vector valued and discrete time
-stochastic process $x(t) \in L^n_2(\Omega);\ t \in \Z$, having the matrix
-valued covariance sequence $R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ and spectra $S(\omega)$:
+Consider a zero mean wide sense stationary, vector valued and discrete
+time stochastic process $x(t) \in L^n_2(\Omega);\ t \in \Z$, having the matrix
+valued covariance sequence
+$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ and spectra $S(\omega)$:
 
@@
 \begin{definition}[Granger Causality]
-  For the WSS series $x(t)$ we say that component $x_j$ \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x(t)$) and write $x_j \rightarrow x_i$ if given Hilbert spaces $\mathcal{H}^{(i)}_{t - 1}$, $\mathcal{H}^{(i, j)}_{t - 1}$ generated by $\{x_i(t - \tau)\}_{\tau = 1}^\infty$ and $\{x_i(t - \tau), x_j(t - s\}_{\tau, s = 1}^\infty$ respectively we have
+  For the WSS series $x(t)$ we say that component $x_j$
+  \textit{Granger-Causes} (GC) component $x_i$ (with respect to
+  $x(t)$) and write $x_j \gc x_i$ if given Hilbert spaces
+  $\mathcal{H}^{(i)}_{t - 1}$, $\mathcal{H}^{(i, j)}_{t - 1}$
+  generated by $\{x_i(t - \tau)\}_{\tau = 1}^\infty$ and
+  $\{x_i(t - \tau), x_j(t - s\}_{\tau, s = 1}^\infty$ respectively we have
 
@@
 
-where $\xi[x \ |\ \mathcal{H}] = \E (x - \hat{\E}[x \ |\ \mathcal{H}])^2$ and $\hat{\E}[x \ |\ \mathcal{H}] = \text{proj}_{\mathcal{H}}(x)$ denotes the (unique) projection onto the Hilbert space $\mathcal{H}$.
+where
+$\xi[x \ |\ \mathcal{H}] = \E (x - \hat{\E}[x \ |\ \mathcal{H}])^2$ and
+$\hat{\E}[x \ |\ \mathcal{H}] = \text{proj}_{\mathcal{H}}(x)$ denotes
+the (unique) projection onto the Hilbert space $\mathcal{H}$.
 \end{definition}
@@
 
+\begin{theorem}[Granger Causality Equivalences]
+  Given W.S.S. vector valued process $x(t)$ the following are equivalent:
+
+  \begin{enumerate}
+    \item{$x_j \gc x_i$}
+    \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$}
+    \item{$\mathcal{H}_t^{(i)} \perp \mathcal{H}_{t - 1}^{(j)}\ |\ \mathcal{H}_{t - 1}^{(i)}$}
+    \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} > 0$}
+    \item{\hl{Wold $A(\tau)$ condition}}
+    \item{Tie in the 0s of $S(\omega)^{-1}$, though this isn't immediately directed.}
+  \end{enumerate}
+\end{theorem}
+
+The following propositions justify various modifications of $x(t)$
+applied in practice to massage $x(t)$ into a form amenable to more
+standard tools.
+
+\begin{theorem}[Invariance Under Invertible and Deterministic Modifications]
+  Let $F(z), G(z)$ describe univariate and invertible
+  linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
+  predictable processes.  Then,
+
+  \begin{equation}
+    x_j \gc x_i \iff F(x_j - f)(t) \gc G(x_i - g)(t)
+  \end{equation}
+\end{theorem}
+
+\begin{theorem}[General Invariance]
+  \hl{Under what conditions is this actually true?}
+
+  Let $\zeta(t)$ be a stationary discrete time stochastic process.  Let
+  $F, G$ be (possibly nonlinear and time varying) invertible filtering
+  operations and $f(t)$, $g(t)$ be perfectly predictable and such that
+  $x_j(t) \defeq F(\zeta_j - f)(t), x_i(t) \defeq G(\zeta_i - g)(t)$ are W.S.S.  Then,
+
+\begin{equation}
+    x_j \gc x_i \iff \zeta_j \te \zeta_i
+  \end{equation}
+\end{theorem}
+
+\begin{theorem}[Tree Recovery]
+  \label{thm:tree_recovery}
+  If the Granger-Causality graph $\mathcal{G}$ for $x(t)$ is a
+  directed forest, then $\mathcal{G}$ can be inferred from pairwise
+  causality tests alone.
+\end{theorem}
+
 \section{Structure Learning}
@@
 \subsection{Bayesian Posterior Thresholding}
+Consider the Bayesian model
+
+\begin{equation}
+  \begin{aligned}
+    (x(t)\ |\ B, \sigma_v^2, \{x(t - \tau)\}_{\tau = 1}^p) &\sim \mathcal{N}(\sum_{\tau = 1}^pB(\tau)x(t - \tau), \sigma_v^2)\\
+    (B_{ij}(\tau)\ |\ G_{ij}(\tau)) &\sim G_{ij}(\tau)\mathcal{N}(0, \sigma_\beta^2) + (1 - G_{ij}(\tau))\delta_0(B_{ij}(\tau))\\
+    (G_{ij}(\tau)) &\sim \mathsf{BER}(p_{ij}(\tau))\\
+    \sigma_v^2 &\sim \Gamma(a_v, b_v)\\
+    \sigma_\beta^2 &\sim \Gamma(a_\beta, b_\beta)
+  \end{aligned}
+\end{equation}
+
+Similar models have been studied by various authors in the context of the linear regression model \hl{[(cite them)]} where it is referred to as \textit{stochastic variable selection}, or referred to as a model for Bayesian variable selection.
+
+Since we have in mind applications where $n$ is large, sampling posterior probabilities can be prohibitively burdensome, so we can instead fit the mean-field variational approximation $p(B, G, \sigma\ |\ X) \approx q_Gq_Bq_\sigma$ and subsequently apply a thresholding operation to the posterior edge inclusion probabilities under $q_G$.
+
 \subsection{Pairwise Minimum Error Spanning Trees}
+Inspired by theorem \ref{thm:tree_recovery}, we propose the following
+structure learning heuristic.
+
+%% This algorithm should perform coordinate descent or jointly refit all the trees.
+
+\begin{enumerate}
+  \item{Set $k = 0$ and $x^0(t) = x(t)$}
+  \item{Compute all pairwise causality measures $\Xi \defeq \big[\ln \frac{\xi_i}{\xi_{ij}} \big]_{i, j}$}
+  \item{Find the maximum spanning arborescence\footnote{An ``arborescence'' is a french word for a tree diagram, and refers to a \textit{directed} tree in graph theory} $\mathcal{T}_k$ of a graph having edge weights $\Xi_{ij}$.}
+  \item{Fit a vector LTI filter $F_k$ having graph defined by $\mathcal{T}_k$ to $x^{k - 1}(t)$ and set $x^k(t) = x^{k - 1}(t) - \hat{x}^{k - 1}(t)$}
+  \item{Repeat over $k$ until a stopping criteria is met}
+\end{enumerate}
+
 \section{Application}

### 2019-03-06 paper.tex
Added some definitions worked on lemma.

@@
 
-\def\gc{\overset{\text{GC}}{\rightarrow}}
-\def\te{\overset{\mathcal{T}}{\rightarrow}}
+\def\gc{\overset{\text{GC}}{\rightarrow}}  % Granger causality arrow
+\def\pwgc{\overset{\text{PW}}{\rightarrow}}  % Granger causality arrow
+\def\te{\overset{\mathcal{T}}{\rightarrow}}  % Transfer entropy arrow
+\def\gcg{\mathcal{G}}  % Granger-causality graph
+\def\VAR{\mathsf{VAR}}  % VAR(p) model
+
+\newcommand{\linE}[2]{\hat{\E}[#1\ |\ #2]}  % Linear projection
+\newcommand{\linEerr}[2]{\xi[#1\ |\ #2]}  % Error of linear projection
+\newcommand{\pa}[1]{P(#1)}  % Parents of a node
+\newcommand{\anc}[1]{\mathcal{A}(#1)}  % Ancestors of a node
 
@@
 valued covariance sequence
-$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ and spectra $S(\omega)$:
+$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$, which we assume to be
+absolutely summable, and spectra $S(\omega)$:
 
@@
   For the WSS series $x(t)$ we say that component $x_j$
-  \textit{Granger-Causes} (GC) component $x_i$ (with respect to
-  $x(t)$) and write $x_j \gc x_i$ if given Hilbert spaces
-  $\mathcal{H}^{(i)}_{t - 1}$, $\mathcal{H}^{(i, j)}_{t - 1}$
-  generated by $\{x_i(t - \tau)\}_{\tau = 1}^\infty$ and
-  $\{x_i(t - \tau), x_j(t - s\}_{\tau, s = 1}^\infty$ respectively we have
+  \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x$)
+  and write $x_j \gc x_i$ if given Hilbert spaces
+  $\mathcal{H}_{t - 1}$, $\mathcal{H}^{-j}_{t - 1}$
+  generated by $\{x_i(t - \tau)\}_{(\tau = 1, i = 1)}^{(\infty, n)}$ and
+  $\{x_i(t - \tau)\}_{(\tau = 1, i \ne j)}^{(\infty, n)}$ respectively we have
 
 \begin{equation}
-  \xi[x_i(t) \ |\ \mathcal{H}^{(i)}_{t - 1}] > \xi[x_i(t) \ |\ \mathcal{H}^{(i, j)}_{t - 1}],
+  \linEerr{x_i(t)}{\mathcal{H}_{t - 1}} < \linEerr{x_i(t)}{\mathcal{H}^{-j}_{t - 1}},
 \end{equation}
@@
 where
-$\xi[x \ |\ \mathcal{H}] = \E (x - \hat{\E}[x \ |\ \mathcal{H}])^2$ and
-$\hat{\E}[x \ |\ \mathcal{H}] = \text{proj}_{\mathcal{H}}(x)$ denotes
+$\xi[x \ |\ \mathcal{H}] = \E (x - \linE{x}{\mathcal{H}})^2$ and
+$\linE{x}{\mathcal{H}} = \text{proj}_{\mathcal{H}}(x)$ denotes
 the (unique) projection onto the Hilbert space $\mathcal{H}$.
@@
 \label{eqn:wold}
-  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) z(t - \tau);\ z(t) \overset{\text{i.i.d.}}{\sim} \mathcal{N}(0, I),
+  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) z(t - \tau),
 \end{equation}
@@
   $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi, \pi]$}, and
-$z(t) \overset{\text{i.i.d.}}{\sim} \mathcal{N}(0, I)$.  We will assume
+$z(t)$ is an uncorrelated sequence having covariance $I$.  We will assume
 that $c(t)$ can be effectively estimated and removed and take
@@
 \begin{equation}
-  x(t) = \sum_{\tau = 1}^\infty B(\tau) x(t - \tau) + v(t);\ v(t) \overset{\text{i.i.d.}}{\sim} \mathcal{N}(0, \Sigma_v).
+  \label{eqn:ar_representation}
+  x(t) = \sum_{\tau = 1}^\infty B(\tau) x(t - \tau) + v(t),
 \end{equation}
 
+where $v(t)$ is an uncorrelated sequence with covariance $\Sigma_v$.
+
 \begin{theorem}[Granger Causality Equivalences]
-  Given W.S.S. vector valued process $x(t)$ the following are equivalent:
+  Let $x(t)$ be a WSS process with absolutely summable covariance
+  sequence, and uniformly bounded spectral density.  Denote by
+  $\xi_{ij} \defeq \linEerr{x_i(t)}{\mathcal{H}^{-j}}$ and
+  $\xi_i \defeq \linEerr{x_i(t)}{\mathcal{H}_t}$.  Then, the following are equivalent:
 
   \begin{enumerate}
-    \item{$x_j \gc x_i$}
+    \item{$x_j \not\gc x_i$}
     \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$}
-    \item{$\mathcal{H}_t^{(i)} \perp \mathcal{H}_{t - 1}^{(j)}\ |\ \mathcal{H}_{t - 1}^{(i)}$}
-    \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} > 0$}
+    \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} = 0$}
+    \item{$\mathcal{H}_t^{i} \perp \mathcal{H}_{t - 1}^{j}\ |\ \mathcal{H}_{t - 1}^{-j}$}
     \item{\hl{Wold $A(\tau)$ condition}}
@@
 
+The equivalence $(1) \iff (3)$ is essentially a restatement of the
+definition, but is in line with the seminal work of Geweke
+\cite{geweke1982}, \cite{geweke1984}.  Statement (4) means that
+$(y - \linE{y}{\mathcal{H}_{t - 1}^{-j}}) \perp (z -
+\linE{z}{\mathcal{H}_{t - 1}^{-j}})\ \forall y \in \mathcal{H}_t^i,\ z \in
+\mathcal{H}_{t - 1}^j$, which is an appealing geometric statement that
+there is no direct feedback from $x_j$ to $x_i$ \cite{lindquist}.
+
+\begin{proof}
+$(1) \implies (2)$: The projection for $x_i(t)$ onto $\mathcal{H}_{t - 1}$ is (by definition) given by a form similar to equation \ref{eqn:ar_representation}:
+
+\[
+  \E|x_i(t) - \linE{x_i(t)}{\mathcal{H}_{t - 1}}|^2 = \E|v_i(t) + \sum_{\tau = 1}^\infty \sum_{k = 1}^n (B_{i, k}(\tau) - \hat{B}_{i, k}(\tau))x_k(t - \tau)|^2.
+\]
+
+ Since $v(t)$ in equation \ref{eqn:ar_representation} is temporally uncorrelated, it follows that the optimal projection is given by the model coefficients themselves.
+This holds similarly for the projection onto $\mathcal{H}_{t - 1}^{-j}$.
+Then by $(1)$ we have
+
+\[
+  \E|x_i(t) - \sum_{\tau = 1}^\infty \sum_{k = 1}^n B_{i, k}(\tau)x_k(t - \tau)|^2 = \E|x_i(t) - \sum_{\tau = 1}^\infty\sum_{k \ne j}B_{i, k}(\tau)x_k(t - \tau)|^2
+\]
+
+By the uniqueness of the projection we must have $\forall \tau\  B_{i, j}(\tau) = 0$.
+
+$(2) \implies (4)$: In computing $(y - \linE{y}{\mathcal{H}_{t = 1}^{-j}})$ for $y \in \mathcal{H}_t^i$ it is sufficient to consider $y = x_i(t)$ since $\mathcal{H}_{t - 1}^i \subseteq \mathcal{H}_{t - 1}^{-j}$, in which case $(x_i(t) - \linE{x_i(t)}{\mathcal{H}_{t = 1}^{-j}}) = v_i(t)$.
+$(4)$ follows since $v_i(t) \perp \mathcal{H}_{t - 1}$ and $\forall z \in \mathcal{H}_{t - 1}^j\ (z - \linE{z}{\mathcal{H}_{t - 1}^{-j}}) \in \mathcal{H}_{t - 1}$
+
+\end{proof}
+
 The following propositions justify various modifications of $x(t)$
@@
 standard tools.
-
-\begin{theorem}[Invariance Under Invertible and Deterministic Modifications]
-  Let $F(z), G(z)$ describe univariate and invertible
-  linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
-  predictable processes.  Then,
-
-  \begin{equation}
-    x_j \gc x_i \iff F(x_j - f)(t) \gc G(x_i - g)(t)
-  \end{equation}
-\end{theorem}
 
@@
 
+\begin{corollary}[Invariance Under Invertible and Deterministic Modifications]
+  Let $F(z), G(z)$ describe univariate and invertible
+  linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
+  predictable processes.  Then,
+
+  \begin{equation}
+    x_j \gc x_i \iff F(x_j - f)(t) \gc G(x_i - g)(t)
+  \end{equation}
+\end{corollary}
+
+\subsection{Pairwise Granger Causality}
+Recall that Granger-causality in general must be understood with
+respect to a particular universe of observations.  If $x_j \gc x_i$
+with respect to $x_{-k}$, it may not hold with respect to $x$.  That
+is, $x_k$ may be a common ancestor that when observed completely
+explains the connection from $x_j$ to $x_i$.  In particular, we will
+study \textit{pairwise} Granger-causality, where $x_i \pwgc x_j$ if
+$x_i$ Granger-causes $x_j$ with respect only to $(x_i, x_j)$.
+
+We establish first some notation.  Remaining in the context of a
+collection of $n$ WSS processes
+$x(t) = \big(x_1(t), \ldots, x_n(t)\big)$, we define the Granger-causality
+graph $\gcg$ to be the directed graph formed on $n$ vertices where an
+edge $(j, i) \in \gcg$ if and only if $x_j \gc x_i$.  We refer to the
+set of parents of $x_i$ in $\gcg$ by $\pa{i}$, that is,
+$j \in \pa{i} \iff x_j \gc x_i$.  The set of \textit{ancestors} of
+$x_i$ in $\gcg$ will be denoted $\anc{i}$, that is, $j \in \anc{i}$ if
+and only if there is a directed path in $\gcg$ from $j$ to $i$.  Even
+though the definitions of Granger-causality naturally suggests that
+each node in the graph should have a self-loop (unless the node is
+simply independent noise) we will not include nodes in their own
+parent sets: $i \not\in \pa{i} \subseteq \anc{i}$.  We note that $\gcg$ need not be a DAG.
+
+The edges of the graph can be given a general notion of ``weight'' by
+associating an edge $(j, i)$ with an LTI filter
+$\{B_{ij}(\tau)\}_{\tau = 1}^\infty$ from the $\VAR$ representation of equation
+\ref{eqn:ar_representation}.
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
+% I'm actually not sure this is true, I think that if there is a common ancestor it is
+% ambiguous.  This may be true for transfer entropy, but for Granger-causality even if
+% the information arriving at i is later than to j, it may be less corrupt and hence
+% still facilitate better predictions about k's future.
+
+Finally, the \textit{propagation delay} from $j$ to $i$, denoted $d(j, i)$, will be defined as
+
+\[
+  d(j, i) = \text{min}\big\{\sum_{k = 0}^{K - 1} d(p_k, p_{k + 1}})\ |\ i = p_0 \rightarrow p_1 \rightarrow \cdots \rightarrow p_K = j \text{ is a path in } \gcg \big\}
+\]
+
+and $d(j, i) \defeq \text{min}\{B_{i, j}(\tau) \ne 0\}$ if $j \in \pa{i}$.
+
+\begin{lemma}
+  \label{lem:pwgc_anc}
+  $x_j \pwgc x_i$ if and only if at least one of the following hold
+
+  \begin{enumerate}
+    \item{$j \in \anc{i}$}
+    \item{$\anc{i}\cap\anc{j} \ne \emptyset$.} %and $d(k, j) < d(k, i)$}
+  \end{enumerate}
+\end{lemma}
+
+% \begin{remark}
+%   This lemma formalizes the intuition of \textit{information flow} in $\gcg$.
+We have $x_j \pwgc x_i$ whenever $i$ is descended from $j$, or if they have a common ancestor (call it $k$) and ``information'' from $k$ reaches $j$ before it reaches $i$.
+% \end{remark}
+
+\begin{proof}
+\end{proof}
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
 \begin{theorem}[Tree Recovery]
   \label{thm:tree_recovery}
-  If the Granger-Causality graph $\mathcal{G}$ for $x(t)$ is a
-  directed forest, then $\mathcal{G}$ can be inferred from pairwise
+  If the Granger-Causality graph $\gcg$ for $x(t)$ is a
+  directed forest, then $\gcg$ can be inferred from pairwise
   causality tests alone.
@@
 \section{Structure Learning}
+\subsection{Local Search Heuristics}
+
 \subsection{Edge-Wise Grouped LASSO}
+Minimize the sparsity promoting regularized least squares problem and choose the hyper-parameters against the Akaike information criteria.
+
+\begin{equation}
+  L(\lambda, p) \defeq \underset{B}{\text{minimize}}\ \frac{1}{T}\sum_{t = 1}^T||x(t) - \sum_{\tau = 1}^pB(\tau)x(t - \tau)||_2^2 + \lambda \sum_{i, j}\big[\alpha||B_{i, j}||_2 + (1 - \alpha)||B_{i, j}||_1\big]
+\end{equation}
+
+\begin{equation}
+  \underset{\lambda, p}{\text{minimize}}\ L(\lambda, p) + \mathsf{AIC}(B^{(\lambda, p)})
+\end{equation}
+
 \subsection{Bayesian Posterior Thresholding}

### 2019-03-10 paper/paper.tex

--- 2019-03-06 paper.tex
@@
 \newcommand{\anc}[1]{\mathcal{A}(#1)}  % Ancestors of a node
+\newcommand{\gcgpath}[4]{#1 \rightarrow #2_1 \rightarrow \cdots \rightarrow #2_{#3} \rightarrow #4}
 
@@
 definition, but is in line with the seminal work of Geweke
-\cite{geweke1982}, \cite{geweke1984}.  Statement (4) means that
+\cite{geweke1982measurement}, \cite{geweke1984}.  Statement (4) means that
 $(y - \linE{y}{\mathcal{H}_{t - 1}^{-j}}) \perp (z -
@@
 respect to a particular universe of observations.  If $x_j \gc x_i$
-with respect to $x_{-k}$, it may not hold with respect to $x$.  That
-is, $x_k$ may be a common ancestor that when observed completely
+with respect to $x_{-k}$, it may not hold with respect to $x$.  For
+example, $x_k$ may be a common ancestor which when observed completely
 explains the connection from $x_j$ to $x_i$.  In particular, we will
@@
 
+% What if we take this as the definition of PWGC?  then, in a simply causal graph
+% the "normal" definition becomes a theorem.  Or, if lem:pwgc_anc is true then
+% maybe that is good enough?
+
+\begin{lemma}
+  \label{lem:pwgc_anc}
+  If $j \in \anc{i}$ then $j \pwgc i$.
+\end{lemma}
+\begin{proof}
+  This is actually not true.  e.g.  $x_j(t) = v_i(t) + x_k(t - 1) + x_{k'}(t - 1)$ with $x_k(t) = v_k(t) + ax_j(t - 1)$ and $x_{k'}(t) = v_{k'}(t) - ax_j(t - 1)$.
+\end{proof}
+
 % - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
@@
 
-Finally, the \textit{propagation delay} from $j$ to $i$, denoted $d(j, i)$, will be defined as
+% Finally, the \textit{propagation delay} from $j$ to $i$, denoted $d(j, i)$, will be defined as
 
-\[
-  d(j, i) = \text{min}\big\{\sum_{k = 0}^{K - 1} d(p_k, p_{k + 1}})\ |\ i = p_0 \rightarrow p_1 \rightarrow \cdots \rightarrow p_K = j \text{ is a path in } \gcg \big\}
-\]
+% \[
+%   d(j, i) = \text{min}\big\{\sum_{k = 0}^{K - 1} d(p_k, p_{k + 1}})\ |\ i = p_0 \rightarrow p_1 \rightarrow \cdots \rightarrow p_K = j \text{ is a path in } \gcg \big\}
+% \]
 
-and $d(j, i) \defeq \text{min}\{B_{i, j}(\tau) \ne 0\}$ if $j \in \pa{i}$.
+% and $d(j, i) \defeq \text{min}\{B_{i, j}(\tau) \ne 0\}$ if $j \in \pa{i}$.
 
@@
   \label{lem:pwgc_anc}
-  $x_j \pwgc x_i$ if and only if at least one of the following hold
+  If $x_j \pwgc x_i$ then one or more of the following hold
 
@@
     \item{$j \in \anc{i}$}
-    \item{$\anc{i}\cap\anc{j} \ne \emptyset$.} %and $d(k, j) < d(k, i)$}
+    \item{$\anc{i}\cap\anc{j} \ne \emptyset$.}
   \end{enumerate}
@@
 \begin{proof}
+  Is this true if we take \ref{lem:pwgc_anc} as an axiom?
 \end{proof}
@@
 
-\begin{theorem}[Tree Recovery]
+We are interested in understanding under what conditions a causality
+graph can be correctly recovered by pairwise testing alone.  It is not
+enough that the graph simply be sparse, as the following example shows.
+
+\begin{example}
+  If the causality graph $\gcg$ of the vector process $x(t)$ is
+  strongly connected\footnote{there is a directed path from each node
+    to every other node}, then the causality graph recovered via
+  pairwise causality tests is the complete graph.
+
+  This follows immediately from the fact (lemma \ref{lem:pwgc_anc})
+  that $j \in \anc{i} \implies j \pwgc i$.
+\end{example}
+
+\begin{definition}[Simply Causal]
+  If the Granger-Causality graph $\gcg$ for $x(t)$ is as a directed
+  acyclic graph such that there is at most one non-overlapping
+  directed path between any two nodes (excluding self-loops) then we
+  will refer to $\gcg$ as being \textit{simply causal}.
+\end{definition}
+
+\begin{example}
+  A directed tree (more generally directed forest) is a simply causal
+  graph.
+\end{example}
+
+\begin{lemma}
+  If $\gcg$ is simply causal then $i \pwgc j$ and $j \in \anc{i}$ are
+  \textit{alternatives}, that is $i \pwgc j \implies j \notin \anc{i}$.
+\end{lemma}
+\begin{proof}
+  Suppose that $\gcg$ is simply causal and that we have both
+  $i \pwgc j$ and $j \in \anc{i}$.  We will show that $i \in \anc{j}$,
+  which produces a contradiction (since $\gcg$ being a DAG is part of
+  the requirements for being simply causal).
+
+  Since $j \in \anc{i}$ there is a path $\gcgpath{j}{a}{p}{i}$.
+  Moreover, since $i \pwgc j$ by lemma \ref{lem:pwgc_anc} there must
+  be a common ancestor $k \in \anc{i} \cap \anc{j}$ (otherwise, we would
+  have $i \in \anc{j}$, contradicting the acyclic property).  The node
+  $k$ has paths to both $j,\ \gcgpath{k}{b}{q}{j}$ and $i,\ \gcgpath{k}{c}{r}{i}$.
+
+\end{proof}
+
+\begin{theorem}[Pairwise Recovery]
   \label{thm:tree_recovery}
-  If the Granger-Causality graph $\gcg$ for $x(t)$ is a
-  directed forest, then $\gcg$ can be inferred from pairwise
-  causality tests alone.
+  If the Granger-Causality graph $\gcg$ for $x(t)$ is simply causal
+  then then $\gcg$ can be inferred from pairwise causality tests
+  alone.
 \end{theorem}
+\begin{proof}
+
+\end{proof}
+
+\begin{example}
+  The condition of simple causality is merely a sufficient condition.
+For example, the complete directed graph with 2 nodes (e.g.
+$B(1) = \left[ \begin{array}{cc} 0 & 1 \\ 1 & 0 \end{array} \right]$) contains a loop but is pairwise recoverable.
+\end{example}
 

### 2019-03-15 paper.tex
Work on paper -- ancestor expansion.

@@
 \def\VAR{\mathsf{VAR}}  % VAR(p) model
+\def\B{\mathsf{B}}  % Filter B
+\def\wtB{\widetilde{\B}}  % General filter B
+\def\A{\mathsf{A}}  % Filter A
 
@@
 \newcommand{\linEerr}[2]{\xi[#1\ |\ #2]}  % Error of linear projection
-\newcommand{\pa}[1]{P(#1)}  % Parents of a node
+\newcommand{\pa}[1]{pa(#1)}  % Parents of a node
 \newcommand{\anc}[1]{\mathcal{A}(#1)}  % Ancestors of a node
+\newcommand{\ancn}[2]{\mathcal{A}_{#1}(#2)}  % nth ancestors of a node
+\newcommand{\gpn}[2]{gp_{#1}(#2)}  # nth generation grandparents
+\newcommand{\gcgpath}[4]{#1 \rightarrow #2_1 \rightarrow \cdots \rightarrow #2_{#3} \rightarrow #4}
 
@@
 \begin{theorem}[Granger Causality Equivalences]
+  \label{thm:gcg}
   Let $x(t)$ be a WSS process with absolutely summable covariance
@@
 respect to a particular universe of observations.  If $x_j \gc x_i$
-with respect to $x_{-k}$, it may not hold with respect to $x$.  That
-is, $x_k$ may be a common ancestor that when observed completely
+with respect to $x_{-k}$, it may not hold with respect to $x$.  For
+example, $x_k$ may be a common ancestor which when observed completely
 explains the connection from $x_j$ to $x_i$.  In particular, we will
-study \textit{pairwise} Granger-causality, where $x_i \pwgc x_j$ if
-$x_i$ Granger-causes $x_j$ with respect only to $(x_i, x_j)$.
+study \textit{pairwise} Granger-causality.
 
-We establish first some notation.  Remaining in the context of a
-collection of $n$ WSS processes
+\begin{definition}[Pairwise Granger-causality]
+  We will say that $x_j$ pairwise Granger-causes $x_i$ and write
+  $x_j \pwgc x_i$ if $x_i$ Granger-causes $x_j$ with respect only to
+  $(x_i, x_j)$.
+\end{definition}
+
+We first need to establish some graph theoretic notation and
+terminology, collected formally in a definition for the reader's
+convenient reference.
+
+\begin{definition}[Causality graph, Parents, etc...]
+
+Remaining in the context of a collection of $n$ WSS processes
 $x(t) = \big(x_1(t), \ldots, x_n(t)\big)$, we define the Granger-causality
 graph $\gcg$ to be the directed graph formed on $n$ vertices where an
-edge $(j, i) \in \gcg$ if and only if $x_j \gc x_i$.  We refer to the
-set of parents of $x_i$ in $\gcg$ by $\pa{i}$, that is,
-$j \in \pa{i} \iff x_j \gc x_i$.  The set of \textit{ancestors} of
-$x_i$ in $\gcg$ will be denoted $\anc{i}$, that is, $j \in \anc{i}$ if
-and only if there is a directed path in $\gcg$ from $j$ to $i$.  Even
-though the definitions of Granger-causality naturally suggests that
-each node in the graph should have a self-loop (unless the node is
-simply independent noise) we will not include nodes in their own
-parent sets: $i \not\in \pa{i} \subseteq \anc{i}$.  We note that $\gcg$ need not be a DAG.
+edge $(j, i) \in \gcg$ if and only if $x_j \gc x_i$.
 
-The edges of the graph can be given a general notion of ``weight'' by
-associating an edge $(j, i)$ with an LTI filter
-$\{B_{ij}(\tau)\}_{\tau = 1}^\infty$ from the $\VAR$ representation of equation
-\ref{eqn:ar_representation}.
+We refer to the set of \textit{parents} of $x_i$ in $\gcg$ by
+$\pa{i}$, that is, $j \in \pa{i} \iff x_j \gc x_i$.
+
+Even though the definitions of Granger-causality naturally suggests
+that each node in the graph should have a self-loop (unless the node
+is simply independent noise) we will not include nodes in their own
+parent sets: $i \not\in \pa{i} \subseteq \anc{i}$.  However, we do not
+immediately assume that $\gcg$ is a DAG, and so we will allow for $i$
+to be it's own \textit{ancestor}.
+
+The set of level $\ell$ \textit{grandparents} of node $i$, denoted
+$\gpn{\ell}{i}$, is the set such that $j \in \gpn{\ell}{i}$ if and only if
+there is a \textit{directed and cycle-free path} of length $\ell$ in
+$\gcg$ from $j$ to $i$ \textit{or} if $j = i$ and there is a cycle of
+length $n$ starting at $i$ and ending back at $i$, that is, $i$ can be
+it's own grandparent.  Clearly, $\pa{i} = \gpn{1}{i}$.
+
+Finally, the set of \textit{level $\ell$ ancestors} of
+$i$: $\ancn{\ell}{i} = \bigcup_{\lambda \le \ell}\gpn{\lambda}{i}$ is the set such that
+$j \in \ancn{\ell}{i}$ if and only if there is a directed and cycle-free
+path of length $\ell$ \textit{or less} in $\gcg$ from $j$ to $i$
+\textit{or} if $j = i$ and there is a directed cycle of length $\ell$ or
+less from $i$ back to $i$ in $\gcg$.  The set of \textit{all
+  ancestors} of $i$ (e.g. $\ancn{n}{i}$) is denoted simply $\anc{i}$.
+\end{definition}
+
+The edges of the Granger-causality graph $\gcg$ can be given a general
+notion of ``weight'' by associating an edge $(j, i)$ with a
+\textit{strictly causal} LTI filter
+$\B_{ij}(z) = \sum_{\tau = 1}^{\infty} B_{ij}(\tau)z^{-\tau}$ from the
+$\VAR$ representation of equation \ref{eqn:ar_representation}.  We
+will represent the action (convolution) of this filter compactly as
+
+\begin{equation}
+  \sum_{\tau = 1}^\infty B_{ij}(\tau)x_j(t - \tau) \defeq \B_{ij}(z)x_j(t).
+\end{equation}
+
+% What if we take this as the definition of PWGC?  then, in a simply causal graph
+% the "normal" definition becomes a theorem.  Or, if lem:pwgc_anc is true then
+% maybe that is good enough?
+
+From the $\VAR$ representation of $x(t)$ there is clearly a tight relationship between each node and it's parent nodes, indeed
+
+\begin{equation}
+  \label{eqn:parent_expanion}
+  x_i(t) = v_i(t) + \B_{ii}(z)x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t).
+\end{equation}
+
+% When is this guaranteed w.r.t. the full system?
+Furthermore, if the filter $\B_{ii}$ is invertible, we can write
+
+\begin{equation*}
+  \begin{aligned}
+    (1 - \B_{ii}(z))x_i(t) &= v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\\
+    \implies x_i(t) &= (1 - \B_{ii}(z))^{-1}\big[v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\big]\\
+    &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)
+  \end{aligned}
+\end{equation*}
+
+In general, we will write $\wtB$ for arbitrary filters maintaining the
+convention that $\wtB_{ii}$ is a filter that modifies the future of
+$x_i$ from it's past and $\wtB_{ij}$ a filter modifying $x_i$ from the
+past of $x_j$.  We will use the convention that the filter
+$\widetilde{\B}$ will represent simply ``a filter'' and is not
+necessarily the same as the filters serving as edge weights in $\gcg$
+(indeed, $\wtB$ will often be $0$), and moreover, we will allow for
+$\widetilde{\B}$ to potentially change from place to place and even
+from line to line in the same sequence of calculations as it's exact
+properties or specification are not important.
+
+We continue to develop the relationship between $x_i$ and it's family tree.
+
+\begin{proposition}
+  \label{prop:parent_expanding}
+  The component $x_i(t)$ of $x(t)$ can be represented as a linear
+  combination of a linear filtering operation acting on the driving
+  noise $v_i$ of $x_i$ as well as the parents of $x_i$:
+
+  \begin{equation}
+    \label{eqn:parent_expansion}
+    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t).
+  \end{equation}
+
+  Moreover, $x_i$ can be expanded in terms of it's ancestors and
+  driving noise only:
+
+  \begin{equation}
+    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \anc{i} \setminus\{i\}}\wtB_{ik}(z)v_k(t).
+  \end{equation}
+\end{proposition}
+\begin{proof}
+  The formula \ref{eqn:parent_expansion} was established in the
+  preceding paragraph and serves as a starting point for the induction
+  hypothesis:
+
+  \begin{equation}
+    \label{eqn:induction}
+    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
+  \end{equation}
+
+  The base case follows by expanding \ref{eqn:parent_expansion}, but we skip the calculations since they are nearly identical to the following induction step:
+
+  \begin{align*}
+    x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)\\
+           &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)\big[\wtB_{kk}v_k(t)\\
+    &\ \ \ \  + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t)\big]\\
+    &\overset{(b)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell + 1}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
+  \end{align*}
+
+  Where set $(a)$ follows by expanding each $x_k$ with equation \ref{eqn:parent_expansion} and $(b)$ requires the inversion of $\big(1 - \sum_{k \in \gpn{\ell}{i}\setminus\{i\}}\wtB_{ik}(z)\wtB_{ki}(z)\big)$ as well as using the fact that $\ancn{\ell}{i} = \ancn{\ell - 1}{i}\cup\gpn{\ell}{i}$.
+We here remind the reader of our earlier established convention that $\wtB$ simply represents the existence of a filter (possibly 0) and can change from place to place.
+
+The induction terminates after $n$ steps since in a graph with $n$ nodes $\gpn{n + 1}{i} = \emptyset$
+
+% Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.
+
+%   \begin{align*}
+%     x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)\\
+%            &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)\big[\wtB_{kk}v_{k}(t) + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t) \big]\\
+%            &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)v_{k}(t) + \sum_{k \in \gpn{2}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)
+%   \end{align*}
+
+% Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.
+
+\end{proof}
+
+\begin{proposition}
+  \label{prop:ancestor_properties}
+  In a general Granger-causality graph $\gcg$, if $j \pwgc i$ then at
+  least one of the following must hold
+
+  \begin{enumerate}
+    \item{$j \in \anc{i}$}
+    \item{$\anc{i} \cap \anc{j} \ne \emptyset$}
+  \end{enumerate}
+\end{proposition}
+\begin{proof}
+\end{proof}
+
+It seems intuitive that a converse of \ref{prop:ancestor_properties}
+would hold, e.g. $j \in \anc{i} \implies j \pwgc i$.  Unfortunately,
+this is not the case in general, as different paths through $\gcg$ can
+lead to cancellation.
+
+\begin{example}
+  Firstly, on $n = 4$ nodes, ``diamond'' shapes can lead to cancellation on paths of length 2:
+
+\begin{equation*}
+  x(t) =
+  \left[
+    \begin{array}{cccc}
+      0 & 0 & 0 & 0\\
+      a & 0 & 0 & 0\\
+      -a & 0 & 0 & 0\\
+      0 & 1 & 1 & 0\\
+    \end{array}
+  \right] x(t - 1) + v(t),
+\end{equation*}
+
+with $\E v(t) = 0,\ \E v(t)v(t - \tau)^\T = \delta_\tau I$.
+
+The graph associated to this system is not strongly causal since there are two paths between $1$ and $4$, namely: $1 \rightarrow 2 \rightarrow 4$ and $1 \rightarrow 3 \rightarrow 4$.
+However, by directly calculating
+
+\begin{align*}
+  x_4(t) &= x_2(t - 1) + x_3(t - 1) + v_4(t)\\
+         &= ax_1(t - 2) + av_2(t - 1) - ax_1(t - 2) -av_3(t - 1) + v_4(t)\\
+         &= a(v_2(t - 1) - v_3(t - 1)) + v_4(t),
+\end{align*}
+
+we see that, since $v(t)$ is spatially and temporally uncorrelated, $1 \not\pwgc 4$.
+\end{example}
+
+\begin{example}
+  A second example on $n = 3$ nodes is also worth examining, in this
+  cancellation is a result of differing time lags:
+
+\begin{equation*}
+  x(t) =
+  \left[
+    \begin{array}{ccc}
+      0 & 0 & 0\\
+      -a & 0 & 0\\
+      0 & 1 & 0\\
+    \end{array}
+  \right] x(t - 1) +
+  \left[
+    \begin{array}{ccc}
+      0 & 0 & 0\\
+      0 & 0 & 0\\
+      a & 0 & 0\\
+    \end{array}
+  \right] x(t - 2) + v(t)
+\end{equation*}
+
+Then
+
+\begin{align*}
+  x_2(t) &= v_2(t) - ax_1(t - 1)\\
+  x_3(t) &= v_3(t) + x_2(t - 1) + ax_1(t - 2)\\
+  \implies x_3(t) &= v_2(t - 1) + v_3(t),
+\end{align*}
+
+and again the graph is not strongly causal and $1 \not\pwgc 3$.
+\end{example}
+
+The previous examples seem rather pathological since we have simply
+constructed a case where different paths in the graph cancel exactly.
+It may be possible to rid ourselves of these pathologies for instance
+by defining a notion of ``robust'' pairwise causality where
+``$j \pwgc _R i$'' if $j \pwgc i$ for any ``small'' perturbation of
+the system matrix.  However, we do not believe such a modification of
+our theory can have useful practical implications since the difference
+between paths that ``nearly cancel'' rather than cancel exactly would
+in practice require inordinate amounts of data to resolve.
+
+Alternatively, we could restrict the coefficients of the system
+matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead, we will
+focus on the defining feature of time series networks, the topology of
+$\gcg$ and consider conditions thereon.
+
+\begin{definition}[Strongly Causal]
+  We will say that a Granger-causality graph $\gcg$ is \textit{strongly causal}
+if $\gcg$ is a DAG and there is at most 1 directed path between any two nodes.
+\end{definition}
+
+\begin{proposition}
+  \label{prop:sc_graph_common_anc}
+  In a strongly causal graph if $j \in \anc{i}$ and $k \in \anc{i} \cap \anc{j}$ then the unique path from $k$ to $i$ contains $j$.
+\end{proposition}
+\begin{proof}
+  Suppose that there is a path from $k$ to $i$ which does not contain $j$.
+In this case, there are multiple paths from $k$ to $i$ (one of which \textit{does} go through $j$ since $j \in \anc{i}$) which contradicts the assumption of strong causality.
+\end{proof}
+
+% Is pwgc transitive?
+
+In light of proposition \ref{prop:sc_graph_common_anc}, the following provides a converse to proposition \ref{prop:ancestor_properties}.
+
+\begin{proposition}
+  \label{lem:pwgc_anc}
+  If $\gcg$ is strongly causal and each self-loop filter
+  $1 - \B_{ii}(z)$ is invertible, then $j \in \anc{i} \implies j \pwgc i$.
+\end{proposition}
+\begin{proof}
+  Since strongly causal graphs are acyclic, this allows us to
+  recursively expand the parents of $x_i$ in $\gcg$ in the
+  $\VAR(\infty)$ representation of $x(t)$ since we will eventually
+  terminate at nodes $d$ such that $\pa{d} = \emptyset$.  To this end,
+  examining $x_i$ in equation \ref{eqn:ar_representation}
+
+  \begin{align*}
+    x_i(t) &= v_i(t) + \sum_{\tau = 1}^\infty \big[ B_{ii}(\tau)x_i(t - \tau) + \sum_{k \in \pa{i}} B_{ik}(\tau)x_k(t - \tau)\big]\\
+    \implies x_i(t) &= (1 - \B_{ii}(z))^{-1} \big[v_i(t) + \sum_{k \in \pa{i}} B_{ik}(\tau)x_k(t - \tau) \big]\\
+    &= \sum_{\tau = 1}^\infty \big[C_{ii}(\tau)v_i(t - \tau) + \sum_{k \in \pa{i}} C_{ik}(\tau)x_k(t - \tau)\big].
+  \end{align*}
+
+  If $\pa{i} = \emptyset$ then we simply have
+  $x_i(t) = \sum_{\tau = 1}^\infty C_{ii}(\tau)v_i(t)$ and
+  $C_{ii} = A_{ii}$ from equation \ref{eqn:wold}.
+
+  Applying this relation recursively we can write
+
+  \begin{equation*}
+    x_i(t) = \sum_{\tau = 1}^\infty\big[C_{ii}(\tau)v_i(t - \tau) + \sum_{k \in \anc{i}}C_{ik}(\tau)v_k(t - \tau) \big].
+  \end{equation*}
+
+\end{proof}
+
+That strong conditions need to be placed on the topology of $\gcg$ in
+order for lemma \ref{lem:pwgc_anc} to be true can be seen through the
+earlier examples.
+
+\begin{lemma}
+  If $\gcg$ is strongly causal, then any $\VAR(p)$ system on $\gcg$ is stable if and only if
+  $B_{ii}(z)$ is stable for every $i = 1, \ldots, n$.
+\end{lemma}
+\begin{proof}
+\end{proof}
+
+% $B(1) = \left[ \begin{array}{cc} 0 & 1 \\ 1 & 0 \end{array} \right]$
 

### 2019-03-15 software/algebra.py

@@ -1,0 +1,76 @@
+from sympy import (Matrix, Symbol, MatrixSymbol, pprint, eye, tensorproduct,
+                   linsolve, simplify, summation, oo)
+
+
+# In the system x(t) = Bx(t - 1) + v(t) where B is lower triangular,
+# is it possible that the pairwise error from x_2 to x_1 is lower than from x_1 to x_2
+# even though there is no direct connection?  What about the GCG likelihood ratio?
+
+xtm1 = MatrixSymbol("x(t - 1)", n=2, m=1)
+vt = MatrixSymbol("v(t)", n=2, m=1)
+
+b1 = Symbol("b_1")
+b2 = Symbol("b_2")
+a = Symbol("a")
+B = Matrix([[b1, 0], [a, b2]])
+
+sv1 = Symbol("sigma_v(1)^2", positive=True)
+sv2 = Symbol("sigma_v(2)^2", positive=True)
+Sv = Matrix([[sv1, 0], [0, sv2]])
+
+Rx0 = MatrixSymbol("R_x(0)", n=2, m=2)
+eq = Rx0 - (Sv + B * Rx0 * B.T)
+
+res = linsolve((eq[0, 0], eq[0, 1], eq[1, 0], eq[1, 1]),
+               Rx0[0, 0], Rx0[0, 1], Rx0[1, 0], Rx0[1, 1])
+
+Rx00, Rx01, Rx10, Rx11 = [simplify(v) for v in res.args[0]]
+Rx0 = Matrix([[Rx00, Rx01], [Rx10, Rx11]])
+Rx1 = simplify(B * Rx0)
+
+# Is this right?
+e00 = simplify(Rx0[0, 0] - (Rx1[0, 1]**2) / Rx0[0, 0])
+e = simplify(Rx0 - Rx1 * (Rx0.inv()) * Rx1.T)
+
+
+# System with 3 variables
+rho = Symbol("rho")
+a = Symbol("a")
+
+B = Matrix([[rho, 0, 0], [a, rho, 0], [a, 0, rho]])
+
+Rx0 = MatrixSymbol("R_x(0)", n=3, m=3)
+sv1 = Symbol("sigma_v(1)^2", positive=True)
+sv2 = Symbol("sigma_v(2)^2", positive=True)
[…]

### 2019-03-18 paper.tex
Some symbolic algebra, not important.

@@
 \newcommand{\ancn}[2]{\mathcal{A}_{#1}(#2)}  % nth ancestors of a node
-\newcommand{\gpn}[2]{gp_{#1}(#2)}  # nth generation grandparents
+\newcommand{\gpn}[2]{gp_{#1}(#2)}  % nth generation grandparents
 \newcommand{\gcgpath}[4]{#1 \rightarrow #2_1 \rightarrow \cdots \rightarrow #2_{#3} \rightarrow #4}
@@
 
-\title{Structure Learning for $VAR(p)$ Models}
+\title{Structure Learning for $\VAR(p)$ Models}
 \author{R. J. Kinnear, R. R. Mazumdar}
@@
 
-  The base case follows by expanding \ref{eqn:parent_expansion}, but we skip the calculations since they are nearly identical to the following induction step:
+  The base case follows by expanding each $x_k$ in equation \ref{eqn:induction} via equation \ref{eqn:parent_expansion}, but we skip the calculations since they are nearly identical to the following induction step:
 
@@
 
-  Where set $(a)$ follows by expanding each $x_k$ with equation \ref{eqn:parent_expansion} and $(b)$ requires the inversion of $\big(1 - \sum_{k \in \gpn{\ell}{i}\setminus\{i\}}\wtB_{ik}(z)\wtB_{ki}(z)\big)$ as well as using the fact that $\ancn{\ell}{i} = \ancn{\ell - 1}{i}\cup\gpn{\ell}{i}$.
+  Where equality $(a)$ follows by expanding each $x_k$ with equation \ref{eqn:parent_expansion} and $(b)$ requires the inversion of $\big(1 - \sum_{k \in \gpn{\ell}{i}\setminus\{i\}}\wtB_{ik}(z)\wtB_{ki}(z)\big)$ as well as using the fact that $\ancn{\ell}{i} = \ancn{\ell - 1}{i}\cup\gpn{\ell}{i}$.
 We here remind the reader of our earlier established convention that $\wtB$ simply represents the existence of a filter (possibly 0) and can change from place to place.

### 2019-03-18 software/algebra.py

@@ -1,76 +1,103 @@
-from sympy import (Matrix, Symbol, MatrixSymbol, pprint, eye, tensorproduct,
-                   linsolve, simplify, summation, oo)
+import numpy as np
+import matplotlib.pyplot as plt
 
+from sympy import (Matrix, Symbol, MatrixSymbol, pprint, eye, tensorproduct,
+                   linsolve, simplify, summation, oo, symbols, lambdify)
 
 # In the system x(t) = Bx(t - 1) + v(t) where B is lower triangular,
 # is it possible that the pairwise error from x_2 to x_1 is lower than from x_1 to x_2
 # even though there is no direct connection?  What about the GCG likelihood ratio?
 
-xtm1 = MatrixSymbol("x(t - 1)", n=2, m=1)
-vt = MatrixSymbol("v(t)", n=2, m=1)
+def mar18():
+    a, b, c = symbols("a, b, c")
+    B = Matrix([[0, 0, c], [a, 0, 0], [0, b, 0]])
 
-b1 = Symbol("b_1")
-b2 = Symbol("b_2")
-a = Symbol("a")
-B = Matrix([[b1, 0], [a, b2]])
+    B = B.subs({b: 0.5, a: 0.5})
+    eig = list(simplify(abs(e)) for e in B.eigenvals().keys())
 
-sv1 = Symbol("sigma_v(1)^2", positive=True)
-sv2 = Symbol("sigma_v(2)^2", positive=True)
-Sv = Matrix([[sv1, 0], [0, sv2]])
+    lmbda = lambdify(c, eig[0])
+    num_a = np.linspace(0, 10, 10000)
+    L = lmbda(num_a)
 
-Rx0 = MatrixSymbol("R_x(0)", n=2, m=2)
-eq = Rx0 - (Sv + B * Rx0 * B.T)
+    plt.plot(num_a, L)
+    plt.hlines(1, num_a[0], num_a[-1])
+    plt.show()
 
-res = linsolve((eq[0, 0], eq[0, 1], eq[1, 0], eq[1, 1]),
-               Rx0[0, 0], Rx0[0, 1], Rx0[1, 0], Rx0[1, 1])
[…]

### 2019-03-19 paper.tex
Progress on proof of pairwise ancestor theorem.

@@
 less from $i$ back to $i$ in $\gcg$.  The set of \textit{all
-  ancestors} of $i$ (e.g. $\ancn{n}{i}$) is denoted simply $\anc{i}$.
+  ancestors} of $i$ (i.e. $\ancn{n}{i}$) is denoted simply $\anc{i}$.
 \end{definition}
@@
 \textit{strictly causal} LTI filter
-$\B_{ij}(z) = \sum_{\tau = 1}^{\infty} B_{ij}(\tau)z^{-\tau}$ from the
-$\VAR$ representation of equation \ref{eqn:ar_representation}.  We
-will represent the action (convolution) of this filter compactly as
+
+\begin{equation}
+  \label{eqn:filter}
+  \B_{ij}(z) = \sum_{\tau = 1}^{\infty} B_{ij}(\tau)z^{-\tau}
+\end{equation}
+
+from the $\VAR$ representation of equation
+\ref{eqn:ar_representation}.  We will represent the action
+(convolution) of this filter compactly by taking $z^{-1}$ as the
+backshift operator when convenient
 
@@
 
-From the $\VAR$ representation of $x(t)$ there is clearly a tight relationship between each node and it's parent nodes, indeed
+From the $\VAR$ representation of $x(t)$ there is clearly a tight
+relationship between each node and it's parent nodes, indeed
 
@@
 
-In general, we will write $\wtB$ for arbitrary filters maintaining the
-convention that $\wtB_{ii}$ is a filter that modifies the future of
-$x_i$ from it's past and $\wtB_{ij}$ a filter modifying $x_i$ from the
-past of $x_j$.  We will use the convention that the filter
-$\widetilde{\B}$ will represent simply ``a filter'' and is not
-necessarily the same as the filters serving as edge weights in $\gcg$
-(indeed, $\wtB$ will often be $0$), and moreover, we will allow for
-$\widetilde{\B}$ to potentially change from place to place and even
-from line to line in the same sequence of calculations as it's exact
-properties or specification are not important.
+% In general, we will write $\wtB$ for arbitrary filters maintaining the
+% convention that $\wtB_{ii}$ is a filter that modifies the future of
+% $x_i$ from it's past and $\wtB_{ij}$ a filter modifying $x_i$ from the
+% past of $x_j$.  We will use the convention that the filter
+% $\widetilde{\B}$ will represent simply ``a filter'' and is not
+% necessarily the same as the filters serving as edge weights in $\gcg$
+% (indeed, $\wtB$ will often be $0$), and moreover, we will allow for
+% $\widetilde{\B}$ to potentially change from place to place and even
+% from line to line in the same sequence of calculations as it's exact
+% properties or specification are not important.
 
@@
 
-\begin{proposition}
+\begin{proposition}[Ancestor Expansion]
   \label{prop:parent_expanding}
-  The component $x_i(t)$ of $x(t)$ can be represented as a linear
-  combination of a linear filtering operation acting on the driving
-  noise $v_i$ of $x_i$ as well as the parents of $x_i$:
+  The component $x_i(t)$ of $x(t)$ can be represented as a sum of
+  linear filtering operations acting on the driving noise $v_i$ of
+  $x_i$ as well as the parents of $x_i$:
 
@@
   \begin{equation}
-    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \anc{i} \setminus\{i\}}\wtB_{ik}(z)v_k(t).
-  \end{equation}
-\end{proposition}
-\begin{proof}
-  The formula \ref{eqn:parent_expansion} was established in the
-  preceding paragraph and serves as a starting point for the induction
-  hypothesis:
-
-  \begin{equation}
-    \label{eqn:induction}
-    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
+    x_i(t) = \A_{ii}(z)v_i(t) + \sum_{k \in \anc{i} \setminus\{i\}}\A_{ik}(z)v_k(t),
   \end{equation}
 
-  The base case follows by expanding each $x_k$ in equation \ref{eqn:induction} via equation \ref{eqn:parent_expansion}, but we skip the calculations since they are nearly identical to the following induction step:
+  where $\A(z) = \sum_{\tau = 0}^\infty A(\tau)$ is the filter from
+  the Wold decomposition representation of $x(t)$, equation
+  (\ref{eqn:wold}).
+\end{proposition}
+
+This statement is ultimately about the sparsity pattern in the Wold
+decomposition matrices $A(\tau)$ since
+$x_i(t) = \sum_{\tau = 0}^\infty \sum_{j = 1}^n a_{ij}(\tau)v_j(t -
+\tau)$.  The proposition states that $a_{ij}(\tau) = 0$ unless
+$j \in \anc{i}$.
+
+% \begin{proof}
+%   The formula (\ref{eqn:parent_expansion}) was established in the
+%   preceding paragraph and serves as a starting point for the induction
+%   hypothesis:
+
+%   \begin{equation}
+%     \label{eqn:induction}
+%     x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
+%   \end{equation}
+
+%   The base case follows by expanding each $x_k$ in equation \ref{eqn:induction} via equation \ref{eqn:parent_expansion}, but we skip the calculations since they are nearly identical to the following induction step:
+
+%   \begin{align*}
+%     x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)\\
+%            &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)\big[\wtB_{kk}v_k(t)\\
+%     &\ \ \ \  + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t)\big]\\
+%     &\overset{(b)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell + 1}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
+%   \end{align*}
+
+%   Where equality $(a)$ follows by expanding each $x_k$ with equation \ref{eqn:parent_expansion} and $(b)$ requires the inversion of $\big(1 - \sum_{k \in \gpn{\ell}{i}\setminus\{i\}}\wtB_{ik}(z)\wtB_{ki}(z)\big)$ as well as using the fact that $\ancn{\ell}{i} = \ancn{\ell - 1}{i}\cup\gpn{\ell}{i}$.
+We here remind the reader of our earlier established convention that $\wtB$ simply represents the existence of a filter (possibly 0) and can change from place to place.
+
+% The induction terminates after $n$ steps since in a graph with $n$ nodes $\gpn{n + 1}{i} = \emptyset$
+
+% % Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.
+
+% %   \begin{align*}
+% %     x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)\\
+% %            &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)\big[\wtB_{kk}v_{k}(t) + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t) \big]\\
+% %            &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)v_{k}(t) + \sum_{k \in \gpn{2}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)
+% %   \end{align*}
+
+% % Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.
+
+% \end{proof}
+% \textbf{this is an alternative proof, I think superior to the one above}.
+
+The matrix $\B(z)$ is analagous to a \textit{weighted adjacency
+  matrix} for the graph $\gcg$.  And, in the same way that the
+$k^{\text{th}}$ power of an adjacency matrix counts the number of
+paths of length $k$ between nodes, $(\B(z)^k)_{ij}$ is a filter
+isolating the ``action'' of $j$ on $i$ at a time lag of $k$ steps.
+Though we only need the sparsity pattern of $\B(z)^k$ for the proof.
+
+We state the following well known theorem as a lemma, proof follows
+easily by induction.
+
+\begin{lemma}
+  \label{lem:adj_matrix}
+  Let $S$ be the transposed adjacency matrix of the Granger-causality
+  graph $\gcg$.  Then, $(S^k)_{ij}$ is the number of paths of length
+  $k$ from node $j$ to node $i$.  Evidently, if
+  $\forall k \in \N,\ (S^k)_{ij} = 0$ then $j \not\in \anc{i}$.
+\end{lemma}
+
+We now prove the proposition.
+
+\begin{proof}
+  From equation (\ref{eqn:ar_representation}), which we are assuming
+  throughout the paper to be invertible, we can write
+
+  \begin{equation*}
+    x(t) = (I - \B(z))^{-1} v(t),
+  \end{equation*}
+
+  where clearly $(I - \B(z))^{-1} = \A(z)$.  Since $\B(z)$ is stable,
+  that is $|\lambda_{\text{max}}(B(z))| < 1$, for every $|z| \le 1$ we
+  have
+
+  \begin{equation*}
+    (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k.
+  \end{equation*}
+
+  Now, let $S$ be a $0-1$ matrix containing the sparsity pattern of
+  $\B(z)$.  Then from lemma (\ref{lem:adj_matrix}) $S$ is the
+  transpose of the adjacency matrix for $\gcg$ and hence $(S^k)_{ij}$
+  is non-zero if and only if $j \in \gpn{k}{i}$, and therefore
+  $\A(z)_{ij}$ is zero if $j \not\in \anc{i}$.
+
+  Thence
 
   \begin{align*}
-    x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)\\
-           &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)\big[\wtB_{kk}v_k(t)\\
-    &\ \ \ \  + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t)\big]\\
-    &\overset{(b)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell + 1}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
+    x_i(t) &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
+    &= \A_{ii}(z) v_i(t) + \sum_{j \in \anc{i} \setminus \{i\}} \A_{ij}(z) v_j(t)\\
   \end{align*}
-
-  Where equality $(a)$ follows by expanding each $x_k$ with equation \ref{eqn:parent_expansion} and $(b)$ requires the inversion of $\big(1 - \sum_{k \in \gpn{\ell}{i}\setminus\{i\}}\wtB_{ik}(z)\wtB_{ki}(z)\big)$ as well as using the fact that $\ancn{\ell}{i} = \ancn{\ell - 1}{i}\cup\gpn{\ell}{i}$.
-We here remind the reader of our earlier established convention that $\wtB$ simply represents the existence of a filter (possibly 0) and can change from place to place.
-
-The induction terminates after $n$ steps since in a graph with $n$ nodes $\gpn{n + 1}{i} = \emptyset$
-
-% Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.
-
-%   \begin{align*}
-%     x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)\\
-%            &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)\big[\wtB_{kk}v_{k}(t) + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t) \big]\\
-%            &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)v_{k}(t) + \sum_{k \in \gpn{2}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)
-%   \end{align*}
-
-% Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.
 
@@
   \label{prop:ancestor_properties}
-  In a general Granger-causality graph $\gcg$, if $j \pwgc i$ then at
-  least one of the following must hold
+  In a Granger-causality graph $\gcg$, if $j \pwgc i$ then at least
+  one of the following must hold
 
@@
 \begin{proof}
+  We will show the contrapositive.
+Assume that $j \not \in \anc{i}$ and $\anc{i} \cap \anc{j} = \emptyset$, then $\forall \tau \in \N,\ \E x_i(t)x_j(t - \tau) = 0$ since from the representation (\ref{eqn:parent_expansion}) we have
+
+  \begin{align*}
+    \E x_i(t)x_j(t - \tau) &= \E \big(A_{ii}(z)v_i(t)\big)\big(A_{jj}(z)v_j(t - \tau)\big)\\
+    &+ \sum_{k \in \anc{i}}\E[\big(a_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{k \in \anc{j}}\E[v_i(t) \big(a_{jk}(z) v_k(t - \tau)\big)]\\
+    &+ \sum_{k \in \anc{i}}\sum_{\ell \in \anc{j}}\E[\big(a_{ik}(z)v_k(t)\big)\big(a_{j\ell}(z)v_\ell(t - \tau)\big)].
+  \end{align*}
+
+  Keeping in mind that $v(t)$ is an isotropic and uncorrelated sequence (i.e.
+$\forall \tau \in \Z,\ k \ne \ell \implies \E v_k(t)v_\ell(t - \tau) = 0$ and $\forall \tau \in \Z\setminus\{0\}\ \E v_i(t)v_i(t - \tau) = 0$), we see that each of these four terms is 0.
+The first term since $i \pwgc i$ is not possible and hence the case $j = i$ is vacuous.
+The second term is 0 since $j \not\in \anc{i}$.
+For the third term, it is possible that $i \in \anc{j}$, however, since $\A_{jk}(z)$ is causal and $\tau \ge 1$, $v_i(t)$ is later in time than $\A_{jk}(z)v_k(t - \tau)$, hence the third term is 0.
+The fourth term is 0 since $\anc{i}\cap\anc{j} = \emptyset$.
+
+The implication $\tau \ge 1 \implies \E x_i(t)x_j(t - \tau) = 0$ is sufficient for
+
 \end{proof}
@@
 It seems intuitive that a converse of \ref{prop:ancestor_properties}
-would hold, e.g. $j \in \anc{i} \implies j \pwgc i$.  Unfortunately,
+would hold, i.e. $j \in \anc{i} \implies j \pwgc i$.  Unfortunately,
 this is not the case in general, as different paths through $\gcg$ can
@@
   The condition of simple causality is merely a sufficient condition.
-For example, the complete directed graph with 2 nodes (e.g.
+For example, the complete directed graph with 2 nodes (i.e.
 $B(1) = \left[ \begin{array}{cc} 0 & 1 \\ 1 & 0 \end{array} \right]$) contains a loop but is pairwise recoverable.

### 2019-03-19 paper.tex
Proof of general pairwise topology implications.

@@
 \begin{theorem}[Granger Causality Equivalences]
-  \label{thm:gcg}
+  \label{thm:granger_causality_equivalences}
   Let $x(t)$ be a WSS process with absolutely summable covariance
@@
 
-  where clearly $(I - \B(z))^{-1} = \A(z)$.  Since $\B(z)$ is stable,
-  that is $|\lambda_{\text{max}}(B(z))| < 1$, for every $|z| \le 1$ we
-  have
+  where $(I - \B(z))^{-1} = \A(z)$ due to the uniqueness of
+  (\ref{eqn:wold}).  Since $\B(z)$ is stable, that is
+  $|\lambda_{\text{max}}(B(z))| < 1$, for every $|z| \le 1$ we have
 
@@
   is non-zero if and only if $j \in \gpn{k}{i}$, and therefore
-  $\A(z)_{ij}$ is zero if $j \not\in \anc{i}$.
+  $\B(z)^k_{ij} = 0$ if $j \not \in \gpn{k}{i}$.
+  $\anc{i} = \cup_k\gpn{k}{i}$ we see that $\A(z)_{ij}$ is zero if
+  $j \not\in \anc{i}$.
 
@@
   \end{align*}
+\end{proof}
 
+Before moving onto our next proposition, we need two lemmas formalizing
+some intuitively apparent facts.
+
+\begin{lemma}
+  \label{lem:subspace_sum_projection}
+  Let $x \in \mathcal{H}$ where $\mathcal{H}$ is a Hilbert space having inner product $\langle \cdot, \cdot \rangle$ and let $\{\mathcal{H}_n\}_{n = 0}^\infty$ be a collection of subspaces of $\mathcal{H}$ such that $x \perp \mathcal{H}_0$.
+Then
+
+  \begin{equation*}
+    \linE{x}{\bigvee_{n = 0}^\infty\mathcal{H}_n} = \linE{x}{\bigvee_{n = 1}^\infty\mathcal{H}_n},
+  \end{equation*}
+
+  where $\mathcal{H}_2 \vee \mathcal{H}_2 = \cl \{\alpha + \beta\ |\ \alpha \in \mathcal{H}_1, \beta \in \mathcal{H}_2 \}$ is the closed sum of subspaces.
+\end{lemma}
+\begin{proof}
+  We can apply the Gram-schmidt process to the collection of subspaces to obtain a sequence of orthogonal subspaces $\widetilde{\mathcal{H}}_0, \widetilde{\mathcal{H}}_1, \ldots$ with $\widetilde{\mathcal{H}}_0 = \mathcal{H}_0$, $\widetilde{\mathcal{H}}_n \subseteq \mathcal{H}_n$, and $\bigvee_{n = 0}^\infty\mathcal{H}_n = \bigoplus_{n = 0}^\infty \widetilde{\mathcal{H}}_n$, where $\oplus$ indicates the orthogonal sum of subspaces (which is always closed).
+Then, we have an explicit representation of the projection
+
+  \begin{align*}
+    \linE{x}{\bigvee_{n = 0}^\infty\mathcal{H}_n} &= \sum_{n = 0}^\infty \langle x, x_n \rangle x_n\\
+    &\overset{(a)}{=} \sum_{n = 1}^\infty \langle x, x_n \rangle x_n,
+  \end{align*}
+
+  where $x_n \in \widetilde{\mathcal{H}}_n \subseteq \mathcal{H}_n$ and $(a)$ follows because of our assumption that $x \perp \mathcal{H}_0$.
+This establishes that the projection is infact a member of the smaller subspace $\bigvee_{n = 1}^\infty\mathcal{H}_n$, and since the error can only increase by projecting onto this smaller subspace, we have the conclusion.
+\end{proof}
+
+The following lemma is reproduced from \cite{lindquist} here for convenience.
+
+\begin{lemma}[\cite{lindquist}]
+  \label{lem:conditional_orthogonality_equivalence}
+  Consider three closed subspaces of a Hilbert space $\mathcal{A}$, $\mathcal{B}$, $\mathcal{X}$.  The following statements are equivalent
+
+  \begin{enumerate}
+    \item{$\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$}
+    \item{$\linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}}\ \forall \beta \in \mathcal{B}$}
+    % \item{$\linE{\beta}{\mathcal{A}} = \linE{\linE{\beta}{\mathcal{X}}}{\mathcal{A}}\ \forall \beta \in \mathcal{B}}$}
+  \end{enumerate}
+\end{lemma}
+\begin{proof}
+  Proceding from the definition of conditional orthogonality we have
+
+  \begin{align*}
+    &\langle \alpha - \linE{\alpha}{\mathcal{X}}, \beta - \linE{\beta}{\mathcal{X}} \rangle = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
+    &{\iff} \langle \alpha + x - \linE{\alpha + x}{\mathcal{X}}, \beta - \linE{\beta}{\mathcal{X}} \rangle = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
+    &\overset{(a)}{\iff} \langle \alpha + x, \beta - \linE{\beta}{\mathcal{X}} \rangle = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
+    &\overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A} \vee \mathcal{X}} = 0\ \forall \beta \in \mathcal{B}\\
+    &\overset{(c)}{\iff} \linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}},
+  \end{align*}
+
+  where $(a)$ follows since $(\beta - \linE{\beta}{\mathcal{X}}) \perp \mathcal{X}$ and $\linE{\alpha + x}{\mathcal{X}} \in \mathcal{X}$, $(b)$ since $z \perp \mathcal{X} \vee \mathcal{A} \iff \linE{z}{\mathcal{X} \vee \mathcal{A}} = 0$, and $(d)$ since $\linE{z}{\mathcal{H}_2} = \linE{\linE{z}{\mathcal{H}_2}}{\mathcal{H}_1}$ for $\mathcal{H}_2 \subseteq \mathcal{H}_1$.
+
+  % This is property (vi) in Lindquist
+  % \begin{align*}
+  %   \langle \alpha - \linE{\alpha}{\mathcal{X}}, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
+  %   \overset{(a)}{\iff} \langle \alpha, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0\ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
+  %   \overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A}} &= 0\ \forall \beta \in \mathcal{B}
+  % \end{align*}
+
+  % where $(a)$ follows because $\linE{\alpha}{\mathcal{X}} \in \mathcal{X}$ and $\beta - \linE{\beta}{\mathcal{X}} \in \mathcal{X}$, and $(b)$ follows since $\beta \perp \mathcal{A} \iff \linE{\beta}{\mathcal{A}} = 0$.
 \end{proof}
@@
   We will show the contrapositive.
-Assume that $j \not \in \anc{i}$ and $\anc{i} \cap \anc{j} = \emptyset$, then $\forall \tau \in \N,\ \E x_i(t)x_j(t - \tau) = 0$ since from the representation (\ref{eqn:parent_expansion}) we have
+Assume that $j \not \in \anc{i}$ and $\anc{i} \cap \anc{j} = \emptyset$, then $\forall \tau \in \Z_+,\ \E x_i(t)x_j(t - \tau) = 0$ since from the representation (\ref{eqn:parent_expansion}) we have
 
   \begin{align*}
-    \E x_i(t)x_j(t - \tau) &= \E \big(A_{ii}(z)v_i(t)\big)\big(A_{jj}(z)v_j(t - \tau)\big)\\
-    &+ \sum_{k \in \anc{i}}\E[\big(a_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{k \in \anc{j}}\E[v_i(t) \big(a_{jk}(z) v_k(t - \tau)\big)]\\
-    &+ \sum_{k \in \anc{i}}\sum_{\ell \in \anc{j}}\E[\big(a_{ik}(z)v_k(t)\big)\big(a_{j\ell}(z)v_\ell(t - \tau)\big)].
+    \E x_i(t)x_j(t - \tau) &= \E \big(\A_{ii}(z)v_i(t)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)\\
+    &+ \sum_{k \in \anc{i}}\E[\big(\A_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{k \in \anc{j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
+    &+ \sum_{k \in \anc{i}}\sum_{\ell \in \anc{j}}\E[\big(\A_{ik}(z)v_k(t)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)]\\
   \end{align*}
@@
   Keeping in mind that $v(t)$ is an isotropic and uncorrelated sequence (i.e.
-$\forall \tau \in \Z,\ k \ne \ell \implies \E v_k(t)v_\ell(t - \tau) = 0$ and $\forall \tau \in \Z\setminus\{0\}\ \E v_i(t)v_i(t - \tau) = 0$), we see that each of these four terms is 0.
+$\forall \tau \in \Z,\ k \ne \ell \implies \E v_k(t)v_\ell(t - \tau) = 0$ and $\forall \tau \in \Z\setminus\{0\}\ \E v_i(t)v_i(t - \tau) = 0$), we will see that each of these above four terms are 0.
 The first term since $i \pwgc i$ is not possible and hence the case $j = i$ is vacuous.
 The second term is 0 since $j \not\in \anc{i}$.
-For the third term, it is possible that $i \in \anc{j}$, however, since $\A_{jk}(z)$ is causal and $\tau \ge 1$, $v_i(t)$ is later in time than $\A_{jk}(z)v_k(t - \tau)$, hence the third term is 0.
-The fourth term is 0 since $\anc{i}\cap\anc{j} = \emptyset$.
+The fourth term above is 0 since $\anc{i} \cap \anc{j} = \emptyset$.
+For the third term, it is possible that (due to cycles) $i \in \anc{j}$ which requires more care, however similarly as before we can at least see that
 
-The implication $\tau \ge 1 \implies \E x_i(t)x_j(t - \tau) = 0$ is sufficient for
+  \begin{equation*}
+    \sum_{k \in \anc{j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)] = \E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)].
+  \end{equation*}
 
+  However, $k \ne \ell \implies \A_{k\ell}(z) = 0$ (i.e.
+$\A(z) = I + A(1)z^{-1} + \cdots$) so $\A_{ji}(z)v_i(t - \tau)$ is strictly earlier in time, for $\tau \ge 0$, than $v_i(t)$ and hence $\E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)] = 0$.
+This establishes the fact that $H_t^{(i)} \perp \mathcal{H}_t^{(j)}$ where $H_t^{(i)} = \{\alpha x_i(t)\ |\ \alpha \in \R\}$ is the Hilbert space generated by the current sample of $x_i$.
+It is clear that $\mathcal{H}_t^{(i)} = \bigvee_{\tau = 0}^\infty H_{t - \tau}^{(i)}$.
+
+  Then for every $s \ge 0$
+
+  \begin{align*}
+    \linE{x_j(t - s)}{\mathcal{H}_t^{(i)}} &= \linE{x_j(t - s)}{\bigvee_{\tau = 0}^\infty H_{t - \tau}^{(i)}}\\
+    &\overset{(a)}{=} \linE{x_j(t - s)}{\bigvee_{\tau = 1}^\infty H_{t - \tau}^{(i)}}\\
+    &= \linE{x_j(t - s)}{\mathcal{H}_{t - 1}^{(i)}},
+  \end{align*}
+
+  where $(a)$ follows from lemma (\ref{lem:subspace_sum_projection}).  Lemma (\ref{lem:conditional_orthogonality_equivalence}) tells us that this is equivalent to
+
+  \begin{equation*}
+    \mathcal{H}_t^{(i)} \perp \mathcal{H}_{t - 1}^{(j)}\ |\ \mathcal{H}_{t - 1}^{(i)},
+  \end{equation*}
+
+  so by theorem (\ref{thm:granger_causality_equivalences}) we conclude $j \not\pwgc i$.
 \end{proof}
+
+We can state some interesting corollaries of this result.
+
+\begin{corollary}
+  If the graph $\gcg$ is acyclic and if $j \pwgc i$ and $i \pwgc j$ then $\anc{i} \cap \anc{j} \ne \emptyset$.
+\end{corollary}
 
@@
 
-The graph associated to this system is not strongly causal since there are two paths between $1$ and $4$, namely: $1 \rightarrow 2 \rightarrow 4$ and $1 \rightarrow 3 \rightarrow 4$.
-However, by directly calculating
+By directly calculating
 
@@
 
-we see that, since $v(t)$ is spatially and temporally uncorrelated, $1 \not\pwgc 4$.
+we see that, since $v(t)$ is isotropic white noise, $1 \not\pwgc 4$.  The problem here is that there are multiple paths from $x_1$ to $x_4$.
 \end{example}
@@
 
-and again the graph is not strongly causal and $1 \not\pwgc 3$.
+and again $1 \not\pwgc 3$.
 \end{example}
@@
 by defining a notion of ``robust'' pairwise causality where
-``$j \pwgc _R i$'' if $j \pwgc i$ for any ``small'' perturbation of
-the system matrix.  However, we do not believe such a modification of
-our theory can have useful practical implications since the difference
-between paths that ``nearly cancel'' rather than cancel exactly would
-in practice require inordinate amounts of data to resolve.
+``$j \pwgc _R i$'' if $j \pwgc i$ almost surely for a ``small'' random
+perturbation of the system matrix.  However, we do not believe such a
+modification of our theory can have useful practical implications
+since the difference between paths that ``nearly cancel'' rather than
+cancel exactly would in practice require inordinate amounts of data to
+resolve.
 
 Alternatively, we could restrict the coefficients of the system
-matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead, we will
-focus on the defining feature of time series networks, the topology of
-$\gcg$ and consider conditions thereon.
+matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead, we
+think it is more meaningful to focus on the defining feature of time
+series networks, that is, the topology of $\gcg$.
 
@@
 \end{definition}
+
+Examples of strongly causal graphs include directed trees (or
+forests), and DAGs where each node has at most one parent.
 
@@
 
-In light of proposition \ref{prop:sc_graph_common_anc}, the following provides a converse to proposition \ref{prop:ancestor_properties}.
+In light of proposition \ref{prop:sc_graph_common_anc}, the following
+provides a converse to proposition \ref{prop:ancestor_properties}.
 
@@
   \label{lem:pwgc_anc}
-  If $\gcg$ is strongly causal and each self-loop filter
-  $1 - \B_{ii}(z)$ is invertible, then $j \in \anc{i} \implies j \pwgc i$.
+  If $\gcg$ is strongly causal then $j \in \anc{i} \implies j \pwgc i$.
 \end{proposition}

### 2019-03-19 software/algebra.py

@@ -8,6 +8,17 @@ from sympy import (Matrix, Symbol, MatrixSymbol, pprint, eye, tensorproduct,
 # is it possible that the pairwise error from x_2 to x_1 is lower than from x_1 to x_2
 # even though there is no direct connection?  What about the GCG likelihood ratio?
 
+def mar19():
+    a, b, c, d = symbols("a, b, c, d")
+    z = Symbol("z")
+    B = Matrix([[0, a, 0],
+                [0, 0, b],
+                [d, c, 0]])
+
+    Bz = B * z
+    A = simplify((eye(3) - Bz).inv())
+    return
+
 def mar18():
     a, b, c = symbols("a, b, c")
     B = Matrix([[0, 0, c], [a, 0, 0], [0, b, 0]])

### 2019-04-09 paper.tex
Some formal updates and prose.

@@
 \def\A{\mathsf{A}}  % Filter A
+\def\H{\mathcal{H}}  % Hilbert space
 
@@
 \label{sec:theory}
-Consider a zero mean wide sense stationary, vector valued and discrete
-time stochastic process $x(t) \in L^n_2(\Omega);\ t \in \Z$, having the matrix
-valued covariance sequence
-$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$, which we assume to be
-absolutely summable, and spectra $S(\omega)$:
+\subsection{Formal Setting}
+Consider the space $L_2(\Omega)$, the usual Hilbert space of finite
+variance random variables over a probability space
+$(\Omega, \mathcal{F}, \mathbb{P})$ having inner product
+$\inner{x}{y} = \E[xy]$.  We will denote the collection vectors of
+such random elements as $L_2^n(\Omega)$ or simply $L_2^n$.
+
+We will work with a discrete time and wide-sense stationary process
+$x(t)$ taking elements in $L_2$.  We suppose that $x(t)$ has zero
+mean, $\E x(t) = 0$, and the matrix valued covariance sequence
+$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ is absolutely summable and has a spectra $S(\omega)$:
 
@@
 
-\begin{definition}[Granger Causality]
-  For the WSS series $x(t)$ we say that component $x_j$
-  \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x$)
-  and write $x_j \gc x_i$ if given Hilbert spaces
-  $\mathcal{H}_{t - 1}$, $\mathcal{H}^{-j}_{t - 1}$
-  generated by $\{x_i(t - \tau)\}_{(\tau = 1, i = 1)}^{(\infty, n)}$ and
-  $\{x_i(t - \tau)\}_{(\tau = 1, i \ne j)}^{(\infty, n)}$ respectively we have
+We will also work frequently with the spaces spanned by the values of
+such a process
 
 \begin{equation}
-  \linEerr{x_i(t)}{\mathcal{H}_{t - 1}} < \linEerr{x_i(t)}{\mathcal{H}^{-j}_{t - 1}},
+  \label{eq:hilbert_space_defn}
+  \begin{aligned}
+    \H^x &= \cl \{\sum_{\tau = -T}^T a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, T \in \N\} \subseteq L_2(\Omega),\\
+    \H_t^x &= \cl \{\sum_{\tau = 0}^T a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, T \in \N\} \subseteq L_2(\Omega),
+  \end{aligned}
 \end{equation}
 
-where
-$\xi[x \ |\ \mathcal{H}] = \E (x - \linE{x}{\mathcal{H}})^2$ and
-$\linE{x}{\mathcal{H}} = \text{proj}_{\mathcal{H}}(x)$ denotes
-the (unique) projection onto the Hilbert space $\mathcal{H}$.
-\end{definition}
-
-This notion captures the idea that the process $x_j$ provides
-information about $x_i$ that is not available from elsewhere.  The
-caveat ``with respect to $x(t)$'' is important in that GC relations
-can change when components are added to or removed from $x(t)$,
-e.g. new GC relations can arise if we remove the observations of a
-common cause, and existing GC relations can disappear if we observe a
-new mediating series.
-
-The notion is closely related to the information theoretic measure of
-transfer entropy, indeed, if everything in Gaussian then they are
-equivalent \cite{barnett2009granger}.
-
-We continue to review some pertinent facts.
+where the closure is naturally in mean-square.  $\H_t^x$ is the space
+spanned by the values of $x(t)$ between the ``infinite past'' and
+``now'', we will often omit the superscript $x$ which should be clear
+from context.  Evidently these spaces are separable and as closed
+subspaces of a Hilbert space they are themselves Hilbert.  We will
+denote the spaces generated in analagous ways by particular components
+of $x$ as e.g. $\H_t^{(i, j)}$, $\H_t^{i}$ or by all but particular
+components as e.g. $\H_t^{-j}$.
 
 As a consequence of the Wold decomposition theorem \cite{lindquist},
-every such sequence has the ``moving average'' $MA(\infty)$ representation
+every such sequence has the ``moving average'' $MA(\infty)$
+representation
 
@@
 \label{eqn:wold}
-  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) z(t - \tau),
+  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) w(t - \tau),
 \end{equation}
@@
   $c(t)$ for every $t$.  For example,
-  $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi, \pi]$}, and
-$z(t)$ is an uncorrelated sequence having covariance $I$.  We will assume
-that $c(t)$ can be effectively estimated and removed and take
-$c(t) = 0$.  Furthermore, we assume that the spectra is uniformly
-bounded away from $0$:
-$\exists \delta:\ \delta I \preceq S(\omega) \preceq \frac{1}{\delta} I\ \forall \omega$ so that the Wold
-representation can be inverted to yield the $VAR(\infty)$ form
+  $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi,
+  \pi]$}, and $w(t)$ is an uncorrelated sequence having a diagonal
+covariance and $A(0) = I$.  We will assume that $c(t) = 0$, which in
+practice is to say that it has been accurately estimated and removed.
+
+Furthermore, we assume that the spectra is uniformly bounded away from
+$0$:
+$\exists \delta:\ \delta I \preceq S(\omega) \preceq \frac{1}{\delta}
+I\ \forall \omega$ so that the Wold representation can be inverted to
+yield the $VAR(\infty)$ form
 
@@
 
+The equations (\ref{eqn:wold}), (\ref{eqn:ar_representation}) can be
+represented as $x(t) = \A(z)w(t) = \B(z)x(t) + v(t)$ via the action of
+the operators (LTI filters)
+$\A(z) \defeq \sum_{\tau = 0}^\infty A(\tau)z^{-\tau}$ and
+$\B(z) \defeq \sum_{\tau = 1}^\infty B(\tau)z^{-\tau}$ where the
+operator $z^{-1}$ is the backshift operator acting on $\H$.  Since
+$||z^{-1}|| = 1$, the inversion formulae
+
+\begin{equation}
+  \label{eqn:lsi_inversion}
+  \A(z) = (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k
+\end{equation}
+
+are valid.
+\subsection{Granger Causality}
+
+\begin{definition}[Granger Causality]
+  For the WSS series $x(t)$ we say that component $x_j$
+  \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x$)
+  and write $x_j \gc x_i$ if given Hilbert spaces
+  $\H_{t - 1}$, $\H^{-j}_{t - 1}$
+
+\begin{equation}
+  \linEerr{x_i(t)}{\H_{t - 1}} < \linEerr{x_i(t)}{\H^{-j}_{t - 1}},
+\end{equation}
+
+where $\xi[x \ |\ \H] = \E (x - \linE{x}{\H})^2$ is the mean squared
+estimation error and $\linE{x}{\H} = \text{proj}_{\H}(x)$ denotes the
+(unique) projection onto the Hilbert space $\H$.
+\end{definition}
+
+This notion captures the idea that the process $x_j$ provides
+information about $x_i$ that is not available from elsewhere.  The
+caveat ``with respect to $x$'' is important in that GC relations can
+change when components are added to or removed from our collection $x$
+of observations, e.g. new GC relations can arise if we remove the
+observations of a common cause, and existing GC relations can
+disappear if we observe a new mediating series.
+
+The notion is closely related to the information theoretic measure of
+transfer entropy, indeed, if everything is Gaussian then they are
+equivalent \cite{barnett2009granger}.
+
 \begin{theorem}[Granger Causality Equivalences]
@@
   Let $x(t)$ be a WSS process with absolutely summable covariance
-  sequence, and uniformly bounded spectral density.  Denote by
-  $\xi_{ij} \defeq \linEerr{x_i(t)}{\mathcal{H}^{-j}}$ and
-  $\xi_i \defeq \linEerr{x_i(t)}{\mathcal{H}_t}$.  Then, the following are equivalent:
+  sequence, and spectral density uniformly bounded above and below.  Denote by
+  $\xi_{ij} \defeq \linEerr{x_i(t)}{\H^{-j}}$ and
+  $\xi_i \defeq \linEerr{x_i(t)}{\H_t}$.  Then, the following are equivalent:
 
@@
     \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} = 0$}
-    \item{$\mathcal{H}_t^{i} \perp \mathcal{H}_{t - 1}^{j}\ |\ \mathcal{H}_{t - 1}^{-j}$}
+    \item{$\H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
     \item{\hl{Wold $A(\tau)$ condition}}
@@
 \cite{geweke1982measurement}, \cite{geweke1984}.  Statement (4) means that
-$(y - \linE{y}{\mathcal{H}_{t - 1}^{-j}}) \perp (z -
-\linE{z}{\mathcal{H}_{t - 1}^{-j}})\ \forall y \in \mathcal{H}_t^i,\ z \in
-\mathcal{H}_{t - 1}^j$, which is an appealing geometric statement that
+$(y - \linE{y}{\H_{t - 1}^{-j}}) \perp (z -
+\linE{z}{\H_{t - 1}^{-j}})\ \forall y \in \H_t^i,\ z \in
+\H_{t - 1}^j$, which is an appealing geometric statement that
 there is no direct feedback from $x_j$ to $x_i$ \cite{lindquist}.
@@
 \begin{proof}
-$(1) \implies (2)$: The projection for $x_i(t)$ onto $\mathcal{H}_{t - 1}$ is (by definition) given by a form similar to equation \ref{eqn:ar_representation}:
+$(1) \implies (2)$: The projection for $x_i(t)$ onto $\H_{t - 1}$ is (by definition) given by a form similar to equation \ref{eqn:ar_representation}:
 
 \[
-  \E|x_i(t) - \linE{x_i(t)}{\mathcal{H}_{t - 1}}|^2 = \E|v_i(t) + \sum_{\tau = 1}^\infty \sum_{k = 1}^n (B_{i, k}(\tau) - \hat{B}_{i, k}(\tau))x_k(t - \tau)|^2.
+  \E|x_i(t) - \linE{x_i(t)}{\H_{t - 1}}|^2 = \E|v_i(t) + \sum_{\tau = 1}^\infty \sum_{k = 1}^n (B_{i, k}(\tau) - \hat{B}_{i, k}(\tau))x_k(t - \tau)|^2.
 \]
@@
  Since $v(t)$ in equation \ref{eqn:ar_representation} is temporally uncorrelated, it follows that the optimal projection is given by the model coefficients themselves.
-This holds similarly for the projection onto $\mathcal{H}_{t - 1}^{-j}$.
+This holds similarly for the projection onto $\H_{t - 1}^{-j}$.
 Then by $(1)$ we have
@@
 
-$(2) \implies (4)$: In computing $(y - \linE{y}{\mathcal{H}_{t = 1}^{-j}})$ for $y \in \mathcal{H}_t^i$ it is sufficient to consider $y = x_i(t)$ since $\mathcal{H}_{t - 1}^i \subseteq \mathcal{H}_{t - 1}^{-j}$, in which case $(x_i(t) - \linE{x_i(t)}{\mathcal{H}_{t = 1}^{-j}}) = v_i(t)$.
-$(4)$ follows since $v_i(t) \perp \mathcal{H}_{t - 1}$ and $\forall z \in \mathcal{H}_{t - 1}^j\ (z - \linE{z}{\mathcal{H}_{t - 1}^{-j}}) \in \mathcal{H}_{t - 1}$
+$(2) \implies (4)$: In computing $(y - \linE{y}{\H_{t = 1}^{-j}})$ for $y \in \H_t^i$ it is sufficient to consider $y = x_i(t)$ since $\H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$, in which case $(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$.
+$(4)$ follows since $v_i(t) \perp \H_{t - 1}$ and $\forall z \in \H_{t - 1}^j\ (z - \linE{z}{\H_{t - 1}^{-j}}) \in \H_{t - 1}$
 
@@
 \subsection{Pairwise Granger Causality}
+\label{sec:pwgc}
 Recall that Granger-causality in general must be understood with
@@
 
+This notion is of interest for a variety of reasons.  From a purely
+theoretical or conceptual standpoint, we will seek to understand when
+and how pairwise causality can capture the idea of ``flow of
+information'' in the underlying graph.  It will also be useful to
+reason about what the conditions under which \textit{unobserved}
+components of $x(t)$ may or may not interfere with inference in the
+actually observed components.  Finally, we will seek to construct
+practical estimation procedures based purely on pairwise causality
+tests and understand their properties.
+
 We first need to establish some graph theoretic notation and
-terminology, collected formally in a definition for the reader's
+terminology, collected formally in definitions for the reader's
 convenient reference.
 
-\begin{definition}[Causality graph, Parents, etc...]
+\begin{definition}[Causality graph, Parents]
 
@@
 $\pa{i}$, that is, $j \in \pa{i} \iff x_j \gc x_i$.
+\end{definition}
 
-Even though the definitions of Granger-causality naturally suggests
-that each node in the graph should have a self-loop (unless the node
-is simply independent noise) we will not include nodes in their own
-parent sets: $i \not\in \pa{i} \subseteq \anc{i}$.  However, we do not
-immediately assume that $\gcg$ is a DAG, and so we will allow for $i$
-to be it's own \textit{ancestor}.
-
+\begin{definition}[Grandparents, ancestors]
 The set of level $\ell$ \textit{grandparents} of node $i$, denoted
@@
 $\gcg$ from $j$ to $i$ \textit{or} if $j = i$ and there is a cycle of
-length $n$ starting at $i$ and ending back at $i$, that is, $i$ can be
+length $\ell$ starting at $i$ and ending back at $i$, that is, $i$ can be
 it's own grandparent.  Clearly, $\pa{i} = \gpn{1}{i}$.
@@
 \end{definition}
+
+Even though the definitions of Granger-causality naturally suggests
+that each node in the graph should have a self-loop (unless the node
+is simply independent noise) we will not include nodes in their own
+parent sets: $i \not\in \pa{i} \subseteq \anc{i}$.  However, we do not
+immediately assume that $\gcg$ is a DAG, and so \textit{we will allow for $i$
+to be it's own ancestor}.
 
@@
 (convolution) of this filter compactly by taking $z^{-1}$ as the
-backshift operator when convenient
+backshift operator on $\ell_2^n(\Omega, \mathcal{F}, \mathbb{P})$ when
+convenient
 
@@
 $x_i(t) = \sum_{\tau = 0}^\infty \sum_{j = 1}^n a_{ij}(\tau)v_j(t -
-\tau)$.  The proposition states that $a_{ij}(\tau) = 0$ unless
+\tau)$.  The proposition states that $\A_{ij}(\tau) = 0$ unless
 $j \in \anc{i}$.
@@
 
-  \begin{equation*}
+  \begin{equation}
+    \label{eqn:resolvant_inv}
     (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k.
-  \end{equation*}
+  \end{equation}
 
   Now, let $S$ be a $0-1$ matrix containing the sparsity pattern of
-  $\B(z)$.  Then from lemma (\ref{lem:adj_matrix}) $S$ is the
-  transpose of the adjacency matrix for $\gcg$ and hence $(S^k)_{ij}$
-  is non-zero if and only if $j \in \gpn{k}{i}$, and therefore
-  $\B(z)^k_{ij} = 0$ if $j \not \in \gpn{k}{i}$.
-  $\anc{i} = \cup_k\gpn{k}{i}$ we see that $\A(z)_{ij}$ is zero if
-  $j \not\in \anc{i}$.
+  $\B(z)$.  Then from lemma \ref{lem:adj_matrix} $S$ is the transpose
+  of the adjacency matrix for $\gcg$ and hence $(S^k)_{ij}$ is
+  non-zero if and only if $j \in \gpn{k}{i}$, and therefore
+  $\B(z)^k_{ij} = 0$ if $j \not \in \gpn{k}{i}$.  The Cayley-Hamilton
+  theorem can then be applied to represent the infinite sum of equation
+  (\ref{eqn:resolvant_inv}) in terms only of \textit{finite} powers of
+  $\B(z)$, so the application of lemma \ref{lem:adj_matrix} in the
+  limit is not necessary.  Finally, since $\anc{i} = \cup_{k = 1}^n\gpn{k}{i}$
+  we see that $\A(z)_{ij}$ is zero if $j \not\in \anc{i}$.
 
@@
   \begin{align*}
-    x_i(t) &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
+    x_i(t) &= (I - \B(z))^{-1}\\
+    &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
     &= \A_{ii}(z) v_i(t) + \sum_{j \in \anc{i} \setminus \{i\}} \A_{ij}(z) v_j(t)\\
@@
   \label{lem:subspace_sum_projection}
-  Let $x \in \mathcal{H}$ where $\mathcal{H}$ is a Hilbert space having inner product $\langle \cdot, \cdot \rangle$ and let $\{\mathcal{H}_n\}_{n = 0}^\infty$ be a collection of subspaces of $\mathcal{H}$ such that $x \perp \mathcal{H}_0$.
-Then
+  Let $x \in \H$ where $\H$ is a Hilbert space having inner product $\inner{\cdot}{\cdot}$ and let $\{\H_n\}_{n = 0}^\infty$ be a collection of subspaces of $\H$ such that $x \perp \H_0$.  Then
 
   \begin{equation*}
-    \linE{x}{\bigvee_{n = 0}^\infty\mathcal{H}_n} = \linE{x}{\bigvee_{n = 1}^\infty\mathcal{H}_n},
+    \linE{x}{\bigvee_{n = 0}^\infty\H_n} = \linE{x}{\bigvee_{n = 1}^\infty\H_n},
   \end{equation*}
 
-  where $\mathcal{H}_2 \vee \mathcal{H}_2 = \cl \{\alpha + \beta\ |\ \alpha \in \mathcal{H}_1, \beta \in \mathcal{H}_2 \}$ is the closed sum of subspaces.
+  where $\H_2 \vee \H_2 = \cl \{\alpha + \beta\ |\ \alpha \in \H_1, \beta \in \H_2 \}$ is the closed sum of subspaces.
 \end{lemma}
 \begin{proof}
-  We can apply the Gram-schmidt process to the collection of subspaces to obtain a sequence of orthogonal subspaces $\widetilde{\mathcal{H}}_0, \widetilde{\mathcal{H}}_1, \ldots$ with $\widetilde{\mathcal{H}}_0 = \mathcal{H}_0$, $\widetilde{\mathcal{H}}_n \subseteq \mathcal{H}_n$, and $\bigvee_{n = 0}^\infty\mathcal{H}_n = \bigoplus_{n = 0}^\infty \widetilde{\mathcal{H}}_n$, where $\oplus$ indicates the orthogonal sum of subspaces (which is always closed).
+  We can apply the Gram-schmidt process to the collection of subspaces to obtain a sequence of orthogonal subspaces $\widetilde{\H}_0, \widetilde{\H}_1, \ldots$ with $\widetilde{\H}_0 = \H_0$, $\widetilde{\H}_n \subseteq \H_n$, and $\bigvee_{n = 0}^\infty\H_n = \bigoplus_{n = 0}^\infty \widetilde{\H}_n$, where $\oplus$ indicates the orthogonal sum of subspaces (which is always closed).
 Then, we have an explicit representation of the projection
@@
   \begin{align*}
-    \linE{x}{\bigvee_{n = 0}^\infty\mathcal{H}_n} &= \sum_{n = 0}^\infty \langle x, x_n \rangle x_n\\
-    &\overset{(a)}{=} \sum_{n = 1}^\infty \langle x, x_n \rangle x_n,
+    \linE{x}{\bigvee_{n = 0}^\infty\H_n} &= \sum_{n = 0}^\infty \inner{x}{x_n} x_n\\
+    &\overset{(a)}{=} \sum_{n = 1}^\infty \inner{x}{x_n} x_n,
   \end{align*}
 
-  where $x_n \in \widetilde{\mathcal{H}}_n \subseteq \mathcal{H}_n$ and $(a)$ follows because of our assumption that $x \perp \mathcal{H}_0$.
-This establishes that the projection is infact a member of the smaller subspace $\bigvee_{n = 1}^\infty\mathcal{H}_n$, and since the error can only increase by projecting onto this smaller subspace, we have the conclusion.
+  where $x_n \in \widetilde{\H}_n \subseteq \H_n$ and $(a)$ follows because of our assumption that $x \perp \H_0$.
+This establishes that the projection is infact a member of the smaller subspace $\bigvee_{n = 1}^\infty\H_n$, and since the error can only increase by projecting onto this smaller subspace, we have the conclusion.
 \end{proof}
@@
   \begin{align*}
-    &\langle \alpha - \linE{\alpha}{\mathcal{X}}, \beta - \linE{\beta}{\mathcal{X}} \rangle = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
-    &{\iff} \langle \alpha + x - \linE{\alpha + x}{\mathcal{X}}, \beta - \linE{\beta}{\mathcal{X}} \rangle = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
-    &\overset{(a)}{\iff} \langle \alpha + x, \beta - \linE{\beta}{\mathcal{X}} \rangle = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
+    &\inner{\alpha - \linE{\alpha}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
+    &{\iff} \inner{\alpha + x - \linE{\alpha + x}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
+    &\overset{(a)}{\iff} \inner{\alpha + x}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
     &\overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A} \vee \mathcal{X}} = 0\ \forall \beta \in \mathcal{B}\\
@@
 
-  where $(a)$ follows since $(\beta - \linE{\beta}{\mathcal{X}}) \perp \mathcal{X}$ and $\linE{\alpha + x}{\mathcal{X}} \in \mathcal{X}$, $(b)$ since $z \perp \mathcal{X} \vee \mathcal{A} \iff \linE{z}{\mathcal{X} \vee \mathcal{A}} = 0$, and $(d)$ since $\linE{z}{\mathcal{H}_2} = \linE{\linE{z}{\mathcal{H}_2}}{\mathcal{H}_1}$ for $\mathcal{H}_2 \subseteq \mathcal{H}_1$.
+  where $(a)$ follows since $(\beta - \linE{\beta}{\mathcal{X}}) \perp \mathcal{X}$ and $\linE{\alpha + x}{\mathcal{X}} \in \mathcal{X}$, $(b)$ since $z \perp \mathcal{X} \vee \mathcal{A} \iff \linE{z}{\mathcal{X} \vee \mathcal{A}} = 0$, and $(d)$ since $\linE{z}{\H_2} = \linE{\linE{z}{\H_2}}{\H_1}$ for $\H_2 \subseteq \H_1$.
 
@@
     &+ \sum_{k \in \anc{i}}\E[\big(\A_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{k \in \anc{j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
-    &+ \sum_{k \in \anc{i}}\sum_{\ell \in \anc{j}}\E[\big(\A_{ik}(z)v_k(t)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)]\\
+    &+ \sum_{k \in \anc{i} \setminus i}\sum_{\ell \in \anc{j} \setminus j}\E[\big(\A_{ik}(z)v_k(t)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)]\\
   \end{align*}
@@
 $\A(z) = I + A(1)z^{-1} + \cdots$) so $\A_{ji}(z)v_i(t - \tau)$ is strictly earlier in time, for $\tau \ge 0$, than $v_i(t)$ and hence $\E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)] = 0$.
-This establishes the fact that $H_t^{(i)} \perp \mathcal{H}_t^{(j)}$ where $H_t^{(i)} = \{\alpha x_i(t)\ |\ \alpha \in \R\}$ is the Hilbert space generated by the current sample of $x_i$.
-It is clear that $\mathcal{H}_t^{(i)} = \bigvee_{\tau = 0}^\infty H_{t - \tau}^{(i)}$.
+This establishes the fact that $H_t^{(i)} \perp \H_t^{(j)}$ where $H_t^{(i)} = \{\alpha x_i(t)\ |\ \alpha \in \R\}$ is the Hilbert space generated by the current sample of $x_i$.
+It is clear that $\H_t^{(i)} = \bigvee_{\tau = 0}^\infty H_{t - \tau}^{(i)}$.
 
@@
   \begin{align*}
-    \linE{x_j(t - s)}{\mathcal{H}_t^{(i)}} &= \linE{x_j(t - s)}{\bigvee_{\tau = 0}^\infty H_{t - \tau}^{(i)}}\\
+    \linE{x_j(t - s)}{\H_t^{(i)}} &= \linE{x_j(t - s)}{\bigvee_{\tau = 0}^\infty H_{t - \tau}^{(i)}}\\
     &\overset{(a)}{=} \linE{x_j(t - s)}{\bigvee_{\tau = 1}^\infty H_{t - \tau}^{(i)}}\\
-    &= \linE{x_j(t - s)}{\mathcal{H}_{t - 1}^{(i)}},
+    &= \linE{x_j(t - s)}{\H_{t - 1}^{(i)}},
   \end{align*}
@@
   \begin{equation*}
-    \mathcal{H}_t^{(i)} \perp \mathcal{H}_{t - 1}^{(j)}\ |\ \mathcal{H}_{t - 1}^{(i)},
+    \H_t^{(i)} \perp \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(i)},
   \end{equation*}
@@
 
-\begin{definition}[Simply Causal]
-  If the Granger-Causality graph $\gcg$ for $x(t)$ is as a directed
-  acyclic graph such that there is at most one non-overlapping
-  directed path between any two nodes (excluding self-loops) then we
-  will refer to $\gcg$ as being \textit{simply causal}.
-\end{definition}
+% \begin{definition}[Simply Causal]
+%   If the Granger-Causality graph $\gcg$ for $x(t)$ is as a directed
+%   acyclic graph such that there is at most one non-overlapping
+%   directed path between any two nodes (excluding self-loops) then we
+%   will refer to $\gcg$ as being \textit{simply causal}.
+% \end{definition}
 
-\begin{example}
-  A directed tree (more generally directed forest) is a simply causal
-  graph.
-\end{example}
+% \begin{example}
+%   A directed tree (more generally directed forest) is a simply causal
+%   graph.
+% \end{example}
 
 \begin{lemma}
-  If $\gcg$ is simply causal then $i \pwgc j$ and $j \in \anc{i}$ are
+  If $\gcg$ is a strongly causal DAG then $i \pwgc j$ and $j \in \anc{i}$ are
   \textit{alternatives}, that is $i \pwgc j \implies j \notin \anc{i}$.
@@
 \begin{proof}
-  Suppose that $\gcg$ is simply causal and that we have both
+  Suppose that $\gcg$ is a strongly causal DAG and that we have both
   $i \pwgc j$ and $j \in \anc{i}$.  We will show that $i \in \anc{j}$,
-  which produces a contradiction (since $\gcg$ being a DAG is part of
-  the requirements for being simply causal).
+  which produces a contradiction (since there must then be a directed cycle).
 
@@
   $k$ has paths to both $j,\ \gcgpath{k}{b}{q}{j}$ and $i,\ \gcgpath{k}{c}{r}{i}$.
+
+  Now, since the graph is strongly causal, path $c$ must overlap with
+  path $a$ from $j$ to $i$ otherwise there would be multiple paths
+  from $k$ to $i$.  By the same reasoning with paths from $k$ to $j$,
+  the path $c$ must also overlap with path $b$ from $k$ to $j$, so in
+  fact
 

### 2019-04-09 paper.tex
Pairwise recovery algorithm and proof WIP.

@@
   \label{lem:pwgc_anc}
-  If $\gcg$ is strongly causal then $j \in \anc{i} \implies j \pwgc i$.
+  If $\gcg$ is a strongly causal DAG then $j \in \anc{i} \implies j \pwgc i$.
 \end{proposition}
@@
   \label{thm:tree_recovery}
-  If the Granger-Causality graph $\gcg$ for $x(t)$ is simply causal
-  then then $\gcg$ can be inferred from pairwise causality tests
-  alone.
+  If the Granger-Causality graph $\gcg$ for $x(t)$ is a strongly
+  causal DAG then $\gcg$ can be inferred from pairwise causality
+  tests alone.  The procedure can be carried out, assuming we have an
+  oracle for pairwise causality, via algorithm (\ref{alg:pwgr}).
+
+  \begin{algorithm}
+    \SetKwInOut{Input}{input}
+    \SetKwInOut{Output}{output}
+    \SetKwInOut{Initialize}{initialize}
+
+    \BlankLine
+    \caption{Pairwise Graph Recovery}
+    \label{alg:pwgr}
+    % \TitleOfAlgo{Pairwise Graph Recovery}
+    \Input{Pairwise Granger-causality relations between $n$ processes whose joint Granger-causality relations are known to form a strongly causal DAG $\gcg$.}
+    \Output{Edges $E = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of the graph $\gcg = ([n], E)$.}
+    \Initialize{$S = [n]$  \texttt{\# unprocessed nodes}\\
+      $E = \emptyset$  \texttt{\# Edges of }$\gcg$\\
+      $k = 0$ \texttt{\# A counter used only for notation}}
+    \BlankLine
+    \While{$S \ne \emptyset$}{
+      $P_k \leftarrow \{i \in S_{k - 1}\ |\ \forall s \in S_{k - 1}\ s \not\pwgc i\}$  \texttt{\# first layer driving nodes}\\
+      $S_k \leftarrow S_{k - 1} \setminus P_k$ \texttt{\# remove drivers from} $S$\\
+      $\widehat{P}_{k + 1} \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ s \not\pwgc i\}$   \texttt{\# second layer driving nodes}\\
+      $D_k \leftarrow \{(i, j) \in P_k \times \widehat{P}_{k + 1}\ |\ i \pwgc j \}$  \texttt{\# first to second layer edges}\\
+      $E_k \leftarrow E_{k - 1} \cup D_k$ \texttt{\# add edges to }$E$\\
+      $k \leftarrow k + 1$
+    }
+    \Return{$E$}
+  \end{algorithm}
 \end{theorem}
+
+We prove the theorem by establishing the correctness of algorithm (\ref{alg:pwgr}).
+The idea is to iteratively ``peel away layers'' of nodes by removing the nodes that have no parents remaining, which always exist since the graph is acyclic.
+The additional requirement of strong causality ensures firstly that all actual edges of $\gcg$ manifest in some way as pairwise relations (by proposition \ref{prop:pwgc_anc}), and secondly, as we will see, that all of the direct descendents of nodes in $P_k$ are in $\widehat{P}_{k + 1}$.
 \begin{proof}
+  Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$, we will proceed by induction.
 
+  On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E_0 = \emptyset$.
+The set $P_0$ (``driving nodes'') is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $P_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
+Now, if $P_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
+If $P_0 \ne [n]$ then $\widehat{P}_1$ must also be non-empty, otherwise there would be driving nodes not included in $P_0$.
+We also see that since nodes in $P_0$ have no parents, we must have $\anc{i}\cap\anc{j} = \emptyset$, and therefore by proposition \ref{prop:ancestor_properties} $i \in \anc{j}$;  moreover, the construction of $\widehat{P}_1$ implies that in fact $i \in \pa{j}$, so $(i, j)$ is an edge of $\gcg$.
+Finally, by similar reasoning, there cannot be an edge $(i, j) \in D_0$ which is not in $\gcg$, since $\anc{i} \cap \anc{j} = \emptyset$.
+
+\hl{WIP}
 \end{proof}
@@
 \begin{example}
-  The condition of simple causality is merely a sufficient condition.
-For example, the complete directed graph with 2 nodes (i.e.
-$B(1) = \left[ \begin{array}{cc} 0 & 1 \\ 1 & 0 \end{array} \right]$) contains a loop but is pairwise recoverable.
+  Being a strongly causal DAG is merely a sufficient condition.
+For example, the complete directed graph with 2 nodes (e.g.
+$B(1) = \left[ \begin{array}{cc} 1/2 & 1 \\ 1 & 1/2 \end{array} \right]$) contains a loop but is pairwise recoverable, though not by algorithm (\ref{alg:pwgr}).
+Clearly, this example is somewhat artificial since when $n = 2$ there is no difference between pairwise Granger-causality and joint Granger-causality amongst all series -- however, one can take any strongly causal DAG, and then add two additional nodes having such a loop without impacting the pairwise recoverability.
 \end{example}

### 2019-04-10 paper.tex
Working on algorithm proof -- have met a snag...

@@
 forests), and DAGs where each node has at most one parent.
+
+\begin{lemma}
+  \label{lem:still_strongly_causal}
+  If $\gcg$ is a strongly causal graph, any subgraph formed by eliminating nodes
+  and all of the in and out edges of the eliminated node is still strongly causal.
+\end{lemma}
+\begin{proof}
+  If $\gcg$ has at most one path between any two nodes, there can only
+  be fewer paths after removing nodes from $\gcg$.
+\end{proof}
 
@@
 \begin{proof}
+% 1. Dispense with trivialities
+% 2. Show that P_k is nonempty
+% 3. Show that P^_{k + 1} is non-empty
+% 4. Show that all edges in D_k are legit in G
+% 5. Show that all outgoing edges of P_k are in D_k
+% 6. Induction
+
   Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$, we will proceed by induction.
@@
 We also see that since nodes in $P_0$ have no parents, we must have $\anc{i}\cap\anc{j} = \emptyset$, and therefore by proposition \ref{prop:ancestor_properties} $i \in \anc{j}$;  moreover, the construction of $\widehat{P}_1$ implies that in fact $i \in \pa{j}$, so $(i, j)$ is an edge of $\gcg$.
-Finally, by similar reasoning, there cannot be an edge $(i, j) \in D_0$ which is not in $\gcg$, since $\anc{i} \cap \anc{j} = \emptyset$.
+
+%Finally, by similar reasoning, there cannot be an edge $(i, j) \in D_0$ which is not in $\gcg$, since $\anc{i} \cap \anc{j} = \emptyset$.
+
+Consider now step $K < n$ of algorithm \ref{alg:pwgr}.
+By lemma \ref{lem:still_strongly_causal} the graph remains strongly causal after having removed nodes $\bigcup_{k = 0}^K P_k$ from $S$, and we assume for induction that $\bigcup_{k = 1}^K D_k$ are all edges of $\gcg$.
 

### 2019-04-10 paper.tex
New algorithm, is it correct now?

@@
     \SetKwInOut{Initialize}{initialize}
+    \DontPrintSemicolon
 
@@
     \Output{Edges $E = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of the graph $\gcg = ([n], E)$.}
-    \Initialize{$S = [n]$  \texttt{\# unprocessed nodes}\\
-      $E = \emptyset$  \texttt{\# Edges of }$\gcg$\\
-      $k = 0$ \texttt{\# A counter used only for notation}}
+    \Initialize{$S_0 = [n]$  \texttt{\# unprocessed nodes}\\
+      $E_0 = \emptyset$  \texttt{\# edges of }$\gcg$\\
+      $P_0 = \emptyset$  \texttt{\# layer by layer driving nodes}\\
+      $k = 0$ \texttt{\# a counter used only for notation}}
     \BlankLine
-    \While{$S \ne \emptyset$}{
-      $P_k \leftarrow \{i \in S_{k - 1}\ |\ \forall s \in S_{k - 1}\ s \not\pwgc i\}$  \texttt{\# first layer driving nodes}\\
-      $S_k \leftarrow S_{k - 1} \setminus P_k$ \texttt{\# remove drivers from} $S$\\
-      $\widehat{P}_{k + 1} \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ s \not\pwgc i\}$   \texttt{\# second layer driving nodes}\\
-      $D_k \leftarrow \{(i, j) \in P_k \times \widehat{P}_{k + 1}\ |\ i \pwgc j \}$  \texttt{\# first to second layer edges}\\
+    \While{$S_k \ne \emptyset$}{
+      $\widehat{P}_k \leftarrow \{i \in S_{k - 1}\ |\ \forall s \in S_{k - 1}\ s \not\pwgc i\}$  \texttt{\# parents of current layer}\\
+      $P_k \leftarrow P_{k - 1} \cup \widehat{P}_k$ \texttt{\# Expand parent set}\\
+      $S_k \leftarrow S_{k - 1} \setminus \widehat{P}_k$ \texttt{\# remove from } $S$\\
+
+      \;
+      $C_k \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ s \not\pwgc i\}$   \texttt{\# potential children of }$P_k$\\
+      $D_k \leftarrow \{(i, j) \in P_k \times C_k\ |\ i \pwgc j \}$  \texttt{\# edges incident on }$C_k$\\
       $E_k \leftarrow E_{k - 1} \cup D_k$ \texttt{\# add edges to }$E$\\
+
+      \;
       $k \leftarrow k + 1$
@@
 The idea is to iteratively ``peel away layers'' of nodes by removing the nodes that have no parents remaining, which always exist since the graph is acyclic.
-The additional requirement of strong causality ensures firstly that all actual edges of $\gcg$ manifest in some way as pairwise relations (by proposition \ref{prop:pwgc_anc}), and secondly, as we will see, that all of the direct descendents of nodes in $P_k$ are in $\widehat{P}_{k + 1}$.
+The additional requirement of strong causality ensures firstly that all actual edges of $\gcg$ manifest in some way as pairwise relations (by proposition \ref{prop:pwgc_anc}), and secondly, as we will see, that parents of $C_k$ are in $P_k$ and can be determined by a pairwise test..
 \begin{proof}
 % 1. Dispense with trivialities
-% 2. Show that P_k is nonempty
-% 3. Show that P^_{k + 1} is non-empty
-% 4. Show that all edges in D_k are legit in G
-% 5. Show that all outgoing edges of P_k are in D_k
+% 2. Show that hat{P}_k is nonempty
+% 3. Show that C_k is nonempty
+% 4. Show that edges in D_k are actual edges in G
+% 5. show that each edge incident on C_k is in D_k
 % 6. Induction

### 2019-04-11 paper.tex
Some slight modifications...

@@
 The idea is to iteratively ``peel away layers'' of nodes by removing the nodes that have no parents remaining, which always exist since the graph is acyclic.
-The additional requirement of strong causality ensures firstly that all actual edges of $\gcg$ manifest in some way as pairwise relations (by proposition \ref{prop:pwgc_anc}), and secondly, as we will see, that parents of $C_k$ are in $P_k$ and can be determined by a pairwise test..
+The additional requirement of strong causality ensures firstly that all actual edges of $\gcg$ manifest in some way as pairwise relations (by proposition \ref{prop:pwgc_anc}), and secondly, as we will see, that parents of $C_k$ are in $P_k$ and can be determined by a pairwise test.
 \begin{proof}
@@
 
-  Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$, we will proceed by induction.
+  Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.
+Suppose then that $n > 1$, we will proceed by induction and show that on each step, every edge incident upon a node in $C_k$ is contained in $D_k$, that every edge in $D_k$ is an edge in $\gcg$, and finally that $\bigcup_{k = 1}^n C_k = [n] \setminus \widehat{P}_0$ and that this implies $\big([n], \bigcup_{k = 1}^n D_k\big) = \gcg$.
 
   On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E_0 = \emptyset$.
-The set $P_0$ (``driving nodes'') is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $P_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
-Now, if $P_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
-If $P_0 \ne [n]$ then $\widehat{P}_1$ must also be non-empty, otherwise there would be driving nodes not included in $P_0$.
-We also see that since nodes in $P_0$ have no parents, we must have $\anc{i}\cap\anc{j} = \emptyset$, and therefore by proposition \ref{prop:ancestor_properties} $i \in \anc{j}$;  moreover, the construction of $\widehat{P}_1$ implies that in fact $i \in \pa{j}$, so $(i, j)$ is an edge of $\gcg$.
+The set $\widehat{P}_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $\widehat{P}_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
+Now, if $\widehat{P}_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
+On the other hand, if $\widehat{P}_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $\widehat{P}_0$.
+Consider now an edge $(i, j) \in D_0$ (again $D_0$ is non-empty for the same reason as $C_0$).
+Since $i$ has no parents, we must have $\anc{i}\cap\anc{j} = \emptyset$ and therefore $i \in \anc{j}$, and due to the construction of $C_0$ we in fact have $i \in \pa{j}$, which establishes that every edge in $D_0$ is also an edge in $\gcg$.
+Finally consider $j \in C_0$, by proposition \ref{prop:pwgc_anc} and since the parents of $C_0$ must be in $P_0$, we see that every edge incident on $j$ must be present in $D_0$.
+
+% We also see that since nodes in $\widehat{P}_0$ have no parents,
+
+% we must have $\anc{i}\cap\anc{j} = \emptyset$, and therefore by proposition \ref{prop:ancestor_properties} $i \in \anc{j}$;  moreover, the construction of $\widehat{P}_1$ implies that in fact $i \in \pa{j}$, so $(i, j)$ is an edge of $\gcg$.
 

### 2019-04-12 paper.tex
Some cleanup.

@@
     \item{$\H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
-    \item{\hl{Wold $A(\tau)$ condition}}
-    \item{Tie in the 0s of $S(\omega)^{-1}$, though this isn't immediately directed.}
+    \item{$\linE{x_i(t)}{\H_{t - 1}^{-j}} = \linE{x_i(t)}{\H_{t - 1}}$}
+    % \item{\hl{Wold $A(\tau)$ condition}}
+    % \item{Tie in the 0s of $S(\omega)^{-1}$, though this isn't immediately directed.}
   \end{enumerate}
@@
 
+Statement (4) means that
+$(y - \linE{y}{\H_{t - 1}^{-j}}) \perp (z - \linE{z}{\H_{t -
+    1}^{-j}})\ \forall y \in \H_t^i,\ z \in \H_{t - 1}^j$, which is an
+appealing geometric statement that there is no direct feedback from
+$x_j$ to $x_i$ \cite{lindquist}.
+
+\begin{proof}
   The equivalence $(1) \iff (3)$ is essentially a restatement of the
   definition, but is in line with the seminal work of Geweke
-\cite{geweke1982measurement}, \cite{geweke1984}.  Statement (4) means that
-$(y - \linE{y}{\H_{t - 1}^{-j}}) \perp (z -
-\linE{z}{\H_{t - 1}^{-j}})\ \forall y \in \H_t^i,\ z \in
-\H_{t - 1}^j$, which is an appealing geometric statement that
-there is no direct feedback from $x_j$ to $x_i$ \cite{lindquist}.
-
-\begin{proof}
-$(1) \implies (2)$: The projection for $x_i(t)$ onto $\H_{t - 1}$ is (by definition) given by a form similar to equation \ref{eqn:ar_representation}:
+  \cite{geweke1982measurement}, \cite{geweke1984}.
+
+  The equivalence $(3) \iff (5)$ is also immediate from the
+  definitions i.e. the projections are equal if and only if the errors
+  are equal.
+
+  $(1) \implies (2)$: The projection for $x_i(t)$ onto $\H_{t - 1}$ is
+  (by definition) given by a form similar to equation
+  \ref{eqn:ar_representation}:
 
@@
 
- Since $v(t)$ in equation \ref{eqn:ar_representation} is temporally uncorrelated, it follows that the optimal projection is given by the model coefficients themselves.
-This holds similarly for the projection onto $\H_{t - 1}^{-j}$.
-Then by $(1)$ we have
+Since $v(t)$ in equation \ref{eqn:ar_representation} is temporally
+uncorrelated, it follows that the optimal projection is given by the
+model coefficients themselves.  This holds similarly for the
+projection onto $\H_{t - 1}^{-j}$.  Then by $(1)$ we have
 
@@
 
-By the uniqueness of the projection we must have $\forall \tau\  B_{i, j}(\tau) = 0$.
-
-$(2) \implies (4)$: In computing $(y - \linE{y}{\H_{t = 1}^{-j}})$ for $y \in \H_t^i$ it is sufficient to consider $y = x_i(t)$ since $\H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$, in which case $(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$.
-$(4)$ follows since $v_i(t) \perp \H_{t - 1}$ and $\forall z \in \H_{t - 1}^j\ (z - \linE{z}{\H_{t - 1}^{-j}}) \in \H_{t - 1}$
+By the uniqueness of the projection we must have
+$\forall \tau\ B_{i, j}(\tau) = 0$.
+
+$(2) \implies (4)$: In computing $(y - \linE{y}{\H_{t = 1}^{-j}})$ for
+$y \in \H_t^i$ it is sufficient to consider $y = x_i(t)$ since
+$\H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$, in which case
+$(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$.  $(4)$ follows
+since $v_i(t) \perp \H_{t - 1}$ and
+$\forall z \in \H_{t - 1}^j\ (z - \linE{z}{\H_{t - 1}^{-j}}) \in \H_{t
+  - 1}$
+
+\hl{TODO: $(4) \implies (1)$}
 
@@
 
-The following propositions justify various modifications of $x(t)$
-applied in practice to massage $x(t)$ into a form amenable to more
-standard tools.
-
-\begin{theorem}[General Invariance]
-  \hl{Under what conditions is this actually true?}
-
-  Let $\zeta(t)$ be a stationary discrete time stochastic process.  Let
-  $F, G$ be (possibly nonlinear and time varying) invertible filtering
-  operations and $f(t)$, $g(t)$ be perfectly predictable and such that
-  $x_j(t) \defeq F(\zeta_j - f)(t), x_i(t) \defeq G(\zeta_i - g)(t)$ are W.S.S.  Then,
-
-\begin{equation}
-    x_j \gc x_i \iff \zeta_j \te \zeta_i
-  \end{equation}
-\end{theorem}
-
-\begin{corollary}[Invariance Under Invertible and Deterministic Modifications]
-  Let $F(z), G(z)$ describe univariate and invertible
+% The following propositions justify various modifications of $x(t)$
+% applied in practice to massage $x(t)$ into a form amenable to more
+% standard tools.
+
+% \begin{theorem}[General Invariance]
+%   \hl{Under what conditions is this actually true?}
+
+%   Let $\zeta(t)$ be a stationary discrete time stochastic process.  Let
+%   $F, G$ be (possibly nonlinear and time varying) invertible filtering
+%   operations and $f(t)$, $g(t)$ be perfectly predictable and such that
+%   $x_j(t) \defeq F(\zeta_j - f)(t), x_i(t) \defeq G(\zeta_i - g)(t)$ are W.S.S.  Then,
+
+% \begin{equation}
+%     x_j \gc x_i \iff \zeta_j \te \zeta_i
+%   \end{equation}
+% \end{theorem}
+
+\begin{proposition}[Invariance Under Invertible and Deterministic Modifications]
+  Let $x(t)$ be a WSS process with absolutely summable covariance
+  sequence, and spectral density uniformly bounded above and below.  Let
+  $F(z), G$ represent univariate and invertible
   linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
@@
   \end{equation}
-\end{corollary}
+
+\begin{proof}
+  \hl{TODO}
+\end{proof}
+\end{proposition}
 
@@
 \begin{equation}
-  \label{eqn:parent_expanion}
+  \label{eqn:parent_expansion}
   x_i(t) = v_i(t) + \B_{ii}(z)x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t).
@@
   \begin{equation}
-    \label{eqn:parent_expansion}
     x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t).
@@
   \begin{equation}
+    \label{eqn:ancestor_expansion}
     x_i(t) = \A_{ii}(z)v_i(t) + \sum_{k \in \anc{i} \setminus\{i\}}\A_{ik}(z)v_k(t),
@@
   \label{prop:ancestor_properties}
+\hl{Need to refine this to exclude $k \rightarrow i \rightarrow j$ from 2.}
+
   In a Granger-causality graph $\gcg$, if $j \pwgc i$ then at least
@@
 
+  \hl{TODO: Separate $H_t^i \perp \H_t^j \implies \H_t^i \perp \H_{t - 1}^j\ |\ \H_{t - 1}^i$ into a lemma and consolidate it with lemma 2, 3.}
+
   Then for every $s \ge 0$
@@
 \begin{example}
+  \label{ex:diamond_cancellation}
   Firstly, on $n = 4$ nodes, ``diamond'' shapes can lead to cancellation on paths of length 2:
@@
 
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{0.45\textwidth}
+    \caption{Graph Corresponding to Example \ref{ex:diamond_cancellation}}
+    \label{fig:diamond_cancellation}
+    \includegraphics[width=\linewidth]{hand_drawn_example1.jpg}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \caption{Graph Corresponding to Example \ref{ex:lag_cancellation}}
+    \label{fig:lag_cancellation}
+    \includegraphics[width=\linewidth]{hand_drawn_example2.jpg}
+  \end{subfigure}
+\end{figure}
+
 \begin{example}
-  A second example on $n = 3$ nodes is also worth examining, in this
+  \label{ex:lag_cancellation}
+  A second example on $n = 3$ nodes is also worth examining, in this case
   cancellation is a result of differing time lags:
@@
   We will say that a Granger-causality graph $\gcg$ is \textit{strongly causal}
-if $\gcg$ is a DAG and there is at most 1 directed path between any two nodes.
+if there is at most 1 directed path between any two nodes.
+
+\hl{TODO: Should DAG be included in this definition?  I think it is
+  not necessary for some theorems.}
 \end{definition}
@@
 
-\begin{proposition}
+\begin{lemma}
   \label{prop:sc_graph_common_anc}
   In a strongly causal graph if $j \in \anc{i}$ and $k \in \anc{i} \cap \anc{j}$ then the unique path from $k$ to $i$ contains $j$.
-\end{proposition}
+\end{lemma}
 \begin{proof}
@@
 \begin{proof}
-  Since strongly causal graphs are acyclic, this allows us to
-  recursively expand the parents of $x_i$ in $\gcg$ in the
-  $\VAR(\infty)$ representation of $x(t)$ since we will eventually
-  terminate at nodes $d$ such that $\pa{d} = \emptyset$.  To this end,
-  examining $x_i$ in equation \ref{eqn:ar_representation}
-
-  \begin{align*}
-    x_i(t) &= v_i(t) + \sum_{\tau = 1}^\infty \big[ B_{ii}(\tau)x_i(t - \tau) + \sum_{k \in \pa{i}} B_{ik}(\tau)x_k(t - \tau)\big]\\
-    \implies x_i(t) &= (1 - \B_{ii}(z))^{-1} \big[v_i(t) + \sum_{k \in \pa{i}} B_{ik}(\tau)x_k(t - \tau) \big]\\
-    &= \sum_{\tau = 1}^\infty \big[C_{ii}(\tau)v_i(t - \tau) + \sum_{k \in \pa{i}} C_{ik}(\tau)x_k(t - \tau)\big].
-  \end{align*}
-
-  If $\pa{i} = \emptyset$ then we simply have
-  $x_i(t) = \sum_{\tau = 1}^\infty C_{ii}(\tau)v_i(t)$ and
-  $C_{ii} = A_{ii}$ from equation \ref{eqn:wold}.
-
-  Applying this relation recursively we can write
+  Using the representation of equation (\ref{eqn:parent_expansion}) we can represent $x_i(t)$ as
 
   \begin{equation*}
-    x_i(t) = \sum_{\tau = 1}^\infty\big[C_{ii}(\tau)v_i(t - \tau) + \sum_{k \in \anc{i}}C_{ik}(\tau)v_k(t - \tau) \big].
+    x_i(t) = v_i(t) + \B_{ii}(z) x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z) x_k(t).
   \end{equation*}
 
+  In order to show that $j \pwgc i$ we can show (due to theorem (\ref{thm:granger_causality_equivalences})) that
+
+  \begin{equation*}
+    \linE{x_i(t)}{\H_{t - 1}^{(i)}} = \linE{x_i(t)}{\H_{t - 1}^{(i, j)}},
+  \end{equation*}
+
+\hl{WIP}
+
+% this allows us to recursively expand the
+%   parents of $x_i$ in $\gcg$ in the $\VAR(\infty)$ representation of
+%   $x(t)$ since we will eventually terminate at nodes $d$ such that
+%   $\pa{d} = \emptyset$.  To this end, examining $x_i$ in equation
+%   \ref{eqn:ar_representation}
+
+%   \begin{align*}
+%     x_i(t) &= v_i(t) + \sum_{\tau = 1}^\infty \big[ B_{ii}(\tau)x_i(t - \tau) + \sum_{k \in \pa{i}} B_{ik}(\tau)x_k(t - \tau)\big]\\
+%     \implies x_i(t) &= (1 - \B_{ii}(z))^{-1} \big[v_i(t) + \sum_{k \in \pa{i}} B_{ik}(\tau)x_k(t - \tau) \big]\\
+%     &= \sum_{\tau = 1}^\infty \big[C_{ii}(\tau)v_i(t - \tau) + \sum_{k \in \pa{i}} C_{ik}(\tau)x_k(t - \tau)\big].
+%   \end{align*}
+
+%   If $\pa{i} = \emptyset$ then we simply have
+%   $x_i(t) = \sum_{\tau = 1}^\infty C_{ii}(\tau)v_i(t)$ and
+%   $C_{ii} = A_{ii}$ from equation \ref{eqn:wold}.
+
+%   Applying this relation recursively we can write
+
+%   \begin{equation*}
+%     x_i(t) = \sum_{\tau = 1}^\infty\big[C_{ii}(\tau)v_i(t - \tau) + \sum_{k \in \anc{i}}C_{ik}(\tau)v_k(t - \tau) \big].
+%   \end{equation*}
+
 \end{proof}
@@
 That strong conditions need to be placed on the topology of $\gcg$ in
-order for lemma \ref{lem:pwgc_anc} to be true can be seen through the
+order for proposition \ref{lem:pwgc_anc} to be true can be seen through the
 earlier examples.
@@
 \begin{proof}
+  \hl{This is certainly true, but I don't think I need to use it anywhere.}
 \end{proof}
@@
 % and $d(j, i) \defeq \text{min}\{B_{i, j}(\tau) \ne 0\}$ if $j \in \pa{i}$.
-
-\begin{lemma}
-  \label{lem:pwgc_anc}
-  If $x_j \pwgc x_i$ then one or more of the following hold
-
-  \begin{enumerate}
-    \item{$j \in \anc{i}$}
-    \item{$\anc{i}\cap\anc{j} \ne \emptyset$.}
-  \end{enumerate}
-\end{lemma}
-
-% \begin{remark}
-%   This lemma formalizes the intuition of \textit{information flow} in $\gcg$.
-We have $x_j \pwgc x_i$ whenever $i$ is descended from $j$, or if they have a common ancestor (call it $k$) and ``information'' from $k$ reaches $j$ before it reaches $i$.
-% \end{remark}
-
-\begin{proof}
-  Is this true if we take \ref{lem:pwgc_anc} as an axiom?
-\end{proof}
 
@@
   the path $c$ must also overlap with path $b$ from $k$ to $j$, so in
-  fact
+  fact path $c$ from $k$ to $i$ is
+  $c = b \rightarrow j \rightarrow a$.
+
+  \hl{TODO: We need the strengthening of prop }(\ref{prop:ancestor_properties})\hl{in order to conclude.}
 
 \end{proof}
+
+\subsection{Recovering $\gcg$ via Pairwise Tests}
+In this section we will show that if $\gcg$ is a strongly causal DAG, then it is possible to recover $\gcg$ via pairwise tests alone.
+This is an interesting fact in and of itself, but also has some implications for applications since pairwise testing is trivially parallelizable.
+In section \ref{sec:structure_learning} we will also analyze the use of pairwise testing as a heuristic for general graphs.
 
@@
     \Initialize{$S_0 = [n]$  \texttt{\# unprocessed nodes}\\
-      $E_0 = \emptyset$  \texttt{\# edges of }$\gcg$\\
-      $P_0 = \emptyset$  \texttt{\# layer by layer driving nodes}\\
+      $E = \emptyset$  \texttt{\# edges of }$\gcg$\\
+      % $P_0 = \emptyset$  \texttt{\# layer by layer driving nodes}\\
       $k = 0$ \texttt{\# a counter used only for notation}}
@@
       $\widehat{P}_k \leftarrow \{i \in S_{k - 1}\ |\ \forall s \in S_{k - 1}\ s \not\pwgc i\}$  \texttt{\# parents of current layer}\\
-      $P_k \leftarrow P_{k - 1} \cup \widehat{P}_k$ \texttt{\# Expand parent set}\\
-      $S_k \leftarrow S_{k - 1} \setminus \widehat{P}_k$ \texttt{\# remove from } $S$\\
+      % $P_k \leftarrow P_{k - 1} \cup \widehat{P}_k$ \texttt{\# Expand parent set}\\
+      $S_{k + 1} \leftarrow S_k \setminus \widehat{P}_k$ \texttt{\# remove from } $S$\\
 
       \;
-      $C_k \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ s \not\pwgc i\}$   \texttt{\# potential children of }$P_k$\\
-      $D_k \leftarrow \{(i, j) \in P_k \times C_k\ |\ i \pwgc j \}$  \texttt{\# edges incident on }$C_k$\\
-      $E_k \leftarrow E_{k - 1} \cup D_k$ \texttt{\# add edges to }$E$\\
+      $C_k \leftarrow \{i \in S_{k + 1}\ |\ \forall s \in S_{k + 1}\ s \not\pwgc i\}$   \texttt{\# potential children of }$P_k$\\
+
+      \For{$r = 0, \ldots, k$}
+      {
+        $D_{kr} \leftarrow \{(i, j) \in \widehat{P}_{k - r} \times C_k\ |\ i \pwgc j,\ \text{no } i \rightarrow j \text{ path in } E \}$\\%  \texttt{\# edges incident on }$C_k$\\
+        $E \leftarrow E \cup D_{kr}$ \texttt{\# add edges to }$E$\\
+      }
 
@@
 The idea is to iteratively ``peel away layers'' of nodes by removing the nodes that have no parents remaining, which always exist since the graph is acyclic.
-The additional requirement of strong causality ensures firstly that all actual edges of $\gcg$ manifest in some way as pairwise relations (by proposition \ref{prop:pwgc_anc}), and secondly, as we will see, that parents of $C_k$ are in $P_k$ and can be determined by a pairwise test.
+The additional requirement of strong causality ensures firstly that all actual edges of $\gcg$ manifest in some way as pairwise relations (by proposition \ref{prop:pwgc_anc}), and secondly, as we will see, that parents of $C_k$ are in $\bigcup_{r = 0}^k\widehat{P}_{k - r}$ and can be determined by a pairwise test.
+
+\begin{example}
+  That we need to proceed backwards through $\widehat{P}_{k - r}$ can be seen from the example of figure \ref{fig:example_fig3}, where if $D_k = \{(i, j) \in \Big(\bigcup_{r = 0}^k \widehat{P}_{k - r}\Big) \times C_k\ |\ i \pwgc j \}$ above we would infer the false positive edge $2 \rightarrow 3$.
+Moreover, the same example shows that if simply $D_k = \{(i, j) \in \widehat{P}_k \times C_k\ |\ i \pwgc j \}$, then the edge $1 \rightarrow 3$ would be missed.
+
+\hl{TODO: Inkscape figure}
+
+  \begin{figure}
+    \centering
+    \label{fig:example_fig3}
+    \caption{Example graph for Algorithm \ref{alg:pwgr}}
+    \includegraphics[width=0.5\linewidth]{hand_drawn_example3.jpg}
+  \end{figure}
+\end{example}
+
 \begin{proof}
@@
 
-  Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.
-Suppose then that $n > 1$, we will proceed by induction and show that on each step, every edge incident upon a node in $C_k$ is contained in $D_k$, that every edge in $D_k$ is an edge in $\gcg$, and finally that $\bigcup_{k = 1}^n C_k = [n] \setminus \widehat{P}_0$ and that this implies $\big([n], \bigcup_{k = 1}^n D_k\big) = \gcg$.
-
-  On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E_0 = \emptyset$.
+  Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$, we will proceed by induction.
+
+On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
+The set $\widehat{P}_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $\widehat{P}_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
+Now, if $\widehat{P}_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
+On the other hand, if $\widehat{P}_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $\widehat{P}_0$.
+Consider now an edge $(i, j) \in D_0$ ($D_0$ is non-empty for the same reason as $C_0$).
+Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).
+Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{lem:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in \widehat{P}_0 \implies u \in S_0 \implies j \not\in C_0$, a contradiction.
+
+% and show that on each step, every edge incident upon a node in $C_k$ is contained in $D_k$, that every edge in $D_k$ is an edge in $\gcg$, and finally that $\bigcup_{k = 1}^n C_k = [n] \setminus \widehat{P}_0$ and that this implies $\big([n], \bigcup_{k = 1}^n D_k\big) = \gcg$.
+
+%  On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E_0 = \emptyset$.
 The set $\widehat{P}_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $\widehat{P}_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
@@
 
-Consider now step $K < n$ of algorithm \ref{alg:pwgr}.
+%Consider now step $K < n$ of algorithm \ref{alg:pwgr}.
 By lemma \ref{lem:still_strongly_causal} the graph remains strongly causal after having removed nodes $\bigcup_{k = 0}^K P_k$ from $S$, and we assume for induction that $\bigcup_{k = 1}^K D_k$ are all edges of $\gcg$.
@@
 \section{Structure Learning}
+\label{sec:structure_learning}
 \subsection{Local Search Heuristics}

### 2019-04-26 paper.tex
Proof (I think) of prop 4.

@@
 \newcommand{\gcgpath}[4]{#1 \rightarrow #2_1 \rightarrow \cdots \rightarrow #2_{#3} \rightarrow #4}
+\newcommand{\wtalpha}[2]{\widetilde{\alpha}(#1, #2)}  % Some notation for lem:pwgc_anc
 
@@
 this is not the case in general, as different paths through $\gcg$ can
-lead to cancellation.
+lead to cancellation (see example \ref{ex:diamond_cancellation}).  In
+fact, we do not even have $j \in \pa{i} \implies j \pwgc i$ (see
+example \ref{ex:lag_cancellation}).
 
@@
   A second example on $n = 3$ nodes is also worth examining, in this case
-  cancellation is a result of differing time lags:
+  cancellation is a result of differing time lags.
 
@@
 \begin{proof}
-  Using the representation of equation (\ref{eqn:parent_expansion}) we can represent $x_i(t)$ as
-
-  \begin{equation*}
-    x_i(t) = v_i(t) + \B_{ii}(z) x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z) x_k(t).
-  \end{equation*}
-
-  In order to show that $j \pwgc i$ we can show (due to theorem (\ref{thm:granger_causality_equivalences})) that
-
-  \begin{equation*}
-    \linE{x_i(t)}{\H_{t - 1}^{(i)}} = \linE{x_i(t)}{\H_{t - 1}^{(i, j)}},
-  \end{equation*}
-
-\hl{WIP}
+  We will show that for some $\psi \in \H_{t - 1}^{(j)}$ we have $\inner{\psi}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} \ne 0$ and therefore that $H_t^{(i)} \not\perp\ \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(i)}$, which by theorem (\ref{thm:granger_causality_equivalences}) and \hl{lemma [name this lemma]} is enough to establish that $j \pwgc i$.
+
+    Firstly, we will establish a representation of $x_i(t)$ that involves $x_j(t)$.
+Denote by $a_{r + 1} \rightarrow a_r \rightarrow \cdots \rightarrow a_1 \rightarrow a_0$ with $a_{r + 1} \defeq j$ and $a_0 \defeq i$ the unique $j \rightarrow i$ path in $\gcg$, we will expand the representation of equation (\ref{eqn:parent_expansion}) backwards along this path:
+
+\begin{align*}
+  x_i(t) &= v_i(t) + \B_{ii}(z) x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z) x_k(t)\\
+         &= \underbrace{v_i(t) + \B_{ii}(z) x_i(t) + \sum_{\substack{k \in \pa{i} \\ k \ne a_1}}\B_{ik}(z) x_k(t)}_{\defeq \wtalpha{a_0}{a_1}} + \B_{a_0a_1}(z)x_{a_1}(t)\\
+         &= \wtalpha{a_0}{a_1} + \B_{a_0a_1}(z)\big[\wtalpha{a_1}{a_2} + \B_{a_1a_2}(z)x_{a_2}(t) \big]\\
+         &= \cdots\\
+         &= \sum_{\ell = 0}^r \underbrace{\Big(\prod_{m = 0}^{\ell - 1} \B_{a_m a_{m + 1}}(z) \Big)}_{\defeq F_\ell(z)} \wtalpha{a_\ell}{a_{\ell + 1}} + \Big(\prod_{m = 0}^{r}\B_{a_m a_{m + 1}}(z)\Big)x_j(t)\\
+         &= \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}} + F_r(z) x_j(t)
+\end{align*}
+
+where we have defined $\prod_{m = 0}^{-1} \bullet \defeq 1$ for notational convenience.
+
+Then,
+
+\begin{align*}
+  \inner{\psi}{x_i(t) - \linE{x_i(t)}{&\H_{t - 1}^{(i)}}} = \inner{\psi}{F_r(z) x_j(t) - \linE{F_r(z) x_j(t)}{\H_{t - 1}^{(i)}}}\\ &+ \sum_{\ell = 0}^r \inner{\psi}{F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}} - \linE{F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}}}{\H_{t - 1}^{(i)}}}\\
+  &\overset{(a)}{=} \inner{\psi}{F_r(z) x_j(t) - \linE{F_r(z) x_j(t)}{\H_{t - 1}^{(i)}}}\\ &-\sum_{\ell = 0}^r \inner{\psi}{\linE{F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}}}{\H_{t - 1}^{(i)}}}\\
+  &= \inner{\psi}{F_r(z)x_j(t)} - \inner{\psi}{\linE{F_r(z)x_j(t) + \sum_{\ell = 0}^r F_{\ell}(z)\wtalpha{a_\ell}{a_{\ell + 1}}}{\H_{t - 1}^{(i)}}}
+\end{align*}
+
+where $(a)$ follows since $\forall \psi \in \H_{t - 1}^{(j)}\ \inner{\psi}{\wtalpha{a_\ell}{a_{\ell + 1}}} = 0$ which follows since for every $\ell = 1, \ldots, r$ we have $a_{\ell} \not\in \anc{j}, j \not\in \anc{a_{\ell}}$ and \hl{an application of lemmas [name these]}.
+Continuing from the above we will choose the particular value $\psi = F_r(z)x_j(t)$, then
+
+\begin{align*}
+  &\inner{F_r(z)x_j(t)}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} = 0\\
+  &\iff \E|F_r(z)x_j(t)|^2 = \inner{F_r(z)x_j(t)}{\linE{F_r(z)x_j(t) + \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}}}{\H_{t - 1}^{(i)}}}\\
+  &\overset{(b)}{\iff} F_r(z)x_j(t) = \linE{F_r(z)x_j(t) + \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}}}{\H_{t - 1}^{(i)}},
+\end{align*}
+
+where $(b)$ follows by the condition for equality in the Cauchy-Schwarz inequality.
+But since $F_r(z)x_j(t) \not \in \H_{t - 1}^{(i)}$, we must have $\inner{\psi}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} \ne 0$ and therefore $j \pwgc i$.
 

### 2019-04-28 paper.tex
Proof of prop 4 and more work on algorithm proof.

--- 2019-04-12 paper.tex
@@
 \newcommand{\gcgpath}[4]{#1 \rightarrow #2_1 \rightarrow \cdots \rightarrow #2_{#3} \rightarrow #4}
+\newcommand{\wtalpha}[2]{\widetilde{\alpha}(#1, #2)}  % Some notation for lem:pwgc_anc
+\newcommand{\dist}[2]{\mathsf{d}(#1, #2)}  % Distance between things
 
@@
 \begin{proposition}
-  \label{lem:pwgc_anc}
+  \label{prop:pwgc_anc}
   If $\gcg$ is a strongly causal DAG then $j \in \anc{i} \implies j \pwgc i$.
@@
 \begin{proof}
-  Using the representation of equation (\ref{eqn:parent_expansion}) we can represent $x_i(t)$ as
+  We will show that for some $\psi \in \H_{t - 1}^{(j)}$ we have
+
+  \begin{equation}
+    \label{eqn:cond_ortho_proof}
+    \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} \ne 0
+  \end{equation}
+
+  and therefore that $H_t^{(i)} \not\perp\ \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(i)}$, which by theorem (\ref{thm:granger_causality_equivalences}) and \hl{lemma [name the lemma]} is enough to establish that $j \pwgc i$.
+
+  Firstly, we will establish a representation of $x_i(t)$ that involves $x_j(t)$.
+Denote by $a_{r + 1} \rightarrow a_r \rightarrow \cdots \rightarrow a_1 \rightarrow a_0$ with $a_{r + 1} \defeq j$ and $a_0 \defeq i$ the \textit{unique} $j \rightarrow i$ path in $\gcg$, we will expand the representation of equation (\ref{eqn:parent_expansion}) backwards along this path:
+
+  \begin{align*}
+    x_i(t) &= v_i(t) + \B_{ii}(z) x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z) x_k(t)\\
+           &= \underbrace{v_{a_0}(t) + \B_{a_0a_0}(z) x_i(t) + \sum_{\substack{k \in \pa{a_0} \\ k \ne a_1}}\B_{a_0 k}(z) x_k(t)}_{\defeq \wtalpha{a_0}{a_1}} + \B_{a_0a_1}(z)x_{a_1}(t)\\
+           &= \wtalpha{a_0}{a_1} + \B_{a_0a_1}(z)\big[\wtalpha{a_1}{a_2} + \B_{a_1a_2}(z)x_{a_2}(t) \big]\\
+           &\overset{(a)}{=} \cdots\\
+           &= \sum_{\ell = 0}^r \underbrace{\Big(\prod_{m = 0}^{\ell - 1} \B_{a_m a_{m + 1}}(z) \Big)}_{\defeq F_\ell(z)} \wtalpha{a_\ell}{a_{\ell + 1}} + \Big(\prod_{m = 0}^{r}\B_{a_m a_{m + 1}}(z)\Big)x_{a_{r + 1}}(t)\\
+           &= \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}} + F_{r + 1}(z) x_j(t)
+  \end{align*}
+
+  where $(a)$ follows by a routine induction argument and where we define $\prod_{m = 0}^{-1} \bullet \defeq 1$ for notational convenience.
+
+  Using this representation to expand equation (\ref{eqn:cond_ortho_proof}), we obtain the following cumbersome expression:
+
+  \begin{align*}
+    &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}}\\
+    &- \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\linE{\sum_{\ell = 0}^r F_\ell(z)\wtalpha{a_\ell}{a_{\ell + 1}}}{\H_{t - 1}^{(i)}}}\\
+    &+ \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\sum_{\ell = 0}^r F_\ell(z)\wtalpha{a_\ell}{a_{\ell + 1}}}.
+  \end{align*}
+
+  Note that by the orthogonality principle, $\psi - \linE{\psi}{\H_{t - 1}^{(i)}} \perp \H_{t - 1}^{(i)}$, the middle term above is $0$.
+Choosing now the particular value $\psi = F_{r + 1}(z)x_j(t) \in \H_{t - 1}^{(j)}$ we arrive at
+
+  \begin{align*}
+    &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}\\
+    &= \E|F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}|^2\\
+    &+ \inner{F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}}{\sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}}},
+  \end{align*}
+
+  which by the Cauchy-Schwarz inequality is $0$ if and only if
 
   \begin{equation*}
-    x_i(t) = v_i(t) + \B_{ii}(z) x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z) x_k(t).
+    \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}} \overset{\text{a.s.}}{=} \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}} - F_{r + 1}(z)x_j(t),
   \end{equation*}
 
-  In order to show that $j \pwgc i$ we can show (due to theorem (\ref{thm:granger_causality_equivalences})) that
+  or by rearranging and applying the representation obtained earlier, if and only if
 
   \begin{equation*}
-    \linE{x_i(t)}{\H_{t - 1}^{(i)}} = \linE{x_i(t)}{\H_{t - 1}^{(i, j)}},
+    x_i(t) \overset{\text{a.s.}}{=} \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}},
   \end{equation*}
 
-\hl{WIP}
-
-% this allows us to recursively expand the
-%   parents of $x_i$ in $\gcg$ in the $\VAR(\infty)$ representation of
-%   $x(t)$ since we will eventually terminate at nodes $d$ such that
-%   $\pa{d} = \emptyset$.  To this end, examining $x_i$ in equation
-%   \ref{eqn:ar_representation}
-
-%   \begin{align*}
-%     x_i(t) &= v_i(t) + \sum_{\tau = 1}^\infty \big[ B_{ii}(\tau)x_i(t - \tau) + \sum_{k \in \pa{i}} B_{ik}(\tau)x_k(t - \tau)\big]\\
-%     \implies x_i(t) &= (1 - \B_{ii}(z))^{-1} \big[v_i(t) + \sum_{k \in \pa{i}} B_{ik}(\tau)x_k(t - \tau) \big]\\
-%     &= \sum_{\tau = 1}^\infty \big[C_{ii}(\tau)v_i(t - \tau) + \sum_{k \in \pa{i}} C_{ik}(\tau)x_k(t - \tau)\big].
-%   \end{align*}
-
-%   If $\pa{i} = \emptyset$ then we simply have
-%   $x_i(t) = \sum_{\tau = 1}^\infty C_{ii}(\tau)v_i(t)$ and
-%   $C_{ii} = A_{ii}$ from equation \ref{eqn:wold}.
-
-%   Applying this relation recursively we can write
-
-%   \begin{equation*}
-%     x_i(t) = \sum_{\tau = 1}^\infty\big[C_{ii}(\tau)v_i(t - \tau) + \sum_{k \in \anc{i}}C_{ik}(\tau)v_k(t - \tau) \big].
-%   \end{equation*}
-
+  but this is impossible since with probability one $v_i(t) \not \in \H_{t - 1}^{(i)}$.
 \end{proof}
 
-That strong conditions need to be placed on the topology of $\gcg$ in
-order for proposition \ref{lem:pwgc_anc} to be true can be seen through the
-earlier examples.
+\begin{corollary}
+  \label{cor:gc_implies_pwgc}
+  If $\gcg$ is a strongly causal DAG then $j \gc i \implies j \pwgc i$.
+
+  \hl{This must be true in general, should include this shortly after the definition ow PWGC}
+\end{corollary}
+
+\begin{remark}
+  It can be seen from the conclusion of the proof that simple facts
+  about Granger causality are indeed reliant on our assumptions about
+  the nature of $x(t)$ laid out in the section \ref{sec:theory}, in
+  particular, the innovations process $v(t)$ must be full rank.
+
+  Moreover, that strong conditions need to be placed on the topology
+  of $\gcg$ in order for proposition \ref{prop:pwgc_anc} to hold can
+  be seen through the earlier examples.
+\end{remark}
 
@@
 
-  This follows immediately from the fact (lemma \ref{lem:pwgc_anc})
+  This follows immediately from the fact (lemma \ref{prop:pwgc_anc})
   that $j \in \anc{i} \implies j \pwgc i$.
@@
   Since $j \in \anc{i}$ there is a path $\gcgpath{j}{a}{p}{i}$.
-  Moreover, since $i \pwgc j$ by lemma \ref{lem:pwgc_anc} there must
+  Moreover, since $i \pwgc j$ by lemma \ref{prop:pwgc_anc} there must
   be a common ancestor $k \in \anc{i} \cap \anc{j}$ (otherwise, we would
@@
       \;
-      $C_k \leftarrow \{i \in S_{k + 1}\ |\ \forall s \in S_{k + 1}\ s \not\pwgc i\}$   \texttt{\# potential children of }$P_k$\\
+      $C_k \leftarrow \{j \in S_{k + 1}\ |\ \forall s \in S_{k + 1}\ s \not\pwgc j\}$   \texttt{\# potential children of }$P_k$\\
 
@@
 
+\begin{lemma}
+  \hl{Some lemma on the inner loop.}
+\end{lemma}
+\begin{proof}
+  Let $j \in C_k$ and $i \in \bigcup_{r = 0}^k P_{k - r}$.
+
+  First, suppose that $i \in \pa{j}$ and that there is no $i \rightarrow j$ path in $E$.
+Then since $i \gc j$ we have $i \pwgc j$ by proposition \ref{prop:pwgc_anc} so at least one of $\{D_{kr} \}_{r = 0}^k$ contains $(i, j)$ and hence $(i, j)$ will be added to $E$ upon the end of the iteration.
+
+  Now, still assuming that $i \in \pa{j}$, suppose instead that there is already an $i \rightarrow j$ path in $E$.
+Using an induction hypothesis that $E$ contains edges if and only if they are in $\gcg$, then either $(i, j) \in E$ correctly, or there is more than one $i \rightarrow j$ path in $\gcg$, which is not possible since $\gcg$ is strongly causal.
+
+  Finally, suppose that $i \not \in \pa{j}$.
+If $i \not \pwgc j$ then we will correctly leave $(i, j)$ absent from $E$, so suppose that $i \pwgc j$.
+Then we must have either $i \in \anc{j}$ or $\anc{i} \cap \anc{j} \ne \emptyset$; suppose first the former.
+Then, there is some $u$ s.t.
+$i \pwgc u \pwgc j$, so by the construction of \hl{it seems} that both $(i, u)$ and $(u, j)$ should have already been added to $E$, so we have an $i \rightarrow j$ path in $E$ \hl{is this sound?}  Suppose then the latter case above, $u \in \anc{i} \cap \anc{j}$ \hl{try to construct a counterexample involving this case}.
+\end{proof}
+
+Proof of the actual theorem...
+
 \begin{proof}
@@
 
-  Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$, we will proceed by induction.
+% Given non-empty P and C...  Can we show all P -> C edges get collected in D?
+  Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$ and let us define
+
+  \begin{equation*}
+    A_k = \{i \in [n]\ |\ \dist{i}{\widehat{P}_0} = k\},
+  \end{equation*}
+
+  where
+
+  \begin{equation*}
+    \dist{i}{P} = \text{max}\{r \in [n]\ |\ \exists (j \rightarrow i) \text{ path of length } r \text{ in } \gcg \text{ for some } j \in P\}.
+  \end{equation*}
+
+  We will proceed by induction and show that when the counter has reached the value $k \ge 1$, it is guaranteed that all nodes in $\bigcup_{\ell \le k}A_\ell$ (nodes having paths of length up to $k$ between themselves and $\widehat{P}_0$) must have all of their incident edges included in $E$, and that every edge in $E$ is an edge of $\gcg$.
+If this is the case, then it can be immediately seen that the algorithm terminates with the correct result, $\gcg = ([n], E)$, once $k = n$; since there can only be paths up to length $n - 1$ in $\gcg$, and there cannot be any nodes or edges that don't involve paths back to $\widehat{P}_0$.
+That is, $\bigcup_{\ell \le n}A_\ell = [n]$.
+
+  % firstly dispensing with some trivialities, and then establishing that every edge in $\gcg$ makes an appearance in some $D_{kr}$ as well as that everything in any $D_{kr}$ is indeed an edge in $\gcg$.
 
   On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
-The set $\widehat{P}_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $\widehat{P}_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
+To quickly dispense with some trivialities: the set $\widehat{P}_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $\widehat{P}_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
 Now, if $\widehat{P}_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
-On the other hand, if $\widehat{P}_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $\widehat{P}_0$.
-Consider now an edge $(i, j) \in D_0$ ($D_0$ is non-empty for the same reason as $C_0$).
+
+  Continuing with the non-trivial case: if $\widehat{P}_0 \ne [n]$ then $C_0$ must be non-empty, otherwise there would be driving nodes not included in $\widehat{P}_0$.
+In this case it is also clear that $A_1 \ne \emptyset$, so let us consider some $j \in A_1$.
+We must have $j \in C_0$, otherwise there would be a path of length $2$ from $\widehat{P}_0$ to $j$.
+Moreover, since $\exists i \in \widehat{P}_0$ s.t.
+$i \gc j$ and therefore $i \pwgc j$ (by proposition \ref{prop:pwgc_anc}) we have $(i, j) \in D_{00}$ and therefore $(i, j) \in E$.
+Therefore $A_1 \subseteq D_{00}$.
+\hl{This is nonsensical because $A_1$ does not contain tuples -- need to show that $A_1 \subseteq C_0$ ?}
+
+  Now consdider some $(i, j) \in D_{00}$ (the same reasoning as for $C_0$ is sufficient to see that $D_{00}$ is non-empty).
 Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).
-Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{lem:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in \widehat{P}_0 \implies u \in S_0 \implies j \not\in C_0$, a contradiction.
+Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in \widehat{P}_0 \implies u \in S_0 \implies j \not\in C_0$, a contradiction.
+This implies every tuple in $D_{00}$ is a bona-fide edge of $\gcg$, as well as that $(i, j) \in A_1$.
+So then $A_1 = D_{00}$, [\hl{nonsensical, see earlier.}] completing the base case for induction.
+
+  Suppose now the algorithm has reached step $k$ and that $E = \bigcup_{\ell \le k - 1}A_\ell$.
+
+% The reasoning here is (I think) correct -- but isn't directly establishing the induction hypothesis.
+%   On the other hand, if $\widehat{P}_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $\widehat{P}_0$.
+Consider now an edge $(i, j) \in D_{00}$ ($D_{00}$ is non-empty for the same reason as $C_0$).
+Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).
+Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in \widehat{P}_0 \implies u \in S_0 \implies j \not\in C_0$, a contradiction.
+So every tuple in $D_{00}$ is a bona-fide edge of $\gcg$.
+Suppose now that for some $j \in C_0$ there is an $i$ s.t.
+$(i, j) \in \gcg$ but $(i, j) \not\in D_{00}$.
+Since $(i, j) \in \gcg$ we have $i \pwgc j$ (corollary \ref{cor:gc_implies_pwgc}), but then if $i \not\in \widehat{P}_0$ then $j \notin C_0$ and if $i \in \widehat{P}_0, j \in C_0$ then necessarily $(i, j) \in D_{00}$ since we cannot at this stage have any paths in $E$.
 

### 2019-04-30 paper.tex
Expanding j ->PW i necessary conditions proposition.

@@
     \label{eqn:ancestor_expansion}
-    x_i(t) = \A_{ii}(z)v_i(t) + \sum_{k \in \anc{i} \setminus\{i\}}\A_{ik}(z)v_k(t),
+    x_i(t) = \A_{ii}(z)v_i(t) + \sum_{\substack{k \in \anc{i} \\ k \ne i}}\A_{ik}(z)v_k(t),
   \end{equation}
@@
     &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
-    &= \A_{ii}(z) v_i(t) + \sum_{j \in \anc{i} \setminus \{i\}} \A_{ij}(z) v_j(t)\\
+    &= \A_{ii}(z) v_i(t) + \sum_{\substack{j \in \anc{i} \\ j \ne i}} \A_{ij}(z) v_j(t)\\
   \end{align*}
@@
     \item{$j \in \anc{i}$}
-    \item{$\anc{i} \cap \anc{j} \ne \emptyset$}
+    \item{$\anc{i} \cap \anc{j} \ne \emptyset$.}
   \end{enumerate}
+
+  \hl{Moreover, if $j \not\in \anc{i}$ then $\exists u \in \anc{i} \cap \anc{j}$ such that $i$ is not contained in at least one $u \rightarrow \cdots \rightarrow j$ path.  [Must still prove this.]}
 \end{proposition}
@@
     &+ \sum_{k \in \anc{i}}\E[\big(\A_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{k \in \anc{j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
-    &+ \sum_{k \in \anc{i} \setminus i}\sum_{\ell \in \anc{j} \setminus j}\E[\big(\A_{ik}(z)v_k(t)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)]\\
+    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ik}(z)v_k(t)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)]\\
   \end{align*}
@@
   so by theorem (\ref{thm:granger_causality_equivalences}) we conclude $j \not\pwgc i$.
-\end{proof}
-
-We can state some interesting corollaries of this result.
+
+  For the second part of the theorem we exclude possibility $(1)$, namely, $j \not \in \anc{i}$ and suppose by way of contradiction that for every $u \in \anc{i} \cap \anc{j}$ 
+\end{proof}
+
+An interesting corollary is the following:
 
@@
 
+\begin{proposition}
+  If $\gcg$ is a strongly causal DAG and $i \pwgc j$ and $j \pwgc i$, then $i \not\in \pa{j}$ and $j \not\in \pa{i}$.
+That is, a pairwise bidirectional edge indicates the absense of any edge in $\gcg$.
+\end{proposition}
+\begin{proof}
+  By way of contradiction, suppose that $i \in \pa{j}$ and note that the case $j \in \pa{i}$ is symmetric.
+We consider the two possibilities allowed by proposition \ref{prop:ancestor_properties} for $j \pwgc i$:  Firstly $j \in \anc{i}$ is impossible since $\gcg$ is assumed to be acyclic; \hl{Secondly} $u \in \anc{i} \cap \anc{j}$ \hl{leads to a contradiction} since there must be multiple paths, namely $u \rightarrow \cdots \rightarrow i \rightarrow \cdots \rightarrow j$ and $u \rightarrow \cdots \rightarrow j$ in $\gcg$, but this isn't possible since $\gcg$ is strongly causal, therefore we cannot have $i \in \pa{j}$.
+\hl{This requires a strengthening of proposition 3!}
+\end{proof}
+
 In light of proposition \ref{prop:sc_graph_common_anc}, the following
@@
 
-  but this is impossible since with probability one $v_i(t) \not \in \H_{t - 1}^{(i)}$.
+  but this is impossible since with probability one $x_i(t) \not \in \H_{t - 1}^{(i)}$.
 \end{proof}

### 2019-04-30 paper.tex
Refining argument for ancestor properties expansion -- need to recurse on C_1 now.

@@
 
-  \hl{Moreover, if $j \not\in \anc{i}$ then $\exists u \in \anc{i} \cap \anc{j}$ such that $i$ is not contained in at least one $u \rightarrow \cdots \rightarrow j$ path.  [Must still prove this.]}
+  \hl{Moreover, if $j \not\in \anc{i}$ then $\exists u \in \anc{i} \cap \anc{j}$ such that $i$ is not contained in at least one $u \rightarrow \cdots \rightarrow j$ path, such an ancestor can be referred to as a \textit{confounder} [Must still prove this.]}
 \end{proposition}
-\begin{proof}
-  We will show the contrapositive.
-Assume that $j \not \in \anc{i}$ and $\anc{i} \cap \anc{j} = \emptyset$, then $\forall \tau \in \Z_+,\ \E x_i(t)x_j(t - \tau) = 0$ since from the representation (\ref{eqn:parent_expansion}) we have
+
+\begin{remark}
+  Considering the contrapositive of the first part of this theorem (which is what we actually prove below) is enlightening: it says that if $i$ is not a decendent of $j$ ($j \not \in \anc{i}$) and that there are no common ancestors ($\anc{i} \cap \anc{j} = \emptyset$) then $j \not \pwgc i$, or $\linE{x_i(t)}{\H_{t - 1}^{(i, j)}} = \linE{x_i(t)}{\H_{t - 1}^{(i)}}$, which is an intuitively natural result.
+
+  The first part is used in part to prove the second (and stronger) statement, which allows common ancestors, but only if the are ``mediated'' by $i$.
+For example, this covers the case of a graph $k \rightarrow i \rightarrow j$, where $\{k\} = \anc{i} \cap \anc{j}$, but clearly $j \not \pwgc i$.
+\end{remark}
+
+\begin{proof}
+  We will first establish the contrapositive of $(1)$ and $(2)$: assume that $j \not \in \anc{i}$ and $\anc{i} \cap \anc{j} = \emptyset$, then we will show that $\forall \tau \in \Z_+,\ \E x_i(t)x_j(t - \tau) = 0$.
+This follows from representation (\ref{eqn:parent_expansion}) since we have
 
@@
 
-  For the second part of the theorem we exclude possibility $(1)$, namely, $j \not \in \anc{i}$ and suppose by way of contradiction that for every $u \in \anc{i} \cap \anc{j}$ 
+  For the second part of the theorem we exclude possibility $(1)$, namely, $j \not \in \anc{i}$ and suppose by way of contradiction that for every $k \in \anc{i} \cap \anc{j}$ every $k \rightarrow \cdots \rightarrow j$ path contains $i$.
+First define the following partition of $\pa{j}$:
+
+  \begin{align*}
+    C_0 &= \{k \in \pa{j}\ |\ i \not\in \anc{k}, \anc{i} \cap \anc{k} = \emptyset, k \ne i\}\\
+    C_1 &= \{k \in \pa{j}\ |\ i \in \anc{k} \text{ or } k = i\}\\
+    C_2 &= \{k \in \pa{j}\ |\ i \not\in \anc{k}, \anc{i} \cap \anc{k} \ne \emptyset, k \ne i\}.
+  \end{align*}
+
+  We notice that $C_2 = \emptyset$ since if $k \in C_2$ then $\exists u \in \anc{i} \cap \anc{k}$ s.t.
+$i \not \in \anc{k}$ and therefore there must be a path $u \rightarrow \cdots \rightarrow k \rightarrow \cdots \rightarrow j$ which does not contain $i$, violating the condition taken in the second part of the theorem's statement.
+We can therefore expand $x_j(t)$ in terms of it's parents, where $\A_{jj}(z) = (1 - \B_{jj}(z))^{-1}$:
+
+  \begin{equation}
+    \label{eqn:xj_partition_expansion}
+    x_j(t) = \A_{jj}(z)\Big(v_j(t) + \sum_{k \in C_0}\B_{ik}(z)x_k(t) + \sum_{k \in C_1}\B_{ik}(z)x_k(t)\Big).
+  \end{equation}
+
+  Using this representation we choose an arbitrary $\Phi(z) x_i(t - 1) \in \H_{t - 1}^{(j)}$ and show that
+
+  \begin{equation}
+    \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}{\Phi(z)x_j(t - 1) - \linE{\Phi(z)x_j(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
+  \end{equation}
+
+  which will imply (\hl{by lemma ???}) that $j \not\pwgc i$ and for which it is equivalent to show that
+
+  \begin{equation}
+    \label{eqn:sufficient_inner_prod}
+    \inner{x_i(t)}{\Phi(z)x_j(t - 1) - \linE{\Phi(z)x_j(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
+  \end{equation}
+
+  by the orthogonality principle since $\linE{x_i(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$.
+Substituting (\ref{eqn:xj_partition_expansion}) into (\ref{eqn:sufficient_inner_prod}) and proceeding term by term we have firstly that
+
+  \begin{equation*}
+    \inner{x_i(t)}{\Phi(z)\A_{jj}(z)v_j(t - 1) - \linE{\Phi(z)\A_{jj}(z)v_j(t - 1)}{\H_{t - 1}^{(i)}}} = 0
+  \end{equation*}
+
+  by expanding $x_i(t)$ with (\ref{eqn:ancestor_expansion}) and using the fact that $j \not \in \anc{i}$.  Secondly we see that $\forall k \in C_0$
+
+  \begin{equation*}
+    \inner{x_i(t)}{\Phi(z)\A_{jj}(z)\B_{ik}(z)x_k(t - 1) - \linE{\Phi(z)\A_{jj}(z)\B_{ik}(z)x_k(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
+  \end{equation*}
+
+  which follows from the first part of the theorem.  Finally, for $k \in C_1$ the case $k = i$ is immediate, so suppose $k \ne i$.  
 \end{proof}

### 2019-04-30 paper.tex
Proof of extension to ancestor properties theorem.

--- 2019-04-30 paper.tex
@@
   \label{prop:ancestor_properties}
-\hl{Need to refine this to exclude $k \rightarrow i \rightarrow j$ from 2.}
-
   In a Granger-causality graph $\gcg$, if $j \pwgc i$ then at least
@@
 
-  \hl{Moreover, if $j \not\in \anc{i}$ then $\exists u \in \anc{i} \cap \anc{j}$ such that $i$ is not contained in at least one $u \rightarrow \cdots \rightarrow j$ path.  [Must still prove this.]}
+  Moreover, if $j \not\in \anc{i}$ then $\exists u \in \anc{i} \cap \anc{j}$ such that $i$ is not contained in at least one $u \rightarrow \cdots \rightarrow j$ path.
 \end{proposition}
+
+\begin{remark}
+  Considering the contrapositive of the first part of this theorem (which is what we actually prove below) is enlightening: it says that if $i$ is not a descendent of $j$ (i.e.
+$j \not \in \anc{i}$) and that there are no common ancestors (i.e.
+$\anc{i} \cap \anc{j} = \emptyset$) then $j \not \pwgc i$, or $\linE{x_i(t)}{\H_{t - 1}^{(i, j)}} = \linE{x_i(t)}{\H_{t - 1}^{(i)}}$, which is an intuitively natural result.
+
+  The first part is used in part to prove the second (and stronger) statement, which allows common ancestors, but only if they are ``mediated'' by $i$.
+For example, this covers the case of a graph $k \rightarrow i \rightarrow j$, where $\{k\} = \anc{i} \cap \anc{j}$, but clearly $j \not \pwgc i$.
+
+  An ancestor such as that in the second part of the theorem can be referred to as a \textit{confounder}, and a simple example is given by the fork $i \leftarrow k \rightarrow j$.
+\end{remark}
+
 \begin{proof}
-  We will show the contrapositive.
-Assume that $j \not \in \anc{i}$ and $\anc{i} \cap \anc{j} = \emptyset$, then $\forall \tau \in \Z_+,\ \E x_i(t)x_j(t - \tau) = 0$ since from the representation (\ref{eqn:parent_expansion}) we have
+  We will first establish the contrapositive of $(1)$ and $(2)$: assume that $j \not \in \anc{i}$ and $\anc{i} \cap \anc{j} = \emptyset$, then we will show that $\forall \tau \in \Z_+,\ \E x_i(t)x_j(t - \tau) = 0$.
+This follows from representation (\ref{eqn:parent_expansion}) since we have
 
@@
     \E x_i(t)x_j(t - \tau) &= \E \big(\A_{ii}(z)v_i(t)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)\\
-    &+ \sum_{k \in \anc{i}}\E[\big(\A_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{k \in \anc{j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
+    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\E[\big(\A_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
     &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ik}(z)v_k(t)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)]\\
@@
 The fourth term above is 0 since $\anc{i} \cap \anc{j} = \emptyset$.
-For the third term, it is possible that (due to cycles) $i \in \anc{j}$ which requires more care, however similarly as before we can at least see that
+For the third term, it is possible that (due to cycles) $i \in \anc{j}$ which requires more care, although similarly as before we can at least see immediately that
 
   \begin{equation*}
-    \sum_{k \in \anc{j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)] = \E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)].
+    \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)] = \E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)].
   \end{equation*}
@@
 
-  For the second part of the theorem we exclude possibility $(1)$, namely, $j \not \in \anc{i}$ and suppose by way of contradiction that for every $u \in \anc{i} \cap \anc{j}$ 
+  For the second part of the theorem we will suppose that $j$ is a node having the following two properties: $(a)$ $j \not \in \anc{i}$ and $(b)$ for every $k \in \anc{i} \cap \anc{j}$ every $k \rightarrow \cdots \rightarrow j$ path contains $i$.
+The first is the negation of property $(1)$ above, and the second is to setup the contradiction $j \not\pwgc i$, which will then establish the theorem.
+
+  Firstly, notice that every $u \in \big(\pa{j} \setminus \{i\}\big)$ necesssarily inherits these same two properties since if $u \in \anc{i}$ then $u \in \anc{i} \cap \anc{j}$ so that every $u \rightarrow \cdots \rightarrow j$ path must contain $i$, but $u \in \pa{j}$, so this is not the case since $u \rightarrow j$ is a path that doesn't contain $i$; moreover, if we consider $w \in \anc{i} \cap \anc{u}$ then we also have $w \in \anc{i} \cap \anc{j}$ so every $w \rightarrow \cdots \rightarrow j$ path must contain $i$.
+These properties therefore extend inductively to every $u \in \big(\anc{j} \setminus \{i\}\big)$.
+
+  In order to deploy a recursive argument, define the following partition of $\pa{u}$, for some node $u$:
+
+  \begin{align*}
+    C_0(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, \anc{i} \cap \anc{k} = \emptyset, k \ne i\}\\
+    C_1(u) &= \{k \in \pa{u}\ |\ i \in \anc{k} \text{ or } k = i\}\\
+    C_2(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, \anc{i} \cap \anc{k} \ne \emptyset, k \ne i\}.
+  \end{align*}
+
+  We notice that for any $u$ having the properties $a, b$ above, we must have $C_2(u) = \emptyset$ since if $k \in C_2(u)$ then $\exists w \in \anc{i} \cap \anc{k}$ s.t.
+$i \not \in \anc{k}$ and therefore there must be a path $w \rightarrow \cdots \rightarrow k \rightarrow \cdots \rightarrow u$ which does not contain $i$.
+
+  Using this partition, we will expand $x_j(t)$ in terms of it's parents, and recursively expand nodes in $C_1$ until we reach a case where $C_1 = \emptyset$.
+For the first step:  %where $\A_{jj}(z) = (1 - \B_{jj}(z))^{-1}$:
+
+  \begin{equation}
+    \label{eqn:xj_partition_expansion}
+    x_j(t) = \A_{jj}(z)\Big(v_j(t) + \sum_{k \in C_0(j)}\B_{ik}(z)x_k(t) + \sum_{k \in C_1(j)}\B_{ik}(z)x_k(t)\Big).
+  \end{equation}
+
+  Using this representation we choose an arbitrary $\Phi(z) x_i(t - 1) \in \H_{t - 1}^{(j)}$ and show that
+
+  \begin{equation}
+    \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}{\Phi(z)x_j(t - 1) - \linE{\Phi(z)x_j(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
+  \end{equation}
+
+  which will imply (\hl{by lemma ???}) that $j \not\pwgc i$ and for which it is equivalent to show that
+
+  \begin{equation}
+    \label{eqn:sufficient_inner_prod}
+    \inner{x_i(t)}{\Phi(z)x_j(t - 1) - \linE{\Phi(z)x_j(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
+  \end{equation}
+
+  by the orthogonality principle since $\linE{x_i(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$.
+Substituting (\ref{eqn:xj_partition_expansion}) into (\ref{eqn:sufficient_inner_prod}) and starting with the first term we have
+
+  \begin{align*}
+    &\inner{x_i(t)}{\Phi(z)\A_{jj}(z)v_j(t - 1) - \linE{\Phi(z)\A_{jj}(z)v_j(t - 1)}{\H_{t - 1}^{(i)}}}\\
+    \overset{(\alpha)}{=}\ &\inner{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}{\Phi(z)\A_{jj}(z)v_j(t - 1)}\\
+    \overset{(\beta)}{=}\ &0,
+  \end{align*}
+
+  where $(\alpha)$ follows by expanding $x_i(t)$ with (\ref{eqn:ancestor_expansion}) and $\linE{\Phi(z)\A_{jj}(z)v_j(t - 1)}{\H_{t - 1}^{(i)}} = 0$ because $\forall \tau, s$
+
+  \begin{equation*}
+    \E v_j(t - \tau) x_i(t - s) = \E v_j(t - \tau) \sum_{k \in \anc{i} \cup \{i\}}\A_{ik}(z)v_k(t - s) = 0,
+  \end{equation*}
+
+  since $j \not \in \anc{i}$; $(\beta)$ follows similarly, that is, $j \not \in \anc{i}$.  Secondly we see that $\forall k \in C_0(j)$
+
+  \begin{equation*}
+    \inner{x_i(t)}{\Phi(z)\A_{jj}(z)\B_{ik}(z)x_k(t - 1) - \linE{\Phi(z)\A_{jj}(z)\B_{ik}(z)x_k(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
+  \end{equation*}
+
+  which follows from the first part of the theorem.
+Finally, for $k \in C_1(j)$ the case $k = i$ is immediate, so suppose $k \ne i$.
+We know from above that $k$ inherits the key properties referred to as $(a), (b)$ and therefore we can recursively expand $k$ in the same way as in equation (\ref{eqn:xj_partition_expansion}).
+Continuing this recursion for each $k \in C_1(j)$ where $k \ne i$ must eventually terminate since $i \in \anc{k}$.
 \end{proof}

### 2019-05-01 paper.tex
Organizing and finishing / cleaning up some other proofs.

@@
 \newcommand{\gpn}[2]{gp_{#1}(#2)}  % nth generation grandparents
-\newcommand{\gcgpath}[4]{#1 \rightarrow #2_1 \rightarrow \cdots \rightarrow #2_{#3} \rightarrow #4}
 \newcommand{\wtalpha}[2]{\widetilde{\alpha}(#1, #2)}  % Some notation for lem:pwgc_anc
 \newcommand{\dist}[2]{\mathsf{d}(#1, #2)}  % Distance between things
+\newcommand{\gcgpath}[2]{#1 \rightarrow \cdots \rightarrow #2}  % A shorter path command
 
@@
 
+\begin{definition}[\hl{Confounder}]
+  A node $k$ will be referred to as a \textit{confounder} of nodes
+  $i, j$ (neither of which are equal to $k$) if
+  $k \in \anc{i} \cap \anc{j}$ and there exists a path
+  $\gcgpath{k}{i}$ not containing $j$, and a path $\gcgpath{k}{i}$
+  not containing $i$.
+\end{definition}
+
 \begin{proposition}
@@
 
-  Moreover, if $j \not\in \anc{i}$ then $\exists u \in \anc{i} \cap \anc{j}$ such that $i$ is not contained in at least one $u \rightarrow \cdots \rightarrow j$ path.
+  Moreover, if $j \not\in \anc{i}$ then $\exists u \in \anc{i} \cap \anc{j}$ which is a confounder, that is, such that $i$ is not contained in at least one $\gcgpath{u}{j}$ path.
 \end{proposition}
@@
 
-It seems intuitive that a converse of \ref{prop:ancestor_properties}
+It seems intuitive that a converse of proposition \ref{prop:ancestor_properties}
 would hold, i.e. $j \in \anc{i} \implies j \pwgc i$.  Unfortunately,
@@
 
-The previous examples seem rather pathological since we have simply
-constructed a case where different paths in the graph cancel exactly.
-It may be possible to rid ourselves of these pathologies for instance
-by defining a notion of ``robust'' pairwise causality where
-``$j \pwgc _R i$'' if $j \pwgc i$ almost surely for a ``small'' random
-perturbation of the system matrix.  However, we do not believe such a
-modification of our theory can have useful practical implications
-since the difference between paths that ``nearly cancel'' rather than
-cancel exactly would in practice require inordinate amounts of data to
-resolve.
+\subsection{Strongly Causal Graphs}
+The examples at the conclusion of the previous section seem rather
+pathological since we have simply constructed a case where different
+paths in the graph cancel exactly.  It may be possible to rid
+ourselves of these pathologies for instance by defining a notion of
+``robust'' pairwise causality where ``$j \pwgc _R i$'' if $j \pwgc i$
+almost surely for a ``small'' random perturbation of the system
+matrix.  However, we do not believe such a modification of our theory
+can have useful practical implications since the difference between
+paths that ``nearly cancel'' rather than cancel exactly would in
+practice require inordinate amounts of data to resolve.
 
@@
 
+\hl{Conjecture: A graph is strongly causal if and only if it contains
+  no confounders.}
+
 Examples of strongly causal graphs include directed trees (or
 forests), and DAGs where each node has at most one parent.
+Fortunately, the strongly causal property is inherited by subgraphs:
 
@@
   If $\gcg$ is a strongly causal graph, any subgraph formed by eliminating nodes
-  and all of the in and out edges of the eliminated node is still strongly causal.
+  as well as all of the in and out edges thereof is still strongly causal.
 \end{lemma}
@@
 
-\begin{lemma}
+We explore a number of properties of strongly causal graphs before
+moving into the main result of section \ref{sec:theory} in section
+\ref{sec:pairwise_algorithm}.
+
+\begin{proposition}
   \label{prop:sc_graph_common_anc}
-  In a strongly causal graph if $j \in \anc{i}$ and $k \in \anc{i} \cap \anc{j}$ then the unique path from $k$ to $i$ contains $j$.
-\end{lemma}
-\begin{proof}
-  Suppose that there is a path from $k$ to $i$ which does not contain $j$.
-In this case, there are multiple paths from $k$ to $i$ (one of which \textit{does} go through $j$ since $j \in \anc{i}$) which contradicts the assumption of strong causality.
+  In a strongly causal graph if $j \in \anc{i}$ then any
+  $k \in \anc{i} \cap \anc{j}$ is not a confounder, that is,
+  the unique path from $k$ to $i$ contains $j$.
+\end{proposition}
+\begin{proof}
+  Suppose that there is a path from $k$ to $i$ which does not contain
+  $j$.  In this case, there are multiple paths from $k$ to $i$ (one of
+  which \textit{does} go through $j$ since $j \in \anc{i}$) which
+  contradicts the assumption of strong causality.
 \end{proof}
@@
 \begin{proposition}
-  If $\gcg$ is a strongly causal DAG and $i \pwgc j$ and $j \pwgc i$, then $i \not\in \pa{j}$ and $j \not\in \pa{i}$.
-That is, a pairwise bidirectional edge indicates the absense of any edge in $\gcg$.
+  If $\gcg$ is a strongly causal DAG then $i \pwgc j$ and $j \in \anc{i}$ are
+  \textit{alternatives}, that is $i \pwgc j \implies j \notin \anc{i}$.
 \end{proposition}
 \begin{proof}
-  By way of contradiction, suppose that $i \in \pa{j}$ and note that the case $j \in \pa{i}$ is symmetric.
-We consider the two possibilities allowed by proposition \ref{prop:ancestor_properties} for $j \pwgc i$:  Firstly $j \in \anc{i}$ is impossible since $\gcg$ is assumed to be acyclic; \hl{Secondly} $u \in \anc{i} \cap \anc{j}$ \hl{leads to a contradiction} since there must be multiple paths, namely $u \rightarrow \cdots \rightarrow i \rightarrow \cdots \rightarrow j$ and $u \rightarrow \cdots \rightarrow j$ in $\gcg$, but this isn't possible since $\gcg$ is strongly causal, therefore we cannot have $i \in \pa{j}$.
-\hl{This requires a strengthening of proposition 3!}
+  Suppose that $i \pwgc j$ and $j \in \anc{i}$.  Then since $\gcg$ is
+  acyclic $i \not\in \anc{j}$, and by proposition
+  \ref{prop:ancestor_properties} there is some
+  $k \in \anc{i}\cap\anc{j}$ which is a confounder.  However, by
+  proposition \ref{prop:sc_graph_common_anc} $k$ cannot be a
+  confounder, a contradiction.
+\end{proof}
+% This is a direct proof
+  % \begin{proof}
+%   Suppose that $\gcg$ is a strongly causal DAG and that we have both
+%   $i \pwgc j$ and $j \in \anc{i}$, which implies that
+%   $i \not \in \anc{j}$ since $\gcg$ is a DAG.  We will establish the
+%   contradiction $i \in \anc{j}$.
+
+%   Since $j \in \anc{i}$ there is a path $\gcgpath{j}{i}$.
+%   Moreover, since $i \pwgc j$ by proposition \ref{prop:pwgc_anc} there
+%   must be a confounding ancestor $k \in \anc{i} \cap \anc{j}$, where the
+%   node $k$ has a path to $j,\ \gcgpath{k}{j}$, as well as a path
+%   to $i,\ \gcgpath{k}{i}$.
+
+%   \hl{Double check that the ancestor properties proposition implies
+%     that BOTH paths do not contain the other node.}
+
+%   Now, since the graph is strongly causal, the only possible
+%   $\gcgpath{k}{i}$ path is the one obtained by concatenating the
+%   $\gcgpath{k}{j}$ path with the $\gcgpath{j}{i}$ path.
+%   However, this is a contradiction since $k$ is a confounder and
+%   should have a $\gcgpath{k}{i}$ path which does not contain $i$.
+% \end{proof}
+
+\begin{proposition}
+  \label{prop:bidirectional_edge}
+  If $\gcg$ is a strongly causal DAG and $i \pwgc j$ and $j \pwgc i$,
+  then $i \not\in \pa{j}$ and $j \not\in \pa{i}$.  That is, a pairwise
+  bidirectional edge indicates the absense of any edge in $\gcg$.
+\end{proposition}
+\begin{proof}
+  \hl{TODO: This is just a corollary of the previous proposition.}
+% This is a direct proof
+% \begin{proof}
+  By way of contradiction, suppose that $i \in \pa{j}$.  We will
+  consider the two possibilities allowed by proposition
+  \ref{prop:ancestor_properties} for $j \pwgc i$.  Firstly
+  $j \in \anc{i}$ is impossible since $\gcg$ is assumed to be acyclic.
+  Secondly, if there is some confounding node
+  $u \in \anc{i} \cap \anc{j}$ with a path
+  $u \rightarrow \cdots \rightarrow j$ which does not contain $i$ we
+  have a contradiction since there must now be multiple
+  $u \rightarrow j$ paths: the aforementioned, and a path
+  $u \rightarrow \cdots \rightarrow i \rightarrow \cdots \rightarrow
+  j$ which \textit{does} contain $i$.  We conclude that $i \in \pa{j}$
+  is impossible, and symmetrically that $j \in \pa{i}$ is as well.
+% \end{proof}
 \end{proof}
@@
   Firstly, we will establish a representation of $x_i(t)$ that involves $x_j(t)$.
-Denote by $a_{r + 1} \rightarrow a_r \rightarrow \cdots \rightarrow a_1 \rightarrow a_0$ with $a_{r + 1} \defeq j$ and $a_0 \defeq i$ the \textit{unique} $j \rightarrow i$ path in $\gcg$, we will expand the representation of equation (\ref{eqn:parent_expansion}) backwards along this path:
+Denote by $a_{r + 1} \rightarrow a_r \rightarrow \cdots \rightarrow a_1 \rightarrow a_0$ with $a_{r + 1} \defeq j$ and $a_0 \defeq i$ the \textit{unique} $\gcgpath{j}{i}$ path in $\gcg$, we will expand the representation of equation (\ref{eqn:parent_expansion}) backwards along this path:
 
@@
 
+We immediately obtain the corollary, which we remind the reader is,
+surprisingly, not true in a general graph.
+
 \begin{corollary}
@@
   If $\gcg$ is a strongly causal DAG then $j \gc i \implies j \pwgc i$.
-
-  \hl{This must be true in general, should include this shortly after the definition ow PWGC}
 \end{corollary}
@@
 
+\hl{We can possibly supply another part of the converse, as well as
+  formalizing the concept described in the second part of proposition} \ref{prop:ancestor_properties}.  \hl{This also ties in nicely with proposition} \ref{prop:bidirectional_edge}
+
+\hl{Finally, that only \textit{confounders} seem to cause problems,
+  but not \textit{colliders} is a feature of the time-based nature of
+  Granger-causality as opposed to Pearl's static notion of causation.}
+
+\begin{proposition}
+  If $\gcg$ is a strongly causal DAG and nodes $i, j$ have a confounder $k$, then
+  $i \pwgc j$ and $j \pwgc i$.
+\end{proposition}
+
+\begin{proof}
+  \hl{I'm not actually sure yet if this is true.}
+\end{proof}
+
 % $B(1) = \left[ \begin{array}{cc} 0 & 1 \\ 1 & 0 \end{array} \right]$
@@
 % - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-We are interested in understanding under what conditions a causality
-graph can be correctly recovered by pairwise testing alone.  It is not
-enough that the graph simply be sparse, as the following example shows.
-
-\begin{example}
-  If the causality graph $\gcg$ of the vector process $x(t)$ is
-  strongly connected\footnote{there is a directed path from each node
-    to every other node}, then the causality graph recovered via
-  pairwise causality tests is the complete graph.
-
-  This follows immediately from the fact (lemma \ref{prop:pwgc_anc})
-  that $j \in \anc{i} \implies j \pwgc i$.
-\end{example}
 
@@
 
-\begin{lemma}
-  If $\gcg$ is a strongly causal DAG then $i \pwgc j$ and $j \in \anc{i}$ are
-  \textit{alternatives}, that is $i \pwgc j \implies j \notin \anc{i}$.
-\end{lemma}
-\begin{proof}
-  Suppose that $\gcg$ is a strongly causal DAG and that we have both
-  $i \pwgc j$ and $j \in \anc{i}$.  We will show that $i \in \anc{j}$,
-  which produces a contradiction (since there must then be a directed cycle).
-
-  Since $j \in \anc{i}$ there is a path $\gcgpath{j}{a}{p}{i}$.
-  Moreover, since $i \pwgc j$ by lemma \ref{prop:pwgc_anc} there must
-  be a common ancestor $k \in \anc{i} \cap \anc{j}$ (otherwise, we would
-  have $i \in \anc{j}$, contradicting the acyclic property).  The node
-  $k$ has paths to both $j,\ \gcgpath{k}{b}{q}{j}$ and $i,\ \gcgpath{k}{c}{r}{i}$.
-
-  Now, since the graph is strongly causal, path $c$ must overlap with
-  path $a$ from $j$ to $i$ otherwise there would be multiple paths
-  from $k$ to $i$.  By the same reasoning with paths from $k$ to $j$,
-  the path $c$ must also overlap with path $b$ from $k$ to $j$, so in
-  fact path $c$ from $k$ to $i$ is
-  $c = b \rightarrow j \rightarrow a$.
-
-  \hl{TODO: We need the strengthening of prop }(\ref{prop:ancestor_properties})\hl{in order to conclude.}
-
-\end{proof}
-
 \subsection{Recovering $\gcg$ via Pairwise Tests}
-In this section we will show that if $\gcg$ is a strongly causal DAG, then it is possible to recover $\gcg$ via pairwise tests alone.
-This is an interesting fact in and of itself, but also has some implications for applications since pairwise testing is trivially parallelizable.
-In section \ref{sec:structure_learning} we will also analyze the use of pairwise testing as a heuristic for general graphs.
+\label{sec:pairwise_algorithm}
+In this section we will show that if $\gcg$ is a strongly causal DAG,
+then it is possible to recover $\gcg$ via pairwise tests alone.  This
+is an interesting fact in and of itself, but also has some
+implications for applications since pairwise testing is trivially
+parallelizable.  In section \ref{sec:structure_learning} we will also
+analyze the use of pairwise testing as a heuristic for general graphs.
+
+In order to recover $\gcg$ via pairwise testing alone, it is not
+enough that the graph simply be sparse, as the following example
+shows.
+
+\begin{example}
+  If the causality graph $\gcg$ of the vector process $x(t)$ is
+  strongly connected\footnote{there is a directed path from each node
+    to every other node}, then the causality graph recovered via
+  pairwise causality tests is the complete graph.
+
+  This follows immediately from the fact (proposition \ref{prop:pwgc_anc})
+  that $j \in \anc{i} \implies j \pwgc i$.
+\end{example}
 

### 2019-05-01 paper/paper.tex
Moved paper to subfolder and added some more remarks.

--- 2019-03-01 paper.tex
@@
 
-\def\gc{\overset{\text{GC}}{\rightarrow}}
-\def\te{\overset{\mathcal{T}}{\rightarrow}}
+\def\gc{\overset{\text{GC}}{\rightarrow}}  % Granger causality arrow
+\def\pwgc{\overset{\text{PW}}{\rightarrow}}  % Granger causality arrow
+\def\te{\overset{\mathcal{T}}{\rightarrow}}  % Transfer entropy arrow
+\def\gcg{\mathcal{G}}  % Granger-causality graph
+\def\VAR{\mathsf{VAR}}  % VAR(p) model
+\def\B{\mathsf{B}}  % Filter B
+\def\wtB{\widetilde{\B}}  % General filter B
+\def\A{\mathsf{A}}  % Filter A
+\def\H{\mathcal{H}}  % Hilbert space
+
+\newcommand{\linE}[2]{\hat{\E}[#1\ |\ #2]}  % Linear projection
+\newcommand{\linEerr}[2]{\xi[#1\ |\ #2]}  % Error of linear projection
+\newcommand{\pa}[1]{pa(#1)}  % Parents of a node
+\newcommand{\anc}[1]{\mathcal{A}(#1)}  % Ancestors of a node
+\newcommand{\ancn}[2]{\mathcal{A}_{#1}(#2)}  % nth ancestors of a node
+\newcommand{\gpn}[2]{gp_{#1}(#2)}  % nth generation grandparents
+\newcommand{\wtalpha}[2]{\widetilde{\alpha}(#1, #2)}  % Some notation for lem:pwgc_anc
+\newcommand{\dist}[2]{\mathsf{d}(#1, #2)}  % Distance between things
+\newcommand{\gcgpath}[2]{#1 \rightarrow \cdots \rightarrow #2}  % A shorter path command
 
@@
 
-\graphicspath{{./figures/}}
+\graphicspath{{../figures/}}
 
-\title{Structure Learning for $VAR(p)$ Models}
+\title{Structure Learning for $\VAR(p)$ Models}
 \author{R. J. Kinnear, R. R. Mazumdar}
@@
 \label{sec:theory}
-Consider a zero mean wide sense stationary, vector valued and discrete
-time stochastic process $x(t) \in L^n_2(\Omega);\ t \in \Z$, having the matrix
-valued covariance sequence
-$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ and spectra $S(\omega)$:
+\subsection{Formal Setting}
+Consider the space $L_2(\Omega)$, the usual Hilbert space of finite
+variance random variables over a probability space
+$(\Omega, \mathcal{F}, \mathbb{P})$ having inner product
+$\inner{x}{y} = \E[xy]$.  We will denote the collection vectors of
+such random elements as $L_2^n(\Omega)$ or simply $L_2^n$.
+
+We will work with a discrete time and wide-sense stationary process
+$x(t)$ taking elements in $L_2$.  We suppose that $x(t)$ has zero
+mean, $\E x(t) = 0$, and the matrix valued covariance sequence
+$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ is absolutely summable and has a spectra $S(\omega)$:
 
@@
 
-\begin{definition}[Granger Causality]
-  For the WSS series $x(t)$ we say that component $x_j$
-  \textit{Granger-Causes} (GC) component $x_i$ (with respect to
-  $x(t)$) and write $x_j \gc x_i$ if given Hilbert spaces
-  $\mathcal{H}^{(i)}_{t - 1}$, $\mathcal{H}^{(i, j)}_{t - 1}$
-  generated by $\{x_i(t - \tau)\}_{\tau = 1}^\infty$ and
-  $\{x_i(t - \tau), x_j(t - s\}_{\tau, s = 1}^\infty$ respectively we have
+We will also work frequently with the spaces spanned by the values of
+such a process
 
 \begin{equation}
-  \xi[x_i(t) \ |\ \mathcal{H}^{(i)}_{t - 1}] > \xi[x_i(t) \ |\ \mathcal{H}^{(i, j)}_{t - 1}],
+  \label{eq:hilbert_space_defn}
+  \begin{aligned}
+    \H^x &= \cl \{\sum_{\tau = -T}^T a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, T \in \N\} \subseteq L_2(\Omega),\\
+    \H_t^x &= \cl \{\sum_{\tau = 0}^T a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, T \in \N\} \subseteq L_2(\Omega),
+  \end{aligned}
 \end{equation}
 
-where
-$\xi[x \ |\ \mathcal{H}] = \E (x - \hat{\E}[x \ |\ \mathcal{H}])^2$ and
-$\hat{\E}[x \ |\ \mathcal{H}] = \text{proj}_{\mathcal{H}}(x)$ denotes
-the (unique) projection onto the Hilbert space $\mathcal{H}$.
-\end{definition}
-
-This notion captures the idea that the process $x_j$ provides
-information about $x_i$ that is not available from elsewhere.  The
-caveat ``with respect to $x(t)$'' is important in that GC relations
-can change when components are added to or removed from $x(t)$,
-e.g. new GC relations can arise if we remove the observations of a
-common cause, and existing GC relations can disappear if we observe a
-new mediating series.
-
-The notion is closely related to the information theoretic measure of
-transfer entropy, indeed, if everything in Gaussian then they are
-equivalent \cite{barnett2009granger}.
-
-We continue to review some pertinent facts.
+where the closure is naturally in mean-square.  $\H_t^x$ is the space
+spanned by the values of $x(t)$ between the ``infinite past'' and
+``now'', we will often omit the superscript $x$ which should be clear
+from context.  Evidently these spaces are separable and as closed
+subspaces of a Hilbert space they are themselves Hilbert.  We will
+denote the spaces generated in analagous ways by particular components
+of $x$ as e.g. $\H_t^{(i, j)}$, $\H_t^{i}$ or by all but particular
+components as e.g. $\H_t^{-j}$.
 
 As a consequence of the Wold decomposition theorem \cite{lindquist},
-every such sequence has the ``moving average'' $MA(\infty)$ representation
+every such sequence has the ``moving average'' $MA(\infty)$
+representation
 
@@
 \label{eqn:wold}
-  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) z(t - \tau);\ z(t) \overset{\text{i.i.d.}}{\sim} \mathcal{N}(0, I),
+  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) w(t - \tau),
 \end{equation}
@@
   $c(t)$ for every $t$.  For example,
-  $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi, \pi]$}, and
-$z(t) \overset{\text{i.i.d.}}{\sim} \mathcal{N}(0, I)$.  We will assume
-that $c(t)$ can be effectively estimated and removed and take
-$c(t) = 0$.  Furthermore, we assume that the spectra is uniformly
-bounded away from $0$:
-$\exists \delta:\ \delta I \preceq S(\omega) \preceq \frac{1}{\delta} I\ \forall \omega$ so that the Wold
-representation can be inverted to yield the $VAR(\infty)$ form
+  $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi,
+  \pi]$}, and $w(t)$ is an uncorrelated sequence having a diagonal
+covariance and $A(0) = I$.  We will assume that $c(t) = 0$, which in
+practice is to say that it has been accurately estimated and removed.
+
+Furthermore, we assume that the spectra is uniformly bounded away from
+$0$:
+$\exists \delta:\ \delta I \preceq S(\omega) \preceq \frac{1}{\delta}
+I\ \forall \omega$ so that the Wold representation can be inverted to
+yield the $VAR(\infty)$ form
 
 \begin{equation}
-  x(t) = \sum_{\tau = 1}^\infty B(\tau) x(t - \tau) + v(t);\ v(t) \overset{\text{i.i.d.}}{\sim} \mathcal{N}(0, \Sigma_v).
+  \label{eqn:ar_representation}
+  x(t) = \sum_{\tau = 1}^\infty B(\tau) x(t - \tau) + v(t),
 \end{equation}
 
+where $v(t)$ is an uncorrelated sequence with covariance $\Sigma_v$.
+
+The equations (\ref{eqn:wold}), (\ref{eqn:ar_representation}) can be
+represented as $x(t) = \A(z)w(t) = \B(z)x(t) + v(t)$ via the action of
+the operators (LTI filters)
+$\A(z) \defeq \sum_{\tau = 0}^\infty A(\tau)z^{-\tau}$ and
+$\B(z) \defeq \sum_{\tau = 1}^\infty B(\tau)z^{-\tau}$ where the
+operator $z^{-1}$ is the backshift operator acting on $\H$.  Since
+$||z^{-1}|| = 1$, the inversion formulae
+
+\begin{equation}
+  \label{eqn:lsi_inversion}
+  \A(z) = (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k
+\end{equation}
+
+are valid.
+\subsection{Granger Causality}
+
+\begin{definition}[Granger Causality]
+  For the WSS series $x(t)$ we say that component $x_j$
+  \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x$)
+  and write $x_j \gc x_i$ if given Hilbert spaces
+  $\H_{t - 1}$, $\H^{-j}_{t - 1}$
+
+\begin{equation}
+  \linEerr{x_i(t)}{\H_{t - 1}} < \linEerr{x_i(t)}{\H^{-j}_{t - 1}},
+\end{equation}
+
+where $\xi[x \ |\ \H] = \E (x - \linE{x}{\H})^2$ is the mean squared
+estimation error and $\linE{x}{\H} = \text{proj}_{\H}(x)$ denotes the
+(unique) projection onto the Hilbert space $\H$.
+\end{definition}
+
+This notion captures the idea that the process $x_j$ provides
+information about $x_i$ that is not available from elsewhere.  The
+caveat ``with respect to $x$'' is important in that GC relations can
+change when components are added to or removed from our collection $x$
+of observations, e.g. new GC relations can arise if we remove the
+observations of a common cause, and existing GC relations can
+disappear if we observe a new mediating series.
+
+The notion is closely related to the information theoretic measure of
+transfer entropy, indeed, if everything is Gaussian then they are
+equivalent \cite{barnett2009granger}.
+
 \begin{theorem}[Granger Causality Equivalences]
-  Given W.S.S. vector valued process $x(t)$ the following are equivalent:
+  \label{thm:granger_causality_equivalences}
+  Let $x(t)$ be a WSS process with absolutely summable covariance
+  sequence, and spectral density uniformly bounded above and below.  Denote by
+  $\xi_{ij} \defeq \linEerr{x_i(t)}{\H^{-j}}$ and
+  $\xi_i \defeq \linEerr{x_i(t)}{\H_t}$.  Then, the following are equivalent:
 
   \begin{enumerate}
-    \item{$x_j \gc x_i$}
+    \item{$x_j \not\gc x_i$}
     \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$}
-    \item{$\mathcal{H}_t^{(i)} \perp \mathcal{H}_{t - 1}^{(j)}\ |\ \mathcal{H}_{t - 1}^{(i)}$}
-    \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} > 0$}
-    \item{\hl{Wold $A(\tau)$ condition}}
-    \item{Tie in the 0s of $S(\omega)^{-1}$, though this isn't immediately directed.}
+    \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} = 0$}
+    \item{$\H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
+    \item{$\linE{x_i(t)}{\H_{t - 1}^{-j}} = \linE{x_i(t)}{\H_{t - 1}}$}
+    % \item{\hl{Wold $A(\tau)$ condition}}
+    % \item{Tie in the 0s of $S(\omega)^{-1}$, though this isn't immediately directed.}
   \end{enumerate}
@@
 
-The following propositions justify various modifications of $x(t)$
-applied in practice to massage $x(t)$ into a form amenable to more
-standard tools.
+Statement (4) means that
+$(y - \linE{y}{\H_{t - 1}^{-j}}) \perp (z - \linE{z}{\H_{t -
+    1}^{-j}})\ \forall y \in \H_t^i,\ z \in \H_{t - 1}^j$, which is an
+appealing geometric statement that there is no direct feedback from
+$x_j$ to $x_i$ \cite{lindquist}.
 
-\begin{theorem}[Invariance Under Invertible and Deterministic Modifications]
-  Let $F(z), G(z)$ describe univariate and invertible
+\begin{proof}
+  The equivalence $(1) \iff (3)$ is essentially a restatement of the
+  definition, but is in line with the seminal work of Geweke
+  \cite{geweke1982measurement}, \cite{geweke1984}.
+
+  The equivalence $(3) \iff (5)$ is also immediate from the
+  definitions i.e. the projections are equal if and only if the errors
+  are equal.
+
+  $(1) \implies (2)$: The projection for $x_i(t)$ onto $\H_{t - 1}$ is
+  (by definition) given by a form similar to equation
+  \ref{eqn:ar_representation}:
+
+\[
+  \E|x_i(t) - \linE{x_i(t)}{\H_{t - 1}}|^2 = \E|v_i(t) + \sum_{\tau = 1}^\infty \sum_{k = 1}^n (B_{i, k}(\tau) - \hat{B}_{i, k}(\tau))x_k(t - \tau)|^2.
+\]
+
+Since $v(t)$ in equation \ref{eqn:ar_representation} is temporally
+uncorrelated, it follows that the optimal projection is given by the
+model coefficients themselves.  This holds similarly for the
+projection onto $\H_{t - 1}^{-j}$.  Then by $(1)$ we have
+
+\[
+  \E|x_i(t) - \sum_{\tau = 1}^\infty \sum_{k = 1}^n B_{i, k}(\tau)x_k(t - \tau)|^2 = \E|x_i(t) - \sum_{\tau = 1}^\infty\sum_{k \ne j}B_{i, k}(\tau)x_k(t - \tau)|^2
+\]
+
+By the uniqueness of the projection we must have
+$\forall \tau\ B_{i, j}(\tau) = 0$.
+
+$(2) \implies (4)$: In computing $(y - \linE{y}{\H_{t = 1}^{-j}})$ for
+$y \in \H_t^i$ it is sufficient to consider $y = x_i(t)$ since
+$\H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$, in which case
+$(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$.  $(4)$ follows
+since $v_i(t) \perp \H_{t - 1}$ and
+$\forall z \in \H_{t - 1}^j\ (z - \linE{z}{\H_{t - 1}^{-j}}) \in \H_{t
+  - 1}$
+
+\hl{TODO: $(4) \implies (1)$}
+
+\end{proof}
+
+% The following propositions justify various modifications of $x(t)$
+% applied in practice to massage $x(t)$ into a form amenable to more
+% standard tools.
+
+% \begin{theorem}[General Invariance]
+%   \hl{Under what conditions is this actually true?}
+
+%   Let $\zeta(t)$ be a stationary discrete time stochastic process.  Let
+%   $F, G$ be (possibly nonlinear and time varying) invertible filtering
+%   operations and $f(t)$, $g(t)$ be perfectly predictable and such that
+%   $x_j(t) \defeq F(\zeta_j - f)(t), x_i(t) \defeq G(\zeta_i - g)(t)$ are W.S.S.  Then,
+
+% \begin{equation}
+%     x_j \gc x_i \iff \zeta_j \te \zeta_i
+%   \end{equation}
+% \end{theorem}
+
+\begin{proposition}[Invariance Under Invertible and Deterministic Modifications]
+  Let $x(t)$ be a WSS process with absolutely summable covariance
+  sequence, and spectral density uniformly bounded above and below.  Let
+  $F(z), G$ represent univariate and invertible
   linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
@@
   \end{equation}
+
+\begin{proof}
+  \hl{TODO}
+\end{proof}
+\end{proposition}
+
+\subsection{Pairwise Granger Causality}
+\label{sec:pwgc}
+Recall that Granger-causality in general must be understood with
+respect to a particular universe of observations.  If $x_j \gc x_i$
+with respect to $x_{-k}$, it may not hold with respect to $x$.  For
+example, $x_k$ may be a common ancestor which when observed completely
+explains the connection from $x_j$ to $x_i$.  In particular, we will
+study \textit{pairwise} Granger-causality.
+
+\begin{definition}[Pairwise Granger-causality]
+  We will say that $x_j$ pairwise Granger-causes $x_i$ and write
+  $x_j \pwgc x_i$ if $x_i$ Granger-causes $x_j$ with respect only to
+  $(x_i, x_j)$.
+\end{definition}
+
+This notion is of interest for a variety of reasons.  From a purely
+theoretical or conceptual standpoint, we will seek to understand when
+and how pairwise causality can capture the idea of ``flow of
+information'' in the underlying graph.  It will also be useful to
+reason about what the conditions under which \textit{unobserved}
+components of $x(t)$ may or may not interfere with inference in the
+actually observed components.  Finally, we will seek to construct
+practical estimation procedures based purely on pairwise causality
+tests and understand their properties.
+
+We first need to establish some graph theoretic notation and
+terminology, collected formally in definitions for the reader's
+convenient reference.
+
+\begin{definition}[Causality graph, Parents]
+
+Remaining in the context of a collection of $n$ WSS processes
+$x(t) = \big(x_1(t), \ldots, x_n(t)\big)$, we define the Granger-causality
+graph $\gcg$ to be the directed graph formed on $n$ vertices where an
+edge $(j, i) \in \gcg$ if and only if $x_j \gc x_i$.
+
+We refer to the set of \textit{parents} of $x_i$ in $\gcg$ by
+$\pa{i}$, that is, $j \in \pa{i} \iff x_j \gc x_i$.
+\end{definition}
+
+\begin{definition}[Grandparents, ancestors]
+The set of level $\ell$ \textit{grandparents} of node $i$, denoted
+$\gpn{\ell}{i}$, is the set such that $j \in \gpn{\ell}{i}$ if and only if
+there is a \textit{directed and cycle-free path} of length $\ell$ in
+$\gcg$ from $j$ to $i$ \textit{or} if $j = i$ and there is a cycle of
+length $\ell$ starting at $i$ and ending back at $i$, that is, $i$ can be
+it's own grandparent.  Clearly, $\pa{i} = \gpn{1}{i}$.
+
+Finally, the set of \textit{level $\ell$ ancestors} of
+$i$: $\ancn{\ell}{i} = \bigcup_{\lambda \le \ell}\gpn{\lambda}{i}$ is the set such that
+$j \in \ancn{\ell}{i}$ if and only if there is a directed and cycle-free
+path of length $\ell$ \textit{or less} in $\gcg$ from $j$ to $i$
+\textit{or} if $j = i$ and there is a directed cycle of length $\ell$ or
+less from $i$ back to $i$ in $\gcg$.  The set of \textit{all
+  ancestors} of $i$ (i.e. $\ancn{n}{i}$) is denoted simply $\anc{i}$.
+\end{definition}
+
+Even though the definitions of Granger-causality naturally suggests
+that each node in the graph should have a self-loop (unless the node
+is simply independent noise) we will not include nodes in their own
+parent sets: $i \not\in \pa{i} \subseteq \anc{i}$.  However, we do not
+immediately assume that $\gcg$ is a DAG, and so \textit{we will allow for $i$
+to be it's own ancestor}.
+
+The edges of the Granger-causality graph $\gcg$ can be given a general
+notion of ``weight'' by associating an edge $(j, i)$ with a
+\textit{strictly causal} LTI filter
+
+\begin{equation}
+  \label{eqn:filter}
+  \B_{ij}(z) = \sum_{\tau = 1}^{\infty} B_{ij}(\tau)z^{-\tau}
+\end{equation}
+
+from the $\VAR$ representation of equation
+\ref{eqn:ar_representation}.  We will represent the action
+(convolution) of this filter compactly by taking $z^{-1}$ as the
+backshift operator on $\ell_2^n(\Omega, \mathcal{F}, \mathbb{P})$ when
+convenient
+
+\begin{equation}
+  \sum_{\tau = 1}^\infty B_{ij}(\tau)x_j(t - \tau) \defeq \B_{ij}(z)x_j(t).
+\end{equation}
+
+% What if we take this as the definition of PWGC?  then, in a simply causal graph
+% the "normal" definition becomes a theorem.  Or, if lem:pwgc_anc is true then
+% maybe that is good enough?
+
+From the $\VAR$ representation of $x(t)$ there is clearly a tight
+relationship between each node and it's parent nodes, indeed
+
+\begin{equation}
+  \label{eqn:parent_expansion}
+  x_i(t) = v_i(t) + \B_{ii}(z)x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t).
+\end{equation}
+
+% When is this guaranteed w.r.t. the full system?
+Furthermore, if the filter $\B_{ii}$ is invertible, we can write
+
+\begin{equation*}
+  \begin{aligned}
+    (1 - \B_{ii}(z))x_i(t) &= v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\\
+    \implies x_i(t) &= (1 - \B_{ii}(z))^{-1}\big[v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\big]\\
+    &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)
+  \end{aligned}
+\end{equation*}
+
+% In general, we will write $\wtB$ for arbitrary filters maintaining the
+% convention that $\wtB_{ii}$ is a filter that modifies the future of
+% $x_i$ from it's past and $\wtB_{ij}$ a filter modifying $x_i$ from the
+% past of $x_j$.  We will use the convention that the filter
+% $\widetilde{\B}$ will represent simply ``a filter'' and is not
+% necessarily the same as the filters serving as edge weights in $\gcg$
+% (indeed, $\wtB$ will often be $0$), and moreover, we will allow for
+% $\widetilde{\B}$ to potentially change from place to place and even
+% from line to line in the same sequence of calculations as it's exact
+% properties or specification are not important.
+
+We continue to develop the relationship between $x_i$ and it's family tree.
+
+\begin{proposition}[Ancestor Expansion]
+  \label{prop:parent_expanding}
+  The component $x_i(t)$ of $x(t)$ can be represented as a sum of
+  linear filtering operations acting on the driving noise $v_i$ of
+  $x_i$ as well as the parents of $x_i$:
+
+  \begin{equation}
+    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t).
+  \end{equation}
+
+  Moreover, $x_i$ can be expanded in terms of it's ancestors and
+  driving noise only:
+
+  \begin{equation}
+    \label{eqn:ancestor_expansion}
+    x_i(t) = \A_{ii}(z)v_i(t) + \sum_{\substack{k \in \anc{i} \\ k \ne i}}\A_{ik}(z)v_k(t),
+  \end{equation}
+
+  where $\A(z) = \sum_{\tau = 0}^\infty A(\tau)$ is the filter from
+  the Wold decomposition representation of $x(t)$, equation
+  (\ref{eqn:wold}).
+\end{proposition}
+
+This statement is ultimately about the sparsity pattern in the Wold
+decomposition matrices $A(\tau)$ since
+$x_i(t) = \sum_{\tau = 0}^\infty \sum_{j = 1}^n a_{ij}(\tau)v_j(t -
+\tau)$.  The proposition states that $\A_{ij}(\tau) = 0$ unless
+$j \in \anc{i}$.
+
+% \begin{proof}
+%   The formula (\ref{eqn:parent_expansion}) was established in the
+%   preceding paragraph and serves as a starting point for the induction
+%   hypothesis:
+
+%   \begin{equation}
+%     \label{eqn:induction}
+%     x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
+%   \end{equation}
+
+%   The base case follows by expanding each $x_k$ in equation \ref{eqn:induction} via equation \ref{eqn:parent_expansion}, but we skip the calculations since they are nearly identical to the following induction step:
+
+%   \begin{align*}
+%     x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)\\
+%            &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)\big[\wtB_{kk}v_k(t)\\
+%     &\ \ \ \  + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t)\big]\\
+%     &\overset{(b)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell + 1}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
+%   \end{align*}
+
+%   Where equality $(a)$ follows by expanding each $x_k$ with equation \ref{eqn:parent_expansion} and $(b)$ requires the inversion of $\big(1 - \sum_{k \in \gpn{\ell}{i}\setminus\{i\}}\wtB_{ik}(z)\wtB_{ki}(z)\big)$ as well as using the fact that $\ancn{\ell}{i} = \ancn{\ell - 1}{i}\cup\gpn{\ell}{i}$.
+We here remind the reader of our earlier established convention that $\wtB$ simply represents the existence of a filter (possibly 0) and can change from place to place.
+
+% The induction terminates after $n$ steps since in a graph with $n$ nodes $\gpn{n + 1}{i} = \emptyset$
+
+% % Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.
+
+% %   \begin{align*}
+% %     x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)\\
+% %            &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)\big[\wtB_{kk}v_{k}(t) + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t) \big]\\
+% %            &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)v_{k}(t) + \sum_{k \in \gpn{2}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)
+% %   \end{align*}
+
+% % Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.
+
+% \end{proof}
+% \textbf{this is an alternative proof, I think superior to the one above}.
+
+The matrix $\B(z)$ is analagous to a \textit{weighted adjacency
+  matrix} for the graph $\gcg$.  And, in the same way that the
+$k^{\text{th}}$ power of an adjacency matrix counts the number of
+paths of length $k$ between nodes, $(\B(z)^k)_{ij}$ is a filter
+isolating the ``action'' of $j$ on $i$ at a time lag of $k$ steps.
+Though we only need the sparsity pattern of $\B(z)^k$ for the proof.
+
+We state the following well known theorem as a lemma, proof follows
+easily by induction.
+
+\begin{lemma}
+  \label{lem:adj_matrix}
+  Let $S$ be the transposed adjacency matrix of the Granger-causality
+  graph $\gcg$.  Then, $(S^k)_{ij}$ is the number of paths of length
+  $k$ from node $j$ to node $i$.  Evidently, if
+  $\forall k \in \N,\ (S^k)_{ij} = 0$ then $j \not\in \anc{i}$.
+\end{lemma}
+
+We now prove the proposition.
+
+\begin{proof}
+  From equation (\ref{eqn:ar_representation}), which we are assuming
+  throughout the paper to be invertible, we can write
+
+  \begin{equation*}
+    x(t) = (I - \B(z))^{-1} v(t),
+  \end{equation*}
+
+  where $(I - \B(z))^{-1} = \A(z)$ due to the uniqueness of
+  (\ref{eqn:wold}).  Since $\B(z)$ is stable, that is
+  $|\lambda_{\text{max}}(B(z))| < 1$, for every $|z| \le 1$ we have
+
+  \begin{equation}
+    \label{eqn:resolvant_inv}
+    (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k.
+  \end{equation}
+
+  Now, let $S$ be a $0-1$ matrix containing the sparsity pattern of
+  $\B(z)$.  Then from lemma \ref{lem:adj_matrix} $S$ is the transpose
+  of the adjacency matrix for $\gcg$ and hence $(S^k)_{ij}$ is
+  non-zero if and only if $j \in \gpn{k}{i}$, and therefore
+  $\B(z)^k_{ij} = 0$ if $j \not \in \gpn{k}{i}$.  The Cayley-Hamilton
+  theorem can then be applied to represent the infinite sum of equation
+  (\ref{eqn:resolvant_inv}) in terms only of \textit{finite} powers of
+  $\B(z)$, so the application of lemma \ref{lem:adj_matrix} in the
+  limit is not necessary.  Finally, since $\anc{i} = \cup_{k = 1}^n\gpn{k}{i}$
+  we see that $\A(z)_{ij}$ is zero if $j \not\in \anc{i}$.
+
+  Thence
+
+  \begin{align*}
+    x_i(t) &= (I - \B(z))^{-1}\\
+    &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
+    &= \A_{ii}(z) v_i(t) + \sum_{\substack{j \in \anc{i} \\ j \ne i}} \A_{ij}(z) v_j(t)\\
+  \end{align*}
+\end{proof}
+
+Before moving onto our next proposition, we need two lemmas formalizing
+some intuitively apparent facts.
+
+\begin{lemma}
+  \label{lem:subspace_sum_projection}
+  Let $x \in \H$ where $\H$ is a Hilbert space having inner product $\inner{\cdot}{\cdot}$ and let $\{\H_n\}_{n = 0}^\infty$ be a collection of subspaces of $\H$ such that $x \perp \H_0$.  Then
+
+  \begin{equation*}
+    \linE{x}{\bigvee_{n = 0}^\infty\H_n} = \linE{x}{\bigvee_{n = 1}^\infty\H_n},
+  \end{equation*}
+
+  where $\H_2 \vee \H_2 = \cl \{\alpha + \beta\ |\ \alpha \in \H_1, \beta \in \H_2 \}$ is the closed sum of subspaces.
+\end{lemma}
+\begin{proof}
+  We can apply the Gram-schmidt process to the collection of subspaces to obtain a sequence of orthogonal subspaces $\widetilde{\H}_0, \widetilde{\H}_1, \ldots$ with $\widetilde{\H}_0 = \H_0$, $\widetilde{\H}_n \subseteq \H_n$, and $\bigvee_{n = 0}^\infty\H_n = \bigoplus_{n = 0}^\infty \widetilde{\H}_n$, where $\oplus$ indicates the orthogonal sum of subspaces (which is always closed).
+Then, we have an explicit representation of the projection
+
+  \begin{align*}
+    \linE{x}{\bigvee_{n = 0}^\infty\H_n} &= \sum_{n = 0}^\infty \inner{x}{x_n} x_n\\
+    &\overset{(a)}{=} \sum_{n = 1}^\infty \inner{x}{x_n} x_n,
+  \end{align*}
+
+  where $x_n \in \widetilde{\H}_n \subseteq \H_n$ and $(a)$ follows because of our assumption that $x \perp \H_0$.
+This establishes that the projection is infact a member of the smaller subspace $\bigvee_{n = 1}^\infty\H_n$, and since the error can only increase by projecting onto this smaller subspace, we have the conclusion.
+\end{proof}
+
+The following lemma is reproduced from \cite{lindquist} here for convenience.
+
+\begin{lemma}[\cite{lindquist}]
+  \label{lem:conditional_orthogonality_equivalence}
+  Consider three closed subspaces of a Hilbert space $\mathcal{A}$, $\mathcal{B}$, $\mathcal{X}$.  The following statements are equivalent
+
+  \begin{enumerate}
+    \item{$\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$}
+    \item{$\linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}}\ \forall \beta \in \mathcal{B}$}
+    % \item{$\linE{\beta}{\mathcal{A}} = \linE{\linE{\beta}{\mathcal{X}}}{\mathcal{A}}\ \forall \beta \in \mathcal{B}}$}
+  \end{enumerate}
+\end{lemma}
+\begin{proof}
+  Proceding from the definition of conditional orthogonality we have
+
+  \begin{align*}
+    &\inner{\alpha - \linE{\alpha}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
+    &{\iff} \inner{\alpha + x - \linE{\alpha + x}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
+    &\overset{(a)}{\iff} \inner{\alpha + x}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
+    &\overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A} \vee \mathcal{X}} = 0\ \forall \beta \in \mathcal{B}\\
+    &\overset{(c)}{\iff} \linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}},
+  \end{align*}
+
+  where $(a)$ follows since $(\beta - \linE{\beta}{\mathcal{X}}) \perp \mathcal{X}$ and $\linE{\alpha + x}{\mathcal{X}} \in \mathcal{X}$, $(b)$ since $z \perp \mathcal{X} \vee \mathcal{A} \iff \linE{z}{\mathcal{X} \vee \mathcal{A}} = 0$, and $(d)$ since $\linE{z}{\H_2} = \linE{\linE{z}{\H_2}}{\H_1}$ for $\H_2 \subseteq \H_1$.
+
+  % This is property (vi) in Lindquist
+  % \begin{align*}
+  %   \langle \alpha - \linE{\alpha}{\mathcal{X}}, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
+  %   \overset{(a)}{\iff} \langle \alpha, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0\ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
+  %   \overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A}} &= 0\ \forall \beta \in \mathcal{B}
+  % \end{align*}
+
+  % where $(a)$ follows because $\linE{\alpha}{\mathcal{X}} \in \mathcal{X}$ and $\beta - \linE{\beta}{\mathcal{X}} \in \mathcal{X}$, and $(b)$ follows since $\beta \perp \mathcal{A} \iff \linE{\beta}{\mathcal{A}} = 0$.
+\end{proof}
+
+\begin{definition}[\hl{Confounder}]
+  A node $k$ will be referred to as a \textit{confounder} of nodes
+  $i, j$ (neither of which are equal to $k$) if
+  $k \in \anc{i} \cap \anc{j}$ and there exists a path
+  $\gcgpath{k}{i}$ not containing $j$, and a path $\gcgpath{k}{j}$
+  not containing $i$.
+\end{definition}
+
+\begin{proposition}
+  \label{prop:ancestor_properties}
+  In a Granger-causality graph $\gcg$, if $j \pwgc i$ then at least
+  one of the following must hold
+
+  \begin{enumerate}
+    \item{$j \in \anc{i}$}
+    \item{$\anc{i} \cap \anc{j} \ne \emptyset$.}
+  \end{enumerate}
+
+  Moreover, if $j \not\in \anc{i}$ then $\exists u \in \anc{i} \cap \anc{j}$ which is a confounder, that is, such that $i$ is not contained in at least one $\gcgpath{u}{j}$ path.
+\end{proposition}
+
+\begin{remark}
+  Considering the contrapositive of the first part of this theorem (which is what we actually prove below) is enlightening: it says that if $i$ is not a descendent of $j$ (i.e.
+$j \not \in \anc{i}$) and that there are no common ancestors (i.e.
+$\anc{i} \cap \anc{j} = \emptyset$) then $j \not \pwgc i$, or $\linE{x_i(t)}{\H_{t - 1}^{(i, j)}} = \linE{x_i(t)}{\H_{t - 1}^{(i)}}$, which is an intuitively natural result.
+
+  The first part is used in part to prove the second (and stronger) statement, which allows common ancestors, but only if they are ``mediated'' by $i$.
+For example, this covers the case of a graph $k \rightarrow i \rightarrow j$, where $\{k\} = \anc{i} \cap \anc{j}$, but clearly $j \not \pwgc i$.
+
+  An ancestor such as that in the second part of the theorem can be referred to as a \textit{confounder}, and a simple example is given by the fork $i \leftarrow k \rightarrow j$.
+\end{remark}
+
+\begin{proof}
+  We will first establish the contrapositive of $(1)$ and $(2)$: assume that $j \not \in \anc{i}$ and $\anc{i} \cap \anc{j} = \emptyset$, then we will show that $\forall \tau \in \Z_+,\ \E x_i(t)x_j(t - \tau) = 0$.
+This follows from representation (\ref{eqn:parent_expansion}) since we have
+
+  \begin{align*}
+    \E x_i(t)x_j(t - \tau) &= \E \big(\A_{ii}(z)v_i(t)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)\\
+    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\E[\big(\A_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
+    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ik}(z)v_k(t)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)]\\
+  \end{align*}
+
+  Keeping in mind that $v(t)$ is an isotropic and uncorrelated sequence (i.e.
+$\forall \tau \in \Z,\ k \ne \ell \implies \E v_k(t)v_\ell(t - \tau) = 0$ and $\forall \tau \in \Z\setminus\{0\}\ \E v_i(t)v_i(t - \tau) = 0$), we will see that each of these above four terms are 0.
+The first term since $i \pwgc i$ is not possible and hence the case $j = i$ is vacuous.
+The second term is 0 since $j \not\in \anc{i}$.
+The fourth term above is 0 since $\anc{i} \cap \anc{j} = \emptyset$.
+For the third term, it is possible that (due to cycles) $i \in \anc{j}$ which requires more care, although similarly as before we can at least see immediately that
+
+  \begin{equation*}
+    \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)] = \E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)].
+  \end{equation*}
+
+  However, $k \ne \ell \implies \A_{k\ell}(z) = 0$ (i.e.
+$\A(z) = I + A(1)z^{-1} + \cdots$) so $\A_{ji}(z)v_i(t - \tau)$ is strictly earlier in time, for $\tau \ge 0$, than $v_i(t)$ and hence $\E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)] = 0$.
+This establishes the fact that $H_t^{(i)} \perp \H_t^{(j)}$ where $H_t^{(i)} = \{\alpha x_i(t)\ |\ \alpha \in \R\}$ is the Hilbert space generated by the current sample of $x_i$.
+It is clear that $\H_t^{(i)} = \bigvee_{\tau = 0}^\infty H_{t - \tau}^{(i)}$.
+
+  \hl{TODO: Separate $H_t^i \perp \H_t^j \implies \H_t^i \perp \H_{t - 1}^j\ |\ \H_{t - 1}^i$ into a lemma and consolidate it with lemma 2, 3.}
+
+  Then for every $s \ge 0$
+
+  \begin{align*}
+    \linE{x_j(t - s)}{\H_t^{(i)}} &= \linE{x_j(t - s)}{\bigvee_{\tau = 0}^\infty H_{t - \tau}^{(i)}}\\
+    &\overset{(a)}{=} \linE{x_j(t - s)}{\bigvee_{\tau = 1}^\infty H_{t - \tau}^{(i)}}\\
+    &= \linE{x_j(t - s)}{\H_{t - 1}^{(i)}},
+  \end{align*}
+
+  where $(a)$ follows from lemma (\ref{lem:subspace_sum_projection}).  Lemma (\ref{lem:conditional_orthogonality_equivalence}) tells us that this is equivalent to
+
+  \begin{equation*}
+    \H_t^{(i)} \perp \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(i)},
+  \end{equation*}
+
+  so by theorem (\ref{thm:granger_causality_equivalences}) we conclude $j \not\pwgc i$.
+
+  For the second part of the theorem we will suppose that $j$ is a node having the following two properties: $(a)$ $j \not \in \anc{i}$ and $(b)$ for every $k \in \anc{i} \cap \anc{j}$ every $k \rightarrow \cdots \rightarrow j$ path contains $i$.
+The first is the negation of property $(1)$ above, and the second is to setup the contradiction $j \not\pwgc i$, which will then establish the theorem.
+
+  Firstly, notice that every $u \in \big(\pa{j} \setminus \{i\}\big)$ necesssarily inherits these same two properties since if $u \in \anc{i}$ then $u \in \anc{i} \cap \anc{j}$ so that every $u \rightarrow \cdots \rightarrow j$ path must contain $i$, but $u \in \pa{j}$, so this is not the case since $u \rightarrow j$ is a path that doesn't contain $i$; moreover, if we consider $w \in \anc{i} \cap \anc{u}$ then we also have $w \in \anc{i} \cap \anc{j}$ so every $w \rightarrow \cdots \rightarrow j$ path must contain $i$.
+These properties therefore extend inductively to every $u \in \big(\anc{j} \setminus \{i\}\big)$.
+
+  In order to deploy a recursive argument, define the following partition of $\pa{u}$, for some node $u$:
+
+  \begin{align*}
+    C_0(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, \anc{i} \cap \anc{k} = \emptyset, k \ne i\}\\
+    C_1(u) &= \{k \in \pa{u}\ |\ i \in \anc{k} \text{ or } k = i\}\\
+    C_2(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, \anc{i} \cap \anc{k} \ne \emptyset, k \ne i\}.
+  \end{align*}
+
+  We notice that for any $u$ having the properties $a, b$ above, we must have $C_2(u) = \emptyset$ since if $k \in C_2(u)$ then $\exists w \in \anc{i} \cap \anc{k}$ s.t.
+$i \not \in \anc{k}$ and therefore there must be a path $w \rightarrow \cdots \rightarrow k \rightarrow \cdots \rightarrow u$ which does not contain $i$.
+
+  Using this partition, we will expand $x_j(t)$ in terms of it's parents, and recursively expand nodes in $C_1$ until we reach a case where $C_1 = \emptyset$.
+For the first step:  %where $\A_{jj}(z) = (1 - \B_{jj}(z))^{-1}$:
+
+  \begin{equation}
+    \label{eqn:xj_partition_expansion}
+    x_j(t) = \A_{jj}(z)\Big(v_j(t) + \sum_{k \in C_0(j)}\B_{ik}(z)x_k(t) + \sum_{k \in C_1(j)}\B_{ik}(z)x_k(t)\Big).
+  \end{equation}
+
+  Using this representation we choose an arbitrary $\Phi(z) x_i(t - 1) \in \H_{t - 1}^{(j)}$ and show that
+
+  \begin{equation}
+    \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}{\Phi(z)x_j(t - 1) - \linE{\Phi(z)x_j(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
+  \end{equation}
+
+  which will imply (\hl{by lemma ???}) that $j \not\pwgc i$ and for which it is equivalent to show that
+
+  \begin{equation}
+    \label{eqn:sufficient_inner_prod}
+    \inner{x_i(t)}{\Phi(z)x_j(t - 1) - \linE{\Phi(z)x_j(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
+  \end{equation}
+
+  by the orthogonality principle since $\linE{x_i(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$.
+Substituting (\ref{eqn:xj_partition_expansion}) into (\ref{eqn:sufficient_inner_prod}) and starting with the first term we have
+
+  \begin{align*}
+    &\inner{x_i(t)}{\Phi(z)\A_{jj}(z)v_j(t - 1) - \linE{\Phi(z)\A_{jj}(z)v_j(t - 1)}{\H_{t - 1}^{(i)}}}\\
+    \overset{(\alpha)}{=}\ &\inner{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}{\Phi(z)\A_{jj}(z)v_j(t - 1)}\\
+    \overset{(\beta)}{=}\ &0,
+  \end{align*}
+
+  where $(\alpha)$ follows by expanding $x_i(t)$ with (\ref{eqn:ancestor_expansion}) and $\linE{\Phi(z)\A_{jj}(z)v_j(t - 1)}{\H_{t - 1}^{(i)}} = 0$ because $\forall \tau, s$
+
+  \begin{equation*}
+    \E v_j(t - \tau) x_i(t - s) = \E v_j(t - \tau) \sum_{k \in \anc{i} \cup \{i\}}\A_{ik}(z)v_k(t - s) = 0,
+  \end{equation*}
+
+  since $j \not \in \anc{i}$; $(\beta)$ follows similarly, that is, $j \not \in \anc{i}$.  Secondly we see that $\forall k \in C_0(j)$
+
+  \begin{equation*}
+    \inner{x_i(t)}{\Phi(z)\A_{jj}(z)\B_{ik}(z)x_k(t - 1) - \linE{\Phi(z)\A_{jj}(z)\B_{ik}(z)x_k(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
+  \end{equation*}
+
+  which follows from the first part of the theorem.
+Finally, for $k \in C_1(j)$ the case $k = i$ is immediate, so suppose $k \ne i$.
+We know from above that $k$ inherits the key properties referred to as $(a), (b)$ and therefore we can recursively expand $k$ in the same way as in equation (\ref{eqn:xj_partition_expansion}).
+Continuing this recursion for each $k \in C_1(j)$ where $k \ne i$ must eventually terminate since $i \in \anc{k}$.
+\end{proof}
+
+An interesting corollary is the following:
+
+\begin{corollary}
+  If the graph $\gcg$ is acyclic and if $j \pwgc i$ and $i \pwgc j$ then $\anc{i} \cap \anc{j} \ne \emptyset$.
+\end{corollary}
+
+It seems intuitive that a converse of proposition \ref{prop:ancestor_properties}
+would hold, i.e. $j \in \anc{i} \implies j \pwgc i$.  Unfortunately,
+this is not the case in general, as different paths through $\gcg$ can
+lead to cancellation (see example \ref{ex:diamond_cancellation}).  In
+fact, we do not even have $j \in \pa{i} \implies j \pwgc i$ (see
+example \ref{ex:lag_cancellation}).
+
+\begin{example}
+  \label{ex:diamond_cancellation}
+  Firstly, on $n = 4$ nodes, ``diamond'' shapes can lead to cancellation on paths of length 2:
+
+\begin{equation*}
+  x(t) =
+  \left[
+    \begin{array}{cccc}
+      0 & 0 & 0 & 0\\
+      a & 0 & 0 & 0\\
+      -a & 0 & 0 & 0\\
+      0 & 1 & 1 & 0\\
+    \end{array}
+  \right] x(t - 1) + v(t),
+\end{equation*}
+
+with $\E v(t) = 0,\ \E v(t)v(t - \tau)^\T = \delta_\tau I$.
+
+By directly calculating
+
+\begin{align*}
+  x_4(t) &= x_2(t - 1) + x_3(t - 1) + v_4(t)\\
+         &= ax_1(t - 2) + av_2(t - 1) - ax_1(t - 2) -av_3(t - 1) + v_4(t)\\
+         &= a(v_2(t - 1) - v_3(t - 1)) + v_4(t),
+\end{align*}
+
+we see that, since $v(t)$ is isotropic white noise, $1 \not\pwgc 4$.  The problem here is that there are multiple paths from $x_1$ to $x_4$.
+\end{example}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{0.45\textwidth}
+    \caption{Graph Corresponding to Example \ref{ex:diamond_cancellation}}
+    \label{fig:diamond_cancellation}
+    \includegraphics[width=\linewidth]{example1.pdf}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \caption{Graph Corresponding to Example \ref{ex:lag_cancellation}}
+    \label{fig:lag_cancellation}
+    \includegraphics[width=\linewidth]{example2.pdf}
+  \end{subfigure}
+\end{figure}
+
+\begin{example}
+  \label{ex:lag_cancellation}
+  A second example on $n = 3$ nodes is also worth examining, in this case
+  cancellation is a result of differing time lags.
+
+\begin{equation*}
+  x(t) =
+  \left[
+    \begin{array}{ccc}
+      0 & 0 & 0\\
+      -a & 0 & 0\\
+      0 & 1 & 0\\
+    \end{array}
+  \right] x(t - 1) +
+  \left[
+    \begin{array}{ccc}
+      0 & 0 & 0\\
+      0 & 0 & 0\\
+      a & 0 & 0\\
+    \end{array}
+  \right] x(t - 2) + v(t)
+\end{equation*}
+
+Then
+
+\begin{align*}
+  x_2(t) &= v_2(t) - ax_1(t - 1)\\
+  x_3(t) &= v_3(t) + x_2(t - 1) + ax_1(t - 2)\\
+  \implies x_3(t) &= v_2(t - 1) + v_3(t),
+\end{align*}
+
+and again $1 \not\pwgc 3$.
+\end{example}
+
+\subsection{Strongly Causal Graphs}
+The examples at the conclusion of the previous section seem rather
+pathological since we have simply constructed a case where different
+paths in the graph cancel exactly.  It may be possible to rid
+ourselves of these pathologies for instance by defining a notion of
+``robust'' pairwise causality where ``$j \pwgc _R i$'' if $j \pwgc i$
+almost surely for a ``small'' random perturbation of the system
+matrix.  However, we do not believe such a modification of our theory
+can have useful practical implications since the difference between
+paths that ``nearly cancel'' rather than cancel exactly would in
+practice require inordinate amounts of data to resolve.
+
+Alternatively, we could restrict the coefficients of the system
+matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead, we
+think it is more meaningful to focus on the defining feature of time
+series networks, that is, the topology of $\gcg$.
+
+\begin{definition}[Strongly Causal]
+  We will say that a Granger-causality graph $\gcg$ is \textit{strongly causal}
+if there is at most 1 directed path between any two nodes.
+
+\hl{TODO: Should DAG be included in this definition?  I think it is
+  not necessary for some theorems.}
+\end{definition}
+
+Examples of strongly causal graphs include directed trees (or
+forests), and DAGs where each node has at most one parent.
+Fortunately, the strongly causal property is inherited by subgraphs:
+
+\begin{lemma}
+  \label{lem:still_strongly_causal}
+  If $\gcg$ is a strongly causal graph, any subgraph formed by eliminating nodes
+  as well as all of the in and out edges thereof is still strongly causal.
+\end{lemma}
+\begin{proof}
+  If $\gcg$ has at most one path between any two nodes, there can only
+  be fewer paths after removing nodes from $\gcg$.
+\end{proof}
+
+We explore a number of properties of strongly causal graphs before
+moving into the main result of section \ref{sec:theory} in section
+\ref{sec:pairwise_algorithm}.  The following important property
+essentially strengthens proposition \ref{prop:ancestor_properties}
+for the case of strongly causal graphs.
+
+\begin{proposition}
+  \label{prop:sc_graph_common_anc}
+  In a strongly causal graph if $j \in \anc{i}$ then any
+  $k \in \anc{i} \cap \anc{j}$ is not a confounder, that is,
+  the unique path from $k$ to $i$ contains $j$.
+\end{proposition}
+\begin{proof}
+  Suppose that there is a path from $k$ to $i$ which does not contain
+  $j$.  In this case, there are multiple paths from $k$ to $i$ (one of
+  which \textit{does} go through $j$ since $j \in \anc{i}$) which
+  contradicts the assumption of strong causality.
+\end{proof}
+
+% Is pwgc transitive?
+
+\begin{corollary}
+  \label{cor:parent_corollary}
+  If $\gcg$ is a strongly causal DAG then $i \pwgc j$ and $j \in \anc{i}$ are
+  \textit{alternatives}, that is $i \pwgc j \implies j \notin \anc{i}$.
+\end{corollary}
+\begin{proof}
+  Suppose that $i \pwgc j$ and $j \in \anc{i}$.  Then since $\gcg$ is
+  acyclic $i \not\in \anc{j}$, and by proposition
+  \ref{prop:ancestor_properties} there is some
+  $k \in \anc{i}\cap\anc{j}$ which is a confounder.  However, by
+  proposition \ref{prop:sc_graph_common_anc} $k$ cannot be a
+  confounder, a contradiction.
+\end{proof}
+% This is a direct proof
+  % \begin{proof}
+%   Suppose that $\gcg$ is a strongly causal DAG and that we have both
+%   $i \pwgc j$ and $j \in \anc{i}$, which implies that
+%   $i \not \in \anc{j}$ since $\gcg$ is a DAG.  We will establish the
+%   contradiction $i \in \anc{j}$.
+
+%   Since $j \in \anc{i}$ there is a path $\gcgpath{j}{i}$.
+%   Moreover, since $i \pwgc j$ by proposition \ref{prop:pwgc_anc} there
+%   must be a confounding ancestor $k \in \anc{i} \cap \anc{j}$, where the
+%   node $k$ has a path to $j,\ \gcgpath{k}{j}$, as well as a path
+%   to $i,\ \gcgpath{k}{i}$.
+
+%   \hl{Double check that the ancestor properties proposition implies
+%     that BOTH paths do not contain the other node.}
+
+%   Now, since the graph is strongly causal, the only possible
+%   $\gcgpath{k}{i}$ path is the one obtained by concatenating the
+%   $\gcgpath{k}{j}$ path with the $\gcgpath{j}{i}$ path.
+%   However, this is a contradiction since $k$ is a confounder and
+%   should have a $\gcgpath{k}{i}$ path which does not contain $i$.
+% \end{proof}
+
+\begin{corollary}
+  \label{cor:bidirectional_edge}
+  If $\gcg$ is a strongly causal DAG such that $i \pwgc j$ and
+  $j \pwgc i$, then $i \not\in \anc{j}$ and $j \not\in \anc{i}$.  In
+  particular, a pairwise bidirectional edge indicates the absense of
+  any edge in $\gcg$.
+\end{corollary}
+\begin{proof}
+  This follows directly from applying proposition
+  \ref{cor:parent_corollary} to $i \pwgc j$ and $j \pwgc i$.
+% This is a direct proof
+% \begin{proof}
+  % By way of contradiction, suppose that $i \in \pa{j}$.  We will
+  % consider the two possibilities allowed by proposition
+  % \ref{prop:ancestor_properties} for $j \pwgc i$.  Firstly
+  % $j \in \anc{i}$ is impossible since $\gcg$ is assumed to be acyclic.
+  % Secondly, if there is some confounding node
+  % $u \in \anc{i} \cap \anc{j}$ with a path
+  % $u \rightarrow \cdots \rightarrow j$ which does not contain $i$ we
+  % have a contradiction since there must now be multiple
+  % $u \rightarrow j$ paths: the aforementioned, and a path
+  % $u \rightarrow \cdots \rightarrow i \rightarrow \cdots \rightarrow
+  % j$ which \textit{does} contain $i$.  We conclude that $i \in \pa{j}$
+  % is impossible, and symmetrically that $j \in \pa{i}$ is as well.
+% \end{proof}
+\end{proof}
+
+In light of proposition \ref{prop:sc_graph_common_anc}, the following
+provides a partial converse to proposition \ref{prop:ancestor_properties}.
+
+\begin{proposition}
+  \label{prop:pwgc_anc}
+  If $\gcg$ is a strongly causal DAG then $j \in \anc{i} \implies j \pwgc i$.
+\end{proposition}
+\begin{proof}
+  We will show that for some $\psi \in \H_{t - 1}^{(j)}$ we have
+
+  \begin{equation}
+    \label{eqn:cond_ortho_proof}
+    \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} \ne 0
+  \end{equation}
+
+  and therefore that $H_t^{(i)} \not\perp\ \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(i)}$, which by theorem (\ref{thm:granger_causality_equivalences}) and \hl{lemma [name the lemma]} is enough to establish that $j \pwgc i$.
+
+  Firstly, we will establish a representation of $x_i(t)$ that involves $x_j(t)$.
+Denote by $a_{r + 1} \rightarrow a_r \rightarrow \cdots \rightarrow a_1 \rightarrow a_0$ with $a_{r + 1} \defeq j$ and $a_0 \defeq i$ the \textit{unique} $\gcgpath{j}{i}$ path in $\gcg$, we will expand the representation of equation (\ref{eqn:parent_expansion}) backwards along this path:
+
+  \begin{align*}
+    x_i(t) &= v_i(t) + \B_{ii}(z) x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z) x_k(t)\\
+           &= \underbrace{v_{a_0}(t) + \B_{a_0a_0}(z) x_i(t) + \sum_{\substack{k \in \pa{a_0} \\ k \ne a_1}}\B_{a_0 k}(z) x_k(t)}_{\defeq \wtalpha{a_0}{a_1}} + \B_{a_0a_1}(z)x_{a_1}(t)\\
+           &= \wtalpha{a_0}{a_1} + \B_{a_0a_1}(z)\big[\wtalpha{a_1}{a_2} + \B_{a_1a_2}(z)x_{a_2}(t) \big]\\
+           &\overset{(a)}{=} \cdots\\
+           &= \sum_{\ell = 0}^r \underbrace{\Big(\prod_{m = 0}^{\ell - 1} \B_{a_m a_{m + 1}}(z) \Big)}_{\defeq F_\ell(z)} \wtalpha{a_\ell}{a_{\ell + 1}} + \Big(\prod_{m = 0}^{r}\B_{a_m a_{m + 1}}(z)\Big)x_{a_{r + 1}}(t)\\
+           &= \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}} + F_{r + 1}(z) x_j(t)
+  \end{align*}
+
+  where $(a)$ follows by a routine induction argument and where we define $\prod_{m = 0}^{-1} \bullet \defeq 1$ for notational convenience.
+
+  Using this representation to expand equation (\ref{eqn:cond_ortho_proof}), we obtain the following cumbersome expression:
+
+  \begin{align*}
+    &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}}\\
+    &- \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\linE{\sum_{\ell = 0}^r F_\ell(z)\wtalpha{a_\ell}{a_{\ell + 1}}}{\H_{t - 1}^{(i)}}}\\
+    &+ \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\sum_{\ell = 0}^r F_\ell(z)\wtalpha{a_\ell}{a_{\ell + 1}}}.
+  \end{align*}
+
+  Note that by the orthogonality principle, $\psi - \linE{\psi}{\H_{t - 1}^{(i)}} \perp \H_{t - 1}^{(i)}$, the middle term above is $0$.
+Choosing now the particular value $\psi = F_{r + 1}(z)x_j(t) \in \H_{t - 1}^{(j)}$ we arrive at
+
+  \begin{align*}
+    &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}\\
+    &= \E|F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}|^2\\
+    &+ \inner{F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}}{\sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}}},
+  \end{align*}
+
+  which by the Cauchy-Schwarz inequality is $0$ if and only if
+
+  \begin{equation*}
+    \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}} \overset{\text{a.s.}}{=} \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}} - F_{r + 1}(z)x_j(t),
+  \end{equation*}
+
+  or by rearranging and applying the representation obtained earlier, if and only if
+
+  \begin{equation*}
+    x_i(t) \overset{\text{a.s.}}{=} \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}},
+  \end{equation*}
+
+  but this is impossible since with probability one $x_i(t) \not \in \H_{t - 1}^{(i)}$.
+\end{proof}
+
+We immediately obtain the corollary, which we remind the reader is,
+surprisingly, not true in a general graph.
+
+\begin{corollary}
+  \label{cor:gc_implies_pwgc}
+  If $\gcg$ is a strongly causal DAG then $j \gc i \implies j \pwgc i$.
+\end{corollary}
+
+\begin{remark}
+  It can be seen from the conclusion of the proof that simple facts
+  about Granger causality are indeed reliant on our assumptions about
+  the nature of $x(t)$ laid out in the section \ref{sec:theory}, in
+  particular, the innovations process $v(t)$ must be full rank.
+
+  Moreover, that strong conditions need to be placed on the topology
+  of $\gcg$ in order for proposition \ref{prop:pwgc_anc} to hold can
+  be seen through the earlier examples.
+\end{remark}
+
+\begin{lemma}
+  If $\gcg$ is strongly causal, then any $\VAR(p)$ system on $\gcg$ is stable if and only if
+  $B_{ii}(z)$ is stable for every $i = 1, \ldots, n$.
+\end{lemma}
+\begin{proof}
+  \hl{This is certainly true, but I don't think I need to use it anywhere.}
+\end{proof}
+
+\begin{remark}
+  \hl{Possibly expand on this thought}
+
+  As in much of statistics, confounding nodes pose challenges for
+  Granger-causality.  However, as opposed to Pearl's causal calculus
+  \cite{pearl2000art}, pairwise Granger-causality does not suffer any
+  difficulty with so-called ``colliders'', that is, the topology
+  $i \rightarrow k \leftarrow j$ will never result in $i \pwgc j$ or
+  $j \pwgc i$.  This is evidently an advantage of the \textit{temporal}
+  nature of Granger causality -- ``information'' cannot flow backwards
+  along the edges of $\gcg$.
+\end{remark}
+
+\begin{example}
+  As a final remark of this subsection we note that a complete
+  converse to proposition \ref{prop:ancestor_properties} is not
+  possible without additional conditions.  Consider the ``fork'' system on $3$
+  nodes (i.e. $2 \leftarrow 1 \rightarrow 3$) defined by
+
+  \begin{equation*}
+    x(t) =
+    \left[
+      \begin{array}{cccc}
+        0 & 0 & 0\\
+        a & 0 & 0\\
+        a & 0 & 0\\
+      \end{array}
+    \right] x(t - 1) + v(t).
+  \end{equation*}
+
+  In this case, node $1$ is a confounder for nodes $2$ and $3$, but
+  $x_3(t) = v_3(t) - v_2(t) + x_2(t)$ and $2 \not \pwgc 3$ (even
+  though $x_2(t)$ and $x_3(t)$ are contemporaneously correlated)
+
+  If we were to augment this system by simply adding an autoregressive
+  component to $x_1(t)$ e.g.  $x_1(t) = v_1(t) + b x_1(t - 1)$ then we
+  \textit{would} have $2 \pwgc 3$ since then
+  $x_3(t) = v_3(t) + av_1(t - 1) - bv_2(t - 1) + bx_2(t - 1)$.
+\end{example}
+
+\subsection{Recovering $\gcg$ via Pairwise Tests}
+\label{sec:pairwise_algorithm}
+In this section we will show that if $\gcg$ is a strongly causal DAG,
+then it is possible to recover $\gcg$ via pairwise tests alone.  This
+is an interesting fact in and of itself, but also has some
+implications for applications since pairwise testing is trivially
+parallelizable.  In section \ref{sec:structure_learning} we will also
+analyze the use of pairwise testing as a heuristic for general graphs.
+
+\begin{theorem}[Pairwise Recovery]
+  \label{thm:tree_recovery}
+  If the Granger-Causality graph $\gcg$ for $x(t)$ is a strongly
+  causal DAG then $\gcg$ can be inferred from pairwise causality
+  tests alone.  The procedure can be carried out, assuming we have an
+  oracle for pairwise causality, via algorithm (\ref{alg:pwgr}).
+
+  \begin{algorithm}
+    \SetKwInOut{Input}{input}
+    \SetKwInOut{Output}{output}
+    \SetKwInOut{Initialize}{initialize}
+    \DontPrintSemicolon
+
+    \BlankLine
+    \caption{Pairwise Graph Recovery}
+    \label{alg:pwgr}
+    % \TitleOfAlgo{Pairwise Graph Recovery}
+    \Input{Pairwise Granger-causality relations between $n$ processes whose joint Granger-causality relations are known to form a strongly causal DAG $\gcg$.}
+    \Output{Edges $E = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of the graph $\gcg = ([n], E)$.}
+    \Initialize{$S_0 = [n]$  \texttt{\# unprocessed nodes}\\
+      $E = \emptyset$  \texttt{\# edges of }$\gcg$\\
+      % $P_0 = \emptyset$  \texttt{\# layer by layer driving nodes}\\
+      $k = 0$ \texttt{\# a counter used only for notation}}
+    \BlankLine
+    $W_0 \leftarrow \{(i, j)\ |\ i \pwgc j, j \not\pwgc i\}$  \texttt{\# potential edges}\\
+    \While{$S_k \ne \emptyset$}{
+      $\widehat{P}_k \leftarrow \{i \in S_{k - 1}\ |\ \forall s \in S_{k - 1}\ (s, i) \not\in W_k\}$  \texttt{\# parents of current layer}\\
+      % $P_k \leftarrow P_{k - 1} \cup \widehat{P}_k$ \texttt{\# Expand parent set}\\
+      $S_{k + 1} \leftarrow S_k \setminus \widehat{P}_k$ \texttt{\# remove from } $S$\\
+
+      \;
+      $C_k \leftarrow \{j \in S_{k + 1}\ |\ \forall s \in S_{k + 1}\ (s, j) \not\in W_k\}$   \texttt{\# potential children of }$P_k$\\
+
+      \For{$r = 0, \ldots, k$}
+      {
+        $D_{kr} \leftarrow \{(i, j) \in \widehat{P}_{k - r} \times C_k\ |\ (i, j) \in W_k,\ \text{no } i \rightarrow j \text{ path in } E \}$\\%  \texttt{\# edges incident on }$C_k$\\
+        $E \leftarrow E \cup D_{kr}$ \texttt{\# add edges to }$E$\\
+        % $W_{k + 1} \leftarrow W_k \setminus D_{kr}$  \texttt{\# remove edges from consideration}\\
+      }
+      $W_{k + 1} \leftarrow W_k \setminus \big(\bigcup_{r = 0}^k D_{kr}\big)$  \texttt{\# remove edges from consideration}\\
+
+      \;
+      $k \leftarrow k + 1$
+    }
+    \Return{$E$}
+  \end{algorithm}
 \end{theorem}
 
-\begin{theorem}[General Invariance]
-  \hl{Under what conditions is this actually true?}
+We prove the theorem by establishing the correctness of algorithm (\ref{alg:pwgr}).
+The idea is to iteratively ``peel away layers'' of nodes by removing the nodes that have no parents remaining, which always exist since the graph is acyclic.
+The additional requirement of strong causality ensures firstly that all actual edges of $\gcg$ manifest in some way as pairwise relations (by proposition \ref{prop:pwgc_anc}), and secondly, as we will see, that parents of $C_k$ are in $\bigcup_{r = 0}^k\widehat{P}_{k - r}$ and can be determined by a pairwise test.
 
-  Let $\zeta(t)$ be a stationary discrete time stochastic process.  Let
-  $F, G$ be (possibly nonlinear and time varying) invertible filtering
-  operations and $f(t)$, $g(t)$ be perfectly predictable and such that
-  $x_j(t) \defeq F(\zeta_j - f)(t), x_i(t) \defeq G(\zeta_i - g)(t)$ are W.S.S.  Then,
+\begin{example}
+  That we need to proceed backwards through $\widehat{P}_{k - r}$ can be seen from the example of figure \ref{fig:example_fig3}, where if $D_k = \{(i, j) \in \Big(\bigcup_{r = 0}^k \widehat{P}_{k - r}\Big) \times C_k\ |\ i \pwgc j \}$ above we would infer the false positive edge $2 \rightarrow 3$.
+Moreover, the same example shows that if simply $D_k = \{(i, j) \in \widehat{P}_k \times C_k\ |\ i \pwgc j \}$, then the edge $1 \rightarrow 3$ would be missed.
+
+\hl{TODO: Inkscape figure}
+
+  \begin{figure}
+    \centering
+    \label{fig:example_fig3}
+    \caption{Example graph for Algorithm \ref{alg:pwgr}}
+    \includegraphics[width=0.5\linewidth]{hand_drawn_example3.jpg}
+  \end{figure}
+\end{example}
+
+\begin{lemma}
+  \hl{Some lemma on the inner loop.}
+\end{lemma}
+\begin{proof}
+  Let $j \in C_k$ and $i \in \bigcup_{r = 0}^k P_{k - r}$.
+
+  First, suppose that $i \in \pa{j}$ and that there is no $i \rightarrow j$ path in $E$.
+Then since $i \gc j$ we have $i \pwgc j$ by proposition \ref{prop:pwgc_anc} so at least one of $\{D_{kr} \}_{r = 0}^k$ contains $(i, j)$ and hence $(i, j)$ will be added to $E$ upon the end of the iteration.
+
+  Now, still assuming that $i \in \pa{j}$, suppose instead that there is already an $i \rightarrow j$ path in $E$.
+Using an induction hypothesis that $E$ contains edges if and only if they are in $\gcg$, then either $(i, j) \in E$ correctly, or there is more than one $i \rightarrow j$ path in $\gcg$, which is not possible since $\gcg$ is strongly causal.
+
+  Finally, suppose that $i \not \in \pa{j}$.
+If $i \not \pwgc j$ then we will correctly leave $(i, j)$ absent from $E$, so suppose that $i \pwgc j$.
+Then we must have either $i \in \anc{j}$ or $\anc{i} \cap \anc{j} \ne \emptyset$; suppose first the former.
+Then, there is some $u$ s.t.
+$i \pwgc u \pwgc j$, so by the construction of \hl{it seems} that both $(i, u)$ and $(u, j)$ should have already been added to $E$, so we have an $i \rightarrow j$ path in $E$ \hl{is this sound?}  Suppose then the latter case above, $u \in \anc{i} \cap \anc{j}$ \hl{try to construct a counterexample involving this case}.
+\end{proof}
+
+Proof of the actual theorem...
+
+\begin{proof}
+% 1. Dispense with trivialities
+% 2. Show that hat{P}_k is nonempty
+% 3. Show that C_k is nonempty
+% 4. Show that edges in D_k are actual edges in G
+% 5. show that each edge incident on C_k is in D_k
+% 6. Induction
+
+% Given non-empty P and C...  Can we show all P -> C edges get collected in D?
+  Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$ and let us define
+
+  \begin{equation*}
+    A_k = \{i \in [n]\ |\ \dist{i}{\widehat{P}_0} = k\},
+  \end{equation*}
+
+  where
+
+  \begin{equation*}
+    \dist{i}{P} = \text{max}\{r \in [n]\ |\ \exists (j \rightarrow i) \text{ path of length } r \text{ in } \gcg \text{ for some } j \in P\}.
+  \end{equation*}
+
+  We will proceed by induction and show that when the counter has reached the value $k \ge 1$, it is guaranteed that all nodes in $\bigcup_{\ell \le k}A_\ell$ (nodes having paths of length up to $k$ between themselves and $\widehat{P}_0$) must have all of their incident edges included in $E$, and that every edge in $E$ is an edge of $\gcg$.
+If this is the case, then it can be immediately seen that the algorithm terminates with the correct result, $\gcg = ([n], E)$, once $k = n$; since there can only be paths up to length $n - 1$ in $\gcg$, and there cannot be any nodes or edges that don't involve paths back to $\widehat{P}_0$.
+That is, $\bigcup_{\ell \le n}A_\ell = [n]$.
+
+  % firstly dispensing with some trivialities, and then establishing that every edge in $\gcg$ makes an appearance in some $D_{kr}$ as well as that everything in any $D_{kr}$ is indeed an edge in $\gcg$.
+
+  On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
+To quickly dispense with some trivialities: the set $\widehat{P}_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $\widehat{P}_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
+Now, if $\widehat{P}_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
+
+  Continuing with the non-trivial case: if $\widehat{P}_0 \ne [n]$ then $C_0$ must be non-empty, otherwise there would be driving nodes not included in $\widehat{P}_0$.
+In this case it is also clear that $A_1 \ne \emptyset$, so let us consider some $j \in A_1$.
+We must have $j \in C_0$, otherwise there would be a path of length $2$ from $\widehat{P}_0$ to $j$.
+Moreover, since $\exists i \in \widehat{P}_0$ s.t.
+$i \gc j$ and therefore $i \pwgc j$ (by proposition \ref{prop:pwgc_anc}) we have $(i, j) \in D_{00}$ and therefore $(i, j) \in E$.
+Therefore $A_1 \subseteq D_{00}$.
+\hl{This is nonsensical because $A_1$ does not contain tuples -- need to show that $A_1 \subseteq C_0$ ?}
+
+  Now consdider some $(i, j) \in D_{00}$ (the same reasoning as for $C_0$ is sufficient to see that $D_{00}$ is non-empty).
+Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).
+Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in \widehat{P}_0 \implies u \in S_0 \implies j \not\in C_0$, a contradiction.
+This implies every tuple in $D_{00}$ is a bona-fide edge of $\gcg$, as well as that $(i, j) \in A_1$.
+So then $A_1 = D_{00}$, [\hl{nonsensical, see earlier.}] completing the base case for induction.
+
+  Suppose now the algorithm has reached step $k$ and that $E = \bigcup_{\ell \le k - 1}A_\ell$.
+
+% The reasoning here is (I think) correct -- but isn't directly establishing the induction hypothesis.
+%   On the other hand, if $\widehat{P}_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $\widehat{P}_0$.
+Consider now an edge $(i, j) \in D_{00}$ ($D_{00}$ is non-empty for the same reason as $C_0$).
+Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).
+Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in \widehat{P}_0 \implies u \in S_0 \implies j \not\in C_0$, a contradiction.
+So every tuple in $D_{00}$ is a bona-fide edge of $\gcg$.
+Suppose now that for some $j \in C_0$ there is an $i$ s.t.
+$(i, j) \in \gcg$ but $(i, j) \not\in D_{00}$.
+Since $(i, j) \in \gcg$ we have $i \pwgc j$ (corollary \ref{cor:gc_implies_pwgc}), but then if $i \not\in \widehat{P}_0$ then $j \notin C_0$ and if $i \in \widehat{P}_0, j \in C_0$ then necessarily $(i, j) \in D_{00}$ since we cannot at this stage have any paths in $E$.
+
+%   and show that on each step, every edge incident upon a node in $C_k$ is contained in $D_k$, that every edge in $D_k$ is an edge in $\gcg$, and finally that $\bigcup_{k = 1}^n C_k = [n] \setminus \widehat{P}_0$ and that this implies $\big([n], \bigcup_{k = 1}^n D_k\big) = \gcg$.
+
+%  On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E_0 = \emptyset$.
+The set $\widehat{P}_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $\widehat{P}_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
+Now, if $\widehat{P}_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
+On the other hand, if $\widehat{P}_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $\widehat{P}_0$.
+Consider now an edge $(i, j) \in D_0$ (again $D_0$ is non-empty for the same reason as $C_0$).
+Since $i$ has no parents, we must have $\anc{i}\cap\anc{j} = \emptyset$ and therefore $i \in \anc{j}$, and due to the construction of $C_0$ we in fact have $i \in \pa{j}$, which establishes that every edge in $D_0$ is also an edge in $\gcg$.
+Finally consider $j \in C_0$, by proposition \ref{prop:pwgc_anc} and since the parents of $C_0$ must be in $P_0$, we see that every edge incident on $j$ must be present in $D_0$.
+
+% We also see that since nodes in $\widehat{P}_0$ have no parents,
+
+% we must have $\anc{i}\cap\anc{j} = \emptyset$, and therefore by proposition \ref{prop:ancestor_properties} $i \in \anc{j}$;  moreover, the construction of $\widehat{P}_1$ implies that in fact $i \in \pa{j}$, so $(i, j)$ is an edge of $\gcg$.
+
+%Finally, by similar reasoning, there cannot be an edge $(i, j) \in D_0$ which is not in $\gcg$, since $\anc{i} \cap \anc{j} = \emptyset$.
+
+%Consider now step $K < n$ of algorithm \ref{alg:pwgr}.
+By lemma \ref{lem:still_strongly_causal} the graph remains strongly causal after having removed nodes $\bigcup_{k = 0}^K P_k$ from $S$, and we assume for induction that $\bigcup_{k = 1}^K D_k$ are all edges of $\gcg$.
+
+\hl{WIP}
+\end{proof}
+
+\begin{example}
+  Being a strongly causal DAG is merely a sufficient condition.
+For example, the complete directed graph with 2 nodes (e.g.
+$B(1) = \left[ \begin{array}{cc} 1/2 & 1 \\ 1 & 1/2 \end{array} \right]$) contains a loop but is pairwise recoverable, though not by algorithm (\ref{alg:pwgr}).
+Clearly, this example is somewhat artificial since when $n = 2$ there is no difference between pairwise Granger-causality and joint Granger-causality amongst all series -- however, one can take any strongly causal DAG, and then add two additional nodes having such a loop without impacting the pairwise recoverability.
+\end{example}
+
+\section{Structure Learning}
+\label{sec:structure_learning}
+\subsection{Local Search Heuristics}
+
+\subsection{Edge-Wise Grouped LASSO}
+Minimize the sparsity promoting regularized least squares problem and choose the hyper-parameters against the Akaike information criteria.
 
 \begin{equation}
-    x_j \gc x_i \iff \zeta_j \te \zeta_i
+  L(\lambda, p) \defeq \underset{B}{\text{minimize}}\ \frac{1}{T}\sum_{t = 1}^T||x(t) - \sum_{\tau = 1}^pB(\tau)x(t - \tau)||_2^2 + \lambda \sum_{i, j}\big[\alpha||B_{i, j}||_2 + (1 - \alpha)||B_{i, j}||_1\big]
 \end{equation}
-\end{theorem}
 
-\begin{theorem}[Tree Recovery]
-  \label{thm:tree_recovery}
-  If the Granger-Causality graph $\mathcal{G}$ for $x(t)$ is a
-  directed forest, then $\mathcal{G}$ can be inferred from pairwise
-  causality tests alone.
-\end{theorem}
+\begin{equation}
+  \underset{\lambda, p}{\text{minimize}}\ L(\lambda, p) + \mathsf{AIC}(B^{(\lambda, p)})
+\end{equation}
 
-\section{Structure Learning}
-\subsection{Edge-Wise Grouped LASSO}
 \subsection{Bayesian Posterior Thresholding}

### 2019-05-02 paper/paper.tex
Some figures and minor updates.

@@
 \def\gc{\overset{\text{GC}}{\rightarrow}}  % Granger causality arrow
-\def\pwgc{\overset{\text{PW}}{\rightarrow}}  % Granger causality arrow
+\def\ngc{\overset{\text{GC}}{\nrightarrow}}  % Negated Granger causality arrow
+\def\pwgc{\overset{\text{PW}}{\rightarrow}}  % Pairwise Granger causality arrow
+\def\npwgc{\overset{\text{PW}}{\nrightarrow}}  % Negated pairwise Granger causality arrow
 \def\te{\overset{\mathcal{T}}{\rightarrow}}  % Transfer entropy arrow
@@
   \begin{enumerate}
-    \item{$x_j \not\gc x_i$}
+    \item{$x_j \ngc x_i$}
     \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$}
@@
 
-  $(1) \implies (2)$: The projection for $x_i(t)$ onto $\H_{t - 1}$ is
+  $(1) \Rightarrow (2)$: The projection for $x_i(t)$ onto $\H_{t - 1}$ is
   (by definition) given by a form similar to equation
@@
 
-$(2) \implies (4)$: In computing $(y - \linE{y}{\H_{t = 1}^{-j}})$ for
+$(2) \Rightarrow (4)$: In computing $(y - \linE{y}{\H_{t = 1}^{-j}})$ for
 $y \in \H_t^i$ it is sufficient to consider $y = x_i(t)$ since
@@
 
-\hl{TODO: $(4) \implies (1)$}
+\hl{TODO: $(4) \Rightarrow (1)$}
 
@@
     (1 - \B_{ii}(z))x_i(t) &= v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\\
-    \implies x_i(t) &= (1 - \B_{ii}(z))^{-1}\big[v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\big]\\
+    \Rightarrow x_i(t) &= (1 - \B_{ii}(z))^{-1}\big[v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\big]\\
     &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)
@@
   Keeping in mind that $v(t)$ is an isotropic and uncorrelated sequence (i.e.
-$\forall \tau \in \Z,\ k \ne \ell \implies \E v_k(t)v_\ell(t - \tau) = 0$ and $\forall \tau \in \Z\setminus\{0\}\ \E v_i(t)v_i(t - \tau) = 0$), we will see that each of these above four terms are 0.
+$\forall \tau \in \Z,\ k \ne \ell \Rightarrow \E v_k(t)v_\ell(t - \tau) = 0$ and $\forall \tau \in \Z\setminus\{0\}\ \E v_i(t)v_i(t - \tau) = 0$), we will see that each of these above four terms are 0.
 The first term since $i \pwgc i$ is not possible and hence the case $j = i$ is vacuous.
@@
 
-  However, $k \ne \ell \implies \A_{k\ell}(z) = 0$ (i.e.
+  However, $k \ne \ell \Rightarrow \A_{k\ell}(z) = 0$ (i.e.
 $\A(z) = I + A(1)z^{-1} + \cdots$) so $\A_{ji}(z)v_i(t - \tau)$ is strictly earlier in time, for $\tau \ge 0$, than $v_i(t)$ and hence $\E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)] = 0$.
@@
 
-  \hl{TODO: Separate $H_t^i \perp \H_t^j \implies \H_t^i \perp \H_{t - 1}^j\ |\ \H_{t - 1}^i$ into a lemma and consolidate it with lemma 2, 3.}
+  \hl{TODO: Separate $H_t^i \perp \H_t^j \Rightarrow \H_t^i \perp \H_{t - 1}^j\ |\ \H_{t - 1}^i$ into a lemma and consolidate it with lemma 2, 3.}
 
@@
 
-  so by theorem (\ref{thm:granger_causality_equivalences}) we conclude $j \not\pwgc i$.
+  so by theorem (\ref{thm:granger_causality_equivalences}) we conclude $j \npwgc i$.
 
   For the second part of the theorem we will suppose that $j$ is a node having the following two properties: $(a)$ $j \not \in \anc{i}$ and $(b)$ for every $k \in \anc{i} \cap \anc{j}$ every $k \rightarrow \cdots \rightarrow j$ path contains $i$.
-The first is the negation of property $(1)$ above, and the second is to setup the contradiction $j \not\pwgc i$, which will then establish the theorem.
+The first is the negation of property $(1)$ above, and the second is to setup the contradiction $j \npwgc i$, which will then establish the theorem.
 
@@
 
-  which will imply (\hl{by lemma ???}) that $j \not\pwgc i$ and for which it is equivalent to show that
+  which will imply (\hl{by lemma ???}) that $j \npwgc i$ and for which it is equivalent to show that
 
@@
 It seems intuitive that a converse of proposition \ref{prop:ancestor_properties}
-would hold, i.e. $j \in \anc{i} \implies j \pwgc i$.  Unfortunately,
+would hold, i.e. $j \in \anc{i} \Rightarrow j \pwgc i$.  Unfortunately,
 this is not the case in general, as different paths through $\gcg$ can
 lead to cancellation (see example \ref{ex:diamond_cancellation}).  In
-fact, we do not even have $j \in \pa{i} \implies j \pwgc i$ (see
+fact, we do not even have $j \in \pa{i} \Rightarrow j \pwgc i$ (see
 example \ref{ex:lag_cancellation}).
@@
 
-we see that, since $v(t)$ is isotropic white noise, $1 \not\pwgc 4$.  The problem here is that there are multiple paths from $x_1$ to $x_4$.
+we see that, since $v(t)$ is isotropic white noise, $1 \npwgc 4$.  The problem here is that there are multiple paths from $x_1$ to $x_4$.
 \end{example}
@@
   \begin{subfigure}[b]{0.45\textwidth}
+    \includegraphics[width=\linewidth]{example1.pdf}
     \caption{Graph Corresponding to Example \ref{ex:diamond_cancellation}}
     \label{fig:diamond_cancellation}
-    \includegraphics[width=\linewidth]{example1.pdf}
+    \centering{$j \in \anc{i} \nRightarrow j \pwgc i$}
   \end{subfigure}
   \begin{subfigure}[b]{0.45\textwidth}
+    \includegraphics[width=\linewidth]{example2.pdf}
     \caption{Graph Corresponding to Example \ref{ex:lag_cancellation}}
     \label{fig:lag_cancellation}
-    \includegraphics[width=\linewidth]{example2.pdf}
+    \centering{$j \in \pa{i} \nRightarrow j \pwgc i$}
   \end{subfigure}
@@
   x_3(t) &= v_3(t) + x_2(t - 1) + ax_1(t - 2)\\
-  \implies x_3(t) &= v_2(t - 1) + v_3(t),
+  \Rightarrow x_3(t) &= v_2(t - 1) + v_3(t),
 \end{align*}
 
-and again $1 \not\pwgc 3$.
+and again $1 \npwgc 3$.
 \end{example}
@@
   If $\gcg$ is a strongly causal DAG then $i \pwgc j$ and $j \in \anc{i}$ are
-  \textit{alternatives}, that is $i \pwgc j \implies j \notin \anc{i}$.
+  \textit{alternatives}, that is $i \pwgc j \Rightarrow j \notin \anc{i}$.
 \end{corollary}
@@
   \label{prop:pwgc_anc}
-  If $\gcg$ is a strongly causal DAG then $j \in \anc{i} \implies j \pwgc i$.
+  If $\gcg$ is a strongly causal DAG then $j \in \anc{i} \Rightarrow j \pwgc i$.
 \end{proposition}
@@
   \label{cor:gc_implies_pwgc}
-  If $\gcg$ is a strongly causal DAG then $j \gc i \implies j \pwgc i$.
+  If $\gcg$ is a strongly causal DAG then $j \gc i \Rightarrow j \pwgc i$.
 \end{corollary}
@@
     \Input{Pairwise Granger-causality relations between $n$ processes whose joint Granger-causality relations are known to form a strongly causal DAG $\gcg$.}
-    \Output{Edges $E = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of the graph $\gcg = ([n], E)$.}
+    \Output{Edges $E = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of the graph $\gcg$.}
     \Initialize{$S_0 = [n]$  \texttt{\# unprocessed nodes}\\
@@
     \BlankLine
-    $W_0 \leftarrow \{(i, j)\ |\ i \pwgc j, j \not\pwgc i\}$  \texttt{\# potential edges}\\
+    $W_0 \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$  \texttt{\# potential edges}\\
     \While{$S_k \ne \emptyset$}{
-      $\widehat{P}_k \leftarrow \{i \in S_{k - 1}\ |\ \forall s \in S_{k - 1}\ (s, i) \not\in W_k\}$  \texttt{\# parents of current layer}\\
-      % $P_k \leftarrow P_{k - 1} \cup \widehat{P}_k$ \texttt{\# Expand parent set}\\
-      $S_{k + 1} \leftarrow S_k \setminus \widehat{P}_k$ \texttt{\# remove from } $S$\\
+      $P_k \leftarrow \{i \in S_{k - 1}\ |\ \forall s \in S_{k - 1}\ (s, i) \not\in W_k\}$  \texttt{\# parents of current layer}\\
+      % $P_k \leftarrow P_{k - 1} \cup P_k$ \texttt{\# Expand parent set}\\
+      $S_{k + 1} \leftarrow S_k \setminus P_k$ \texttt{\# remove from } $S$\\
 
@@
       {
-        $D_{kr} \leftarrow \{(i, j) \in \widehat{P}_{k - r} \times C_k\ |\ (i, j) \in W_k,\ \text{no } i \rightarrow j \text{ path in } E \}$\\%  \texttt{\# edges incident on }$C_k$\\
+        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times C_k\ |\ (i, j) \in W_k,\ \text{no } i \rightarrow j \text{ path in } E \}$\\%  \texttt{\# edges incident on }$C_k$\\
         $E \leftarrow E \cup D_{kr}$ \texttt{\# add edges to }$E$\\
@@
 The idea is to iteratively ``peel away layers'' of nodes by removing the nodes that have no parents remaining, which always exist since the graph is acyclic.
-The additional requirement of strong causality ensures firstly that all actual edges of $\gcg$ manifest in some way as pairwise relations (by proposition \ref{prop:pwgc_anc}), and secondly, as we will see, that parents of $C_k$ are in $\bigcup_{r = 0}^k\widehat{P}_{k - r}$ and can be determined by a pairwise test.
+The additional requirement of strong causality ensures firstly that all actual edges of $\gcg$ manifest in some way as pairwise relations (by proposition \ref{prop:pwgc_anc}), and secondly, as we will see, that parents of $C_k$ are in $\bigcup_{r = 0}^kP_{k - r}$ and can be determined by a pairwise test.
 
 \begin{example}
-  That we need to proceed backwards through $\widehat{P}_{k - r}$ can be seen from the example of figure \ref{fig:example_fig3}, where if $D_k = \{(i, j) \in \Big(\bigcup_{r = 0}^k \widehat{P}_{k - r}\Big) \times C_k\ |\ i \pwgc j \}$ above we would infer the false positive edge $2 \rightarrow 3$.
-Moreover, the same example shows that if simply $D_k = \{(i, j) \in \widehat{P}_k \times C_k\ |\ i \pwgc j \}$, then the edge $1 \rightarrow 3$ would be missed.
+
+  That we need to proceed backwards through $P_{k - r}$ can be seen from the example of figure \ref{fig:example_fig3}, where if $D_k = \{(i, j) \in \Big(\bigcup_{r = 0}^k P_{k - r}\Big) \times C_k\ |\ i \pwgc j \}$ above we would infer the false positive edge $2 \rightarrow 3$.
+Moreover, the same example shows that if simply $D_k = \{(i, j) \in P_k \times C_k\ |\ i \pwgc j \}$, then the edge $1 \rightarrow 3$ would be missed.
 
@@
     \caption{Example graph for Algorithm \ref{alg:pwgr}}
-    \includegraphics[width=0.5\linewidth]{hand_drawn_example3.jpg}
+    \footnotesize{Black arrows indicate true parent-child
+      relations.  Red dotted arrows indicate pairwise causality (due to
+      ancestor relations), green dash-dotted arrow indicates
+      bidirectional pairwise causality (due to the confounding node
+      $1$).  Blue groupings indicate each $P_k$ in Algorithm
+      \ref{alg:pwgr}.  Repetition is for aesthetics.}
+
+    \begin{subfigure}[b]{0.45\textwidth}
+      \includegraphics[width=\linewidth]{example_algorithm.pdf}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.45\textwidth}
+      \includegraphics[width=\linewidth]{example_algorithm2.pdf}
+    \end{subfigure}
   \end{figure}
@@
 
-  First, suppose that $i \in \pa{j}$ and that there is no $i \rightarrow j$ path in $E$.
-Then since $i \gc j$ we have $i \pwgc j$ by proposition \ref{prop:pwgc_anc} so at least one of $\{D_{kr} \}_{r = 0}^k$ contains $(i, j)$ and hence $(i, j)$ will be added to $E$ upon the end of the iteration.
-
-  Now, still assuming that $i \in \pa{j}$, suppose instead that there is already an $i \rightarrow j$ path in $E$.
-Using an induction hypothesis that $E$ contains edges if and only if they are in $\gcg$, then either $(i, j) \in E$ correctly, or there is more than one $i \rightarrow j$ path in $\gcg$, which is not possible since $\gcg$ is strongly causal.
-
-  Finally, suppose that $i \not \in \pa{j}$.
-If $i \not \pwgc j$ then we will correctly leave $(i, j)$ absent from $E$, so suppose that $i \pwgc j$.
-Then we must have either $i \in \anc{j}$ or $\anc{i} \cap \anc{j} \ne \emptyset$; suppose first the former.
-Then, there is some $u$ s.t.
-$i \pwgc u \pwgc j$, so by the construction of \hl{it seems} that both $(i, u)$ and $(u, j)$ should have already been added to $E$, so we have an $i \rightarrow j$ path in $E$ \hl{is this sound?}  Suppose then the latter case above, $u \in \anc{i} \cap \anc{j}$ \hl{try to construct a counterexample involving this case}.
+  First, suppose that $i \in \pa{j}$ and that there is no
+  $i \rightarrow j$ path in $E$.  Then since $i \gc j$ we have
+  $i \pwgc j$ by proposition \ref{prop:pwgc_anc} so at least one of
+  $\{D_{kr} \}_{r = 0}^k$ contains $(i, j)$ and hence $(i, j)$ will be
+  added to $E$ upon the end of the iteration.
+
+  Now, still assuming that $i \in \pa{j}$, suppose instead that there
+  is already an $i \rightarrow j$ path in $E$.  Using an induction
+  hypothesis that $E$ contains edges if and only if they are in
+  $\gcg$, then either $(i, j) \in E$ correctly, or there is more than
+  one $i \rightarrow j$ path in $\gcg$, which is not possible since
+  $\gcg$ is strongly causal.
+
+  Finally, suppose that $i \not \in \pa{j}$.  If $i \not \pwgc j$ then
+  we will correctly leave $(i, j)$ absent from $E$, so suppose that
+  $i \pwgc j$.  Then we must have either $i \in \anc{j}$ or
+  $\anc{i} \cap \anc{j} \ne \emptyset$; suppose first the former.
+  Then, there is some $u$ s.t. $i \pwgc u \pwgc j$, so by the
+  construction of \hl{it seems} that both $(i, u)$ and $(u, j)$ should
+  have already been added to $E$, so we have an $i \rightarrow j$ path
+  in $E$ \hl{is this sound?}  Suppose then the latter case above,
+  $u \in \anc{i} \cap \anc{j}$ \hl{try to construct a counterexample
+    involving this case}.
 \end{proof}
@@
 \begin{proof}
+  Denote the edges of $\gcg$ by $\mathcal{E}$, we endevour to show
+  that upon the termination of Algorithm \ref{alg:pwgr} we have
+  $E = \mathcal{E}$.  On the first iteration of the algorithm
+  ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
+
+  We will first dispense with some trivialities: if $n = 0$ or $n = 1$
+  the algorithm is correct, returning on the first iteration with
+  $E = \emptyset$.  Suppose then that $n > 1$.  The set
+  $P_0$ is non-empty since $\gcg$ is a DAG and therefore
+  there are nodes in $\gcg$ without parents, thence if
+  $P_0 = \emptyset$ proposition
+  \ref{prop:ancestor_properties} would be contradicted.  Now, if
+  $P_0 = [n]$ then the graph necessarily contains no edges
+  and we will correctly return $E = \emptyset$.
+
+  Moving on from any trivial cases, notice that
+  $\mathcal{E} \subseteq W_0$.  This follows since by Proposition
+  \ref{prop:pwgc_anc} we have $j \in \pa{i} \Rightarrow j \pwgc i$ and by
+  Corollary \ref{cor:bidirectional_edge} there are no bidirectional
+  edges in $\gcg$.
+
 % 1. Dispense with trivialities
@@
 % Given non-empty P and C...  Can we show all P -> C edges get collected in D?
-  Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$ and let us define
-
-  \begin{equation*}
-    A_k = \{i \in [n]\ |\ \dist{i}{\widehat{P}_0} = k\},
-  \end{equation*}
-
-  where
-
-  \begin{equation*}
-    \dist{i}{P} = \text{max}\{r \in [n]\ |\ \exists (j \rightarrow i) \text{ path of length } r \text{ in } \gcg \text{ for some } j \in P\}.
-  \end{equation*}
-
-  We will proceed by induction and show that when the counter has reached the value $k \ge 1$, it is guaranteed that all nodes in $\bigcup_{\ell \le k}A_\ell$ (nodes having paths of length up to $k$ between themselves and $\widehat{P}_0$) must have all of their incident edges included in $E$, and that every edge in $E$ is an edge of $\gcg$.
-If this is the case, then it can be immediately seen that the algorithm terminates with the correct result, $\gcg = ([n], E)$, once $k = n$; since there can only be paths up to length $n - 1$ in $\gcg$, and there cannot be any nodes or edges that don't involve paths back to $\widehat{P}_0$.
+  % Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$ and let us define
+
+  % \begin{equation*}
+  %   A_k = \{i \in [n]\ |\ \dist{i}{P_0} = k\},
+  % \end{equation*}
+
+  % where
+
+  % \begin{equation*}
+  %   \dist{i}{P} = \text{max}\{r \in [n]\ |\ \exists (j \rightarrow i) \text{ path of length } r \text{ in } \gcg \text{ for some } j \in P\}.
+  % \end{equation*}
+
+  % We will proceed by induction and show that when the counter has reached the value $k \ge 1$, it is guaranteed that all nodes in $\bigcup_{\ell \le k}A_\ell$ (nodes having paths of length up to $k$ between themselves and $P_0$) must have all of their incident edges included in $E$, and that every edge in $E$ is an edge of $\gcg$.
+If this is the case, then it can be immediately seen that the algorithm terminates with the correct result, $\gcg = ([n], E)$, once $k = n$; since there can only be paths up to length $n - 1$ in $\gcg$, and there cannot be any nodes or edges that don't involve paths back to $P_0$.
 That is, $\bigcup_{\ell \le n}A_\ell = [n]$.
 
-  % firstly dispensing with some trivialities, and then establishing that every edge in $\gcg$ makes an appearance in some $D_{kr}$ as well as that everything in any $D_{kr}$ is indeed an edge in $\gcg$.
-
-  On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
-To quickly dispense with some trivialities: the set $\widehat{P}_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $\widehat{P}_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
-Now, if $\widehat{P}_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
-
-  Continuing with the non-trivial case: if $\widehat{P}_0 \ne [n]$ then $C_0$ must be non-empty, otherwise there would be driving nodes not included in $\widehat{P}_0$.
+  % % firstly dispensing with some trivialities, and then establishing that every edge in $\gcg$ makes an appearance in some $D_{kr}$ as well as that everything in any $D_{kr}$ is indeed an edge in $\gcg$.
+
+  % On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
+To quickly dispense with some trivialities: the set $P_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $P_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
+Now, if $P_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
+
+  % Continuing with the non-trivial case: if $P_0 \ne [n]$ then $C_0$ must be non-empty, otherwise there would be driving nodes not included in $P_0$.
 In this case it is also clear that $A_1 \ne \emptyset$, so let us consider some $j \in A_1$.
-We must have $j \in C_0$, otherwise there would be a path of length $2$ from $\widehat{P}_0$ to $j$.
-Moreover, since $\exists i \in \widehat{P}_0$ s.t.
+We must have $j \in C_0$, otherwise there would be a path of length $2$ from $P_0$ to $j$.
+Moreover, since $\exists i \in P_0$ s.t.
 $i \gc j$ and therefore $i \pwgc j$ (by proposition \ref{prop:pwgc_anc}) we have $(i, j) \in D_{00}$ and therefore $(i, j) \in E$.
@@
 
-  Now consdider some $(i, j) \in D_{00}$ (the same reasoning as for $C_0$ is sufficient to see that $D_{00}$ is non-empty).
+  % Now consdider some $(i, j) \in D_{00}$ (the same reasoning as for $C_0$ is sufficient to see that $D_{00}$ is non-empty).
 Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).
-Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in \widehat{P}_0 \implies u \in S_0 \implies j \not\in C_0$, a contradiction.
+Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in P_0 \Rightarrow u \in S_0 \Rightarrow j \not\in C_0$, a contradiction.
 This implies every tuple in $D_{00}$ is a bona-fide edge of $\gcg$, as well as that $(i, j) \in A_1$.
@@
 
-  Suppose now the algorithm has reached step $k$ and that $E = \bigcup_{\ell \le k - 1}A_\ell$.
+  % Suppose now the algorithm has reached step $k$ and that $E = \bigcup_{\ell \le k - 1}A_\ell$.
 
 % The reasoning here is (I think) correct -- but isn't directly establishing the induction hypothesis.
-%   On the other hand, if $\widehat{P}_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $\widehat{P}_0$.
+%   On the other hand, if $P_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $P_0$.
 Consider now an edge $(i, j) \in D_{00}$ ($D_{00}$ is non-empty for the same reason as $C_0$).
 Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).
-Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in \widehat{P}_0 \implies u \in S_0 \implies j \not\in C_0$, a contradiction.
+Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in P_0 \Rightarrow u \in S_0 \Rightarrow j \not\in C_0$, a contradiction.
 So every tuple in $D_{00}$ is a bona-fide edge of $\gcg$.
@@
 $(i, j) \in \gcg$ but $(i, j) \not\in D_{00}$.
-Since $(i, j) \in \gcg$ we have $i \pwgc j$ (corollary \ref{cor:gc_implies_pwgc}), but then if $i \not\in \widehat{P}_0$ then $j \notin C_0$ and if $i \in \widehat{P}_0, j \in C_0$ then necessarily $(i, j) \in D_{00}$ since we cannot at this stage have any paths in $E$.
-
-%   and show that on each step, every edge incident upon a node in $C_k$ is contained in $D_k$, that every edge in $D_k$ is an edge in $\gcg$, and finally that $\bigcup_{k = 1}^n C_k = [n] \setminus \widehat{P}_0$ and that this implies $\big([n], \bigcup_{k = 1}^n D_k\big) = \gcg$.
+Since $(i, j) \in \gcg$ we have $i \pwgc j$ (corollary \ref{cor:gc_implies_pwgc}), but then if $i \not\in P_0$ then $j \notin C_0$ and if $i \in P_0, j \in C_0$ then necessarily $(i, j) \in D_{00}$ since we cannot at this stage have any paths in $E$.
+
+%   and show that on each step, every edge incident upon a node in $C_k$ is contained in $D_k$, that every edge in $D_k$ is an edge in $\gcg$, and finally that $\bigcup_{k = 1}^n C_k = [n] \setminus P_0$ and that this implies $\big([n], \bigcup_{k = 1}^n D_k\big) = \gcg$.
 
 %  On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E_0 = \emptyset$.
-The set $\widehat{P}_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $\widehat{P}_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
-Now, if $\widehat{P}_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
-On the other hand, if $\widehat{P}_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $\widehat{P}_0$.
+The set $P_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $P_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
+Now, if $P_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
+On the other hand, if $P_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $P_0$.
 Consider now an edge $(i, j) \in D_0$ (again $D_0$ is non-empty for the same reason as $C_0$).
@@
 
-% We also see that since nodes in $\widehat{P}_0$ have no parents,
-
-% we must have $\anc{i}\cap\anc{j} = \emptyset$, and therefore by proposition \ref{prop:ancestor_properties} $i \in \anc{j}$;  moreover, the construction of $\widehat{P}_1$ implies that in fact $i \in \pa{j}$, so $(i, j)$ is an edge of $\gcg$.
+% We also see that since nodes in $P_0$ have no parents,
+
+% we must have $\anc{i}\cap\anc{j} = \emptyset$, and therefore by proposition \ref{prop:ancestor_properties} $i \in \anc{j}$;  moreover, the construction of $P_1$ implies that in fact $i \in \pa{j}$, so $(i, j)$ is an edge of $\gcg$.
 

### 2019-05-02 paper/paper.tex
Work on algorithm proof.

@@
 $j \not \in \anc{i}$) and that there are no common ancestors (i.e.
-$\anc{i} \cap \anc{j} = \emptyset$) then $j \not \pwgc i$, or $\linE{x_i(t)}{\H_{t - 1}^{(i, j)}} = \linE{x_i(t)}{\H_{t - 1}^{(i)}}$, which is an intuitively natural result.
+$\anc{i} \cap \anc{j} = \emptyset$) then $j \npwgc i$, or $\linE{x_i(t)}{\H_{t - 1}^{(i, j)}} = \linE{x_i(t)}{\H_{t - 1}^{(i)}}$, which is an intuitively natural result.
 
   The first part is used in part to prove the second (and stronger) statement, which allows common ancestors, but only if they are ``mediated'' by $i$.
-For example, this covers the case of a graph $k \rightarrow i \rightarrow j$, where $\{k\} = \anc{i} \cap \anc{j}$, but clearly $j \not \pwgc i$.
+For example, this covers the case of a graph $k \rightarrow i \rightarrow j$, where $\{k\} = \anc{i} \cap \anc{j}$, but clearly $j \npwgc i$.
 
@@
 Examples of strongly causal graphs include directed trees (or
-forests), and DAGs where each node has at most one parent.
-Fortunately, the strongly causal property is inherited by subgraphs:
+forests), DAGs where each node has at most one parent, and each of the
+figures of this paper.  Fortunately, the strongly causal property is
+inherited by subgraphs:
 
@@
   In this case, node $1$ is a confounder for nodes $2$ and $3$, but
-  $x_3(t) = v_3(t) - v_2(t) + x_2(t)$ and $2 \not \pwgc 3$ (even
+  $x_3(t) = v_3(t) - v_2(t) + x_2(t)$ and $2 \npwgc 3$ (even
   though $x_2(t)$ and $x_3(t)$ are contemporaneously correlated)
@@
     \BlankLine
-    $W_0 \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$  \texttt{\# potential edges}\\
+    $W \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$  \texttt{\# potential edges}\\
     \While{$S_k \ne \emptyset$}{
-      $P_k \leftarrow \{i \in S_{k - 1}\ |\ \forall s \in S_{k - 1}\ (s, i) \not\in W_k\}$  \texttt{\# parents of current layer}\\
+      $P_k \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W\}$  \texttt{\# parents of current layer}\\
       % $P_k \leftarrow P_{k - 1} \cup P_k$ \texttt{\# Expand parent set}\\
@@
       \;
-      $C_k \leftarrow \{j \in S_{k + 1}\ |\ \forall s \in S_{k + 1}\ (s, j) \not\in W_k\}$   \texttt{\# potential children of }$P_k$\\
+      $C_k \leftarrow \{j \in S_{k + 1}\ |\ \forall s \in S_{k + 1}\ (s, j) \not\in W\}$   \texttt{\# potential children of }$P_k$\\
 
@@
       {
-        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times C_k\ |\ (i, j) \in W_k,\ \text{no } i \rightarrow j \text{ path in } E \}$\\%  \texttt{\# edges incident on }$C_k$\\
+        \label{alg:inner_loop}
+        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times C_k\ |\ (i, j) \in W,\ \text{no } i \rightarrow j \text{ path in } E \}$\\%  \texttt{\# edges incident on }$C_k$\\
         $E \leftarrow E \cup D_{kr}$ \texttt{\# add edges to }$E$\\
@@
       }
-      $W_{k + 1} \leftarrow W_k \setminus \big(\bigcup_{r = 0}^k D_{kr}\big)$  \texttt{\# remove edges from consideration}\\
+      % $W_{k + 1} \leftarrow W_k \setminus \big(\bigcup_{r = 0}^k D_{kr}\big)$  \texttt{\# remove edges from consideration}\\
 
@@
 \begin{example}
-
-  That we need to proceed backwards through $P_{k - r}$ can be seen from the example of figure \ref{fig:example_fig3}, where if $D_k = \{(i, j) \in \Big(\bigcup_{r = 0}^k P_{k - r}\Big) \times C_k\ |\ i \pwgc j \}$ above we would infer the false positive edge $2 \rightarrow 3$.
-Moreover, the same example shows that if simply $D_k = \{(i, j) \in P_k \times C_k\ |\ i \pwgc j \}$, then the edge $1 \rightarrow 3$ would be missed.
-
-\hl{TODO: Inkscape figure}
+  The set $W$ collects all of the edges that may be edges of $\gcg$.
+  In reference to figure \ref{fig:example_fig3}, each of the solid
+  black edges, as well as the dotted red edges will be included, but
+  \textit{not} the bidirectional green dash-dotted edge.  That we exclude
+  bidirectional edges is important since it will provide guarantees
+  that $P_k$ is non-empty.  The groupings $P_0, \ldots, P_3$
+  are also indicated in figure \ref{fig:example_fig3}.
+
+  The algorithm proceeds first with the parentless nodes $1, 2$ on the
+  initial iteration where the edge $(1, 3)$ is added to $E$.  On the
+  next iteration, the edges $(3, 4), (2, 4)$ are added, and the false
+  edge $(1, 4)$ is excluded due to the path
+  $1 \rightarrow 3 \rightarrow 4$ in E.  Finally, edges
+  $(3, 5), (4, 6)$ are added, and the false
+  $(1, 5), (1, 6), (3, 6), (2, 6)$ edges are similarly excluded due to
+  the ordering of the inner loop.
 
@@
     \centering
-    \label{fig:example_fig3}
     \caption{Example graph for Algorithm \ref{alg:pwgr}}
@@
       \ref{alg:pwgr}.  Repetition is for aesthetics.}
+    \label{fig:example_fig3}
 
@@
   \end{figure}
+
+  That we need to proceed backwards through $P_{k - r}$ as in the
+  inner loop of line \ref{alg:inner_loop} can also be seen from this
+  example, where if instead
+  $D_k = \{(i, j) \in \Big(\bigcup_{r = 0}^k P_{k - r}\Big) \times
+  C_k\ |\ i \pwgc j \}$ we would infer the false positive edge
+  $1 \rightarrow 4$.  Moreover, the same example shows that if simply
+  $D_k = \{(i, j) \in P_k \times C_k\ |\ i \pwgc j \}$, then the edge
+  $1 \rightarrow 3$ would be missed.
 \end{example}
@@
 \begin{lemma}
-  \hl{Some lemma on the inner loop.}
+  \label{lem:inner_loop_lemma}
+  At any given iteration, $k$, of Algorithm \ref{alg:pwgr}, if we have
+  $e \in E$ if and only if $e \in \mathcal{E}$, then the inner loop (line \ref{alg:inner_loop})
+  will add $(i, j) \in \big(\bigcup_{r = 0}^kP_{k - r}\big) \times C_k$ to $E$
+  if $(i, j) \in \mathcal{E}$.  i.e. any real edge will be included.
 \end{lemma}
 \begin{proof}
-  Let $j \in C_k$ and $i \in \bigcup_{r = 0}^k P_{k - r}$.
-
-  First, suppose that $i \in \pa{j}$ and that there is no
-  $i \rightarrow j$ path in $E$.  Then since $i \gc j$ we have
-  $i \pwgc j$ by proposition \ref{prop:pwgc_anc} so at least one of
+  Let $j \in C_k$, $i \in \bigcup_{r = 0}^k P_{k - r}$, and assume
+  $i \in \pa{j}$.
+
+  First, suppose that there is no $i \rightarrow j$ path in $E$.  Then
+  since $i \gc j$ we have $i \pwgc j$ and $j \npwgc i$ (i.e.
+  $(i, j) \in W$) by Proposition \ref{prop:pwgc_anc} and Corollary
+  \ref{cor:bidirectional_edge} so at least one of
   $\{D_{kr} \}_{r = 0}^k$ contains $(i, j)$ and hence $(i, j)$ will be
@@
 
-  Now, still assuming that $i \in \pa{j}$, suppose instead that there
-  is already an $i \rightarrow j$ path in $E$.  Using an induction
-  hypothesis that $E$ contains edges if and only if they are in
-  $\gcg$, then either $(i, j) \in E$ correctly, or there is more than
-  one $i \rightarrow j$ path in $\gcg$, which is not possible since
-  $\gcg$ is strongly causal.
-
-  Finally, suppose that $i \not \in \pa{j}$.  If $i \not \pwgc j$ then
-  we will correctly leave $(i, j)$ absent from $E$, so suppose that
-  $i \pwgc j$.  Then we must have either $i \in \anc{j}$ or
-  $\anc{i} \cap \anc{j} \ne \emptyset$; suppose first the former.
-  Then, there is some $u$ s.t. $i \pwgc u \pwgc j$, so by the
-  construction of \hl{it seems} that both $(i, u)$ and $(u, j)$ should
-  have already been added to $E$, so we have an $i \rightarrow j$ path
-  in $E$ \hl{is this sound?}  Suppose then the latter case above,
-  $u \in \anc{i} \cap \anc{j}$ \hl{try to construct a counterexample
-    involving this case}.
-\end{proof}
+  Now, suppose instead that there is already an $i \rightarrow j$ path
+  in $E$.  Given our assumption that $E$ contains edges if and only if
+  they are edges of $\gcg$, then either we already have $(i, j) \in E$
+  correctly, or there is more than one $i \rightarrow j$ path in
+  $\gcg$, which is not possible since $\gcg$ is strongly causal.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:W_subset_E}
+  $W \subseteq \mathcal{E}$
+\end{lemma}
+\begin{proof}
+  By proposition \ref{prop:pwgc_anc} we have
+  $j \in \pa{i} \Rightarrow j \pwgc i$ and by Corollary \ref{cor:bidirectional_edge}
+  there are no bidirectional edges in $\gcg$.
+\end{proof}
+
+\hl{Must show that Algorithm terminates in $n$ steps.}
+\begin{lemma}
+  For any $i \in P_k$ we have $i \in C_{k - 1}$.  \hl{[Can't yet prove
+    this part] Moreover, for each $i \in \bigcup_{k = 0}^n P_k$, if
+    $(s, i) \in W$ then $(s, i) \in E$ upon the termination of
+    Algorithm} \ref{alg:pwgr}.
+\end{lemma}
+\begin{proof}
+  We prove the second part by induction, the first part being
+  established along the way.  By construction $P_0$ has no parent
+  nodes in $\gcg$ so the base case is immediate.  Fix now some
+  $k \ge 1$, $i \in P_k$, and let $s$ be such that $(s, i) \in W$.
+  For every $s$ such that $(s, i) \in W$ we have $s \not\in S_k$ since
+  otherwise $i \not\in P_k$.  Therefore
+  $s \in \bigcup_{r = 0}^{k - 1}P_{k - r}$, which implies
+  $s \not\in S_k$, and finally that $i \in C_{k - 1}$.  Then by Lemma
+  \ref{lem:W_subset_E} every edge in $\mathcal{E}$ is also present in
+  $W$, and by Lemma \ref{lem:inner_loop_lemma} \hl{[We can't conclude
+    here because of the strong conditions required on lemma}
+    \ref{lem:inner_loop_lemma}]
+\end{proof}
+
+%   Finally, suppose that $i \not \in \pa{j}$.  If $i \npwgc j$ then
+%   we will correctly leave $(i, j)$ absent from $E$, so suppose that
+%   $i \pwgc j$.  Then we must have either $i \in \anc{j}$ or a confounding 
+%   $k \in \anc{i} \cap \anc{j}$; suppose first the former.
+%   Then, there is some $u$ s.t. $i \pwgc u \pwgc j$, so by the
+%   construction of \hl{it seems} that both $(i, u)$ and $(u, j)$ should
+%   have already been added to $E$, so we have an $i \rightarrow j$ path
+%   in $E$ \hl{is this sound?}  Suppose then the latter case above,
+%   $u \in \anc{i} \cap \anc{j}$ \hl{try to construct a counterexample
+%     involving this case}.
+% \end{proof}
 
@@
   $P_0$ is non-empty since $\gcg$ is a DAG and therefore
-  there are nodes in $\gcg$ without parents, thence if
+  there are nodes in $\gcg$ without parents (if
   $P_0 = \emptyset$ proposition
-  \ref{prop:ancestor_properties} would be contradicted.  Now, if
+  \ref{prop:ancestor_properties} would be contradicted).  Now, if
   $P_0 = [n]$ then the graph necessarily contains no edges
@@
   Moving on from any trivial cases, notice that
-  $\mathcal{E} \subseteq W_0$.  This follows since by Proposition
+  $\mathcal{E} \subseteq W$.  This follows since by Proposition
   \ref{prop:pwgc_anc} we have $j \in \pa{i} \Rightarrow j \pwgc i$ and by

### 2019-05-13 paper/paper.tex
Significant work on completing a converse of the main ancestor_properties result.  Introduced notion of a persistent system.

@@
 \documentclass[12pt]{article}
+
+% TODO: read this https://lemire.me/blog/rules-to-write-a-good-research-paper/
+
+% TODO:
+% -- define "DAG"
+% -- remove speculative statements
+% --
 
@@
 
-\begin{proposition}[Invariance Under Invertible and Deterministic Modifications]
-  Let $x(t)$ be a WSS process with absolutely summable covariance
-  sequence, and spectral density uniformly bounded above and below.  Let
-  $F(z), G$ represent univariate and invertible
-  linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
-  predictable processes.  Then,
-
-  \begin{equation}
-    x_j \gc x_i \iff F(x_j - f)(t) \gc G(x_i - g)(t)
-  \end{equation}
-
-\begin{proof}
-  \hl{TODO}
-\end{proof}
-\end{proposition}
+% \begin{proposition}[Invariance Under Invertible and Deterministic Modifications]
+%   Let $x(t)$ be a WSS process with absolutely summable covariance
+%   sequence, and spectral density uniformly bounded above and below.  Let
+%   $F(z), G$ represent univariate and invertible
+%   linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
+%   predictable processes.  Then,
+
+%   \begin{equation}
+%     x_j \gc x_i \iff F(x_j - f)(t) \gc G(x_i - g)(t)
+%   \end{equation}
+
+% \begin{proof}
+%   \hl{TODO}
+% \end{proof}
+% \end{proposition}
 
@@
 
+\begin{definition}[Graph Theory Review]
+  A \textit{graph} $\gcg = (\mathcal{V}, \mathcal{E})$ is simply a
+  tuple of sets respectively called \textit{nodes} and \textit{edges}.
+  Throughout this paper, we have in all cases
+  $V = [n] \defeq \{1, 2, \ldots, n\}$.  We will also focus solely on
+  \textit{directed} graphs, where the edges
+  $\mathcal{E} \subseteq V \times V$ are \textit{ordered} pairs.
+
+  A (directed) \textit{path} (of length $r$) between nodes $i, j$,
+  denoted $\gcgpath{i}{j}$, is a sequence
+  $a_0, a_1, \ldots, a_{r - 1}, a_r$ with $a_0 = i$ and $a_r = j$ such
+  that $\forall\ 0 \le k \le r\ (a_k, a_{k + 1}) \in \mathcal{E}$.
+
+  A \textit{cycle} is a path of length $2$ or more between a node and
+  itself.  An edge between a node and itself $(i, i)$ (which is not a
+  cycle) is referred to as a \textit{loop}.
+
+  A graph $\gcg$ is a \textit{directed acyclic graph} (DAG) if it is a
+  directed graph and does not contain any cycles.
+\end{definition}
+
 \begin{definition}[Causality graph, Parents]
@@
   The set of level $\ell$ \textit{grandparents} of node $i$, denoted
-$\gpn{\ell}{i}$, is the set such that $j \in \gpn{\ell}{i}$ if and only if
-there is a \textit{directed and cycle-free path} of length $\ell$ in
-$\gcg$ from $j$ to $i$ \textit{or} if $j = i$ and there is a cycle of
-length $\ell$ starting at $i$ and ending back at $i$, that is, $i$ can be
-it's own grandparent.  Clearly, $\pa{i} = \gpn{1}{i}$.
-
-Finally, the set of \textit{level $\ell$ ancestors} of
-$i$: $\ancn{\ell}{i} = \bigcup_{\lambda \le \ell}\gpn{\lambda}{i}$ is the set such that
-$j \in \ancn{\ell}{i}$ if and only if there is a directed and cycle-free
-path of length $\ell$ \textit{or less} in $\gcg$ from $j$ to $i$
-\textit{or} if $j = i$ and there is a directed cycle of length $\ell$ or
-less from $i$ back to $i$ in $\gcg$.  The set of \textit{all
-  ancestors} of $i$ (i.e. $\ancn{n}{i}$) is denoted simply $\anc{i}$.
+  $\gpn{\ell}{i}$, is the set such that $j \in \gpn{\ell}{i}$ if and
+  only if there is a \textit{directed path} of length $\ell$ in $\gcg$
+  from $j$ to $i$.  Note that a node can be it's own grandparent
+  unless $\gcg$ is a DAG.  Clearly, $\pa{i} = \gpn{1}{i}$.
+
+  Finally, the set of \textit{level $\ell$ ancestors} of $i$:
+  $\ancn{\ell}{i} = \bigcup_{\lambda \le \ell}\gpn{\lambda}{i}$ is the
+  set such that $j \in \ancn{\ell}{i}$ if and only if there is a
+  directed path of length $\ell$ \textit{or less} in $\gcg$ from $j$
+  to $i$.  The set of \textit{all ancestors} of $i$
+  (i.e. $\ancn{n}{i}$) is denoted simply $\anc{i}$.
+
+  Note that since loops are not considered paths, it is not possible
+  for a node to be it's own parent.
 \end{definition}
-
-Even though the definitions of Granger-causality naturally suggests
-that each node in the graph should have a self-loop (unless the node
-is simply independent noise) we will not include nodes in their own
-parent sets: $i \not\in \pa{i} \subseteq \anc{i}$.  However, we do not
-immediately assume that $\gcg$ is a DAG, and so \textit{we will allow for $i$
-to be it's own ancestor}.
 
@@
   \begin{equation}
-    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t).
-  \end{equation}
+    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t),
+  \end{equation}
+
+  where $\wtB_{ii}(z) = (1 - \B_{ii}(z))^{-1}$ and
+  $\wtB_{ik}(z) = \wtB_{ii}(z) B_{ik}(z)$.
 
@@
 
-  where $\A(z) = \sum_{\tau = 0}^\infty A(\tau)$ is the filter from
+  where $\A(z) = \sum_{\tau = 0}^\infty A(\tau)z^{-\tau}$ is the filter from
   the Wold decomposition representation of $x(t)$, equation
@@
   \begin{align*}
-    x_i(t) &= (I - \B(z))^{-1}\\
+    x_i(t) &= (I - \B(z))^{-1}v(t)\\
     &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
-    &= \A_{ii}(z) v_i(t) + \sum_{\substack{j \in \anc{i} \\ j \ne i}} \A_{ij}(z) v_j(t)\\
+    &= \A_{ii}(z) v_i(t) + \sum_{\substack{j \in \anc{i} \\ j \ne i}} \A_{ij}(z) v_j(t)
   \end{align*}
@@
 
-  where $\H_2 \vee \H_2 = \cl \{\alpha + \beta\ |\ \alpha \in \H_1, \beta \in \H_2 \}$ is the closed sum of subspaces.
+  where $\H_1 \vee \H_2 = \cl \{\alpha + \beta\ |\ \alpha \in \H_1, \beta \in \H_2 \}$ is the closed sum of subspaces.
 \end{lemma}
@@
   not containing $i$.
+
+  A simple example is furnished by the ``fork'' graph
+  $i \leftarrow k \rightarrow j$.
 \end{definition}
@@
   \label{prop:ancestor_properties}
+  \hl{Add the stronger statements on uncorrelatedness for use on confounders later.}
+
   In a Granger-causality graph $\gcg$, if $j \pwgc i$ then at least
@@
 For example, this covers the case of a graph $k \rightarrow i \rightarrow j$, where $\{k\} = \anc{i} \cap \anc{j}$, but clearly $j \npwgc i$.
-
-  An ancestor such as that in the second part of the theorem can be referred to as a \textit{confounder}, and a simple example is given by the fork $i \leftarrow k \rightarrow j$.
 \end{remark}
@@
 \subsection{Strongly Causal Graphs}
-The examples at the conclusion of the previous section seem rather
-pathological since we have simply constructed a case where different
-paths in the graph cancel exactly.  It may be possible to rid
-ourselves of these pathologies for instance by defining a notion of
-``robust'' pairwise causality where ``$j \pwgc _R i$'' if $j \pwgc i$
-almost surely for a ``small'' random perturbation of the system
-matrix.  However, we do not believe such a modification of our theory
-can have useful practical implications since the difference between
-paths that ``nearly cancel'' rather than cancel exactly would in
-practice require inordinate amounts of data to resolve.
-
-Alternatively, we could restrict the coefficients of the system
-matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead, we
-think it is more meaningful to focus on the defining feature of time
-series networks, that is, the topology of $\gcg$.
+\label{sec:strongly_causal_graphs}
+% The examples at the conclusion of the previous section seem rather
+% pathological since we have simply constructed a case where different
+% paths in the graph cancel exactly.  It may be possible to rid
+% ourselves of these pathologies for instance by defining a notion of
+% ``robust'' pairwise causality where ``$j \pwgc _R i$'' if $j \pwgc i$
+% almost surely for a ``small'' random perturbation of the system
+% matrix.  However, we do not believe such a modification of our theory
+% can have useful practical implications since the difference between
+% paths that ``nearly cancel'' rather than cancel exactly would in
+% practice require inordinate amounts of data to resolve.
+
+In this section and the next we will seek to understand when converse
+statements of Proposition \ref{prop:ancestor_properties} \textit{do}
+hold.  One possibility is to restrict to coefficients of the system
+matrix matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead,
+we think it is more meaningful to focus on the defining feature of
+time series networks, that is, the topology of $\gcg$.
 
 \begin{definition}[Strongly Causal]
-  We will say that a Granger-causality graph $\gcg$ is \textit{strongly causal}
-if there is at most 1 directed path between any two nodes.
-
-\hl{TODO: Should DAG be included in this definition?  I think it is
-  not necessary for some theorems.}
+  \label{def:strongly_causal}
+  We will say that a Granger-causality graph $\gcg$ is
+  \textit{strongly causal} if there is at most 1 directed path between
+  any two nodes.
 \end{definition}
@@
 Examples of strongly causal graphs include directed trees (or
-forests), DAGs where each node has at most one parent, and each of the
-figures of this paper.  Fortunately, the strongly causal property is
-inherited by subgraphs:
+forests), DAGs where each node has at most one parent, and figure
+\ref{fig:example_fig3} of this paper.  The persistence condition is
+used to ensure that $\linE{x_i(t)}{\H_{t - 1}^{(i)}} \ne 0$ i.e. that
+the each process has at least some amount of ``memory'' -- this can
+occur either due to cycles, or non-zero $\B_{ii}(z)$ terms in equation
+\eqref{eqn:ar_representation}.  It is evident that the strong causal property
+is inherited by subgraphs, but persistence need not be, unless $\gcg$ is a DAG.
+
+\hl{Not sure these are needed, but were important in my own thinking.}
 
@@
   be fewer paths after removing nodes from $\gcg$.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:still_persistent}
+  If $\gcg$ is a persistent DAG, any subgraph formed by eliminating nodes
+  as well as all of the in and out edges thereof is still persistent.
+\end{lemma}
+\begin{proof}
+  \hl{TODO -- is this needed?}
 \end{proof}
@@
 
+\begin{proposition}
+  \label{prop:independent_confounders}
+  If $\gcg$ is strongly causal we have distinct
+  $k_1, k_2 \in \anc{i} \cap \anc{j}$ are confounders of $i, j$, then
+  $k_1 \not\in \anc{k_2}$, $k_2 \not\in \anc{k_1}$ and
+  $\not\exists \ell \in \anc{k_1} \cap \anc{k_2}$ confounding
+  $k_1, k_2$.
+\end{proposition}
+\begin{proof}
+  \hl{TODO}
+\end{proof}
+
+\begin{corollary}
+  If $\gcg$ is strongly causal, then distinct confounders are
+  uncorrelated.
+\end{corollary}
+\begin{proof}
+  This follows directly by combining proposition
+  \ref{prop:independent_confounders} with proposition \ref{prop:ancestor_properties}.  \hl{I need to update prop} \ref{prop:ancestor_properties} \hl{to note the strong non-correlation property.}
+\end{proof}
+
+% \begin{lemma}
+%   If $\gcg$ is a strongly causal DAG and $j \in \anc{i}$ then 
+% \end{lemma}
+
 In light of proposition \ref{prop:sc_graph_common_anc}, the following
@@
 
+  % Should this be written as a lemma?
   \begin{align*}
@@
 
-  but this is impossible since with probability one $x_i(t) \not \in \H_{t - 1}^{(i)}$.
+  but this is impossible since $x_i(t) \not \in \H_{t - 1}^{(i)}$.
 \end{proof}
@@
 
+\subsection{Persistent Systems}
+\label{sec:persistent_systems}
+In section \ref{sec:strongly_causal_graphs} we obtained a converse to
+the part $(a)$ of proposition \ref{prop:ancestor_properties} via the
+notion of a strongly causal graph topology.  In this section, we
+complete a converse by adding the additional requirement we refer to
+as ``persistence''.  This condition requires that each node maintains
+some ``memory'' of the past.
+
+\begin{definition}[Lag Function]
+  Given a caual filter $\B(z) = \sum_{\tau = 0}^\infty b(\tau)z^{-\tau}$
+  define
+
+  \begin{align}
+    \tau_0(\B) &= \text{min}\{\tau \in \Z_+\ |\ b(\tau) \ne 0\},\\
+    \tau_{\infty}(\B) &= \text{sup}\{\tau \in \Z_+\ |\ b(\tau) \ne 0\}.\\
+  \end{align}
+
+  i.e. the ``first'' and ``last'' coefficients of the filter $\B(z)$,
+  where $\tau_\infty(\B) = \infty$ if the filter has an infinite
+  impulse response.
+\end{definition}
+
+\begin{definition}[Persistent]
+  We will say that a WSS process $x(t)$ with Granger-causality graph
+  $\gcg$ is \textit{persistent} if for every $i \in [n]$, every
+  $k \in \anc{i}$ we have $\tau_\infty(\A_{ik}) = \infty$.
+\end{definition}
+
+\begin{remark}
+  There are a multitude of conditions which would guarantee $x(t)$ is
+  persistent, for innstance, any ARMA system with a non-zero
+  autoregressive component is likely to be persistent due to recursion
+  in the coefficients naturally lead to an $\mathsf{MA}(\infty)$
+  representation.
+
+  In particular, any system with a finite impulse response
+  (i.e. $\mathsf{VAR}(q)$ models) is likely to be persistent.  These
+  models are common in applications of Granger-causality.
+
+  Moreover, persistence is not the weakest condition necessary for the
+  results of this section, but we opt to avoid unnecessarily
+  complicated definitions.
+\end{remark}
+
+\begin{lemma}
+  \label{lem:time_lag_cancellation}
+  Suppose $v(t)$ is a scalar sequence with unit variance and zero
+  autocorrelation and let $\A(z), \B(z)$ be nonzero and strictly
+  causal (i.e. $\tau_0(\A) \ge 1$) linear filters.  Then,
+
+  \begin{equation}
+    \inner{F(z)\A(z)v(t)}{\B(z)v(t)} = 0\ \forall \text{ strictly causal filters } F(z)
+  \end{equation}
+
+  if and only if $\tau_0(\A) \ge \tau_\infty(\B)$.
+\end{lemma}
+\begin{proof}
+  We have
+
+  \begin{align}
+    \inner{\A(z)v(t)}{\B(z)v(t)} &= \sum_{\tau = 1}^\infty \sum_{s = 1}^\infty a(\tau)b(s)\E[v(t - s)v(t - \tau)]\\
+    &= \sum_{\tau = \text{max}(\tau_0(\A), \tau_0(\B))}^{\text{min}(\tau_\infty(\A), \tau_\infty(\B))} a(\tau) b(\tau)\\
+  \end{align}
+
+  due to the uncorrelatedness assumptions on $v(t)$.  This expression
+  is $0$ if and only if $\tau_0(\A) \ge 1 + \tau_\infty(\B)$ or if
+  $\tau_0(\B) \ge 1 + \tau_\infty(\A)$ or if the coefficients are
+  orthogonal along the common support.
+
+  Specializing this fact to $\inner{F(z)\A(z)v(t)}{\B(z)v(t)} = 0$ we
+  see that the coefficients cannot be orthogonal for every choice of
+  $F$, and that $\text{sup}_F \tau_\infty(F\A) = \infty$, leaving only
+  the possibility that
+
+  \begin{align*}
+    \tau_0(F\A) \ge 1 + \tau_\infty(\B) \forall F &\overset{(a)}{\iff} \tau_0(\A) \ge 1 + \tau_\infty(\B) - \underset{F}{\text{min }} \tau_0(F)\\
+    &\overset{(b)}{\iff} \tau_0(\A) \ge \tau_\infty(\B),
+  \end{align*}
+
+  where $(a)$ follows since $\tau_0(F\A) = \tau_0(F) + \tau_0(\A)$,
+  and $(b)$ since $\text{min}_F\ \tau_0(F) = 1$.
+\end{proof}
+
+\begin{corollary}
+  \label{cor:time_lag_cancellation}
+  For $k \in \anc{i} \cap \anc{j}$ we have
+
+  \begin{align*}
+    \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} &= 0\ \forall \text{ strictly causal } F(z)\\
+    \iff \inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} &= 0\ \forall \text{ strictly causal } F(z)\\
+    \iff \tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})
+  \end{align*}
+\end{corollary}
+\begin{proof}
+  The final equivalence follows immediately from the Lemma.  For the first equivalence we have
+
+  \begin{align*}
+    \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} &= 0\ \forall \text{ strictly causal } F(z)\\
+    \iff \inner{F(z)A_{jk}(z)v_k(t)}{x_i(t - \tau)} &= 0\ \forall \tau \ge 1, \text{ strictly causal } F(z),
+  \end{align*}
+
+  which can be expanded by equation \eqref{eqn:ancestor_expansion} to
+  obtain (after cancellinng all ancestors of $i$ other than $k$)
+
+  \begin{equation*}
+    \inner{F(z)A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t - \tau)} = 0\ \forall \tau \ge 1, \text{ strictly causal } F(z),
+  \end{equation*}
+
+  which by the Lemma is equivalent to $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$ as stated.
+\end{proof}
+
+\begin{proposition}
+  Suppose $\gcg$ is a strongly causal DAG and that $x(t)$ is persistent DAG, then if $k$
+  confounds $i, j$ we have $i \pwgc j$ and $j \pwgc i$.
+\end{proposition}
+\begin{proof}
+  We will show that $j \pwgc i$, the other being symmetric.  First
+  note also that by proposition \ref{prop:sc_graph_common_anc} we
+  cannot have $i \in \anc{j}$ or $j \in \anc{i}$ and therefore every
+  $k \in \anc{i}\cap\anc{j}$ will be a confounder.
+
+  It is sufficient to show that $\exists \psi \in \H_{t - 1}^{(j)}$
+  such that
+
+  \begin{equation*}
+    \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} \ne 0.
+  \end{equation*}
+
+  To this end, let $F(z)$ be an arbitrary but strictly causal linear,
+  and equation \eqref{eqn:ancestor_expansion} to $x_i(t)$ and
+  $\psi \defeq F(z)x_j(t)$:
+
+  \begin{align*}
+    &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}\\
+    &\overset{(a)}{=} \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
+    &\overset{(b)}{=} \inner{\sum_{k \in \anc{j}}\big(F(z)\A_{jk}(z)v_k(t) - \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}\big)}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
+    &\overset{(c)}{=} \sum_{k \in \anc{i}\cap\anc{j}}\Big(\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} - \inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{ii}v_i(t)}\\
+    &- \sum_{\ell \in \anc{i}}\inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{i\ell}(z)v_\ell(t)}\Big)
+  \end{align*}
+
+  where in $(a)$ we have removed the $\linE{x_i(t)}{\H_{t - 1}^{(i)}}$
+  term via the orthogonality principle, in $(b)$ there is no
+  $F(z)\A_{jj}(z)v_j(t)$ term since due to $j \not\in \anc{i}$ it is
+  orthogonal to $\H_t^{(i)}$.  Finally, $(c)$ follows by applying
+  orthogonality properties of $v(t)$, as well as the fact that
+  $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$ for
+  $k \not \in \anc{i}$.  Note that
+  $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$
+  and thence there is in general no cancellation in the final term
+  above for $\ell \in \anc{i}$.
+
+  Evidently, (\hl{This is clearly not immediately evident}) this is
+  $0$ for each $F$ if and only if
+  $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$ and
+  $\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} = 0$, which by
+  Collary \ref{cor:time_lag_cancellation} occurs if and only if
+  $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$, which is impossible
+  since in general $\tau_0(\A_{jk}) < \infty$ and in a persistent
+  graph $\tau_\infty(\A_{ik}) = \infty$.
+\end{proof}
+
 \subsection{Recovering $\gcg$ via Pairwise Tests}
@@
     \Initialize{$S_0 = [n]$  \texttt{\# unprocessed nodes}\\
-      $E = \emptyset$  \texttt{\# edges of }$\gcg$\\
+      $E_0 = \emptyset$  \texttt{\# edges of }$\gcg$\\
       % $P_0 = \emptyset$  \texttt{\# layer by layer driving nodes}\\
@@
         \label{alg:inner_loop}
-        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times C_k\ |\ (i, j) \in W,\ \text{no } i \rightarrow j \text{ path in } E \}$\\%  \texttt{\# edges incident on }$C_k$\\
-        $E \leftarrow E \cup D_{kr}$ \texttt{\# add edges to }$E$\\
+        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times C_k\ |\ (i, j) \in W,\ \text{no } i \rightarrow j \text{ path in } E_k \}$\\%  \texttt{\# edges incident on }$C_k$\\
+        $E_k \leftarrow E_k \cup D_{kr}$ \texttt{\# add edges to }$E_k$\\
         % $W_{k + 1} \leftarrow W_k \setminus D_{kr}$  \texttt{\# remove edges from consideration}\\
@@
     }
+    $E \leftarrow E_{k + 1}$\\
     \Return{$E$}
@@
   initial iteration where the edge $(1, 3)$ is added to $E$.  On the
-  next iteration, the edges $(3, 4), (2, 4)$ are added, and the false
-  edge $(1, 4)$ is excluded due to the path
-  $1 \rightarrow 3 \rightarrow 4$ in E.  Finally, edges
-  $(3, 5), (4, 6)$ are added, and the false
-  $(1, 5), (1, 6), (3, 6), (2, 6)$ edges are similarly excluded due to
-  the ordering of the inner loop.
+  next iteration, the edges $(3, 4), (2, 4), (3, 5)$ are added, and
+  the false edges $(1, 4), (1, 5)$ is excluded due to the paths
+  $1 \rightarrow 3 \rightarrow 4$ and $1 \rightarrow 3 \rightarrow 5$
+  in E.  Finally, edge $(4, 6)$ are added, and the false
+  $(1, 6), (3, 6), (2, 6)$ edges are similarly excluded due to the
+  ordering of the inner loop.
 
@@
       relations.  Red dotted arrows indicate pairwise causality (due to
-      ancestor relations), green dash-dotted arrow indicates
+      ancestor relations), green dash-dotted arrows indicates
       bidirectional pairwise causality (due to the confounding node
       $1$).  Blue groupings indicate each $P_k$ in Algorithm
-      \ref{alg:pwgr}.  Repetition is for aesthetics.}
+      \ref{alg:pwgr}.}
     \label{fig:example_fig3}
@@
 
+\begin{lemma}[Loop Invariant]
+  \label{lem:loop_invariant}
+  $\forall k \ge 0,\ e \in E_k \Rightarrow e \in \mathcal{E}$.
+  \hl{Not sure if this is useful / the right approach.}
+\end{lemma}
+\begin{proof}
+  We will proceed by induction.  Firstly, the base case
+  $E_0 = \emptyset$ is trivial, so fix some $k \ge 1$ and consider an
+  edge $(i, j) \in \big(\bigcup_{r = 0}^k P_{k - r} \big) \times C_k$
+  such that $(i, j) \in W$ (if the edge is not in $W$ then it is not
+  in $\mathcal{E}$ by Lemma \ref{lem:W_subset_E}).
+
+  \hl{WIP}
+\end{proof}
+
+\begin{definition}[Depth]
+  For our present purposes we will define the \textit{depth} $d(j)$ of
+  a node $j$ in $\gcg$ to be the length of the \textit{longest} path
+  from a node in $P_0$ to $j$, where $d(j) = 0$ if $j \in P_0$.  It is
+  apparent (\hl{right?}) that such a path will always exist in a
+  strongly causal graph.  For example, in Figure
+  \ref{fig:example_fig3} we have $d(3) = 1$ and $d(4) = 2$.
+\end{definition}
+
+\begin{lemma}
+  $i \in P_k \iff d(i) = k$ and $j \in S_k \iff d(j) \ge k$.
+\end{lemma}
+\begin{proof}
+  We proceed by induction.  The base case $j \in P_0 \iff d(j) = 0$ is
+  by definition, and $j \in S_0 \iff d(j) \ge 0$ is trivial since
+  $S_0 = [n]$.  So suppose that the lemma is true up to $k - 1$.
+
+  For $i \in P_k$ we have by construction that there is no $j \in S_k$
+  such that $(j, i) \in W$ and therefore no $j \in S_k$ such that
+  $j \in \pa{i}$.  Therefore, any parent of $i$ can only be in
+  $S_{k - r}$ for $r \ge 1$ and by the induction hypothesis which
+  implies that $d(i) \le k$.  But if $d(i) < k$ then $i \in P_{k - r}$
+  for some $r \ge 1$ (again by the induction hypothesis) thence
+  $d(i) = k$ since $P_k \cap P_{k - r} = \emptyset$ by construction.
+
+  Now for $s \in S_k$ we have by induction that and the fact that
+  $s \in S_k \implies s \in S_{k - 1}$ by construction we have
+  $d(s) \ge k - 1$, but again by induction (this time on $P_{k - 1}$)
+  we have $d(s) \ne k$ since $S_k = S_{k - 1} \setminus P_{k - 1}$
+  and therefore $d(s) \ge k$.
+
+  Conversely, suppose $i \in [n]$ is such that $d(i) = k$.  Then
+  $i \in S_{k - 1}$ by the hypothesis, but also $i \not\in P_{k - 1}$
+  so then $i \in S_k = S_{k - 1} \setminus P_{k - 1}$.  Now, recalling
+  the definition of $P_k$
+
+  \begin{equation*}
+    P_k = \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W \},
+  \end{equation*}
+
+  if $s \in S_k$ is such that $s \pwgc i$ then by Proposition
+  \ref{prop:ancestor_properties} $(a) s \in \anc{i}$ or
+  $(b) \exists u \in \anc{i} \cap \anc{j}$ which is a confounder.  If
+  we have $s \in \anc{i}$ then since $d(s) \ge k$ we must have
+  $d(i) > k$, which is a contradiction so $s \not\in \anc{i}$.  On the
+  other hand, if we have a confounding $u \in \anc{i} \cap \anc{s}$
+  then we must have $d(u) < k$ (since $s \in \anc{i}$ and $d(i) = k$)
+  which implies by induction that $u \in P_{k - r}$ for some $r \ge 1$,
+  and therefore that $u \not\in ???$  \hl{This is wrong.}
+
+  Now let $j \in S_k$, then $d(j) \ge k - 1$ since we
+  also have $j \in S_{k - 1}$ (i.e. $S_k$ is a decreasing sequence of
+  sets) \hl{Can't conclude unless $i \in P_k \iff d(i) = k$, in which
+    case, we have $S_k = S_{k - 1} \setminus P_{k - 1}$ and therefore
+    $d(j) \ge k$.}
+\end{proof}
+
+\begin{lemma}[Loop Invariant]
+  For every $(i, j) \in \mathcal{E}$ such that $d(j) \le k$ we have
+  $(i, j) \in E_k$.  Conversely, for every $(i, j) \in E_{k}$ we have
+  $d(j) \le k$ and $(i, j) \in \mathcal{E}$.  That is, at iteration
+  $k, E_k$ and $\mathcal{E}$ agree on the set of edges whose
+  terminating node is at most $k$ steps away from $P_0$.
+\end{lemma}
+\begin{proof}
+  We will proceed by induction.  The base case $E_0 = \emptyset$ is
+  trivial, so fix some $k \ge 1$, and suppose that the lemma holds for
+  all nodes of depth less than $k$.
+\end{proof}
+
 %   Finally, suppose that $i \not \in \pa{j}$.  If $i \npwgc j$ then

### 2019-05-14 paper/paper.tex
Completed correctness proofs for Algorithm 1.

@@
 \def\gcg{\mathcal{G}}  % Granger-causality graph
+\def\gcge{\mathcal{E}}  % Graph edges
 \def\VAR{\mathsf{VAR}}  % VAR(p) model
@@
 \begin{definition}[Graph Theory Review]
-  A \textit{graph} $\gcg = (\mathcal{V}, \mathcal{E})$ is simply a
+  A \textit{graph} $\gcg = (\mathcal{V}, \gcge)$ is simply a
   tuple of sets respectively called \textit{nodes} and \textit{edges}.
@@
   \textit{directed} graphs, where the edges
-  $\mathcal{E} \subseteq V \times V$ are \textit{ordered} pairs.
+  $\gcge \subseteq V \times V$ are \textit{ordered} pairs.
 
@@
   $a_0, a_1, \ldots, a_{r - 1}, a_r$ with $a_0 = i$ and $a_r = j$ such
-  that $\forall\ 0 \le k \le r\ (a_k, a_{k + 1}) \in \mathcal{E}$.
+  that $\forall\ 0 \le k \le r\ (a_k, a_{k + 1}) \in \gcge$.
 
@@
 
-\begin{lemma}
-  \label{lem:still_persistent}
-  If $\gcg$ is a persistent DAG, any subgraph formed by eliminating nodes
-  as well as all of the in and out edges thereof is still persistent.
-\end{lemma}
-\begin{proof}
-  \hl{TODO -- is this needed?}
-\end{proof}
+% \begin{lemma}
+%   \label{lem:still_persistent}
+%   If $\gcg$ is a persistent DAG, any subgraph formed by eliminating nodes
+%   as well as all of the in and out edges thereof is still persistent.
+% \end{lemma}
+% \begin{proof}
+%   \hl{TODO -- is this needed?}
+% \end{proof}
 
@@
 
-\begin{proposition}
-  \label{prop:independent_confounders}
-  If $\gcg$ is strongly causal we have distinct
-  $k_1, k_2 \in \anc{i} \cap \anc{j}$ are confounders of $i, j$, then
-  $k_1 \not\in \anc{k_2}$, $k_2 \not\in \anc{k_1}$ and
-  $\not\exists \ell \in \anc{k_1} \cap \anc{k_2}$ confounding
-  $k_1, k_2$.
-\end{proposition}
-\begin{proof}
-  \hl{TODO}
-\end{proof}
-
-\begin{corollary}
-  If $\gcg$ is strongly causal, then distinct confounders are
-  uncorrelated.
-\end{corollary}
-\begin{proof}
-  This follows directly by combining proposition
-  \ref{prop:independent_confounders} with proposition \ref{prop:ancestor_properties}.  \hl{I need to update prop} \ref{prop:ancestor_properties} \hl{to note the strong non-correlation property.}
-\end{proof}
+% \begin{proposition}
+%   \label{prop:independent_confounders}
+%   If $\gcg$ is strongly causal we have distinct
+%   $k_1, k_2 \in \anc{i} \cap \anc{j}$ are confounders of $i, j$, then
+%   $k_1 \not\in \anc{k_2}$, $k_2 \not\in \anc{k_1}$ and
+%   $\not\exists \ell \in \anc{k_1} \cap \anc{k_2}$ confounding
+%   $k_1, k_2$.
+% \end{proposition}
+% \begin{proof}
+%   \hl{TODO}
+% \end{proof}
+
+% \begin{corollary}
+%   If $\gcg$ is strongly causal, then distinct confounders are
+%   uncorrelated.
+% \end{corollary}
+% \begin{proof}
+%   This follows directly by combining proposition
+%   \ref{prop:independent_confounders} with proposition \ref{prop:ancestor_properties}.  \hl{I need to update prop} \ref{prop:ancestor_properties} \hl{to note the strong non-correlation property.}
+% \end{proof}
 
@@
 \begin{proposition}
-  Suppose $\gcg$ is a strongly causal DAG and that $x(t)$ is persistent DAG, then if $k$
-  confounds $i, j$ we have $i \pwgc j$ and $j \pwgc i$.
+  \label{prop:persistence_converse}
+  Suppose $\gcg$ is a strongly causal DAG and that $x(t)$ is persistent, then if $k$
+  confounds $(i, j)$ we have $i \pwgc j$ and $j \pwgc i$.
 \end{proposition}
@@
 
-  To this end, let $F(z)$ be an arbitrary but strictly causal linear,
-  and equation \eqref{eqn:ancestor_expansion} to $x_i(t)$ and
-  $\psi \defeq F(z)x_j(t)$:
+  To this end, let $F(z)$ be an arbitrary but strictly causal linear
+  filter.  We apply equation \eqref{eqn:ancestor_expansion} to
+  $x_i(t)$ and $\psi \defeq F(z)x_j(t)$:
 
@@
   $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$, which is impossible
-  since in general $\tau_0(\A_{jk}) < \infty$ and in a persistent
-  graph $\tau_\infty(\A_{ik}) = \infty$.
+  since in general $\tau_0(\A_{jk}) < \infty$ and for a persistent system
+  $\tau_\infty(\A_{ik}) = \infty$.
 \end{proof}
@@
   \label{thm:tree_recovery}
-  If the Granger-Causality graph $\gcg$ for $x(t)$ is a strongly
-  causal DAG then $\gcg$ can be inferred from pairwise causality
-  tests alone.  The procedure can be carried out, assuming we have an
-  oracle for pairwise causality, via algorithm (\ref{alg:pwgr}).
+  If the Granger-causality graph $\gcg$ for persistent process $x(t)$
+  is a strongly causal DAG then $\gcg$ can be inferred from pairwise
+  causality tests alone.  The procedure can be carried out, assuming
+  we have an oracle for pairwise causality, via algorithm
+  (\ref{alg:pwgr}).
 
@@
     % \TitleOfAlgo{Pairwise Graph Recovery}
-    \Input{Pairwise Granger-causality relations between $n$ processes whose joint Granger-causality relations are known to form a strongly causal DAG $\gcg$.}
-    \Output{Edges $E = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of the graph $\gcg$.}
+    \Input{Pairwise Granger-causality relations between a persistent
+      processes of dimension $n$ whose joint Granger-causality
+      relations are known to form a strongly causal DAG $\gcg$.}
+    \Output{Edges $\gcge = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of
+      the graph $\gcg$.}
     \Initialize{$S_0 = [n]$  \texttt{\# unprocessed nodes}\\
@@
       % $P_0 = \emptyset$  \texttt{\# layer by layer driving nodes}\\
-      $k = 0$ \texttt{\# a counter used only for notation}}
+      $k = 1$ \texttt{\# a counter used only for notation}}
     \BlankLine
-    $W \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$  \texttt{\# potential edges}\\
-    \While{$S_k \ne \emptyset$}{
-      $P_k \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W\}$  \texttt{\# parents of current layer}\\
-      % $P_k \leftarrow P_{k - 1} \cup P_k$ \texttt{\# Expand parent set}\\
-      $S_{k + 1} \leftarrow S_k \setminus P_k$ \texttt{\# remove from } $S$\\
-
+    $W \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$  \texttt{\# candidate edges}\\
+    $P_0 \leftarrow \{i \in S_0\ |\ \forall s \in S_0\ (s, i) \not\in W\}$  \texttt{\# parentless nodes}\\
+    \While{$S_{k - 1} \ne \emptyset$}{
+      $S_k \leftarrow S_{k - 1} \setminus P_{k - 1}$ \texttt{\# remove nodes with depth }$k - 1$\\
+      $P_k \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W\}$   \texttt{\# candidate children of }$P_{k - 1}$\\
       \;
-      $C_k \leftarrow \{j \in S_{k + 1}\ |\ \forall s \in S_{k + 1}\ (s, j) \not\in W\}$   \texttt{\# potential children of }$P_k$\\
-
-      \For{$r = 0, \ldots, k$}
+
+      $D_{k0} \leftarrow \emptyset$\\
+      \For{$r = 1, \ldots, k$} 
       {
-        \label{alg:inner_loop}
-        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times C_k\ |\ (i, j) \in W,\ \text{no } i \rightarrow j \text{ path in } E_k \}$\\%  \texttt{\# edges incident on }$C_k$\\
-        $E_k \leftarrow E_k \cup D_{kr}$ \texttt{\# add edges to }$E_k$\\
+        $Q \leftarrow E_{k - 1} \cup \big(\bigcup_{\ell = 0}^{r - 1} D_{k\ell}\big)$ \texttt{\# currently known edges}\\
+        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times P_k\ |\ (i, j) \in W,\ \text{no } \gcgpath{i}{j} \text{ path in } Q\}$
+        % $E_k \leftarrow E_{k - 1} \cup D_{kr}$ \texttt{\# add edges to }$E_k$ \label{alg:inner_loop_end}\\
         % $W_{k + 1} \leftarrow W_k \setminus D_{kr}$  \texttt{\# remove edges from consideration}\\
       }
+      $E_k \leftarrow E_{k - 1} \cup \big(\bigcup_{r = 1}^k D_{kr}\big)$ \texttt{\# update } $E_k$ \texttt{ with new edges}\\
+      % \label{alg_line:inner_loop_end}\\
       % $W_{k + 1} \leftarrow W_k \setminus \big(\bigcup_{r = 0}^k D_{kr}\big)$  \texttt{\# remove edges from consideration}\\
-
       \;
@@
     }
-    $E \leftarrow E_{k + 1}$\\
-    \Return{$E$}
+    \Return{$E_{k - 1}$}
   \end{algorithm}
@@
 
-We prove the theorem by establishing the correctness of algorithm (\ref{alg:pwgr}).
-The idea is to iteratively ``peel away layers'' of nodes by removing the nodes that have no parents remaining, which always exist since the graph is acyclic.
-The additional requirement of strong causality ensures firstly that all actual edges of $\gcg$ manifest in some way as pairwise relations (by proposition \ref{prop:pwgc_anc}), and secondly, as we will see, that parents of $C_k$ are in $\bigcup_{r = 0}^kP_{k - r}$ and can be determined by a pairwise test.
+We will shortly prove the theorem by establishing the correctness of
+algorithm (\ref{alg:pwgr}).  The idea is to iteratively ``peel away
+layers'' of nodes by removing the nodes that have no parents
+remaining, which always exist since the graph is acyclic.  The
+requirement of strong causality ensures that all actual edges of
+$\gcg$ manifest in some way as pairwise relations (by proposition
+\ref{prop:pwgc_anc}), and the requirement of persistence allows
+confounding to be eliminated by removing bidirectional edges.  Without
+persistence, each confounded pair would give rise to $4$ possible
+pairwise topologies consistent with $\gcg$, one for each type of
+pairwise edge (no edge, unidirectional, bidirectional).
 
 \begin{example}
-  The set $W$ collects all of the edges that may be edges of $\gcg$.
-  In reference to figure \ref{fig:example_fig3}, each of the solid
-  black edges, as well as the dotted red edges will be included, but
-  \textit{not} the bidirectional green dash-dotted edge.  That we exclude
-  bidirectional edges is important since it will provide guarantees
-  that $P_k$ is non-empty.  The groupings $P_0, \ldots, P_3$
-  are also indicated in figure \ref{fig:example_fig3}.
+  The set $W$ collects ancestor relations in $\gcg$ (see Lemma
+  \ref{lem:W_subset_E}).  In reference to figure
+  \ref{fig:example_fig3}, each of the solid black edges, as well as
+  the dotted red edges will be included in $W$, but \textit{not} the
+  bidirectional green dash-dotted edges, which we are able to exclude
+  by appealing to Corollary \ref{cor:bidirectional_edge}.  The
+  groupings $P_0, \ldots, P_3$ are also indicated in figure
+  \ref{fig:example_fig3}.
 
@@
   $1 \rightarrow 3 \rightarrow 4$ and $1 \rightarrow 3 \rightarrow 5$
-  in E.  Finally, edge $(4, 6)$ are added, and the false
+  already present.  Finally, edge $(4, 6)$ is added, and the false
   $(1, 6), (3, 6), (2, 6)$ edges are similarly excluded due to the
@@
       relations.  Red dotted arrows indicate pairwise causality (due to
-      ancestor relations), green dash-dotted arrows indicates
+      non-parent relations), green dash-dotted arrows indicates
       bidirectional pairwise causality (due to the confounding node
@@
   That we need to proceed backwards through $P_{k - r}$ as in the
-  inner loop of line \ref{alg:inner_loop} can also be seen from this
-  example, where if instead
-  $D_k = \{(i, j) \in \Big(\bigcup_{r = 0}^k P_{k - r}\Big) \times
-  C_k\ |\ i \pwgc j \}$ we would infer the false positive edge
-  $1 \rightarrow 4$.  Moreover, the same example shows that if simply
-  $D_k = \{(i, j) \in P_k \times C_k\ |\ i \pwgc j \}$, then the edge
-  $1 \rightarrow 3$ would be missed.
+  inner loop of on $r$ can also be seen from this example, where if
+  instead we simply added the set
+
+  \begin{equation*}
+    D_k' = \{(i, j) \in \Big(\bigcup_{r = 1}^k P_{k - r}\Big) \times P_k\ |\ i \pwgc j \}
+  \end{equation*}
+
+  to $E_k$ then we would infer the false positive edge
+  $1 \rightarrow 4$.  Moreover, the same example shows that simply
+  using the set
+
+  \begin{equation*}
+    D_k'' = \{(i, j) \in P_{k - 1} \times P_k\ |\ i \pwgc j \}  ,
+  \end{equation*}
+
+  causes the edge $1 \rightarrow 3$ to be missed.
 \end{example}
 
-\begin{lemma}
-  \label{lem:inner_loop_lemma}
-  At any given iteration, $k$, of Algorithm \ref{alg:pwgr}, if we have
-  $e \in E$ if and only if $e \in \mathcal{E}$, then the inner loop (line \ref{alg:inner_loop})
-  will add $(i, j) \in \big(\bigcup_{r = 0}^kP_{k - r}\big) \times C_k$ to $E$
-  if $(i, j) \in \mathcal{E}$.  i.e. any real edge will be included.
+% \begin{lemma}
+%   \label{lem:inner_loop_lemma}
+%   At any given iteration, $k$, of Algorithm \ref{alg:pwgr}, if we have
+%   $e \in E$ if and only if $e \in \gcge$, then the inner loop (line \ref{alg:inner_loop})
+%   will add $(i, j) \in \big(\bigcup_{r = 0}^kP_{k - r}\big) \times C_k$ to $E$
+%   if $(i, j) \in \gcge$.  i.e. any real edge will be included.
+% \end{lemma}
+% \begin{proof}
+%   Let $j \in C_k$, $i \in \bigcup_{r = 0}^k P_{k - r}$, and assume
+%   $i \in \pa{j}$.
+
+%   First, suppose that there is no $i \rightarrow j$ path in $E$.  Then
+%   since $i \gc j$ we have $i \pwgc j$ and $j \npwgc i$ (i.e.
+%   $(i, j) \in W$) by Proposition \ref{prop:pwgc_anc} and Corollary
+%   \ref{cor:bidirectional_edge} so at least one of
+%   $\{D_{kr} \}_{r = 0}^k$ contains $(i, j)$ and hence $(i, j)$ will be
+%   added to $E$ upon the end of the iteration.
+
+%   Now, suppose instead that there is already an $i \rightarrow j$ path
+%   in $E$.  Given our assumption that $E$ contains edges if and only if
+%   they are edges of $\gcg$, then either we already have $(i, j) \in E$
+%   correctly, or there is more than one $i \rightarrow j$ path in
+%   $\gcg$, which is not possible since $\gcg$ is strongly causal.
+% \end{proof}
+
+% \begin{lemma}
+%   \label{lem:W_subset_E}
+%   $W \subseteq \gcge$
+% \end{lemma}
+% \begin{proof}
+%   By proposition \ref{prop:pwgc_anc} we have
+%   $j \in \pa{i} \Rightarrow j \pwgc i$ and by Corollary \ref{cor:bidirectional_edge}
+%   there are no bidirectional edges in $\gcg$.
+% \end{proof}
+
+% \hl{Must show that Algorithm terminates in $n$ steps.}
+% \begin{lemma}
+%   For any $i \in P_k$ we have $i \in C_{k - 1}$.  \hl{[Can't yet prove
+%     this part] Moreover, for each $i \in \bigcup_{k = 0}^n P_k$, if
+%     $(s, i) \in W$ then $(s, i) \in E$ upon the termination of
+%     Algorithm} \ref{alg:pwgr}.
+% \end{lemma}
+% \begin{proof}
+%   We prove the second part by induction, the first part being
+%   established along the way.  By construction $P_0$ has no parent
+%   nodes in $\gcg$ so the base case is immediate.  Fix now some
+%   $k \ge 1$, $i \in P_k$, and let $s$ be such that $(s, i) \in W$.
+%   For every $s$ such that $(s, i) \in W$ we have $s \not\in S_k$ since
+%   otherwise $i \not\in P_k$.  Therefore
+%   $s \in \bigcup_{r = 0}^{k - 1}P_{k - r}$, which implies
+%   $s \not\in S_k$, and finally that $i \in C_{k - 1}$.  Then by Lemma
+%   \ref{lem:W_subset_E} every edge in $\gcge$ is also present in
+%   $W$, and by Lemma \ref{lem:inner_loop_lemma} \hl{[We can't conclude
+%     here because of the strong conditions required on lemma}
+%     \ref{lem:inner_loop_lemma}]
+% \end{proof}
+
+% \begin{lemma}[Loop Invariant]
+%   \label{lem:loop_invariant}
+%   $\forall k \ge 0,\ e \in E_k \Rightarrow e \in \gcge$.
+%   \hl{Not sure if this is useful / the right approach.}
+% \end{lemma}
+% \begin{proof}
+%   We will proceed by induction.  Firstly, the base case
+%   $E_0 = \emptyset$ is trivial, so fix some $k \ge 1$ and consider an
+%   edge $(i, j) \in \big(\bigcup_{r = 0}^k P_{k - r} \big) \times C_k$
+%   such that $(i, j) \in W$ (if the edge is not in $W$ then it is not
+%   in $\gcge$ by Lemma \ref{lem:W_subset_E}).
+
+%   \hl{WIP}
+% \end{proof}
+
+Our proof proceeds in 5 steps stated formally as lemmas.  Firstly, we
+characterize the sets $W$ and $P_k$.  Then we establish a correctness
+result for the inner loop on $r$, a correctness result for the outer
+loop on $k$, and finally that the algorithm terminates in a finite
+number of steps.
+
+\begin{lemma}[$W$ Represents Ancestor Relations]
+  \label{lem:W_subset_E}
+  In Algorithm \ref{alg:pwgr} we have
+  $(i, j) \in W$ if and only if $i \in \anc{j}$.  In particular,
+  $W \subseteq \gcge$.
 \end{lemma}
 \begin{proof}
-  Let $j \in C_k$, $i \in \bigcup_{r = 0}^k P_{k - r}$, and assume
-  $i \in \pa{j}$.
-
-  First, suppose that there is no $i \rightarrow j$ path in $E$.  Then
-  since $i \gc j$ we have $i \pwgc j$ and $j \npwgc i$ (i.e.
-  $(i, j) \in W$) by Proposition \ref{prop:pwgc_anc} and Corollary
-  \ref{cor:bidirectional_edge} so at least one of
-  $\{D_{kr} \}_{r = 0}^k$ contains $(i, j)$ and hence $(i, j)$ will be
-  added to $E$ upon the end of the iteration.
-
-  Now, suppose instead that there is already an $i \rightarrow j$ path
-  in $E$.  Given our assumption that $E$ contains edges if and only if
-  they are edges of $\gcg$, then either we already have $(i, j) \in E$
-  correctly, or there is more than one $i \rightarrow j$ path in
-  $\gcg$, which is not possible since $\gcg$ is strongly causal.
-\end{proof}
-
-\begin{lemma}
-  \label{lem:W_subset_E}
-  $W \subseteq \mathcal{E}$
-\end{lemma}
-\begin{proof}
-  By proposition \ref{prop:pwgc_anc} we have
-  $j \in \pa{i} \Rightarrow j \pwgc i$ and by Corollary \ref{cor:bidirectional_edge}
-  there are no bidirectional edges in $\gcg$.
-\end{proof}
-
-\hl{Must show that Algorithm terminates in $n$ steps.}
-\begin{lemma}
-  For any $i \in P_k$ we have $i \in C_{k - 1}$.  \hl{[Can't yet prove
-    this part] Moreover, for each $i \in \bigcup_{k = 0}^n P_k$, if
-    $(s, i) \in W$ then $(s, i) \in E$ upon the termination of
-    Algorithm} \ref{alg:pwgr}.
-\end{lemma}
-\begin{proof}
-  We prove the second part by induction, the first part being
-  established along the way.  By construction $P_0$ has no parent
-  nodes in $\gcg$ so the base case is immediate.  Fix now some
-  $k \ge 1$, $i \in P_k$, and let $s$ be such that $(s, i) \in W$.
-  For every $s$ such that $(s, i) \in W$ we have $s \not\in S_k$ since
-  otherwise $i \not\in P_k$.  Therefore
-  $s \in \bigcup_{r = 0}^{k - 1}P_{k - r}$, which implies
-  $s \not\in S_k$, and finally that $i \in C_{k - 1}$.  Then by Lemma
-  \ref{lem:W_subset_E} every edge in $\mathcal{E}$ is also present in
-  $W$, and by Lemma \ref{lem:inner_loop_lemma} \hl{[We can't conclude
-    here because of the strong conditions required on lemma}
-    \ref{lem:inner_loop_lemma}]
-\end{proof}
-
-\begin{lemma}[Loop Invariant]
-  \label{lem:loop_invariant}
-  $\forall k \ge 0,\ e \in E_k \Rightarrow e \in \mathcal{E}$.
-  \hl{Not sure if this is useful / the right approach.}
-\end{lemma}
-\begin{proof}
-  We will proceed by induction.  Firstly, the base case
-  $E_0 = \emptyset$ is trivial, so fix some $k \ge 1$ and consider an
-  edge $(i, j) \in \big(\bigcup_{r = 0}^k P_{k - r} \big) \times C_k$
-  such that $(i, j) \in W$ (if the edge is not in $W$ then it is not
-  in $\mathcal{E}$ by Lemma \ref{lem:W_subset_E}).
-
-  \hl{WIP}
+  Let $j \in [n]$ and suppose that $i \in \anc{j}$.  Then $i \pwgc j$
+  by Proposition \ref{prop:pwgc_anc}.  Proposition
+  \ref{prop:sc_graph_common_anc} ensures that $(i, j)$ are not
+  confounded and Corollary \ref{cor:parent_corollary} that
+  $j \not\in \anc{i}$ so $j \npwgc i$ and thence by Proposition
+  \ref{prop:ancestor_properties} $(i, j) \in W$.
+
+  Convsersely, suppose $(i, j) \in W$.  Then since $j \npwgc i$
+  Proposition \ref{prop:persistence_converse} ensures that $(j, i)$
+  are not confounded and so by Proposition \ref{prop:ancestor_properties}
+  we must have $i \in \anc{j}$.
 \end{proof}
@@
 
-\begin{lemma}
+\begin{lemma}[Depth Characterization of $P_k$]
+  \label{lem:depth_lemma}
   $i \in P_k \iff d(i) = k$ and $j \in S_k \iff d(j) \ge k$.
@@
 \begin{proof}
-  We proceed by induction.  The base case $j \in P_0 \iff d(j) = 0$ is
-  by definition, and $j \in S_0 \iff d(j) \ge 0$ is trivial since
-  $S_0 = [n]$.  So suppose that the lemma is true up to $k - 1$.
-
-  For $i \in P_k$ we have by construction that there is no $j \in S_k$
-  such that $(j, i) \in W$ and therefore no $j \in S_k$ such that
-  $j \in \pa{i}$.  Therefore, any parent of $i$ can only be in
-  $S_{k - r}$ for $r \ge 1$ and by the induction hypothesis which
-  implies that $d(i) \le k$.  But if $d(i) < k$ then $i \in P_{k - r}$
-  for some $r \ge 1$ (again by the induction hypothesis) thence
-  $d(i) = k$ since $P_k \cap P_{k - r} = \emptyset$ by construction.
-
-  Now for $s \in S_k$ we have by induction that and the fact that
-  $s \in S_k \implies s \in S_{k - 1}$ by construction we have
+  We proceed by induction, noting that $P_0$ is non-empty since $\gcg$
+  is acyclic and therefore $\gcg$ contains nodes without parents.  The
+  base case $i \in P_0 \iff d(i) = 0$ is by definition, and
+  $j \in S_0 \iff d(j) \ge 0$ is trivial since $S_0 = [n]$.  So
+  suppose that the lemma is true up to $k - 1$.
+
+  ($i \in P_k \implies d(i) = k$): Let $i \in P_k$.  Suppose that
+  $d(i) \ge k + 1$, then $\exists j \in \pa{i}$ such that
+  $j \not\in \cup_{r \ge 1}P_{k - r}$ (otherwise $d(i) \le k$), this
+  implies that $j \in S_k$ with $(j, i) \in W$ (by Lemma
+  \ref{lem:W_subset_E}) which is not possible due to the construction of
+  $P_k$ and therefore $d(i) \le k$.  Moreover,
+  $P_k \subseteq S_k \subseteq S_{k - 1}$ implies that
+  $d(i) \ge k - 1$ by the induction hypothesis, but if $d(i) = k - 1$
+  then $i \in P_{k - 1}$ again by induction which is impossible since
+  $i \in P_k$ and therefore $d(i) = k$.
+
+  ($s \in S_k \implies d(s) \ge k$): Let
+  $s \in S_k \subseteq S_{k - 1}$.  We have by induction that
   $d(s) \ge k - 1$, but again by induction (this time on $P_{k - 1}$)
-  we have $d(s) \ne k$ since $S_k = S_{k - 1} \setminus P_{k - 1}$
+  we have $d(s) \ne k - 1$ since $S_k = S_{k - 1} \setminus P_{k - 1}$
   and therefore $d(s) \ge k$.
 
-  Conversely, suppose $i \in [n]$ is such that $d(i) = k$.  Then
-  $i \in S_{k - 1}$ by the hypothesis, but also $i \not\in P_{k - 1}$
-  so then $i \in S_k = S_{k - 1} \setminus P_{k - 1}$.  Now, recalling
-  the definition of $P_k$
+  ($d(i) = k \implies i \in P_k$): Suppose $i \in [n]$ is such that
+  $d(i) = k$.  Then $i \in S_{k - 1}$ by the hypothesis, but also
+  $i \not\in P_{k - 1}$ so then
+  $i \in S_k = S_{k - 1} \setminus P_{k - 1}$ and thus $d(i) \ge k$.
+  Now, recalling the definition of $P_k$
 
@@
 
-  if $s \in S_k$ is such that $s \pwgc i$ then by Proposition
-  \ref{prop:ancestor_properties} $(a) s \in \anc{i}$ or
-  $(b) \exists u \in \anc{i} \cap \anc{j}$ which is a confounder.  If
-  we have $s \in \anc{i}$ then since $d(s) \ge k$ we must have
-  $d(i) > k$, which is a contradiction so $s \not\in \anc{i}$.  On the
-  other hand, if we have a confounding $u \in \anc{i} \cap \anc{s}$
-  then we must have $d(u) < k$ (since $s \in \anc{i}$ and $d(i) = k$)
-  which implies by induction that $u \in P_{k - r}$ for some $r \ge 1$,
-  and therefore that $u \not\in ???$  \hl{This is wrong.}
-
-  Now let $j \in S_k$, then $d(j) \ge k - 1$ since we
-  also have $j \in S_{k - 1}$ (i.e. $S_k$ is a decreasing sequence of
-  sets) \hl{Can't conclude unless $i \in P_k \iff d(i) = k$, in which
-    case, we have $S_k = S_{k - 1} \setminus P_{k - 1}$ and therefore
-    $d(j) \ge k$.}
+  if $s \in S_k$ is such that $(s, i) \in W$ then $s \pwgc i$ and
+  $i \npwgc s$ so that by persistence and Proposition
+  \ref{prop:persistence_converse} there cannot be a counfounder of
+  $(s, i)$ (otherwise $i \pwgc s$) so then by Proposition
+  \ref{prop:ancestor_properties} we have $s \in \anc{i}$.  We have
+  shown that $s \in S_k \implies d(s) \ge k$ and so we must have
+  $d(i) > k$, a contradiction, thence $s \not\in \anc{i}$,
+  $s \npwgc i$, $(s, i) \not\in W$ and $i \in P_k$.
+
+  ($d(j) \ge k \implies j \in S_k$): Let $j \in [n]$ such that
+  $d(j) \ge k$, then by induction we have $j \in S_{k - 1}$.  This
+  implies by the construction of $S_k$ that $j \not\in S_k$ only if
+  $j \in P_{k - 1}$, but we have shown that this only occurs when
+  $d(j) = k - 1$, but $d(j) > k - 1$ so $j \in S_k$.
 \end{proof}
 
-\begin{lemma}[Loop Invariant]
-  For every $(i, j) \in \mathcal{E}$ such that $d(j) \le k$ we have
-  $(i, j) \in E_k$.  Conversely, for every $(i, j) \in E_{k}$ we have
-  $d(j) \le k$ and $(i, j) \in \mathcal{E}$.  That is, at iteration
-  $k, E_k$ and $\mathcal{E}$ agree on the set of edges whose
-  terminating node is at most $k$ steps away from $P_0$.
+\begin{lemma}[Inner Loop]
+  \label{lem:inner_loop_lemma}
+  Fix an integer $k \ge 1$ and suppose that $(i, j) \in E_{k - 1}$ if
+  and only if $(i, j) \in \gcge$ and $d(j) \le k - 1$.  Then, we have
+  $(i, j) \in D_{kr}$ if and only if $(i, j) \in \gcge $, $d(j) = k$,
+  and $d(i) = k - r$.
+\end{lemma}
+\begin{proof}
+  We prove by induction on $r$, keeping in mind the results of Lemmas
+  \ref{lem:W_subset_E} and \ref{lem:depth_lemma}.  For the base case,
+  let $r = 1$ and suppose that $(i, j) \in \gcge$ with $d(j) = k$ and
+  $d(i) = k - 1$.  Then, $(i, j) \in W$ and by our assumptions on
+  $E_{k - 1}$ there is no $\gcgpath{i}{j}$ path in $E_{k - 1}$
+  and therefore $(i, j) \in D_{k1}$.  Conversely, suppose that
+  $(i, j) \in D_{k1}$.  Then, $d(i) = k - 1$ and $d(j) = k$ which, since
+  $(i, j) \in W \implies i \in \anc{j}$ implies that
+  $i \in \pa{j}$ and $(i, j) \in \gcge$.
+
+  Now, fix $r > 1$ and suppose that the result holds up to $r - 1$.
+  Let $(i, j) \in \gcge$ with $d(j) = k$ and $d(i) = k - r$.  Then,
+  $(i, j) \in W$ and by induction and strong causality there cannot
+  already be an $\gcgpath{i}{j}$ path in
+  $E_{k - 1} \cup \big(\bigcup_{\ell = 0}^{r - 1} D_{kr}\big)$,
+  therefore $(i, j) \in D_{kr}$.  Conversely, suppose
+  $(i, j) \in D_{kr}$.  Then we have $d(i) = k - r$, $d(j) = k$, and
+  $i \in \anc{j}$.  Suppose by way of contradiction that
+  $i \not\in \pa{j}$, then there must be some $u \in \pa{j}$ such that
+  $i \in \anc{u}$.  But, this implies that $d(i) < d(u)$ and by
+  induction that $(u, j) \in \bigcup_{\ell = 1}^{r - 1}D_{k\ell}$.
+  Morevoer, sice $d(u) < k$ (otherwise $d(j) > k$) each edge in
+  the $\gcgpath{i}{u}$ path must already be in $E_{k - 1}$, and so
+  there must be an $\gcgpath{i}{j}$ path in
+  $E_{k - 1}\cup\big(\bigcup_{\ell = 0}^{r - 1}D_{kr}\big)$, which is
+  a contradiction since we assumed $(i, j) \in D_{kr}$.  Therefore
+  $i \in \pa{j}$ and $(i, j) \in \gcge$.
+\end{proof}
+
+\begin{lemma}[Outer Loop]
+  \label{lem:outer_loop_lemma}
+  We have $(i, j) \in E_k$ if and only if $(i, j) \in \gcge$ and
+  $d(j) \le k$.  That is, at iteration $k, E_k$ and $\gcge$ agree on
+  the set of edges whose terminating node is at most $k$ steps away
+  from $P_0$.
 \end{lemma}
@@
   all nodes of depth less than $k$.
+
+  Suppose that
+  $(i, j) \in E_k = E_{k - 1}\cup \big(\bigcup_{r = 1}^k D_{rk}
+  \big)$.  Then clearly there is some $1 \le r \le k$ such that
+  $(i, j) \in D_{kr}$ so that by Lemma \ref{lem:inner_loop_lemma} we
+  have $(i, j) \in \gcge$ and $d(j) = k$.
+
+  Conversely, suppose that $(i, j) \in \gcge$ and $d(j) \le k$.  If
+  $d(j) < k$ then by induction $(i, j) \in E_{k - 1} \subseteq E_k$ so
+  suppose further than $d(j) = k$.  Since $i \in \pa{j}$ we must have
+  $d(i) < k$ (else $d(j) > k$) and again by Lemma
+  \ref{lem:inner_loop_lemma} $(i, j) \in \bigcup_{r = 1}^k D_{kr}$
+  which implies that $(i, j) \in E_k$.  The result follows.
 \end{proof}
 
-%   Finally, suppose that $i \not \in \pa{j}$.  If $i \npwgc j$ then
-%   we will correctly leave $(i, j)$ absent from $E$, so suppose that
-%   $i \pwgc j$.  Then we must have either $i \in \anc{j}$ or a confounding 
-%   $k \in \anc{i} \cap \anc{j}$; suppose first the former.
-%   Then, there is some $u$ s.t. $i \pwgc u \pwgc j$, so by the
-%   construction of \hl{it seems} that both $(i, u)$ and $(u, j)$ should
-%   have already been added to $E$, so we have an $i \rightarrow j$ path
-%   in $E$ \hl{is this sound?}  Suppose then the latter case above,
-%   $u \in \anc{i} \cap \anc{j}$ \hl{try to construct a counterexample
-%     involving this case}.
-% \end{proof}
-
-Proof of the actual theorem...
-
+\begin{lemma}[Finite Termination]
+  Algorithm \ref{alg:pwgr} terminates and returns $E_{k^\star - 1} = \gcge$
+  for some $k^\star \le n$.
+\end{lemma}
 \begin{proof}
-  Denote the edges of $\gcg$ by $\mathcal{E}$, we endevour to show
-  that upon the termination of Algorithm \ref{alg:pwgr} we have
-  $E = \mathcal{E}$.  On the first iteration of the algorithm
-  ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
-
-  We will first dispense with some trivialities: if $n = 0$ or $n = 1$
-  the algorithm is correct, returning on the first iteration with
-  $E = \emptyset$.  Suppose then that $n > 1$.  The set
-  $P_0$ is non-empty since $\gcg$ is a DAG and therefore
-  there are nodes in $\gcg$ without parents (if
-  $P_0 = \emptyset$ proposition
-  \ref{prop:ancestor_properties} would be contradicted).  Now, if
-  $P_0 = [n]$ then the graph necessarily contains no edges
-  and we will correctly return $E = \emptyset$.
-
-  Moving on from any trivial cases, notice that
-  $\mathcal{E} \subseteq W$.  This follows since by Proposition
-  \ref{prop:pwgc_anc} we have $j \in \pa{i} \Rightarrow j \pwgc i$ and by
-  Corollary \ref{cor:bidirectional_edge} there are no bidirectional
-  edges in $\gcg$.
+  If $n = 1$, the algorithm is clearly correct, returning on the first
+  iteration with $E_1 = \emptyset$.  When $n > 1$ Lemma
+  \ref{lem:outer_loop_lemma} ensures that $E_k$ coincides with
+  $\{(i, j) \in \gcge\ |\ d(j) \le k\}$ and since $d(j) \le n - 1$ for
+  any $j \in [n]$ there is some $k^\star \le n$ such that
+  $E_{k^\star - 1} = \gcge$.  We must have $S_{k^\star} = \emptyset$
+  since $j \in S_{k^\star} \iff d(j) \ge k^\star$ (if $d(j) > k - 1$ then
+  $E_{k^\star - 1} \ne \gcge$) and therefore the algorithm terminates.
+\end{proof}
+
+% \begin{proof}
+%   Denote the edges of $\gcg$ by $\gcge$, we endevour to show
+%   that upon the termination of Algorithm \ref{alg:pwgr} we have
+%   $E = \gcge$.  On the first iteration of the algorithm
+%   ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
+
+%   Moving on from any trivial cases, notice that
+%   $\gcge \subseteq W$.  This follows since by Proposition
+%   \ref{prop:pwgc_anc} we have $j \in \pa{i} \Rightarrow j \pwgc i$ and by
+%   Corollary \ref{cor:bidirectional_edge} there are no bidirectional
+%   edges in $\gcg$.
 
@@
 
-\hl{WIP}
-\end{proof}
+% \hl{WIP}
+% \end{proof}
 
 \begin{example}
-  Being a strongly causal DAG is merely a sufficient condition.
-For example, the complete directed graph with 2 nodes (e.g.
-$B(1) = \left[ \begin{array}{cc} 1/2 & 1 \\ 1 & 1/2 \end{array} \right]$) contains a loop but is pairwise recoverable, though not by algorithm (\ref{alg:pwgr}).
-Clearly, this example is somewhat artificial since when $n = 2$ there is no difference between pairwise Granger-causality and joint Granger-causality amongst all series -- however, one can take any strongly causal DAG, and then add two additional nodes having such a loop without impacting the pairwise recoverability.
+  We close this section by noting that the conditions of persistence
+  and strong causality are only sufficient conditions.  For example,
+  the complete directed graph with 2 nodes i.e.
+
+  \begin{equation*}
+    B(1) = \left[ \begin{array}{cc} 1/2 & 1 \\ 1 & 1/2 \end{array}\right]
+  \end{equation*}
+
+  contains a loop but is pairwise recoverable, though not by algorithm
+  (\ref{alg:pwgr}).  Clearly, this example is somewhat artificial
+  since when $n = 2$ there is no difference between pairwise
+  Granger-causality and joint Granger-causality amongst all series --
+  however, one can add any number of nodes having no parents or
+  children to a graph containing a length 2 cycle, in which case the
+  graph clearly remains pairwise recoverable.
 \end{example}

### 2019-05-16 paper/paper.tex
Updated figures.

@@
 \begin{definition}[Graph Theory Review]
-  A \textit{graph} $\gcg = (\mathcal{V}, \gcge)$ is simply a
+  A \textit{graph} $\gcg = (V, \gcge)$ is simply a
   tuple of sets respectively called \textit{nodes} and \textit{edges}.

### 2019-05-16 paper/paper.tex
Broke ancestor_properties into 2 props.  Must fix references to them.

@@
 \usepackage[backend=bibtex,sorting=none]{biblatex}
+\usepackage{enumitem}
 \bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
@@
     \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} = 0$}
-    \item{$\H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
+    \item{$\H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$ \hl{Modify to look just at $H_t$}}
     \item{$\linE{x_i(t)}{\H_{t - 1}^{-j}} = \linE{x_i(t)}{\H_{t - 1}}$}
@@
   \begin{equation}
+    \label{eqn:parent_expansion}
     x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t),
@@
   where $\wtB_{ii}(z) = (1 - \B_{ii}(z))^{-1}$ and
-  $\wtB_{ik}(z) = \wtB_{ii}(z) B_{ik}(z)$.
+  $\wtB_{ik}(z) = \wtB_{ii}(z) B_{ik}(z)$.  \hl{Is this true: }$\wtB_{ii} = \A_{ii}$ ?
 
@@
 
-\begin{definition}[\hl{Confounder}]
+\begin{proposition}
+  \label{prop:ancestor_properties}
+  Consider two distinct nodes $i, j$ of a Granger-causality graph
+  $\gcg$.  Then, if
+
+  \begin{enumerate}[label=(\alph*)]
+    \item{$j \not\in \anc{i}$}
+    \item{$\anc{i}\cap\anc{j} = \emptyset$}
+  \end{enumerate}
+
+  then $H_t^{(i)} \perp \H_t^{(j)}$, that is,
+  $\forall \tau \in \Z_+\ \E[x_i(t)x_j(t - \tau)] = 0$.  Moreover,
+  this implies that $j \npwgc i$.
+\end{proposition}
+
+\begin{remark}
+  Note that $H_t^{(i)} \perp \H_t^{(j)}$ states that the
+  \textit{present} of $x_i(t)$ is uncorrelated with the \textit{past}
+  of $x_j(t)$, but we need not have $\H_t^{(i)} \perp \H_t^{(j)}$
+  since $i \in \anc{j}$ is not excluded.
+
+  We also point out the contrapositive that if $j \pwgc i$ then we
+  must have either $j \in \anc{i}$ or $\anc{i} \cap \anc{j} \ne 0$.
+\end{remark}
+
+\begin{proof}
+  We show directly that
+  $\forall \tau \in \Z_+\ \E[x_i(t)x_j(t - \tau)] = 0$.  To this end,
+  fix $\tau \ge 0$, then by expanding with equation \eqref{eqn:parent_expansion} we have
+
+  \begin{align*}
+    \E x_i(t)x_j(t - \tau) &= \E \big(\A_{ii}(z)v_i(t)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)\\
+    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\E[\big(\A_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
+    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ik}(z)v_k(t)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)].
+  \end{align*}
+
+% (i.e.
+%   $\forall \tau \in \Z,\ k \ne \ell \Rightarrow \E v_k(t)v_\ell(t -
+%   \tau) = 0$ and
+%   $\forall \tau \in \Z\setminus\{0\}\ \E v_i(t)v_i(t - \tau) = 0$),
+
+  Keeping in mind that $v(t)$ is an isotropic and uncorrelated
+  sequence we will see that each of these above four terms are 0.  The
+  first term since $i \ne j$.  The second since $j \not\in \anc{i}$.
+  The fourth since $\anc{i} \cap \anc{j} = \emptyset$.  For the third
+  term, it is possible that $i \in \anc{j}$, which requires more care.
+
+  Similarly as before we can see immediately that
+
+  \begin{equation*}
+    \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)] = \E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)].
+  \end{equation*}
+
+  However, $k \ne \ell \Rightarrow A_{k\ell}(0) = 0$ (i.e.
+  $\A(z) = I + A(1)z^{-1} + \cdots$) so $\A_{ji}(z)v_i(t - \tau)$ is
+  strictly earlier in time (for $\tau \ge 0$) than $v_i(t)$ and hence
+  $\E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)] = 0$.
+\end{proof}
+
+The previous proposition can be strengthened significantly for the
+case of Granger-causality.  Note that it is possible to have some
+$k \in \anc{i} \cap \anc{j}$ where still $j \npwgc i$, for example
+consider the three node graph $k \rightarrow i \rightarrow j$ where
+clearly $k \in \anc{i}\cap\anc{j}$ but $j \npwgc i$.  We must
+introduce the concept of a \textit{confounding} variable, which
+effectively eliminates the possibility presented in this example.
+
+\begin{definition}[Confounder]
   A node $k$ will be referred to as a \textit{confounder} of nodes
@@
 \begin{proposition}
-  \label{prop:ancestor_properties}
-  \hl{Add the stronger statements on uncorrelatedness for use on confounders later.}
-
-  In a Granger-causality graph $\gcg$, if $j \pwgc i$ then at least
-  one of the following must hold
-
-  \begin{enumerate}
-    \item{$j \in \anc{i}$}
-    \item{$\anc{i} \cap \anc{j} \ne \emptyset$.}
-  \end{enumerate}
-
-  Moreover, if $j \not\in \anc{i}$ then $\exists u \in \anc{i} \cap \anc{j}$ which is a confounder, that is, such that $i$ is not contained in at least one $\gcgpath{u}{j}$ path.
+  If in a Granger-causality graph $\gcg$ we have $j \pwgc i$ then
+  $j \in \anc{i}$ or $\exists k \in \anc{i} \cap\anc{j}$ which is a
+  confounder of $(i, j)$.
 \end{proposition}
 
-\begin{remark}
-  Considering the contrapositive of the first part of this theorem (which is what we actually prove below) is enlightening: it says that if $i$ is not a descendent of $j$ (i.e.
-$j \not \in \anc{i}$) and that there are no common ancestors (i.e.
-$\anc{i} \cap \anc{j} = \emptyset$) then $j \npwgc i$, or $\linE{x_i(t)}{\H_{t - 1}^{(i, j)}} = \linE{x_i(t)}{\H_{t - 1}^{(i)}}$, which is an intuitively natural result.
-
-  The first part is used in part to prove the second (and stronger) statement, which allows common ancestors, but only if they are ``mediated'' by $i$.
-For example, this covers the case of a graph $k \rightarrow i \rightarrow j$, where $\{k\} = \anc{i} \cap \anc{j}$, but clearly $j \npwgc i$.
-\end{remark}
-
 \begin{proof}
-  We will first establish the contrapositive of $(1)$ and $(2)$: assume that $j \not \in \anc{i}$ and $\anc{i} \cap \anc{j} = \emptyset$, then we will show that $\forall \tau \in \Z_+,\ \E x_i(t)x_j(t - \tau) = 0$.
-This follows from representation (\ref{eqn:parent_expansion}) since we have
-
-  \begin{align*}
-    \E x_i(t)x_j(t - \tau) &= \E \big(\A_{ii}(z)v_i(t)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)\\
-    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\E[\big(\A_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
-    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ik}(z)v_k(t)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)]\\
-  \end{align*}
-
-  Keeping in mind that $v(t)$ is an isotropic and uncorrelated sequence (i.e.
-$\forall \tau \in \Z,\ k \ne \ell \Rightarrow \E v_k(t)v_\ell(t - \tau) = 0$ and $\forall \tau \in \Z\setminus\{0\}\ \E v_i(t)v_i(t - \tau) = 0$), we will see that each of these above four terms are 0.
-The first term since $i \pwgc i$ is not possible and hence the case $j = i$ is vacuous.
-The second term is 0 since $j \not\in \anc{i}$.
-The fourth term above is 0 since $\anc{i} \cap \anc{j} = \emptyset$.
-For the third term, it is possible that (due to cycles) $i \in \anc{j}$ which requires more care, although similarly as before we can at least see immediately that
-
-  \begin{equation*}
-    \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)] = \E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)].
-  \end{equation*}
-
-  However, $k \ne \ell \Rightarrow \A_{k\ell}(z) = 0$ (i.e.
-$\A(z) = I + A(1)z^{-1} + \cdots$) so $\A_{ji}(z)v_i(t - \tau)$ is strictly earlier in time, for $\tau \ge 0$, than $v_i(t)$ and hence $\E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)] = 0$.
-This establishes the fact that $H_t^{(i)} \perp \H_t^{(j)}$ where $H_t^{(i)} = \{\alpha x_i(t)\ |\ \alpha \in \R\}$ is the Hilbert space generated by the current sample of $x_i$.
-It is clear that $\H_t^{(i)} = \bigvee_{\tau = 0}^\infty H_{t - \tau}^{(i)}$.
-
-  \hl{TODO: Separate $H_t^i \perp \H_t^j \Rightarrow \H_t^i \perp \H_{t - 1}^j\ |\ \H_{t - 1}^i$ into a lemma and consolidate it with lemma 2, 3.}
-
-  Then for every $s \ge 0$
-
-  \begin{align*}
-    \linE{x_j(t - s)}{\H_t^{(i)}} &= \linE{x_j(t - s)}{\bigvee_{\tau = 0}^\infty H_{t - \tau}^{(i)}}\\
-    &\overset{(a)}{=} \linE{x_j(t - s)}{\bigvee_{\tau = 1}^\infty H_{t - \tau}^{(i)}}\\
-    &= \linE{x_j(t - s)}{\H_{t - 1}^{(i)}},
-  \end{align*}
-
-  where $(a)$ follows from lemma (\ref{lem:subspace_sum_projection}).  Lemma (\ref{lem:conditional_orthogonality_equivalence}) tells us that this is equivalent to
-
-  \begin{equation*}
-    \H_t^{(i)} \perp \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(i)},
-  \end{equation*}
-
-  so by theorem (\ref{thm:granger_causality_equivalences}) we conclude $j \npwgc i$.
-
-  For the second part of the theorem we will suppose that $j$ is a node having the following two properties: $(a)$ $j \not \in \anc{i}$ and $(b)$ for every $k \in \anc{i} \cap \anc{j}$ every $k \rightarrow \cdots \rightarrow j$ path contains $i$.
-The first is the negation of property $(1)$ above, and the second is to setup the contradiction $j \npwgc i$, which will then establish the theorem.
-
-  Firstly, notice that every $u \in \big(\pa{j} \setminus \{i\}\big)$ necesssarily inherits these same two properties since if $u \in \anc{i}$ then $u \in \anc{i} \cap \anc{j}$ so that every $u \rightarrow \cdots \rightarrow j$ path must contain $i$, but $u \in \pa{j}$, so this is not the case since $u \rightarrow j$ is a path that doesn't contain $i$; moreover, if we consider $w \in \anc{i} \cap \anc{u}$ then we also have $w \in \anc{i} \cap \anc{j}$ so every $w \rightarrow \cdots \rightarrow j$ path must contain $i$.
-These properties therefore extend inductively to every $u \in \big(\anc{j} \setminus \{i\}\big)$.
-
-  In order to deploy a recursive argument, define the following partition of $\pa{u}$, for some node $u$:
+  We will prove by way of contradiction.  To this end, suppose that
+  $j$ is a node such that: $(a)$ $j \not \in \anc{i}$ and $(b)$ for
+  every $k \in \anc{i} \cap \anc{j}$ every
+  $k \rightarrow \cdots \rightarrow j$ path contains $i$.
+
+  Firstly, notice that every $u \in \big(\pa{j} \setminus \{i\}\big)$
+  necesssarily inherits these same two properties.  This follows since
+  if we also had $u \in \anc{i}$ then $u \in \anc{i} \cap \anc{j}$ so that every
+  $u \rightarrow \cdots \rightarrow j$ path must contain $i$, but
+  $u \in \pa{j}$, so this is not the case since $u \rightarrow j$ is a
+  path that doesn't contain $i$; moreover, if we consider
+  $w \in \anc{i} \cap \anc{u}$ then we also have
+  $w \in \anc{i} \cap \anc{j}$ so every
+  $w \rightarrow \cdots \rightarrow j$ path must contain $i$.  These
+  properties therefore extend inductively to every
+  $u \in \big(\anc{j} \setminus \{i\}\big)$.
+
+  In order to deploy a recursive argument, define the following
+  partition of $\pa{u}$, for some node $u$: (\hl{TODO: Draw a figure depicting these sets})
 
@@
 
-  We notice that for any $u$ having the properties $a, b$ above, we must have $C_2(u) = \emptyset$ since if $k \in C_2(u)$ then $\exists w \in \anc{i} \cap \anc{k}$ s.t.
-$i \not \in \anc{k}$ and therefore there must be a path $w \rightarrow \cdots \rightarrow k \rightarrow \cdots \rightarrow u$ which does not contain $i$.
-
-  Using this partition, we will expand $x_j(t)$ in terms of it's parents, and recursively expand nodes in $C_1$ until we reach a case where $C_1 = \emptyset$.
-For the first step:  %where $\A_{jj}(z) = (1 - \B_{jj}(z))^{-1}$:
+  We notice that for any $u$ having the properties $a, b$ above, we
+  must have $C_2(u) = \emptyset$ since if $k \in C_2(u)$ then
+  $\exists w \in \anc{i} \cap \anc{k}$ s.t. $i \not \in \anc{k}$ and
+  therefore there must be a path $\gcgpath{w}{k} \rightarrow u$ which
+  does not contain $i$.
+
+  Using this partition, we will expand $x_j(t)$ in terms of it's
+  parents, and recursively expand nodes in $C_1$ until we reach a case
+  where $C_1 = \emptyset$.  For the first step equation
+  \eqref{eqn:parent_expansion} gives us:
 
@@
 
-  which will imply (\hl{by lemma ???}) that $j \npwgc i$ and for which it is equivalent to show that
+  which will imply (by Theorem \ref{thm:granger_causality_equivalences}) that $j \npwgc i$ and for which it is equivalent to show that
 
@@
 
-  which follows from the first part of the theorem.
-Finally, for $k \in C_1(j)$ the case $k = i$ is immediate, so suppose $k \ne i$.
-We know from above that $k$ inherits the key properties referred to as $(a), (b)$ and therefore we can recursively expand $k$ in the same way as in equation (\ref{eqn:xj_partition_expansion}).
-Continuing this recursion for each $k \in C_1(j)$ where $k \ne i$ must eventually terminate since $i \in \anc{k}$.
+  which follows from Proposition \ref{prop:ancestor_properties}.
+  Finally, for $k \in C_1(j)$ the case $k = i$ is immediate (since the
+  error in estimating $\Phi(z)\B_{ii}(z)x_i(t)$ given
+  $\H_{t - 1}^{(i)}$ is $0$), so suppose $k \ne i$.  We know from
+  above that $k$ inherits the key properties referred to as $a, b$ and
+  therefore we can recursively expand $k$ in the same way as in
+  equation (\ref{eqn:xj_partition_expansion}).  Continuing this
+  recursion for each $k \in C_1(j)$ (where $k \ne i$) must eventually
+  terminate since $i \in \anc{k}$.
 \end{proof}

### 2019-05-17 paper/paper.tex
Shifted around various sections and removed some superfluous statements.

@@
 \usepackage{enumitem}
+\usepackage{etex,etoolbox}
 \bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
@@
 We will work with a discrete time and wide-sense stationary process
-$x(t)$ taking elements in $L_2$.  We suppose that $x(t)$ has zero
-mean, $\E x(t) = 0$, and the matrix valued covariance sequence
-$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ is absolutely summable and has a spectra $S(\omega)$:
+$x(t)$ taking elements in $L_2$.  We suppose that $x(t)$ (with
+$t \in \Z$) has zero mean, $\E x(t) = 0$, and the matrix valued
+covariance sequence
+$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ is absolutely
+summable and has a spectra $S(\omega)$:
 
@@
   \begin{aligned}
-    \H^x &= \cl \{\sum_{\tau = -T}^T a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, T \in \N\} \subseteq L_2(\Omega),\\
-    \H_t^x &= \cl \{\sum_{\tau = 0}^T a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, T \in \N\} \subseteq L_2(\Omega),
+    % \H^x &= \cl \{\sum_{\tau = -T}^T a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, T \in \N\} \subseteq L_2(\Omega),\\
+    \H_t^x &= \cl \{\sum_{\tau = 0}^T a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, T \in \N\} \subseteq L_2(\Omega)\\
+    H_t^x &= \{a x(t)\ |\ a \in \R\} \subseteq L_2(\Omega),
   \end{aligned}
@@
 ``now'', we will often omit the superscript $x$ which should be clear
-from context.  Evidently these spaces are separable and as closed
+from context.  Evidently these spaces are separable, and as closed
 subspaces of a Hilbert space they are themselves Hilbert.  We will
@@
 
+We require two technical facts we refer to in the sequel, the notion
+of conditional orthogonality is used throughout.
+
+\begin{lemma}[\cite{lindquist} Proposition 2.4.2]
+  \label{lem:conditional_orthogonality_equivalence}
+  Consider three closed subspaces of a Hilbert space $\mathcal{A}$,
+  $\mathcal{B}$, $\mathcal{X}$.  The following statements are
+  equivalent
+
+  \begin{enumerate}
+    \item{$\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$}
+    \item{$\linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}}\ \forall \beta \in \mathcal{B}.$}
+    % \item{$\linE{\beta}{\mathcal{A}} = \linE{\linE{\beta}{\mathcal{X}}}{\mathcal{A}}\ \forall \beta \in \mathcal{B}}$}
+    \end{enumerate}
+
+    Where $\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$ denotes
+    conditional orthogonality:
+
+    \begin{equation*}
+      \inner{a - \linE{a}{\mathcal{X}}}{b - \linE{b}{\mathcal{X}}} = 0\ \forall a \in \mathcal{A}, b \in \mathcal{B}.
+    \end{equation*}
+\end{lemma}
+% \begin{proof}
+%   Proceding from the definition of conditional orthogonality we have
+
+%   \begin{align*}
+%     &\inner{\alpha - \linE{\alpha}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
+%     &{\iff} \inner{\alpha + x - \linE{\alpha + x}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
+%     &\overset{(a)}{\iff} \inner{\alpha + x}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
+%     &\overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A} \vee \mathcal{X}} = 0\ \forall \beta \in \mathcal{B}\\
+%     &\overset{(c)}{\iff} \linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}},
+%   \end{align*}
+
+%   where $(a)$ follows since $(\beta - \linE{\beta}{\mathcal{X}}) \perp \mathcal{X}$ and $\linE{\alpha + x}{\mathcal{X}} \in \mathcal{X}$, $(b)$ since $z \perp \mathcal{X} \vee \mathcal{A} \iff \linE{z}{\mathcal{X} \vee \mathcal{A}} = 0$, and $(d)$ since $\linE{z}{\H_2} = \linE{\linE{z}{\H_2}}{\H_1}$ for $\H_2 \subseteq \H_1$.
+
+%   % This is property (vi) in Lindquist
+%   % \begin{align*}
+%   %   \langle \alpha - \linE{\alpha}{\mathcal{X}}, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
+%   %   \overset{(a)}{\iff} \langle \alpha, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0\ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
+%   %   \overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A}} &= 0\ \forall \beta \in \mathcal{B}
+%   % \end{align*}
+
+%   % where $(a)$ follows because $\linE{\alpha}{\mathcal{X}} \in \mathcal{X}$ and $\beta - \linE{\beta}{\mathcal{X}} \in \mathcal{X}$, and $(b)$ follows since $\beta \perp \mathcal{A} \iff \linE{\beta}{\mathcal{A}} = 0$.
+% \end{proof}
+\begin{lemma}
+  \label{lem:subspace_sum_projection}
+  Let $x \in \H$ where $\H$ is a separable Hilbert space having inner product
+  $\inner{\cdot}{\cdot}$ and let $\{\H_n\}_{n = 0}^N$ be a
+  collection of subspaces of $\H$ such that $x \perp \H_0$.  Then
+
+  \begin{equation*}
+    \linE{x}{\bigvee_{n = 0}^N\H_n} = \linE{x}{\bigvee_{n = 1}^N\H_n},
+  \end{equation*}
+
+  where
+  $\H_1 \vee \H_2 = \cl \{\alpha + \beta\ |\ \alpha \in \H_1, \beta
+  \in \H_2 \}$ is the closed sum of subspaces.
+\end{lemma}
+\begin{proof}
+  Apply the Gram-schmidt process and directly calculate the projections.
+\end{proof}
+
 \begin{theorem}[Granger Causality Equivalences]
@@
   Let $x(t)$ be a WSS process with absolutely summable covariance
-  sequence, and spectral density uniformly bounded above and below.  Denote by
-  $\xi_{ij} \defeq \linEerr{x_i(t)}{\H^{-j}}$ and
-  $\xi_i \defeq \linEerr{x_i(t)}{\H_t}$.  Then, the following are equivalent:
+  sequence, and spectral density uniformly bounded above and below.
+  Denote by $\xi_{ij} \defeq \linEerr{x_i(t)}{\H^{-j}}$ and
+  $\xi_i \defeq \linEerr{x_i(t)}{\H_t}$ for distinct $i, j$.  Then,
+  the following are equivalent:
 
@@
     \item{$x_j \ngc x_i$}
-    \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$}
+    \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$ i.e. $\B_{ij}(z) = 0$}
     \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} = 0$}
-    \item{$\H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$ \hl{Modify to look just at $H_t$}}
+    \item{$\H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j} \iff H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
     \item{$\linE{x_i(t)}{\H_{t - 1}^{-j}} = \linE{x_i(t)}{\H_{t - 1}}$}
@@
 
-Statement (4) means that
-$(y - \linE{y}{\H_{t - 1}^{-j}}) \perp (z - \linE{z}{\H_{t -
-    1}^{-j}})\ \forall y \in \H_t^i,\ z \in \H_{t - 1}^j$, which is an
-appealing geometric statement that there is no direct feedback from
-$x_j$ to $x_i$ \cite{lindquist}.
-
 \begin{proof}
@@
 
+  We have $(4) \iff (5)$ as a result of Lemma
+  \ref{lem:conditional_orthogonality_equivalence}, the additional
+  equivalence in $(4)$ is apparent since $\H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$.
+
   $(1) \Rightarrow (2)$: The projection for $x_i(t)$ onto $\H_{t - 1}$ is
@@
 
-$(2) \Rightarrow (4)$: In computing $(y - \linE{y}{\H_{t = 1}^{-j}})$ for
-$y \in \H_t^i$ it is sufficient to consider $y = x_i(t)$ since
-$\H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$, in which case
-$(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$.  $(4)$ follows
+$(2) \Rightarrow (4)$: In computing $(y - \linE{y}{\H_{t - 1}^{-j}})$ for
+$y \in H_t^i$ it is sufficient to consider $y = x_i(t)$ by linearity, then since
+$H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$ we have
+$(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$ since $\B_{ij}(z) = 0$.  $(4)$ then follows
 since $v_i(t) \perp \H_{t - 1}$ and
@@
   - 1}$
-
-\hl{TODO: $(4) \Rightarrow (1)$}
-
 \end{proof}
@@
 
-\subsection{Pairwise Granger Causality}
-\label{sec:pwgc}
-Recall that Granger-causality in general must be understood with
-respect to a particular universe of observations.  If $x_j \gc x_i$
-with respect to $x_{-k}$, it may not hold with respect to $x$.  For
-example, $x_k$ may be a common ancestor which when observed completely
-explains the connection from $x_j$ to $x_i$.  In particular, we will
-study \textit{pairwise} Granger-causality.
-
-\begin{definition}[Pairwise Granger-causality]
-  We will say that $x_j$ pairwise Granger-causes $x_i$ and write
-  $x_j \pwgc x_i$ if $x_i$ Granger-causes $x_j$ with respect only to
-  $(x_i, x_j)$.
-\end{definition}
-
-This notion is of interest for a variety of reasons.  From a purely
-theoretical or conceptual standpoint, we will seek to understand when
-and how pairwise causality can capture the idea of ``flow of
-information'' in the underlying graph.  It will also be useful to
-reason about what the conditions under which \textit{unobserved}
-components of $x(t)$ may or may not interfere with inference in the
-actually observed components.  Finally, we will seek to construct
-practical estimation procedures based purely on pairwise causality
-tests and understand their properties.
-
+\subsection{Granger Causality Graphs}
 We first need to establish some graph theoretic notation and
@@
 
-\begin{definition}[Causality graph, Parents]
-
-Remaining in the context of a collection of $n$ WSS processes
-$x(t) = \big(x_1(t), \ldots, x_n(t)\big)$, we define the Granger-causality
-graph $\gcg$ to be the directed graph formed on $n$ vertices where an
-edge $(j, i) \in \gcg$ if and only if $x_j \gc x_i$.
-
-We refer to the set of \textit{parents} of $x_i$ in $\gcg$ by
-$\pa{i}$, that is, $j \in \pa{i} \iff x_j \gc x_i$.
-\end{definition}
-
 \begin{definition}[Grandparents, ancestors]
@@
 
+Our principle object of study will be a graph determined by
+Granger-causality relations as follows.
+
+\begin{definition}[Causality graph, Parents]
+
+Remaining in the context of a collection of $n$ WSS processes
+$x(t) = \big(x_1(t), \ldots, x_n(t)\big)$, we define the Granger-causality
+graph $\gcg$ to be the directed graph formed on $n$ vertices where an
+edge $(j, i) \in \gcg$ if and only if $x_j \gc x_i$.
+
+We refer to the set of \textit{parents} of $x_i$ in $\gcg$ by
+$\pa{i}$, that is, $j \in \pa{i} \iff x_j \gc x_i$.
+\end{definition}
+
 The edges of the Granger-causality graph $\gcg$ can be given a general
@@
 \begin{equation}
-  \sum_{\tau = 1}^\infty B_{ij}(\tau)x_j(t - \tau) \defeq \B_{ij}(z)x_j(t).
+  \B_{ij}(z)x_j(t) \defeq \sum_{\tau = 1}^\infty B_{ij}(\tau)x_j(t - \tau).
 \end{equation}
@@
 
-We continue to develop the relationship between $x_i$ and it's family tree.
+We continue to develop the relationship between $x_i$ and it's family
+tree, where the following proposition is analagous to the definitions
+of feedback free processes of \cite{caines1975feedback}.
 
@@
   where $\wtB_{ii}(z) = (1 - \B_{ii}(z))^{-1}$ and
-  $\wtB_{ik}(z) = \wtB_{ii}(z) B_{ik}(z)$.  \hl{Is this true: }$\wtB_{ii} = \A_{ii}$ ?
+  $\wtB_{ik}(z) = \wtB_{ii}(z) B_{ik}(z)$.
 
@@
 decomposition matrices $A(\tau)$ since
-$x_i(t) = \sum_{\tau = 0}^\infty \sum_{j = 1}^n a_{ij}(\tau)v_j(t -
-\tau)$.  The proposition states that $\A_{ij}(\tau) = 0$ unless
+$x_i(t) = \sum_{\tau = 0}^\infty \sum_{j = 1}^n A_{ij}(\tau)v_j(t -
+\tau)$.  The proposition states that $\A_{ij}(z) = 0$ unless
 $j \in \anc{i}$.
@@
 \begin{proof}
+  Equation \eqref{eqn:parent_expansion} was derived prior to the
+  statement of the theorem.  We are left to demonstrate equation
+  \eqref{eqn:ancestor_expansion}.
+
   From equation (\ref{eqn:ar_representation}), which we are assuming
@@
     \label{eqn:resolvant_inv}
-    (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k.
+    (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k,
   \end{equation}
 
-  Now, let $S$ be a $0-1$ matrix containing the sparsity pattern of
-  $\B(z)$.  Then from lemma \ref{lem:adj_matrix} $S$ is the transpose
-  of the adjacency matrix for $\gcg$ and hence $(S^k)_{ij}$ is
-  non-zero if and only if $j \in \gpn{k}{i}$, and therefore
-  $\B(z)^k_{ij} = 0$ if $j \not \in \gpn{k}{i}$.  The Cayley-Hamilton
-  theorem can then be applied to represent the infinite sum of equation
-  (\ref{eqn:resolvant_inv}) in terms only of \textit{finite} powers of
-  $\B(z)$, so the application of lemma \ref{lem:adj_matrix} in the
-  limit is not necessary.  Finally, since $\anc{i} = \cup_{k = 1}^n\gpn{k}{i}$
-  we see that $\A(z)_{ij}$ is zero if $j \not\in \anc{i}$.
+  where we can apply the Cayley-Hamilton theorem to write the infinite
+  sum of \eqref{eqn:resolvant_inv} in terms of \textit{finite} powers
+  of $\B$.
+
+  Let $S$ be a $0-1$ matrix containing the sparsity pattern of
+  $\B(z)$, from lemma \ref{lem:adj_matrix} $S$ is the transpose of the
+  adjacency matrix for $\gcg$ and hence $(S^k)_{ij}$ is non-zero if
+  and only if $j \in \gpn{k}{i}$, and therefore $\B(z)^k_{ij} = 0$ if
+  $j \not \in \gpn{k}{i}$.  Finally, since
+  $\anc{i} = \bigcup_{k = 1}^n\gpn{k}{i}$ we see that $\A_{ij}(z)$ is
+  zero if $j \not\in \anc{i}$.
 
@@
 
-Before moving onto our next proposition, we need two lemmas formalizing
-some intuitively apparent facts.
-
-\begin{lemma}
-  \label{lem:subspace_sum_projection}
-  Let $x \in \H$ where $\H$ is a Hilbert space having inner product $\inner{\cdot}{\cdot}$ and let $\{\H_n\}_{n = 0}^\infty$ be a collection of subspaces of $\H$ such that $x \perp \H_0$.  Then
-
-  \begin{equation*}
-    \linE{x}{\bigvee_{n = 0}^\infty\H_n} = \linE{x}{\bigvee_{n = 1}^\infty\H_n},
-  \end{equation*}
-
-  where $\H_1 \vee \H_2 = \cl \{\alpha + \beta\ |\ \alpha \in \H_1, \beta \in \H_2 \}$ is the closed sum of subspaces.
-\end{lemma}
-\begin{proof}
-  We can apply the Gram-schmidt process to the collection of subspaces to obtain a sequence of orthogonal subspaces $\widetilde{\H}_0, \widetilde{\H}_1, \ldots$ with $\widetilde{\H}_0 = \H_0$, $\widetilde{\H}_n \subseteq \H_n$, and $\bigvee_{n = 0}^\infty\H_n = \bigoplus_{n = 0}^\infty \widetilde{\H}_n$, where $\oplus$ indicates the orthogonal sum of subspaces (which is always closed).
-Then, we have an explicit representation of the projection
-
-  \begin{align*}
-    \linE{x}{\bigvee_{n = 0}^\infty\H_n} &= \sum_{n = 0}^\infty \inner{x}{x_n} x_n\\
-    &\overset{(a)}{=} \sum_{n = 1}^\infty \inner{x}{x_n} x_n,
-  \end{align*}
-
-  where $x_n \in \widetilde{\H}_n \subseteq \H_n$ and $(a)$ follows because of our assumption that $x \perp \H_0$.
-This establishes that the projection is infact a member of the smaller subspace $\bigvee_{n = 1}^\infty\H_n$, and since the error can only increase by projecting onto this smaller subspace, we have the conclusion.
-\end{proof}
-
-The following lemma is reproduced from \cite{lindquist} here for convenience.
-
-\begin{lemma}[\cite{lindquist}]
-  \label{lem:conditional_orthogonality_equivalence}
-  Consider three closed subspaces of a Hilbert space $\mathcal{A}$, $\mathcal{B}$, $\mathcal{X}$.  The following statements are equivalent
-
-  \begin{enumerate}
-    \item{$\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$}
-    \item{$\linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}}\ \forall \beta \in \mathcal{B}$}
-    % \item{$\linE{\beta}{\mathcal{A}} = \linE{\linE{\beta}{\mathcal{X}}}{\mathcal{A}}\ \forall \beta \in \mathcal{B}}$}
-  \end{enumerate}
-\end{lemma}
-\begin{proof}
-  Proceding from the definition of conditional orthogonality we have
-
-  \begin{align*}
-    &\inner{\alpha - \linE{\alpha}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
-    &{\iff} \inner{\alpha + x - \linE{\alpha + x}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
-    &\overset{(a)}{\iff} \inner{\alpha + x}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
-    &\overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A} \vee \mathcal{X}} = 0\ \forall \beta \in \mathcal{B}\\
-    &\overset{(c)}{\iff} \linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}},
-  \end{align*}
-
-  where $(a)$ follows since $(\beta - \linE{\beta}{\mathcal{X}}) \perp \mathcal{X}$ and $\linE{\alpha + x}{\mathcal{X}} \in \mathcal{X}$, $(b)$ since $z \perp \mathcal{X} \vee \mathcal{A} \iff \linE{z}{\mathcal{X} \vee \mathcal{A}} = 0$, and $(d)$ since $\linE{z}{\H_2} = \linE{\linE{z}{\H_2}}{\H_1}$ for $\H_2 \subseteq \H_1$.
-
-  % This is property (vi) in Lindquist
-  % \begin{align*}
-  %   \langle \alpha - \linE{\alpha}{\mathcal{X}}, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
-  %   \overset{(a)}{\iff} \langle \alpha, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0\ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
-  %   \overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A}} &= 0\ \forall \beta \in \mathcal{B}
-  % \end{align*}
-
-  % where $(a)$ follows because $\linE{\alpha}{\mathcal{X}} \in \mathcal{X}$ and $\beta - \linE{\beta}{\mathcal{X}} \in \mathcal{X}$, and $(b)$ follows since $\beta \perp \mathcal{A} \iff \linE{\beta}{\mathcal{A}} = 0$.
-\end{proof}
+\subsection{Pairwise Granger Causality}
+\label{sec:pwgc}
+Recall that Granger-causality in general must be understood with
+respect to a particular universe of observations.  If $x_j \gc x_i$
+with respect to $x_{-k}$, it may not hold with respect to $x$.  For
+example, $x_k$ may be a common ancestor which when observed completely
+explains the connection from $x_j$ to $x_i$.  In particular, we will
+study \textit{pairwise} Granger-causality.
+
+\begin{definition}[Pairwise Granger-causality]
+  We will say that $x_j$ pairwise Granger-causes $x_i$ and write
+  $x_j \pwgc x_i$ if $x_j$ Granger-causes $x_i$ with respect only to
+  $(x_i, x_j)$.
+\end{definition}
+
+This notion is of interest for a variety of reasons.  From a purely
+theoretical or conceptual standpoint, we will seek to understand when
+and how pairwise causality can capture the idea of ``flow of
+information'' in the underlying graph.  It may also be useful for
+reasoning about what the conditions under which \textit{unobserved}
+components of $x(t)$ may or may not interfere with inference in the
+actually observed components.  Finally, motivated from a practical
+standpoint to analyze causation in large systems, we will seek to
+construct practical estimation procedures based purely on pairwise
+causality tests which can result in reduced computational complexity.
 
 \begin{proposition}
-  \label{prop:ancestor_properties}
-  Consider two distinct nodes $i, j$ of a Granger-causality graph
+  \label{prop:ancestor_uncorrelated}
+  Consider distinct nodes $i, j$ in a Granger-causality graph
   $\gcg$.  Then, if
@@
   Note that $H_t^{(i)} \perp \H_t^{(j)}$ states that the
-  \textit{present} of $x_i(t)$ is uncorrelated with the \textit{past}
-  of $x_j(t)$, but we need not have $\H_t^{(i)} \perp \H_t^{(j)}$
-  since $i \in \anc{j}$ is not excluded.
+  \textit{present} of $x_i(t)$ is uncorrelated with $x_j(t)$,
+  but we need not have $\H_t^{(i)} \perp \H_t^{(j)}$ since
+  $i \in \anc{j}$ is not excluded.
 
@@
 case of Granger-causality.  Note that it is possible to have some
-$k \in \anc{i} \cap \anc{j}$ where still $j \npwgc i$, for example
-consider the three node graph $k \rightarrow i \rightarrow j$ where
+$k \in \anc{i} \cap \anc{j}$ where still $j \npwgc i$, consider for example
+the three node graph $k \rightarrow i \rightarrow j$ where
 clearly $k \in \anc{i}\cap\anc{j}$ but $j \npwgc i$.  We must
@@
 \begin{proposition}
+  \label{prop:ancestor_properties}
   If in a Granger-causality graph $\gcg$ we have $j \pwgc i$ then
@@
   In order to deploy a recursive argument, define the following
-  partition of $\pa{u}$, for some node $u$: (\hl{TODO: Draw a figure depicting these sets})
+  partition of $\pa{u}$, for some node $u$:
 
@@
 
-  which follows from Proposition \ref{prop:ancestor_properties}.
+  which follows from Proposition \ref{prop:ancestor_uncorrelated}.
   Finally, for $k \in C_1(j)$ the case $k = i$ is immediate (since the
@@
 \begin{corollary}
-  If the graph $\gcg$ is acyclic and if $j \pwgc i$ and $i \pwgc j$ then $\anc{i} \cap \anc{j} \ne \emptyset$.
+  If the graph $\gcg$ is acyclic and if $j \pwgc i$ and $i \pwgc j$ then $\exists k \in \anc{i} \cap \anc{j}$ confounding $(i, j)$.
 \end{corollary}
 
-It seems intuitive that a converse of proposition \ref{prop:ancestor_properties}
-would hold, i.e. $j \in \anc{i} \Rightarrow j \pwgc i$.  Unfortunately,
-this is not the case in general, as different paths through $\gcg$ can
-lead to cancellation (see example \ref{ex:diamond_cancellation}).  In
-fact, we do not even have $j \in \pa{i} \Rightarrow j \pwgc i$ (see
-example \ref{ex:lag_cancellation}).
+It seems reasonable to expect a converse of proposition
+\ref{prop:ancestor_properties} to hold, i.e.
+$j \in \anc{i} \Rightarrow j \pwgc i$.  Unfortunately, this is not the
+case in general, as different paths through $\gcg$ can lead to
+cancellation (see example \ref{ex:diamond_cancellation}).  In fact, we
+do not even have $j \in \pa{i} \Rightarrow j \pwgc i$ (see example
+\ref{ex:lag_cancellation}).
 
@@
 statements of Proposition \ref{prop:ancestor_properties} \textit{do}
-hold.  One possibility is to restrict to coefficients of the system
-matrix matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead,
+hold.  One possibility is to restrict the coefficients of the system
+matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead,
 we think it is more meaningful to focus on the defining feature of
@@
 forests), DAGs where each node has at most one parent, and figure
-\ref{fig:example_fig3} of this paper.  The persistence condition is
-used to ensure that $\linE{x_i(t)}{\H_{t - 1}^{(i)}} \ne 0$ i.e. that
-the each process has at least some amount of ``memory'' -- this can
-occur either due to cycles, or non-zero $\B_{ii}(z)$ terms in equation
-\eqref{eqn:ar_representation}.  It is evident that the strong causal property
-is inherited by subgraphs, but persistence need not be, unless $\gcg$ is a DAG.
-
-\hl{Not sure these are needed, but were important in my own thinking.}
-
-\begin{lemma}
-  \label{lem:still_strongly_causal}
-  If $\gcg$ is a strongly causal graph, any subgraph formed by eliminating nodes
-  as well as all of the in and out edges thereof is still strongly causal.
-\end{lemma}
-\begin{proof}
-  If $\gcg$ has at most one path between any two nodes, there can only
-  be fewer paths after removing nodes from $\gcg$.
-\end{proof}
+\ref{fig:example_fig3} of this paper.  It is evident that the strong
+causal property is inherited by subgraphs.
 
 % \begin{lemma}
-%   \label{lem:still_persistent}
-%   If $\gcg$ is a persistent DAG, any subgraph formed by eliminating nodes
-%   as well as all of the in and out edges thereof is still persistent.
+%   \label{lem:still_strongly_causal}
+%   If $\gcg$ is a strongly causal graph, any subgraph formed by eliminating nodes
+%   as well as all of the in and out edges thereof is still strongly causal.
 % \end{lemma}
 % \begin{proof}
-%   \hl{TODO -- is this needed?}
+%   If $\gcg$ has at most one path between any two nodes, there can only
+%   be fewer paths after removing nodes from $\gcg$.
 % \end{proof}
 
-We explore a number of properties of strongly causal graphs before
-moving into the main result of section \ref{sec:theory} in section
-\ref{sec:pairwise_algorithm}.  The following important property
-essentially strengthens proposition \ref{prop:ancestor_properties}
-for the case of strongly causal graphs.
+For later use, and to get a feel for the topological implications of
+strong causality We explore a number of properties of such graphs
+before moving into the main result of this section.  The following
+important property essentially strengthens proposition
+\ref{prop:ancestor_properties} for the case of strongly causal graphs.
 
@@
 
-  and therefore that $H_t^{(i)} \not\perp\ \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(i)}$, which by theorem (\ref{thm:granger_causality_equivalences}) and \hl{lemma [name the lemma]} is enough to establish that $j \pwgc i$.
+  and therefore that $H_t^{(i)} \not\perp\ \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(i)}$, which by theorem (\ref{thm:granger_causality_equivalences}) is enough to establish that $j \pwgc i$.
 
@@
            &= \wtalpha{a_0}{a_1} + \B_{a_0a_1}(z)\big[\wtalpha{a_1}{a_2} + \B_{a_1a_2}(z)x_{a_2}(t) \big]\\
-           &\overset{(a)}{=} \cdots\\
-           &= \sum_{\ell = 0}^r \underbrace{\Big(\prod_{m = 0}^{\ell - 1} \B_{a_m a_{m + 1}}(z) \Big)}_{\defeq F_\ell(z)} \wtalpha{a_\ell}{a_{\ell + 1}} + \Big(\prod_{m = 0}^{r}\B_{a_m a_{m + 1}}(z)\Big)x_{a_{r + 1}}(t)\\
+           &\overset{(a)}{=} \sum_{\ell = 0}^r \underbrace{\Big(\prod_{m = 0}^{\ell - 1} \B_{a_m a_{m + 1}}(z) \Big)}_{\defeq F_\ell(z)} \wtalpha{a_\ell}{a_{\ell + 1}} + \Big(\prod_{m = 0}^{r}\B_{a_m a_{m + 1}}(z)\Big)x_{a_{r + 1}}(t)\\
            &= \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}} + F_{r + 1}(z) x_j(t)
@@
 
+% \begin{remark}
+%   It can be seen from the conclusion of the proof that simple facts
+%   about Granger causality are indeed reliant on our assumptions about
+%   the nature of $x(t)$ laid out in the section \ref{sec:theory}, in
+%   particular, the innovations process $v(t)$ must be full rank.
+
+%   Moreover, that strong conditions need to be placed on the topology
+%   of $\gcg$ in order for proposition \ref{prop:pwgc_anc} to hold can
+%   be seen through the earlier examples.
+% \end{remark}
+
+% This holds in any DAG
+% \begin{lemma}
+%   If $\gcg$ is strongly causal, then any $\VAR(p)$ system on $\gcg$ is stable if and only if
+%   $B_{ii}(z)$ is stable for every $i = 1, \ldots, n$.
+% \end{lemma}
+% \begin{proof}
+%   \hl{This is certainly true, but I don't think I need to use it anywhere.}
+% \end{proof}
+
 \begin{remark}
-  It can be seen from the conclusion of the proof that simple facts
-  about Granger causality are indeed reliant on our assumptions about
-  the nature of $x(t)$ laid out in the section \ref{sec:theory}, in
-  particular, the innovations process $v(t)$ must be full rank.
-
-  Moreover, that strong conditions need to be placed on the topology
-  of $\gcg$ in order for proposition \ref{prop:pwgc_anc} to hold can
-  be seen through the earlier examples.
-\end{remark}
-
-\begin{lemma}
-  If $\gcg$ is strongly causal, then any $\VAR(p)$ system on $\gcg$ is stable if and only if
-  $B_{ii}(z)$ is stable for every $i = 1, \ldots, n$.
-\end{lemma}
-\begin{proof}
-  \hl{This is certainly true, but I don't think I need to use it anywhere.}
-\end{proof}
-
-\begin{remark}
-  \hl{Possibly expand on this thought}
-
-  As in much of statistics, confounding nodes pose challenges for
-  Granger-causality.  However, as opposed to Pearl's causal calculus
-  \cite{pearl2000art}, pairwise Granger-causality does not suffer any
-  difficulty with so-called ``colliders'', that is, the topology
-  $i \rightarrow k \leftarrow j$ will never result in $i \pwgc j$ or
-  $j \pwgc i$.  This is evidently an advantage of the \textit{temporal}
-  nature of Granger causality -- ``information'' cannot flow backwards
-  along the edges of $\gcg$.
+  As we have seen and as is true in much of statistics, confounding
+  nodes pose challenges for Granger-causality.  However, as opposed to
+  Pearl's causal calculus \cite{pearl2000art}, pairwise
+  Granger-causality does not suffer any difficulty with so-called
+  ``colliders'', that is, the topology $i \rightarrow k \leftarrow j$
+  will never result in $i \pwgc j$ or $j \pwgc i$.  This is evidently
+  an advantage of the \textit{temporal} nature of Granger causality --
+  ``information'' cannot flow backwards along the edges of $\gcg$.
 \end{remark}
@@
   If we were to augment this system by simply adding an autoregressive
-  component to $x_1(t)$ e.g.  $x_1(t) = v_1(t) + b x_1(t - 1)$ then we
-  \textit{would} have $2 \pwgc 3$ since then
-  $x_3(t) = v_3(t) + av_1(t - 1) - bv_2(t - 1) + bx_2(t - 1)$.
+  component (i.e. some ``memory'') to $x_1(t)$ e.g.
+  $x_1(t) = v_1(t) + b x_1(t - 1)$ then we \textit{would} have
+  $2 \pwgc 3$ since then
+  $x_3(t) = v_3(t) + av_1(t - 1) - bv_2(t - 1) + bx_2(t - 1)$.  We
+  develop this idea further in the next section.
 \end{example}
@@
 In section \ref{sec:strongly_causal_graphs} we obtained a converse to
-the part $(a)$ of proposition \ref{prop:ancestor_properties} via the
+part $(a)$ of proposition \ref{prop:ancestor_properties} via the
 notion of a strongly causal graph topology.  In this section, we
@@
 \end{remark}
+
+\hl{TODO: Show formally that any $\VAR(p)$ model is persistent.}
 
@@
 \label{sec:pairwise_algorithm}
-In this section we will show that if $\gcg$ is a strongly causal DAG,
-then it is possible to recover $\gcg$ via pairwise tests alone.  This
-is an interesting fact in and of itself, but also has some
-implications for applications since pairwise testing is trivially
-parallelizable.  In section \ref{sec:structure_learning} we will also
-analyze the use of pairwise testing as a heuristic for general graphs.
+In this section we will show that if the $\gcg$ of a persistent
+process is a strongly causal DAG, then it is possible to recover
+$\gcg$ via pairwise tests alone.
+
+% This is an interesting fact in and
+% of itself, but also has some implications for applications since
+% pairwise testing is trivially parallelizable.  In section
+% \ref{sec:structure_learning} we will also analyze the use of pairwise
+% testing as a heuristic for general graphs.
 
@@
   is a strongly causal DAG then $\gcg$ can be inferred from pairwise
-  causality tests alone.  The procedure can be carried out, assuming
-  we have an oracle for pairwise causality, via algorithm
+  causality tests.  The procedure can be carried out, assuming
+  we have an oracle for pairwise causality, via Algorithm
   (\ref{alg:pwgr}).
@@
       relations.  Red dotted arrows indicate pairwise causality (due to
-      non-parent relations), green dash-dotted arrows indicates
+      non-parent relations), green dash-dotted arrows indicate
       bidirectional pairwise causality (due to the confounding node

### 2019-05-19 paper/paper.tex
Heuristic algorithm and a bit of code.

@@
 forests), DAGs where each node has at most one parent, and figure
-\ref{fig:example_fig3} of this paper.  It is evident that the strong
-causal property is inherited by subgraphs.
+\ref{fig:example_fig3} of this paper.  A complete bipartite graph with
+$2n$ nodes is also strongly causal, demonstrating that the number of
+edges of such a graph can still scale quadratically with the number of
+nodes.  It is evident that the strong causal property is inherited by
+subgraphs.
 
@@
     \Input{Pairwise Granger-causality relations between a persistent
-      processes of dimension $n$ whose joint Granger-causality
+      processe of dimension $n$ whose joint Granger-causality
       relations are known to form a strongly causal DAG $\gcg$.}
@@
 
-\subsection{Bayesian Posterior Thresholding}
-Consider the Bayesian model
-
-\begin{equation}
-  \begin{aligned}
-    (x(t)\ |\ B, \sigma_v^2, \{x(t - \tau)\}_{\tau = 1}^p) &\sim \mathcal{N}(\sum_{\tau = 1}^pB(\tau)x(t - \tau), \sigma_v^2)\\
-    (B_{ij}(\tau)\ |\ G_{ij}(\tau)) &\sim G_{ij}(\tau)\mathcal{N}(0, \sigma_\beta^2) + (1 - G_{ij}(\tau))\delta_0(B_{ij}(\tau))\\
-    (G_{ij}(\tau)) &\sim \mathsf{BER}(p_{ij}(\tau))\\
-    \sigma_v^2 &\sim \Gamma(a_v, b_v)\\
-    \sigma_\beta^2 &\sim \Gamma(a_\beta, b_\beta)
-  \end{aligned}
-\end{equation}
-
-Similar models have been studied by various authors in the context of the linear regression model \hl{[(cite them)]} where it is referred to as \textit{stochastic variable selection}, or referred to as a model for Bayesian variable selection.
-
-Since we have in mind applications where $n$ is large, sampling posterior probabilities can be prohibitively burdensome, so we can instead fit the mean-field variational approximation $p(B, G, \sigma\ |\ X) \approx q_Gq_Bq_\sigma$ and subsequently apply a thresholding operation to the posterior edge inclusion probabilities under $q_G$.
-
-\subsection{Pairwise Minimum Error Spanning Trees}
-Inspired by theorem \ref{thm:tree_recovery}, we propose the following
-structure learning heuristic.
-
-%% This algorithm should perform coordinate descent or jointly refit all the trees.
-
-\begin{enumerate}
-  \item{Set $k = 0$ and $x^0(t) = x(t)$}
-  \item{Compute all pairwise causality measures $\Xi \defeq \big[\ln \frac{\xi_i}{\xi_{ij}} \big]_{i, j}$}
-  \item{Find the maximum spanning arborescence\footnote{An ``arborescence'' is a french word for a tree diagram, and refers to a \textit{directed} tree in graph theory} $\mathcal{T}_k$ of a graph having edge weights $\Xi_{ij}$.}
-  \item{Fit a vector LTI filter $F_k$ having graph defined by $\mathcal{T}_k$ to $x^{k - 1}(t)$ and set $x^k(t) = x^{k - 1}(t) - \hat{x}^{k - 1}(t)$}
-  \item{Repeat over $k$ until a stopping criteria is met}
-\end{enumerate}
+% \subsection{Bayesian Posterior Thresholding}
+% Consider the Bayesian model
+
+% \begin{equation}
+%   \begin{aligned}
+%     (x(t)\ |\ B, \sigma_v^2, \{x(t - \tau)\}_{\tau = 1}^p) &\sim \mathcal{N}(\sum_{\tau = 1}^pB(\tau)x(t - \tau), \sigma_v^2)\\
+%     (B_{ij}(\tau)\ |\ G_{ij}(\tau)) &\sim G_{ij}(\tau)\mathcal{N}(0, \sigma_\beta^2) + (1 - G_{ij}(\tau))\delta_0(B_{ij}(\tau))\\
+%     (G_{ij}(\tau)) &\sim \mathsf{BER}(p_{ij}(\tau))\\
+%     \sigma_v^2 &\sim \Gamma(a_v, b_v)\\
+%     \sigma_\beta^2 &\sim \Gamma(a_\beta, b_\beta)
+%   \end{aligned}
+% \end{equation}
+
+% Similar models have been studied by various authors in the context of the linear regression model \hl{[(cite them)]} where it is referred to as \textit{stochastic variable selection}, or referred to as a model for Bayesian variable selection.
+
+% Since we have in mind applications where $n$ is large, sampling posterior probabilities can be prohibitively burdensome, so we can instead fit the mean-field variational approximation $p(B, G, \sigma\ |\ X) \approx q_Gq_Bq_\sigma$ and subsequently apply a thresholding operation to the posterior edge inclusion probabilities under $q_G$.
+
+% \subsection{Pairwise Minimum Error Spanning Trees}
+% Inspired by theorem \ref{thm:tree_recovery}, we propose the following
+% structure learning heuristic.
+
+% %% This algorithm should perform coordinate descent or jointly refit all the trees.
+
+% \begin{enumerate}
+%   \item{Set $k = 0$ and $x^0(t) = x(t)$}
+%   \item{Compute all pairwise causality measures $\Xi \defeq \big[\ln \frac{\xi_i}{\xi_{ij}} \big]_{i, j}$}
+%   \item{Find the maximum spanning arborescence\footnote{An ``arborescence'' is a french word for a tree diagram, and refers to a \textit{directed} tree in graph theory} $\mathcal{T}_k$ of a graph having edge weights $\Xi_{ij}$.}
+%   \item{Fit a vector LTI filter $F_k$ having graph defined by $\mathcal{T}_k$ to $x^{k - 1}(t)$ and set $x^k(t) = x^{k - 1}(t) - \hat{x}^{k - 1}(t)$}
+%   \item{Repeat over $k$ until a stopping criteria is met}
+% \end{enumerate}
+
+\subsection{Layered Strongly Causal Graphs}
+It is clearn that Algorithm \ref{alg:pwgr} is a purely theoretical construct, and cannot be applied directly to real data.
+Instead, we convert the algorithm into an heuristic and observe it's performance experimentally.
+We first consider the heuristic version of Algorithm \ref{alg:pwgr} in Algorithm \ref{alg:pwgr_heuristic} and show that is a well defined procedure:
+
+\begin{proposition}
+  Algorithm \ref{alg:pwgr_heuristic} terminates in at most $n$ iterations and returns a strongly causal graph $\widehat{\gcg}$.
+\end{proposition}
+\begin{proof}
+  \hl{TODO}
+\end{proof}
+
+A complete estimation procedure for a sparse filter $\widehat{\B}(z)$
+used to estimate $x(t + 1)$ from $\H_{t - 1}$ is then given in Algorithm \ref{alg:Bz_estimate}.
+
+\begin{algorithm}
+  \SetKwInOut{Input}{input}
+  \SetKwInOut{Output}{output}
+  \SetKwInOut{Initialize}{initialize}
+  \DontPrintSemicolon
+
+  \BlankLine
+  \caption{Graph Recovery Heuristic $\mathsf{pw\_scg}(F, \delta, b, R)$}
+  \label{alg:pwgr_heuristic}
+  % \TitleOfAlgo{Pairwise Graph Recovery}
+  \Input{Estimates of pairwise Granger-causaltiy statistics $F_{ij} = \ln \frac{\xi_{ij}}{\xi_i}$\\
+    Sparsity parameters ($\delta, b, R$)}
+  \Output{A strongly causal graph $\widehat{\gcg}$}
+  \Initialize{$S = [n]$  \texttt{\# unprocessed nodes}\\
+    $E = \emptyset$  \texttt{\# edges of }$\widehat{\gcg}$\\
+    $k = 1$ \texttt{\# a counter used only for notation}}
+
+  \BlankLine
+
+  $W_\delta \leftarrow \{(i, j)\ |\ F_{ji} \ge \delta, F_{ji} > F_{ij}\}$  \texttt{\# candidate edges}\\
+  $\mathcal{I}_0 \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# total incident strength for each node}\\
+  $P_0 \leftarrow \mathsf{arg\_select\_min\_N}(\mathcal{I}_0, N=R)$  \texttt{\# select} $R$ \texttt{root nodes}\\
+  \BlankLine
+
+  \While{$S \ne \emptyset$}{
+    $S \leftarrow S \setminus P_{k - 1}$ \texttt{\# remove processed nodes}\\
+    $\mathcal{I}_k \leftarrow \big(\sum_{j \in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# intra-}$S$ \texttt{incident strength}\\
+    $P_k \leftarrow \mathsf{arg\_select\_min\_N}(\mathcal{I}_k, N=b)$  \texttt{\# select} $b$ \texttt{branching nodes}\\
+    \;
+    \texttt{\# add strongest} $P_{k - r} \rightarrow P_k$ \texttt{edges and maintain strong causality}\\
+    \For{$r = 1, \ldots, k$} 
+    {
+      \For{$(i, j) \in \mathsf{sort}\Big(\{(i, j) \in P_{k - r} \times P_k\ |\ (i, j) \in W_\delta\} \mathsf{\ by\ descending\ } F_{ji}\Big)$}{
+        \If{\textit{no }$\gcgpath{i}{j}$\textit{ path in }$E$}{
+          $E \leftarrow E \cup \{(i, j)\}$
+        }
+      }
+    }
+    $k \leftarrow k + 1$
+  }
+  \Return{$([n], E)$}
+\end{algorithm}
+
+\begin{algorithm}
+  \SetKwInOut{Input}{input}
+  \SetKwInOut{Output}{output}
+  \SetKwInOut{Initialize}{initialize}
+  \DontPrintSemicolon
+
+  \BlankLine
+  \caption{Filter Estimator}
+  \label{alg:Bz_estimate}
+  \Input{Wide sense stationary process $x(t)$\\
+    Maximum iterations $M$\\
+    Sparsity parameters $\{(\delta_m\, b_m, R_m)\}_{m = 1}^M$}
+  \Output{Linear one-step-ahead estimator $\widehat{\B}(z)$}
+  \Initialize{$\epsilon^{(0)}(t) = x(t)$ \texttt{\# initial error process}\\
+    $m = 1$ \texttt{\# iteration counter}}
+  \BlankLine
+  \For{$m = 1, \ldots, M$}{
+    $F \leftarrow \mathsf{pwgc}(\epsilon^{(m - 1)}(t))$ \texttt{\# estimate pairwise Granger-causality statistics}\\
+    $\widehat{\gcg}_m \leftarrow \mathsf{pw\_scg}(F, \delta_m, b_m, R_m)$ \texttt{\# graph estimate (Algorithm \ref{alg:pwgr_heuristic})}\\
+    $\widehat{\B}^{(m)}(z) \leftarrow \mathsf{estimate\_B}(\epsilon^{(m - 1)}(t), \widehat{\gcg}_m)$ \texttt{\# Estimate filter with topology} $\widehat{\gcg}_m$\\
+    $\epsilon^{(m)} \leftarrow \epsilon^{(m - 1)}(t) - \widehat{\B}^{(m)}(z)\epsilon^{(m - 1)}(t)$ \texttt{\# updated error process}\\
+  }
+  \BlankLine
+  $\widehat{\B}(z) \leftarrow \widehat{\B}^{(M)}(z) \prod_{k = 1}^{M - 1}\big(I - \widehat{\B}^{(M - k)}(z)\big)$ \texttt{\# compute total filter}\\
+  \Return{$\widehat{\B}(z)$}
+\end{algorithm}
 

### 2019-05-19 software/gc_methods.py

@@ -1,0 +1,72 @@
+import networkx as nx
+import numpy as np
+
+from sklearn.linear_model import LinearRegression, LassoLarsIC
+
+
+def estimate_b_lstsqr(y, X):
+    """
+    Simple direct estimate of a linear filter y_hat = Xb.
+    """
+    lr = LinearRegression(fit_intercept=False, normalize=False,
+                          copy_X=True)
+    lr.fit(X, y)
+    return lr.coef_
+
+
+def estimate_b_lasso(y, X, criterion="bic"):
+    """
+    Estimate y_hat = Xb via LASSO and using BIC to choose regularizers.
+    """
+    lasso = LassoLarsIC(criterion=criterion, fit_intercept=False,
+                        normalize=False)
+    lr = LinearRegression(fit_intercept=False, normalize=False,
+                          copy_X=False)
+
+    lasso.fit(X, y)
+    w = lasso.coef_
+
+    # Use lasso only to select the support
+    X_lasso = X[:, [i for i, wi in enumerate(w) if wi != 0]]
+    w_lr = lr.fit(X_lasso, y).coef_
+    w[[i for i, wi in enumerate(w) if wi != 0]] = w_lr
+    return w
+
+
+def estimate_B(X, G, max_lag=10):
+    """
+    Estimate x(t) = B(z)x(t) respecting the topology of G.
+    """
+    
+    return
+
+
+def example_graph():
+    N = list(range(1, 7))
+    E = [(1, 3), (2, 4), (3, 4), (3, 5), (4, 6)]
+
+    G = nx.DiGraph()
+    G.add_nodes_from(N)
+    G.add_edges_from(E)  # This will add nodes if they aren't already there.
+
+    # set(G.predecessors(i)): set of parents of i in G
+    # set(nx.ancestors(G, i)): set of ancestors of i in G
+    return
+
+
+
+def example():
+
+    def sample_x():
+        T = 2000
+        th = 2.25 * np.pi
+        a = 2 * np.cos(th)
+        e = np.random.normal(size=T)
+        x = np.zeros(T + 2)
+        for t in range(T):
+            x[t] = a * x[t - 1] - x[t - 2] + e[t]
+        return x
+
+    [plt.plot(sample_x(), linewidth=.75, alpha=0.8) for _ in range(20)]
+    plt.show()
+    return

### 2019-05-21 paper/paper.tex
Added some citations and updated intro of paper.

@@
 understanding of the notion of Granger-Causality
-\cite{granger1969investigating} as a tool for answering these questions.
-
-Though the notion of causality is a philosophically slippery concept,
-it is fundamental to the way we understand the world and to the
-progress of science in general.  Indeed, without faith in the
-consistency of causal interactions the results of experimental science
-could not be generalized or applied in any meaningful way.  In the
-case of Granger-Causality, we state that if an event '$A$' provides us
-with unique (that is, not available anywhere else) information about a
-later event $B$, then $A$ must have a causal impact on $B$.  As
-opposed to the notion of Causation promoted by Pearl
-\cite{pearl2000art}, this is an entirely model-free notion of cause,
-and instead leverages the intuition that a cause must precede it's
-effect.
+\cite{granger1969investigating} \cite{Granger1980329} as a tool for
+answering these questions.
+
+Heuristically, we state that if an event '$A$' occuring strictly prior
+in time to an even '$B$' and $A$ provides us with exclusive (that is,
+not available anywhere else) information about $B$, then $A$ must have
+had a causal impact on $B$.  This is an entirely model-free notion of
+cause, and instead leverages the intuition that a cause must precede
+it's effect.
 
@@
 for \textit{true} causation, since our statements about causation are
-highly dependent upon the data that we are able to observe.  We prefer
-instead to interpret Granger causality as a means of uncovering a flow
-of ``information'' or ``energy'' through some underlying graph of
+highly dependent upon the data that we are able to observe.  Rather,
+the interpretation of Granger causality is as a means of uncovering a
+flow of ``information'' or ``energy'' through some underlying graph of
 interactions.  Though this graph cannot be observed directly, we will
 infer it's presence as a latent structure among our observed time
-series data.
+series data.  This intuition is levereged in a variety of applications
+e.g. in Neuroscience as a means of recovering interactions amongst
+brain regions \cite{bressler2011wiener}, \cite{anna_paper2008},
+\cite{david2008identifying}; in the study of the dependence and
+connectedness of financial institutions \cite{NBERw16223}; gene
+expression networks \cite{Fujita2007}; and power system design
+\cite{Misyrlis2016450}, \cite{yuan2014root}.
 
 Finding the ``best'' graph structure consistent with observed data is
-generally an extremely challenging problem, though the comparison of
-quality between different structures, and hence the notion of
-``best'', needs to be quantified.  In applications where we are
-interested merely in minimizing the mean squared error of a linear
+generally an extremely challenging problem (i.e. this is a best subset
+selection problem, see \cite{bss_mio}, \cite{hastie_bss_comp}), though
+the comparison of quality between different structures, and hence the
+notion of ``best'', needs to be quantified.  In applications where we
+are interested merely in minimizing the mean squared error of a linear
 one-step-ahead predictor, then we will naturally desire an entirely
@@
 In \cite{bach2004learning} the authors apply a local search heuristic
-to an AIC penalized approximation of the likelihood where at each
-iteration an edge is either added, removed, or reversed.  This is a
-common approach to combinatorial optimization due to it's simplicity,
-but is liable to get stuck in shallow local minima.
+to the Whittle likelihood with an AIC penalization.  The local search
+heuristic where at each iteration an edge is either added, removed, or
+reversed is a common approach to combinatorial optimization due to
+it's simplicity, but is liable to get stuck in shallow local minima.
 
@@
 
-From a Bayesian perspective, spike-and-slab priors can be used to
-estimate the probability of an edge's inclusion, after which a graph
-can be chosen heuristically by thresholding or searching for the MAP
-graph structure.
-
-Our contribution is to provide an alternative heuristic based on
-minimum spanning trees, inspired by our result in section
-\ref{sec:theory} that the true causality graph can be recovered from
-pairwise information alone in the case that the underlying graph is a
-directed forest.  We compare this algorithm to a grouped LASSO
-\cite{yuan2006model}, local search, and a Bayesian posterior
-probability thresholding scheme.
+The contributions of this paper are as follows: firstly, in section
+\ref{sec:theory} we study \textit{pairwise} Granger-causality
+relations, providing novel theorems connecting the structure of the
+causality graph to the pairwise ``causality flow'' in the system, as
+well as a characterization of the sparsity pattern of matrices arising
+in the Wold decomposition.  We establish sufficient conditions
+(sections \ref{sec:strongly_causal_graphs},
+\ref{sec:persistent_systems}) under which a fully conditional
+Granger-causality graph can be recovered from pairwise tests alone
+(section \ref{sec:pairwise_algorithm}).  Secondly, we propose in
+section \ref{sec:structure_learning} alternative graph search
+heuristics inspired by our theoretical results.  Our heuristics are
+compared against a standard grouped LASSO algorithm
+\cite{yuan2006model} as well as local search hill climbing (\hl{state
+  here the results}).  In section \ref{sec:application} we develop an
+example application (\hl{more detail...}).  And provide concluding
+remarks on further open problems in our conclusion (section
+\ref{sec:conclusion}).
 
@@
 $(\Omega, \mathcal{F}, \mathbb{P})$ having inner product
-$\inner{x}{y} = \E[xy]$.  We will denote the collection vectors of
-such random elements as $L_2^n(\Omega)$ or simply $L_2^n$.
-
-We will work with a discrete time and wide-sense stationary process
-$x(t)$ taking elements in $L_2$.  We suppose that $x(t)$ (with
-$t \in \Z$) has zero mean, $\E x(t) = 0$, and the matrix valued
-covariance sequence
-$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ is absolutely
-summable and has a spectra $S(\omega)$:
+$\inner{x}{y} = \E[xy]$.  We will work with a discrete time and
+wide-sense stationary (WSS) vector valued process $x(t)$ (with
+$t \in \Z$) with elements taking values in $L_2$.  We suppose that
+$x(t)$ has zero mean, $\E x(t) = 0$, and has absolutely summable
+matrix valued covariance sequence
+$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ with an absolutely
+continuous spectral density $S(\omega)$:
 
@@
 \end{equation}
+
+% and that the spectra is bounded uniformly away from 0:
+
+% \begin{equation*}
+%   \int_{-\pi}^\pi \ln\det S(\omega) \d\omega > -\infty.
+% \end{equation*}
 
@@
 
-where the closure is naturally in mean-square.  $\H_t^x$ is the space
-spanned by the values of $x(t)$ between the ``infinite past'' and
-``now'', we will often omit the superscript $x$ which should be clear
-from context.  Evidently these spaces are separable, and as closed
-subspaces of a Hilbert space they are themselves Hilbert.  We will
-denote the spaces generated in analagous ways by particular components
-of $x$ as e.g. $\H_t^{(i, j)}$, $\H_t^{i}$ or by all but particular
-components as e.g. $\H_t^{-j}$.
+where the closure is naturally in mean-square.  We will often omit the
+superscript $x$ which should be clear from context.  Evidently these
+spaces are separable, and as closed subspaces of a Hilbert space they
+are themselves Hilbert.  We will denote the spaces generated in
+analagous ways by particular components of $x$ as e.g.
+$\H_t^{(i, j)}$, $\H_t^{i}$ or by all but particular components as
+e.g. $\H_t^{-j}$.
 
 As a consequence of the Wold decomposition theorem \cite{lindquist},
-every such sequence has the ``moving average'' $MA(\infty)$
+every WSS sequence has the moving average $MA(\infty)$
 representation
@@
 \label{eqn:wold}
-  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) w(t - \tau),
+  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) v(t - \tau),
 \end{equation}
 
-where $c(t)$ is a perfectly predictable sequence\footnote{$c(t)$ is
-  perfectly predictable if any sample $c(t_0)$ is enough to determine
-  $c(t)$ for every $t$.  For example,
+where $c(t)$ is a purely deterministic sequence\footnote{the purely
+  deterministic sequence $c(t)$ is one which lies in the remote past
+  $\bigcap_{\tau=1}^\infty \H_{t - \tau}^x$ of the process.  For such
+  processes a single sample $c(t_0)$ is enough to determine $c(t)$ for
+  every $t$.  For example,
   $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi,
-  \pi]$}, and $w(t)$ is an uncorrelated sequence having a diagonal
-covariance and $A(0) = I$.  We will assume that $c(t) = 0$, which in
-practice is to say that it has been accurately estimated and removed.
-
-Furthermore, we assume that the spectra is uniformly bounded away from
-$0$:
-$\exists \delta:\ \delta I \preceq S(\omega) \preceq \frac{1}{\delta}
-I\ \forall \omega$ so that the Wold representation can be inverted to
-yield the $VAR(\infty)$ form
+  \pi]$}, and $w(t)$ is an uncorrelated sequence and $A(0) = I$.  We
+will assume that $c(t) = 0$, which in practice is to say that it has
+been accurately estimated and removed.  We additionally require that
+this representation can be inverted to yield the $VAR(\infty)$ form
 
@@
   \label{eqn:ar_representation}
-  x(t) = \sum_{\tau = 1}^\infty B(\tau) x(t - \tau) + v(t),
+  x(t) = \sum_{\tau = 1}^\infty B(\tau) x(t - \tau) + v(t).
 \end{equation}
 
-where $v(t)$ is an uncorrelated sequence with covariance $\Sigma_v$.
-
 The equations (\ref{eqn:wold}), (\ref{eqn:ar_representation}) can be
-represented as $x(t) = \A(z)w(t) = \B(z)x(t) + v(t)$ via the action of
-the operators (LTI filters)
+represented as $x(t) = \A(z)v(t) = \B(z)x(t) + v(t)$ via the action
+(convolution) of the operators (LTI filters)
 $\A(z) \defeq \sum_{\tau = 0}^\infty A(\tau)z^{-\tau}$ and
 $\B(z) \defeq \sum_{\tau = 1}^\infty B(\tau)z^{-\tau}$ where the
-operator $z^{-1}$ is the backshift operator acting on $\H$.  Since
-$||z^{-1}|| = 1$, the inversion formulae
+operator $z^{-1}$ is the backshift operator acting on
+$\ell_2^n(\Omega, \mathcal{F}, \mathbb{P})$, that is:
+
+\begin{equation}
+  \label{eqn:filter_action}
+  \B_{ij}(z)x_j(t) \defeq \sum_{\tau = 1}^\infty B_{ij}(\tau)x_j(t - \tau).
+\end{equation}
+
+Finally, since $||z^{-1}|| = 1$ we have the inversion formula
 
@@
   \label{eqn:lsi_inversion}
-  \A(z) = (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k
+  \A(z) = (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k.
 \end{equation}
 
-are valid.
+The aforementioned assumptions are quite weak.  The strongest
+assumption we require is finally that $\Sigma_v$ is a diagonal matrix,
+which is referred to as a lack of instantaneous feedback in $x(t)$.
+We formally state our setup to conclude this section:
+
+\begin{definition}[Basic Setup]
+  The process $x(t)$ is an $n$ dimensional wide sense stationary
+  process having invertible $VAR(\infty)$ representation
+  \eqref{eqn:ar_representation} where $v(t)$ is sequentially
+  uncorrelated and has a diagonal covariance matrix.  The $MA(\infty)$
+  representation of equation \eqref{eqn:wold} has $c(t) = 0$ and
+  $A(0) = I$.
+\end{definition}
+
 \subsection{Granger Causality}
@@
 The notion is closely related to the information theoretic measure of
-transfer entropy, indeed, if everything is Gaussian then they are
-equivalent \cite{barnett2009granger}.
+transfer entropy, indeed, if the distribution of $v(t)$ is known to be
+Gaussian then they are equivalent \cite{barnett2009granger}.
 
@@
 \begin{definition}[Causality graph, Parents]
-
-Remaining in the context of a collection of $n$ WSS processes
-$x(t) = \big(x_1(t), \ldots, x_n(t)\big)$, we define the Granger-causality
-graph $\gcg$ to be the directed graph formed on $n$ vertices where an
-edge $(j, i) \in \gcg$ if and only if $x_j \gc x_i$.
-
-We refer to the set of \textit{parents} of $x_i$ in $\gcg$ by
-$\pa{i}$, that is, $j \in \pa{i} \iff x_j \gc x_i$.
+  We define the Granger-causality graph $\gcg = ([n], \gcge)$ to be the directed
+  graph formed on $n$ vertices where an edge $(j, i) \in \gcge$ if and
+  only if $x_j$ Granger-causes $x_i$ (with respect to $x$).  That is,
+  $$(j, i) \in \gcge \iff j \in \pa{i} \iff x_j \gc x_i.$$
 \end{definition}
@@
 The edges of the Granger-causality graph $\gcg$ can be given a general
-notion of ``weight'' by associating an edge $(j, i)$ with a
-\textit{strictly causal} LTI filter
-
-\begin{equation}
-  \label{eqn:filter}
-  \B_{ij}(z) = \sum_{\tau = 1}^{\infty} B_{ij}(\tau)z^{-\tau}
-\end{equation}
-
-from the $\VAR$ representation of equation
-\ref{eqn:ar_representation}.  We will represent the action
-(convolution) of this filter compactly by taking $z^{-1}$ as the
-backshift operator on $\ell_2^n(\Omega, \mathcal{F}, \mathbb{P})$ when
-convenient
-
-\begin{equation}
-  \B_{ij}(z)x_j(t) \defeq \sum_{\tau = 1}^\infty B_{ij}(\tau)x_j(t - \tau).
-\end{equation}
+notion of ``weight'' by associating an edge $(j, i)$ with the
+\textit{strictly causal} LTI filter $\B_{ij}(z)$ (see eqn
+\eqref{eqn:filter_action}).  Thence, the matrix $\B(z)$ is analagous
+to a \textit{weighted adjacency matrix}\footnote{We are using the
+  convention that $\B_{ij}(z)$ is a filter with input $x_j$ and output
+  $x_i$ so as to write the action of the system as $\B(z)x(t)$ with
+  $x(t)$ as a column vector.  This competes with the usual convention
+  for adjacency matrices where $A_{ij} = 1$ if there is an edge
+  $(i, j)$.  In our case, the sparsity pattern of $\B_{ij}$ is the
+  \textit{transposed} adjacency matrix.} for the graph $\gcg$.  And,
+in the same way that the $k^{\text{th}}$ power of an adjacency matrix
+counts the number of paths of length $k$ between nodes,
+$(\B(z)^k)_{ij}$ is a filter isolating the ``action'' of $j$ on $i$ at
+a time lag of $k$ steps, this is exemplified in the inversion formula
+\ref{eqn:lsi_inversion}.
+
+An elementary theorem connecting the adjacency matrix with paths will
+allow us to deduce the sparsity pattern of $\A(z)$.  We omit proof,
+though it follows easily by induction:
+
+\begin{lemma}
+  \label{lem:adj_matrix}
+  Let $S$ be the transposed adjacency matrix of the Granger-causality
+  graph $\gcg$.  Then, $(S^k)_{ij}$ is the number of paths of length
+  $k$ from node $j$ to node $i$.  Evidently, if
+  $\forall k \in \N,\ (S^k)_{ij} = 0$ then $j \not\in \anc{i}$.
+\end{lemma}
 
@@
 From the $\VAR$ representation of $x(t)$ there is clearly a tight
-relationship between each node and it's parent nodes, indeed
-
-\begin{equation}
-  \label{eqn:parent_expansion}
-  x_i(t) = v_i(t) + \B_{ii}(z)x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t).
-\end{equation}
-
-% When is this guaranteed w.r.t. the full system?
-Furthermore, if the filter $\B_{ii}$ is invertible, we can write
-
-\begin{equation*}
-  \begin{aligned}
-    (1 - \B_{ii}(z))x_i(t) &= v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\\
-    \Rightarrow x_i(t) &= (1 - \B_{ii}(z))^{-1}\big[v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\big]\\
-    &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)
-  \end{aligned}
-\end{equation*}
+relationship between each node and it's parent nodes, the following
+proposition is analagous to the definition of feedback free processes
+of \cite{caines1975feedback}:
+
+% % When is this guaranteed w.r.t. the full system?
+% Furthermore, if the filter $\B_{ii}$ is invertible, we can write
+
+% \begin{equation*}
+%   \begin{aligned}
+%     (1 - \B_{ii}(z))x_i(t) &= v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\\
+%     \Rightarrow x_i(t) &= (1 - \B_{ii}(z))^{-1}\big[v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\big]\\
+%     &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)
+%   \end{aligned}
+% \end{equation*}
 
@@
 
-We continue to develop the relationship between $x_i$ and it's family
-tree, where the following proposition is analagous to the definitions
-of feedback free processes of \cite{caines1975feedback}.
+% We continue to develop the relationship between $x_i$ and it's family
+% tree, where the following proposition is analagous to the definitions
+% of feedback free processes of \cite{caines1975feedback}.
 
@@
   \label{prop:parent_expanding}
-  The component $x_i(t)$ of $x(t)$ can be represented as a sum of
-  linear filtering operations acting on the driving noise $v_i$ of
-  $x_i$ as well as the parents of $x_i$:
+  The component $x_i(t)$ of $x(t)$ can be represented in terms of it's
+  parents in $\gcg$:
 
@@
     \label{eqn:parent_expansion}
-    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t),
+    x_i(t) = v_i(t) + \B_{ii}(z)x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t).
   \end{equation}
 
-  where $\wtB_{ii}(z) = (1 - \B_{ii}(z))^{-1}$ and
-  $\wtB_{ik}(z) = \wtB_{ii}(z) B_{ik}(z)$.
-
-  Moreover, $x_i$ can be expanded in terms of it's ancestors and
-  driving noise only:
+  Moreover, $x_i$ can be expanded in terms of it's ancestor's $v(t)$
+  components only:
 
@@
 
-The matrix $\B(z)$ is analagous to a \textit{weighted adjacency
-  matrix} for the graph $\gcg$.  And, in the same way that the
-$k^{\text{th}}$ power of an adjacency matrix counts the number of
-paths of length $k$ between nodes, $(\B(z)^k)_{ij}$ is a filter
-isolating the ``action'' of $j$ on $i$ at a time lag of $k$ steps.
-Though we only need the sparsity pattern of $\B(z)^k$ for the proof.
-
-We state the following well known theorem as a lemma, proof follows
-easily by induction.
-
-\begin{lemma}
-  \label{lem:adj_matrix}
-  Let $S$ be the transposed adjacency matrix of the Granger-causality
-  graph $\gcg$.  Then, $(S^k)_{ij}$ is the number of paths of length
-  $k$ from node $j$ to node $i$.  Evidently, if
-  $\forall k \in \N,\ (S^k)_{ij} = 0$ then $j \not\in \anc{i}$.
-\end{lemma}
-
-We now prove the proposition.
-
 \begin{proof}
-  Equation \eqref{eqn:parent_expansion} was derived prior to the
-  statement of the theorem.  We are left to demonstrate equation
-  \eqref{eqn:ancestor_expansion}.
+  Equation \eqref{eqn:parent_expansion} is immediate from the
+  $VAR(\infty)$ representation of \eqref{eqn:ar_representation} and
+  Theorem \ref{thm:granger_causality_equivalences}, we are left to
+  demonstrate \eqref{eqn:ancestor_expansion}.
 
@@
 example, $x_k$ may be a common ancestor which when observed completely
-explains the connection from $x_j$ to $x_i$.  In particular, we will
-study \textit{pairwise} Granger-causality.
+explains the connection from $x_j$ to $x_i$.  In this paper we study
+\textit{pairwise} Granger-causality, and seek to understand when
+knowledge of pairwise relations is sufficient to deduce the true fully
+conditional relations of $\gcg$.
 
@@
 This notion is of interest for a variety of reasons.  From a purely
-theoretical or conceptual standpoint, we will seek to understand when
-and how pairwise causality can capture the idea of ``flow of
-information'' in the underlying graph.  It may also be useful for
-reasoning about what the conditions under which \textit{unobserved}
-components of $x(t)$ may or may not interfere with inference in the
-actually observed components.  Finally, motivated from a practical
-standpoint to analyze causation in large systems, we will seek to
-construct practical estimation procedures based purely on pairwise
-causality tests which can result in reduced computational complexity.
-
-\begin{proposition}
-  \label{prop:ancestor_uncorrelated}
+conceptual standpoint, we will see how the notion can in some sense
+capture the idea of ``flow of information'' in the underlying graph.
+It may also be useful for reasoning about what the conditions under
+which \textit{unobserved} components of $x(t)$ may or may not
+interfere with inference in the actually observed components.
+Finally, motivated from a practical standpoint to analyze causation in
+large systems, we will seek to construct practical estimation
+procedures based purely on pairwise causality tests since the
+computation of such pairwise relations is somewhat easier.
+
+The following is a fundamental building block:
+
+\begin{lemma}
+  \label{lem:ancestor_uncorrelated}
   Consider distinct nodes $i, j$ in a Granger-causality graph
@@
   this implies that $j \npwgc i$.
-\end{proposition}
+\end{lemma}
 
@@
   $i \in \anc{j}$ is not excluded.
-
-  We also point out the contrapositive that if $j \pwgc i$ then we
-  must have either $j \in \anc{i}$ or $\anc{i} \cap \anc{j} \ne 0$.
 \end{remark}
@@
 
-  which follows from Proposition \ref{prop:ancestor_uncorrelated}.
+  which follows from Lemma \ref{lem:ancestor_uncorrelated}.
   Finally, for $k \in C_1(j)$ the case $k = i$ is immediate (since the
@@
 matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead,
-we think it is more meaningful to focus on the defining feature of
+we think it more meaningful to focus on the defining feature of
 time series networks, that is, the topology of $\gcg$.
@@
   from a node in $P_0$ to $j$, where $d(j) = 0$ if $j \in P_0$.  It is
-  apparent (\hl{right?}) that such a path will always exist in a
-  strongly causal graph.  For example, in Figure
+  apparent that such a path will always exist.  For example, in Figure
   \ref{fig:example_fig3} we have $d(3) = 1$ and $d(4) = 2$.
@@
 
+\begin{proposition}
+  The error sequence in Algorithm \ref{alg:Bz_estimate} is a zero mean
+  WSS process $\E\epsilon^{(m)}(t) = 0$ with monotonically decreasing
+  variance: $\E|\epsilon^{(m + 1)}(t)|^2 \le \E|\epsilon^{(m)}(t)|^2$.
+\end{proposition}
+\begin{proof}
+  \hl{TODO -- this is probably not even true}
+\end{proof}
+
 \begin{algorithm}
@@
   \label{alg:Bz_estimate}
-  \Input{Wide sense stationary process $x(t)$\\
+  \Input{$T$ samples from a WSS process $x: \{x(t)\}_{t = 1}^T$\\
     Maximum iterations $M$\\
@@
     $\widehat{\B}^{(m)}(z) \leftarrow \mathsf{estimate\_B}(\epsilon^{(m - 1)}(t), \widehat{\gcg}_m)$ \texttt{\# Estimate filter with topology} $\widehat{\gcg}_m$\\
-    $\epsilon^{(m)} \leftarrow \epsilon^{(m - 1)}(t) - \widehat{\B}^{(m)}(z)\epsilon^{(m - 1)}(t)$ \texttt{\# updated error process}\\
+    $\epsilon^{(m)}(t) \leftarrow \epsilon^{(m - 1)}(t) - \widehat{\B}^{(m)}(z)\epsilon^{(m - 1)}(t)$ \texttt{\# updated error process}\\
   }
@@
 \section{Application}
+\label{sec:application}
 \section{Conclusion}
+\label{sec:conclusion}
+Extensions to transfer entropy?.  Case of time varying statistics?.  Develop sparse estimators for SS models (\cite{barnett2015granger})?.  Apply MIO for the Whittle relaxation?
 

### 2019-05-21 software/gc_methods.py

@@ -4,19 +4,25 @@ import numpy as np
 from sklearn.linear_model import LinearRegression, LassoLarsIC
 
 
-def estimate_b_lstsqr(y, X):
+def estimate_b_lstsqr(X, y):
     """
     Simple direct estimate of a linear filter y_hat = Xb.
+
+    y: np.array (T)
+    X: np.array (T x s)
     """
     lr = LinearRegression(fit_intercept=False, normalize=False,
-                          copy_X=True)
+                          copy_X=False)
     lr.fit(X, y)
     return lr.coef_
 
 
-def estimate_b_lasso(y, X, criterion="bic"):
+def estimate_b_lasso(X, y, criterion="bic"):
     """
     Estimate y_hat = Xb via LASSO and using BIC to choose regularizers.
+
+    y: np.array (T)
+    X: np.array (T x s)
     """
     lasso = LassoLarsIC(criterion=criterion, fit_intercept=False,
                         normalize=False)
@@ -25,6 +31,9 @@ def estimate_b_lasso(y, X, criterion="bic"):
 
     lasso.fit(X, y)
     w = lasso.coef_
+    if np.all(w == 0):
+        # All zero support
+        return w
 
     # Use lasso only to select the support
     X_lasso = X[:, [i for i, wi in enumerate(w) if wi != 0]]
@@ -33,11 +42,52 @@ def estimate_b_lasso(y, X, criterion="bic"):
     return w
 
 
+def form_Xy(X, i, a, max_lag=10):
+    """
+    Builds appropriate (X, y) matrices for autoregressive prediction.
+
+    X: np.array (T, n)
+    i: index of X to model
+    a: length n boolean array indicating which variables to include
+      - denote s = sum(a) i.e. "sparsity"
+    max_lag: the number of lagged values to include in AR predictions.
+
+    returns: y_i, _X
+    y_i: np.array (T - max_lag, 1)
+    _X: np.array (T - max_lag, max_lag * s)
+    """
+    T, s = X[:, a].shape
+    p = max_lag
+
+    y = X[p:, i]
+    _X = np.hstack([X[p - tau: -tau, a] for tau in range(1, p + 1)])
+    assert _X.shape == (T - max_lag, s * max_lag)
+    assert len(y) == T - max_lag
+    return _X, y
+
+
 def estimate_B(X, G, max_lag=10):
     """
[…]

### 2019-05-22 software/gc_methods.py
Looking at stock and sentiment data.

@@ -90,7 +90,6 @@ def estimate_B(X, G, max_lag=10):
         b = estimate_b_lasso(X_i, y_i)
     return
 
-
 def example_graph():
     N = list(range(1, 7))
     E = [(1, 3), (2, 4), (3, 4), (3, 5), (4, 6)]

### 2019-05-22 software/stock_data/study_stocks.py

@@ -1,0 +1,332 @@
+import seaborn.apionly as sns
+import statsmodels as sm
+import statsmodels.tsa as tsa
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import os
+import scipy.stats as sps
+import xarray as xr
+import pandas_market_calendars as mcal
+
+from itertools import chain
+from statsmodels.tsa.stattools import (adfuller, acf, pacf,
+                                       grangercausalitytests)
+
+DATA_DIR = "data/"
+
+def get_stock_name(f):
+    return f.split(".")[0]
+
+
+def read_stock_csv(directory, file_name, cols=None):
+    """
+    Intended to read files formatted as "<ticker>.us.txt" with data
+    (Date, Open, High, Low, Volume, OpenInt)
+    
+    """
+    if cols is None:
+        cols = ["Close", "Volume"]
+    if "Date" not in cols:
+        cols = cols + ["Date"]
+
+    name = get_stock_name(file_name)
+    try:
+        # index_col then reset_index is to ensure date types are parsed
+        D = pd.read_csv(directory + file_name, parse_dates=True,
+                        infer_datetime_format=True,
+                        usecols=cols, dtype={"Volume": float},
+                        index_col="Date").reset_index()
+        D["Symbol"] = name
+    except pd.errors.EmptyDataError:
+        return None
+    else:
+        return D
+
+
+def read_sent_csv(file_path):
+    """
+    Reads the sentiment analysis data
+    """
+    sent = pd.read_csv(file_path, index_col=0,
+                       infer_datetime_format=True,
+                       parse_dates=["date"])\
+        .rename({"symbol": "Symbol",
+                 "sentiment_signal": "sent",
+                 "date": "Date"},
+                axis="columns")
+    sent["Symbol"] = sent["Symbol"].str.lower()
+    sent = sent.set_index(["Symbol", "Date"])\
+        .sort_index()
+    return sent
+
+
+def stock_filter(D, min_average_volume=50000,
+                 min_date=None, max_date=None,
+                 inclusive_dates=False):
+    """
+    Filter out stocks on dates and volumes.
+    """
+    def conv_date(d):
+        if isinstance(d, str):
+            return pd.Timestamp(d)
+        else:
+            return d
+
+    if D is None:
+        return None
+
[…]

### 2019-05-29 paper/paper.tex
Some results and more code for simulations.

@@
 \usepackage{etex,etoolbox}
+\usepackage{hyperref}
+\usepackage{todonotes}
 \bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
-\usepackage{hyperref}
 
@@
 
-\begin{proposition}
-  Algorithm \ref{alg:pwgr_heuristic} terminates in at most $n$ iterations and returns a strongly causal graph $\widehat{\gcg}$.
-\end{proposition}
-\begin{proof}
-  \hl{TODO}
-\end{proof}
-
-A complete estimation procedure for a sparse filter $\widehat{\B}(z)$
-used to estimate $x(t + 1)$ from $\H_{t - 1}$ is then given in Algorithm \ref{alg:Bz_estimate}.
-
-\begin{proposition}
-  The error sequence in Algorithm \ref{alg:Bz_estimate} is a zero mean
-  WSS process $\E\epsilon^{(m)}(t) = 0$ with monotonically decreasing
-  variance: $\E|\epsilon^{(m + 1)}(t)|^2 \le \E|\epsilon^{(m)}(t)|^2$.
-\end{proposition}
-\begin{proof}
-  \hl{TODO -- this is probably not even true}
-\end{proof}
+% \begin{proposition}
+%   Algorithm \ref{alg:pwgr_heuristic} terminates in at most $n$ iterations and returns a strongly causal graph $\widehat{\gcg}$.
+% \end{proposition}
+% \begin{proof}
+%   \hl{TODO}
+% \end{proof}
+
+% A complete estimation procedure for a sparse filter $\widehat{\B}(z)$
+% used to estimate $x(t + 1)$ from $\H_{t - 1}$ is then given in Algorithm \ref{alg:Bz_estimate}.
+
+% \begin{proposition}
+%   The error sequence in Algorithm \ref{alg:Bz_estimate} is a zero mean
+%   WSS process $\E\epsilon^{(m)}(t) = 0$ with monotonically decreasing
+%   variance: $\E|\epsilon^{(m + 1)}(t)|^2 \le \E|\epsilon^{(m)}(t)|^2$.
+% \end{proposition}
+% \begin{proof}
+%   \hl{TODO -- this is probably not even true}
+% \end{proof}
 
@@
       \For{$(i, j) \in \mathsf{sort}\Big(\{(i, j) \in P_{k - r} \times P_k\ |\ (i, j) \in W_\delta\} \mathsf{\ by\ descending\ } F_{ji}\Big)$}{
-        \If{\textit{no }$\gcgpath{i}{j}$\textit{ path in }$E$}{
+        \If{no $\gcgpath{u}{j}$ path in $E$, for any $u \in \anc{i}$}{
           $E \leftarrow E \cup \{(i, j)\}$
@@
 
+\section{Empirical Evaluation}
+We have implemented simple empirical experiments in Python
+\cite{scipy}, in particular we leverage the LASSO implementation from
+\texttt{sklearn} \cite{sklearn} and the random graph generators from
+\texttt{networkx} \cite{networkx}.  We run the same experiment on
+three separate graph topologies having $n = 75$ nodes: a tree (which
+is strongly causal), a random directed acyclic graph, and an Erdos
+Renyi random graph with edge inclusion probability of $2 / n$.
+
+We populate each of the edges (including self loops) with random
+linear filters constructed by placing $5$ transfer function poles
+uniformly at random in a disc of radius $3 / 4$ (which guarantees
+stability for acyclic graphs).  The resulting system is driven by
+i.i.d. Gaussian random vectors, each component having random variance
+$\sigma_i^2 = 1/2 + r_i$ where $r_i \sim \text{exp}(1/2)$.  Results
+are obtained for $50$ random iterations per each of $6$ dataset
+lengths $T \in \{100, 500, 1000, 2500, 5000, 10000\}$.
+
+The pairwise causality function $\mathsf{pwgc}$ of Algorithm
+\ref{alg:Bz_estimate} are obtained via the residuals
+$\mathbf{\epsilon}_{ij}$ of the model
+$\mathbf{x}_i = \mathbf{X}_{ij}\widehat{\beta_{}} +
+\mathbf{\epsilon}_{ij}$ estimated via ordinary least squares, where
+$\mathbf{X}_{ij}$ includes $10$ lagged values of processes $i$ and
+$j$.
+
+The function $\mathsf{estimate_B}$ in Algorithm \ref{alg:Bz_estimate}
+is carried out via the LASSO \cite{tibshirani2015statistical}, with
+the regularization parameter being chosen to minimize the Bayesian
+information criteria, a more conservative heuristic than the AIC.
+
+\begin{figure}
+  \centering
+  \caption{Simulation Results and Representitive Random Graph Topologies}
+  \label{fig:simulation_results}
+
+  \begin{subfigure}[b]{0.30\textwidth}
+    \caption{Random Tree}
+    \includegraphics[width=\linewidth]{example_tree.pdf}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.30\textwidth}
+    \caption{Random DAG}
+    \includegraphics[width=\linewidth]{example_dag.pdf}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.30\textwidth}
+    \caption{Erdos-Renyi Graph}
+    \includegraphics[width=\linewidth]{example_gnp.pdf}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.30\textwidth}
+    \includegraphics[width=\linewidth]{random_tree_simulation.pdf}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.30\textwidth}
+    \includegraphics[width=\linewidth]{random_dag_simulation.pdf}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.30\textwidth}
+    % \includegraphics[width=\linewidth]{random_gnp_simulation.pdf}
+    \missingfigure{}
+  \end{subfigure}
+\end{figure}
+
 \section{Application}
@@
 \label{sec:conclusion}
-Extensions to transfer entropy?.  Case of time varying statistics?.  Develop sparse estimators for SS models (\cite{barnett2015granger})?.  Apply MIO for the Whittle relaxation?
+% Extensions to transfer entropy?.  Case of time varying statistics?.  Develop sparse estimators for SS models (\cite{barnett2015granger})?.  Apply MIO for the Whittle relaxation?
 

### 2019-05-29 software/gc_methods.py

@@ -1,12 +1,15 @@
 import networkx as nx
 import numpy as np
-from scipy import linalg
+import sympy as sp
 
+from scipy import linalg
 from sklearn.linear_model import LinearRegression, LassoLarsIC
-from multiprocessing import Pool
 
 from itertools import product
 from math import ceil
+
+from var_system import (attach_node_prop, add_self_loops,
+                        remove_zero_filters)
 
 def estimate_b_lstsqr(X, y):
     """
@@ -15,10 +18,8 @@ def estimate_b_lstsqr(X, y):
     y: np.array (T)
     X: np.array (T x s)
     """
-    lr = LinearRegression(fit_intercept=False, normalize=False,
-                          copy_X=False, n_jobs=1)
-    lr.fit(X, y)
-    return lr.coef_
+    w = linalg.cho_solve(linalg.cho_factor(X.T @ X), X.T @ y)
+    return w
 
 
 def estimate_b_lasso(X, y, criterion="bic"):
@@ -30,8 +31,6 @@ def estimate_b_lasso(X, y, criterion="bic"):
     """
     lasso = LassoLarsIC(criterion=criterion, fit_intercept=False,
                         normalize=False, precompute=True)
-    lr = LinearRegression(fit_intercept=False, normalize=False,
-                          copy_X=False)
 
     # NOTE: This takes 5 - 10 ms for (5000, 125) matrix X
     lasso.fit(X, y)
@@ -42,7 +41,7 @@ def estimate_b_lasso(X, y, criterion="bic"):
 
     # Use lasso only to select the support
     X_lasso = X[:, [i for i, wi in enumerate(w) if wi != 0]]
-    w_lr = lr.fit(X_lasso, y).coef_
+    w_lr = estimate_b_lstsqr(X_lasso, y)
     w[[i for i, wi in enumerate(w) if wi != 0]] = w_lr
     return w
 
@@ -123,10 +122,16 @@ def estimate_B(G, max_lag=10, copy_G=False,
         # Estimate
         b = estimate_b_lasso(X[:max_T], y[:max_T], criterion=ic)
 
+        # Compute residuals
+        r = y - X @ b
+
         # Add the filters as properties to G
         for grouping, j in enumerate(a_i):  # Edge j --b(z)--> i
             G[j][i]["b_hat(z)"] = b[grouping * max_lag: (grouping + 1) * max_lag]
[…]

### 2019-05-29 software/var_system.py

@@ -1,7 +1,7 @@
 import numpy as np
 import networkx as nx
 from LSI_filters import random_arma
-
+from itertools import combinations
 from scipy.linalg import solve_discrete_lyapunov
 
 # TODO: It would probably be wise to do this with scipy sparse matrices!
@@ -270,8 +270,7 @@ def get_estimation_errors(G, G_hat):
             for e in all_edges]
 
 
-def drive_gcg(G, T, sigma2_v, filter_attr="b(z)",
-              stabilize=0.80):
+def drive_gcg(G, T, sigma2_v, filter_attr="b(z)"):
     """
     Drives the gcg G with T samples of Gaussian noise.  sigma_v should
     be a sequence of variances for the noise.
@@ -282,8 +281,6 @@ def drive_gcg(G, T, sigma2_v, filter_attr="b(z)",
     n = len(G.nodes)
     V = np.random.normal(size=(T, n)) * np.sqrt(sigma2_v)
     var = gcg_to_var(G, filter_attr=filter_attr, assert_stable=False)
-    if stabilize:
-        var.ad_hoc_stabilize(rho=stabilize)
     X = var.drive(V)
 
     for i, node_i in enumerate(G.nodes):
@@ -357,9 +354,12 @@ class VAR:
         Stabilizes the system by iteratively dividing down all the
         constituent coefficients.
         """
-        while not self.is_stable(margin=1 - rho):
-            i = np.random.choice(range(self.p))
-            self.B[i] /= 1.25
+        # This doesn't work
+        # while not self.is_stable(margin=1 - rho):
+        #     i = np.random.choice(range(self.p))
+        #     self.B[i] /= 1.25
+
+        self.B = rho * (self.B / (self.get_rho()))
         return
 
     def is_stable(self, margin=1e-6):
@@ -510,3 +510,39 @@ class VAR:
             self._z[:n] = y
         self.x = self._z[:n]  # System output
         return Y
+
+
+def _symmetric_sum(A_list, k):
+    """
+    Calculates a symmetric sum of square matrices in A
+    e_k(A) = sum_{1 <= i1 <= i2 <= ... <= ik <= p} A[i1]A[i2]...A[ik]
+    where p = len(A_list)
+    """
+    p = len(A_list)
+    n = A_list[0].shape[0]
+    A = np.zeros((n, n))
+
+    if k == 1:
+        for j in range(p):
+            A = A + A_list[j]
+    else:
[…]

### 2019-05-31 paper/paper.tex
Removed todo notes in paper.
Working on simulation methods -- results are terrible.  pw_scg is also not returning strongly causal graphs.

@@
 % TODO: read this https://lemire.me/blog/rules-to-write-a-good-research-paper/
-
-% TODO:
-% -- define "DAG"
-% -- remove speculative statements
-% --
 
@@
 \begin{proof}
-  The equivalence $(1) \iff (3)$ is essentially a restatement of the
-  definition, but is in line with the seminal work of Geweke
-  \cite{geweke1982measurement}, \cite{geweke1984}.
-
-  The equivalence $(3) \iff (5)$ is also immediate from the
+  % Geweke uses the log of the ratio of the determinants of the residual variances
+  % 
+  % The equivalence $(1) \iff (3)$ is essentially a restatement of the
+  % definition, but is in line with the seminal work of Geweke
+  % \cite{geweke1982measurement}, \cite{geweke1984}.
+
+  $(1) \iff (3)$ is a tautology, and $(3) \iff (5)$ is immediate from the
   definitions i.e. the projections are equal if and only if the errors
-  are equal.
-
-  We have $(4) \iff (5)$ as a result of Lemma
-  \ref{lem:conditional_orthogonality_equivalence}, the additional
-  equivalence in $(4)$ is apparent since $\H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$.
+  are equal.  We have $(4) \iff (5)$ as a result of Lemma
+  \ref{lem:conditional_orthogonality_equivalence}
+
+% , the additional
+%   equivalence in $(4)$ is apparent since $\H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$.
 

### 2019-05-31 software/LSI_filters.py

@@ -228,14 +228,19 @@ def plot_filter_pz(b, a):
     return
 
 
-def plot_filter(b, a, dB = True, plot_gd = False):
+def plot_filter(b, a, dB=True, plot_gd=False, ret_fig=False,
+                ax_fig=None):
     """
     Plots a filter's response
     """
     w, h = freqz(b, a)
-    fig = plt.figure()
-    ax1 = fig.add_subplot(1,1,1)
-    ax1.set_title("frequency response")
+
+    if ax_fig:
+        ax1, fig = ax_fig
+    else:
+        fig = plt.figure()
+        ax1 = fig.add_subplot(1,1,1)
+        ax1.set_title("frequency response")
 
     if dB:
         ax1.plot(w, 20*np.log10(abs(h)), "b", linewidth = 2)
@@ -253,8 +258,11 @@ def plot_filter(b, a, dB = True, plot_gd = False):
         ax2.plot(w, gd, "g", linewidth = 2)
         ax2.set_ylabel("Group Delay [s]", color = "g")
 
-    fig.show()
-    return
+    if ret_fig:
+        return ax1, fig
+    else:
+        fig.show()
+        return
 
 #-------------------------------------------------------------------------------
 def plot_filter_PSD(b, a, dB = True, title = "PSD"):

### 2019-05-31 software/gc_methods.py

@@ -3,13 +3,14 @@ import numpy as np
 import sympy as sp
 
 from scipy import linalg
+from scipy.linalg import toeplitz
 from sklearn.linear_model import LinearRegression, LassoLarsIC
 
 from itertools import product
 from math import ceil
 
 from var_system import (attach_node_prop, add_self_loops,
-                        remove_zero_filters)
+                        remove_zero_filters, get_X)
 
 def estimate_b_lstsqr(X, y):
     """
@@ -20,6 +21,15 @@ def estimate_b_lstsqr(X, y):
     """
     w = linalg.cho_solve(linalg.cho_factor(X.T @ X), X.T @ y)
     return w
+
+
+def estimate_b_lstsqr_cov(R, r):
+    try:
+        return linalg.cho_solve(linalg.cho_factor(R), r)
+    except linalg.LinAlgError:
+        # print("WARNING: R matrix is not PSD!  lambda_min(R) = {}"
+        #       "".format(np.min(linalg.eigvals(R))))
+        return linalg.solve(R, r, check_finite=True, lower=True)
 
 
 def estimate_b_lasso(X, y, criterion="bic"):
@@ -75,6 +85,21 @@ def form_Xy(X_raw, y_raw, p=10):
     assert X.shape == (T - p, s * p), "Wrong _X shape!"
     assert len(y) == T - p, "wrong y shape!"
     return X, y
+
+
+def block_toeplitz(left_col, top_row=None):
+    p = len(left_col)
+    assert len(top_row) == p
+    assert np.allclose(left_col[0], top_row[0])
+    if top_row is None:
+        top_row = left_col
+
+    f = left_col + top_row[::-1][:-1]
+
+    # return np.block([top_list] +
+    #                 [left_list[tau::-1] +
+    #                  top_list[tau:] for tau in range(1, p)])
+    return np.block([[f[i - j] for j in range(p)] for i in range(p)])
 
 
 def estimate_B(G, max_lag=10, copy_G=False,
@@ -145,15 +170,17 @@ def univariate_AR_error(x, max_lag=10, criterion="bic",
 def bivariate_AR_error(y, x, max_lag=10, criterion="bic",
                        method="lasso"):
     X, y =  form_Xy(X_raw=np.hstack((y[:, None], x[:, None])),
-                    y_raw=y)
+                    y_raw=y, p=max_lag)
     return _compute_AR_error(X, y, criterion, method)
 
 
[…]

### 2019-05-31 software/var_system.py

@@ -280,8 +280,10 @@ def drive_gcg(G, T, sigma2_v, filter_attr="b(z)"):
     """
     n = len(G.nodes)
     V = np.random.normal(size=(T, n)) * np.sqrt(sigma2_v)
+    W = np.random.normal(size=(int(np.sqrt(T)), n)) * np.sqrt(sigma2_v)
     var = gcg_to_var(G, filter_attr=filter_attr, assert_stable=False)
-    X = var.drive(V)
+    var.drive(W)  # Burn in 
+    X = var.drive(V)  # Actual output
 
     for i, node_i in enumerate(G.nodes):
         G.nodes[node_i]["x"] = X[:, i]

### 2019-05-31 software/gc_methods.py
Working on improving pw_scg -- it isn't producing strongly causal graphs.

@@ -259,6 +259,7 @@ def fast_compute_pairwise_gc(X, max_lag=10):
     only difference is that it handles just plain least squares.
 
     TODO: However there are significant discrepancies!
+    TODO: There is something clearly wrong about this implementation.
     """
     xi_i, xi_ij = fast_compute_xi(X, max_lag)
     return compute_gc_score(xi_i, xi_ij)
@@ -299,6 +300,28 @@ def pw_scg(F, delta=None, b=None, R=None):
         else:
             return True
 
+    def get_paths(G, i, j):
+        return list(nx.algorithms.simple_paths.all_simple_paths(G, i, j))
+
+    def is_strongly_causal(G):
+        for j in G.nodes:
+            for i in nx.ancestors(G, j):
+                num_paths = len(get_paths(G, i, j))
+                if num_paths == 0:
+                    raise AssertionError("{} is not an ancestor of {}!"
+                                         "".format(i, j))
+                elif num_paths > 1:
+                    return False
+        return True
+
+    def maintains_strong_causality(G, i, j):
+        """
+        Whether adding edge i, j to G maintains strong causality.
+        """
+        common_ancestors = len(nx.ancestors(G, i) & nx.ancestors(G, j))
+        common_descendants = len(nx.descendants(G, i) & nx.descendants(G, j))
+        return common_ancestors == 0 and common_descendants == 0
+
     n = F.shape[0]
     S = set(range(n))
 
@@ -329,37 +352,42 @@ def pw_scg(F, delta=None, b=None, R=None):
     # incident strength of each node
     I = compute_incident_strength(F, S, W_pred)
     P = arg_select_min_N(I, N=R)
-    P_k = P
+    P_k = [P]
 
     # ------------ Iterations -------------------
     while len(S) != 0:
-        S = S - P_k
+        S = S - P_k[-1]
         I = compute_incident_strength(F, S, W_pred)
-        P_k = arg_select_min_N(I, N=b * len(P))
+        P_k_next = arg_select_min_N(I, N=b * len(P_k[-1]))
+        P_k.append(P_k_next)
 
-        for i, j in sort_edges_by_F(P, P_k, F, W_set):
-            if not any((has_path(G, a, j) for a in G.predecessors(i))):
-                G.add_edge(i, j)
-        P = P | P_k
+        for r in range(1, k + 1):
[…]

### 2019-05-31 paper/paper.tex
Working on statistical tests for pairwise GC.

@@
 
+\section{Finite Sample Graph Recovery}
+\label{sec:structure_learning}
+\hl{TODO: Citation for this scheme}
+In performing pairwise estimates $x_j \pwgc x_i$ we follow the simple
+scheme of estimating the following two linear models via ordinary
+least squares:
+
+\begin{align}
+  H_0:&\ \widehat{x}_i^{(i)}(t) = \sum_{\tau = 1}^{p_i} b_{ii}(\tau)x_i(t - \tau),\\
+  H_1:&\ \widehat{x}_i^{(i, j)}(t) = \sum_{\tau = 1}^{p_i} b_{ii}(\tau)x_i(t - \tau) + \sum_{\tau = 1}^{p_j}b_{ij}(\tau)x_j(t - \tau).
+\end{align}
+
+Under the null hypothesis $H_0$ we can form the asymptotically
+$\chi^2(p_j)$ distributed test statistic
+
+\begin{equation}
+  F_{ij} = \frac{T}{p_j}\Big(\frac{\xi_i}{\xi_{ij}} - 1\Big),
+\end{equation}
+
+where $\xi_i$ is the sample mean square of the residuals $x_i(t) - \widehat{x}^{(i)}_i(t)$, and similarly for $\xi_{ij}$.
+Given additional assumptions on $x(t)$ (\hl{TODO: What assumptions?}) we have the convergence (\hl{a.s.
+or in P?})
+
+\begin{equation}
+  \underset{T \rightarrow \infty}{\text{lim }}F_{ij} =
+  \left\{
+    \begin{array}{ll}
+      0;\ x_j \npwgc x_i\\
+      \infty;\ x_j \pwgc x_i
+    \end{array}
+  \right.
+\end{equation}
+
+Therefore, we can make the statement that Algorithm \ref{alg:pwgr_heuristic} asymptotically implements Algorithm \ref{alg:pwgr}.  \hl{TODO: Modify Algorithm 2 to make this true.}
+
+It is known (\hl{citation}) that $F_{ij}$ is sensitive to the choice of $p_i$ and $p_j$.
+
 \section{Structure Learning}

### 2019-05-31 software/gc_methods.py

@@ -199,11 +199,17 @@ def _fast_compute_AR_error(R, r):
     return max(0, R[0, 0] - w @ r)
 
 
-def compute_gc_score(xi_i, xi_ij):
-    F = -np.log(xi_ij / xi_i[:, None])
-    F[F < 0] = 0.0
-    F[np.isnan(F)] = 0.0
+def compute_gc_score(xi_i, xi_ij, T, p_lags):
+    assert T > 1 + 2 * p_lags,\
+        "T = {} is too small for p_lags = {}".format(T, p_lags)
+    F = T * (xi_i[:, None] / xi_ij - 1)  # This is chi2(p) under the null
+    # F = F * (T - 2 * p_lags - 1) / p_lags
     return F
+
+    # F = -np.log(xi_ij / xi_i[:, None])
+    # F[F < 0] = 0.0
+    # F[np.isnan(F)] = 0.0
+    # return F
 
 
 def compute_xi(X, max_lag, criterion="bic", method="lstsqr"):
@@ -249,8 +255,14 @@ def fast_compute_xi(X, max_lag=10, reg_delta=0.0):
 
 
 def compute_pairwise_gc(X, max_lag=10, criterion="bic", method="lasso"):
+    T, _, p = *X.shape, max_lag
     xi_i, xi_ij = compute_xi(X, max_lag, criterion, method)
-    return compute_gc_score(xi_i, xi_ij)
+    return compute_gc_score(xi_i, xi_ij, T, p)
+
+
+def normalize_gc_score(F, p):
+    F = sps.chi2.cdf(F, p)
+    return F
 
 
 def fast_compute_pairwise_gc(X, max_lag=10):
@@ -261,8 +273,9 @@ def fast_compute_pairwise_gc(X, max_lag=10):
     TODO: However there are significant discrepancies!
     TODO: There is something clearly wrong about this implementation.
     """
+    T, _, p = *X.shape, max_lag
     xi_i, xi_ij = fast_compute_xi(X, max_lag)
-    return compute_gc_score(xi_i, xi_ij)
+    return compute_gc_score(xi_i, xi_ij, T, p)
 
 
 def pw_scg(F, delta=None, b=None, R=None):
@@ -286,7 +299,7 @@ def pw_scg(F, delta=None, b=None, R=None):
         return set(arg_sorted[:N])
 
     def compute_incident_strength(F, S, W_pred):
[…]

### 2019-06-10 paper/paper.tex
Updated paper with hypothesis test section and algorithm to match code.

--- 2019-05-31 paper/paper.tex
@@
 \begin{definition}[Granger Causality]
+  \label{def:granger_causality}
   For the WSS series $x(t)$ we say that component $x_j$
@@
 
+\section{Finite Sample Graph Recovery}
+\label{sec:structure_learning}
+In this section we provide a review of our methods for implementing
+Algorithm 1 given a \textit{finite} sample of $T$ data points.  We
+apply the simplest reasonable methods in order to maintain a focus on
+our main contributions (i.e. Algorithm \ref{alg:pwgr}), more
+sophisticated schemes can only serve to improve the results.  Textbook
+reviews of the following concepts are provided e.g. by
+\cite{all_of_statistics}, \cite{murphy_mlp}, and elsewhere.
+
+We proceed first by defining pairwise Granger-causality hypothesis
+tests, second a model order selection criteria, and finally the method
+for choosing a $P-$value cutoff.
+
+\subsection{Hypothesis Testing}
+In performing pairwise checks for Granger-causality $x_j \pwgc x_i$ we
+follow the simple scheme of estimating the two linear models via
+ordinary least squares:
+
+\begin{align}
+  H_0:&\ \widehat{x}_i^{(p_i)}(t) = \sum_{\tau = 1}^{p_i} b_{ii}(\tau)x_i(t - \tau),\\
+  H_1:&\ \widehat{x}_i^{(p_i, p_j)}(t) = \sum_{\tau = 1}^{p_i} b_{ii}(\tau)x_i(t - \tau) + \sum_{\tau = 1}^{p_j}b_{ij}(\tau)x_j(t - \tau).
+\end{align}
+
+Under the null hypothesis $H_0$ (i.e. that $x_j \npwgc x_i$) we can
+form the asymptotically $\chi^2(p_j)$ distributed test statistic
+(\hl{TODO: Citation that this is chi2})
+
+\begin{equation}
+  \label{eqn:gc_statistics}
+  F_{ij}(p_i, p_j) = \frac{T}{p_j}\Big(\frac{\xi_i(p_i)}{\xi_{ij}(p_i, p_j))} - 1\Big),
+\end{equation}
+
+where $\xi_i(p_i)$ is the sample mean square of the residuals\footnote{This
+  quantity is often denoted $\widehat{\sigma}$, but we maintain
+  notation from Definition \ref{def:granger_causality}.}
+$x_i(t) - \widehat{x}^{(p_i)}_i(t)$
+
+\begin{equation*}
+  \xi_i(p_i) = \frac{1}{T - p_i}\sum_{t = p_i + 1}^T (x_i(t) - \widehat{x}_i^{(p_i)}(t))^2
+\end{equation*}
+
+Given additional assumptions on $x(t)$ (\hl{TODO: What assumptions? --
+  this must also depend on choice of $p_i, p_j$}) we have the
+convergence (\hl{a.s. or in P?  -- I think in P})
+
+\begin{equation}
+  \underset{T \rightarrow \infty}{\text{lim }}F_{ij} =
+  \left\{
+    \begin{array}{ll}
+      0;\ x_j \npwgc x_i\\
+      \infty;\ x_j \pwgc x_i
+    \end{array}
+  \right.
+\end{equation}
+
+Therefore, we can make the statement that Algorithm
+\ref{alg:pwgr_heuristic} asymptotically implements Algorithm
+\ref{alg:pwgr}.  \hl{TODO: Modify Algorithm 2 to make this true.}
+\hl{TODO: Enclose this into a theorem}
+
+\subsection{Model Order Selection}
+\label{sec:model_order_selection}
+There are a variety of methods to choose the filter orders $p_i$ and
+$p_j$ \hl{TODO: cite some more}, we will focus in particular on the
+Bayesian Information Criteria (BIC).  The BIC is substantially more
+conservative than the Akaiake Information Criteria, and since we are
+searching for \textit{sparse graphs}, we therefore prefer the BIC. For
+the bivariate case we focus for simplicity on $p_i = p_j$ and write
+simply\footnote{We use the same $p$ for the bivariate case, but this
+  need not be the same as the corresponding univariate estimate}
+$x_{ij}^{(p)}$, in which case the BIC, which we seek to
+\textit{maximize} over $p$ is given by:
+
+\begin{equation}
+  \label{eqn:bic}
+  \begin{aligned}
+    BIC_{\text{univariate}}(p) &= -(T - p - 1) \ln \xi_i(p) - p\ln T,\\
+    BIC_{\text{bivariate}}(p) &= -(T - p - 1) \ln \xi_{ij}(p) - 2p\ln T.\\
+  \end{aligned}
+\end{equation}
+
+We carry this out via the forward search described in Algorithm
+\ref{alg:maximize_bic}.  Since $BIC(p)$ need not be a convex function,
+this scheme is not guaranteed to find the true maximum of the BIC, but
+this approximation suffices for our purposes \hl{TODO: is this
+  sufficient?  A full search is not so much harder.}
+
+\begin{algorithm}
+  \caption{BIC Model Selection Algorithm}
+  \label{alg:maximize_bic}
+
+  \SetKwInOut{Input}{input}
+  \SetKwInOut{Output}{output}
+  \SetKwInOut{Initialize}{initialize}
+  \DontPrintSemicolon
+
+  \BlankLine
+  \Input{$T$ samples of WSS processes $x_i(t), x_j(t)$.  A maximum possible lag $p_{\text{max}}$.}
+  \Output{Model orders $p_i$ and $p$ for models in Hypotheses $H_0$ and $H_1$ respectively.}
+  \Initialize{
+    $BIC_{\text{univariate}} = -\infty$\\
+    $BIC_{\text{bivariate}} = -\infty$\\
+  }
+
+  \# \texttt{Find $p_i$}\\
+  \For{$p \in \{1, 2, \ldots, p_{\text{max}}\}$}{
+    $\xi_i(p) = $\ \texttt{fit\_univariate}$(x_i(t), p)$ \texttt{ \# Fit model for $H_0$}\\
+    $bic_p = -(T - p - 1) \ln \xi_i(p) - p\ln(T - p - 1)$\\
+    \If{$bic_p > BIC_{\text{univariate}}$}{
+      $BIC_{\text{univariate}} = bic_p$\\
+      \texttt{continue}\\
+    } \Else{
+      $p_i = p - 1$\\
+      \texttt{break}\\
+    }
+  }
+
+  \BlankLine
+  \# \texttt{Find $p$}\\
+  \For{$p \in \{1, 2, \ldots, p_{\text{max}}\}$}{
+    $\xi_{ij}(p) = $\ \texttt{fit\_biivariate}$(x_i(t), x_j(t), p)$ \texttt{ \# Fit model for $H_1$}\\
+    $bic_p = -(T - p - 1) \ln \xi_{ij}(p) - 2p\ln(T - p - 1)$\\
+    \If{$bic_p > BIC_{\text{bivariate}}$}{
+      $BIC_{\text{biivariate}} = bic_p$\\
+      \texttt{continue}\\
+    }\Else{
+      $p = p - 1$\\
+      \texttt{break}\\
+    }
+  }
+    \Return{$p_i, p$}
+\end{algorithm}
+
+\subsection{Edge Probabilities and Error Rate Controls}
+Denote $F_{ij}$ the Granger-causality statistic of equation
+\ref{eqn:gc_statistics} with model orders chosen by the methods of
+Section \ref{sec:model_order_selection}.  Note that this statistic is
+asymptotically $\chi^2(p)$ distributed, and denote by $G$ the
+$\chi^2(p)$ cumulative distribution function.  We will define the matrix
+
+\begin{equation}
+  \label{eqn:edge_inclusion_probability}
+  P_{ij} = G(F_{ij}),
+\end{equation}
+
+to be the matrix of pairwise edge inclusion probabilities.  This is
+motivated by the hypothesis test where the hypothesis $H_0$ will be
+rejected (and thence we will conclude that $x_j \pwgc x_i$) if
+$P_{ij} > 1 - \delta$.
+
+The value $\delta$ can be chosen by a variety of methods, in our case
+we apply the Benjamini Hochberg criteria \cite{benjamini_hochberg}
+\cite{all_of_statistics} to control the false discovery rate of
+pairwise edges to a level $\alpha$ (where we generally take
+$\alpha = 0.05$).
+
 \section{Structure Learning}
@@
 Instead, we convert the algorithm into an heuristic and observe it's performance experimentally.
-We first consider the heuristic version of Algorithm \ref{alg:pwgr} in Algorithm \ref{alg:pwgr_heuristic} and show that is a well defined procedure:
+We first consider the heuristic version of Algorithm \ref{alg:pwgr} in Algorithm \ref{alg:pwgr_heuristic} and show that is a well defined procedure (note that $\ceil{x}$ denotes the integer ceiling function):
 
@@
   \BlankLine
-  \caption{Graph Recovery Heuristic $\mathsf{pw\_scg}(F, \delta, b, R)$}
+  \caption{Graph Recovery Heuristic $\mathsf{pw\_scg}(F, P, \alpha)$}
   \label{alg:pwgr_heuristic}
-  % \TitleOfAlgo{Pairwise Graph Recovery}
-  \Input{Estimates of pairwise Granger-causaltiy statistics $F_{ij} = \ln \frac{\xi_{ij}}{\xi_i}$\\
-    Sparsity parameters ($\delta, b, R$)}
+  \Input{Estimates of pairwise Granger-causaltiy statistics $F_{ij}$ (eqn \ref{eqn:gc_statistics}), Matrix of edge probabilities $P_{ij}$ (eqn \ref{eqn:edge_inclusion_probability}), threshold $\delta$.}
   \Output{A strongly causal graph $\widehat{\gcg}$}
@@
 
-  $W_\delta \leftarrow \{(i, j)\ |\ F_{ji} \ge \delta, F_{ji} > F_{ij}\}$  \texttt{\# candidate edges}\\
-  $\mathcal{I}_0 \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# total incident strength for each node}\\
-  $P_0 \leftarrow \mathsf{arg\_select\_min\_N}(\mathcal{I}_0, N=R)$  \texttt{\# select} $R$ \texttt{root nodes}\\
+  $W_\delta \leftarrow \{(i, j)\ |\ P_{ji} > 1 - \delta, F_{ji} > F_{ij}\}$  \texttt{\# candidate edges}\\
+  $\mathcal{I}_0 \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# total node incident probability}\\
+  $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) < \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Nodes with fewest incident edges}\\
+  \If{$P_0 = \emptyset$}{
+    $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) \le \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Ensure non-empty}
+  }
   \BlankLine
@@
     $S \leftarrow S \setminus P_{k - 1}$ \texttt{\# remove processed nodes}\\
-    $\mathcal{I}_k \leftarrow \big(\sum_{j \in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# intra-}$S$ \texttt{incident strength}\\
-    $P_k \leftarrow \mathsf{arg\_select\_min\_N}(\mathcal{I}_k, N=b)$  \texttt{\# select} $b$ \texttt{branching nodes}\\
+    % $\mathcal{I}_k \leftarrow \big(\sum_{j \in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# intra-}$S$ \texttt{incident strength}\\
+    $\mathcal{I}_k \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$\\
+    $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) < \ceil{\text{min}(\mathcal{I}_k)}\}$\\
+    \If{$P_k = \emptyset$}{
+      $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) \le \ceil{\text{min}(\mathcal{I}_k)}\}$
+    }
     \;
-    \texttt{\# add strongest} $P_{k - r} \rightarrow P_k$ \texttt{edges and maintain strong causality}\\
-    \For{$r = 1, \ldots, k$} 
-    {
-      \For{$(i, j) \in \mathsf{sort}\Big(\{(i, j) \in P_{k - r} \times P_k\ |\ (i, j) \in W_\delta\} \mathsf{\ by\ descending\ } F_{ji}\Big)$}{
-        \If{no $\gcgpath{u}{j}$ path in $E$, for any $u \in \anc{i}$}{
+    \texttt{\# add strongest edges, maintaining strong causality}\\
+    $U_k \leftarrow \bigcup_{r = 1}^k P_{k - r}$ \texttt{\# Include all forward edges}\\
+    \For{$(i, j) \in \mathsf{sort}\Big(\{(i, j) \in U_k \times P_k\ |\ (i, j) \in W_\delta\} \mathsf{\ by\ descending\ } F_{ji}\Big)$} {
+      \If{$\mathsf{is\_strongly\_causal}(E \cup \{(i, j)\})$} {
         $E \leftarrow E \cup \{(i, j)\}$
@@
     }
-    }
-    $k \leftarrow k + 1$
+    $k \leftarrow k + 1$\\
   }

### 2019-06-10 software/gc_methods.py

@@ -3,14 +3,17 @@ import numpy as np
 import sympy as sp
 
 from scipy import linalg
+from scipy import stats as sps
 from scipy.linalg import toeplitz
 from sklearn.linear_model import LinearRegression, LassoLarsIC
 
 from itertools import product
+from functools import reduce
 from math import ceil
 
 from var_system import (attach_node_prop, add_self_loops,
                         remove_zero_filters, get_X)
+from stat_util import benjamini_hochberg
 
 def estimate_b_lstsqr(X, y):
     """
@@ -103,7 +106,7 @@ def block_toeplitz(left_col, top_row=None):
 
 
 def estimate_B(G, max_lag=10, copy_G=False,
-               max_T=None, ic="bic"):
+               max_T=None, ic="bic", method="lasso"):
     """
     Estimate x(t) = B(z)x(t) respecting the topology of G.
 
@@ -145,7 +148,12 @@ def estimate_B(G, max_lag=10, copy_G=False,
         X, y = form_Xy(X_raw, y_raw, p=max_lag)
 
         # Estimate
-        b = estimate_b_lasso(X[:max_T], y[:max_T], criterion=ic)
+        if method == "lasso":
+            b = estimate_b_lasso(X[:max_T], y[:max_T], criterion=ic)
+        elif method == "lstsqr":
+            b = estimate_b_lstsqr(X[:max_T], y[:max_T])
+        else:
+            raise AssertionError("Bad method but should deal with it earlier!")
 
         # Compute residuals
         r = y - X @ b
@@ -162,16 +170,52 @@ def estimate_B(G, max_lag=10, copy_G=False,
 
 # NOTE: forming the Xy matrices consumes only microseconds
 def univariate_AR_error(x, max_lag=10, criterion="bic",
-                        method="lasso"):
-    X, y = form_Xy(X_raw=x[:, None], y_raw=x, p=max_lag)
-    return _compute_AR_error(X, y, criterion, method)
+                        method="lasso", ret_p_opt=False):
+    # Try larger model orders until we don't increase the BIC
+    if criterion != "bic":
+        raise NotImplementedError("Only BIC criterion is supported")
+
+    bic = -np.infty
+    T = len(x)
+    for p in range(1, max_lag + 1):
+        X_, y_ = form_Xy(X_raw=x[:, None], y_raw=x, p=p)
+        sv2_p = _compute_AR_error(X_, y_, criterion="bic",
[…]

### 2019-06-10 software/gc_methods.py
A bit of code cleanup, removing unused options etc.

@@ -169,8 +169,7 @@ def estimate_B(G, max_lag=10, copy_G=False,
 
 
 # NOTE: forming the Xy matrices consumes only microseconds
-def univariate_AR_error(x, max_lag=10, criterion="bic",
-                        method="lasso", ret_p_opt=False):
+def univariate_AR_error(x, max_lag=10, criterion="bic"):
     # Try larger model orders until we don't increase the BIC
     if criterion != "bic":
         raise NotImplementedError("Only BIC criterion is supported")
@@ -179,8 +178,7 @@ def univariate_AR_error(x, max_lag=10, criterion="bic",
     T = len(x)
     for p in range(1, max_lag + 1):
         X_, y_ = form_Xy(X_raw=x[:, None], y_raw=x, p=p)
-        sv2_p = _compute_AR_error(X_, y_, criterion="bic",
-                                  method="lstsqr")
+        sv2_p = _compute_AR_error(X_, y_, criterion="bic")
         bic_p = -(T - p - 1) * np.log(sv2_p) - p * np.log(T - p - 1)
         if bic_p > bic:
             bic = bic_p
@@ -188,14 +186,10 @@ def univariate_AR_error(x, max_lag=10, criterion="bic",
         else:
             p = p - 1
             break
-    if ret_p_opt:
-        return sv2_p, p
-    else:
-        return sv2_p
+    return sv2_p, p
 
 
-def bivariate_AR_error(y, x, max_lag=10, criterion="bic",
-                       method="lstsqr", ret_p_opt=False):
+def bivariate_AR_error(y, x, max_lag=10, criterion="bic"):
     if criterion != "bic":
         raise NotImplementedError("Only BIC criterion is supported")
 
@@ -204,7 +198,7 @@ def bivariate_AR_error(y, x, max_lag=10, criterion="bic",
     for p in range(1, max_lag + 1):
         X_, y_ =  form_Xy(X_raw=np.hstack((y[:, None], x[:, None])),
                           y_raw=y, p=p)
-        sv2_p = _compute_AR_error(X_, y_, criterion="bic", method="lstsqr")
+        sv2_p = _compute_AR_error(X_, y_, criterion="bic")
         bic_p = -(T - p - 1) * np.log(sv2_p) - 2 * p * np.log(T - p - 1)
         if bic_p > bic:
             bic = bic_p
@@ -212,10 +206,7 @@ def bivariate_AR_error(y, x, max_lag=10, criterion="bic",
         else:
[…]

### 2019-06-10 software/gc_methods.py
Removed all the criterion options -- use only BIC.

@@ -35,14 +35,14 @@ def estimate_b_lstsqr_cov(R, r):
         return linalg.solve(R, r, check_finite=True, lower=True)
 
 
-def estimate_b_lasso(X, y, criterion="bic"):
+def estimate_b_lasso(X, y):
     """
     Estimate y_hat = Xb via LASSO and using BIC to choose regularizers.
 
     y: np.array (T)
     X: np.array (T x s)
     """
-    lasso = LassoLarsIC(criterion=criterion, fit_intercept=False,
+    lasso = LassoLarsIC(criterion="bic", fit_intercept=False,
                         normalize=False, precompute=True)
 
     # NOTE: This takes 5 - 10 ms for (5000, 125) matrix X
@@ -106,7 +106,7 @@ def block_toeplitz(left_col, top_row=None):
 
 
 def estimate_B(G, max_lag=10, copy_G=False,
-               max_T=None, ic="bic", method="lasso"):
+               max_T=None, method="lasso"):
     """
     Estimate x(t) = B(z)x(t) respecting the topology of G.
 
@@ -149,7 +149,7 @@ def estimate_B(G, max_lag=10, copy_G=False,
 
         # Estimate
         if method == "lasso":
-            b = estimate_b_lasso(X[:max_T], y[:max_T], criterion=ic)
+            b = estimate_b_lasso(X[:max_T], y[:max_T])
         elif method == "lstsqr":
             b = estimate_b_lstsqr(X[:max_T], y[:max_T])
         else:
@@ -169,16 +169,13 @@ def estimate_B(G, max_lag=10, copy_G=False,
 
 
 # NOTE: forming the Xy matrices consumes only microseconds
-def univariate_AR_error(x, max_lag=10, criterion="bic"):
+def univariate_AR_error(x, max_lag=10):
     # Try larger model orders until we don't increase the BIC
-    if criterion != "bic":
-        raise NotImplementedError("Only BIC criterion is supported")
-
     bic = -np.infty
     T = len(x)
     for p in range(1, max_lag + 1):
         X_, y_ = form_Xy(X_raw=x[:, None], y_raw=x, p=p)
-        sv2_p = _compute_AR_error(X_, y_, criterion="bic")
+        sv2_p = _compute_AR_error(X_, y_)
         bic_p = -(T - p - 1) * np.log(sv2_p) - p * np.log(T - p - 1)
         if bic_p > bic:
             bic = bic_p
[…]

### 2019-06-10 software/gc_methods.py
Notes.

@@ -204,11 +204,14 @@ def bivariate_AR_error(y, x, max_lag=10):
 
 
 def compute_covariances(X, p, symmetrize=False):
+    # We do not do any normalizing
+    # I believe this should ensure positive-definite ness.
+    # We are effectively /windowing/ the signal.
     T = X.shape[0]
     R = np.dstack(
         [X[p:, :].T @ X[p:, :]] + 
         [X[p:, :].T @ X[p - tau: -tau, :]
-         for tau in range(1, p + 1)]) / (T - p + 1)
+         for tau in range(1, p + 1)])
     return R
 
 

### 2019-06-11 software/gc_methods.py
Updated compute_covariance method.

@@ -209,8 +209,8 @@ def compute_covariances(X, p, symmetrize=False):
     # We are effectively /windowing/ the signal.
     T = X.shape[0]
     R = np.dstack(
-        [X[p:, :].T @ X[p:, :]] + 
-        [X[p:, :].T @ X[p - tau: -tau, :]
+        [X.T @ X] +
+        [X[tau:, :].T @ X[0: -tau, :]
          for tau in range(1, p + 1)])
     return R
 

### 2019-06-11 paper/paper.tex
Fixing "_fast" computation methods.

@@
 \bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
+% \bibliography{\string~global_bib.bib}
 
@@
 \input{\string~/Documents/academics/global_academics/latex_preamble}
+% \input{\string~latex_preamble}
 
@@
 
-We carry this out via the forward search described in Algorithm
-\ref{alg:maximize_bic}.  Since $BIC(p)$ need not be a convex function,
-this scheme is not guaranteed to find the true maximum of the BIC, but
-this approximation suffices for our purposes \hl{TODO: is this
-  sufficient?  A full search is not so much harder.}
-
-\begin{algorithm}
-  \caption{BIC Model Selection Algorithm}
-  \label{alg:maximize_bic}
-
-  \SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{output}
-  \SetKwInOut{Initialize}{initialize}
-  \DontPrintSemicolon
-
-  \BlankLine
-  \Input{$T$ samples of WSS processes $x_i(t), x_j(t)$.  A maximum possible lag $p_{\text{max}}$.}
-  \Output{Model orders $p_i$ and $p$ for models in Hypotheses $H_0$ and $H_1$ respectively.}
-  \Initialize{
-    $BIC_{\text{univariate}} = -\infty$\\
-    $BIC_{\text{bivariate}} = -\infty$\\
-  }
-
-  \# \texttt{Find $p_i$}\\
-  \For{$p \in \{1, 2, \ldots, p_{\text{max}}\}$}{
-    $\xi_i(p) = $\ \texttt{fit\_univariate}$(x_i(t), p)$ \texttt{ \# Fit model for $H_0$}\\
-    $bic_p = -(T - p - 1) \ln \xi_i(p) - p\ln(T - p - 1)$\\
-    \If{$bic_p > BIC_{\text{univariate}}$}{
-      $BIC_{\text{univariate}} = bic_p$\\
-      \texttt{continue}\\
-    } \Else{
-      $p_i = p - 1$\\
-      \texttt{break}\\
-    }
-  }
-
-  \BlankLine
-  \# \texttt{Find $p$}\\
-  \For{$p \in \{1, 2, \ldots, p_{\text{max}}\}$}{
-    $\xi_{ij}(p) = $\ \texttt{fit\_biivariate}$(x_i(t), x_j(t), p)$ \texttt{ \# Fit model for $H_1$}\\
-    $bic_p = -(T - p - 1) \ln \xi_{ij}(p) - 2p\ln(T - p - 1)$\\
-    \If{$bic_p > BIC_{\text{bivariate}}$}{
-      $BIC_{\text{biivariate}} = bic_p$\\
-      \texttt{continue}\\
-    }\Else{
-      $p = p - 1$\\
-      \texttt{break}\\
-    }
-  }
-    \Return{$p_i, p$}
-\end{algorithm}
+We carry this out by a simple direct search on each model order
+between $1$ and some prescribed $p_\text{max}$.  In practice it is
+sufficient to pick $p_\text{max}$ ad-hoc or via some simple heuristic
+(e.g. plotting the sequence $BIC(p)$ over $p$), and we will review in
+Section \ref{sec:efficient_model_estimation} an algorithm to carry out
+this search efficiently.
+
+% \begin{algorithm}
+%   \caption{BIC Model Selection Algorithm}
+%   \label{alg:maximize_bic}
+
+%   \SetKwInOut{Input}{input}
+%   \SetKwInOut{Output}{output}
+%   \SetKwInOut{Initialize}{initialize}
+%   \DontPrintSemicolon
+
+%   \BlankLine
+%   \Input{$T$ samples of WSS processes $x_i(t), x_j(t)$.  A maximum possible lag $p_{\text{max}}$.}
+%   \Output{Model orders $p_i$ and $p$ for models in Hypotheses $H_0$ and $H_1$ respectively.}
+%   \Initialize{
+%     $BIC_{\text{univariate}} = -\infty$\\
+%     $BIC_{\text{bivariate}} = -\infty$\\
+%   }
+
+%   \# \texttt{Find $p_i$}\\
+%   \For{$p \in \{1, 2, \ldots, p_{\text{max}}\}$}{
+%     $\xi_i(p) = $\ \texttt{fit\_univariate}$(x_i(t), p)$ \texttt{ \# Fit model for $H_0$}\\
+%     $bic_p = -(T - p - 1) \ln \xi_i(p) - p\ln(T - p - 1)$\\
+%     \If{$bic_p > BIC_{\text{univariate}}$}{
+%       $BIC_{\text{univariate}} = bic_p$\\
+%       \texttt{continue}\\
+%     } \Else{
+%       $p_i = p - 1$\\
+%       \texttt{break}\\
+%     }
+%   }
+
+%   \BlankLine
+%   \# \texttt{Find $p$}\\
+%   \For{$p \in \{1, 2, \ldots, p_{\text{max}}\}$}{
+%     $\xi_{ij}(p) = $\ \texttt{fit\_biivariate}$(x_i(t), x_j(t), p)$ \texttt{ \# Fit model for $H_1$}\\
+%     $bic_p = -(T - p - 1) \ln \xi_{ij}(p) - 2p\ln(T - p - 1)$\\
+%     \If{$bic_p > BIC_{\text{bivariate}}$}{
+%       $BIC_{\text{biivariate}} = bic_p$\\
+%       \texttt{continue}\\
+%     }\Else{
+%       $p = p - 1$\\
+%       \texttt{break}\\
+%     }
+%   }
+%     \Return{$p_i, p$}
+% \end{algorithm}
+
+\subsection{Efficient Model Estimation}
+\label{sec:efficient_model_estimation}
+In practice, the vast majority of computational effort involved in
+implementing our estimation algorithm is spent calculating the error
+estimates $\xi_i(p_i)$ and $\xi_{ij}(p, p)$.  This requires fitting a
+total of $n^2p_{\text{max}}$ autoregressive models, where the most
+naive algorithm (e.g. solving a least squares problem for each model)
+for this task will consume $O(n^2p_{\text{max}}^4T)$ time, it is
+possible to carry out this task in a much more modest
+$O(n^2p_{\text{max}}^2 ) + O(n^2p_{\text{max}}T)$ time.
+
+Firstly, we estimate the sequence of Covariance
+matrices\footnote{The particular indexing and normalization given in
+  equation \ref{eqn:covariance_estimate} is critical to ensure
+  $\widehat{R}$ is positive semidefinite.  The estimate can be viewed
+  as calculating the covariance sequence of a signal multiplied by a
+  rectangular window.}
+
+\begin{equation}
+  \label{eqn:covariance_estimate}
+  \widehat{R}_x(\tau) = \frac{1}{T}\sum_{t = \tau + 1}^T x(t) x(t - \tau)^\T;\ \tau = 0, \ldots, p_{\text{max}},
+\end{equation}
+
+an $O(n^2p_{\text{max}}T)$ operation.  Secondly, the variance
+estimates $\xi_i(p)$ for $p = 1, \ldots, p_{\text{max}}$ can be
+evaluated in $O(p_{\text{max}}^2)$ time each by applying the
+Levinson-Durbin recursion to $\widehat{R}_{ii}(tau)$, which
+effectively estimates a sequence of $AR$ models, producing $\xi_i(p)$
+as a side-effect (see
+\cite{hayes_statistical_digital_signal_processing} and
+\cite{levinson_durbin_recursion}).
+
+Similarly, the variance estimates $\xi_{ij}$ and $\xi_{ji}$ can be
+obtained by estimating $\frac{(n + 1)n}{2}$ bivariate AR models, again
+in $O(p_{\text{max}}^2)$ via Whittle's generalized LD recursion
+\cite{whittle_generalized_levinson_durbin}.
 

### 2019-06-11 software/gc_methods.py

@@ -27,12 +27,7 @@ def estimate_b_lstsqr(X, y):
 
 
 def estimate_b_lstsqr_cov(R, r):
-    try:
-        return linalg.cho_solve(linalg.cho_factor(R), r)
-    except linalg.LinAlgError:
-        # print("WARNING: R matrix is not PSD!  lambda_min(R) = {}"
-        #       "".format(np.min(linalg.eigvals(R))))
-        return linalg.solve(R, r, check_finite=True, lower=True)
+    return linalg.cho_solve(linalg.cho_factor(R), r)
 
 
 def estimate_b_lasso(X, y):
@@ -215,10 +210,11 @@ def compute_covariances(X, p, symmetrize=False):
     # I believe this should ensure positive-definite ness.
     # We are effectively /windowing/ the signal.
     T = X.shape[0]
-    R = np.dstack(
-        [X.T @ X] +
-        [X[tau:, :].T @ X[0: -tau, :]
-         for tau in range(1, p + 1)])
+    R = np.stack(
+        [X.T @ X / T] +
+        [X[tau:, :].T @ X[: -tau, :] / T
+         for tau in range(1, p + 1)],
+         axis=0)
     return R
 
 
@@ -254,6 +250,32 @@ def _fast_compute_AR_error(R, r):
     w = estimate_b_lstsqr_cov(R, r)
     return max(0, R[0, 0] - w @ r)
 
+
+def _fast_univariate_AR_error(r, T):
+    # TODO: Apply levinson recursion
+    p_max = len(r) - 1
+    eps = np.nan * np.zeros(p_max + 1)
+    eps[0] = r[0]
+    for p in range(1, p_max + 1):
+        # R = toeplitz(r[:-1])
+        # _r = r[1:]
+        R = toeplitz(r[:p])
+        _r = r[1: p + 1]
+        w = estimate_b_lstsqr_cov(R, _r)
+        eps[p] = r[0] - w @ _r
+
+    bic = compute_bic(eps, T, s=1)
+    p_opt = np.argmax(bic)
+    return eps[p_opt], p_opt
+
+
+def compute_bic(eps, T, s=1):
+    """
+    set s = 1 for univariate case and s = 2 for bivariate case.
+    """
+    bic = [-T * np.log(eps_p) - s * p * np.log(T)
+           for p, eps_p in enumerate(eps)]
+    return bic
 
 def compute_gc_score(xi_i, xi_ij, T, p_lags):
[…]

### 2019-06-11 software/gc_methods.py
Minor update on using lev_durb.

@@ -253,6 +253,9 @@ def _fast_compute_AR_error(R, r):
 
 def _fast_univariate_AR_error(r, T):
     # TODO: Apply levinson recursion
+    # TODO: How to import the levinson project?
+    # _, _, eps = lev_durb(r)
+
     p_max = len(r) - 1
     eps = np.nan * np.zeros(p_max + 1)
     eps[0] = r[0]

### 2019-06-11 software/simulations.py
Some untracked but should-be-tracked files.

@@ -1,0 +1,172 @@
+import seaborn.apionly as sns
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+
+from matplotlib import rc as mpl_rc
+font = {"family" : "normal",
+        "weight" : "bold",
+        "size"   : 22}
+
+mpl_rc("font", **font)
+mpl_rc("text", usetex=True)
+
+from var_system import (random_gnp_dag, drive_gcg, get_errors,
+                        get_estimation_errors, random_tree_dag,
+                        get_X, attach_node_prop, gcg_to_var)
+from gc_methods import (compute_pairwise_gc, estimate_B, pw_scg,
+                        estimate_graph, full_filter_estimator,
+                        compute_MCC_score)
+
+
+def full_single_pass_experiment(simulation_name, graph_type, M_passes=1):
+    n_nodes, p_lags, max_lags = 50, 5, 10
+    N_iters = 50
+
+    T_max = 11000
+    # T_iters = [100, 500, 1000, 2500, 5000, 10000]
+    T_iters = [100, 500, 1000, 2500, 10000]
+
+    if graph_type == "Tree":  # TODO: This doesn't actually produce trees!
+        random_graph = lambda: random_tree_dag(
+            n_nodes, p_lags, pole_rad=0.75)
+    elif graph_type == "DAG":
+        random_graph = lambda: random_gnp_dag(
+            n_nodes, p_lags, pole_rad=0.75,
+            edge_prob=2. / n_nodes)
+    elif graph_type == "ErdosRenyi":
+        random_graph = lambda: random_gnp(
+            n_nodes, p_lags, pole_rad=0.50,
+            edge_prob=2. / n_nodes)
+    else:
+        raise NotImplementedError("graph_type '{}' is not understood"
+                                  "".format(graph_type))
+
+    b_errs = np.zeros((N_iters, len(T_iters)))
+    errs = np.zeros_like(b_errs)
+    errs_vec = np.zeros((N_iters, len(T_iters), n_nodes))
+    true_errs = np.zeros_like(b_errs)
+    true_errs_vec = np.zeros_like(errs_vec)
+    rhos = np.zeros_like(b_errs)
+    N_cross_edges = np.zeros_like(b_errs)
+    N_hat_cross_edges = np.zeros_like(b_errs)
+    N_cross_intersect_edges = np.zeros_like(b_errs)
+
+    # TODO: There is a ton of repeated work going on all over
[…]

### 2019-06-13 software/pwgc/gc_methods.py
Some file rearranging.

@@ -1,11 +1,14 @@
+"""
+Methods which implement granger-causality computations.
+"""
+
 import networkx as nx
 import numpy as np
-import sympy as sp
 
 from scipy import linalg
 from scipy import stats as sps
 from scipy.linalg import toeplitz
-from sklearn.linear_model import LinearRegression, LassoLarsIC
+from sklearn.linear_model import LassoLarsIC
 
 from itertools import product
 from functools import reduce
@@ -13,7 +16,11 @@ from math import ceil
 
 from var_system import (attach_node_prop, add_self_loops,
                         remove_zero_filters, get_X)
-from stat_util import benjamini_hochberg
+try:
+    from .stat_util import benjamini_hochberg
+except ImportError:
+    from stat_util import benjamini_hochberg
+
 
 def estimate_b_lstsqr(X, y):
     """
@@ -627,15 +634,15 @@ def estimate_graph(X, G, max_lags=10, method="lasso"):
 
 
 def compute_tp_tn(N_edges, N_hat_edges, N_intersect_edges, n_nodes):
-    P = N_edges  # Positives
+    # P = N_edges  # Positives
     TP = N_intersect_edges  # True positives
     FP = N_hat_edges - N_intersect_edges  # False positives
-    TPR = TP / P  # True positive rate (Sensitivity)
+    # TPR = TP / P  # True positive rate (Sensitivity)
 
     N = n_nodes**2 - N_edges  # Negatives
     FN = N_edges - N_intersect_edges  # False negatives
     TN = N - FN  # True negatives
-    TNR = TN / N  # True negative rate
+    # TNR = TN / N  # True negative rate
     return TP, TN, FP, FN
 
 
@@ -667,4 +674,3 @@ def example_graph():
     # set(nx.ancestors(G, i)): set of ancestors of i in G
     # A = nx.adjacency_matrix(G).todense(): Find adj matrix
     return
-

### 2019-06-13 software/pwgc/draw_graphs.py
New files etc.

@@ -1,0 +1,115 @@
+import networkx as nx
+import matplotlib.pyplot as plt
+
+from networkx.drawing.nx_agraph import write_dot, graphviz_layout
+
+from var_system import (random_gnp_dag, random_tree_dag)
+
+def draw_graph(G, file_name=None):
+    pos = graphviz_layout(G, prog="dot")
+    nx.draw(G, pos, arrowsize=30,
+            with_labels=True, linewidth=2,
+            font_size=24, font_family="serif",
+            font_weight="bold", linewidths=2,
+            alpha=0.9, node_color="lightsalmon")
+
+    if file_name is not None:
+        if isinstance(file_name, str):
+            plt.savefig(file_name)
+        else:  # Allow passing a list
+            [plt.savefig(f) for f in file_name]
+    plt.show()
+    return
+
+
+def draw_graph_estimates(G, G_hat, file_name=None,
+                         count_self_loops=False):
+    # To ensure we don't mutate them
+    _G = G.copy()
+    _G_hat = G_hat.copy()
+
+    nodes = _G.nodes
+    if not count_self_loops:
+        _G.remove_edges_from([(i, i) for i in nodes if (i, i) in G.edges])
+        _G_hat.remove_edges_from([(i, i) for i in G_hat.nodes
+                                  if (i, i) in G_hat.edges])
+
+    pos = graphviz_layout(_G, prog="dot")
+
+    G_correct = nx.DiGraph()
+    G_correct.add_nodes_from(nodes)
+    G_correct.add_edges_from(set(_G.edges) & set(_G_hat.edges))
+
+    G_missed = nx.DiGraph()
+    G_missed.add_nodes_from(nodes)
+    G_missed.add_edges_from(set(_G.edges) - set(_G_hat.edges))
+
+    G_extra = nx.DiGraph()
+    G_extra.add_nodes_from(nodes)
+    G_extra.add_edges_from(set(_G_hat.edges) - set(_G.edges))
+
+    n_edges = len(_G.edges)
+    n_correct = len(G_correct.edges)
+    n_missed = len(G_missed.edges)
+    n_extra = len(G_extra.edges)
+    assert n_missed == n_edges - n_correct
+    
+
+    nx.draw(G_correct, pos, arrowsize=20,
+            with_labels=True, linewidth=2,
+            font_size=24, font_family="serif",
+            font_weight="bold",
+            alpha=0.9, node_color="lightblue",
+            edge_color="g")
+
+    nx.draw(G_missed, pos, arrowsize=20,
[…]

### 2019-06-13 software/pwgc/stat_util.py

@@ -1,0 +1,32 @@
+import scipy.stats as sps
+import numpy as np
+
+
+def benjamini_hochberg(P, alpha=0.05, independent=True):
+    """
+    alpha is the desired bound on the false discovery rate
+    (i.e. proportion of false positives against total positives.
+
+    m is the total number of hypotheses that we test, and P is an
+    array of P-values.
+
+    Use independent = False if the p-values are not independent of one
+    and other.
+
+    NOTE: P must be a 1-dimensional array of p-values, one for each test.
+
+    Return: The BH rejection threshold.  One should reject all P-values s.t.
+    P < th_bh.
+    """
+    P = np.sort(P)
+    m = len(P)
+
+    if independent:
+        C = 1
+    else:
+        C = sum(1. / k for k in range(1, m + 1))
+
+    ell = np.arange(1, m + 1) * alpha / (C * m)
+    
+    R = np.argmax(np.diff(P < ell))
+    return P[R]

### 2019-06-13 software/test/test_gc_methods.py

@@ -1,0 +1,95 @@
+import seaborn as sns
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+
+from matplotlib import rc as mpl_rc
+font = {"family" : "normal",
+        "weight" : "bold",
+        "size"   : 22}
+
+mpl_rc("font", **font)
+mpl_rc("text", usetex=False)
+
+from var_system import (random_gnp_dag, drive_gcg, get_errors,
+                        get_estimation_errors, random_tree_dag,
+                        get_X, attach_node_prop, gcg_to_var)
+from gc_methods import (compute_pairwise_gc, estimate_B, pw_scg,
+                        estimate_graph, full_filter_estimator,
+                        compute_covariances, compute_xi, fast_compute_xi,
+                        normalize_gc_score)
+from draw_graphs import draw_graph_estimates
+from stat_util import benjamini_hochberg
+
+
+def test_fast_computation():
+    """
+    Test the "fast" mode computations.
+
+    TODO: They appear to still produce terrible results for the
+    TODO: complete simulations even though the results are similar
+.   TODO: here.  The covariance estimates are frequently indefinite.
+    """
+    n_nodes, p_lags, max_lags = 75, 5, 10
+    random_graph = lambda: random_tree_dag(
+        n_nodes, p_lags, pole_rad=0.25)
+
+    T = 500
+    sv2_true = 0.5 + np.random.exponential(0.5, size=n_nodes)
+    G = random_graph()
+    drive_gcg(G, T, sv2_true, filter_attr="b(z)")
+    X = get_X(G)
+
+    xi_i_corr, xi_ij_corr, _, p_ij = compute_xi(X, max_lag=max_lags)
+
+    fig, axes = plt.subplots(2, 1, sharex=True, sharey=True)
+
+    for delta in [0, 1e-3, 1e-2]:
+        xi_i_delta, xi_ij_delta = fast_compute_xi(X, max_lag=max_lags,
+                                                  reg_delta=delta)
+
+        axes[0].scatter(xi_i_corr, xi_i_delta,
+                        label="delta = {}".format(delta))
+        axes[0].set_title("xi_i")
+
+        axes[1].scatter(xi_ij_corr, xi_ij_delta,
[…]

### 2019-06-13 software/pwgc/gc_methods.py
Some formatting.

@@ -84,7 +84,8 @@ def form_Xy(X_raw, y_raw, p=10):
 
     # Stack together lags of the same variables
     X = np.hstack(
-        [np.hstack([X_raw[p - tau: -tau, j][:, None] for tau in range(1, p + 1)])
+        [np.hstack([X_raw[p - tau: -tau, j][:, None]
+                    for tau in range(1, p + 1)])
          for j in range(s)])
 
     assert X.shape == (T - p, s * p), "Wrong _X shape!"
@@ -169,7 +170,8 @@ def estimate_B(G, max_lag=10, copy_G=False,
 
         # Add the filters as properties to G
         for grouping, j in enumerate(a_i):  # Edge j --b(z)--> i
-            G[j][i]["b_hat(z)"] = b[grouping * max_lag: (grouping + 1) * max_lag]
+            G[j][i]["b_hat(z)"] = b[grouping * max_lag:
+                                    (grouping + 1) * max_lag]
 
         # As well as the residuals
         G.nodes[i]["sv^2_hat"] = np.var(r)
@@ -199,8 +201,8 @@ def bivariate_AR_error(y, x, max_lag=10):
     bic = -np.infty
     T = len(x)
     for p in range(1, max_lag + 1):
-        X_, y_ =  form_Xy(X_raw=np.hstack((y[:, None], x[:, None])),
-                          y_raw=y, p=p)
+        X_, y_ = form_Xy(X_raw=np.hstack((y[:, None], x[:, None])),
+                         y_raw=y, p=p)
         sv2_p = _compute_AR_error(X_, y_)
         bic_p = -(T - p - 1) * np.log(sv2_p) - 2 * p * np.log(T - p - 1)
         if bic_p > bic:
@@ -221,7 +223,7 @@ def compute_covariances(X, p, symmetrize=False):
         [X.T @ X / T] +
         [X[tau:, :].T @ X[: -tau, :] / T
          for tau in range(1, p + 1)],
-         axis=0)
+        axis=0)
     return R
 
 
@@ -251,7 +253,7 @@ def compute_covariances(X, p, symmetrize=False):
 def _compute_AR_error(X, y):
     w = estimate_b_lstsqr(X, y)
     return np.var(y - X @ w)
-    
+
 
 def _fast_compute_AR_error(R, r):
     w = estimate_b_lstsqr_cov(R, r)
@@ -287,11 +289,13 @@ def compute_bic(eps, T, s=1):
            for p, eps_p in enumerate(eps)]
     return bic
 
+
[…]

### 2019-06-13 software/pwgc/test_gc_methods.py

@@ -1,12 +1,10 @@
-import seaborn as sns
-import matplotlib.pyplot as plt
-import pandas as pd
 import numpy as np
 
+import matplotlib.pyplot as plt
 from matplotlib import rc as mpl_rc
-font = {"family" : "normal",
-        "weight" : "bold",
-        "size"   : 22}
+font = {"family": "normal",
+        "weight": "bold",
+        "size": 22}
 
 mpl_rc("font", **font)
 mpl_rc("text", usetex=False)

### 2019-06-13 software/min_spanning.py
fast xi calculation and numba for gc_methods.

@@ -1,29 +1,0 @@
-import networkx as nx
-import numpy as np
-import matplotlib.pyplot as plt
-
-from networkx.drawing.nx_agraph import graphviz_layout
-
-
-def complete_digraph(n=10, wgt_f=lambda: np.random.standard_t(df=3)):
-    G = nx.DiGraph()
-    for i in range(n):
-        [G.add_edge(i, j, weight=wgt_f()) for j in range(n)
-         if j != i]
-    return G
-
-
-G = complete_digraph(n=25,
-                     wgt_f=lambda: np.random.standard_t(df=1))
-
-print("Searching for spanning arborescence")
-T = nx.minimum_spanning_arborescence(G, attr="weight",
-                                     default=np.inf)
-print("Done.")
-
-pos = graphviz_layout(G, prog="dot")
-nx.draw(T, pos=pos, node_color="b", edgelist=list(T.edges),
-        edge_color=[T.get_edge_data(u, v)["weight"] for u, v in T.edges],
-        width=2, edge_cmap=plt.cm.viridis)
-# plt.colorbar()
-plt.show()

### 2019-06-13 software/pwgc/__init__.py

@@ -1,0 +1,1 @@
+name = "pwgc"

### 2019-06-13 software/pwgc/gc_methods.py

@@ -4,6 +4,8 @@ Methods which implement granger-causality computations.
 
 import networkx as nx
 import numpy as np
+import numba
+from levinson import (lev_durb, whittle_lev_durb)
 
 from scipy import linalg
 from scipy import stats as sps
@@ -14,12 +16,17 @@ from itertools import product
 from functools import reduce
 from math import ceil
 
-from var_system import (attach_node_prop, add_self_loops,
-                        remove_zero_filters, get_X)
+# TODO: What is the proper way to do this?  I want packaging,
+# TODO: but I also want to be able to C-c C-c this file into
+# TODO: ipython while I am working...
 try:
     from .stat_util import benjamini_hochberg
+    from .var_system import (attach_node_prop, add_self_loops,
+                             remove_zero_filters, get_X)
 except ImportError:
     from stat_util import benjamini_hochberg
+    from var_system import (attach_node_prop, add_self_loops,
+                            remove_zero_filters, get_X)
 
 
 def estimate_b_lstsqr(X, y):
@@ -180,6 +187,7 @@ def estimate_B(G, max_lag=10, copy_G=False,
 
 
 # NOTE: forming the Xy matrices consumes only microseconds
+# NOTE: But is the majority of time spent...
 def univariate_AR_error(x, max_lag=10):
     # Try larger model orders until we don't increase the BIC
     bic = -np.infty
@@ -214,41 +222,23 @@ def bivariate_AR_error(y, x, max_lag=10):
     return sv2_p, p
 
 
-def compute_covariances(X, p, symmetrize=False):
-    # We do not do any normalizing
-    # I believe this should ensure positive-definite ness.
-    # We are effectively /windowing/ the signal.
-    T = X.shape[0]
-    R = np.stack(
-        [X.T @ X / T] +
-        [X[tau:, :].T @ X[: -tau, :] / T
-         for tau in range(1, p + 1)],
-        axis=0)
+@numba.jit(nopython=True, cache=True)
+def compute_covariances(X, p):
+    # Recall that we must ensure the result is positive definite --
+    # calculate the cov of a windowed signal.
[…]

### 2019-06-13 software/pwgc/test_gc_methods.py

@@ -90,4 +90,4 @@ def test_pw_scg():
     return
 
 
-test_pw_scg()
+# test_pw_scg()

### 2019-06-13 software/pwgc/var_system.py

@@ -1,8 +1,12 @@
 import numpy as np
 import networkx as nx
-from LSI_filters import random_arma
 from itertools import combinations
 from scipy.linalg import solve_discrete_lyapunov
+
+try:
+    from .LSI_filters import random_arma
+except ImportError:
+    from LSI_filters import random_arma
 
 # TODO: It would probably be wise to do this with scipy sparse matrices!
 

### 2019-06-14 software/pwgc/gc_methods.py
Trying to understand difference between levinson and lstsqr.

@@ -195,7 +195,7 @@ def univariate_AR_error(x, max_lag=10):
     for p in range(1, max_lag + 1):
         X_, y_ = form_Xy(X_raw=x[:, None], y_raw=x, p=p)
         sv2_p = _compute_AR_error(X_, y_)
-        bic_p = -(T - p - 1) * np.log(sv2_p) - p * np.log(T - p - 1)
+        bic_p = -T * np.log(sv2_p) - p * np.log(T)
         if bic_p > bic:
             bic = bic_p
             continue
@@ -212,7 +212,7 @@ def bivariate_AR_error(y, x, max_lag=10):
         X_, y_ = form_Xy(X_raw=np.hstack((y[:, None], x[:, None])),
                          y_raw=y, p=p)
         sv2_p = _compute_AR_error(X_, y_)
-        bic_p = -(T - p - 1) * np.log(sv2_p) - 2 * p * np.log(T - p - 1)
+        bic_p = -T * np.log(sv2_p) - 4 * p * np.log(T)
         if bic_p > bic:
             bic = bic_p
             continue
@@ -232,11 +232,6 @@ def compute_covariances(X, p):
     R[0] = X.T @ X
     for tau in range(1, p + 1):
         R[tau] = X[tau:, :].T @ X[:-tau, :]
-    # R = np.stack(
-    #     [X.T @ X / T] +
-    #     [X[tau:, :].T @ X[: -tau, :] / T
-    #      for tau in range(1, p + 1)],
-    #     axis=0)
     return R
 
 
@@ -278,17 +273,13 @@ def compute_bic(eps, T, s=1):
     bic = np.empty(len(eps))
     logT = np.log(T)
     for p, log_eps_p in enumerate(np.log(eps)):
-        bic[p] = -T * log_eps_p - s * p * logT
+        bic[p] = -log_eps_p - s * p * logT / T
     return bic
 
 
 def compute_gc_score(xi_i, xi_ij, T, p_lags):
-    assert T > 1 + 2 * np.max(p_lags),\
-        "T = {} is too small for max(p_lags) = {}".format(T, np.max(p_lags))
-    # TODO: Should this be T / p_j ??
-    # F is is chi2(p) under the null
     F = T * (xi_i[:, None] / xi_ij - 1) / p_lags
-    # F = F * (T - 2 * p_lags - 1) / p_lags
+    F[p_lags == 0] = 0
     return F
 
 
@@ -337,12 +328,13 @@ def fast_compute_xi(X, max_lag=10):
[…]

### 2019-06-14 software/pwgc/test_gc_methods.py

@@ -80,13 +80,21 @@ def test_pw_scg():
     X = get_X(G)
     X = X - np.mean(X, axis=0)[None, :]
 
-    # TODO: Use max_lags and compute lag length by bic
-    F, P = compute_pairwise_gc(X, max_lag=p_lags)
-    P_edges = normalize_gc_score(F, P)  # p-values are just 1 - F
-    P_values = 1 - P_edges[~np.eye(n_nodes, dtype=bool)].ravel()
-    t_bh = benjamini_hochberg(P_values, alpha=alpha, independent=False)
-    G_hat = pw_scg(F, P_edges, t_bh)
-    draw_graph_estimates(G, G_hat)
+    def est_graph(X, fast=False):
+        if fast:
+            F, P = fast_compute_pairwise_gc(X, p_lags)
+        else:
+            F, P = compute_pairwise_gc(X, max_lag=p_lags)
+
+        P_edges = normalize_gc_score(F, P)  # p-values are just 1 - F
+        P_values = 1 - P_edges[~np.eye(n_nodes, dtype=bool)].ravel()
+        t_bh = benjamini_hochberg(P_values, alpha=alpha, independent=False)
+        G_hat = pw_scg(F, P_edges, t_bh)
+        draw_graph_estimates(G, G_hat)
+        return
+
+    est_graph(X, False)
+    est_graph(X, True)
     return
 
 

### 2019-06-14 software/pwgc/draw_graphs.py
Simulations and nicer figures.

@@ -3,7 +3,8 @@ import matplotlib.pyplot as plt
 
 from networkx.drawing.nx_agraph import write_dot, graphviz_layout
 
-from var_system import (random_gnp_dag, random_tree_dag)
+from .var_system import (random_gnp_dag, random_scg)
+
 
 def draw_graph(G, file_name=None):
     pos = graphviz_layout(G, prog="dot")
@@ -93,23 +94,4 @@ def draw_graph_estimates(G, G_hat, file_name=None,
         else:  # Allow passing a list
             [plt.savefig(f) for f in file_name]
     plt.show()
-    return
-
-
-def draw_example_graphs():
-    n_nodes = 75
-    edge_prob = 2. / n_nodes
-
-    G_tree = random_tree_dag(n_nodes, p_lags, pole_rad=0.75)
-    G_dag = random_gnp_dag(n_nodes, p_lags, pole_rad=0.75,
-                           edge_prob=edge_prob)
-    G_gnp = random_gnp(n_nodes, p_lags, pole_rad=0.75,
-                       edge_prob=edge_prob)
-
-    draw_graph(G_tree, ["../figures/example_tree.pdf",
-                        "../figures/jpgs_pngs/example_tree.png"])
-    draw_graph(G_dag, ["../figures/example_dag.pdf",
-                       "../figures/jpgs_pngs/example_dag.png"])
-    draw_graph(G_gnp, ["../figures/example_gnp.pdf",
-                       "../figures/jpgs_pngs/example_gnp.png"])
     return

### 2019-06-14 software/pwgc/gc_methods.py

@@ -278,7 +278,7 @@ def compute_bic(eps, T, s=1):
 
 
 def compute_gc_score(xi_i, xi_ij, T, p_lags):
-    F = T * (xi_i[:, None] / xi_ij - 1) / p_lags
+    F = T * ((xi_i[:, None] / xi_ij) - 1) / p_lags
     F[p_lags == 0] = 0
     return F
 
@@ -624,7 +624,8 @@ def estimate_graph(X, G, max_lags=10, method="lasso", alpha=0.05):
     T, n = X.shape
 
     # Compute the pairwise errors and filter sizes
-    F, P = compute_pairwise_gc(X, max_lag=max_lags)
+    # F, P = compute_pairwise_gc(X, max_lag=max_lags)
+    F, P = fast_compute_pairwise_gc(X, max_lag=max_lags)
 
     # Screen edges via benjamini hochberg criterion
     P_edges = normalize_gc_score(F, P)  # p-values are just 1 - F

### 2019-06-14 software/pwgc/test_gc_methods.py

@@ -65,10 +65,10 @@ def test_fast_computation():
 
 
 def test_pw_scg():
-    n_nodes, p_lags, max_lags = 50, 5, 10
+    n_nodes, p_lags, p_max = 50, 5, 15
     alpha = 0.05
 
-    T_max = 500
+    T_max = 5000
 
     random_graph = lambda: random_tree_dag(
         n_nodes, p_lags, pole_rad=0.75)

### 2019-06-14 software/pwgc/var_system.py

@@ -71,6 +71,7 @@ def _graph_to_dag(G):
     n_nodes = len(G.nodes)
     G = nx.DiGraph([(u, v)
                     for (u, v) in G.edges() if u < v])
+
     [G.add_node(v) for v in set(range(n_nodes)) - set(G.nodes)]    
     assert nx.is_directed_acyclic_graph(G)
     return G
@@ -126,23 +127,21 @@ def random_gnp(n_nodes, p_lags, edge_prob=0.3, pole_rad=0.95):
         return "b(z)", -random_arma(p=p_lags, q=0, k=1, p_radius=pole_rad)[1][1:]
 
     # Add self loops only with probability edge_prob
-    self_loops = np.random.binomial(n=1, p=edge_prob,
-                                    size=len(G.nodes))
-    for add, i in zip(self_loops, G.nodes):
-        if add:
-            G.add_edge(i, i)
+    G = add_self_loops(G, copy_G=False)
     G = _attach_edge_properties(G, get_edge)
+
+    stabilize_system(G)
     
     G.graph["true_lags"] = p_lags
     return G
 
 
-def random_tree_dag(n_nodes, p_lags, pole_rad=0.95):
+def random_scg(n_nodes, p_lags, pole_rad=0.95):
     """
     I THINK STRONGLY CAUSAL
 
-    Produces a random directed tree with random ARMA filters having
-    pole radius `pole_rad` on the edges.  The tree will have `n_nodes`
+    Produces a random strongly causal graph with random ARMA filters having
+    pole radius `pole_rad` on the edges.  The graph will have `n_nodes`
     nodes and each node will have `r_degree` children.
 
     # NOTE: This will always have the same number of edges (why?)
@@ -293,6 +292,11 @@ def drive_gcg(G, T, sigma2_v, filter_attr="b(z)"):
         G.nodes[node_i]["x"] = X[:, i]
         G.nodes[node_i]["sv2"] = sigma2_v[i]
     return G
+
+
+def stabilize_system(G):
+    nx.simple_cycles(G)
+    return
 
 
 class VAR:

### 2019-06-14 software/simulations.py

@@ -3,6 +3,10 @@ import matplotlib.pyplot as plt
 import pandas as pd
 import numpy as np
 
+import statsmodels.api as sm
+from statsmodels.sandbox.regression.predstd import wls_prediction_std
+
+
 from matplotlib import rc as mpl_rc
 font = {"family" : "normal",
         "weight" : "bold",
@@ -11,33 +15,39 @@ font = {"family" : "normal",
 mpl_rc("font", **font)
 mpl_rc("text", usetex=True)
 
-from var_system import (random_gnp_dag, drive_gcg, get_errors,
-                        get_estimation_errors, random_tree_dag,
-                        get_X, attach_node_prop, gcg_to_var)
-from gc_methods import (compute_pairwise_gc, estimate_B, pw_scg,
-                        estimate_graph, full_filter_estimator,
-                        compute_MCC_score)
+from pwgc.var_system import (random_gnp_dag, drive_gcg, get_errors,
+                             get_estimation_errors, random_scg,
+                             get_X, attach_node_prop, gcg_to_var)
+from pwgc.gc_methods import (compute_pairwise_gc, estimate_B, pw_scg,
+                             estimate_graph, full_filter_estimator,
+                             compute_MCC_score)
 
 
-def full_single_pass_experiment(simulation_name, graph_type, M_passes=1):
-    n_nodes, p_lags, max_lags = 50, 5, 10
-    N_iters = 50
+def full_single_pass_experiment(simulation_name, graph_type):
+    np.random.seed(0)
+    n_nodes, p_lags, p_max = 50, 5, 20
+    alpha, N_iters = 0.05, 3
 
-    T_max = 11000
-    # T_iters = [100, 500, 1000, 2500, 5000, 10000]
-    T_iters = [100, 500, 1000, 2500, 10000]
+    T_iters = [50 + 50 * k for k in range(1, 99)]
+    T_max = T_iters[-1] + 100
 
-    if graph_type == "Tree":  # TODO: This doesn't actually produce trees!
-        random_graph = lambda: random_tree_dag(
+    # Tune the expected number of edges to match N_edges
+    N_edges = (len(random_scg(n_nodes, p_lags, pole_rad=0.75).edges) -
+               n_nodes)  # Due to self loops
+    edge_prob = 2 * N_edges / (n_nodes * (n_nodes - 1))
+
+    if graph_type == "SCG":
+        random_graph = lambda: random_scg(
[…]

### 2019-06-14 software/examples.py
Graph example file.

@@ -1,0 +1,29 @@
+"""
+Generates some example graph topologies to include in paper.
+"""
+
+from pwgc.var_system import (random_scg, random_gnp_dag, random_gnp)
+from pwgc.draw_graphs import draw_graph
+
+
+def draw_example_graphs():
+    n_nodes, p_lags = 75, 5
+    edge_prob = 2. / n_nodes
+
+    G_scg = random_scg(n_nodes, p_lags, pole_rad=0.75)
+    G_dag = random_gnp_dag(n_nodes, p_lags, pole_rad=0.75,
+                           edge_prob=edge_prob)
+    G_gnp = random_gnp(n_nodes, p_lags, pole_rad=0.75,
+                       edge_prob=edge_prob)
+
+    draw_graph(G_scg, ["../figures/example_scg.pdf",
+                        "../figures/jpgs_pngs/example_tree.png"])
+    draw_graph(G_dag, ["../figures/example_dag.pdf",
+                       "../figures/jpgs_pngs/example_dag.png"])
+    draw_graph(G_gnp, ["../figures/example_gnp.pdf",
+                       "../figures/jpgs_pngs/example_gnp.png"])
+    return
+
+
+if __name__ == "__main__":
+    draw_example_graphs()

### 2019-06-15 software/pwgc/gc_methods.py
Minor updates to how errors are computed.

@@ -173,7 +173,7 @@ def estimate_B(G, max_lag=10, copy_G=False,
             raise AssertionError("Bad method but should deal with it earlier!")
 
         # Compute residuals
-        r = y - X @ b
+        r = y[max_T:] - X[max_T:] @ b
 
         # Add the filters as properties to G
         for grouping, j in enumerate(a_i):  # Edge j --b(z)--> i

### 2019-06-15 software/simulations.py

@@ -29,7 +29,12 @@ def full_single_pass_experiment(simulation_name, graph_type):
     alpha, N_iters = 0.05, 3
 
     T_iters = [50 + 50 * k for k in range(1, 99)]
-    T_max = T_iters[-1] + 100
+
+    # NOTE: We use only X[T:] to estimate the error
+    # NOTE: so it is important to have a significant trailing
+    # NOTE: set of data in order to make honest estimates of
+    # NOTE: the out of sample predictions.
+    T_max = T_iters[-1] * 2
 
     # Tune the expected number of edges to match N_edges
     N_edges = (len(random_scg(n_nodes, p_lags, pole_rad=0.75).edges) -

### 2019-06-15 software/pwgc/draw_graphs.py
method to produce complete digraph.

@@ -3,7 +3,10 @@ import matplotlib.pyplot as plt
 
 from networkx.drawing.nx_agraph import write_dot, graphviz_layout
 
-from .var_system import (random_gnp_dag, random_scg)
+try:
+    from .var_system import (random_gnp_dag, random_scg)
+except ImportError:
+    from var_system import (random_gnp_dag, random_scg)
 
 
 def draw_graph(G, file_name=None):

### 2019-06-15 software/pwgc/gc_methods.py

@@ -611,6 +611,13 @@ def get_residual_graph(G_hat):
     return G_hat
 
 
+def estimate_dense_graph(X, max_T=None, method="lasso"):
+    T, n = X.shape
+    G = nx.complete_graph(n)
+    G = attach_node_prop(
+    
+    return
+
 def estimate_graph(X, G, max_lags=10, method="lasso", alpha=0.05):
     """
     Produce an estimated graph from the data X.
@@ -632,31 +639,22 @@ def estimate_graph(X, G, max_lags=10, method="lasso", alpha=0.05):
     P_values = 1 - P_edges[~np.eye(n, dtype=bool)].ravel()
     t_bh = benjamini_hochberg(P_values, alpha=alpha, independent=False)
 
-    # F = fast_compute_pairwise_gc(X, max_lag=max_lags)
-
     # Estimate a strongly causal graph
     G_hat = pw_scg(F, P_edges, t_bh)
-    # draw_graph_estimates(G, G_hat)
     G_hat = attach_node_prop(G_hat, G, prop_attach="x", prop_from="x")
     G_hat = add_self_loops(G_hat, copy_G=False)
-    # draw_graph_estimates(G, G_hat)
 
-    # Restimates the full graph via the LASSO procedure -- this tends
-    # to remove many of the false positive edges, but also at the expense
-    # of true positives.
-    # TODO: Improve upon this step
-
+    # Restimate the coefficients
     if method == "lasso":
-        G_hat = estimate_B(G_hat, max_lags, copy_G=False, max_T=X.shape[0],
+        G_hat = estimate_B(G_hat, max_lags, copy_G=False, max_T=T,
                            method=method)
         G_hat = remove_zero_filters(G_hat, "b_hat(z)", copy_G=False)
     elif method == "lstsqr":
-        G_hat = estimate_B(G_hat, max_lags, copy_G=False, max_T=X.shape[0],
+        G_hat = estimate_B(G_hat, max_lags, copy_G=False, max_T=T,
                            method=method)
     else:
         raise NotImplementedError("The fitting method {} is not available!"
                                   "".format(method))
-    # draw_graph_estimates(G, G_hat)
     return G_hat
 
 

### 2019-06-15 software/pwgc/var_system.py

@@ -72,7 +72,7 @@ def _graph_to_dag(G):
     G = nx.DiGraph([(u, v)
                     for (u, v) in G.edges() if u < v])
 
-    [G.add_node(v) for v in set(range(n_nodes)) - set(G.nodes)]    
+    _ = [G.add_node(v) for v in set(range(n_nodes)) - set(G.nodes)]    
     assert nx.is_directed_acyclic_graph(G)
     return G
 
@@ -85,6 +85,14 @@ def _attach_edge_properties(G, gen_func):
     for u, v in G.edges:
         prop, value = gen_func()
         G[u][v][prop] = value
+    return G
+
+
+def make_complete_digraph(n_nodes):
+    G = nx.complete_graph(n_nodes)
+    G = _graph_to_dag(G)
+    G.add_edges_from([(v, u) for (u, v) in G.edges])
+    G = add_self_loops(G)
     return G
 
 
@@ -101,12 +109,13 @@ def random_gnp_dag(n_nodes, p_lags, edge_prob=0.3, pole_rad=0.95):
     G = nx.gnp_random_graph(n_nodes, p=edge_prob, directed=True)
 
     def get_edge():
-        return "b(z)", -random_arma(p=p_lags, q=0, k=1, p_radius=pole_rad)[1][1:]
+        return "b(z)", -random_arma(p=p_lags, q=0,
+                                    k=1, p_radius=pole_rad)[1][1:]
 
     G = _graph_to_dag(G)
     G = add_self_loops(G, copy_G=False)
     G = _attach_edge_properties(G, get_edge)
-    
+
     G.graph["true_lags"] = p_lags
     return G
 
@@ -124,14 +133,15 @@ def random_gnp(n_nodes, p_lags, edge_prob=0.3, pole_rad=0.95):
     G = nx.gnp_random_graph(n_nodes, p=edge_prob, directed=True)
 
     def get_edge():
-        return "b(z)", -random_arma(p=p_lags, q=0, k=1, p_radius=pole_rad)[1][1:]
+        return "b(z)", -random_arma(p=p_lags, q=0,
+                                    k=1, p_radius=pole_rad)[1][1:]
 
     # Add self loops only with probability edge_prob
     G = add_self_loops(G, copy_G=False)
     G = _attach_edge_properties(G, get_edge)
 
     stabilize_system(G)
-    
+
     G.graph["true_lags"] = p_lags
     return G
 
[…]

### 2019-06-15 software/simulations.py

@@ -83,12 +83,6 @@ def full_single_pass_experiment(simulation_name, graph_type):
             # TODO: I am using T_max samples on every run!
             G_hat = estimate_graph(X, G, max_lags=p_max,
                                    method="lstsqr", alpha=alpha)
-            # draw_graph_estimates(G, G_hat)
-            # G_hat = full_filter_estimator(G, M_passes=1, T_max=T)
-            # X = get_X(G, prop="x")
-
-            # # We pass in G only to attach "x" to G_hat
-            # G_hat = estimate_graph(X[:T], G)
 
             b_err = np.mean(list(map(np.var, get_estimation_errors(G, G_hat))))
             err = np.sum(get_errors(G_hat))

### 2019-06-16 software/pwgc/gc_methods.py
Correct a syntax error!

@@ -621,7 +621,7 @@ def estimate_dense_graph(X, max_lag=10,
     assert np.all(X == get_X(G))
 
     estimate_B(G, max_lag, copy_G=copy_G, max_T=max_T, method=method)
-    G = remove_zero_filters(G), "b_hat(z)", copy_G=False)
+    G = remove_zero_filters(G, "b_hat(z)", copy_G=False)
     return G
 
 

### 2019-06-16 software/pwgc/gc_methods.py
Fix another syntax issue.

@@ -620,7 +620,7 @@ def estimate_dense_graph(X, max_lag=10,
     G = attach_X(G, X)
     assert np.all(X == get_X(G))
 
-    estimate_B(G, max_lag, copy_G=copy_G, max_T=max_T, method=method)
+    estimate_B(G, max_lag, copy_G=False, max_T=max_T, method=method)
     G = remove_zero_filters(G, "b_hat(z)", copy_G=False)
     return G
 

### 2019-06-16 software/simulations.py

@@ -133,9 +133,14 @@ def full_single_pass_experiment(simulation_name, graph_type):
             print("N[{} / {}] -- T[{} / {}]\r".format(
                 N_iter + 1, N_iters, T_iter + 1, len(T_iters)))
 
+            # ~252ms
             G, X, sv2_true = make_test_data(T)
+
+            # ~824ms
             G_hat_pwgc = estimate_graph(X, G, max_lags=p_max,
                                         method="lstsqr", alpha=alpha)
+
+            # ~
             G_hat_lasso = estimate_dense_graph(X, max_lag=p_max,
                                                max_T=T,
                                                method="lasso")

### 2019-06-16 software/pwgc/gc_methods.py
Cleaned up some warnings.

@@ -174,6 +174,14 @@ def estimate_B(G, max_lag=10, copy_G=False,
         else:
             raise AssertionError("Bad method but should deal with it earlier!")
 
+        if len(y[max_T:]) <= 100:
+            raise ValueError(
+                "len(y) = {} is not long "
+                "enough for using max_T = {} " "samples for estimation and "
+                "the remaining for the " "out of sample error calculation. "
+                "Ensure that we have " "max_T <= len(y) - 100."
+                "".format(len(y), max_T))
+
         # Compute residuals
         r = y[max_T:] - X[max_T:] @ b
 
@@ -280,8 +288,9 @@ def compute_bic(eps, T, s=1):
 
 
 def compute_gc_score(xi_i, xi_ij, T, p_lags):
-    F = T * ((xi_i[:, None] / xi_ij) - 1) / p_lags
-    F[p_lags == 0] = 0
+    with np.errstate(divide="ignore", invalid="ignore"):
+        F = T * ((xi_i[:, None] / xi_ij) - 1) / p_lags
+    F[p_lags ==0] = 0
     return F
 
 
@@ -364,6 +373,7 @@ def compute_pairwise_gc(X, max_lag=10):
 
 def normalize_gc_score(F, p):
     F = sps.chi2.cdf(F, p)
+    F[np.isnan(F)] = 0.0
     return F
 
 

### 2019-06-16 software/simulations.py

@@ -83,7 +83,7 @@ class TrackErrors:
 
 def full_single_pass_experiment(simulation_name, graph_type):
     np.random.seed(0)
-    n_nodes, p_lags, p_max = 50, 5, 20
+    n_nodes, p_lags, p_max = 50, 5, 15
     alpha, N_iters = 0.05, 3
 
     T_iters = [50 + 50 * k for k in range(1, 99)]
@@ -115,13 +115,13 @@ def full_single_pass_experiment(simulation_name, graph_type):
         raise NotImplementedError("graph_type '{}' is not understood"
                                   "".format(graph_type))
 
-    def make_test_data(T):
+    def make_test_data():
         sv2_true = 0.5 + np.random.exponential(0.5, size=n_nodes)
         G = random_graph()
         drive_gcg(G, T_max, sv2_true, filter_attr="b(z)")
         X = get_X(G)
         X = X - np.mean(X, axis=0)[None, :]
-        X = X[:T]
+        X = X
         return G, X, sv2_true
 
     errs_pwgc = TrackErrors(N_iters, T_iters, n_nodes)
@@ -134,16 +134,17 @@ def full_single_pass_experiment(simulation_name, graph_type):
                 N_iter + 1, N_iters, T_iter + 1, len(T_iters)))
 
             # ~252ms
-            G, X, sv2_true = make_test_data(T)
+            G, X, sv2_true = make_test_data()
 
             # ~824ms
-            G_hat_pwgc = estimate_graph(X, G, max_lags=p_max,
+            G_hat_pwgc = estimate_graph(X[:T], G, max_lags=p_max,
                                         method="lstsqr", alpha=alpha)
 
-            # ~
-            G_hat_lasso = estimate_dense_graph(X, max_lag=p_max,
-                                               max_T=T,
-                                               method="lasso")
+            # ~37s
+            if N_iter == 0:
+                G_hat_lasso = estimate_dense_graph(X, max_lag=p_max,
+                                                   max_T=T,
+                                                   method="lasso")
 
             errs_pwgc.update(G, G_hat_pwgc, sv2_true, N_iter, T_iter)
             errs_lasso.update(G, G_hat_lasso, sv2_true, N_iter, T_iter)

### 2019-06-17 paper/paper.tex
Simulations and nicer figures.
Comparison simulations for LASSO and PWGC.

@@
 even in the case of time series data \cite{basu2015}
-\cite{wong2016lasso}.
+\cite{wong2016lasso}, variations on which have been applied elsewhere
+\cite{DBLP:journals/corr/HallacPBL17} \cite{haufe2008sparse}
+\cite{bolstad2011causal} \cite{he2013stationary}.
 
@@
 (section \ref{sec:pairwise_algorithm}).  Secondly, we propose in
-section \ref{sec:structure_learning} alternative graph search
-heuristics inspired by our theoretical results.  Our heuristics are
-compared against a standard grouped LASSO algorithm
-\cite{yuan2006model} as well as local search hill climbing (\hl{state
-  here the results}).  In section \ref{sec:application} we develop an
-example application (\hl{more detail...}).  And provide concluding
-remarks on further open problems in our conclusion (section
-\ref{sec:conclusion}).
-
-%% -Can we also formulate an SDP relaxation?
-%% -Fit trees jointly or via coordinate descent?
+section \ref{sec:structure_learning} a graph search heuristic which
+implements our theoretical results to finite data samples.  Our
+heuristics are compared against the LASSO algorithm in Section
+\ref{sec:empirical_evaluation}.  We stress the scalability of our
+algorithm which is capable of comfortably handling hundreds or
+thousands of nodes on a single machine.  In section
+\ref{sec:application} we develop an example application (\hl{more
+  detail...}).  And provide concluding remarks on further open
+problems in our conclusion (section \ref{sec:conclusion}).
+
+% \cite{yuan2006model} as well as local search hill
+% climbing (\hl{state here the results}).
 
@@
     \item{$\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$}
-    \item{$\linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}}\ \forall \beta \in \mathcal{B}.$}
+    \item{$\linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}}\ \forall \beta \in \mathcal{B}$.}
     % \item{$\linE{\beta}{\mathcal{A}} = \linE{\linE{\beta}{\mathcal{X}}}{\mathcal{A}}\ \forall \beta \in \mathcal{B}}$}
@@
 $x_{ij}^{(p)}$, in which case the BIC, which we seek to
-\textit{maximize} over $p$ is given by:
+\textit{minimize} over $p$ is given by:
 
@@
   \begin{aligned}
-    BIC_{\text{univariate}}(p) &= -(T - p - 1) \ln \xi_i(p) - p\ln T,\\
-    BIC_{\text{bivariate}}(p) &= -(T - p - 1) \ln \xi_{ij}(p) - 2p\ln T.\\
+    BIC_{\text{univariate}}(p) &= T \ln \xi_i(p) + p\ln T,\\
+    BIC_{\text{bivariate}}(p) &= T \ln \xi_{ij}(p) + 2p\ln T.\\
   \end{aligned}
@@
 
-\section{Structure Learning}
-\label{sec:structure_learning}
-\subsection{Local Search Heuristics}
-
-\subsection{Edge-Wise Grouped LASSO}
-Minimize the sparsity promoting regularized least squares problem and choose the hyper-parameters against the Akaike information criteria.
-
-\begin{equation}
-  L(\lambda, p) \defeq \underset{B}{\text{minimize}}\ \frac{1}{T}\sum_{t = 1}^T||x(t) - \sum_{\tau = 1}^pB(\tau)x(t - \tau)||_2^2 + \lambda \sum_{i, j}\big[\alpha||B_{i, j}||_2 + (1 - \alpha)||B_{i, j}||_1\big]
-\end{equation}
-
-\begin{equation}
-  \underset{\lambda, p}{\text{minimize}}\ L(\lambda, p) + \mathsf{AIC}(B^{(\lambda, p)})
-\end{equation}
+% \section{Structure Learning}
+% \label{sec:structure_learning}
+% \subsection{Local Search Heuristics}
+
+% \subsection{Edge-Wise Grouped LASSO}
+% Minimize the sparsity promoting regularized least squares problem and choose the hyper-parameters against the Akaike information criteria.
+
+% \begin{equation}
+%   L(\lambda, p) \defeq \underset{B}{\text{minimize}}\ \frac{1}{T}\sum_{t = 1}^T||x(t) - \sum_{\tau = 1}^pB(\tau)x(t - \tau)||_2^2 + \lambda \sum_{i, j}\big[\alpha||B_{i, j}||_2 + (1 - \alpha)||B_{i, j}||_1\big]
+% \end{equation}
+
+% \begin{equation}
+%   \underset{\lambda, p}{\text{minimize}}\ L(\lambda, p) + \mathsf{AIC}(B^{(\lambda, p)})
+% \end{equation}
 
@@
 
-\subsection{Layered Strongly Causal Graphs}
-It is clearn that Algorithm \ref{alg:pwgr} is a purely theoretical construct, and cannot be applied directly to real data.
+% \subsection{Layered Strongly Causal Graphs}
+% It is clearn that Algorithm \ref{alg:pwgr} is a purely theoretical construct, and cannot be applied directly to real data.
 Instead, we convert the algorithm into an heuristic and observe it's performance experimentally.
@@
 
-\begin{algorithm}
-  \SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{output}
-  \SetKwInOut{Initialize}{initialize}
-  \DontPrintSemicolon
-
-  \BlankLine
-  \caption{Graph Recovery Heuristic $\mathsf{pw\_scg}(F, P, \alpha)$}
-  \label{alg:pwgr_heuristic}
-  \Input{Estimates of pairwise Granger-causaltiy statistics $F_{ij}$ (eqn \ref{eqn:gc_statistics}), Matrix of edge probabilities $P_{ij}$ (eqn \ref{eqn:edge_inclusion_probability}), threshold $\delta$.}
-  \Output{A strongly causal graph $\widehat{\gcg}$}
-  \Initialize{$S = [n]$  \texttt{\# unprocessed nodes}\\
-    $E = \emptyset$  \texttt{\# edges of }$\widehat{\gcg}$\\
-    $k = 1$ \texttt{\# a counter used only for notation}}
-
-  \BlankLine
-
-  $W_\delta \leftarrow \{(i, j)\ |\ P_{ji} > 1 - \delta, F_{ji} > F_{ij}\}$  \texttt{\# candidate edges}\\
-  $\mathcal{I}_0 \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# total node incident probability}\\
-  $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) < \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Nodes with fewest incident edges}\\
-  \If{$P_0 = \emptyset$}{
-    $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) \le \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Ensure non-empty}
-  }
-  \BlankLine
-
-  \While{$S \ne \emptyset$}{
-    $S \leftarrow S \setminus P_{k - 1}$ \texttt{\# remove processed nodes}\\
-    % $\mathcal{I}_k \leftarrow \big(\sum_{j \in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# intra-}$S$ \texttt{incident strength}\\
-    $\mathcal{I}_k \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$\\
-    $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) < \ceil{\text{min}(\mathcal{I}_k)}\}$\\
-    \If{$P_k = \emptyset$}{
-      $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) \le \ceil{\text{min}(\mathcal{I}_k)}\}$
-    }
-    \;
-    \texttt{\# add strongest edges, maintaining strong causality}\\
-    $U_k \leftarrow \bigcup_{r = 1}^k P_{k - r}$ \texttt{\# Include all forward edges}\\
-    \For{$(i, j) \in \mathsf{sort}\Big(\{(i, j) \in U_k \times P_k\ |\ (i, j) \in W_\delta\} \mathsf{\ by\ descending\ } F_{ji}\Big)$} {
-      \If{$\mathsf{is\_strongly\_causal}(E \cup \{(i, j)\})$} {
-        $E \leftarrow E \cup \{(i, j)\}$
-      }
-    }
-    $k \leftarrow k + 1$\\
-  }
-  \Return{$([n], E)$}
-\end{algorithm}
-
-\begin{algorithm}
-  \SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{output}
-  \SetKwInOut{Initialize}{initialize}
-  \DontPrintSemicolon
-
-  \BlankLine
-  \caption{Filter Estimator}
-  \label{alg:Bz_estimate}
-  \Input{$T$ samples from a WSS process $x: \{x(t)\}_{t = 1}^T$\\
-    Maximum iterations $M$\\
-    Sparsity parameters $\{(\delta_m\, b_m, R_m)\}_{m = 1}^M$}
-  \Output{Linear one-step-ahead estimator $\widehat{\B}(z)$}
-  \Initialize{$\epsilon^{(0)}(t) = x(t)$ \texttt{\# initial error process}\\
-    $m = 1$ \texttt{\# iteration counter}}
-  \BlankLine
-  \For{$m = 1, \ldots, M$}{
-    $F \leftarrow \mathsf{pwgc}(\epsilon^{(m - 1)}(t))$ \texttt{\# estimate pairwise Granger-causality statistics}\\
-    $\widehat{\gcg}_m \leftarrow \mathsf{pw\_scg}(F, \delta_m, b_m, R_m)$ \texttt{\# graph estimate (Algorithm \ref{alg:pwgr_heuristic})}\\
-    $\widehat{\B}^{(m)}(z) \leftarrow \mathsf{estimate\_B}(\epsilon^{(m - 1)}(t), \widehat{\gcg}_m)$ \texttt{\# Estimate filter with topology} $\widehat{\gcg}_m$\\
-    $\epsilon^{(m)}(t) \leftarrow \epsilon^{(m - 1)}(t) - \widehat{\B}^{(m)}(z)\epsilon^{(m - 1)}(t)$ \texttt{\# updated error process}\\
-  }
-  \BlankLine
-  $\widehat{\B}(z) \leftarrow \widehat{\B}^{(M)}(z) \prod_{k = 1}^{M - 1}\big(I - \widehat{\B}^{(M - k)}(z)\big)$ \texttt{\# compute total filter}\\
-  \Return{$\widehat{\B}(z)$}
-\end{algorithm}
+% \begin{algorithm}
+%   \SetKwInOut{Input}{input}
+%   \SetKwInOut{Output}{output}
+%   \SetKwInOut{Initialize}{initialize}
+%   \DontPrintSemicolon
+
+%   \BlankLine
+%   \caption{Graph Recovery Heuristic $\mathsf{pw\_scg}(F, P, \alpha)$}
+%   \label{alg:pwgr_heuristic}
+%   \Input{Estimates of pairwise Granger-causaltiy statistics $F_{ij}$ (eqn \ref{eqn:gc_statistics}), Matrix of edge probabilities $P_{ij}$ (eqn \ref{eqn:edge_inclusion_probability}), threshold $\delta$.}
+%   \Output{A strongly causal graph $\widehat{\gcg}$}
+%   \Initialize{$S = [n]$  \texttt{\# unprocessed nodes}\\
+%     $E = \emptyset$  \texttt{\# edges of }$\widehat{\gcg}$\\
+%     $k = 1$ \texttt{\# a counter used only for notation}}
+
+%   \BlankLine
+
+%   $W_\delta \leftarrow \{(i, j)\ |\ P_{ji} > 1 - \delta, F_{ji} > F_{ij}\}$  \texttt{\# candidate edges}\\
+%   $\mathcal{I}_0 \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# total node incident probability}\\
+%   $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) < \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Nodes with fewest incident edges}\\
+%   \If{$P_0 = \emptyset$}{
+%     $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) \le \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Ensure non-empty}
+%   }
+%   \BlankLine
+
+%   \While{$S \ne \emptyset$}{
+%     $S \leftarrow S \setminus P_{k - 1}$ \texttt{\# remove processed nodes}\\
+%     % $\mathcal{I}_k \leftarrow \big(\sum_{j \in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# intra-}$S$ \texttt{incident strength}\\
+%     $\mathcal{I}_k \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$\\
+%     $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) < \ceil{\text{min}(\mathcal{I}_k)}\}$\\
+%     \If{$P_k = \emptyset$}{
+%       $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) \le \ceil{\text{min}(\mathcal{I}_k)}\}$
+%     }
+%     \;
+%     \texttt{\# add strongest edges, maintaining strong causality}\\
+%     $U_k \leftarrow \bigcup_{r = 1}^k P_{k - r}$ \texttt{\# Include all forward edges}\\
+%     \For{$(i, j) \in \mathsf{sort}\Big(\{(i, j) \in U_k \times P_k\ |\ (i, j) \in W_\delta\} \mathsf{\ by\ descending\ } F_{ji}\Big)$} {
+%       \If{$\mathsf{is\_strongly\_causal}(E \cup \{(i, j)\})$} {
+%         $E \leftarrow E \cup \{(i, j)\}$
+%       }
+%     }
+%     $k \leftarrow k + 1$\\
+%   }
+%   \Return{$([n], E)$}
+% \end{algorithm}
+
+% \begin{algorithm}
+%   \SetKwInOut{Input}{input}
+%   \SetKwInOut{Output}{output}
+%   \SetKwInOut{Initialize}{initialize}
+%   \DontPrintSemicolon
+
+%   \BlankLine
+%   \caption{Filter Estimator}
+%   \label{alg:Bz_estimate}
+%   \Input{$T$ samples from a WSS process $x: \{x(t)\}_{t = 1}^T$\\
+%     Maximum iterations $M$\\
+%     Sparsity parameters $\{(\delta_m\, b_m, R_m)\}_{m = 1}^M$}
+%   \Output{Linear one-step-ahead estimator $\widehat{\B}(z)$}
+%   \Initialize{$\epsilon^{(0)}(t) = x(t)$ \texttt{\# initial error process}\\
+%     $m = 1$ \texttt{\# iteration counter}}
+%   \BlankLine
+%   \For{$m = 1, \ldots, M$}{
+%     $F \leftarrow \mathsf{pwgc}(\epsilon^{(m - 1)}(t))$ \texttt{\# estimate pairwise Granger-causality statistics}\\
+%     $\widehat{\gcg}_m \leftarrow \mathsf{pw\_scg}(F, \delta_m, b_m, R_m)$ \texttt{\# graph estimate (Algorithm \ref{alg:pwgr_heuristic})}\\
+%     $\widehat{\B}^{(m)}(z) \leftarrow \mathsf{estimate\_B}(\epsilon^{(m - 1)}(t), \widehat{\gcg}_m)$ \texttt{\# Estimate filter with topology} $\widehat{\gcg}_m$\\
+%     $\epsilon^{(m)}(t) \leftarrow \epsilon^{(m - 1)}(t) - \widehat{\B}^{(m)}(z)\epsilon^{(m - 1)}(t)$ \texttt{\# updated error process}\\
+%   }
+%   \BlankLine
+%   $\widehat{\B}(z) \leftarrow \widehat{\B}^{(M)}(z) \prod_{k = 1}^{M - 1}\big(I - \widehat{\B}^{(M - k)}(z)\big)$ \texttt{\# compute total filter}\\
+%   \Return{$\widehat{\B}(z)$}
+% \end{algorithm}
 
 \section{Empirical Evaluation}
-We have implemented simple empirical experiments in Python
-\cite{scipy}, in particular we leverage the LASSO implementation from
+\label{sec:empirical_evaluation}
+We have implemented our empirical experiments in Python \cite{scipy},
+in particular we leverage the LASSO implementation from
 \texttt{sklearn} \cite{sklearn} and the random graph generators from
-\texttt{networkx} \cite{networkx}.  We run the same experiment on
-three separate graph topologies having $n = 75$ nodes: a tree (which
-is strongly causal), a random directed acyclic graph, and an Erdos
-Renyi random graph with edge inclusion probability of $2 / n$.
+\texttt{networkx} \cite{networkx}.  We run the same experiment on two
+separate graph topologies having $n = 50$ nodes: a strongly causal
+graph and a directed acyclic graph.  These are generated respectively
+by drawing a random tree and a random Erdos Renyi graph, then
+directing edges from lower numbered nodes to higher numbered nodes,
+the edge probabilities in the Erdos Renyi graph are tuned to result in
+the same number of edges in expectation.
 
@@
 linear filters constructed by placing $5$ transfer function poles
-uniformly at random in a disc of radius $3 / 4$ (which guarantees
-stability for acyclic graphs).  The resulting system is driven by
-i.i.d. Gaussian random vectors, each component having random variance
-$\sigma_i^2 = 1/2 + r_i$ where $r_i \sim \text{exp}(1/2)$.  Results
-are obtained for $50$ random iterations per each of $6$ dataset
-lengths $T \in \{100, 500, 1000, 2500, 5000, 10000\}$.
-
-The pairwise causality function $\mathsf{pwgc}$ of Algorithm
-\ref{alg:Bz_estimate} are obtained via the residuals
-$\mathbf{\epsilon}_{ij}$ of the model
-$\mathbf{x}_i = \mathbf{X}_{ij}\widehat{\beta_{}} +
-\mathbf{\epsilon}_{ij}$ estimated via ordinary least squares, where
-$\mathbf{X}_{ij}$ includes $10$ lagged values of processes $i$ and
-$j$.
-
-The function $\mathsf{estimate_B}$ in Algorithm \ref{alg:Bz_estimate}
-is carried out via the LASSO \cite{tibshirani2015statistical}, with
-the regularization parameter being chosen to minimize the Bayesian
-information criteria, a more conservative heuristic than the AIC.
+(i.e. $p = 5$) uniformly at random in a disc of radius $3 / 4$ (which
+guarantees stability for acyclic graphs).  The resulting system is
+driven by i.i.d. Gaussian random vectors, each component having random
+variance $\sigma_i^2 = 1/2 + r_i$ where $r_i \sim \text{exp}(1/2)$.
+We set $p_{\text{max}} = 15$ and collect results for $300$ trials with $T$
+varying between $50$ and $5000$.
+
+Results of our Algorithm are compared against the LASSO:
+
+\begin{equation}
+  \begin{aligned}
+  \xi_{\text{LASSO}}^{(\lambda)} &= \underset{B}{\text{min}}\ \frac{1}{T}\sum_{t = p + 1}^T||x(t) - \sum_{\tau = 1}^p B(\tau) x(t - \tau)||_2^2 + \lambda \sum_{\tau = 1}^p ||\matvec B(\tau)||_1\\
+  \xi_{\text{LASSO}} &= \underset{\lambda \ge 0}{\text{min}}\ \xi_{\text{LASSO}}^{(\lambda)} + \mathsf{BIC}(B_{\text{LASSO}}^{(\lambda)})\\
+  \end{aligned}
+\end{equation}
+
+where $||\matvec B||_1$ emphasizes the application of a vector norm, and
+$\lambda$ is the regularization parameter chosen via the BIC.
+
+% The pairwise causality function $\mathsf{pwgc}$ of Algorithm
+% \ref{alg:Bz_estimate} are obtained via the residuals
+% $\mathbf{\epsilon}_{ij}$ of the model
+% $\mathbf{x}_i = \mathbf{X}_{ij}\widehat{\beta_{}} +
+% \mathbf{\epsilon}_{ij}$ estimated via ordinary least squares, where
+% $\mathbf{X}_{ij}$ includes $10$ lagged values of processes $i$ and
+% $j$.
+
+% The function $\mathsf{estimate_B}$ in Algorithm \ref{alg:Bz_estimate}
+% is carried out via the LASSO \cite{tibshirani2015statistical}, with
+% the regularization parameter being chosen to minimize the Bayesian
+% information criteria, a more conservative heuristic than the AIC.
 
@@
 
-  \begin{subfigure}[b]{0.30\textwidth}
-    \caption{Random Tree}
-    \includegraphics[width=\linewidth]{example_tree.pdf}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \caption{Random SCG}
+    \includegraphics[width=\linewidth]{example_scg.pdf}
   \end{subfigure}
-  \begin{subfigure}[b]{0.30\textwidth}
+  \begin{subfigure}[b]{0.45\textwidth}
     \caption{Random DAG}
@@
   \end{subfigure}
-  \begin{subfigure}[b]{0.30\textwidth}
-    \caption{Erdos-Renyi Graph}
-    \includegraphics[width=\linewidth]{example_gnp.pdf}
+
+  \begin{subfigure}[b]{0.45\textwidth}
+    \includegraphics[width=\linewidth]{random_scg_simulation.pdf}
+    % \missingfigure{}
   \end{subfigure}
-
-  \begin{subfigure}[b]{0.30\textwidth}
-    \includegraphics[width=\linewidth]{random_tree_simulation.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.30\textwidth}
+  \begin{subfigure}[b]{0.45\textwidth}
     \includegraphics[width=\linewidth]{random_dag_simulation.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.30\textwidth}
-    % \includegraphics[width=\linewidth]{random_gnp_simulation.pdf}
-    \missingfigure{}
+    % \missingfigure{}
   \end{subfigure}
 \end{figure}
+
+In reference to figure \ref{fig:simulation_results} it should not be
+overly surprising that our PWGC algorithm performs better than the
+LASSO for the case of a strongly causal graph, since in this case the
+assumptions which guarantee the correctness of Algorithm
+\ref{alg:pwgr_heuristic} hold.  However, the performance is still
+markedly superior in the case of a more general DAG.  Finally, we
+remark that the PWGC algorithm can be extended to apply to more
+general graphs via sequential application.  We will also see in
+section \ref{sec:application} that our algorithm can scale effectively
+to large problems.
 

### 2019-06-17 software/simulations.py

@@ -1,3 +1,6 @@
+"""
+"""
+
 import seaborn.apionly as sns
 import matplotlib.pyplot as plt
 import pandas as pd
@@ -141,19 +144,19 @@ def full_single_pass_experiment(simulation_name, graph_type):
                                         method="lstsqr", alpha=alpha)
 
             # ~37s
-            if N_iter == 0:
-                G_hat_lasso = estimate_dense_graph(X, max_lag=p_max,
-                                                   max_T=T,
-                                                   method="lasso")
+            G_hat_lasso = estimate_dense_graph(X, max_lag=p_max,
+                                               max_T=T,
+                                               method="lasso")
 
             errs_pwgc.update(G, G_hat_pwgc, sv2_true, N_iter, T_iter)
             errs_lasso.update(G, G_hat_lasso, sv2_true, N_iter, T_iter)
 
-    D_errs_pwgc, D_true_errs_pwgc, D_MCC_pwgc = errs_pwgc.get_results()
-    D_errs_lasso, D_true_errs_lasso, D_MCC_lasso = errs_lasso.get_results()
+    D_errs_pwgc, D_true_errs, D_MCC_pwgc = errs_pwgc.get_results()
+    D_errs_lasso, _, D_MCC_lasso = errs_lasso.get_results()
 
     plot_results(
-        D_MCC_pwgc, D_errs_pwgc, D_true_errs_pwgc, T_iters,
+        D_MCC_pwgc, D_errs_pwgc, D_MCC_lasso, D_errs_lasso,
+        D_true_errs,
         title=("Test Errors against T (Random {} graph on "
                "$n = {}$ nodes)".format(graph_type, n_nodes)),
         save_file=["../figures/{}_simulation.pdf".format(simulation_name),
@@ -173,9 +176,11 @@ def _fit_sqrt_model(x, y):
     return y_low, y_hat, y_high
 
 
-def _plot_sqrt_model(x, y, y_low, y_hat, y_high, ax, color="b",
+def _plot_sqrt_model(x, y, y_low, y_hat, y_high, ax,
+                     color="b", marker="o",
                      fill_between=True):
-    ax.scatter(x, y, color=color, marker="o")
+    ax.scatter(x, y, color=color, marker=marker, alpha=0.75,
+               linewidths=0.0, edgecolors=color)
     ax.plot(x, y_hat, color=color, linewidth=2)
[…]

### 2019-06-17 software/lasso_comparison.py
Docstring.

@@ -1,4 +1,5 @@
 """
+Comparisons for PWGC and LASSO.
 """
 
 import seaborn.apionly as sns

### 2019-06-18 paper/paper.tex
Parallel implementation of pairwise GC stats.

@@
 remark that the PWGC algorithm can be extended to apply to more
-general graphs via sequential application.  We will also see in
-section \ref{sec:application} that our algorithm can scale effectively
-to large problems.
+general graphs via sequential application.
+
+In regards scalability, we have observed that performing the $O(n^2)$
+pairwise Granger-causality calculations consumes the vast majority
+($> 90\%$) of the computation time.  Since this step is trivially
+parallelizable\footnote{We admit that we are using only a
+  single-threaded implementation}, our algorithm should scale well
+with multiple cores or multiple machines.
+
+% We will also see in section \ref{sec:application} that our algorithm
+% can scale effectively to large problems.
 

### 2019-06-18 software/lasso_comparison.py

@@ -2,7 +2,6 @@
 Comparisons for PWGC and LASSO.
 """
 
-import seaborn.apionly as sns
 import matplotlib.pyplot as plt
 import pandas as pd
 import numpy as np
@@ -35,9 +34,7 @@ class TrackErrors:
 
         self.b_errs = np.zeros((N_iters, N_T_iters))
         self.errs = np.zeros(self.b_errs.shape)
-        self.errs_vec = np.zeros((N_iters, N_T_iters, n_nodes))
         self.true_errs = np.zeros(self.b_errs.shape)
-        self.true_errs_vec = np.zeros(self.errs_vec.shape)
         self.rhos = np.zeros(self.b_errs.shape)
         self.N_cross_edges = np.zeros(self.b_errs.shape)
         self.N_hat_cross_edges = np.zeros(self.b_errs.shape)
@@ -48,18 +45,14 @@ class TrackErrors:
         b_err = np.mean(list(map(np.var,
                                  get_estimation_errors(G, G_hat))))
         err = np.sum(get_errors(G_hat))
-        err_vec = np.sort(get_errors(G_hat))
         true_err = np.sum(sv2_true)
-        true_err_vec = np.sort(sv2_true)
 
         self.rhos[N_iter, T_iter] = gcg_to_var(
             G_hat, "b_hat(z)", assert_stable=False).get_rho()
 
         self.b_errs[N_iter, T_iter] = b_err
         self.errs[N_iter, T_iter] = err
-        self.errs_vec[N_iter, T_iter] = err_vec
         self.true_errs[N_iter, T_iter] = true_err
-        self.true_errs_vec[N_iter, T_iter] = true_err_vec
         self.N_cross_edges[N_iter, T_iter] = len(
             set((i, j) for i, j in G.edges if i != j))
         self.N_hat_cross_edges[N_iter, T_iter] = len(set(
@@ -85,7 +78,7 @@ class TrackErrors:
         return D_errs, D_true_errs, D_MCC
 
 
-def full_single_pass_experiment(simulation_name, graph_type):
+def lasso_comparison(simulation_name, graph_type):
     np.random.seed(0)
     n_nodes, p_lags, p_max = 50, 5, 15
     alpha, N_iters = 0.05, 3
@@ -243,4 +236,5 @@ def plot_results(D_MCC_pwgc, D_errs_pwgc,
         else:  # Allow passing a list
             _ = [fig.savefig(f) for f in save_file]
     plt.show()
+
     return

### 2019-06-18 software/pwgc/gc_methods.py

@@ -5,16 +5,18 @@ Methods which implement granger-causality computations.
 import networkx as nx
 import numpy as np
 import numba
-from levinson import (lev_durb, whittle_lev_durb)
+
+from multiprocessing import Pool
+from itertools import product
+from functools import reduce
+from math import ceil
 
 from scipy import linalg
 from scipy import stats as sps
 from scipy.linalg import toeplitz
 from sklearn.linear_model import LassoLarsIC
 
-from itertools import product
-from functools import reduce
-from math import ceil
+from levinson import (lev_durb, whittle_lev_durb)
 
 # TODO: What is the proper way to do this?  I want packaging,
 # TODO: but I also want to be able to C-c C-c this file into
@@ -335,19 +337,26 @@ def compute_xi(X, max_lag):
 # Numba parallelism seems much slower unfortunately
 # @numba.jit(nopython=True, cache=True, parallel=True)
 @numba.jit(nopython=True, cache=True)
-def fast_compute_xi(X, max_lag=10):
-    T, n = X.shape
-    R = compute_covariances(X, max_lag)
+def fast_compute_xi(R, T, n_min=0, n_max=np.inf):
+    """
+    Calculates Xi[n_min: n_max, :]
+    """
+    _, n, _ = R.shape
+    n_min = max(n_min, 0)
+    n_max = min(n_max, n)
 
-    xi_i = np.zeros(n)
-    xi_ij, p_ij = np.zeros((n, n)), np.zeros((n, n))
+    xi_i = np.nan * np.zeros(n)
+    xi_ij, p_ij = np.nan * np.zeros((n, n)), np.nan * np.zeros((n, n))
 
-    for i in range(n):
+    for i in range(n_min, n_max):
         xi_i[i], p_i = _fast_univariate_AR_error(R[:, i, i], T)
         xi_ij[i, i] = xi_i[i]
         p_ij[i, i] = p_i
 
-    for i in numba.prange(n):
+    # Recall that we are computing a horizontal band
+    # as well as the transposed vertical band
+
+    for i in range(n_min, n_max):
         for j in range(i):
             Rij = form_bivariate_covariance(R, i, j)
             ei, ej, _p = _fast_bivariate_AR_error(Rij, T)
[…]

### 2019-06-18 software/pwgc/test_gc_methods.py

@@ -10,7 +10,7 @@ mpl_rc("font", **font)
 mpl_rc("text", usetex=False)
 
 from var_system import (random_gnp_dag, drive_gcg, get_errors,
-                        get_estimation_errors, random_tree_dag,
+                        get_estimation_errors, random_scg,
                         get_X, attach_node_prop, gcg_to_var)
 from gc_methods import (compute_pairwise_gc, estimate_B, pw_scg,
                         estimate_graph, full_filter_estimator,
@@ -70,7 +70,7 @@ def test_pw_scg():
 
     T_max = 5000
 
-    random_graph = lambda: random_tree_dag(
+    random_graph = lambda: random_scg(
         n_nodes, p_lags, pole_rad=0.75)
     # random_graph = lambda: random_gnp_dag(
     #     n_nodes, p_lags, pole_rad=0.75, edge_prob=2./n_nodes)

### 2019-06-19 paper/paper.tex
Simulation on increasing edge probability, more cleanup on paper.

--- 2019-06-17 paper/paper.tex
@@
 
-We proceed first by defining pairwise Granger-causality hypothesis
-tests, second a model order selection criteria, and finally the method
-for choosing a $P-$value cutoff.
-
-\subsection{Hypothesis Testing}
+In subsection \ref{sec:pairwise_hypothesis_testing} we define pairwise
+Granger-causality hypothesis tests, in subsection
+\ref{sec:model_order_selection} a model order selection criteria, in
+subsection \ref{sec:efficient_model_estimation} an efficient
+estimation algorithm, in subsection \ref{sec:error_rate_control} the
+method for choosing an hypothesis testing threshold, and finally in
+subsection \ref{sec:finite_pwgc} the unified finite sample algorithm.
+
+\subsection{Pairwise Hypothesis Testing}
+\label{sec:pairwise_hypothesis_testing}
 In performing pairwise checks for Granger-causality $x_j \pwgc x_i$ we
-follow the simple scheme of estimating the two linear models via
-ordinary least squares:
+follow the simple scheme of estimating the following two linear models:
 
@@
   H_0:&\ \widehat{x}_i^{(p_i)}(t) = \sum_{\tau = 1}^{p_i} b_{ii}(\tau)x_i(t - \tau),\\
-  H_1:&\ \widehat{x}_i^{(p_i, p_j)}(t) = \sum_{\tau = 1}^{p_i} b_{ii}(\tau)x_i(t - \tau) + \sum_{\tau = 1}^{p_j}b_{ij}(\tau)x_j(t - \tau).
+  H_1:&\ \widehat{x}_i^{(p_{ij})}(t) = \sum_{\tau = 1}^{p_{ij}} b_{ii}(\tau)x_i(t - \tau) + \sum_{\tau = 1}^{p_{ij}}b_{ij}(\tau)x_j(t - \tau).
 \end{align}
@@
 Under the null hypothesis $H_0$ (i.e. that $x_j \npwgc x_i$) we can
-form the asymptotically $\chi^2(p_j)$ distributed test statistic
+form the asymptotically $\chi^2(p_{ij})$ distributed test statistic
 (\hl{TODO: Citation that this is chi2})
@@
   \label{eqn:gc_statistics}
-  F_{ij}(p_i, p_j) = \frac{T}{p_j}\Big(\frac{\xi_i(p_i)}{\xi_{ij}(p_i, p_j))} - 1\Big),
+  F_{ij}(p_i, p_{ij}) = \frac{T}{p_{ij}}\Big(\frac{\xi_i(p_i)}{\xi_{ij}(p_{ij})} - 1\Big),
 \end{equation}
@@
 \begin{equation*}
-  \xi_i(p_i) = \frac{1}{T - p_i}\sum_{t = p_i + 1}^T (x_i(t) - \widehat{x}_i^{(p_i)}(t))^2
+  \xi_i(p_i) = \frac{1}{T - p_i}\sum_{t = p_i + 1}^T (x_i(t) - \widehat{x}_i^{(p_i)}(t))^2,
 \end{equation*}
 
-Given additional assumptions on $x(t)$ (\hl{TODO: What assumptions? --
-  this must also depend on choice of $p_i, p_j$}) we have the
-convergence (\hl{a.s. or in P?  -- I think in P})
+and similarly for $\xi_{ij}(p_{ij})$.  Under moment conditions $x(t)$
+(\hl{TODO: What assumptions? -- this must also depend on choice of
+  $p_i, p_j$}) we have the convergence in probability:
 
 \begin{equation}
-  \underset{T \rightarrow \infty}{\text{lim }}F_{ij} =
+  F_{ij}(p_i, p_{ij}) \overset{\P}{\rightarrow}
   \left\{
@@
     \end{array}
-  \right.
+  \right. \text{ as } T \rightarrow \infty.  % the '.' after \right is necessary.
 \end{equation}
@@
 There are a variety of methods to choose the filter orders $p_i$ and
-$p_j$ \hl{TODO: cite some more}, we will focus in particular on the
-Bayesian Information Criteria (BIC).  The BIC is substantially more
-conservative than the Akaiake Information Criteria, and since we are
-searching for \textit{sparse graphs}, we therefore prefer the BIC. For
-the bivariate case we focus for simplicity on $p_i = p_j$ and write
-simply\footnote{We use the same $p$ for the bivariate case, but this
-  need not be the same as the corresponding univariate estimate}
-$x_{ij}^{(p)}$, in which case the BIC, which we seek to
-\textit{minimize} over $p$ is given by:
+$p_j$ (see e.g. \cite{lutkepohl2005new}), but we will focus in
+particular on the Bayesian Information Criteria (BIC).  The BIC is
+substantially more conservative than the popular alternative Akaiake
+Information Criteria (the BIC is also asymptotically consistent), and
+since we are searching for \textit{sparse graphs}, we therefore prefer
+the BIC, where we seek to \textit{minimize} over $p$:
 
@@
   \begin{aligned}
-    BIC_{\text{univariate}}(p) &= T \ln \xi_i(p) + p\ln T,\\
-    BIC_{\text{bivariate}}(p) &= T \ln \xi_{ij}(p) + 2p\ln T.\\
+    BIC_{\text{univariate}}(p) &= \ln\ \xi_i(p) + p\frac{\ln T}{T},\\
+    BIC_{\text{bivariate}}(p) &= \ln \det \widehat{\Sigma}_{ij}(p) + 4p\frac{\ln T}{T},\\
   \end{aligned}
@@
 
+where $\widehat{\Sigma}_{ij}(p)$ is the $2 \times 2$ residual
+covariance matrix for the $VAR(p)$ model of $(x_i(t), x_j(t))$.  The
+bivariate errors $\xi_{ij}(p)$ and $\xi_{ji}(p)$ are the diagonal
+entries of $\widehat{\Sigma}_{ij}(p)$.
+
 We carry this out by a simple direct search on each model order
-between $1$ and some prescribed $p_\text{max}$.  In practice it is
+between $0$ and some prescribed $p_\text{max}$.  In practice it is
 sufficient to pick $p_\text{max}$ ad-hoc or via some simple heuristic
-(e.g. plotting the sequence $BIC(p)$ over $p$), and we will review in
-Section \ref{sec:efficient_model_estimation} an algorithm to carry out
-this search efficiently.
-
-% \begin{algorithm}
-%   \caption{BIC Model Selection Algorithm}
-%   \label{alg:maximize_bic}
-
-%   \SetKwInOut{Input}{input}
-%   \SetKwInOut{Output}{output}
-%   \SetKwInOut{Initialize}{initialize}
-%   \DontPrintSemicolon
-
-%   \BlankLine
-%   \Input{$T$ samples of WSS processes $x_i(t), x_j(t)$.  A maximum possible lag $p_{\text{max}}$.}
-%   \Output{Model orders $p_i$ and $p$ for models in Hypotheses $H_0$ and $H_1$ respectively.}
-%   \Initialize{
-%     $BIC_{\text{univariate}} = -\infty$\\
-%     $BIC_{\text{bivariate}} = -\infty$\\
-%   }
-
-%   \# \texttt{Find $p_i$}\\
-%   \For{$p \in \{1, 2, \ldots, p_{\text{max}}\}$}{
-%     $\xi_i(p) = $\ \texttt{fit\_univariate}$(x_i(t), p)$ \texttt{ \# Fit model for $H_0$}\\
-%     $bic_p = -(T - p - 1) \ln \xi_i(p) - p\ln(T - p - 1)$\\
-%     \If{$bic_p > BIC_{\text{univariate}}$}{
-%       $BIC_{\text{univariate}} = bic_p$\\
-%       \texttt{continue}\\
-%     } \Else{
-%       $p_i = p - 1$\\
-%       \texttt{break}\\
-%     }
-%   }
-
-%   \BlankLine
-%   \# \texttt{Find $p$}\\
-%   \For{$p \in \{1, 2, \ldots, p_{\text{max}}\}$}{
-%     $\xi_{ij}(p) = $\ \texttt{fit\_biivariate}$(x_i(t), x_j(t), p)$ \texttt{ \# Fit model for $H_1$}\\
-%     $bic_p = -(T - p - 1) \ln \xi_{ij}(p) - 2p\ln(T - p - 1)$\\
-%     \If{$bic_p > BIC_{\text{bivariate}}$}{
-%       $BIC_{\text{biivariate}} = bic_p$\\
-%       \texttt{continue}\\
-%     }\Else{
-%       $p = p - 1$\\
-%       \texttt{break}\\
-%     }
-%   }
-%     \Return{$p_i, p$}
-% \end{algorithm}
+e.g. plotting the sequence $BIC(p)$ over $p$, though it is not
+technically possible to guarantee that the optimal $p$ is chosen
+(since there can in general be arbitrarily long lags from one variable
+to another).
 
@@
 implementing our estimation algorithm is spent calculating the error
-estimates $\xi_i(p_i)$ and $\xi_{ij}(p, p)$.  This requires fitting a
-total of $n^2p_{\text{max}}$ autoregressive models, where the most
+estimates $\xi_i(p_i)$ and $\xi_{ij}(p_{ij})$.  This requires fitting
+a total of $n^2p_{\text{max}}$ autoregressive models, where the most
 naive algorithm (e.g. solving a least squares problem for each model)
@@
 possible to carry out this task in a much more modest
-$O(n^2p_{\text{max}}^2 ) + O(n^2p_{\text{max}}T)$ time.
-
-Firstly, we estimate the sequence of Covariance
-matrices\footnote{The particular indexing and normalization given in
-  equation \ref{eqn:covariance_estimate} is critical to ensure
-  $\widehat{R}$ is positive semidefinite.  The estimate can be viewed
-  as calculating the covariance sequence of a signal multiplied by a
-  rectangular window.}
+$O(n^2p_{\text{max}}^2 ) + O(n^2p_{\text{max}}T)$ time via the
+autocorrelation method which deploys the following autocovariance
+estimates in the Yule-Walker equations:\footnote{The particular
+  indexing and normalization given in equation
+  \ref{eqn:covariance_estimate} is critical to ensure $\widehat{R}$ is
+  positive semidefinite.  The estimate can be viewed as calculating
+  the covariance sequence of a signal multiplied by a rectangular
+  window.}
 
@@
 
-an $O(n^2p_{\text{max}}T)$ operation.  Secondly, the variance
-estimates $\xi_i(p)$ for $p = 1, \ldots, p_{\text{max}}$ can be
-evaluated in $O(p_{\text{max}}^2)$ time each by applying the
-Levinson-Durbin recursion to $\widehat{R}_{ii}(tau)$, which
-effectively estimates a sequence of $AR$ models, producing $\xi_i(p)$
-as a side-effect (see
+These covariance estimates constitute the $O(n^2p_{\text{max}}T)$
+operation.  Given these particular estimates, the variances $\xi_i(p)$
+for $p = 1, \ldots, p_{\text{max}}$ can be evaluated in
+$O(p_{\text{max}}^2)$ time each by applying the Levinson-Durbin
+recursion to $\widehat{R}_{ii}(\tau)$, which effectively estimates a
+sequence of $AR$ models, producing $\xi_i(p)$ as a side-effect (see
 \cite{hayes_statistical_digital_signal_processing} and
@@
 
-Similarly, the variance estimates $\xi_{ij}$ and $\xi_{ji}$ can be
-obtained by estimating $\frac{(n + 1)n}{2}$ bivariate AR models, again
-in $O(p_{\text{max}}^2)$ via Whittle's generalized LD recursion
+Similarly, the variance estimates $\widehat{\Sigma}_{ij}(p)$ (which
+include $\xi_{ij}$ and $\xi_{ji}$) can be obtained by estimating
+$\frac{(n + 1)n}{2}$ bivariate AR models, again in
+$O(p_{\text{max}}^2)$ time via Whittle's generalized Levinson-Durbin
+recursion\footnote{We have made use of standalone tailor made
+  implementations of these algorithms, available at
+  \textsf{github.com/RJTK/Levinson-Durbin-Recursion}.}
 \cite{whittle_generalized_levinson_durbin}.
@@
 \subsection{Edge Probabilities and Error Rate Controls}
+\label{sec:error_rate_control}
 Denote $F_{ij}$ the Granger-causality statistic of equation
@@
 Section \ref{sec:model_order_selection}.  Note that this statistic is
-asymptotically $\chi^2(p)$ distributed, and denote by $G$ the
-$\chi^2(p)$ cumulative distribution function.  We will define the matrix
+asymptotically $\chi^2(p_{ij})$ distributed, and denote by $G$ the
+cumulative distribution function thereof.  We will define the matrix
 
@@
 pairwise edges to a level $\alpha$ (where we generally take
-$\alpha = 0.05$).
+$\alpha = 0.05$).  Applying the extremely simple Bonferroni method is
+overly conservative for this application.
+
+\subsection{Finite Sample Recovery Algorithm}
+\label{sec:finite_pwgc}
+
+\begin{algorithm}
+    \SetKwInOut{Input}{input}
+    \SetKwInOut{Output}{output}
+    \SetKwInOut{Initialize}{initialize}
+    \DontPrintSemicolon
+
+    \BlankLine
+    \caption{Finite Sample Pairwise Graph Recovery (PWGC)}
+    \label{alg:finite_pwgc}
+
+\end{algorithm}
 
@@
 \texttt{sklearn} \cite{sklearn} and the random graph generators from
-\texttt{networkx} \cite{networkx}.  We run the same experiment on two
-separate graph topologies having $n = 50$ nodes: a strongly causal
-graph and a directed acyclic graph.  These are generated respectively
-by drawing a random tree and a random Erdos Renyi graph, then
-directing edges from lower numbered nodes to higher numbered nodes,
-the edge probabilities in the Erdos Renyi graph are tuned to result in
-the same number of edges in expectation.
+\texttt{networkx} \cite{networkx}.  We run experiments using two
+separate graph topologies having $n$ nodes: a strongly causal graph
+(SCG) and a directed acyclic graph (DAG).  These are generated
+respectively by drawing a random tree and a random Erdos Renyi graph
+(with edge probability $q = \frac{2}{n}$ resulting in approximately
+the same number of edges for the SCG as for the DAG), then creating a
+directed graph by directing edges from lower numbered nodes to higher
+numbered nodes.
 
@@
 guarantees stability for acyclic graphs).  The resulting system is
-driven by i.i.d. Gaussian random vectors, each component having random
+driven by i.i.d. Gaussian random noise, each component having random
 variance $\sigma_i^2 = 1/2 + r_i$ where $r_i \sim \text{exp}(1/2)$.
-We set $p_{\text{max}} = 15$ and collect results for $300$ trials with $T$
-varying between $50$ and $5000$.
-
-Results of our Algorithm are compared against the LASSO:
+We set $p_{\text{max}} = 15$.  Results and representitive graphs are
+collected in Figure \ref{fig:simulation_results}.
+
+We compare our results against the LASSO as a baseline, estimating the
+filter $\B(z)$ that minimizes $\xi_{\text{LASSO}}$ in the following:
 
@@
 
-% The pairwise causality function $\mathsf{pwgc}$ of Algorithm
-% \ref{alg:Bz_estimate} are obtained via the residuals
-% $\mathbf{\epsilon}_{ij}$ of the model
-% $\mathbf{x}_i = \mathbf{X}_{ij}\widehat{\beta_{}} +
-% \mathbf{\epsilon}_{ij}$ estimated via ordinary least squares, where
-% $\mathbf{X}_{ij}$ includes $10$ lagged values of processes $i$ and
-% $j$.
-
-% The function $\mathsf{estimate_B}$ in Algorithm \ref{alg:Bz_estimate}
-% is carried out via the LASSO \cite{tibshirani2015statistical}, with
-% the regularization parameter being chosen to minimize the Bayesian
-% information criteria, a more conservative heuristic than the AIC.
-
 \begin{figure}
@@
 
-  \begin{subfigure}[b]{0.45\textwidth}
+  {\scriptsize Comparison of PWGC and LASSO for $VAR(p)$ model
+    estimation.  The MCC is measured on the left axis in blue with
+    circle markers, and the relative log mean-squared prediction error
+    $\frac{\ln\ \tr \widehat{\Sigma}_v}{\ln\ \tr \Sigma_v}$ on the
+    right axis in red with triangle markers.  We use $q$ to denote the
+    edge probability in a random Erdos-Reny graph (modified to form a
+    DAG).  Solid lines indicate ad-hoc regression estimates for
+    visualization purposes only.
+
+    Matthew's Correlation Coefficient (MCC)
+    \cite{matthews1975comparison} synthesizes the confusion matrix
+    into a single score appropriate for unbalanced labels and is
+    calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
+    performance, $0$ being the performance of random guessing, and
+    $-1$ being perfectly opposed.
+  }
+
+  \begin{subfigure}[b]{0.3\textwidth}
     \caption{Random SCG}
@@
   \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{Random DAG}
+  \begin{subfigure}[b]{0.3\textwidth}
+    \caption{Random DAG $(q = \frac{2}{n})$}
     \includegraphics[width=\linewidth]{example_dag.pdf}
   \end{subfigure}
+  \begin{subfigure}[b]{0.3\textwidth}
+    \caption{Random DAG $(q = \frac{4}{n})$}
+    \includegraphics[width=\linewidth]{example_dag_q.pdf}
+  \end{subfigure}
 
   \begin{subfigure}[b]{0.45\textwidth}
+    \caption{PWGC vs LASSO (SCG)}
+    \label{fig:simulation_results_comparison1}
     \includegraphics[width=\linewidth]{random_scg_simulation.pdf}
@@
   \begin{subfigure}[b]{0.45\textwidth}
+    \caption{PWGC vs LASSO (DAG, $q = \frac{2}{n}$)}
+    \label{fig:simulation_results_comparison2}
     \includegraphics[width=\linewidth]{random_dag_simulation.pdf}
@@
   \end{subfigure}
+
+  \begin{subfigure}[b]{0.45\textwidth}
+    \caption{Fixed $T$, increasing $n$ (SCG)}
+    \label{fig:simulation_results_scaling}
+    \includegraphics[width=\linewidth]{increasing_n_simulation.pdf}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \caption{Fixed $T, n$, increasing edges $q$ (DAG)}
+    \label{fig:simulation_results_dense}
+    \includegraphics[width=\linewidth]{dag_increasing_q_simulation.pdf}
+  \end{subfigure}
 \end{figure}
 
-In reference to figure \ref{fig:simulation_results} it should not be
-overly surprising that our PWGC algorithm performs better than the
-LASSO for the case of a strongly causal graph, since in this case the
-assumptions which guarantee the correctness of Algorithm
-\ref{alg:pwgr_heuristic} hold.  However, the performance is still
-markedly superior in the case of a more general DAG.  Finally, we
-remark that the PWGC algorithm can be extended to apply to more
-general graphs via sequential application.  We will also see in
-section \ref{sec:application} that our algorithm can scale effectively
-to large problems.
+In reference to figure \ref{fig:simulation_results_comparison1} it
+should not be overly surprising that our PWGC algorithm performs
+better than the LASSO for the case of a strongly causal graph, since
+in this case the assumptions which guarantee the correctness of
+Algorithm \ref{alg:pwgr_heuristic} hold.  However, the performance is
+still markedly superior in the case of a more general DAG.
+
+In regards scalability, we have observed that performing the $O(n^2)$
+pairwise Granger-causality calculations consumes the vast majority
+($> 90\%$) of the computation time.  Since this step is trivially
+parallelizable, our algorithm also scales well with multiple cores or
+multiple machines.  Figure \ref{fig:simulation_results_scaling} is a
+demonstrating of this scalability, where we are able to estimate
+graphs having over $1500$ nodes (over $2.25 \times 10 ^6$ possible
+edges) using only $T = 500$ data points, granted, an SCG on this many
+nodes is extremely sparse.  Application to denser graphs is explored
+in Figure \ref{fig:simulation_results_dense}, where $n$ and $T$ are
+held fixed while increasing the edge probability $q$, thereby creating
+denser graphs.  For larger values of $q$ in this plot there is a
+comparable number of edges as there are data samples $T$.
 

### 2019-06-19 software/lasso_comparison.py

@@ -83,13 +83,13 @@ def lasso_comparison(simulation_name, graph_type):
     n_nodes, p_lags, p_max = 50, 5, 15
     alpha, N_iters = 0.05, 3
 
-    T_iters = [50 + 50 * k for k in range(1, 99)]
+    T_iters = list(map(int, np.logspace(2, 4.7, 100)))
+    # T_iters = [50 + 250 * k for k in range(1, 149)]
 
     # NOTE: We use only X[T:] to estimate the error
     # NOTE: so it is important to have a significant trailing
     # NOTE: set of data in order to make honest estimates of
     # NOTE: the out of sample predictions.
-    T_max = T_iters[-1] * 2
 
     # Tune the expected number of edges to match N_edges
     N_edges = (len(random_scg(n_nodes, p_lags, pole_rad=0.75).edges) -
@@ -112,10 +112,10 @@ def lasso_comparison(simulation_name, graph_type):
         raise NotImplementedError("graph_type '{}' is not understood"
                                   "".format(graph_type))
 
-    def make_test_data():
+    def make_test_data(T):
         sv2_true = 0.5 + np.random.exponential(0.5, size=n_nodes)
         G = random_graph()
-        drive_gcg(G, T_max, sv2_true, filter_attr="b(z)")
+        drive_gcg(G, 500 + 2 * T, sv2_true, filter_attr="b(z)")
         X = get_X(G)
         X = X - np.mean(X, axis=0)[None, :]
         X = X
@@ -131,7 +131,7 @@ def lasso_comparison(simulation_name, graph_type):
                 N_iter + 1, N_iters, T_iter + 1, len(T_iters)))
 
             # ~252ms
-            G, X, sv2_true = make_test_data()
+            G, X, sv2_true = make_test_data(T)
 
             # ~824ms
             G_hat_pwgc = estimate_graph(X[:T], G, max_lags=p_max,

### 2019-06-19 software/pwgc/gc_methods.py

@@ -40,7 +40,18 @@ def estimate_b_lstsqr(X, y):
     y: np.array (T)
     X: np.array (T x s)
     """
-    w = linalg.cho_solve(linalg.cho_factor(X.T @ X), X.T @ y)
+    T, n = X.shape
+    R = X.T @ X / T
+    r = X.T @ y / T
+    try:
+        w = linalg.cho_solve(linalg.cho_factor(R), r)
+    except linalg.LinAlgError:
+        R = R + np.min(np.diag(R)) *  np.eye(R.shape[0])
+
+    try:
+        w = linalg.cho_solve(linalg.cho_factor(R), r)
+    except linalg.LinAlgError:
+        w, _, _, _ = linalg.lstsq(R, r)
     return w
 
 
@@ -250,11 +261,6 @@ def compute_covariances(X, p):
 def _compute_AR_error(X, y):
     w = estimate_b_lstsqr(X, y)
     return np.var(y - X @ w)
-
-
-def _fast_compute_AR_error(R, r):
-    w = estimate_b_lstsqr_cov(R, r)
-    return max(0, R[0, 0] - w @ r)
 
 
 @numba.jit(nopython=True, cache=True)

### 2019-06-19 software/simulation_increasing_n.py

@@ -1,0 +1,164 @@
+"""
+Simulations with T held fixed and n increasing.
+"""
+
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import statsmodels.api as sm
+
+from matplotlib import rc as mpl_rc
+font = {"family" : "normal",
+        "weight" : "bold",
+        "size"   : 22}
+
+mpl_rc("font", **font)
+mpl_rc("text", usetex=True)
+
+
+from pwgc.var_system import (random_gnp_dag, drive_gcg, get_errors,
+                             get_estimation_errors, random_scg,
+                             get_X, attach_node_prop, gcg_to_var)
+from pwgc.gc_methods import (compute_pairwise_gc, estimate_B, pw_scg,
+                             estimate_graph, full_filter_estimator,
+                             compute_MCC_score, estimate_dense_graph)
+
+
+class TrackErrors:
+    def __init__(self, n_iters, N_reps):
+        self.n_iters = n_iters
+        self.N_reps = N_reps
+        self.b_errs = np.zeros((len(self.n_iters), N_reps))
+        self.errs = np.zeros(self.b_errs.shape)
+        self.true_errs = np.zeros(self.b_errs.shape)
+        self.N_cross_edges = np.zeros(self.b_errs.shape)
+        self.N_hat_cross_edges = np.zeros(self.b_errs.shape)
+        self.N_cross_intersect_edges = np.zeros(self.b_errs.shape)
+        return
+
+    def update(self, G, G_hat, sv2_true, n_iter, rep):
+        b_err = np.mean(list(map(np.var,
+                                 get_estimation_errors(G, G_hat))))
+        err = np.sum(get_errors(G_hat))
+        true_err = np.sum(sv2_true)
+
+        self.b_errs[n_iter, rep] = b_err
+        self.errs[n_iter, rep] = err
+        self.true_errs[n_iter, rep] = true_err
+
+        self.N_cross_edges[n_iter, rep] = len(
+            set((i, j) for i, j in G.edges if i != j))
+        self.N_hat_cross_edges[n_iter, rep] = len(set(
+            (i, j) for i, j in G_hat.edges if i != j))
+        self.N_cross_intersect_edges[n_iter, rep] = len(
+            (set(G.edges) & set(G_hat.edges)) - set((i, i)
+                                                    for i in G.nodes))
+        return
[…]

### 2019-06-19 software/simulation_increasing_p.py

@@ -1,0 +1,191 @@
+"""
+Simulation with both T and n held fixed, but with increasing
+ER edge probability in the DAG generator.
+"""
+
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import statsmodels.api as sm
+
+from sklearn.kernel_ridge import KernelRidge
+from sklearn.model_selection import GridSearchCV
+
+from matplotlib import rc as mpl_rc
+font = {"family" : "normal",
+        "weight" : "bold",
+        "size"   : 22}
+
+mpl_rc("font", **font)
+mpl_rc("text", usetex=True)
+
+
+from pwgc.var_system import (random_gnp_dag, drive_gcg, get_errors,
+                             get_estimation_errors, random_scg,
+                             get_X, attach_node_prop, gcg_to_var)
+from pwgc.gc_methods import (compute_pairwise_gc, estimate_B, pw_scg,
+                             estimate_graph, full_filter_estimator,
+                             compute_MCC_score, estimate_dense_graph)
+
+
+class TrackErrors:
+    def __init__(self, q_iters, N_reps, n):
+        self.q_iters = q_iters
+        self.n = n
+        self.N_reps = N_reps
+        self.b_errs = np.zeros((len(self.q_iters), N_reps))
+        self.errs = np.zeros(self.b_errs.shape)
+        self.true_errs = np.zeros(self.b_errs.shape)
+        self.N_cross_edges = np.zeros(self.b_errs.shape)
+        self.N_hat_cross_edges = np.zeros(self.b_errs.shape)
+        self.N_cross_intersect_edges = np.zeros(self.b_errs.shape)
+        return
+
+    def update(self, G, G_hat, sv2_true, n_iter, rep):
+        b_err = np.mean(list(map(np.var,
+                                 get_estimation_errors(G, G_hat))))
+        err = np.sum(get_errors(G_hat))
+        true_err = np.sum(sv2_true)
+
+        self.b_errs[n_iter, rep] = b_err
+        self.errs[n_iter, rep] = err
+        self.true_errs[n_iter, rep] = true_err
+
+        self.N_cross_edges[n_iter, rep] = len(
+            set((i, j) for i, j in G.edges if i != j))
+        self.N_hat_cross_edges[n_iter, rep] = len(set(
[…]

### 2019-06-21 paper/paper.tex
More references for paper plus some prose.

@@
 highly dependent upon the data that we are able to observe.  Rather,
-the interpretation of Granger causality is as a means of uncovering a
+the interpretation of Granger causality is as a means of uncovering a plausible
 flow of ``information'' or ``energy'' through some underlying graph of
@@
 connectedness of financial institutions \cite{NBERw16223}; gene
-expression networks \cite{Fujita2007}; and power system design
-\cite{Misyrlis2016450}, \cite{yuan2014root}.
+expression networks \cite{Fujita2007},
+\cite{grouped_graphical_granger_modelling_for_gene_expression_regulatory_networks_discovery},
+\cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty},
+and power system design \cite{Misyrlis2016450}, \cite{yuan2014root}.
+
+\hl{TODO: From the gene network papers, it looks like these gene
+  networks may actually have strongly-causal structure.  i.e. we can
+  highlight the fact that our deeper understanding of a special
+  problem case may in fact have significant applicability in these
+  application areas.}
+
+Other citations: support recovery consistency for the weighted lasso
+in univariate AR(p) models:
 
@@
 the comparison of quality between different structures, and hence the
-notion of ``best'', needs to be quantified.  In applications where we
+notion of ``best'' needs to be quantified.  In applications where we
 are interested merely in minimizing the mean squared error of a linear
@@
 convex relaxation to penalizing the count of the non-zero edges.  The
-LASSO enjoys strong theoretical guarantees \cite{wainwright2009sharp},
-even in the case of time series data \cite{basu2015}
-\cite{wong2016lasso}, variations on which have been applied elsewhere
-\cite{DBLP:journals/corr/HallacPBL17} \cite{haufe2008sparse}
-\cite{bolstad2011causal} \cite{he2013stationary}.
+LASSO enjoys fairly strong theoretical guarantees
+\cite{wainwright2009sharp}, extending largely to the case of
+stationary time series data with a sufficiently fast rate of
+dependence decay \cite{basu2015} \cite{wong2016lasso}
+\cite{autoregressive_process_modelling_via_the_lasso_procedure}.
+Variations on the LASSO have been applied in a number of different
+contexts \cite{DBLP:journals/corr/HallacPBL17} \cite{haufe2008sparse}
+\cite{bolstad2011causal} \cite{he2013stationary}
+\cite{grouped_graphical_granger_modelling_for_gene_expression_regulatory_networks_discovery}.
+One of the key improvements to the original LASSO algorithm is the
+adaptive (i.e. weighted) adaLASSO \cite{adaptive_lasso_zou2006}, for
+which oracle results are established.  The crux of the matter for time
+series analysis is to craft a good weighting scheme, see e.g.
+\cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty}.
+Success of the LASSO has lead to the famous ``bet on sparsity''
+principle -- that is, for large problems and finite sources of data,
+you might as well bet that a sparse model is correct, because if it
+isn't, there is no hope.
+
+In the context of time series data, sparsity assumptions remain
+important, but there is significant additional structure that may
+arise as a result of considering the topology of the underlying
+Granger-causality graph, which to our knowledge remains largely
+unexplored.  The focus of this paper is to shed light on some of these
+topological questions, in particular, we study a particularly simple
+notion of causality graph topology which we term ``strongly causal''
+and show that $VAR$ models having this structure satisfy naturaal
+intuitive notions of ``information flow'' through the graph.
+Moreover, we show that such graphs are perfectly (asymptotically)
+recoverable with only \textit{pairwise} Granger-causality tests.  Our
+finite sample results are based on simulations, where we show
+promising results for these graphs topologies where our algorithm
+performs substantially better in our simulation setup than do
+competing LASSO algorithms, even for graphs that do not exactly
+satisfy our strongly-causal topology assumptions.
+
+\todo{What graphs in practice may be strongly causal -- see the gene network papers.}
 
@@
 problems in our conclusion (section \ref{sec:conclusion}).
+
+\todo{Simulations with very small T (i.e. T = 20)}
 
@@
 form the asymptotically $\chi^2(p_{ij})$ distributed test statistic
-(\hl{TODO: Citation that this is chi2})
+\todo{Is this actually $\chi^2$ when $p_{ij}$ is chosen from the data?  This is adaptive!}
 
@@
 Therefore, we can make the statement that Algorithm
-\ref{alg:pwgr_heuristic} asymptotically implements Algorithm
+\ref{alg:finite_pwgc} asymptotically implements Algorithm
 \ref{alg:pwgr}.  \hl{TODO: Modify Algorithm 2 to make this true.}
@@
 
+After the graph topology $\widehat{\gcg}$ has been estimated via
+Algorithm \ref{alg:finite_pwgc}, we refit the entire model with the
+specified sparsity pattern directly via ordinary least squares.
+Producing graph estimates which are not strongly causal can be
+achieved out by performing sequential estimates
+$\widehat{x}_1(t), \widehat{x}_2(t), \ldots$ by estimating a strongly
+causal graph with the residuals of the previous model as input, and
+then refitting on the estimated sparsity pattern.  However we reserve
+any analysis of this scheme for future work.
+
 \begin{algorithm}
@@
     \label{alg:finite_pwgc}
-
+    \Input{Estimates of pairwise Granger-causality statistics $F_{ij}$
+      (eqn \ref{eqn:gc_statistics}).
+Matrix of edge probabilities $P_{ij}$ (eqn \ref{eqn:edge_inclusion_probability}).
+Hypothesis testing threshold $\delta$ chosen via the Benjamini-Hochberg criterion (Section \ref{sec:error_rate_control})}
+    \Output{A strongly causal graph $\widehat{\gcg}$}
+    \Initialize{$S = [n]$  \texttt{\# unprocessed nodes}\\
+      $E = \emptyset$  \texttt{\# edges of }$\widehat{\gcg}$\\
+      $k = 1$ \texttt{\# a counter used only for notation}}
+
+    \BlankLine
+
+    $W_\delta \leftarrow \{(i, j)\ |\ P_{ji} > 1 - \delta, F_{ji} > F_{ij}\}$  \texttt{\# candidate edges}\\
+    $\mathcal{I}_0 \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# total node incident probability}\\
+    $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) < \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Nodes with fewest incident edges}\\
+    \If{$P_0 = \emptyset$}{
+      $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) \le \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Ensure non-empty}
+    }
+    \BlankLine
+
+    \While{$S \ne \emptyset$}{
+      $S \leftarrow S \setminus P_{k - 1}$ \texttt{\# remove processed nodes}\\
+      % $\mathcal{I}_k \leftarrow \big(\sum_{j \in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# intra-}$S$ \texttt{incident strength}\\
+      $\mathcal{I}_k \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$\\
+      $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) < \ceil{\text{min}(\mathcal{I}_k)}\}$\\
+      \If{$P_k = \emptyset$}{
+        $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) \le \ceil{\text{min}(\mathcal{I}_k)}\}$
+      }
+      \;
+      \texttt{\# add strongest edges, maintaining strong causality}\\
+      $U_k \leftarrow \bigcup_{r = 1}^k P_{k - r}$ \texttt{\# Include all forward edges}\\
+      \For{$(i, j) \in \mathsf{sort}\Big(\{(i, j) \in U_k \times P_k\ |\ (i, j) \in W_\delta\} \mathsf{\ by\ descending\ } F_{ji}\Big)$} {
+        \If{$\mathsf{is\_strongly\_causal}(E \cup \{(i, j)\})$} {
+          \texttt{\# }$\mathsf{is\_strongly\_causal}$ \texttt{can be implemented by keeping}\\
+          \texttt{\# track of ancestor / descendant relationships}\\
+          $E \leftarrow E \cup \{(i, j)\}$
+        }
+      }
+      $k \leftarrow k + 1$\\
+    }
+    \Return{$([n], E)$}
 \end{algorithm}
@@
 in this case the assumptions which guarantee the correctness of
-Algorithm \ref{alg:pwgr_heuristic} hold.  However, the performance is
+Algorithm \ref{alg:finite_pwgc} hold.  However, the performance is
 still markedly superior in the case of a more general DAG.
@@
 
+\clearpage
 \printbibliography

### 2019-06-21 software/lasso_comparison.py
Toying with lasso comparisons -- PWGC is doingly poorly for very small amounts of data (probably due to my crude statistical tests).

@@ -83,8 +83,10 @@ def lasso_comparison(simulation_name, graph_type):
     n_nodes, p_lags, p_max = 50, 5, 15
     alpha, N_iters = 0.05, 3
 
-    T_iters = list(map(int, np.logspace(2, 4.7, 100)))
-    # T_iters = [50 + 250 * k for k in range(1, 149)]
+    # T_iters = list(map(int, np.logspace(2, 3, 20)))
+    # T_iters = list(map(int, np.linspace(500, 5000, 20)))
+    # T_iters = [50 + 100 * k for k in range(1, 99)]
+    T_iters = [20, 40, 80, 100, 150, 250, 500]
 
     # NOTE: We use only X[T:] to estimate the error
     # NOTE: so it is important to have a significant trailing
@@ -135,12 +137,27 @@ def lasso_comparison(simulation_name, graph_type):
 
             # ~824ms
             G_hat_pwgc = estimate_graph(X[:T], G, max_lags=p_max,
-                                        method="lstsqr", alpha=alpha)
+                                        method="lstsqr", alpha=alpha,
+                                        fast_mode=False)
 
             # ~37s
+            # G_hat_lasso = estimate_dense_graph(X, max_lag=p_max,
+            #                                    max_T=T,
+            #                                    method="lasso")
+            # G_hat_lasso = estimate_dense_graph(X, max_lag=p_max,
+            #                                    max_T=T,
+            #                                    method="glasso")
+            # G_hat_lasso = estimate_dense_graph(X, max_lag=p_max,
+            #                                    max_T=T,
+            #                                    method="alasso")
+
+            # TODO: Notice that I'm giving it the /real/ p value
             G_hat_lasso = estimate_dense_graph(X, max_lag=p_max,
                                                max_T=T,
-                                               method="lasso")
+                                               method="alasso")
+            # G_hat_lasso = estimate_dense_graph(X, max_lag=p_max,
+            #                                    max_T=T,
+            #                                    method="lasso")
 
             errs_pwgc.update(G, G_hat_pwgc, sv2_true, N_iter, T_iter)
             errs_lasso.update(G, G_hat_lasso, sv2_true, N_iter, T_iter)
@@ -153,9 +170,10 @@ def lasso_comparison(simulation_name, graph_type):
         D_true_errs,
         title=("Test Errors against T (Random {} graph on "
[…]

### 2019-06-21 software/pwgc/gc_methods.py

@@ -14,7 +14,8 @@ from math import ceil
 from scipy import linalg
 from scipy import stats as sps
 from scipy.linalg import toeplitz
-from sklearn.linear_model import LassoLarsIC
+from sklearn.linear_model import LassoLarsIC, LassoLarsCV
+from spams import fistaFlat
 
 from levinson import (lev_durb, whittle_lev_durb)
 
@@ -68,6 +69,9 @@ def estimate_b_lasso(X, y):
     """
     lasso = LassoLarsIC(criterion="bic", fit_intercept=False,
                         normalize=False, precompute=True)
+    # lasso = LassoLarsCV(fit_intercept=False, verbose=False,
+    #                     normalize=True, precompute=True,
+    #                     cv=3, n_jobs=3)
 
     # NOTE: This takes 5 - 10 ms for (5000, 125) matrix X
     lasso.fit(X, y)
@@ -77,10 +81,90 @@ def estimate_b_lasso(X, y):
         return w
 
     # Use lasso only to select the support
-    X_lasso = X[:, [i for i, wi in enumerate(w) if wi != 0]]
+    eps = 1e-9
+    X_lasso = X[:, [i for i, wi in enumerate(w) if abs(wi) >= eps]]
     w_lr = estimate_b_lstsqr(X_lasso, y)
-    w[[i for i, wi in enumerate(w) if wi != 0]] = w_lr
+    w[[i for i, wi in enumerate(w) if abs(wi) >= eps]] = w_lr
     return w
+
+
+def estimate_b_alasso(X, y):
+    """
+    Estimate y_hat = Xb via Adaptive-LASSO
+    and using BIC to choose regularizers.
+
+    y: np.array (T)
+    X: np.array (T x s)
+    """
+    nu = 2.0  # We approximate l_q regression where q = 1 - nu
+
+    # This is an adaptive scheme that uses "piloting weights" to improve
+    # the selection capacity of the lasso.
+    b0 = estimate_b_lstsqr(X, y)
+    w = 1. / np.abs(b0)**nu
+    X_wgt = X / w[None, :]
+
+    lasso = LassoLarsIC(criterion="bic", fit_intercept=False,
+                        normalize=False, precompute=True)
+    # lasso = LassoLarsCV(fit_intercept=False, verbose=False,
+    #                     normalize=True, precompute=True,
+    #                     cv=3, n_jobs=3)
+
+    # NOTE: This takes 5 - 10 ms for (5000, 125) matrix X
[…]

### 2019-06-25 paper/paper.tex
Cleanup on statistical testing portion of paper.

@@
 \begin{align}
-  H_0:&\ \widehat{x}_i^{(p_i)}(t) = \sum_{\tau = 1}^{p_i} b_{ii}(\tau)x_i(t - \tau),\\
-  H_1:&\ \widehat{x}_i^{(p_{ij})}(t) = \sum_{\tau = 1}^{p_{ij}} b_{ii}(\tau)x_i(t - \tau) + \sum_{\tau = 1}^{p_{ij}}b_{ij}(\tau)x_j(t - \tau).
+  H_0:&\ \widehat{x}_i^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau),\\
+  H_1:&\ \widehat{x}_i^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau) + \sum_{\tau = 1}^pb_{ij}(\tau)x_j(t - \tau).
 \end{align}
 
-Under the null hypothesis $H_0$ (i.e. that $x_j \npwgc x_i$) we can
-form the asymptotically $\chi^2(p_{ij})$ distributed test statistic
-\todo{Is this actually $\chi^2$ when $p_{ij}$ is chosen from the data?  This is adaptive!}
+We formulate the statistic
 
@@
   \label{eqn:gc_statistics}
-  F_{ij}(p_i, p_{ij}) = \frac{T}{p_{ij}}\Big(\frac{\xi_i(p_i)}{\xi_{ij}(p_{ij})} - 1\Big),
+  F_{ij}(p) = \frac{T}{p}\Big(\frac{\xi_i(p)}{\xi_{ij}(p)} - 1\Big),
 \end{equation}
 
-where $\xi_i(p_i)$ is the sample mean square of the residuals\footnote{This
-  quantity is often denoted $\widehat{\sigma}$, but we maintain
-  notation from Definition \ref{def:granger_causality}.}
-$x_i(t) - \widehat{x}^{(p_i)}_i(t)$
+where $\xi_i(p)$ is the sample mean square of the
+residuals\footnote{This quantity is often denoted $\widehat{\sigma}$,
+  but we maintain notation from Definition
+  \ref{def:granger_causality}.}  $x_i(t) - \widehat{x}^{(p)}_i(t)$,
 
 \begin{equation*}
-  \xi_i(p_i) = \frac{1}{T - p_i}\sum_{t = p_i + 1}^T (x_i(t) - \widehat{x}_i^{(p_i)}(t))^2,
+  \xi_i(p) = \frac{1}{T - p}\sum_{t = p + 1}^T (x_i(t) - \widehat{x}_i^{(p)}(t))^2,
 \end{equation*}
 
-and similarly for $\xi_{ij}(p_{ij})$.  Under moment conditions $x(t)$
-(\hl{TODO: What assumptions? -- this must also depend on choice of
-  $p_i, p_j$}) we have the convergence in probability:
+and similarly for $\xi_{ij}(p)$.  We test $F_{ij}(p)$ against a
+$\chi^2(p)$ distribution.
+
+If disturbances are Gaussian, we will have the convergence
 
 \begin{equation}
-  F_{ij}(p_i, p_{ij}) \overset{\P}{\rightarrow}
+  F_{ij}(p) \overset{\P}{\rightarrow}
   \left\{
@@
 
-Therefore, we can make the statement that Algorithm
-\ref{alg:finite_pwgc} asymptotically implements Algorithm
-\ref{alg:pwgr}.  \hl{TODO: Modify Algorithm 2 to make this true.}
-\hl{TODO: Enclose this into a theorem}
-
+In our finite sample implementation (see Algorithm
+\ref{alg:finite_pwgc}) we add edges to $\widehat{\gcg}$ in order of
+the decreasing magnitude of $F_{ij}$ instead of proceeding backwards
+through $P_{k - r}$ in Algorithm \ref{alg:pwgr}.  This makes greater
+use of the information provided by the test statistic $F_{ij}$,
+moreover, if $x_i \gc x_j$ and $x_j \gc x_k$, it is expected that
+$F_{kj} > F_{ji}$, thereby providing the same effect as proceeding
+backwards through $P_{k - r}$.
 \subsection{Model Order Selection}
@@
 
-We compare our results against the LASSO as a baseline, estimating the
-filter $\B(z)$ that minimizes $\xi_{\text{LASSO}}$ in the following:
+We compare our results against the LASSO.  Motivated by scaling, we
+split the squared error term into separate terms, one for each group of
+incident edges on a node, and estimate the collection of $n$ incident
+filters $\big\{\B_{ij}(z)\big\}_{j = 1}^n$ that minimizes
+$\xi_i^{\text{LASSO}}$ in the following:
 
@@
   \begin{aligned}
-  \xi_{\text{LASSO}}^{(\lambda)} &= \underset{B}{\text{min}}\ \frac{1}{T}\sum_{t = p + 1}^T||x(t) - \sum_{\tau = 1}^p B(\tau) x(t - \tau)||_2^2 + \lambda \sum_{\tau = 1}^p ||\matvec B(\tau)||_1\\
-  \xi_{\text{LASSO}} &= \underset{\lambda \ge 0}{\text{min}}\ \xi_{\text{LASSO}}^{(\lambda)} + \mathsf{BIC}(B_{\text{LASSO}}^{(\lambda)})\\
+  \xi_i^{\text{LASSO}}(\lambda) &= \underset{B}{\text{min}}\ \frac{1}{T}\sum_{t = p + 1}^T\big(x_i(t) - \sum_{\tau = 1}^p\sum_{j = 1}^n B_{i, j}(\tau) x(t - \tau)\big)^2 + \lambda \sum_{\tau = 1}^p \sum_{j = 1}^n |B_{ij}(\tau)|\\
+  \xi_i^{\text{LASSO}} &= \underset{\lambda \ge 0}{\text{min}}\ \xi_i^{\text{LASSO}}(\lambda) + \mathsf{BIC}\big(B_i^{\text{LASSO}}(\lambda)\big)\\
   \end{aligned}
@@
 
-where $||\matvec B||_1$ emphasizes the application of a vector norm, and
-$\lambda$ is the regularization parameter chosen via the BIC.
+where we are choosing $\lambda$, the regularization parameter, via the BIC.
 
@@
 Algorithm \ref{alg:finite_pwgc} hold.  However, the performance is
-still markedly superior in the case of a more general DAG.
+still markedly superior in the case of a more general DAG.  We would
+conjecture that a DAG having a similar degree of sparsity as an SCG is
+likely to be ``close'' to an SCG.  Figure
+\ref{fig:simulation_results_dense} illustrates the severe (expected)
+degredation in performance as the number of DAG edges increases while
+the number of data samples $T$ remains fixed.  For larger values $q$
+in this plot, the number of edges in the graph is comparable to the
+number of data samples.
 
@@
 edges) using only $T = 500$ data points, granted, an SCG on this many
-nodes is extremely sparse.  Application to denser graphs is explored
-in Figure \ref{fig:simulation_results_dense}, where $n$ and $T$ are
-held fixed while increasing the edge probability $q$, thereby creating
-denser graphs.  For larger values of $q$ in this plot there is a
-comparable number of edges as there are data samples $T$.
+nodes is extremely sparse.  However, it is certainly possible to
+improve the scalability of the LASSO as applied to time series data by
+taking advantage of the block-Toeplitz covariance structure.
 

### 2019-06-25 software/lasso_comparison.py

@@ -81,12 +81,12 @@ class TrackErrors:
 def lasso_comparison(simulation_name, graph_type):
     np.random.seed(0)
     n_nodes, p_lags, p_max = 50, 5, 15
-    alpha, N_iters = 0.05, 1
+    alpha, N_iters = 0.05, 3
 
     # T_iters = list(map(int, np.logspace(2, 3, 20)))
     # T_iters = list(map(int, np.linspace(30, 5000, 20)))
     # T_iters = [50 + 100 * k for k in range(1, 99)]
-    T_iters = [50 + 100 * k for k in range(1, 99, 2)]
+    T_iters = [50 + 100 * k for k in range(1, 99, 1)]
     # T_iters = [20, 40, 80, 100, 150, 250, 500]
 
     # NOTE: We use only X[T:] to estimate the error
@@ -173,10 +173,11 @@ def lasso_comparison(simulation_name, graph_type):
         D_true_errs,
         title=("Test Errors against T (Random {} graph on "
                "$n = {}$ nodes)".format(graph_type, n_nodes)),
-        lasso_title="Adaptive LASSO")
-        # save_file=["../figures/{}_simulation.pdf".format(simulation_name),
-        #            "../figures/jpgs_pngs/"
-        #            "{}_simulation.png".format(simulation_name)])
+        lasso_title="Adaptive LASSO",
+        save_file=["../figures/{}_simulation.pdf".format(simulation_name),
+                   "../figures/jpgs_pngs/"
+                   "{}_simulation.png".format(simulation_name)],
+        show_results=False)
     return
 
 
@@ -231,7 +232,8 @@ def plot_results(D_MCC_pwgc, D_errs_pwgc,
                  D_MCC_lasso, D_errs_lasso,
                  D_true_errs,
                  save_file=None, title="",
-                 lasso_title="LASSO"):
+                 lasso_title="LASSO",
+                 show_results=True):
     true_errs = np.array(D_true_errs["value"], dtype=float)
     t = np.array(D_true_errs["variable"], dtype=float)
 
@@ -257,6 +259,6 @@ def plot_results(D_MCC_pwgc, D_errs_pwgc,
             fig.savefig(save_file)
         else:  # Allow passing a list
             _ = [fig.savefig(f) for f in save_file]
-    plt.show()
-
+    if show_results:
+        plt.show()
     return

### 2019-06-25 software/pwgc/gc_methods.py

@@ -15,6 +15,7 @@ from scipy import linalg
 from scipy import stats as sps
 from scipy.linalg import toeplitz, cho_solve, cho_factor
 from sklearn.linear_model import LassoLarsIC, LassoLarsCV
+from sklearn.preprocessing import StandardScaler
 from spams import fistaFlat
 
 from levinson import (lev_durb, whittle_lev_durb)
@@ -84,6 +85,19 @@ def _estimate_b_lasso(G, X, y):
     return w
 
 
+def _standardize_X(f):
+    def _f(X, *args, **kwargs):
+        _X = X - np.mean(X, axis=0)[None, :]
+        s = np.std(_X, axis=0)[None, :]
+        _X = _X / s
+
+        w = f(_X, *args, **kwargs)
+        w = w / s.ravel()
+        return w
+    return _f
+
+
+@_standardize_X
 def estimate_b_lasso(X, y):
     """
     Estimate y_hat = Xb via LASSO and using BIC to choose regularizers.
@@ -96,7 +110,8 @@ def estimate_b_lasso(X, y):
     return w
 
 
-def estimate_b_alasso(X, y, lmbda_lstsqr=2.0, nu=0.5):
+@_standardize_X
+def estimate_b_alasso(X, y, lmbda_lstsqr=2.0, nu=2.0):
     """
     Estimate y_hat = Xb via Adaptive-LASSO
     and using BIC to choose regularizer.
@@ -397,7 +412,7 @@ def compute_gc_score(xi_i, xi_ij, T, p_lags, F_distr=False):
             F = (T - p_lags) * ((xi_i[:, None] / xi_ij) - 1) / p_lags
         else:
             F = T * ((xi_i[:, None] / xi_ij) - 1) / p_lags
-    F[p_lags ==0] = 0
+    F[p_lags == 0] = 0
     return F
 
 

### 2019-06-25 software/test/test_gc_methods.py

@@ -46,8 +46,8 @@ class TrackErrors:
 
 def example_linear_regression():
     np.random.seed(0)
-    p = 100
-    p0 = 15
+    p = 20
+    p0 = 8
 
     sigma = 2.5
     N_iters = 250
@@ -55,8 +55,8 @@ def example_linear_regression():
     errs_lstsqr, errs_lasso, errs_alasso = [TrackErrors(N_iters, p, p0)
                                             for _ in range(3)]
 
-    sample_points = np.array(list(map(int, np.linspace(10, 1e5, N_iters))))
-    n_test = 1000
+    sample_points = np.array(list(map(int, np.linspace(10, 5e5, N_iters))))
+    n_test = 10000
     for i, n in enumerate(sample_points):
         print("i = {} / {}, n = {}\r".format(i + 1, N_iters, n))
         noise = sigma * np.random.normal(size=n + n_test)
@@ -65,9 +65,7 @@ def example_linear_regression():
         # L = np.random.standard_t(size=(p, p), df=3)
         L = np.random.normal(size=(p, p))  # Make X dependent
         X = np.random.normal(size=(n + n_test, p))
-        X[1:] += 0.5 * np.random.normal() * X[:-1]
         X[2:] += 0.5 * np.random.normal() * X[:-2]
-        X[3:] += 0.5 * np.random.normal() * X[:-3]
         X = X @ L
 
         X = X / np.std(X)

### 2019-06-25 paper/paper.tex
Added some figures.

@@
   \textit{strongly causal} if there is at most 1 directed path between
-  any two nodes.
+  any two nodes.  Strongly Causal Graphs will be referred to as ``SCGs''.
 \end{definition}
@@
 subgraphs.
+
+\begin{example}
+  Though examples of SCGs are easy to construct in theory, should
+  practitioners expect SCGs to arise in application?  While a positive
+  answer to this question is not \textit{necessary} for the concept to
+  be useful, it is certainly sufficient.  Though the answer is likely
+  to depend upon the particular application area, examples appear to
+  be available in biology, in particular, the authors of
+  \cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty}
+  cite an example of the so called ``transcription regulatory network
+  of \textit{E.coli}'', and
+  \cite{learning_genome_scale_regulatory_networks} study a much larger
+  regulatory network of \textit{Saccharomyces cerevisiae}.  These
+  networks, which we reproduce\footnote{Figure \ref{fig:gene_network1}
+    is reproduced under the Creative Commons Attribution
+    Non-Commercial License
+    (\url{http://creativecommons.org/licenses/by-nc/2.5}) and Figure
+    \ref{fig:gene_network2} under the Creative Commons Attribution
+    License (\url{https://creativecommons.org/licenses/by/4.0/})} in
+  figure \ref{fig:gene_networks} appear to have at most a small number
+  of edges which violate the strong-causality condition.
+
+  \begin{figure}[h]
+    \centering
+    \caption{Transcription Regulatory Networks}
+    \label{fig:gene_networks}
+    \begin{subfigure}[b]{0.45\textwidth}
+      \caption{\textit{E.Coli} Network of
+        \cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty}}
+      \label{fig:gene_network1}
+      \includegraphics[width=\linewidth, height=\linewidth]{ecoli_regulatory_network.png}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.45\textwidth}
+      \caption{\textit{Saccharomyces cerevisiae} Network of
+        \cite{learning_genome_scale_regulatory_networks}}
+      \label{fig:gene_network2}
+      \includegraphics[width=\linewidth, height=\linewidth]{huge_gene_network.png}
+    \end{subfigure}
+  \end{figure}
+\end{example}
 

### 2019-06-28 paper/paper.tex
Substantial work on prose in paper + substantial changes to simulation figures.

@@
 \bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
+% \usepackage{apxproof}
+
 % \bibliography{\string~global_bib.bib}
@@
 
-\title{Structure Learning for $\VAR(p)$ Models}
-\author{R. J. Kinnear, R. R. Mazumdar}
+\title{Graph Topological Aspects of Granger Causal Network Learning }
+\author{R. J. Kinnear \\
+  \small\href{mailto:ryan@kinnear.ca}{ryan@kinnear.ca} \\
+  \small\url{https://github.com/RJTK} \and R. R. Mazumdar \\
+  \small\href{mailto:mazum@uwaterloo.ca}{mazum@uwaterloo.ca} }
 
@@
 \maketitle
-\abstract{We study Granger Causality and propose a structure learning
-  heuristic for uncovering a parsimonious representation of large
-  $\mathsf{VAR}(p)$ models.}
+\abstract{We study Granger-causality in the context of wide-sense
+  stationary time series, where our focus is on the topological
+  aspects of the underlying causality graph.  We establish sufficient
+  conditions (in particular, we develop the notion of a ``strongly
+  causal'' graph topology) under which the true causality graph can be
+  recovered via pairwise causality testing alone, and provide examples
+  from the gene regulatory network literature suggesting that our
+  concept of a strongly causal graph may be applicable to this field.
+  We implement and detail finite-sample heuristics derived from our
+  theory, and establish through simulation the efficiency gains (both
+  statistical and computational) which can be obtained (in comparison
+  to LASSO algorithms) when structural assumptions are met.}
+
+\keywords{Granger-causality, causality graph, time series, vector autoregression,
+  LASSO, gene regulatory networks, network learning}
+
+\paragraph{Acknowledgement}
+
+We acknowledge the support of the Natural Sciences and Engineering Research Council of Canada (NSERC), [funding reference number 518418-2018].
+Cette recherche a été financée par le Conseil de recherches en sciences naturelles et en génie du Canada (CRSNG), [numéro de référence 518418-2018].
+
+\clearpage
+
+% \tableofcontents
+% \clearpage
 
@@
 \label{sec:introduction}
-Consider a collection of stochastic processes producing observations
-at discrete time intervals.  Are the underlying processes dependent?
-Can we quantify any of the underlying relationships?  Can the arrow of
-time help us to distinguish a directionality or flow of dependence
-among our observed series?  In this paper we contribute to the
-understanding of the notion of Granger-Causality
-\cite{granger1969investigating} \cite{Granger1980329} as a tool for
-answering these questions.
-
-Heuristically, we state that if an event '$A$' occuring strictly prior
-in time to an even '$B$' and $A$ provides us with exclusive (that is,
-not available anywhere else) information about $B$, then $A$ must have
-had a causal impact on $B$.  This is an entirely model-free notion of
-cause, and instead leverages the intuition that a cause must precede
-it's effect.
-
-In practice, Granger's notion of causation it not a convincing test
-for \textit{true} causation, since our statements about causation are
-highly dependent upon the data that we are able to observe.  Rather,
-the interpretation of Granger causality is as a means of uncovering a plausible
-flow of ``information'' or ``energy'' through some underlying graph of
-interactions.  Though this graph cannot be observed directly, we will
-infer it's presence as a latent structure among our observed time
-series data.  This intuition is levereged in a variety of applications
+% Consider a collection of stochastic processes producing observations
+% at discrete time intervals.  Are the underlying processes dependent?
+% Can we quantify any of the underlying relationships?  Can the arrow of
+% time help us to distinguish a directionality or flow of dependence
+% among our observed series?  In this paper we contribute to the
+% understanding of the notion of Granger-Causality
+% \cite{granger1969investigating} \cite{Granger1980329} as a tool for
+% answering these questions.
+
+% Heuristically, we state that if an event '$A$' occuring strictly prior
+% in time to an even '$B$' and $A$ provides us with exclusive (that is,
+% not available anywhere else) information about $B$, then $A$ must have
+% had a causal impact on $B$.  This is an entirely model-free notion of
+% cause, and instead leverages the intuition that a cause must precede
+% it's effect.
+
+In this paper we study the notion of Granger-causality
+\cite{granger1969investigating} \cite{Granger1980329} as a means of
+uncovering an underlying causal structure in multivariate time series.
+Though the underlying causality graph cannot be observed directly, we
+will infer it's presence as a latent structure among our observed time
+series data.  This notion is leveraged in a variety of applications
 e.g. in Neuroscience as a means of recovering interactions amongst
@@
 expression networks \cite{Fujita2007},
+\cite{methods_for_inferring_gene_regulatory_networks_from_time_series_expression_data},
 \cite{grouped_graphical_granger_modelling_for_gene_expression_regulatory_networks_discovery},
-\cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty},
+\cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty};
 and power system design \cite{Misyrlis2016450}, \cite{yuan2014root}.
 
-\hl{TODO: From the gene network papers, it looks like these gene
-  networks may actually have strongly-causal structure.  i.e. we can
-  highlight the fact that our deeper understanding of a special
-  problem case may in fact have significant applicability in these
-  application areas.}
-
-Other citations: support recovery consistency for the weighted lasso
-in univariate AR(p) models:
-
-Finding the ``best'' graph structure consistent with observed data is
-generally an extremely challenging problem (i.e. this is a best subset
-selection problem, see \cite{bss_mio}, \cite{hastie_bss_comp}), though
-the comparison of quality between different structures, and hence the
-notion of ``best'' needs to be quantified.  In applications where we
-are interested merely in minimizing the mean squared error of a linear
-one-step-ahead predictor, then we will naturally desire an entirely
-dense graph of connections, since each edge can only serve to reduce
-estimation error.  However, since the number of edges scales
-quadratically in $n$, it is imperative to infer a sparse causality
-graph for large systems, both to avoid overfitting observed data, as
-well as to aid the interpretability of the results.
-
-In \cite{bach2004learning} the authors apply a local search heuristic
-to the Whittle likelihood with an AIC penalization.  The local search
-heuristic where at each iteration an edge is either added, removed, or
-reversed is a common approach to combinatorial optimization due to
-it's simplicity, but is liable to get stuck in shallow local minima.
+Granger-causality can generally be formulated by searching for the
+``best'' graph structure consistent with observed data, which is in
+general an extremely challenging problem (i.e. it may be framed as a
+best subset selection problem, see \cite{bss_mio},
+\cite{hastie_bss_comp}), moreover, the comparison of quality between
+different structures, and hence the notion of ``best'' needs
+qualification.  In applications where we are interested merely in
+minimizing the mean squared error of a linear one-step-ahead
+predictor, then we will be satisfied with an entirely dense graph of
+connections, since each edge can only serve to reduce estimation
+error.  However, since the number of edges scales quadratically in $n$
+(the number of nodes) it becomes imperative to infer a sparse
+causality graph for large systems, both to avoid overfitting observed
+data, as well as to aid the interpretability of the results.
+
+A fairly early approach to the problem in the context of large systems
+is provided by \cite{bach2004learning}, where the authors apply a
+local search heuristic to the Whittle likelihood with an AIC
+penalization.  The local search heuristic where at each iteration an
+edge is either added, removed, or reversed is a common approach to
+combinatorial optimization due to it's simplicity, but is liable to
+get stuck in shallow local minima.
 
@@
 dependence decay \cite{basu2015} \cite{wong2016lasso}
-\cite{autoregressive_process_modelling_via_the_lasso_procedure}.
-Variations on the LASSO have been applied in a number of different
-contexts \cite{DBLP:journals/corr/HallacPBL17} \cite{haufe2008sparse}
+\cite{autoregressive_process_modelling_via_the_lasso_procedure}, and
+variations on the LASSO have been applied in a number of different
+time series contexts as well as Granger-causality
+\cite{DBLP:journals/corr/HallacPBL17} \cite{haufe2008sparse}
 \cite{bolstad2011causal} \cite{he2013stationary}
@@
 One of the key improvements to the original LASSO algorithm is the
-adaptive (i.e. weighted) adaLASSO \cite{adaptive_lasso_zou2006}, for
-which oracle results are established.  The crux of the matter for time
-series analysis is to craft a good weighting scheme, see e.g.
-\cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty}.
-Success of the LASSO has lead to the famous ``bet on sparsity''
-principle -- that is, for large problems and finite sources of data,
-you might as well bet that a sparse model is correct, because if it
-isn't, there is no hope.
-
+adaptive (i.e. weighted) ``adaLASSO'' \cite{adaptive_lasso_zou2006},
+for which oracle results (i.e. asymptotic support recovery) are
+established under less restrictive conditions than for the vanilla
+LASSO \cite{wainwright2009sharp}.%   Success of the LASSO has lead to
+% the famous ``bet on sparsity'' principle ``Use a procedure that does
+% well in sparse problems, since no procedure does well in dense
+% problems.''  \cite{tibshirani2015statistical}.
+
+\subsection{Contributions}
 In the context of time series data, sparsity assumptions remain
@@
 notion of causality graph topology which we term ``strongly causal''
-and show that $VAR$ models having this structure satisfy naturaal
+and show that $\VAR$ models having this structure satisfy natural
 intuitive notions of ``information flow'' through the graph.
-Moreover, we show that such graphs are perfectly (asymptotically)
-recoverable with only \textit{pairwise} Granger-causality tests.  Our
-finite sample results are based on simulations, where we show
-promising results for these graphs topologies where our algorithm
-performs substantially better in our simulation setup than do
-competing LASSO algorithms, even for graphs that do not exactly
-satisfy our strongly-causal topology assumptions.
-
-\todo{What graphs in practice may be strongly causal -- see the gene network papers.}
-
-The contributions of this paper are as follows: firstly, in section
-\ref{sec:theory} we study \textit{pairwise} Granger-causality
+Moreover, we show that such graphs are perfectly recoverable with only
+\textit{pairwise} Granger-causality tests, which would otherwise suffer
+from serious confounding problems.  Our finite sample results are
+based on simulations, where we show promising results for these graphs
+topologies where our algorithm performs substantially better in our
+simulation setup than do competing LASSO algorithms, even for graphs
+that do not exactly satisfy our strongly-causal topology assumptions.
+
+In the case of gene expression networks, we show examples from the
+literature which suggest our concept of a ``strongly causal graph''
+topology may have application in this field (see Section
+\ref{sec:strongly_causal_graphs})
+
+The principle contributions of this paper are as follows: firstly, in
+section \ref{sec:theory} we study \textit{pairwise} Granger-causality
 relations, providing novel theorems connecting the structure of the
@@
 well as a characterization of the sparsity pattern of matrices arising
-in the Wold decomposition.  We establish sufficient conditions
-(sections \ref{sec:strongly_causal_graphs},
+in the Wold decomposition, generalizing in some sense the notion of
+``feedback-free'' processes studied by \cite{caines1975feedback} in
+close connection with Granger-causality.  We establish sufficient
+conditions (sections \ref{sec:strongly_causal_graphs},
 \ref{sec:persistent_systems}) under which a fully conditional
 Granger-causality graph can be recovered from pairwise tests alone
-(section \ref{sec:pairwise_algorithm}).  Secondly, we propose in
-section \ref{sec:structure_learning} a graph search heuristic which
-implements our theoretical results to finite data samples.  Our
-heuristics are compared against the LASSO algorithm in Section
+(sec \ref{sec:pairwise_algorithm}).  Secondly, we propose in section
+\ref{sec:structure_learning} a graph search heuristic which implements
+our theoretical results to finite data samples, specifying and
+summarizing appropriate methods for hypothesis testing (section
+\ref{sec:pairwise_hypothesis_testing}), model order selection (section
+\ref{sec:model_order_selection}), computationally efficient estimation
+(section \ref{sec:efficient_model_estimation}), and error rate
+controls (section \ref{sec:error_rate_control}).  Our heuristics are
+compared against the LASSO algorithm in Section
 \ref{sec:empirical_evaluation}.  We stress the scalability of our
 algorithm which is capable of comfortably handling hundreds or
-thousands of nodes on a single machine.  In section
-\ref{sec:application} we develop an example application (\hl{more
-  detail...}).  And provide concluding remarks on further open
-problems in our conclusion (section \ref{sec:conclusion}).
-
-\todo{Simulations with very small T (i.e. T = 20)}
+thousands of nodes on a single machine, as opposed to standard LASSO
+algorithms which do not take advantage of the special structure
+associated with stationary time series data.  In section
+\ref{sec:application} we develop an example application.  And provide
+concluding remarks on further open problems in our concluding section
+\ref{sec:conclusion}.  \todo{do an application}
 
@@
 are themselves Hilbert.  We will denote the spaces generated in
-analagous ways by particular components of $x$ as e.g.
+analogous ways by particular components of $x$ as e.g.
 $\H_t^{(i, j)}$, $\H_t^{i}$ or by all but particular components as
@@
   $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi,
-  \pi]$}, and $w(t)$ is an uncorrelated sequence and $A(0) = I$.  We
+  \pi]$}, $v(t)$ is an uncorrelated sequence and $A(0) = I$.  We
 will assume that $c(t) = 0$, which in practice is to say that it has
 been accurately estimated and removed.  We additionally require that
-this representation can be inverted to yield the $VAR(\infty)$ form
+this representation can be inverted to yield the $\VAR(\infty)$ form
 
@@
 $\B(z) \defeq \sum_{\tau = 1}^\infty B(\tau)z^{-\tau}$ where the
-operator $z^{-1}$ is the backshift operator acting on
+operator $z^{-1}$ is the back shift operator acting on
 $\ell_2^n(\Omega, \mathcal{F}, \mathbb{P})$, that is:
@@
   The process $x(t)$ is an $n$ dimensional wide sense stationary
-  process having invertible $VAR(\infty)$ representation
+  process having invertible $\VAR(\infty)$ representation
   \eqref{eqn:ar_representation} where $v(t)$ is sequentially
@@
 \begin{proof}
-  Apply the Gram-schmidt process and directly calculate the projections.
+  Apply the Gram-Schmidt process and directly calculate the projections.
 \end{proof}
@@
 
-\begin{definition}[Grandparents, ancestors]
+\begin{definition}[Parents, Grandparents, Ancestors]
+  A node $j$ is a \textit{parent} of node $i$ if $(j, i) \in \gcge$.
+  The set of all $i$'s parents will be denoted $\pa{i}$.
+
   The set of level $\ell$ \textit{grandparents} of node $i$, denoted
@@
   only if there is a \textit{directed path} of length $\ell$ in $\gcg$
-  from $j$ to $i$.  Note that a node can be it's own grandparent
-  unless $\gcg$ is a DAG.  Clearly, $\pa{i} = \gpn{1}{i}$.
+  from $j$ to $i$.  Clearly, $\pa{i} = \gpn{1}{i}$.
 
@@
   Note that since loops are not considered paths, it is not possible
-  for a node to be it's own parent.
+  for a node to be it's own parent.  However, unless $\gcg$ is a DAG,
+  a node can be it's own ancestor.  We will ocassionally need to
+  explicitly exclude $i$ from $\anc{i}$, in which case we will write
+  $\anc{i}\setminus \{i\}$.
 \end{definition}
@@
 
-\begin{definition}[Causality graph, Parents]
+\begin{definition}[Causality graph]
   We define the Granger-causality graph $\gcg = ([n], \gcge)$ to be the directed
@@
 \textit{strictly causal} LTI filter $\B_{ij}(z)$ (see eqn
-\eqref{eqn:filter_action}).  Thence, the matrix $\B(z)$ is analagous
+\eqref{eqn:filter_action}).  Thence, the matrix $\B(z)$ is analogous
 to a \textit{weighted adjacency matrix}\footnote{We are using the
@@
   $(i, j)$.  In our case, the sparsity pattern of $\B_{ij}$ is the
-  \textit{transposed} adjacency matrix.} for the graph $\gcg$.  And,
+  \textit{transposed} conventional adjacency matrix.} for the graph $\gcg$.  And,
 in the same way that the $k^{\text{th}}$ power of an adjacency matrix
@@
 relationship between each node and it's parent nodes, the following
-proposition is analagous to the definition of feedback free processes
+proposition is analogous to the definition of feedback free processes
 of \cite{caines1975feedback}:
@@
 $x_i(t) = \sum_{\tau = 0}^\infty \sum_{j = 1}^n A_{ij}(\tau)v_j(t -
-\tau)$.  The proposition states that $\A_{ij}(z) = 0$ unless
-$j \in \anc{i}$.
+\tau)$.  The proposition states that if $j \not \in \anc{i}$ then
+$\A_{ij}(z) = 0$, however the implication
+$j \in \anc{i} \implies \A_{ij}(z)$ need not hold.
 
@@
   Equation \eqref{eqn:parent_expansion} is immediate from the
-  $VAR(\infty)$ representation of \eqref{eqn:ar_representation} and
+  $\VAR(\infty)$ representation of \eqref{eqn:ar_representation} and
   Theorem \ref{thm:granger_causality_equivalences}, we are left to
@@
   Firstly, notice that every $u \in \big(\pa{j} \setminus \{i\}\big)$
-  necesssarily inherits these same two properties.  This follows since
+  necessarily inherits these same two properties.  This follows since
   if we also had $u \in \anc{i}$ then $u \in \anc{i} \cap \anc{j}$ so that every
@@
   $j \pwgc i$, then $i \not\in \anc{j}$ and $j \not\in \anc{i}$.  In
-  particular, a pairwise bidirectional edge indicates the absense of
+  particular, a pairwise bidirectional edge indicates the absence of
   any edge in $\gcg$.
@@
 % \begin{lemma}
-%   If $\gcg$ is strongly causal, then any $\VAR(p)$ system on $\gcg$ is stable if and only if
+%   If $\gcg$ is strongly causal, then any $\\VAR(p)$ system on $\gcg$ is stable if and only if
 %   $B_{ii}(z)$ is stable for every $i = 1, \ldots, n$.
@@
 \begin{definition}[Lag Function]
-  Given a caual filter $\B(z) = \sum_{\tau = 0}^\infty b(\tau)z^{-\tau}$
+  Given a causal filter $\B(z) = \sum_{\tau = 0}^\infty b(\tau)z^{-\tau}$
   define
@@
   i.e. the ``first'' and ``last'' coefficients of the filter $\B(z)$,
-  where $\tau_\infty(\B) = \infty$ if the filter has an infinite
+  where $\tau_\infty(\B) \defeq \infty$ if the filter has an infinite
   impulse response.
@@
 \begin{remark}
-  There are a multitude of conditions which would guarantee $x(t)$ is
-  persistent, for innstance, any ARMA system with a non-zero
-  autoregressive component is likely to be persistent due to recursion
-  in the coefficients naturally lead to an $\mathsf{MA}(\infty)$
-  representation.
-
-  In particular, any system with a finite impulse response
-  (i.e. $\mathsf{VAR}(q)$ models) is likely to be persistent.  These
-  models are common in applications of Granger-causality.
+  In the context of Granger-causality, ``most'' systems should be
+  persistent.  In particular, $\mathsf{\VAR}(p)$ models are likely to
+  be persistent since these naturally result in equivalent an
+  $\mathsf{MA}(\infty)$ representation, except for the pathological
+  case where $\B(z)$ is Nilpotent.
 
@@
 
-\hl{TODO: Show formally that any $\VAR(p)$ model is persistent.}
+\begin{example}
+  process $x(t)$ generated by the $\VAR(1)$ model\footnote{Recall that
+    any $\VAR(p)$ model with $p < \infty$ can be written as a
+    $\VAR(1)$ model, so we lose little generality in considering this
+    case.}  having $\B(z) = Bz^{-1}$ where $B$, we will examine
+  conditions that guarantee persistence.  Pick any
+  $i \in [n], j \in \anc{i} \setminus \{i\}$, then the stability of
+  $B$ allows us to write
+
+  \begin{equation*}
+    \A(z) = \sum_{k = 0}^\infty B^k z^{-k},
+  \end{equation*}
+
+  whereby we see that $\exists k > 0$ such that $[B^k]_{ij} \ne 0$
+  (since $j \in \anc{i}$).  Then consider
+
+  \begin{equation*}
+    \begin{aligned}
+      e_i^\T B^{rk} e_j &\overset{(a)}{=} \big((P^\T e_i)^\T J^{rk} P^{-1}e_j\big)\\
+      &= \tr [(P^\T e_i)^\T J^{rk} P^{-1}e_j]\\
+      &\overset{(b)}{=} \tr [(J^{rk}) (v u^\T)],
+    \end{aligned}
+  \end{equation*}
+
+  where $(a)$ utilizes the Jordan Normal Form of $B$, and $(b)$
+  denotes $u = P^\T e_i$ and $v = P^{-1}e_j$.  In order for
+  $\tau_\infty(\A_{ij}) < \infty$, there must be some $N > 1$ such
+  that $\forall r \ge N$, the above term is $0$.  This may be the case
+  for instance if $B$ is a nilpotent matrix.
+
+  Let us suppose now that $B$ is diagonalizable (i.e. $J$ is a
+  diagonal matrix) with at least $2$ distinct eigenvalues; in this
+  case $B$ is also \textit{not} nilpotent.  We can then rewrite the
+  above as
+
+  \begin{equation*}
+    f(r) \defeq \tr [(J^{rk}) (v u^\T)] = \sum_{\nu = 1}^n \lambda_\nu ^{rk} v_\nu u_\nu \defeq \sum_{\nu = 1}^n \lambda_\nu^{rk} \beta_\nu
+  \end{equation*}
+
+  where $\lambda_\nu$ denotes the eigenvalues of $B$ and
+  $\beta_\nu = u_\nu v_\nu$.  Note that $f(0) = 0$ since $i \ne j$,
+  $u$ is a row of $P$ and $v$ is a column of $P^{-1}$.  Moreover,
+  $f(1) \ne 0$ by hypothesis.  But, in order for
+  $f(r) = 0\ \forall r \ge N$, it would need to be the case that
+
+  \begin{equation*}
+    \Dg(\lambda)^r \lambda = Vz
+  \end{equation*}
+
+  had a solution in $z$ for every $r \ge N$, where $V$ is an
+  $n \times n - 1$ full-rank matrix spanning the nullspace of $\beta$.
+  That is, iterates of $\Dg(\lambda)$ applied to $\lambda$ would need
+  to remain inside $\mathcal{N}(\beta)$.  This would imply that
+
+  \begin{equation*}
+    VV^\dagger \lambda^{r + 1} = \lambda^{r + 1},
+  \end{equation*}
+
+  i.e. that $\lambda^{r + 1}$ is an eigenvector of $VV^\dagger$ for an
+  infinite number of integers $r$.  However, since there can only be a
+  finite number of (unit length) eigenvectors, this cannot be the case
+  unless every eigenvalue $(\lambda_1, \ldots, \lambda_n)$ were equal.
+
+  \todo{This may remain true for any matrix with at least 2 distinct eigenvalues}
+  We can conclude by stating that a $\VAR(1)$ system corresponding to
+  matrix $B$ is guaranteed to be persistent if $B$ is diagonalizable
+  and has more than $1$ distinct eigenvalue.
+\end{example}
 
@@
   which can be expanded by equation \eqref{eqn:ancestor_expansion} to
-  obtain (after cancellinng all ancestors of $i$ other than $k$)
+  obtain (after cancelling all ancestors of $i$ other than $k$)
 
@@
 
-  Evidently, (\hl{This is clearly not immediately evident}) this is
-  $0$ for each $F$ if and only if
+  \todo{This is clearly not immediately evident}
+
+  Evidently, this is $0$ for each $F$ if and only if
   $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$ and
   $\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} = 0$, which by
-  Collary \ref{cor:time_lag_cancellation} occurs if and only if
+  Corollary \ref{cor:time_lag_cancellation} occurs if and only if
   $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$, which is impossible
-  since in general $\tau_0(\A_{jk}) < \infty$ and for a persistent system
-  $\tau_\infty(\A_{ik}) = \infty$.
+  since in general $\tau_0(\A_{jk}) < \infty$ and for a persistent
+  system $\tau_\infty(\A_{ik}) = \infty$.
 \end{proof}
@@
     \Input{Pairwise Granger-causality relations between a persistent
-      processe of dimension $n$ whose joint Granger-causality
+      process of dimension $n$ whose joint Granger-causality
       relations are known to form a strongly causal DAG $\gcg$.}
@@
     $W \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$  \texttt{\# candidate edges}\\
-    $P_0 \leftarrow \{i \in S_0\ |\ \forall s \in S_0\ (s, i) \not\in W\}$  \texttt{\# parentless nodes}\\
+    $P_0 \leftarrow \{i \in S_0\ |\ \forall s \in S_0\ (s, i) \not\in W\}$  \texttt{\# parent-less nodes}\\
     \While{$S_{k - 1} \ne \emptyset$}{
@@
 
-  Convsersely, suppose $(i, j) \in W$.  Then since $j \npwgc i$
+  Conversely, suppose $(i, j) \in W$.  Then since $j \npwgc i$
   Proposition \ref{prop:persistence_converse} ensures that $(j, i)$
@@
   $i \npwgc s$ so that by persistence and Proposition
-  \ref{prop:persistence_converse} there cannot be a counfounder of
+  \ref{prop:persistence_converse} there cannot be a confounder of
   $(s, i)$ (otherwise $i \pwgc s$) so then by Proposition
@@
   induction that $(u, j) \in \bigcup_{\ell = 1}^{r - 1}D_{k\ell}$.
-  Morevoer, sice $d(u) < k$ (otherwise $d(j) > k$) each edge in
+  Moreover, since $d(u) < k$ (otherwise $d(j) > k$) each edge in
   the $\gcgpath{i}{u}$ path must already be in $E_{k - 1}$, and so
@@
 where $\widehat{\Sigma}_{ij}(p)$ is the $2 \times 2$ residual
-covariance matrix for the $VAR(p)$ model of $(x_i(t), x_j(t))$.  The
+covariance matrix for the $\VAR(p)$ model of $(x_i(t), x_j(t))$.  The
 bivariate errors $\xi_{ij}(p)$ and $\xi_{ji}(p)$ are the diagonal
@@
 variance $\sigma_i^2 = 1/2 + r_i$ where $r_i \sim \text{exp}(1/2)$.
-We set $p_{\text{max}} = 15$.  Results and representitive graphs are
+We set $p_{\text{max}} = 15$.  Results and representative graphs are
 collected in Figure \ref{fig:simulation_results}.
@@
   \centering
-  \caption{Simulation Results and Representitive Random Graph Topologies}
-  \label{fig:simulation_results}
-
-  {\scriptsize Comparison of PWGC and LASSO for $VAR(p)$ model
-    estimation.  The MCC is measured on the left axis in blue with
-    circle markers, and the relative log mean-squared prediction error
-    $\frac{\ln\ \tr \widehat{\Sigma}_v}{\ln\ \tr \Sigma_v}$ on the
-    right axis in red with triangle markers.  We use $q$ to denote the
-    edge probability in a random Erdos-Reny graph (modified to form a
-    DAG).  Solid lines indicate ad-hoc regression estimates for
-    visualization purposes only.
-
-    Matthew's Correlation Coefficient (MCC)
-    \cite{matthews1975comparison} synthesizes the confusion matrix
-    into a single score appropriate for unbalanced labels and is
-    calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
-    performance, $0$ being the performance of random guessing, and
-    $-1$ being perfectly opposed.
-  }
-
+  \caption{Representative Random Graph Topologies}
+  \label{fig:random_graph_topologies}
   \begin{subfigure}[b]{0.3\textwidth}
@@
   \end{subfigure}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \caption{Simulation Results}
+  \label{fig:simulation_results}
+
+  {\scriptsize Comparison of PWGC and LASSO for $\VAR(p)$ model
+    estimation.  The MCC is measured on the left axis in blue with
+    circle markers, and the relative log mean-squared prediction error
+    $\frac{\ln\ \tr \widehat{\Sigma}_v}{\ln\ \tr \Sigma_v}$ on the
+    right axis in red with triangle markers.  We use $q$ to denote the
+    edge probability in a random Erdos-Reny graph (modified to form a
+    DAG).  Solid lines indicate ad-hoc regression estimates for
+    visualization purposes only.
+
+    Matthew's Correlation Coefficient (MCC)
+    \cite{matthews1975comparison} synthesizes the confusion matrix
+    into a single score appropriate for unbalanced labels and is
+    calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
+    performance, $0$ being the performance of random guessing, and
+    $-1$ being perfectly opposed.
+  }
 
   \begin{subfigure}[b]{0.45\textwidth}
-    \caption{PWGC vs LASSO (SCG)}
+    \caption{PWGC vs adaLASSO (SCG)}
     \label{fig:simulation_results_comparison1}
-    \includegraphics[width=\linewidth]{random_scg_simulation.pdf}
+    \includegraphics[width=\linewidth]{new_lasso_comparison_scg_simulation.pdf}
     % \missingfigure{}
@@
   \begin{subfigure}[b]{0.45\textwidth}
-    \caption{PWGC vs LASSO (DAG, $q = \frac{2}{n}$)}
+    \caption{PWGC vs adaLASSO (DAG, $q = \frac{2}{n}$)}
     \label{fig:simulation_results_comparison2}
-    \includegraphics[width=\linewidth]{random_dag_simulation.pdf}
+    \includegraphics[width=\linewidth]{new_lasso_comparison_dag_simulation.pdf}
     % \missingfigure{}
@@
     \label{fig:simulation_results_dense}
-    \includegraphics[width=\linewidth]{dag_increasing_q_simulation.pdf}
+    \includegraphics[width=\linewidth]{dag_increasing_q_small_T_alasso_simulation.pdf}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.45\textwidth}
+    \caption{MCC Comparison for $T \le 100$}
+    \label{fig:mcc_comparison}
+    \includegraphics[width=\linewidth]{new_mcc_comparison000.pdf}
   \end{subfigure}
@@
 \ref{fig:simulation_results_dense} illustrates the severe (expected)
-degredation in performance as the number of DAG edges increases while
+degradation in performance as the number of DAG edges increases while
 the number of data samples $T$ remains fixed.  For larger values $q$
@@
 multiple machines.  Figure \ref{fig:simulation_results_scaling} is a
-demonstrating of this scalability, where we are able to estimate
+demonstration of this scalability, where we are able to estimate
 graphs having over $1500$ nodes (over $2.25 \times 10 ^6$ possible
@@
 \label{sec:conclusion}
-% Extensions to transfer entropy?.  Case of time varying statistics?.  Develop sparse estimators for SS models (\cite{barnett2015granger})?.  Apply MIO for the Whittle relaxation?
+In this paper we have argued that considering particular topological
+properties of Granger-causality networks can provide substantial
+theoretical insights as well as computational benefits.  In
+particular, the notion of a strongly-causal graph has been exploited
+to establish conditions under which pairwise causality testing alone
+is sufficient for recovering a complete Granger-causality graph.
+Moreover, examples from the literature suggest that such topological
+assumptions may be reasonable in some applications and secondly, even
+when the strong-causality assumption is not met, we have provided some
+empirical evidence to suggest that our pairwise testing algorithm PWGC
+can still outperform the LASSO and adaLASSO, both of which are
+commonly employed in applications.  Application to more general graphs
+may also be within reach via an iterative graph building process,
+i.e. estimating a sequence of SCGs via PWGC applied to the residuals
+of the previous model.
+
+We emphasize that the causality graph topology is one of the key
+defining features of time series analysis in comparison to standard
+multivariate regression and therefore advocate for further study of
+how different topological assumptions may impact the recovery of
+causality graphs.  For example, are there provable guarantees on the
+error rate of PWGC when applied to non strongly-causal graphs?
+Computationally, can constraint systems or cunning adaptive weighting
+schemes impose useful prior knowledge about graph topology for the
+LASSO algorithm?  In another direction, it is known that the more
+general notion of transfer entropy reduces to Granger-causality in the
+case of Gaussian data \cite{barnett2009granger}, do our results extend
+to causality networks defined via transfer entropy?  Finally, the work
+of \cite{barnett2015granger} has established the superiority of
+Granger-causality testing in state space models (as opposed to pure
+autoregressions) in many cases.  Combining this work with our PWGC
+algorithm (by modifying the approach described in Section
+\ref{sec:pairwise_hypothesis_testing}) therefore is likely to enable
+application to very large networks of time series data which are not
+well approximated by finite $\VAR(p)$ models.
 

### 2019-06-28 software/lasso_comparison.py

@@ -8,7 +8,7 @@ import numpy as np
 
 import statsmodels.api as sm
 from statsmodels.sandbox.regression.predstd import wls_prediction_std
-
+from plotting_helpers import COLOR1, COLOR2, plotting_model
 
 from matplotlib import rc as mpl_rc
 font = {"family" : "normal",
@@ -16,7 +16,7 @@ font = {"family" : "normal",
         "size"   : 22}
 
 mpl_rc("font", **font)
-mpl_rc("text", usetex=True)
+mpl_rc("text", usetex=True)  # TODO: Fix my fonts
 
 from pwgc.var_system import (random_gnp_dag, drive_gcg, get_errors,
                              get_estimation_errors, random_scg,
@@ -78,15 +78,151 @@ class TrackErrors:
         return D_errs, D_true_errs, D_MCC
 
 
+def simple_lasso():
+    np.random.seed(0)
+    n_nodes, p_lags, p_max = 25, 2, 3
+
+    T_iters = np.array(list(map(int, np.linspace(50, 20000, 150))))
+
+    N_edges = (len(random_scg(n_nodes, p_lags, pole_rad=0.75).edges) -
+               n_nodes)  # Due to self loops
+    edge_prob = 2 * N_edges / (n_nodes * (n_nodes - 1))
+
+    random_graph = lambda: random_gnp_dag(
+        n_nodes, p_lags, pole_rad=0.75,
+        edge_prob=edge_prob)
+
+    def make_test_data(T):
+        sv2_true = 0.5 + np.random.exponential(0.5, size=n_nodes)
+        G = random_graph()
+        drive_gcg(G, 500 + min(2 * T, T + 15000), sv2_true, filter_attr="b(z)")
+        X = get_X(G)
+        X = X - np.mean(X, axis=0)[None, :]
+        X = X
+        return G, X, sv2_true
+
+    errs_lasso = TrackErrors(1, T_iters, n_nodes)
+
+    N_iter = 0
+    for T_iter, T in enumerate(T_iters):
+        print("T[{} / {}]".format(
+            T_iter + 1, len(T_iters)),
+            end="\r")
+        G, X, sv2_true = make_test_data(T)
+        G_hat_lasso = estimate_dense_graph(X, max_lag=p_max,
+                                           max_T=T,
+                                           method="alasso")
+        errs_lasso.update(G, G_hat_lasso, sv2_true, N_iter, T_iter)
+
[…]

### 2019-06-28 software/pwgc/gc_methods.py

@@ -14,7 +14,7 @@ from math import ceil
 from scipy import linalg
 from scipy import stats as sps
 from scipy.linalg import toeplitz, cho_solve, cho_factor
-from sklearn.linear_model import LassoLarsIC, LassoLarsCV
+from sklearn.linear_model import LassoLarsIC, LassoLarsCV, Lasso
 from sklearn.preprocessing import StandardScaler
 from spams import fistaFlat
 
@@ -286,15 +286,17 @@ def estimate_B(G, max_lag=10, copy_G=False,
         # Form linear regression matrices
         X, y = form_Xy(X_raw, y_raw, p=max_lag)
 
+        X_maxT, y_maxT = X[:max_T], y[:max_T]
+
         # Estimate
         if method == "lasso":
-            b = estimate_b_lasso(X[:max_T], y[:max_T])
+            b = estimate_b_lasso(X_maxT, y_maxT)
         elif method == "glasso":
-            b = estimate_b_glasso(X, y, p=max_lag, T_train=max_T)
+            b = estimate_b_glasso(X_maxT, y_maxT, p=max_lag, T_train=max_T)
         elif method == "alasso":
-            b = estimate_b_alasso(X, y)
+            b = estimate_b_alasso(X_maxT, y_maxT, nu=1.5)
         elif method == "lstsqr":
-            b = estimate_b_lstsqr(X[:max_T], y[:max_T])
+            b = estimate_b_lstsqr(X_maxT, y_maxT)
         else:
             raise AssertionError("Bad method but should deal with it earlier!")
 
@@ -504,7 +506,7 @@ def normalize_gc_score(F, p, T=None, F_distr=False):
     if F_distr:
         if T is None:
             raise ValueError("If F_distr=True, we also require T.")
-        sps.f.cdf(F, p, T - p)
+        F = sps.f.cdf(F, p, T - p)
     else:
         F = sps.chi2.cdf(F, p)
     F[np.isnan(F)] = 0.0
@@ -803,6 +805,29 @@ def estimate_dense_graph(X, max_lag=10,
     estimate_B(G, max_lag, copy_G=False, max_T=max_T, method=method)
     G = remove_zero_filters(G, "b_hat(z)", copy_G=False)
     return G
+
+
+def pwgc_estimate_graph(X, max_lags=10, alpha=0.05,
+                        method="lstsqr"):
+    T, n = X.shape
+
[…]

### 2019-06-28 software/simulation_increasing_n.py

@@ -7,6 +7,9 @@ import pandas as pd
 import numpy as np
 import statsmodels.api as sm
 
+from plotting_helpers import (COLOR3, COLOR4, COLOR5,
+                              plotting_model)
+
 from matplotlib import rc as mpl_rc
 font = {"family" : "normal",
         "weight" : "bold",
@@ -63,88 +66,103 @@ class TrackErrors:
         return MCC, self.b_errs, self.errs, self.true_errs
 
 
-def pwgc_increasing_n(T, simulation_name, graph_type):
+def pwgc_increasing_n(T, simulation_name):
     np.random.seed(0)
     alpha, p_lags, p_max = 0.05, 5, 15
     T_max = 2 * T
 
-    n_iters = np.arange(10, 2000, 20)
-    N_reps = 3
+    n_iters = np.arange(10, 1500, 10)
+    N_reps = 2
 
-    if graph_type == "SCG":
-        def random_graph(n):
-            return random_scg(n, p_lags, pole_rad=0.75)
-    elif graph_type == "DAG":
-        def random_graph(n):
-            N_edges = (len(random_scg(n, p_lags,
-                                      pole_rad=0.75).edges) -
-                       n)  # Due to self loops
-            edge_prob = 2 * N_edges / (n * (n - 1))
-            return random_gnp_dag(n, p_lags, pole_rad=0.75,
-                                  edge_prob=edge_prob)
-    else:
-        raise NotImplementedError("Only 'SCG' and 'DAG' graph "
-                                  "types are available")
+    def random_graph_scg(n):
+        return random_scg(n, p_lags, pole_rad=0.75)
 
-    def make_test_data(n):
+    def random_graph_dag_q2(n):
+        N_edges = (len(random_scg(n, p_lags,
+                                  pole_rad=0.75).edges) -
+                   n)  # Due to self loops
+        edge_prob_q2 = 2 * N_edges / (n * (n - 1))
+        return random_gnp_dag(n, p_lags, pole_rad=0.75,
+                              edge_prob=edge_prob_q2)
+
+    def random_graph_dag_q4(n):
+        N_edges = (len(random_scg(n, p_lags,
+                                  pole_rad=0.75).edges) -
+                   n)  # Due to self loops
[…]

### 2019-06-28 software/simulation_increasing_p.py

@@ -8,8 +8,7 @@ import pandas as pd
 import numpy as np
 import statsmodels.api as sm
 
-from sklearn.kernel_ridge import KernelRidge
-from sklearn.model_selection import GridSearchCV
+from plotting_helpers import (COLOR1, COLOR2, plotting_model)
 
 from matplotlib import rc as mpl_rc
 font = {"family" : "normal",
@@ -68,16 +67,16 @@ class TrackErrors:
         return MCC, self.b_errs, self.errs, self.true_errs
 
 
-def pwgc_increasing_n(T):
-    simulation_name="dag_increasing_q_small_T"
+def pwgc_increasing_q(T):
+    simulation_name="dag_increasing_q_small_T_alasso"
     
     np.random.seed(0)
-    alpha, p_lags, p_max, T = 0.05, 5, 15, 500
+    alpha, p_lags, p_max, T = 0.05, 3, 5, 500
     T_max = 2 * T
 
     n = 50
-    edge_probs = np.linspace(2. / n**2, 1. / np.log(n), 100)
-    N_reps = 3
+    edge_probs = np.linspace(2. / n**2, 1. / np.log(n), 200)
+    N_reps = 2
 
     def random_graph(q):
         return random_gnp_dag(n, p_lags, pole_rad=0.75,
@@ -98,13 +97,14 @@ def pwgc_increasing_n(T):
 
     for q_iter, q in enumerate(edge_probs):
         for rep in range(N_reps):
-            print("q[{} / {}]\r".format(q_iter + 1, len(edge_probs)))
+            print("q[{} / {}]".format(q_iter + 1, len(edge_probs)),
+                  end="\r")
 
             G, X, sv2_true = make_test_data(q)
             G_hat_pwgc = estimate_graph(X[:T], G, max_lags=p_max,
                                         method="lstsqr", alpha=alpha)
             G_hat_lasso = estimate_dense_graph(X, max_lag=p_lags,
-                                               max_T=T, method="lasso")
+                                               max_T=T, method="alasso")
             errs_pwgc.update(G, G_hat_pwgc, sv2_true, q_iter, rep)
             errs_lasso.update(G, G_hat_lasso, sv2_true, q_iter, rep)
 
@@ -126,43 +126,44 @@ def _plot_results(q_iters, mcc_pwgc, errs_pwgc,
                   mcc_lasso, errs_lasso,
[…]

### 2019-06-30 paper/paper.tex
Some descriptions for figures and minor updates.

@@
 \begin{theorem}[Pairwise Recovery]
-  \label{thm:tree_recovery}
+  \label{thm:scg_recovery}
   If the Granger-causality graph $\gcg$ for persistent process $x(t)$
@@
 % \subsection{Pairwise Minimum Error Spanning Trees}
-% Inspired by theorem \ref{thm:tree_recovery}, we propose the following
+% Inspired by theorem \ref{scg_recovery}, we propose the following
 % structure learning heuristic.
@@
 
-where we are choosing $\lambda$, the regularization parameter, via the BIC.
+where we are choosing $\lambda$, the regularization parameters, via the BIC.
 
@@
 
+\begin{remark}[Graph Topologies]
+  We depict in Figure \ref{fig:random_graph_topologies} the topologies
+  of random graphs used in our empirical evaluation.  For values of
+  $q$ close to $\frac{2}{n}$, the resulting random graphs tend to have
+  a topology which is, at least qualitatively, close to the SCG.  As
+  the value of $q$ increases, the random graphs deviate farther from
+  the SCG topology, and we therefore expect the LASSO to outperform
+  PWGC for larger values of $q$.  This can be observed in Figure
+  \ref{fig:simulation_results_dense}, at least where performance is
+  measured by the support recovery (i.e. via the MCC).
+\end{remark}
+
+\begin{remark}[MCC as a Support Recovery Measurement]
+  We apply Matthew's Correlation Coefficient (MCC)
+  \cite{matthews1975comparison} as a statistic for measuring support
+  recovery performance.  This statistic synthesizes the confusion
+  matrix into a single score appropriate for unbalanced labels and is
+  calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
+  performance, $0$ being the performance of random guessing, and $-1$
+  being perfectly opposed.
+\end{remark}
+
+\begin{remark}[Error Measurement]
+  We we estimate the 1-step ahead prediction error by forming the variance matrix estimate
+
+  \begin{equation*}
+    \widehat{\Sigma}_v \defeq \frac{1}{T_{\text{out}}} \sum_{t = 1}^{T_{\text{out}}} (x(t) - \widehat{x}(t))(x(t) - \widehat{x}(t))^\T
+  \end{equation*}
+
+  on a long stream of out-of-sample data.  We then report the quantity
+
+  \begin{equation*}
+    \frac{\ln \tr \widehat{\Sigma}_v}{\ln \tr \Sigma_v}
+  \end{equation*}
+
+  where $\widehat{\Sigma}_v = \Sigma_v$ is the best possible performance.
+\end{remark}
+
 \begin{figure}
   \centering
-  \caption{Simulation Results}
+  \caption{PWGC Compared Against AdaLASSO \cite{adaptive_lasso_zou2006}}
   \label{fig:simulation_results}
-
-  {\scriptsize Comparison of PWGC and LASSO for $\VAR(p)$ model
-    estimation.  The MCC is measured on the left axis in blue with
-    circle markers, and the relative log mean-squared prediction error
-    $\frac{\ln\ \tr \widehat{\Sigma}_v}{\ln\ \tr \Sigma_v}$ on the
-    right axis in red with triangle markers.  We use $q$ to denote the
-    edge probability in a random Erdos-Reny graph (modified to form a
-    DAG).  Solid lines indicate ad-hoc regression estimates for
-    visualization purposes only.
-
-    Matthew's Correlation Coefficient (MCC)
-    \cite{matthews1975comparison} synthesizes the confusion matrix
-    into a single score appropriate for unbalanced labels and is
-    calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
-    performance, $0$ being the performance of random guessing, and
-    $-1$ being perfectly opposed.
-  }
 
@@
 
+  {\scriptsize Comparison of PWGC and LASSO for $\VAR(p)$ model
+    estimation.  We make comparisons against both the MCC and the
+    relative log mean-squared prediction error
+    $\frac{\ln\tr \widehat{\Sigma}_v}{\ln\tr \Sigma_v}$.  Results
+    in Figure \ref{fig:simulation_results_comparison1} are for systems
+    guaranteed to satisfy the assumptions required for Theorem
+    \ref{thm:scg_recovery}.  Figure
+    \ref{fig:simulation_results_comparison2} provides results for
+    systems which do not guarantee the assumptions of Theorem
+    \ref{thm:scg_recovery}, though the graph has a similar level of
+    sparsity.}
+\end{figure}
+
+\begin{figure}
+  \centering
   \begin{subfigure}[b]{0.45\textwidth}
@@
     \label{fig:simulation_results_scaling}
-    \includegraphics[width=\linewidth]{increasing_n_simulation.pdf}
+    \includegraphics[width=\linewidth]{new_increasing_n_simulation.pdf}
   \end{subfigure}
   \begin{subfigure}[b]{0.45\textwidth}
+    \caption{MCC Comparison for $T \le 100$}
+    \label{fig:small_T_comparison}
+    \includegraphics[width=\linewidth]{new_mcc_comparison000.pdf}
+  \end{subfigure}
+
+  {\scriptsize Figure \ref{fig:simulation_results_scaling} depicts
+    support recovery performance as the number of nodes $n$ increases,
+    and the edge proportion as well as the number of samples $T$ is
+    held fixed.  Remarkably, performance appears to degrade slowly in
+    $n$ -- it is the graph topology (SCG or non-SCG) as well as the
+    level of sparsity (measured by $q$) which is the main factor for
+    support recovery performance.
+
+    Figure \ref{fig:small_T_comparison} provides a support recovery
+    comparison for very small values of $T$ typical of some
+    applications.}
+\end{figure}
+
+\begin{figure}
   \caption{Fixed $T, n$, increasing edges $q$ (DAG)}
@@
   \includegraphics[width=\linewidth]{dag_increasing_q_small_T_alasso_simulation.pdf}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{MCC Comparison for $T \le 100$}
-    \label{fig:mcc_comparison}
-    \includegraphics[width=\linewidth]{new_mcc_comparison000.pdf}
-  \end{subfigure}
+
+  {\scriptsize Figure \ref{fig:simulation_results_dense} provides a
+    comparison between PWGC and AdaLASSO as the density of graph edges
+    (as measured by $q$) increases.  For reference,
+    $\frac{2}{n} = 0.04$ has approximately the same level of sparsity
+    as the SCGs we simulated.  As $q$ increases, the AdaLASSO
+    outperforms PWGC as measured by the MCC.  However, PWGC maintains
+    superior performance for 1-step-ahead prediction.  We speculate
+    that this is a result of fitting the sparsity pattern recovered by
+    PWGC via OLS which directly seeks to optimize this metric, whereas
+    the LASSO is encumbered by the sparsity inducing penalty.}
 \end{figure}

### 2019-06-30 software/lasso_comparison.py

@@ -128,7 +128,7 @@ def simple_lasso():
 
 def lasso_mcc_comparison():
     np.random.seed(0)
-    n_nodes, p_lags, p_max = 50, 3, 5
+    n_nodes, p_lags, p_max = 50, 5, 15
     alpha = 0.05
 
     N_edges = (len(random_scg(n_nodes, p_lags, pole_rad=0.75).edges) -
@@ -205,16 +205,15 @@ def lasso_mcc_comparison():
                      title=("MCC Comparison on $n = {}$ nodes"
                             "".format(n_nodes)),
                      lasso_title="Adaptive LASSO", show_results=True,
-                     save_file=["../figures/new_mcc_comparison000.pdf",
-                                "../figures/jpgs_pngs/new_mcc_comparison000.png"]
+                     save_file=["../figures/new_mcc_comparison001.pdf",
+                                "../figures/jpgs_pngs/new_mcc_comparison001.png"]
                     )
     return
 
 
 def lasso_comparison(simulation_name, graph_type):
     np.random.seed(0)
-    # n_nodes, p_lags, p_max = 50, 5, 15
-    n_nodes, p_lags, p_max = 50, 3, 5
+    n_nodes, p_lags, p_max = 50, 5, 15
     alpha, N_iters = 0.05, 2
 
     # T_iters = list(map(int, np.logspace(2, 3, 20)))
@@ -451,6 +450,6 @@ def plot_mcc_results(t, mcc_pwgc_scg, mcc_lasso_scg,
 
 
 if __name__ == "__main__":
-    # lasso_comparison("lasso_comparison_scg", "SCG")
-    # lasso_comparison("lasso_comparison_dag", "DAG")
+    lasso_comparison("lasso_comparison_scg_pmax15", "SCG")
+    lasso_comparison("lasso_comparison_dag_pmax15", "DAG")
     lasso_mcc_comparison()

### 2019-06-30 software/simulation_increasing_p.py

@@ -68,10 +68,10 @@ class TrackErrors:
 
 
 def pwgc_increasing_q(T):
-    simulation_name="dag_increasing_q_small_T_alasso"
-    
+    simulation_name = "dag_increasing_q_small_T_alasso_pmax15"
+
     np.random.seed(0)
-    alpha, p_lags, p_max, T = 0.05, 3, 5, 500
+    alpha, p_lags, p_max, T = 0.05, 5, 15, 500
     T_max = 2 * T
 
     n = 50

### 2019-06-30 paper/paper.tex
Slight figure rearranging.

@@
   \centering
-  \caption{PWGC Compared Against AdaLASSO \cite{adaptive_lasso_zou2006}}
-  \label{fig:simulation_results}
-
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{PWGC vs adaLASSO (SCG)}
+  \caption{PWGC Compared Against AdaLASSO \cite{adaptive_lasso_zou2006} (SCG)}
   \label{fig:simulation_results_comparison1}
-    \includegraphics[width=\linewidth]{new_lasso_comparison_scg_simulation.pdf}
-    % \missingfigure{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{PWGC vs adaLASSO (DAG, $q = \frac{2}{n}$)}
-    \label{fig:simulation_results_comparison2}
-    \includegraphics[width=\linewidth]{new_lasso_comparison_dag_simulation.pdf}
-    % \missingfigure{}
-  \end{subfigure}
+  \includegraphics[width=0.95\textwidth]{new_lasso_comparison_scg_simulation.pdf}
 
@@
     guaranteed to satisfy the assumptions required for Theorem
-    \ref{thm:scg_recovery}.  Figure
-    \ref{fig:simulation_results_comparison2} provides results for
-    systems which do not guarantee the assumptions of Theorem
-    \ref{thm:scg_recovery}, though the graph has a similar level of
-    sparsity.}
+    \ref{thm:scg_recovery}.}
+\end{figure}
+
+\begin{figure}
+  \caption{PWGC vs adaLASSO (DAG, $q = \frac{2}{n}$)}
+  \label{fig:simulation_results_comparison2}
+  \includegraphics[width=0.95\textwidth]{new_lasso_comparison_dag_simulation.pdf}
+  {\scriptsize Figure \ref{fig:simulation_results_comparison2}
+    provides results for systems which do not guarantee the
+    assumptions of Theorem \ref{thm:scg_recovery}, though the graph
+    has a similar level of sparsity.}
 \end{figure}

### 2019-07-05 paper/paper.tex
Some updates on paper prose.

@@
 % TODO: read this https://lemire.me/blog/rules-to-write-a-good-research-paper/
+
+% TODO: Are we, in this setting, immune to confounders?  We are removing
+%       all bi-directional edges, so it seems likely that we are?  Can we state
+%       any theorem about this?
 
@@
 established under less restrictive conditions than for the vanilla
-LASSO \cite{wainwright2009sharp}.%   Success of the LASSO has lead to
+LASSO.  Our experimental comparisons in Section
+\ref{sec:empirical_evaluation} are against the
+adaLASSO. % Success of the LASSO has lead to
 % the famous ``bet on sparsity'' principle ``Use a procedure that does
@@
 causality graph to the pairwise ``causality flow'' in the system, as
-well as a characterization of the sparsity pattern of matrices arising
-in the Wold decomposition, generalizing in some sense the notion of
-``feedback-free'' processes studied by \cite{caines1975feedback} in
-close connection with Granger-causality.  We establish sufficient
-conditions (sections \ref{sec:strongly_causal_graphs},
-\ref{sec:persistent_systems}) under which a fully conditional
-Granger-causality graph can be recovered from pairwise tests alone
-(sec \ref{sec:pairwise_algorithm}).  Secondly, we propose in section
-\ref{sec:structure_learning} a graph search heuristic which implements
-our theoretical results to finite data samples, specifying and
-summarizing appropriate methods for hypothesis testing (section
-\ref{sec:pairwise_hypothesis_testing}), model order selection (section
-\ref{sec:model_order_selection}), computationally efficient estimation
-(section \ref{sec:efficient_model_estimation}), and error rate
-controls (section \ref{sec:error_rate_control}).  Our heuristics are
-compared against the LASSO algorithm in Section
+well as an interpretation in terms of the graph topology of the
+sparsity pattern of matrices arising in the Wold decomposition,
+generalizing in some sense the notion of ``feedback-free'' processes
+studied by \cite{caines1975feedback} in close connection with
+Granger-causality.  We establish sufficient conditions (sections
+\ref{sec:strongly_causal_graphs}, \ref{sec:persistent_systems}) under
+which a fully conditional Granger-causality graph can be recovered
+from pairwise tests alone (sec \ref{sec:pairwise_algorithm}).
+Secondly, we propose in section \ref{sec:structure_learning} a graph
+search heuristic which implements our theoretical results to finite
+data samples, specifying and summarizing appropriate methods for
+hypothesis testing (section \ref{sec:pairwise_hypothesis_testing}),
+model order selection (section \ref{sec:model_order_selection}),
+computationally efficient estimation (section
+\ref{sec:efficient_model_estimation}), and error rate controls
+(section \ref{sec:error_rate_control}).  Our heuristics are compared
+against the adaLASSO algorithm in Section
 \ref{sec:empirical_evaluation}.  We stress the scalability of our
@@
 associated with stationary time series data.  In section
-\ref{sec:application} we develop an example application.  And provide
-concluding remarks on further open problems in our concluding section
+\ref{sec:application} we develop an example application.  Concluding
+remarks on further open problems and extentions are provided in Section
 \ref{sec:conclusion}.  \todo{do an application}
@@
 $\inner{x}{y} = \E[xy]$.  We will work with a discrete time and
-wide-sense stationary (WSS) vector valued process $x(t)$ (with
-$t \in \Z$) with elements taking values in $L_2$.  We suppose that
-$x(t)$ has zero mean, $\E x(t) = 0$, and has absolutely summable
-matrix valued covariance sequence
+wide-sense stationary (WSS) $n$-dimensional vector valued process
+$x(t)$ (with $t \in \Z$) where the $n$ elements take values in $L_2$.  We
+suppose that $x(t)$ has zero mean, $\E x(t) = 0$, and has absolutely
+summable matrix valued covariance sequence
 $R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ with an absolutely
@@
     % \H^x &= \cl \{\sum_{\tau = -T}^T a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, T \in \N\} \subseteq L_2(\Omega),\\
-    \H_t^x &= \cl \{\sum_{\tau = 0}^T a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, T \in \N\} \subseteq L_2(\Omega)\\
+    \H_t^x &= \cl \{\sum_{\tau = 0}^p a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, p \in \N\} \subseteq L_2(\Omega)\\
     H_t^x &= \{a x(t)\ |\ a \in \R\} \subseteq L_2(\Omega),
@@
 $\H_t^{(i, j)}$, $\H_t^{i}$ or by all but particular components as
-e.g. $\H_t^{-j}$.
+$\H_t^{-j}$.
 
@@
   $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi,
-  \pi]$}, $v(t)$ is an uncorrelated sequence and $A(0) = I$.  We
-will assume that $c(t) = 0$, which in practice is to say that it has
-been accurately estimated and removed.  We additionally require that
-this representation can be inverted to yield the $\VAR(\infty)$ form
+  \pi]$}, $v(t)$ is an uncorrelated sequence and $A(0) = I$.  We will
+assume that $c(t) = 0$, which in practice is to say that the process
+has been detrended.  We additionally require that this representation
+can be inverted to yield the $\VAR(\infty)$ form
 
@@
 \begin{definition}[Basic Setup]
+  \label{def:basic_setup}
   The process $x(t)$ is an $n$ dimensional wide sense stationary
@@
   \label{def:granger_causality}
-  For the WSS series $x(t)$ we say that component $x_j$
+  For the WSS series $x(t)$ satisfying the assumptions of Definition
+  \ref{def:basic_setup} we will say that component $x_j$
   \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x$)
-  and write $x_j \gc x_i$ if given Hilbert spaces
-  $\H_{t - 1}$, $\H^{-j}_{t - 1}$
+  and write $x_j \gc x_i$ if given Hilbert spaces $\H_{t - 1}$,
+  $\H^{-j}_{t - 1}$
 
@@
     \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$ i.e. $\B_{ij}(z) = 0$}
-    \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} = 0$}
+    \item{$F_{ij} \defeq \big(\frac{\xi_i}{\xi_{ij}} - 1\big) = 0$}
     \item{$\H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j} \iff H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
@@
 
-  A (directed) \textit{path} (of length $r$) between nodes $i, j$,
-  denoted $\gcgpath{i}{j}$, is a sequence
+  A (directed) \textit{path} (of length $r$) from node $i$ to node
+  $j$, denoted $\gcgpath{i}{j}$, is a sequence
   $a_0, a_1, \ldots, a_{r - 1}, a_r$ with $a_0 = i$ and $a_r = j$ such
-  that $\forall\ 0 \le k \le r\ (a_k, a_{k + 1}) \in \gcge$.
+  that $\forall\ 0 \le k \le r\ (a_k, a_{k + 1}) \in \gcge$, and where
+  $(a_k, a_{k - 1})$ are \textit{distinct} for $0 \le k < r$.
 
   A \textit{cycle} is a path of length $2$ or more between a node and
-  itself.  An edge between a node and itself $(i, i)$ (which is not a
-  cycle) is referred to as a \textit{loop}.
+  itself.  An edge between a node and itself $(i, i) \in \gcge$ (which
+  is not a cycle) is referred to as a \textit{loop}.
 
@@
   A node $j$ is a \textit{parent} of node $i$ if $(j, i) \in \gcge$.
-  The set of all $i$'s parents will be denoted $\pa{i}$.
+  The set of all $i$'s parents will be denoted $\pa{i}$, and we
+  explicitly exclude loops as a special case, that is,
+  $i \not\in \pa{i}$ even if $(i, i) \in \gcge$.
 
@@
 
-  Note that since loops are not considered paths, it is not possible
-  for a node to be it's own parent.  However, unless $\gcg$ is a DAG,
-  a node can be it's own ancestor.  We will ocassionally need to
-  explicitly exclude $i$ from $\anc{i}$, in which case we will write
-  $\anc{i}\setminus \{i\}$.
+  Recall that we do not allow a node to be it's own parent, however
+  unless $\gcg$ is a DAG, a node can be it's own ancestor.  We will
+  ocassionally need to explicitly exclude $i$ from $\anc{i}$, in which
+  case we will write $\anc{i}\setminus \{i\}$.
 \end{definition}
@@
 An elementary theorem connecting the adjacency matrix with paths will
-allow us to deduce the sparsity pattern of $\A(z)$.  We omit proof,
-though it follows easily by induction:
+allow us to deduce the sparsity pattern of $\A(z)$.  Proof follows
+easily by induction:
 
@@
 From the $\VAR$ representation of $x(t)$ there is clearly a tight
-relationship between each node and it's parent nodes, the following
-proposition is analogous to the definition of feedback free processes
-of \cite{caines1975feedback}:
+relationship between each node and it's parent nodes, the relationship
+is quantified through the sparsity pattern of $B(z)$.  Similarly, the
+following proposition is analogous to the definition of feedback free
+processes of \cite{caines1975feedback} and provides an interpretation
+of the sparsity pattern of $A(z)$ (from the MA representation of
+$x(t)$) in terms of the causality graph $\gcg$.
 
@@
 \tau)$.  The proposition states that if $j \not \in \anc{i}$ then
-$\A_{ij}(z) = 0$, however the implication
-$j \in \anc{i} \implies \A_{ij}(z)$ need not hold.
+$\A_{ij}(z) = 0$.
+
+\todo{Do we have $A_{ij}(z) \ne 0 \implies j \in \anc{i}$?}
+
+% TODO: ???
+% , however the implication $j \in \anc{i} \implies \A_{ij}(z) \ne 0$ need not hold.
 
@@
 with respect to $x_{-k}$, it may not hold with respect to $x$.  For
-example, $x_k$ may be a common ancestor which when observed completely
-explains the connection from $x_j$ to $x_i$.  In this paper we study
+example, $x_k$ may be a common ancestor which when observed, completely
+explains the connection from $x_j$ to $x_i$.  In this section we study
 \textit{pairwise} Granger-causality, and seek to understand when
@@
 computation of such pairwise relations is somewhat easier.
+\todo{Reconsider this paragraph}
 
@@
   \label{prop:ancestor_properties}
-  If in a Granger-causality graph $\gcg$ we have $j \pwgc i$ then
+  If in a Granger-causality graph $\gcg$ where $j \pwgc i$ then
   $j \in \anc{i}$ or $\exists k \in \anc{i} \cap\anc{j}$ which is a
@@
 \end{proposition}
+
+\begin{remark}
+  The interpretation of this proposition is that for $j \pwgc i$ then
+  there must either be ``causal flow'' from $j$ to $i$
+  ($j \in \anc{i}$) or there must be a confounder $k$ through which
+  common information is received.
+\end{remark}
 
@@
 \begin{corollary}
-  If the graph $\gcg$ is acyclic and if $j \pwgc i$ and $i \pwgc j$ then $\exists k \in \anc{i} \cap \anc{j}$ confounding $(i, j)$.
+  If the graph $\gcg$ is a DAG then $j \pwgc i, i \pwgc j \implies \exists k \in \anc{i} \cap \anc{j}$ confounding $(i, j)$.
 \end{corollary}
@@
   \textit{strongly causal} if there is at most 1 directed path between
-  any two nodes.  Strongly Causal Graphs will be referred to as ``SCGs''.
+  any two nodes.  Strongly Causal Graphs will be referred to as SCGs.
 \end{definition}
@@
 In light of proposition \ref{prop:sc_graph_common_anc}, the following
-provides a partial converse to proposition \ref{prop:ancestor_properties}.
+provides a partial converse to proposition
+\ref{prop:ancestor_properties}, and supports the intuition of ``causal
+flow'' through paths in $\gcg$.
 
@@
   will never result in $i \pwgc j$ or $j \pwgc i$.  This is evidently
-  an advantage of the \textit{temporal} nature of Granger causality --
-  ``information'' cannot flow backwards along the edges of $\gcg$.
+  an advantage of the \textit{temporal} nature of Granger -- there is
+  no backwards causal flow along the edges of $\gcg$.
 \end{remark}
@@
   where $\tau_\infty(\B) \defeq \infty$ if the filter has an infinite
-  impulse response.
+  length.
 \end{definition}
@@
   persistent.  In particular, $\mathsf{\VAR}(p)$ models are likely to
-  be persistent since these naturally result in equivalent an
+  be persistent since these naturally result in an equivalent
   $\mathsf{MA}(\infty)$ representation, except for the pathological
@@
   Moreover, persistence is not the weakest condition necessary for the
-  results of this section, but we opt to avoid unnecessarily
-  complicated definitions.
+  results of this section, the condition
+  $\tau_0(\A_{jk}) < \tau_\infty(\A_{ik})$ for each $i, j, k$ such
+  that $k \in \anc{i} \cap \anc{j}$ is enough.
+  \todo{Write something on the intuition here.}
 \end{remark}
@@
     $\VAR(1)$ model, so we lose little generality in considering this
-    case.}  having $\B(z) = Bz^{-1}$ where $B$, we will examine
+    case.}  having $\B(z) = Bz^{-1}$, we will examine
   conditions that guarantee persistence.  Pick any
@@
   \begin{equation*}
-    \Dg(\lambda)^r \lambda = Vz
+    \Dg(\bm{\lambda})^r \bm{\lambda} = Vz
   \end{equation*}
@@
   had a solution in $z$ for every $r \ge N$, where $V$ is an
-  $n \times n - 1$ full-rank matrix spanning the nullspace of $\beta$.
-  That is, iterates of $\Dg(\lambda)$ applied to $\lambda$ would need
-  to remain inside $\mathcal{N}(\beta)$.  This would imply that
+  $n \times n - 1$ full-rank matrix spanning the nullspace of $\beta$,
+  and $\bm{\lambda} = (\lambda_1, \ldots, \lambda_n)$. That is,
+  iterates of $\Dg(\bm{\lambda})$ applied to $\bm{\lambda}$ would need to remain
+  inside $\beta$'s nullspace.  This would imply that
 
   \begin{equation*}
-    VV^\dagger \lambda^{r + 1} = \lambda^{r + 1},
+    VV^\dagger \bm{\lambda}^{r + 1} = \bm{\lambda}^{r + 1},
   \end{equation*}
 
-  i.e. that $\lambda^{r + 1}$ is an eigenvector of $VV^\dagger$ for an
-  infinite number of integers $r$.  However, since there can only be a
-  finite number of (unit length) eigenvectors, this cannot be the case
-  unless every eigenvalue $(\lambda_1, \ldots, \lambda_n)$ were equal.
+  i.e. that $\bm{\lambda}^{r + 1}$ is an eigenvector of $VV^\dagger$
+  for an infinite number of integers $r$ (the exponentiation is to be
+  understood as a pointwise operation).  However, since there can only
+  be a finite number of (unit length) eigenvectors, this cannot be the
+  case unless every eigenvalue $(\lambda_1, \ldots, \lambda_n)$ were
+  equal.
 
@@
 We set $p_{\text{max}} = 15$.  Results and representative graphs are
-collected in Figure \ref{fig:simulation_results}.
+collected in Figures \ref{fig:random_graph_topologies},
+\ref{fig:simulation_results_comparison1},
+\ref{fig:simulation_results_comparison2},
+\ref{fig:simulation_results_scaling_and_small_T},
+\ref{fig:simulation_results_dense}.
 
@@
   \label{fig:simulation_results_comparison1}
-  \includegraphics[width=0.95\textwidth]{new_lasso_comparison_scg_simulation.pdf}
+  \includegraphics[width=0.95\textwidth]{new_lasso_comparison_scg_pmax15_simulation.pdf}
 
@@
   \label{fig:simulation_results_comparison2}
-  \includegraphics[width=0.95\textwidth]{new_lasso_comparison_dag_simulation.pdf}
+  \includegraphics[width=0.95\textwidth]{new_lasso_comparison_dag_pmax15_simulation.pdf}
   {\scriptsize Figure \ref{fig:simulation_results_comparison2}
@@
   \centering
+  \caption{PWGC Scaling and Small Sample Performance}
+  \label{fig:simulation_results_scaling_and_small_T}
   \begin{subfigure}[b]{0.45\textwidth}
@@
     \label{fig:small_T_comparison}
-    \includegraphics[width=\linewidth]{new_mcc_comparison000.pdf}
+    \includegraphics[width=\linewidth]{new_mcc_comparison001.pdf}
   \end{subfigure}
 
-  {\scriptsize Figure \ref{fig:simulation_results_scaling} depicts
+  {\scriptsize Figure \ref{fig:simulation_results_scaling} measures
     support recovery performance as the number of nodes $n$ increases,
     and the edge proportion as well as the number of samples $T$ is
-    held fixed.  Remarkably, performance appears to degrade slowly in
-    $n$ -- it is the graph topology (SCG or non-SCG) as well as the
-    level of sparsity (measured by $q$) which is the main factor for
-    support recovery performance.
+    held fixed.  Remarkably, the degradation as $n$ increases is
+    limited, it is primarily the graph topology (SCG or non-SCG) as
+    well as the level of sparsity (measured by $q$) which are the
+    determining factors for support recovery performance.
 
     Figure \ref{fig:small_T_comparison} provides a support recovery
-    comparison for very small values of $T$ typical of some
+    comparison for very small values of, $T$ typical for many
     applications.}

### 2019-07-05 software/lasso_comparison.py

@@ -126,7 +126,7 @@ def simple_lasso():
     return
 
 
-def lasso_mcc_comparison():
+def lasso_mcc_comparison(showfig=False):
     np.random.seed(0)
     n_nodes, p_lags, p_max = 50, 5, 15
     alpha = 0.05
@@ -204,14 +204,14 @@ def lasso_mcc_comparison():
                      mcc_pwgc_dag_q4, mcc_lasso_dag_q4,
                      title=("MCC Comparison on $n = {}$ nodes"
                             "".format(n_nodes)),
-                     lasso_title="Adaptive LASSO", show_results=True,
+                     lasso_title="Adaptive LASSO", show_results=showfig,
                      save_file=["../figures/new_mcc_comparison001.pdf",
                                 "../figures/jpgs_pngs/new_mcc_comparison001.png"]
                     )
     return
 
 
-def lasso_comparison(simulation_name, graph_type):
+def lasso_comparison(simulation_name, graph_type, showfig=False):
     np.random.seed(0)
     n_nodes, p_lags, p_max = 50, 5, 15
     alpha, N_iters = 0.05, 2
@@ -284,7 +284,7 @@ def lasso_comparison(simulation_name, graph_type):
 
     plot_results(
         D_MCC_pwgc, D_errs_pwgc, D_MCC_lasso, D_errs_lasso,
-        D_true_errs,
+        D_true_errs, show_results=showfig,
         title=("Test Errors against T (Random {} graph on "
                "$n = {}$ nodes)".format(graph_type, n_nodes)),
         lasso_name="Adaptive LASSO",

### 2019-07-05 software/pwgc/gc_methods.py

@@ -797,6 +797,8 @@ def get_residual_graph(G_hat):
 
 def estimate_dense_graph(X, max_lag=10,
                          max_T=None, method="lasso"):
+    # TODO: Replace this with ts_lasso
+
     _, n = X.shape
     G = make_complete_digraph(n)
     G = attach_X(G, X)

### 2019-07-05 software/test/test_gc_methods.py
Test files.

@@ -1,0 +1,181 @@
+import unittest
+import numpy as np
+import matplotlib.pyplot as plt
+
+from pwgc.gc_methods import (_fast_univariate_AR_error,
+                             _fast_bivariate_AR_error,
+                             compute_covariances,
+                             form_bivariate_covariance,
+                             fast_compute_xi, fast_compute_pairwise_gc)
+from pwgc.var_system import (random_tree_dag, get_X, drive_gcg)
+
+
+class TestARFitting(unittest.TestCase):
+    def SetUp(self):
+        np.random.seed(0)
+        return
+
+    def _create_data(self):
+        n_nodes, p_lags, max_lags = 30, 5, 10
+
+        T_max = 2000
+
+        random_graph = lambda: random_tree_dag(
+            n_nodes, p_lags, pole_rad=0.75)
+        sv2_true = 0.5 + np.random.exponential(0.5, size=n_nodes)
+        G = random_graph()
+        drive_gcg(G, T_max, sv2_true, filter_attr="b(z)")
+        X = get_X(G)
+        X = X - np.mean(X, axis=0)[None, :]
+        return X
+
+    def test_univariate_AR_error001(self):
+        # This only really checks that it runs without an error
+        X = self._create_data()
+        p_max = 10
+        T, n = X.shape
+
+        R = compute_covariances(X, p=p_max)
+        for i in range(n):
+            r = R[:, i, i]
+            err, p_opt = _fast_univariate_AR_error(r, T)
+        return
+
+    def test_univariate_AR_error002(self):
+        # This only really checks that it runs without an error
+        X = self._create_data()
+        p_max = 15
+        T, n = X.shape
+
+        Eps = np.zeros((p_max + 1, n))  # From levinson
+        SV2 = np.zeros((p_max + 1, n))  # Direct lstsqr
+
+        R = compute_covariances(X, p=p_max)
+        for i in range(n):
+            r = R[:, i, i]
+            _, _, eps = lev_durb(r)
+            Eps[:, i] = eps
+            bic = compute_bic(eps, T, s=1)
+            p_opt = np.argmax(bic)
+            plt.plot(bic, color="b", alpha=0.5)
+            plt.scatter(p_opt, bic[p_opt], color="r", marker="o")
+
+        for i in range(n):
+            bic = np.zeros(p_max + 1)
+            for p in range(p_max + 1):
+                if p == 0:
[…]

### 2019-07-05 software/intraday_sp500/analysis.py
Some bug fixes plus a bit of progress on the sp500 data.

@@ -47,27 +47,48 @@ def main():
     X = D_close.to_numpy()
 
     graph_estimates = []
-    G_e = pwgc_estimate_graph(X, max_lags=20, method="alasso")
+    max_lag = 2
+    G_e = pwgc_estimate_graph(X, max_lags=max_lag, method="alasso")
     graph_estimates.append(G_e)
     for i in range(3):
         G_r = get_residual_graph(graph_estimates[-1])
         X_r = get_X(G_r)
-        G_e = pwgc_estimate_graph(X_r, max_lags=20, method="alasso",
+        G_e = pwgc_estimate_graph(X_r, max_lags=max_lag, method="alasso",
                                   alpha=0.01)
         graph_estimates.append(G_e)
 
     G_sp500_hat = combine_graphs(graph_estimates)
-    # draw_graph(G_sp500_hat)
-    nx.readwrite.gexf.write_gexf(G_sp500_hat, path="./G_sp500_hat.gexf")
+
+    G_sp500_hat = attach_X(G_sp500_hat, X)
+
+    T_max = 750
+    G_sp500_hat = estimate_B(G_sp500_hat, max_lag=max_lag, method="alasso",
+                             max_T=T_max)
+
+    len_r = len(X) - T_max - max_lag
+    # for i in G_sp500_hat.nodes:
+        
+    X_r = get_X(G_sp500_hat, prop="r")
 
     G_sp500_hat = nx.relabel_nodes(G_sp500_hat,
                                    {i: sp500[i] for i in range(len(sp500))})
-    # G_sp500_hat.remove_edges_from([(i, j) for i, j in G_sp500_hat.edges
-    #                               if i == j])
 
-    G_sp500_hat = attach_X(G_sp500_hat, X)
-    G_sp500_hat = estimate_B(G_sp500_hat, max_lag=10, method="alasso")
-    X_r = get_X(G_r, prop="x")
+
+    # draw_graph(G_sp500_hat)
+    nx.readwrite.gexf.write_gexf(G_sp500_hat, path="./G_sp500_hat.gexf")
+
+    i = 10
+    plt.plot(X[6:, i], color="b", alpha=0.75, linewidth=2)
+    plt.plot(X_r[:, i], color="r", alpha=0.75)
+    plt.show()
+
+    plt.plot(X[6:, i] - X_r[:, i])
+    plt.show()
+
+    i = 12
[…]

### 2019-07-05 software/pwgc/gc_methods.py

@@ -279,6 +279,7 @@ def estimate_B(G, max_lag=10, copy_G=False,
 
         if len(a_i) == 0:
             G.nodes[i]["sv^2_hat"] = np.var(y_raw)
+            G.nodes[i]["r"] = G.nodes[i]["x"][max_T + max_lag:]
             continue
         else:
             X_raw = np.hstack([G.nodes[j]["x"][:, None] for j in a_i])

### 2019-07-05 software/pwgc/var_system.py

@@ -169,17 +169,17 @@ def random_scg(n_nodes, p_lags, pole_rad=0.95):
     return G
 
 
-def sort_X_by_nodes(G, X):
+def sort_X_by_nodes(G, X, prop="x"):
     nodes = list(G.nodes)
     nodes_inv = [nodes.index(i) for i in range(len(nodes))]
     X_inv = X[:, nodes_inv]
-    assert np.all(X_inv[:, 0] == G.nodes[0]["x"])
+    assert np.all(X_inv[:, 0] == G.nodes[0][prop])
     return X_inv
 
 
 def get_X(G, prop="x"):
     X = get_node_property_vector(G, prop).T
-    X = sort_X_by_nodes(G, X)
+    X = sort_X_by_nodes(G, X, prop)
     return X
 
 

### 2019-07-05 software/intraday_sp500/analysis.py
Variance comparison.

@@ -62,12 +62,9 @@ def main():
     G_sp500_hat = attach_X(G_sp500_hat, X)
 
     T_max = 750
-    G_sp500_hat = estimate_B(G_sp500_hat, max_lag=max_lag, method="alasso",
+    G_sp500_hat = estimate_B(G_sp500_hat, max_lag=max_lag, method="lstsqr",
                              max_T=T_max)
 
-    len_r = len(X) - T_max - max_lag
-    # for i in G_sp500_hat.nodes:
-        
     X_r = get_X(G_sp500_hat, prop="r")
 
     G_sp500_hat = nx.relabel_nodes(G_sp500_hat,
@@ -89,6 +86,8 @@ def main():
     plt.plot(X[-247:, i], color="b", alpha=0.75, linewidth=2)
     plt.plot(X_r[:, i], color="r", alpha=0.75, linewidth=1)
     plt.show()
+
+    V = [np.var(X[-247:, i]) - np.var(X_r[:, i]) for i in range(500)]
 
     deg_seq = [d for n, d in G_sp500_hat.degree()]
 

### 2019-07-09 paper/paper.tex
Hacking away at classifying graphs...  A / C is extremely difficult, but there appear to be much clearer separations between individuals.

@@
 \label{sec:application}
+In this section we apply our methods to a real set of EEG data
+obtained from the ``EEG Database Data
+Set''\footnote{http://archive.ics.uci.edu/ml/datasets/EEG+Database}
+\cite{zhang1995event} on the UCI machine learning repository
+\cite{uci_mlr}.  This dataset contains 1 second long measurements of
+(64 channel) EEG signals from patients who are given visual stimuli
+and are broken into two groups: control and alcoholic.  The dataset
+then consists of
+$\mathcal{D} = \{\big(x^{(i)}(t)\big)_{t = 1}^T, y^{(i)} \}_{i = 1}^N$
+where $T = 256$, $x^{(i)}(t) \in \R^{64}$,
+$y^{(i)} \in \{\text{A}, \text{C}\}$, and \todo{fill in N} $N = ????$.
+
+We have constructed a simple pipeline for classifying each each trial
+as either ``A'' or ``C'' by first applying an iterative PWGC algorithm
+(see Algorithm \ref{alg:iterated_pwgc}) directly to the EEG data
+$x^{(i)}$ to obtain a Granger-causality graph $\widehat{G}^{(i)}$ then
+simply feeding the vectorized adjacency matrix of $\widehat{G}^{(i)}$
+through a Kernel PCA stage and then a Support Vector Machine with
+parameters fit by cross validation.
+
+While it is almost certainly possible to achieve much greater
+classification accuracy on this dataset by constructing a classifier
+to act directly on $x^{(i)}(t)$, our purpose is to demonstrate that
+relevant information is being captured by the graph estimates
+$\widehat{G}^{(i)}$.
+
+Figures \ref{fig:ar_eeg_example} and \ref{fig:eeg_oos_eeror}
+illustrate the appropriateness of modelling $x(t)$ with $\VAR$ models,
+as opposed simply to a collection of $n$ unidimensional
+autoregressiive models.  The conclusion being that the
+Granger-causality graph estimates are not purely spurious.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{0.45\textwidth}
+    \caption{$\mathsf{AR}(p)$ Model Example}
+    \label{fig:ar_eeg_example}
+    \includegraphics[width=\linewidth]{eeg_estimate_example.pdf}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \caption{OOS Error for a Single Subject}
+    \label{fig:eeg_oos_error}
+    \includegraphics[width=\linewidth]{eeg_oos_demonstration.pdf}
+  \end{subfigure}
+
+  {\scriptsize Figure \ref{fig:ar_eeg_example} provides an example of an
+  autoregressive model fir via the Levinson-Durbin algorithm (choosing
+  the system order via the BIC) on a single EEG channel.  Figure
+  \ref{fig:eeg_oos_error} demonstrates the improvement of a single
+  $\VAR$ model over independent $\mathsf{AR}$ models.
+
+  Out-of-Sample estimates are performed on data that were not used in
+  fitting the models.}
+
+\end{figure}
+
 \section{Conclusion}

### 2019-07-09 software/eeg_analysis.py

@@ -32,6 +32,9 @@ channels = [
     "P2", "P3", "P4", "P5", "P6", "P7", "P8", "PO1", "PO2", "PO7", "PO8",
     "POZ", "PZ", "T7", "T8", "TP7", "TP8", "X", "Y", "nd"]
 
+# X, Y are "EOG" (Electro-oculogram) for eyes
+# nd is a reference electrode
+
 thread_lock = Lock()
 
 
@@ -45,9 +48,9 @@ def compute_all_networks():
 
 def compute_and_save_networks(subject):
     print(subject)
-    if subject[:4] == "co2a":
+    if subject[3] == "a":
         pass
-    elif subject[:4] == "co2c":
+    elif subject[3] == "c":
         pass
     else:
         return
@@ -142,12 +145,11 @@ def make_adj_mat(all_G_hat):
     return G_adj
 
 
-def oos_error_demonstration():
+def oos_error_demonstration(subject="co2a0000364"):
     all_V0 = []
     all_V = []
     all_V_bl = []
     all_G_hat = []
-    subject = "co2a0000364"
     folder = "eeg/" + subject + "/"
     for file_name in os.listdir(folder):
         V, V_bl, V0, G_hat = estimate_eeg_graph(folder + file_name,
@@ -305,5 +307,5 @@ def read_trial(file_name):
 if __name__== "__main__":
     # oos_example()
     # oos_error_demonstration()
-    # compute_all_networks()
+    compute_all_networks()
     pass

### 2019-07-09 software/eeg_graph_analysis.py

@@ -4,6 +4,7 @@ import seaborn as sns
 import networkx as nx
 import os
 import grakel
+import pandas as pd
 
 from grakel import GraphKernel  # Graph kernels
 from datetime import datetime
@@ -13,13 +14,26 @@ from eeg_analysis import channels
 
 from sklearn.svm import SVC
 from sklearn.model_selection import (StratifiedShuffleSplit, StratifiedKFold,
-                                     GridSearchCV, learning_curve)
+                                     GridSearchCV, learning_curve,
+                                     RandomizedSearchCV)
 # from sklearn.learning_curves import learning_curve
 from sklearn.metrics import matthews_corrcoef
-from sklearn.decomposition import PCA
+from sklearn.decomposition import PCA, KernelPCA
 from sklearn.cross_decomposition import PLSRegression
 from sklearn.manifold import Isomap, TSNE
 from sklearn.discriminant_analysis import LinearDiscriminantAnalysis
+from sklearn.decomposition import NMF
+from sklearn.naive_bayes import GaussianNB
+from sklearn.pipeline import Pipeline
+from sklearn.preprocessing import PowerTransformer, QuantileTransformer
+from sklearn.neighbors import KNeighborsClassifier
+from sklearn.tree import DecisionTreeClassifier
+from sklearn.ensemble import AdaBoostClassifier
+from sklearn.cross_decomposition import PLSRegression
+from sklearn.metrics.pairwise import chi2_kernel, polynomial_kernel, rbf_kernel
+from sklearn.kernel_approximation import Nystroem
+from sklearn.preprocessing import LabelEncoder
+from sklearn.linear_model import LogisticRegressionCV
 
 from matplotlib import rc as mpl_rc
 font = {"family": "normal",
@@ -30,46 +44,248 @@ mpl_rc("font", **font)
 mpl_rc("text", usetex=False)
 
 
+def main_Xy2():
+    X, y, z = load_subject_Xyz()
+    le = LabelEncoder()
+    y = le.fit_transform(y)
+
+    # 2
+    for i in range(10):
+        xform = Pipeline(
+            steps=[("pca", KernelPCA(n_components=5,
+                                     kernel="rbf",
+                                     gamma=0.75,
[…]

### 2019-07-09 software/eeg_graph_analysis.py
Logistic Regression Results.

@@ -33,7 +33,9 @@ from sklearn.cross_decomposition import PLSRegression
 from sklearn.metrics.pairwise import chi2_kernel, polynomial_kernel, rbf_kernel
 from sklearn.kernel_approximation import Nystroem
 from sklearn.preprocessing import LabelEncoder
-from sklearn.linear_model import LogisticRegressionCV
+from sklearn.linear_model import LogisticRegression
+from sklearn.manifold import Isomap, TSNE, MDS, SpectralEmbedding
+from sklearn.metrics import f1_score, matthews_corrcoef
 
 from matplotlib import rc as mpl_rc
 font = {"family": "normal",
@@ -41,7 +43,7 @@ font = {"family": "normal",
         "size": 22}
 
 mpl_rc("font", **font)
-mpl_rc("text", usetex=False)
+mpl_rc("text", usetex=True)
 
 
 def main_Xy2():
@@ -50,37 +52,25 @@ def main_Xy2():
     y = le.fit_transform(y)
 
     # 2
-    for i in range(10):
-        xform = Pipeline(
-            steps=[("pca", KernelPCA(n_components=5,
-                                     kernel="rbf",
-                                     gamma=0.75,
-                                     eigen_solver="arpack",
-                                     random_state=i,
-                                     )),
-                   ("power_transform", PowerTransformer(method="yeo-johnson",
-                                                        standardize=True))])
-        X_pca = xform.fit_transform(X)
-        D_pca = pd.DataFrame(X_pca)
-        D_pca["label"] = ["Alcoholic" if yi == 1 else "Control" for yi in y]
-        pp = sns.pairplot(D_pca.iloc[::2, ], hue="label",
-                          vars=[c for c in D_pca.columns if c != "label"],
-                          plot_kws={"alpha": 0.25})
-        plt.title(i)
-        plt.show()
-
-    # ss = StratifiedShuffleSplit(n_splits=1, test_size=0.2)
-    # train_ix, valid_ix = next(ss.split(X, y))
-    # X_valid, y_valid = X[valid_ix], y[valid_ix]
-    # X_train, y_train = X[train_ix], y[train_ix]
-
-    # K = Nystroem(kernel="sigmoid", gamma=0.75, n_components=200).fit_transform(K)
-    # pt = PowerTransformer(method="yeo-johnson",
-    #                       standardize=True)
[…]

### 2019-07-09 paper/paper.tex
Some final figures for the eeg analysis.  Should I provide a comparison against LASSO?

@@
 \usepackage{todonotes}
+\usepackage{fullpage}
 \bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
@@
   statistical and computational) which can be obtained (in comparison
-  to LASSO algorithms) when structural assumptions are met.}
+  to LASSO algorithms) when structural assumptions are met.  Finally,
+  we provide an example application where we accurately discriminate
+  between subjects in an EEG study based purely on the
+  Granger-causality graphs inferred from data by our algorithms,
+  demonstrating that meaningful features are captured by the
+  Granger-causal graph topology.}
 
 \keywords{Granger-causality, causality graph, time series, vector autoregression,
-  LASSO, gene regulatory networks, network learning}
+  LASSO, gene regulatory networks, network learning, EEG}
 
@@
 from serious confounding problems.  Our finite sample results are
-based on simulations, where we show promising results for these graphs
+based on simulations, where we show promising results for these graph
 topologies where our algorithm performs substantially better in our
@@
 associated with stationary time series data.  In section
-\ref{sec:application} we develop an example application.  Concluding
-remarks on further open problems and extentions are provided in Section
-\ref{sec:conclusion}.  \todo{do an application}
-
-% \cite{yuan2006model} as well as local search hill
-% climbing (\hl{state here the results}).
+\ref{sec:application} we develop an example application where a
+classifier is constructed to accurately discriminate between subjects
+in an EEG study based only on the Granger-causality graph topology
+inferred by our algorithms.  Concluding remarks on further open
+problems and extentions are provided in Section \ref{sec:conclusion}.
 
@@
     \label{alg:finite_pwgc}
+
     \Input{Estimates of pairwise Granger-causality statistics $F_{ij}$
@@
 \begin{remark}[Error Measurement]
-  We we estimate the 1-step ahead prediction error by forming the variance matrix estimate
+  We estimate the 1-step ahead prediction error by forming the variance matrix estimate
 
@@
 \cite{uci_mlr}.  This dataset contains 1 second long measurements of
-(64 channel) EEG signals from patients who are given visual stimuli
-and are broken into two groups: control and alcoholic.  The dataset
-then consists of
+(64 channel) EEG signals from patients who are given visual stimuli.
+The subjects in the study are labeled as being either ``control'' or
+``alcoholic'', however, we will ignore this label and instead focus on
+distinguishing between subjects based on the Granger-causality graphs
+inferred from the subject's trials.  Our reasoning is to focus on the
+underlying question: ``Does our PWGC algorithm uncover meaningful
+Granger-causal connections from EEG data?''.  Focusing only on the
+subject label allows us to answer this question without simultaneously
+grappeling with the physiological question of whether alcoholic
+subjects as a group have discernable differences in their EEG
+readings\footnote{Some exploratory analysis actually suggests
+  that alcoholic Granger-causality graphs are not substantively
+  different}, which is a stronger requirement than simply that there
+are meaningful distinctions between the EEG readings of subjects
+generally.
+
+The dataset consists of
 $\mathcal{D} = \{\big(x^{(i)}(t)\big)_{t = 1}^T, y^{(i)} \}_{i = 1}^N$
 where $T = 256$, $x^{(i)}(t) \in \R^{64}$,
-$y^{(i)} \in \{\text{A}, \text{C}\}$, and \todo{fill in N} $N = ????$.
-
-We have constructed a simple pipeline for classifying each each trial
-as either ``A'' or ``C'' by first applying an iterative PWGC algorithm
-(see Algorithm \ref{alg:iterated_pwgc}) directly to the EEG data
-$x^{(i)}$ to obtain a Granger-causality graph $\widehat{G}^{(i)}$ then
-simply feeding the vectorized adjacency matrix of $\widehat{G}^{(i)}$
-through a Kernel PCA stage and then a Support Vector Machine with
+$y^{(i)} \in [N_{subjects}]$ (with $N_{subjects} = 119$), and
+$N = 10723$.  There are on average 90 trials for each subject,
+ranging between 30 and 119.
+
+We have constructed a simple pipeline for discriminating between
+subjects by first applying an iterative PWGC algorithm (see Algorithm
+\ref{alg:iterated_pwgc}) directly to the EEG data $x^{(i)}$ to obtain
+a Granger-causality graph $\widehat{G}^{(i)}$.  We then feed the
+vectorized adjacency matrix of $\widehat{G}^{(i)}$ through a
+polynomial Kernel multinomial logistic regression model with
 parameters fit by cross validation.
+
+The iterated application of PWGC is an example of a heuristic by which
+graph estimates that aren't constrained to being strongly causal can
+be obtained.  Theoretical analysis of more sophisticated heuristics
+than PWGC (Algorithm \ref{alg:finite_pwgc}) is deferred to future work.
+
+\begin{algorithm}
+  \SetKwInOut{Input}{input}
+  \SetKwInOut{Output}{output}
+  \SetKwInOut{Initialize}{initialize}
+  \DontPrintSemicolon
+
+  \BlankLine
+  \caption{Iterated PWGC Heuristic}
+  \label{alg:iterated_pwgc}
+
+  \Input{Data $x(t)$, maximum number of iterations $N$,
+    threshold $r \in (0, 1)$.}
+  \Output{Granger-causality graph estimate $\widehat{G}$ obtaind from
+    iterated application of PWGC Algorithm \ref{alg:finite_pwgc}}
+  \Initialize{$i = 1$, $\widehat{x}_0(t) \leftarrow x(t)$,
+    $\sigma_0 \leftarrow Var\ x(t)$}
+
+  \While{$i \le N$} {
+    Estimate $\VAR$ model via PWGC on $x_{i - 1}(t)$ to obtain $\widehat{x}_{i - 1}(t)$ and $\widehat{G}_{i - 1}$\\
+    $\epsilon_i(t) \leftarrow x_{i - 1}(t) - \widehat{x}_{i - 1}(t)$ \texttt{\# Compute residuals}\\
+    $\sigma_i^2 \leftarrow \text{MSE}\big(\epsilon_i(t)\big)$ \texttt{\# Compute MSE}\\
+    \If{$\sigma_i^2 > r \sigma_{i - 1}^2$}{
+      \texttt{break  \# Insignificant Improvement}
+    }
+    $i \leftarrow i + 1$
+  }
+  $\widehat{G} \leftarrow \bigcup_{i} \widehat{G}_{i}$ \texttt{\# Combine PWGC graph estimates}\\
+  \Return{$\widehat{G}$}
+\end{algorithm}
 
@@
 
-Figures \ref{fig:ar_eeg_example} and \ref{fig:eeg_oos_eeror}
+Figures \ref{fig:ar_eeg_example} and \ref{fig:eeg_oos_error}
 illustrate the appropriateness of modelling $x(t)$ with $\VAR$ models,
@@
 
-  {\scriptsize Figure \ref{fig:ar_eeg_example} provides an example of an
-  autoregressive model fir via the Levinson-Durbin algorithm (choosing
-  the system order via the BIC) on a single EEG channel.  Figure
-  \ref{fig:eeg_oos_error} demonstrates the improvement of a single
-  $\VAR$ model over independent $\mathsf{AR}$ models.
+  {\scriptsize Figure \ref{fig:ar_eeg_example} provides an example of
+    an autoregressive model fir via the Levinson-Durbin algorithm
+    (choosing the system order via the BIC) on a single EEG channel.
+    Figure \ref{fig:eeg_oos_error} demonstrates the improvement of a
+    single $\VAR$ model over independent $\mathsf{AR}$ models,
+    providing evidence that intra-node edges inferred by PWGC are not
+    simply spurious.
 
@@
   fitting the models.}
+\end{figure}
+
+We provide the final results in Figure
+\ref{fig:logistic_regression_results} where the classification
+accuracy is quantified by the (multiclass) MCC of $\approx 0.20$ on a
+held out validation set consisting of $20\%$ of the data.  This MCC
+score exceeds by a modest margin the performance of randomly guessing,
+lending strong evidence to the assertion that PWGC successfully
+recovers some meaningful differences between the Granger-causality
+graphs of different subjects, particularly if we recall that there are
+generally fewer examples for each subject than there are subjects
+overall.  Moreover, distinguishing only between two particular
+subjects (as opposed to distinguishing between one subject and $118$
+other possibilities) is substantially easier -- an illustration is
+provided in Figure \ref{fig:logistic_regression_2pair} where we embed
+the data into the plane through supervised dimensionality reduction.
+In general, fitting a classifier between two subjects achieves an
+$MCC$ in excess of $0.6$ and up to $0.9$.
+
+\begin{figure}
+  \centering
+  \caption{Subject Classification from Granger-causality Graphs}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \caption{Distinguishing Individual Subjects}
+    \label{fig:logistic_regression_2pair}
+    \includegraphics[width=\linewidth]{logistic_regression_separation.pdf}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \caption{Complete Multiclass Classifier}
+    \label{fig:logistic_regression_results}
+    \includegraphics[width=\linewidth]{logistic_regression_confusion_matrix.pdf}
+
+    {\scriptsize }
+  \end{subfigure}
+
+  {\scriptsize Figure \ref{fig:logistic_regression_2pair} illustrates
+    discriminating between two particular subjects based on their
+    EEG Granger-causality graphs.  In this case $MCC = 0.75$ on held
+    out data.  Visualization is constructed by supervised
+    dimensionality reduction and is purely illustrative.
+
+    Figure \ref{fig:logistic_regression_results} provides the
+    row-normalized confusion matrix (computed on held out validation
+    data) of a multiclass logistic regression classifier used to
+    classify subjects based on their EEG Granger-causality graphs.  $MCC = 0.20$}
 
@@
 assumptions may be reasonable in some applications and secondly, even
-when the strong-causality assumption is not met, we have provided some
-empirical evidence to suggest that our pairwise testing algorithm PWGC
-can still outperform the LASSO and adaLASSO, both of which are
-commonly employed in applications.  Application to more general graphs
-may also be within reach via an iterative graph building process,
-i.e. estimating a sequence of SCGs via PWGC applied to the residuals
-of the previous model.
+when the strong-causality assumption is not met, we have provided
+simulation evidence to suggest that our pairwise testing algorithm
+PWGC can still outperform the LASSO and adaLASSO, both of which are
+commonly employed in applications.  Application to inference of more
+general graphs may also be within reach via an iterative graph
+building process as suggested in Section \ref{sec:application}.
+
+Our application Section \ref{sec:application} provides strong evidence
+that PWGC is capable of uncovering meaningful features from networks
+of time series data by constructing a classifier which accurately
+discriminates between subjects in an EEG experiment based purely on
+the Granger-causality graph topology inferred by the iterated
+application of PWGC.
 
@@
 algorithm (by modifying the approach described in Section
-\ref{sec:pairwise_hypothesis_testing}) therefore is likely to enable
-application to very large networks of time series data which are not
-well approximated by finite $\VAR(p)$ models.
+\ref{sec:pairwise_hypothesis_testing} to instead utilize state-space
+Granger-causality testing) therefore is likely to enable application
+to very large networks of time series data which are not well
+approximated by finite $\VAR(p)$ models.
 

### 2019-07-09 software/eeg_analysis.py

@@ -261,7 +261,15 @@ def baseline_estimates(X, max_T, max_lag, plot_example=False):
     return V
 
 
-def full_graph_estimates(X, max_T, max_lag, N_iters=4, post_estimate="lstsqr"):
+def full_graph_estimates(X, max_T, max_lag, N_iters=4,
+                         r=0.9, post_estimate="lstsqr"):
+    """
+    Iterates the PWGC algorithm up to N_iters times, then performs a
+    global estimate using the method specified by post_estimate.
+    Using post_estimate="alasso" will result in post-filtering the
+    edges, using post_estimate="lstsqr" will keep the entire
+    support inferred by each iteration of PWGC.
+    """
     graph_estimates = []
 
     def _estimate(X_data):
@@ -278,7 +286,7 @@ def full_graph_estimates(X, max_T, max_lag, N_iters=4, post_estimate="lstsqr"):
         G_e, X_r = _estimate(X_r)
         graph_estimates.append(G_e)
         sv2_r = np.var(X_r)
-        if sv2_r / sv2_r_prev > 0.9:
+        if sv2_r / sv2_r_prev > r:
             break
         else:
             sv2_r_prev = sv2_r

### 2019-07-09 software/eeg_graph_analysis.py

@@ -46,47 +46,14 @@ mpl_rc("font", **font)
 mpl_rc("text", usetex=True)
 
 
-def main_Xy2():
+def logistic_regression_clf():
+    """
+    Compute and quantify the quality of a multinomial logistic regression
+    model between all subjects.
+    """
     X, y, z = load_subject_Xyz()
     le = LabelEncoder()
     y = le.fit_transform(y)
-
-    # 2
-    xform = Pipeline(
-        steps=[("pca", KernelPCA(n_components=5,
-                                 kernel="rbf",
-                                 gamma=0.75,
-                                 eigen_solver="arpack",
-                                 random_state=1,
-                                 )),
-               ("power_transform", PowerTransformer(method="yeo-johnson",
-                                                    standardize=True))])
-    X_pca = xform.fit_transform(X)
-    D_pca = pd.DataFrame(X_pca)
-    D_pca["label"] = ["Alcoholic" if yi == "A" else "Control" for yi in z]
-    pp = sns.pairplot(D_pca.iloc[::2, ], hue="label",
-                      vars=[c for c in D_pca.columns if c != "label"],
-                      plot_kws={"alpha": 0.25})
-    plt.show()
-
-    subjects = np.unique(y)
-    s0 = subjects[3]
-    s1 = subjects[-1]
-
-    sel = np.logical_or(y == s0, y == s1)
-    _y = y[sel]
-    _X = X[sel]
-    qt = QuantileTransformer(output_distribution="normal")
-    K = Nystroem(kernel="poly", gamma=0.75, degree=4, coef0=1.0, n_components=500,
-                 random_state=0)\
-                 .fit_transform(_X)
-    K = qt.fit_transform(K)
-
-    pls = PLSRegression(n_components=2, scale=False)
-    pls.fit(K, _y)
-    X_pls = pls.x_scores_
-    plt.scatter(X_pls[:, 0], X_pls[:, 1], c=_y)
-    plt.show()
 
     ss = StratifiedShuffleSplit(n_splits=1, test_size=0.2)
     train_ix, valid_ix = next(ss.split(X, y))
@@ -109,7 +76,7 @@ def main_Xy2():
                   "logistic__penalty": ["l2"]}
 
     cv = RandomizedSearchCV(estimator=clf, param_distributions=param_grid,
-                            n_iter=50, n_jobs=3, cv=3, refit=True)
[…]

### 2019-07-10 software/lasso_comparison.py
A variety of updates.  ts-lasso does not improve upon earlier results.

@@ -11,9 +11,9 @@ from statsmodels.sandbox.regression.predstd import wls_prediction_std
 from plotting_helpers import COLOR1, COLOR2, plotting_model
 
 from matplotlib import rc as mpl_rc
-font = {"family" : "normal",
-        "weight" : "bold",
-        "size"   : 22}
+font = {"family": "normal",
+        "weight": "bold",
+        "size": 22}
 
 mpl_rc("font", **font)
 mpl_rc("text", usetex=True)  # TODO: Fix my fonts
@@ -23,7 +23,8 @@ from pwgc.var_system import (random_gnp_dag, drive_gcg, get_errors,
                              get_X, attach_node_prop, gcg_to_var)
 from pwgc.gc_methods import (compute_pairwise_gc, estimate_B, pw_scg,
                              estimate_graph, full_filter_estimator,
-                             compute_MCC_score, estimate_dense_graph)
+                             compute_MCC_score, estimate_dense_graph,
+                             alasso_fista_estimate_dense_graph)
 
 
 class TrackErrors:
@@ -126,7 +127,7 @@ def simple_lasso():
     return
 
 
-def lasso_mcc_comparison(showfig=False):
+def lasso_mcc_comparison(simulation_name, showfig=False, fista=False):
     np.random.seed(0)
     n_nodes, p_lags, p_max = 50, 5, 15
     alpha = 0.05
@@ -144,7 +145,6 @@ def lasso_mcc_comparison(showfig=False):
     random_dag_graph_q4 = lambda: random_gnp_dag(
         n_nodes, p_lags, pole_rad=0.75,
         edge_prob=edge_prob_q4)
-
 
     def make_test_data(T, graph_type="DAG"):
         sv2_true = 0.5 + np.random.exponential(0.5, size=n_nodes)
@@ -184,9 +184,12 @@ def lasso_mcc_comparison(showfig=False):
             G_hat_pwgc = estimate_graph(X[:T], G, max_lags=p_max,
                                         method="lstsqr", alpha=alpha,
                                         fast_mode=fast_mode, F_distr=True)
-            G_hat_lasso = estimate_dense_graph(X, max_lag=p_max,
-                                               max_T=T,
-                                               method="alasso")
+            if fista:
+                G_hat_lasso = alasso_fista_estimate_dense_graph(
+                    X, max_lag=p_max, max_T=T)
+            else:
[…]

### 2019-07-10 software/pwgc/gc_methods.py

@@ -17,6 +17,7 @@ from scipy.linalg import toeplitz, cho_solve, cho_factor
 from sklearn.linear_model import LassoLarsIC, LassoLarsCV, Lasso
 from sklearn.preprocessing import StandardScaler
 from spams import fistaFlat
+from ts_lasso.ts_lasso import fit_VAR
 
 from levinson import (lev_durb, whittle_lev_durb)
 
@@ -27,12 +28,14 @@ try:
     from .stat_util import benjamini_hochberg
     from .var_system import (attach_node_prop, add_self_loops,
                              remove_zero_filters, get_X,
-                             make_complete_digraph, attach_X)
+                             make_complete_digraph, attach_X,
+                             gcg_to_var)
 except ImportError:
     from stat_util import benjamini_hochberg
     from var_system import (attach_node_prop, add_self_loops,
                             remove_zero_filters, get_X,
-                            make_complete_digraph, attach_X)
+                            make_complete_digraph, attach_X,
+                            gcg_to_var)
 
 
 def estimate_b_lstsqr(X, y, ret_G=False, lmbda=np.inf):
@@ -803,9 +806,40 @@ def estimate_dense_graph(X, max_lag=10,
     _, n = X.shape
     G = make_complete_digraph(n)
     G = attach_X(G, X)
-    assert np.all(X == get_X(G))
 
     estimate_B(G, max_lag, copy_G=False, max_T=max_T, method=method)
+    G = remove_zero_filters(G, "b_hat(z)", copy_G=False)
+    return G
+
+
+def alasso_fista_estimate_dense_graph(X, max_lag=10, max_T=None, nu=1.25,
+                                      eps=1e-3, full_path=False):
+    T, n = X.shape
+    if max_T is None:
+        max_T = T
+
+    B, _, _, _ = fit_VAR(X[:max_T, :], max_lag, nu=nu, eps=eps,
+                         full_path=full_path)
+
+    G = make_complete_digraph(n)
+    G = attach_X(G, X)
+
+    for i in range(n):
+        for j in range(n):
+            b_hatz = B[:, i, j]
+            G[j][i]["b_hat(z)"] = b_hatz
+
+    if max_T < T:
+        p = len(B)
+        z0 = np.hstack([X[max_T - tau] for tau in range(1, p + 1)])
+        var_sys = gcg_to_var(G, filter_attr="b_hat(z)", assert_stable=False)
[…]

### 2019-07-10 software/pwgc/test_gc_methods.py

@@ -15,7 +15,7 @@ from var_system import (random_gnp_dag, drive_gcg, get_errors,
 from gc_methods import (compute_pairwise_gc, estimate_B, pw_scg,
                         estimate_graph, full_filter_estimator,
                         compute_covariances, compute_xi, fast_compute_xi,
-                        normalize_gc_score)
+                        normalize_gc_score, fast_compute_pairwise_gc)
 from draw_graphs import draw_graph_estimates
 from stat_util import benjamini_hochberg
 

### 2019-07-10 software/pwgc/var_system.py

@@ -531,12 +531,32 @@ class VAR:
         # Output matrix to be returned
         Y = np.empty((T, n))
         for t in range(T):
-            y = np.dot(self._B, self._z) + u[t, :]  # Next output
+            y = self.estimate_next_step() + u[t, :]
             Y[t, :] = y
-            self._z = np.roll(self._z, n)  # Update system state
-            self._z[:n] = y
+            self.update_state(y)
         self.x = self._z[:n]  # System output
         return Y
+
+    def estimate_next_step(self):
+        return np.dot(self._B, self._z)
+
+    def update_state(self, x):
+        self._z = np.roll(self._z, self.n)
+        self._z[:self.n] = x
+        return
+
+    def compute_residuals(self, X):
+        T, n = X.shape
+        assert n == self.n
+
+        R = np.zeros((T, n))
+        for t in range(T):
+            x = X[t, :]
+            x_hat = self.estimate_next_step()
+            r = x - x_hat
+            R[t, :] = r
+            self.update_state(x)
+        return R
 
 
 def _symmetric_sum(A_list, k):

### 2019-07-10 paper/paper.tex
typo.

@@
 For later use, and to get a feel for the topological implications of
-strong causality We explore a number of properties of such graphs
+strong causality, we explore a number of properties of such graphs
 before moving into the main result of this section.  The following

### 2019-07-11 paper/paper.tex
Minor updates, corrects, etc.

@@
 \usepackage{hyperref}
+\usepackage{fullpage}
+\setlength{\marginparwidth}{2cm}
 \usepackage{todonotes}
-\usepackage{fullpage}
 \bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
@@
 
-\todo{Do we have $A_{ij}(z) \ne 0 \implies j \in \anc{i}$?}
-
 % TODO: ???
@@
   where $(I - \B(z))^{-1} = \A(z)$ due to the uniqueness of
-  (\ref{eqn:wold}).  Since $\B(z)$ is stable, that is
-  $|\lambda_{\text{max}}(B(z))| < 1$, for every $|z| \le 1$ we have
+  (\ref{eqn:wold}).  Since $\B(z)$ is stable we have
 
@@
     \label{eqn:resolvant_inv}
-    (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k,
+    (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k.
   \end{equation}
 
-  where we can apply the Cayley-Hamilton theorem to write the infinite
-  sum of \eqref{eqn:resolvant_inv} in terms of \textit{finite} powers
-  of $\B$.
-
-  Let $S$ be a $0-1$ matrix containing the sparsity pattern of
-  $\B(z)$, from lemma \ref{lem:adj_matrix} $S$ is the transpose of the
-  adjacency matrix for $\gcg$ and hence $(S^k)_{ij}$ is non-zero if
-  and only if $j \in \gpn{k}{i}$, and therefore $\B(z)^k_{ij} = 0$ if
-  $j \not \in \gpn{k}{i}$.  Finally, since
-  $\anc{i} = \bigcup_{k = 1}^n\gpn{k}{i}$ we see that $\A_{ij}(z)$ is
-  zero if $j \not\in \anc{i}$.
+  Invoking the Cayley-Hamilton theorem allows writing the infinite sum
+  of \eqref{eqn:resolvant_inv} in terms of \textit{finite} powers of
+  $\B$.
+
+  Let $S$ be a matrix with elements in $\{0, 1\}$ which represents the
+  sparsity pattern of $\B(z)$, from lemma \ref{lem:adj_matrix} $S$ is
+  the transpose of the adjacency matrix for $\gcg$ and hence
+  $(S^k)_{ij}$ is non-zero if and only if $j \in \gpn{k}{i}$, and
+  therefore $\B(z)^k_{ij} = 0$ if $j \not \in \gpn{k}{i}$.  Finally,
+  since $\anc{i} = \bigcup_{k = 1}^n\gpn{k}{i}$ we see that
+  $\A_{ij}(z)$ is zero if $j \not\in \anc{i}$.
 
@@
   \begin{align*}
-    x_i(t) &= (I - \B(z))^{-1}v(t)\\
+    x_i(t) &= [(I - \B(z))^{-1}v(t)]_i\\
     &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
@@
 conceptual standpoint, we will see how the notion can in some sense
-capture the idea of ``flow of information'' in the underlying graph.
-It may also be useful for reasoning about what the conditions under
-which \textit{unobserved} components of $x(t)$ may or may not
-interfere with inference in the actually observed components.
-Finally, motivated from a practical standpoint to analyze causation in
-large systems, we will seek to construct practical estimation
-procedures based purely on pairwise causality tests since the
-computation of such pairwise relations is somewhat easier.
-\todo{Reconsider this paragraph}
+capture the idea of ``flow of information'' in the underlying graph,
+in the sense that if $j \in \anc{i}$ we expect that $j \pwgc i$.  It
+may also be useful for reasoning about the conditions under which
+\textit{unobserved} components of $x(t)$ may or may not interfere with
+inference in the actually observed components.  Finally, motivated
+from a practical standpoint to analyze causation in large systems, we
+will seek to construct practical estimation procedures based purely on
+pairwise causality tests since the computation of such pairwise
+relations is somewhat easier.
 
@@
 \begin{remark}
-  Note that $H_t^{(i)} \perp \H_t^{(j)}$ states that the
-  \textit{present} of $x_i(t)$ is uncorrelated with $x_j(t)$,
-  but we need not have $\H_t^{(i)} \perp \H_t^{(j)}$ since
-  $i \in \anc{j}$ is not excluded.
+  Note that $H_t^{(i)} \perp \H_t^{(j)}$ states that the \textit{present}
+  of $x_i(t)$ is uncorrelated with the whole history of $x_j(t)$, but
+  we need not have $\H_t^{(i)} \perp \H_t^{(j)}$ since $i \in \anc{j}$ is
+  not excluded.
 \end{remark}
@@
 
-The previous proposition can be strengthened significantly for the
-case of Granger-causality.  Note that it is possible to have some
-$k \in \anc{i} \cap \anc{j}$ where still $j \npwgc i$, consider for example
-the three node graph $k \rightarrow i \rightarrow j$ where
-clearly $k \in \anc{i}\cap\anc{j}$ but $j \npwgc i$.  We must
-introduce the concept of a \textit{confounding} variable, which
-effectively eliminates the possibility presented in this example.
+The previous proposition can be strengthened significantly; notice
+that it is possible to have some $k \in \anc{i} \cap \anc{j}$ where still
+$j \npwgc i$, an example is furnished by the three node graph
+$k \rightarrow i \rightarrow j$ where clearly
+$k \in \anc{i}\cap\anc{j}$ but $j \npwgc i$.  We must introduce the concept
+of a \textit{confounding} variable, which effectively eliminates the
+possibility presented in this example.
 
@@
 
-  We notice that for any $u$ having the properties $a, b$ above, we
+  We notice that for any $u$ having the properties $(a), (b)$ above, we
   must have $C_2(u) = \emptyset$ since if $k \in C_2(u)$ then
@@
 
-  which follows from Lemma \ref{lem:ancestor_uncorrelated}.
-  Finally, for $k \in C_1(j)$ the case $k = i$ is immediate (since the
-  error in estimating $\Phi(z)\B_{ii}(z)x_i(t)$ given
-  $\H_{t - 1}^{(i)}$ is $0$), so suppose $k \ne i$.  We know from
-  above that $k$ inherits the key properties referred to as $a, b$ and
-  therefore we can recursively expand $k$ in the same way as in
-  equation (\ref{eqn:xj_partition_expansion}).  Continuing this
-  recursion for each $k \in C_1(j)$ (where $k \ne i$) must eventually
-  terminate since $i \in \anc{k}$.
+  which follows from Lemma \ref{lem:ancestor_uncorrelated}.  Finally,
+  for $k \in C_1(j)$ the case $k = i$ is immediate (since the error in
+  estimating $\Phi(z)\B_{ii}(z)x_i(t)$ given $\H_{t - 1}^{(i)}$ is
+  $0$), so suppose $k \ne i$.  We know from above that $k$ inherits the
+  key properties referred to as $(a)$ and $(b)$ above and therefore we
+  can recursively expand $k$ in the same way as in equation
+  (\ref{eqn:xj_partition_expansion}).  Continuing this recursion for
+  each $k \in C_1(j)$ (where $k \ne i$) must eventually terminate since
+  $i \in \anc{k}$.
 \end{proof}
@@
   will never result in $i \pwgc j$ or $j \pwgc i$.  This is evidently
-  an advantage of the \textit{temporal} nature of Granger -- there is
+  an advantage of the \textit{temporal} nature of Granger-causality -- there is
   no backwards causal flow along the edges of $\gcg$.
@@
 complete a converse by adding the additional requirement we refer to
-as ``persistence''.  This condition requires that each node maintains
-some ``memory'' of the past.
+as ``persistence''.
 
@@
   where $\tau_\infty(\B) \defeq \infty$ if the filter has an infinite
-  length.
+  length, and $\tau_0(\B) \defeq \infty$ if $\B(z) = 0$.
 \end{definition}
 
+This interpretation of the following persistence condition is that
+each node stores some ``memory'' of the past.
+
 \begin{definition}[Persistent]
-  We will say that a WSS process $x(t)$ with Granger-causality graph
-  $\gcg$ is \textit{persistent} if for every $i \in [n]$, every
-  $k \in \anc{i}$ we have $\tau_\infty(\A_{ik}) = \infty$.
+  We will say that the process $x(t)$ with Granger-causality graph
+  $\gcg$ is \textit{persistent} if for every $i \in [n]$ and every
+  $k \in \anc{i}$ we have $\tau_0(\A_{ik}) < \infty$ and $\tau_\infty(\A_{ik}) = \infty$.
 \end{definition}
@@
   results of this section, the condition
-  $\tau_0(\A_{jk}) < \tau_\infty(\A_{ik})$ for each $i, j, k$ such
-  that $k \in \anc{i} \cap \anc{j}$ is enough.
-  \todo{Write something on the intuition here.}
+  $\tau_0(\A_{jk}) < \tau_\infty(\A_{ik})$ for each $i, j, k$ such that
+  $k \in \anc{i} \cap \anc{j}$ is enough.  The intuition being that nodes
+  $i$ and $j$ are not receiving temporally disjoint information from
+  $k$.
+
 \end{remark}
@@
   had a solution in $z$ for every $r \ge N$, where $V$ is an
-  $n \times n - 1$ full-rank matrix spanning the nullspace of $\beta$,
+  $n \times n - 1$ full-rank matrix whose columns span the nullspace of $\beta$,
   and $\bm{\lambda} = (\lambda_1, \ldots, \lambda_n)$. That is,
@@
 
-  \todo{This may remain true for any matrix with at least 2 distinct eigenvalues}
-  We can conclude by stating that a $\VAR(1)$ system corresponding to
-  matrix $B$ is guaranteed to be persistent if $B$ is diagonalizable
-  and has more than $1$ distinct eigenvalue.
+  We see from this example that the collection of $\VAR(1)$ systems
+  which are not persistent are pathological, in the sense that their
+  system matrices have zero measure when viewed as a subset of $\R^{n^2}$.
 \end{example}
@@
 \begin{proof}
-  The final equivalence follows immediately from the Lemma.  For the first equivalence we have
+  The final equivalence follows immediately from Lemma \ref{lem:time_lag_cancellation}.  For the first equivalence we have
 
@@
   \label{prop:persistence_converse}
-  Suppose $\gcg$ is a strongly causal DAG and that $x(t)$ is persistent, then if $k$
-  confounds $(i, j)$ we have $i \pwgc j$ and $j \pwgc i$.
+  Suppose $\gcg$ is a strongly causal DAG and that $x(t)$ is
+  persistent, then if there exists a $k$ which confounds $(i, j)$ we
+  have $i \pwgc j$ and $j \pwgc i$.
 \end{proposition}
@@
 
-  Evidently, this is $0$ for each $F$ if and only if
-  $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$ and
-  $\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} = 0$, which by
+  This is $0$ for every $F$ if and only if for all $F$ and
+  $\forall k \in \anc{i} \cap \anc{j}$ we have
+  $$\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$$ and
+  $$\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} = 0,$$ which by
   Corollary \ref{cor:time_lag_cancellation} occurs if and only if
-  $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$, which is impossible
-  since in general $\tau_0(\A_{jk}) < \infty$ and for a persistent
-  system $\tau_\infty(\A_{ik}) = \infty$.
+  $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$, which is impossible since by persistence
+  $\tau_0(\A_{jk}) < \infty$ and  $\tau_\infty(\A_{ik}) = \infty$.
 \end{proof}
@@
 process is a strongly causal DAG, then it is possible to recover
-$\gcg$ via pairwise tests alone.
+$\gcg$ via pairwise tests alone.  This is the main conclusion of the
+theoretical analysis in this paper.
 
@@
   next iteration, the edges $(3, 4), (2, 4), (3, 5)$ are added, and
-  the false edges $(1, 4), (1, 5)$ is excluded due to the paths
+  the false edges $(1, 4), (1, 5)$ are excluded due to the paths
   $1 \rightarrow 3 \rightarrow 4$ and $1 \rightarrow 3 \rightarrow 5$
-  already present.  Finally, edge $(4, 6)$ is added, and the false
+  already being present.  Finally, edge $(4, 6)$ is added, and the false
   $(1, 6), (3, 6), (2, 6)$ edges are similarly excluded due to the
@@
 
-If disturbances are Gaussian, we will have the convergence
+If the estimation procedure is consistent, we will have the following
+convergence (in $\P$ or a.s.):
 
 \begin{equation}
-  F_{ij}(p) \overset{\P}{\rightarrow}
+  F_{ij}(p) \rightarrow
   \left\{
@@
 \label{sec:model_order_selection}
-There are a variety of methods to choose the filter orders $p_i$ and
-$p_j$ (see e.g. \cite{lutkepohl2005new}), but we will focus in
-particular on the Bayesian Information Criteria (BIC).  The BIC is
-substantially more conservative than the popular alternative Akaiake
-Information Criteria (the BIC is also asymptotically consistent), and
-since we are searching for \textit{sparse graphs}, we therefore prefer
-the BIC, where we seek to \textit{minimize} over $p$:
+There are a variety of methods to choose the filter order $p$ (see
+e.g. \cite{lutkepohl2005new}), but we will focus in particular on the
+Bayesian Information Criteria (BIC).  The BIC is substantially more
+conservative than the popular alternative Akaiake Information Criteria
+(the BIC is also asymptotically consistent), and since we are
+searching for \textit{sparse graphs}, we therefore prefer the BIC,
+where we seek to \textit{minimize} over $p$:
 
@@
 We carry this out by a simple direct search on each model order
-between $0$ and some prescribed $p_\text{max}$.  In practice it is
+between $0$ and some prescribed $p_\text{max}$, resulting in a
+collection $p_{ij}$ of model order estimates.  In practice, it is
 sufficient to pick $p_\text{max}$ ad-hoc or via some simple heuristic
 e.g. plotting the sequence $BIC(p)$ over $p$, though it is not
-technically possible to guarantee that the optimal $p$ is chosen
-(since there can in general be arbitrarily long lags from one variable
-to another).
+technically possible to guarantee that the optimal $p$ is less than
+the chosen $p_\text{max}$ (since there can in general be arbitrarily
+long lags from one variable to another).
 
@@
 implementing our estimation algorithm is spent calculating the error
-estimates $\xi_i(p_i)$ and $\xi_{ij}(p_{ij})$.  This requires fitting
-a total of $n^2p_{\text{max}}$ autoregressive models, where the most
+estimates $\xi_i(p_i)$ and $\xi_{ij}(p_{ij})$.  This requires fitting a
+total of $n^2p_{\text{max}}$ autoregressive models, where the most
 naive algorithm (e.g. solving a least squares problem for each model)
@@
 $O(n^2p_{\text{max}}^2 ) + O(n^2p_{\text{max}}T)$ time via the
-autocorrelation method which deploys the following autocovariance
-estimates in the Yule-Walker equations:\footnote{The particular
-  indexing and normalization given in equation
-  \ref{eqn:covariance_estimate} is critical to ensure $\widehat{R}$ is
-  positive semidefinite.  The estimate can be viewed as calculating
-  the covariance sequence of a signal multiplied by a rectangular
-  window.}
+autocorrelation method
+\cite{hayes_statistical_digital_signal_processing} which substitutes
+the following autocovariance estimates in the Yule-Walker
+equations:\footnote{The particular indexing and normalization given in
+  equation \ref{eqn:covariance_estimate} is critical to ensure
+  $\widehat{R}$ is positive semidefinite.  The estimate can be viewed
+  as calculating the covariance sequence of a signal multiplied by a
+  rectangular window.}
 
@@
 \end{equation}
+
+It is imperative that the first index in the summation is $\tau + 1$, as
+opposed perhaps to $p_\text{max}$ and that the normalization is
+$1 / T$, as opposed perhaps to $1 / (T - p_\text{max})$, in order to
+guarantee that $\widehat{R}_x(\tau)$ forms a valid (i.e. positive
+definite) covariance sequence.  This results in some bias, however the
+dramatic computational speedup is worth it for our purposes.
 
@@
 \ref{eqn:gc_statistics} with model orders chosen by the methods of
-Section \ref{sec:model_order_selection}.  Note that this statistic is
-asymptotically $\chi^2(p_{ij})$ distributed, and denote by $G$ the
-cumulative distribution function thereof.  We will define the matrix
+Section \ref{sec:model_order_selection}.  We assume that this
+statistic is asymptotically $\chi^2(p_{ij})$ distributed (the
+disturbances are Gaussian), and denote by $G$ the cumulative
+distribution function thereof.  We will define the matrix
 
@@
 
-to be the matrix of pairwise edge inclusion probabilities.  This is
+to be the matrix of pairwise edge inclusion P-values.  This is
 motivated by the hypothesis test where the hypothesis $H_0$ will be
@@
 pairwise edges to a level $\alpha$ (where we generally take
-$\alpha = 0.05$).  Applying the extremely simple Bonferroni method is
-overly conservative for this application.
+$\alpha = 0.05$).
 
@@
 specified sparsity pattern directly via ordinary least squares.
-Producing graph estimates which are not strongly causal can be
-achieved out by performing sequential estimates
-$\widehat{x}_1(t), \widehat{x}_2(t), \ldots$ by estimating a strongly
-causal graph with the residuals of the previous model as input, and
-then refitting on the estimated sparsity pattern.  However we reserve
-any analysis of this scheme for future work.
+
+We note that producing graph estimates which are not strongly causal
+can potentially be achieved by performing sequential estimates
+$\widehat{x}_1(t), \widehat{x}_2(t), \ldots$ estimating a strongly causal
+graph with the residuals of the previous model as input, and then
+refitting on the combined sparsity pattern.  We experiment with this
+heuristic in our example application of Section \ref{sec:application},
+but reserve theoretical analysis for future work.
 
@@
 
-We compare our results against the LASSO.  Motivated by scaling, we
-split the squared error term into separate terms, one for each group of
-incident edges on a node, and estimate the collection of $n$ incident
-filters $\big\{\B_{ij}(z)\big\}_{j = 1}^n$ that minimizes
+We compare our results against the adaptive LASSO
+\cite{adaptive_lasso_zou2006}, which outperformed substantially both
+the LASSO and the grouped LASSO.  Motivated by scaling, we split the
+squared error term into separate terms, one for each group of incident
+edges on a node, and estimate the collection of $n$ incident filters
+$\big\{\B_{ij}(z)\big\}_{j = 1}^n$ that minimizes
 $\xi_i^{\text{LASSO}}$ in the following:
@@
   \includegraphics[width=0.95\textwidth]{new_lasso_comparison_dag_pmax15_simulation.pdf}
+
   {\scriptsize Figure \ref{fig:simulation_results_comparison2}
@@
 \ref{fig:simulation_results_dense} illustrates the severe (expected)
-degradation in performance as the number of DAG edges increases while
+degradation in performance as the number of edges increases while
 the number of data samples $T$ remains fixed.  For larger values $q$
@@
 number of data samples.
+
+We have also paid close attention to the performance of PWGC in the
+very small sample ($T \le 100$) regime (see Figure
+\ref{fig:small_T_comparison}), as this is the regime many applications
+must contend with.
 
@@
 demonstration of this scalability, where we are able to estimate
-graphs having over $1500$ nodes (over $2.25 \times 10 ^6$ possible
-edges) using only $T = 500$ data points, granted, an SCG on this many
-nodes is extremely sparse.  However, it is certainly possible to
-improve the scalability of the LASSO as applied to time series data by
-taking advantage of the block-Toeplitz covariance structure.
+graphs having over $1500$ nodes (over $2.25 \times 10 ^6$ possible edges)
+using only $T = 500$ data points, granted, an SCG on this many nodes
+is extremely sparse.
 
@@
 ``alcoholic'', however, we will ignore this label and instead focus on
-distinguishing between subjects based on the Granger-causality graphs
-inferred from the subject's trials.  Our reasoning is to focus on the
-underlying question: ``Does our PWGC algorithm uncover meaningful
-Granger-causal connections from EEG data?''.  Focusing only on the
-subject label allows us to answer this question without simultaneously
-grappeling with the physiological question of whether alcoholic
-subjects as a group have discernable differences in their EEG
-readings\footnote{Some exploratory analysis actually suggests
-  that alcoholic Granger-causality graphs are not substantively
-  different}, which is a stronger requirement than simply that there
-are meaningful distinctions between the EEG readings of subjects
-generally.
+\textit{distinguishing between subjects} based on the
+Granger-causality graphs inferred from the subject's trials.  Our
+reasoning is to focus on the underlying question: ``Does our PWGC
+algorithm uncover meaningful Granger-causal connections from EEG
+data?''.  Focusing only on the subject label allows us to answer this
+question without simultaneously grappeling with the physiological
+question of whether alcoholic subjects as a group have discernable
+differences in their EEG readings\footnote{Some exploratory analysis
+  actually suggests that alcoholic Granger-causality graphs are not
+  substantively different}, which is a stronger requirement than
+simply that there are meaningful distinctions between the EEG readings
+of subjects generally.
 
@@
 
-The iterated application of PWGC is an example of a heuristic by which
-graph estimates that aren't constrained to being strongly causal can
-be obtained.  Theoretical analysis of more sophisticated heuristics
-than PWGC (Algorithm \ref{alg:finite_pwgc}) is deferred to future work.
+\begin{remark}
+  The iterated application of PWGC is an example of a heuristic by
+  which graph estimates that aren't constrained to being strongly
+  causal can be obtained.  We defer to future work the theoretical
+  analysis of more sophisticated heuristics, or algorithms appropriate
+  for different topological assumptions.
+\end{remark}
 
@@
 relevant information is being captured by the graph estimates
-$\widehat{G}^{(i)}$.
+$\widehat{G}^{(i)}$.  That is, that the latent structure
+$\widehat{G}^{(i)}$ may provide scientifically relevant insights, as
+opposed to simply being a feature for classification tasks.
 
@@
     (choosing the system order via the BIC) on a single EEG channel.
-    Figure \ref{fig:eeg_oos_error} demonstrates the improvement of a
-    single $\VAR$ model over independent $\mathsf{AR}$ models,
-    providing evidence that intra-node edges inferred by PWGC are not
-    simply spurious.
+    Linear autoregressive models appear qualitatively to be adequate
+    for this data.  Figure \ref{fig:eeg_oos_error} demonstrates the
+    improvement of a unified $\VAR$ model over independent
+    $\mathsf{AR}$ models, providing evidence that intra-node edges
+    inferred by PWGC are not simply spurious.
 
@@
 \ref{fig:logistic_regression_results} where the classification
-accuracy is quantified by the (multiclass) MCC of $\approx 0.20$ on a
-held out validation set consisting of $20\%$ of the data.  This MCC
-score exceeds by a modest margin the performance of randomly guessing,
+accuracy is quantified by the (multiclass) MCC of $\approx 0.20$ on a held
+out validation set consisting of $20\%$ of the data.  This MCC score
+exceeds by a modest margin the performance of randomly guessing,
 lending strong evidence to the assertion that PWGC successfully
@@
 subjects (as opposed to distinguishing between one subject and $118$
-other possibilities) is substantially easier -- an illustration is
-provided in Figure \ref{fig:logistic_regression_2pair} where we embed
-the data into the plane through supervised dimensionality reduction.
-In general, fitting a classifier between two subjects achieves an
-$MCC$ in excess of $0.6$ and up to $0.9$.
+others) is substantially easier -- an illustration is provided in
+Figure \ref{fig:logistic_regression_2pair} where we embed the data
+into the plane through supervised dimensionality reduction.  In
+general, fitting a simple classifier between two subjects achieves an
+$MCC$ in excess of $0.6$ and up to $0.9$, even when the same
+hyperparameters are carried over for different pairs of subjects.
 
@@
 
-Our application Section \ref{sec:application} provides strong evidence
-that PWGC is capable of uncovering meaningful features from networks
-of time series data by constructing a classifier which accurately
-discriminates between subjects in an EEG experiment based purely on
-the Granger-causality graph topology inferred by the iterated
-application of PWGC.
+Our application in Section \ref{sec:application} provides strong
+evidence that PWGC is capable of uncovering meaningful features from
+networks of time series data by constructing a classifier which
+accurately discriminates between subjects in an EEG experiment based
+purely on the Granger-causality graph topology inferred by the
+iterated application of PWGC.
 

### 2019-07-12 paper/paper.tex
Remove todo note.
Alphabetized keywords.
Amend keywords.
Breaking paper into separate pieces of work.  Also fixed significant error in one of the Lemmas.

@@
 \documentclass[12pt]{article}
-
-% TODO: read this https://lemire.me/blog/rules-to-write-a-good-research-paper/
 
@@
 
-\keywords{Granger-causality, causality graph, time series, vector autoregression,
-  LASSO, gene regulatory networks, network learning, EEG}
+\keywords{causality graph, EEG, gene regulatory networks,
+  Granger-causality, LASSO, network learning, time series, vector
+  autoregression}
 
@@
 notion of causality graph topology which we term ``strongly causal''
-and show that $\VAR$ models having this structure satisfy natural
-intuitive notions of ``information flow'' through the graph.
-Moreover, we show that such graphs are perfectly recoverable with only
-\textit{pairwise} Granger-causality tests, which would otherwise suffer
-from serious confounding problems.  Our finite sample results are
-based on simulations, where we show promising results for these graph
-topologies where our algorithm performs substantially better in our
-simulation setup than do competing LASSO algorithms, even for graphs
-that do not exactly satisfy our strongly-causal topology assumptions.
+and show that stationary times series whose underlying causality graph
+has this structure satisfy natural intuitive notions of ``information
+flow'' through the graph.  Moreover, we show that such graphs are
+perfectly recoverable with only \textit{pairwise} Granger-causality
+tests, which would otherwise suffer from serious confounding problems.
+Our finite sample results are based on simulations, where we show
+promising results for these graph topologies where our algorithm
+performs substantially better in our simulation setup than do
+competing LASSO algorithms, even for graphs that do not exactly
+satisfy our strongly-causal topology assumptions.
 
@@
 topology may have application in this field (see Section
-\ref{sec:strongly_causal_graphs})
+\ref{sec:strongly_causal_graphs}).
 

### 2019-07-12 paper_theory/paper_theory.tex
Breaking paper into separate pieces of work.  Also fixed significant error in one of the Lemmas.

--- 2019-07-12 paper/paper.tex
@@
   from the gene regulatory network literature suggesting that our
-  concept of a strongly causal graph may be applicable to this field.
-  We implement and detail finite-sample heuristics derived from our
-  theory, and establish through simulation the efficiency gains (both
-  statistical and computational) which can be obtained (in comparison
-  to LASSO algorithms) when structural assumptions are met.  Finally,
-  we provide an example application where we accurately discriminate
-  between subjects in an EEG study based purely on the
-  Granger-causality graphs inferred from data by our algorithms,
-  demonstrating that meaningful features are captured by the
-  Granger-causal graph topology.}
-
-\keywords{causality graph, EEG, gene regulatory networks,
-  Granger-causality, LASSO, network learning, time series, vector
+  concept of a strongly causal graph may be applicable to this field.}
+
+\keywords{causality graph, gene regulatory networks,
+  Granger-causality, network learning, time series, vector
   autoregression}
@@
 \label{sec:introduction}
-% Consider a collection of stochastic processes producing observations
-% at discrete time intervals.  Are the underlying processes dependent?
-% Can we quantify any of the underlying relationships?  Can the arrow of
-% time help us to distinguish a directionality or flow of dependence
-% among our observed series?  In this paper we contribute to the
-% understanding of the notion of Granger-Causality
-% \cite{granger1969investigating} \cite{Granger1980329} as a tool for
-% answering these questions.
-
-% Heuristically, we state that if an event '$A$' occuring strictly prior
-% in time to an even '$B$' and $A$ provides us with exclusive (that is,
-% not available anywhere else) information about $B$, then $A$ must have
-% had a causal impact on $B$.  This is an entirely model-free notion of
-% cause, and instead leverages the intuition that a cause must precede
-% it's effect.
-
 In this paper we study the notion of Granger-causality
@@
 established under less restrictive conditions than for the vanilla
-LASSO.  Our experimental comparisons in Section
-\ref{sec:empirical_evaluation} are against the
-adaLASSO. % Success of the LASSO has lead to
-% the famous ``bet on sparsity'' principle ``Use a procedure that does
-% well in sparse problems, since no procedure does well in dense
-% problems.''  \cite{tibshirani2015statistical}.
+LASSO.  We study a novel perspective on the problem by establishing
+conditions under which simple \textit{pairwise} Granger-causality
+tests are sufficient to recover the entire graph.
+
+%   Prior assumptions
+% on the \textit{topology} of the underlying Granger-causality graph can
+% result in tests more powerful than more generally applicable
+% algorithms, and we provide examples from the literature where
+% simplifying topological assumptions may be met.
 
@@
 tests, which would otherwise suffer from serious confounding problems.
-Our finite sample results are based on simulations, where we show
-promising results for these graph topologies where our algorithm
-performs substantially better in our simulation setup than do
-competing LASSO algorithms, even for graphs that do not exactly
-satisfy our strongly-causal topology assumptions.
+Aside from being an interesting theoretical perspective, prior
+assumptions about the underlying graph (similarly to sparsity
+assumptions) can greatly improve upon the statistical power of
+causality graph recovery algorithms when the assumptions are met.
 
@@
 which a fully conditional Granger-causality graph can be recovered
-from pairwise tests alone (sec \ref{sec:pairwise_algorithm}).
-Secondly, we propose in section \ref{sec:structure_learning} a graph
-search heuristic which implements our theoretical results to finite
-data samples, specifying and summarizing appropriate methods for
-hypothesis testing (section \ref{sec:pairwise_hypothesis_testing}),
-model order selection (section \ref{sec:model_order_selection}),
-computationally efficient estimation (section
-\ref{sec:efficient_model_estimation}), and error rate controls
-(section \ref{sec:error_rate_control}).  Our heuristics are compared
-against the adaLASSO algorithm in Section
-\ref{sec:empirical_evaluation}.  We stress the scalability of our
-algorithm which is capable of comfortably handling hundreds or
-thousands of nodes on a single machine, as opposed to standard LASSO
-algorithms which do not take advantage of the special structure
-associated with stationary time series data.  In section
-\ref{sec:application} we develop an example application where a
-classifier is constructed to accurately discriminate between subjects
-in an EEG study based only on the Granger-causality graph topology
-inferred by our algorithms.  Concluding remarks on further open
-problems and extentions are provided in Section \ref{sec:conclusion}.
-
-\section{Theory}
+from pairwise tests alone (sec \ref{sec:pairwise_algorithm}).  We
+report \hl{elsewhere}\todo{where?  Or can I fit in a brief summary?}
+the results of extensive finite sample simulations as well as an
+application to EEG data, where in addition we summarize appropriate
+methods for hypothesis testing, model order selection, computationally
+efficient estimations, and error rate controls in practice.  Our
+simulation results establish significant potential for improvement
+over existing methods.  Concluding remarks on further open problems
+and extentions are provided in Section \ref{sec:conclusion}.
+
+\section{Graph Topological Aspects of Granger-causality}
 \label{sec:theory}
@@
 wide-sense stationary (WSS) $n$-dimensional vector valued process
-$x(t)$ (with $t \in \Z$) where the $n$ elements take values in $L_2$.  We
-suppose that $x(t)$ has zero mean, $\E x(t) = 0$, and has absolutely
-summable matrix valued covariance sequence
-$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ with an absolutely
-continuous spectral density $S(\omega)$:
-
-\begin{equation}
-  \label{eqn:fourier_pair}
-  \begin{aligned}
-    R(\tau) &= \frac{1}{2\pi}\int_{-\pi}^\pi S(\omega) e^{j\tau\omega}\d \omega,\\
-    S(\omega) &= \sum_{\tau=-\infty}^\infty R(\tau)e^{-j\tau\omega}.
-  \end{aligned}
-\end{equation}
-
-% and that the spectra is bounded uniformly away from 0:
-
-% \begin{equation*}
-%   \int_{-\pi}^\pi \ln\det S(\omega) \d\omega > -\infty.
-% \end{equation*}
+$x(t)$ (with $t \in \Z$) where the $n$ elements take values in $L_2$.
+We suppose that $x(t)$ has zero mean, $\E x(t) = 0$, and has
+absolutely summable matrix valued covariance sequence
+$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$, and an absolutely continuous
+spectral density.
 
@@
 
-where $c(t)$ is a purely deterministic sequence\footnote{the purely
-  deterministic sequence $c(t)$ is one which lies in the remote past
-  $\bigcap_{\tau=1}^\infty \H_{t - \tau}^x$ of the process.  For such
-  processes a single sample $c(t_0)$ is enough to determine $c(t)$ for
-  every $t$.  For example,
-  $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi,
-  \pi]$}, $v(t)$ is an uncorrelated sequence and $A(0) = I$.  We will
-assume that $c(t) = 0$, which in practice is to say that the process
-has been detrended.  We additionally require that this representation
-can be inverted to yield the $\VAR(\infty)$ form
+where $c(t)$ is a purely deterministic sequence$v(t)$ is an
+uncorrelated sequence and $A(0) = I$.  We will assume that $c(t) = 0$.
+We additionally require that this representation can be inverted to
+yield the $\VAR(\infty)$ form
 
@@
 
-Finally, since $||z^{-1}|| = 1$ we have the inversion formula
+Finally, we have the inversion formula
 
@@
 which is referred to as a lack of instantaneous feedback in $x(t)$.
-We formally state our setup to conclude this section:
+We formally state our setup as a definition, which is the setup for
+the remainder of the paper:
 
@@
 observations of a common cause, and existing GC relations can
-disappear if we observe a new mediating series.
-
-The notion is closely related to the information theoretic measure of
-transfer entropy, indeed, if the distribution of $v(t)$ is known to be
-Gaussian then they are equivalent \cite{barnett2009granger}.
-
-We require two technical facts we refer to in the sequel, the notion
-of conditional orthogonality is used throughout.
+disappear if we observe a new mediating series. The notion is closely
+related to the information theoretic measure of transfer entropy,
+indeed, if the distribution of $v(t)$ is known to be Gaussian then
+they are equivalent \cite{barnett2009granger}.
+
+The notion of conditional orthogonality is used throughout, as well as
+some equivalent representations for Granger-causality.
 
@@
 
+  \todo{Do I even use (a) anymore?  I think I can just put the definition of conditional orthogonality into Theorem 1.}
+
   \begin{enumerate}
@@
 \end{lemma}
-% \begin{proof}
-%   Proceding from the definition of conditional orthogonality we have
-
-%   \begin{align*}
-%     &\inner{\alpha - \linE{\alpha}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
-%     &{\iff} \inner{\alpha + x - \linE{\alpha + x}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
-%     &\overset{(a)}{\iff} \inner{\alpha + x}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
-%     &\overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A} \vee \mathcal{X}} = 0\ \forall \beta \in \mathcal{B}\\
-%     &\overset{(c)}{\iff} \linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}},
-%   \end{align*}
-
-%   where $(a)$ follows since $(\beta - \linE{\beta}{\mathcal{X}}) \perp \mathcal{X}$ and $\linE{\alpha + x}{\mathcal{X}} \in \mathcal{X}$, $(b)$ since $z \perp \mathcal{X} \vee \mathcal{A} \iff \linE{z}{\mathcal{X} \vee \mathcal{A}} = 0$, and $(d)$ since $\linE{z}{\H_2} = \linE{\linE{z}{\H_2}}{\H_1}$ for $\H_2 \subseteq \H_1$.
-
-%   % This is property (vi) in Lindquist
-%   % \begin{align*}
-%   %   \langle \alpha - \linE{\alpha}{\mathcal{X}}, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
-%   %   \overset{(a)}{\iff} \langle \alpha, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0\ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
-%   %   \overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A}} &= 0\ \forall \beta \in \mathcal{B}
-%   % \end{align*}
-
-%   % where $(a)$ follows because $\linE{\alpha}{\mathcal{X}} \in \mathcal{X}$ and $\beta - \linE{\beta}{\mathcal{X}} \in \mathcal{X}$, and $(b)$ follows since $\beta \perp \mathcal{A} \iff \linE{\beta}{\mathcal{A}} = 0$.
-% \end{proof}
-\begin{lemma}
-  \label{lem:subspace_sum_projection}
-  Let $x \in \H$ where $\H$ is a separable Hilbert space having inner product
-  $\inner{\cdot}{\cdot}$ and let $\{\H_n\}_{n = 0}^N$ be a
-  collection of subspaces of $\H$ such that $x \perp \H_0$.  Then
-
-  \begin{equation*}
-    \linE{x}{\bigvee_{n = 0}^N\H_n} = \linE{x}{\bigvee_{n = 1}^N\H_n},
-  \end{equation*}
-
-  where
-  $\H_1 \vee \H_2 = \cl \{\alpha + \beta\ |\ \alpha \in \H_1, \beta
-  \in \H_2 \}$ is the closed sum of subspaces.
-\end{lemma}
-\begin{proof}
-  Apply the Gram-Schmidt process and directly calculate the projections.
-\end{proof}
 
@@
   \label{thm:granger_causality_equivalences}
-  Let $x(t)$ be a WSS process with absolutely summable covariance
-  sequence, and spectral density uniformly bounded above and below.
-  Denote by $\xi_{ij} \defeq \linEerr{x_i(t)}{\H^{-j}}$ and
-  $\xi_i \defeq \linEerr{x_i(t)}{\H_t}$ for distinct $i, j$.  Then,
-  the following are equivalent:
+  The following are equivalent:
 
@@
     \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$ i.e. $\B_{ij}(z) = 0$}
-    \item{$F_{ij} \defeq \big(\frac{\xi_i}{\xi_{ij}} - 1\big) = 0$}
-    \item{$\H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j} \iff H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
+    \item{$H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
     \item{$\linE{x_i(t)}{\H_{t - 1}^{-j}} = \linE{x_i(t)}{\H_{t - 1}}$}
-    % \item{\hl{Wold $A(\tau)$ condition}}
-    % \item{Tie in the 0s of $S(\omega)^{-1}$, though this isn't immediately directed.}
   \end{enumerate}
@@
 
-  $(1) \iff (3)$ is a tautology, and $(3) \iff (5)$ is immediate from the
-  definitions i.e. the projections are equal if and only if the errors
-  are equal.  We have $(4) \iff (5)$ as a result of Lemma
-  \ref{lem:conditional_orthogonality_equivalence}
-
-% , the additional
-%   equivalence in $(4)$ is apparent since $\H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$.
-
-  $(1) \Rightarrow (2)$: The projection for $x_i(t)$ onto $\H_{t - 1}$ is
-  (by definition) given by a form similar to equation
-  \ref{eqn:ar_representation}:
-
-\[
-  \E|x_i(t) - \linE{x_i(t)}{\H_{t - 1}}|^2 = \E|v_i(t) + \sum_{\tau = 1}^\infty \sum_{k = 1}^n (B_{i, k}(\tau) - \hat{B}_{i, k}(\tau))x_k(t - \tau)|^2.
-\]
-
-Since $v(t)$ in equation \ref{eqn:ar_representation} is temporally
-uncorrelated, it follows that the optimal projection is given by the
-model coefficients themselves.  This holds similarly for the
-projection onto $\H_{t - 1}^{-j}$.  Then by $(1)$ we have
-
-\[
-  \E|x_i(t) - \sum_{\tau = 1}^\infty \sum_{k = 1}^n B_{i, k}(\tau)x_k(t - \tau)|^2 = \E|x_i(t) - \sum_{\tau = 1}^\infty\sum_{k \ne j}B_{i, k}(\tau)x_k(t - \tau)|^2
-\]
-
-By the uniqueness of the projection we must have
-$\forall \tau\ B_{i, j}(\tau) = 0$.
-
-$(2) \Rightarrow (4)$: In computing $(y - \linE{y}{\H_{t - 1}^{-j}})$ for
-$y \in H_t^i$ it is sufficient to consider $y = x_i(t)$ by linearity, then since
+  $(1) \Rightarrow (2)$ follows as a result of the uniqueness of orthogonal
+  projection (i.e. the best estimate is necessarily the coefficients
+  of the model).  $(2) \Rightarrow (3)$ follows since in computing
+  $(y - \linE{y}{\H_{t - 1}^{-j}})$ for $y \in H_t^i$ it is sufficient
+  to consider $y = x_i(t)$ by linearity, then since
   $H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$ we have
-$(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$ since $\B_{ij}(z) = 0$.  $(4)$ then follows
-since $v_i(t) \perp \H_{t - 1}$ and
-$\forall z \in \H_{t - 1}^j\ (z - \linE{z}{\H_{t - 1}^{-j}}) \in \H_{t
-  - 1}$
+  $(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$ since
+  $\B_{ij}(z) = 0$ and $v_i(t) \perp \H_{t - 1}$.  $(3) \iff (4)$ is a result
+  of Lemma \ref{lem:conditional_orthogonality_equivalence}.  And,
+  $(4) \implies (1)$ follows directly from the Definition.
 \end{proof}
 
-% The following propositions justify various modifications of $x(t)$
-% applied in practice to massage $x(t)$ into a form amenable to more
-% standard tools.
-
-% \begin{theorem}[General Invariance]
-%   \hl{Under what conditions is this actually true?}
-
-%   Let $\zeta(t)$ be a stationary discrete time stochastic process.  Let
-%   $F, G$ be (possibly nonlinear and time varying) invertible filtering
-%   operations and $f(t)$, $g(t)$ be perfectly predictable and such that
-%   $x_j(t) \defeq F(\zeta_j - f)(t), x_i(t) \defeq G(\zeta_i - g)(t)$ are W.S.S.  Then,
-
-% \begin{equation}
-%     x_j \gc x_i \iff \zeta_j \te \zeta_i
-%   \end{equation}
-% \end{theorem}
-
-% \begin{proposition}[Invariance Under Invertible and Deterministic Modifications]
-%   Let $x(t)$ be a WSS process with absolutely summable covariance
-%   sequence, and spectral density uniformly bounded above and below.  Let
-%   $F(z), G$ represent univariate and invertible
-%   linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
-%   predictable processes.  Then,
-
-%   \begin{equation}
-%     x_j \gc x_i \iff F(x_j - f)(t) \gc G(x_i - g)(t)
-%   \end{equation}
-
-% \begin{proof}
-%   \hl{TODO}
-% \end{proof}
-% \end{proposition}
-
 \subsection{Granger Causality Graphs}
-We first need to establish some graph theoretic notation and
-terminology, collected formally in definitions for the reader's
-convenient reference.
+We establish some graph theoretic notation and terminology, collected
+formally in definitions for the reader's convenient reference.
 
@@
   Recall that we do not allow a node to be it's own parent, however
-  unless $\gcg$ is a DAG, a node can be it's own ancestor.  We will
+  unless $\gcg$ is a DAG a node \textit{can} be it's own ancestor.  We will
   ocassionally need to explicitly exclude $i$ from $\anc{i}$, in which
@@
 
-% What if we take this as the definition of PWGC?  then, in a simply causal graph
-% the "normal" definition becomes a theorem.  Or, if lem:pwgc_anc is true then
-% maybe that is good enough?
-
 From the $\VAR$ representation of $x(t)$ there is clearly a tight
@@
 
-% % When is this guaranteed w.r.t. the full system?
-% Furthermore, if the filter $\B_{ii}$ is invertible, we can write
-
-% \begin{equation*}
-%   \begin{aligned}
-%     (1 - \B_{ii}(z))x_i(t) &= v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\\
-%     \Rightarrow x_i(t) &= (1 - \B_{ii}(z))^{-1}\big[v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\big]\\
-%     &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)
-%   \end{aligned}
-% \end{equation*}
-
-% In general, we will write $\wtB$ for arbitrary filters maintaining the
-% convention that $\wtB_{ii}$ is a filter that modifies the future of
-% $x_i$ from it's past and $\wtB_{ij}$ a filter modifying $x_i$ from the
-% past of $x_j$.  We will use the convention that the filter
-% $\widetilde{\B}$ will represent simply ``a filter'' and is not
-% necessarily the same as the filters serving as edge weights in $\gcg$
-% (indeed, $\wtB$ will often be $0$), and moreover, we will allow for
-% $\widetilde{\B}$ to potentially change from place to place and even
-% from line to line in the same sequence of calculations as it's exact
-% properties or specification are not important.
-
-% We continue to develop the relationship between $x_i$ and it's family
-% tree, where the following proposition is analagous to the definitions
-% of feedback free processes of \cite{caines1975feedback}.
-
 \begin{proposition}[Ancestor Expansion]
@@
 $\A_{ij}(z) = 0$.
-
-% TODO: ???
-% , however the implication $j \in \anc{i} \implies \A_{ij}(z) \ne 0$ need not hold.
-
-% \begin{proof}
-%   The formula (\ref{eqn:parent_expansion}) was established in the
-%   preceding paragraph and serves as a starting point for the induction
-%   hypothesis:
-
-%   \begin{equation}
-%     \label{eqn:induction}
-%     x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
-%   \end{equation}
-
-%   The base case follows by expanding each $x_k$ in equation \ref{eqn:induction} via equation \ref{eqn:parent_expansion}, but we skip the calculations since they are nearly identical to the following induction step:
-
-%   \begin{align*}
-%     x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)\\
-%            &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)\big[\wtB_{kk}v_k(t)\\
-%     &\ \ \ \  + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t)\big]\\
-%     &\overset{(b)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell + 1}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
-%   \end{align*}
-
-%   Where equality $(a)$ follows by expanding each $x_k$ with equation \ref{eqn:parent_expansion} and $(b)$ requires the inversion of $\big(1 - \sum_{k \in \gpn{\ell}{i}\setminus\{i\}}\wtB_{ik}(z)\wtB_{ki}(z)\big)$ as well as using the fact that $\ancn{\ell}{i} = \ancn{\ell - 1}{i}\cup\gpn{\ell}{i}$.
-We here remind the reader of our earlier established convention that $\wtB$ simply represents the existence of a filter (possibly 0) and can change from place to place.
-
-% The induction terminates after $n$ steps since in a graph with $n$ nodes $\gpn{n + 1}{i} = \emptyset$
-
-% % Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.
-
-% %   \begin{align*}
-% %     x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)\\
-% %            &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)\big[\wtB_{kk}v_{k}(t) + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t) \big]\\
-% %            &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)v_{k}(t) + \sum_{k \in \gpn{2}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)
-% %   \end{align*}
-
-% % Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.
-
-% \end{proof}
-% \textbf{this is an alternative proof, I think superior to the one above}.
 
@@
   \begin{align*}
-    x_i(t) &= [(I - \B(z))^{-1}v(t)]_i\\
-    &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
+    x_i(t) &= [(I - \B(z))^{-1}v(t)]_i = \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
     &= \A_{ii}(z) v_i(t) + \sum_{\substack{j \in \anc{i} \\ j \ne i}} \A_{ij}(z) v_j(t)
@@
 capture the idea of ``flow of information'' in the underlying graph,
-in the sense that if $j \in \anc{i}$ we expect that $j \pwgc i$.  It
-may also be useful for reasoning about the conditions under which
+in the sense that if $j \in \anc{i}$ we expect that $j \pwgc i$.  It may
+also be useful for reasoning about the conditions under which
 \textit{unobserved} components of $x(t)$ may or may not interfere with
 inference in the actually observed components.  Finally, motivated
-from a practical standpoint to analyze causation in large systems, we
-will seek to construct practical estimation procedures based purely on
-pairwise causality tests since the computation of such pairwise
-relations is somewhat easier.
-
-The following is a fundamental building block:
+from a practical standpoint to analyze causation in large systems,
+practical estimation procedures based purely on pairwise causality
+tests are of interest since the computation of such pairwise relations
+is substantially easier..
+
+\begin{lemma}
+  \label{lem:separated_ancestor_uncorrelated}
+  Consider distinct nodes $i, j$ in a Granger-causality graph
+  $\gcg$.  If
+
+  \begin{enumerate}[label=(\alph*)]
+    \item{$j \not\in \anc{i}$ and $i \not\in \anc{j}$}
+    \item{$\anc{i}\cap\anc{j} = \emptyset$}
+  \end{enumerate}
+
+  then $\H_t^{(i)} \perp \H_t^{(j)}$, that is,
+  $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.  Moreover,
+  this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_{t - 1}^i} = 0$.
+\end{lemma}
+
+\begin{remark}
+  The possibility that there exist nodes sharing both $i$ and $j$ as
+  ancestors is not excluded, the point being that the temporal nature
+  of Granger-causality eliminates the problems caused by ``colliders''
+  (in the language of Pearl's causal calculus
+  \cite{pearl2009causality}) in static causal inference.
+\end{remark}
+
+\begin{proof}
+  We show directly that
+  $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.  To this end,
+  fix $s, \tau \ge 0$, then by expanding with equation \eqref{eqn:ancestor_expansion} we have
+
+  \begin{align*}
+    \E x_i(t - s)x_j(t - \tau)
+    &= \E \big(\A_{ii}(z)v_i(t - s)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)\\
+    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\E[\big(\A_{ik}(z)v_k(t - s)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)]\\
+    &+ \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[\big(\A_{ii}(z)v_i(t - s)\big) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
+    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ik}(z)v_k(t - s)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)].
+  \end{align*}
+
+  Keeping in mind that $v(t)$ is an isotropic and uncorrelated
+  sequence we see that each of these above four terms are 0: the
+  first term since $i \ne j$, the second and third since
+  $j \not\in \anc{i}$ and $i \not\in \anc{j}$ and finally the fourth since
+  $\anc{i} \cap \anc{j} = \emptyset$.
+\end{proof}
+
+It is possible for components of $x(t)$ to be correlated at some time
+lags without resulting in pairwise causality:
 
@@
   \label{lem:ancestor_uncorrelated}
-  Consider distinct nodes $i, j$ in a Granger-causality graph
-  $\gcg$.  Then, if
+  Consider distinct nodes $i, j$ in a Granger-causality graph $\gcg$.
+  If
 
@@
 
-  then $H_t^{(i)} \perp \H_t^{(j)}$, that is,
-  $\forall \tau \in \Z_+\ \E[x_i(t)x_j(t - \tau)] = 0$.  Moreover,
-  this implies that $j \npwgc i$.
+  then $j \npwgc i$.
 \end{lemma}
-
-\begin{remark}
-  Note that $H_t^{(i)} \perp \H_t^{(j)}$ states that the \textit{present}
-  of $x_i(t)$ is uncorrelated with the whole history of $x_j(t)$, but
-  we need not have $\H_t^{(i)} \perp \H_t^{(j)}$ since $i \in \anc{j}$ is
-  not excluded.
-\end{remark}
-
 \begin{proof}
-  We show directly that
-  $\forall \tau \in \Z_+\ \E[x_i(t)x_j(t - \tau)] = 0$.  To this end,
-  fix $\tau \ge 0$, then by expanding with equation \eqref{eqn:parent_expansion} we have
+  By Theorem \ref{thm:granger_causality_equivalences} it suffices to show that
+
+  \begin{equation*}
+    \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^i}} = 0,
+  \end{equation*}
+
+  which by the orthogonality principle is equivalent to
+
+  \begin{equation}
+    \label{eqn:proof_inner0}
+    \inner{x_i(t)}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^i}} = 0.
+  \end{equation}
+
+  Define the disjoint sets
 
   \begin{align*}
-    \E x_i(t)x_j(t - \tau) &= \E \big(\A_{ii}(z)v_i(t)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)\\
-    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\E[\big(\A_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
-    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ik}(z)v_k(t)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)].
+    C_0(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, k \ne i \}\\
+    C_1(u) &= \{k \in \pa{u}\ |\ i \in \anc{k} \text{ or } k = i \}.
   \end{align*}
 
-% (i.e.
-%   $\forall \tau \in \Z,\ k \ne \ell \Rightarrow \E v_k(t)v_\ell(t -
-%   \tau) = 0$ and
-%   $\forall \tau \in \Z\setminus\{0\}\ \E v_i(t)v_i(t - \tau) = 0$),
-
-  Keeping in mind that $v(t)$ is an isotropic and uncorrelated
-  sequence we will see that each of these above four terms are 0.  The
-  first term since $i \ne j$.  The second since $j \not\in \anc{i}$.
-  The fourth since $\anc{i} \cap \anc{j} = \emptyset$.  For the third
-  term, it is possible that $i \in \anc{j}$, which requires more care.
-
-  Similarly as before we can see immediately that
-
-  \begin{equation*}
-    \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)] = \E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)].
-  \end{equation*}
-
-  However, $k \ne \ell \Rightarrow A_{k\ell}(0) = 0$ (i.e.
-  $\A(z) = I + A(1)z^{-1} + \cdots$) so $\A_{ji}(z)v_i(t - \tau)$ is
-  strictly earlier in time (for $\tau \ge 0$) than $v_i(t)$ and hence
-  $\E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)] = 0$.
+  We can then expand the parents of $j$ using Equation \ref{eqn:parent_expansion} as
+
+  \begin{equation*}
+    x_j(t) = v_j(t) + \sum_{k \in C_0(j)} B_{jk}(z) x_k(t) + \sum_{k \in C_1(j)} B_{jk}(z) x_k(t),
+  \end{equation*}
+
+  which when substituted into the left hand side of Equation
+  \ref{eqn:proof_inner0} results (by Lemma
+  \ref{lem:separated_ancestor_uncorrelated}) in
+
+  \begin{equation*}
+    \inner{x_i(t)}{\sum_{k \in C_1(j)} B_{jk}(z) x_k(t) - \sum_{k \in C_1(j)} \linE{B_{jk}(z) x_k(t)}{\H_{t - 1}^i}}.
+  \end{equation*}
+
+  We can continue this process recursively (i.e. split each
+  $k \in C_1(j)$ into $C_0(k)$ and $C_1(k)$) which must eventually
+  terminate with $C_1(u) = \{i\}$.  Therefore, there exists some
+  causal filter $\Phi(z)$ such that Equation \ref{eqn:proof_inner0} is
+  equivalent to
+
+  \begin{equation}
+    \label{eqn:proof_inner0}
+    \inner{x_i(t)}{\Phi(z)x_i(t) - \linE{\Phi(z)x_i(t)}{\H_{t - 1}^i}},
+  \end{equation}
+
+  which is $0$ since $\linE{\Phi(z)x_i(t)}{\H_{t - 1}^i} = \Phi(z)x_i(t)$.
 \end{proof}
 
-The previous proposition can be strengthened significantly; notice
+The previous Lemma can be strengthened significantly; notice
 that it is possible to have some $k \in \anc{i} \cap \anc{j}$ where still
@@
 \label{sec:strongly_causal_graphs}
-% The examples at the conclusion of the previous section seem rather
-% pathological since we have simply constructed a case where different
-% paths in the graph cancel exactly.  It may be possible to rid
-% ourselves of these pathologies for instance by defining a notion of
-% ``robust'' pairwise causality where ``$j \pwgc _R i$'' if $j \pwgc i$
-% almost surely for a ``small'' random perturbation of the system
-% matrix.  However, we do not believe such a modification of our theory
-% can have useful practical implications since the difference between
-% paths that ``nearly cancel'' rather than cancel exactly would in
-% practice require inordinate amounts of data to resolve.
 
@@
 
-% \begin{lemma}
-%   \label{lem:still_strongly_causal}
-%   If $\gcg$ is a strongly causal graph, any subgraph formed by eliminating nodes
-%   as well as all of the in and out edges thereof is still strongly causal.
-% \end{lemma}
-% \begin{proof}
-%   If $\gcg$ has at most one path between any two nodes, there can only
-%   be fewer paths after removing nodes from $\gcg$.
-% \end{proof}
-
 For later use, and to get a feel for the topological implications of
@@
 \end{proof}
-
-% Is pwgc transitive?
 
@@
 
-% \begin{proposition}
-%   \label{prop:independent_confounders}
-%   If $\gcg$ is strongly causal we have distinct
-%   $k_1, k_2 \in \anc{i} \cap \anc{j}$ are confounders of $i, j$, then
-%   $k_1 \not\in \anc{k_2}$, $k_2 \not\in \anc{k_1}$ and
-%   $\not\exists \ell \in \anc{k_1} \cap \anc{k_2}$ confounding
-%   $k_1, k_2$.
-% \end{proposition}
-% \begin{proof}
-%   \hl{TODO}
-% \end{proof}
-
-% \begin{corollary}
-%   If $\gcg$ is strongly causal, then distinct confounders are
-%   uncorrelated.
-% \end{corollary}
-% \begin{proof}
-%   This follows directly by combining proposition
-%   \ref{prop:independent_confounders} with proposition \ref{prop:ancestor_properties}.  \hl{I need to update prop} \ref{prop:ancestor_properties} \hl{to note the strong non-correlation property.}
-% \end{proof}
-
-% \begin{lemma}
-%   If $\gcg$ is a strongly causal DAG and $j \in \anc{i}$ then 
-% \end{lemma}
-
 In light of proposition \ref{prop:sc_graph_common_anc}, the following
@@
 \end{corollary}
-
-% \begin{remark}
-%   It can be seen from the conclusion of the proof that simple facts
-%   about Granger causality are indeed reliant on our assumptions about
-%   the nature of $x(t)$ laid out in the section \ref{sec:theory}, in
-%   particular, the innovations process $v(t)$ must be full rank.
-
-%   Moreover, that strong conditions need to be placed on the topology
-%   of $\gcg$ in order for proposition \ref{prop:pwgc_anc} to hold can
-%   be seen through the earlier examples.
-% \end{remark}
-
-% This holds in any DAG
-% \begin{lemma}
-%   If $\gcg$ is strongly causal, then any $\\VAR(p)$ system on $\gcg$ is stable if and only if
-%   $B_{ii}(z)$ is stable for every $i = 1, \ldots, n$.
-% \end{lemma}
-% \begin{proof}
-%   \hl{This is certainly true, but I don't think I need to use it anywhere.}
-% \end{proof}
 
@@
 theoretical analysis in this paper.
-
-% This is an interesting fact in and
-% of itself, but also has some implications for applications since
-% pairwise testing is trivially parallelizable.  In section
-% \ref{sec:structure_learning} we will also analyze the use of pairwise
-% testing as a heuristic for general graphs.
 
@@
 
-% \begin{lemma}
-%   \label{lem:inner_loop_lemma}
-%   At any given iteration, $k$, of Algorithm \ref{alg:pwgr}, if we have
-%   $e \in E$ if and only if $e \in \gcge$, then the inner loop (line \ref{alg:inner_loop})
-%   will add $(i, j) \in \big(\bigcup_{r = 0}^kP_{k - r}\big) \times C_k$ to $E$
-%   if $(i, j) \in \gcge$.  i.e. any real edge will be included.
-% \end{lemma}
-% \begin{proof}
-%   Let $j \in C_k$, $i \in \bigcup_{r = 0}^k P_{k - r}$, and assume
-%   $i \in \pa{j}$.
-
-%   First, suppose that there is no $i \rightarrow j$ path in $E$.  Then
-%   since $i \gc j$ we have $i \pwgc j$ and $j \npwgc i$ (i.e.
-%   $(i, j) \in W$) by Proposition \ref{prop:pwgc_anc} and Corollary
-%   \ref{cor:bidirectional_edge} so at least one of
-%   $\{D_{kr} \}_{r = 0}^k$ contains $(i, j)$ and hence $(i, j)$ will be
-%   added to $E$ upon the end of the iteration.
-
-%   Now, suppose instead that there is already an $i \rightarrow j$ path
-%   in $E$.  Given our assumption that $E$ contains edges if and only if
-%   they are edges of $\gcg$, then either we already have $(i, j) \in E$
-%   correctly, or there is more than one $i \rightarrow j$ path in
-%   $\gcg$, which is not possible since $\gcg$ is strongly causal.
-% \end{proof}
-
-% \begin{lemma}
-%   \label{lem:W_subset_E}
-%   $W \subseteq \gcge$
-% \end{lemma}
-% \begin{proof}
-%   By proposition \ref{prop:pwgc_anc} we have
-%   $j \in \pa{i} \Rightarrow j \pwgc i$ and by Corollary \ref{cor:bidirectional_edge}
-%   there are no bidirectional edges in $\gcg$.
-% \end{proof}
-
-% \hl{Must show that Algorithm terminates in $n$ steps.}
-% \begin{lemma}
-%   For any $i \in P_k$ we have $i \in C_{k - 1}$.  \hl{[Can't yet prove
-%     this part] Moreover, for each $i \in \bigcup_{k = 0}^n P_k$, if
-%     $(s, i) \in W$ then $(s, i) \in E$ upon the termination of
-%     Algorithm} \ref{alg:pwgr}.
-% \end{lemma}
-% \begin{proof}
-%   We prove the second part by induction, the first part being
-%   established along the way.  By construction $P_0$ has no parent
-%   nodes in $\gcg$ so the base case is immediate.  Fix now some
-%   $k \ge 1$, $i \in P_k$, and let $s$ be such that $(s, i) \in W$.
-%   For every $s$ such that $(s, i) \in W$ we have $s \not\in S_k$ since
-%   otherwise $i \not\in P_k$.  Therefore
-%   $s \in \bigcup_{r = 0}^{k - 1}P_{k - r}$, which implies
-%   $s \not\in S_k$, and finally that $i \in C_{k - 1}$.  Then by Lemma
-%   \ref{lem:W_subset_E} every edge in $\gcge$ is also present in
-%   $W$, and by Lemma \ref{lem:inner_loop_lemma} \hl{[We can't conclude
-%     here because of the strong conditions required on lemma}
-%     \ref{lem:inner_loop_lemma}]
-% \end{proof}
-
-% \begin{lemma}[Loop Invariant]
-%   \label{lem:loop_invariant}
-%   $\forall k \ge 0,\ e \in E_k \Rightarrow e \in \gcge$.
-%   \hl{Not sure if this is useful / the right approach.}
-% \end{lemma}
-% \begin{proof}
-%   We will proceed by induction.  Firstly, the base case
-%   $E_0 = \emptyset$ is trivial, so fix some $k \ge 1$ and consider an
-%   edge $(i, j) \in \big(\bigcup_{r = 0}^k P_{k - r} \big) \times C_k$
-%   such that $(i, j) \in W$ (if the edge is not in $W$ then it is not
-%   in $\gcge$ by Lemma \ref{lem:W_subset_E}).
-
-%   \hl{WIP}
-% \end{proof}
-
 Our proof proceeds in 5 steps stated formally as lemmas.  Firstly, we
@@
 
-% \begin{proof}
-%   Denote the edges of $\gcg$ by $\gcge$, we endevour to show
-%   that upon the termination of Algorithm \ref{alg:pwgr} we have
-%   $E = \gcge$.  On the first iteration of the algorithm
-%   ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
-
-%   Moving on from any trivial cases, notice that
-%   $\gcge \subseteq W$.  This follows since by Proposition
-%   \ref{prop:pwgc_anc} we have $j \in \pa{i} \Rightarrow j \pwgc i$ and by
-%   Corollary \ref{cor:bidirectional_edge} there are no bidirectional
-%   edges in $\gcg$.
-
-% 1. Dispense with trivialities
-% 2. Show that hat{P}_k is nonempty
-% 3. Show that C_k is nonempty
-% 4. Show that edges in D_k are actual edges in G
-% 5. show that each edge incident on C_k is in D_k
-% 6. Induction
-
-% Given non-empty P and C...  Can we show all P -> C edges get collected in D?
-  % Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$ and let us define
-
-  % \begin{equation*}
-  %   A_k = \{i \in [n]\ |\ \dist{i}{P_0} = k\},
-  % \end{equation*}
-
-  % where
-
-  % \begin{equation*}
-  %   \dist{i}{P} = \text{max}\{r \in [n]\ |\ \exists (j \rightarrow i) \text{ path of length } r \text{ in } \gcg \text{ for some } j \in P\}.
-  % \end{equation*}
-
-  % We will proceed by induction and show that when the counter has reached the value $k \ge 1$, it is guaranteed that all nodes in $\bigcup_{\ell \le k}A_\ell$ (nodes having paths of length up to $k$ between themselves and $P_0$) must have all of their incident edges included in $E$, and that every edge in $E$ is an edge of $\gcg$.
-If this is the case, then it can be immediately seen that the algorithm terminates with the correct result, $\gcg = ([n], E)$, once $k = n$; since there can only be paths up to length $n - 1$ in $\gcg$, and there cannot be any nodes or edges that don't involve paths back to $P_0$.
-That is, $\bigcup_{\ell \le n}A_\ell = [n]$.
-
-  % % firstly dispensing with some trivialities, and then establishing that every edge in $\gcg$ makes an appearance in some $D_{kr}$ as well as that everything in any $D_{kr}$ is indeed an edge in $\gcg$.
-
-  % On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
-To quickly dispense with some trivialities: the set $P_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $P_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
-Now, if $P_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
-
-  % Continuing with the non-trivial case: if $P_0 \ne [n]$ then $C_0$ must be non-empty, otherwise there would be driving nodes not included in $P_0$.
-In this case it is also clear that $A_1 \ne \emptyset$, so let us consider some $j \in A_1$.
-We must have $j \in C_0$, otherwise there would be a path of length $2$ from $P_0$ to $j$.
-Moreover, since $\exists i \in P_0$ s.t.
-$i \gc j$ and therefore $i \pwgc j$ (by proposition \ref{prop:pwgc_anc}) we have $(i, j) \in D_{00}$ and therefore $(i, j) \in E$.
-Therefore $A_1 \subseteq D_{00}$.
-\hl{This is nonsensical because $A_1$ does not contain tuples -- need to show that $A_1 \subseteq C_0$ ?}
-
-  % Now consdider some $(i, j) \in D_{00}$ (the same reasoning as for $C_0$ is sufficient to see that $D_{00}$ is non-empty).
-Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).
-Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in P_0 \Rightarrow u \in S_0 \Rightarrow j \not\in C_0$, a contradiction.
-This implies every tuple in $D_{00}$ is a bona-fide edge of $\gcg$, as well as that $(i, j) \in A_1$.
-So then $A_1 = D_{00}$, [\hl{nonsensical, see earlier.}] completing the base case for induction.
-
-  % Suppose now the algorithm has reached step $k$ and that $E = \bigcup_{\ell \le k - 1}A_\ell$.
-
-% The reasoning here is (I think) correct -- but isn't directly establishing the induction hypothesis.
-%   On the other hand, if $P_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $P_0$.
-Consider now an edge $(i, j) \in D_{00}$ ($D_{00}$ is non-empty for the same reason as $C_0$).
-Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).
-Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in P_0 \Rightarrow u \in S_0 \Rightarrow j \not\in C_0$, a contradiction.
-So every tuple in $D_{00}$ is a bona-fide edge of $\gcg$.
-Suppose now that for some $j \in C_0$ there is an $i$ s.t.
-$(i, j) \in \gcg$ but $(i, j) \not\in D_{00}$.
-Since $(i, j) \in \gcg$ we have $i \pwgc j$ (corollary \ref{cor:gc_implies_pwgc}), but then if $i \not\in P_0$ then $j \notin C_0$ and if $i \in P_0, j \in C_0$ then necessarily $(i, j) \in D_{00}$ since we cannot at this stage have any paths in $E$.
-
-%   and show that on each step, every edge incident upon a node in $C_k$ is contained in $D_k$, that every edge in $D_k$ is an edge in $\gcg$, and finally that $\bigcup_{k = 1}^n C_k = [n] \setminus P_0$ and that this implies $\big([n], \bigcup_{k = 1}^n D_k\big) = \gcg$.
-
-%  On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E_0 = \emptyset$.
-The set $P_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $P_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
-Now, if $P_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
-On the other hand, if $P_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $P_0$.
-Consider now an edge $(i, j) \in D_0$ (again $D_0$ is non-empty for the same reason as $C_0$).
-Since $i$ has no parents, we must have $\anc{i}\cap\anc{j} = \emptyset$ and therefore $i \in \anc{j}$, and due to the construction of $C_0$ we in fact have $i \in \pa{j}$, which establishes that every edge in $D_0$ is also an edge in $\gcg$.
-Finally consider $j \in C_0$, by proposition \ref{prop:pwgc_anc} and since the parents of $C_0$ must be in $P_0$, we see that every edge incident on $j$ must be present in $D_0$.
-
-% We also see that since nodes in $P_0$ have no parents,
-
-% we must have $\anc{i}\cap\anc{j} = \emptyset$, and therefore by proposition \ref{prop:ancestor_properties} $i \in \anc{j}$;  moreover, the construction of $P_1$ implies that in fact $i \in \pa{j}$, so $(i, j)$ is an edge of $\gcg$.
-
-%Finally, by similar reasoning, there cannot be an edge $(i, j) \in D_0$ which is not in $\gcg$, since $\anc{i} \cap \anc{j} = \emptyset$.
-
-%Consider now step $K < n$ of algorithm \ref{alg:pwgr}.
-By lemma \ref{lem:still_strongly_causal} the graph remains strongly causal after having removed nodes $\bigcup_{k = 0}^K P_k$ from $S$, and we assume for induction that $\bigcup_{k = 1}^K D_k$ are all edges of $\gcg$.
-
-% \hl{WIP}
-% \end{proof}
-
 \begin{example}
@@
 
-\section{Finite Sample Graph Recovery}
-\label{sec:structure_learning}
-In this section we provide a review of our methods for implementing
-Algorithm 1 given a \textit{finite} sample of $T$ data points.  We
-apply the simplest reasonable methods in order to maintain a focus on
-our main contributions (i.e. Algorithm \ref{alg:pwgr}), more
-sophisticated schemes can only serve to improve the results.  Textbook
-reviews of the following concepts are provided e.g. by
-\cite{all_of_statistics}, \cite{murphy_mlp}, and elsewhere.
-
-In subsection \ref{sec:pairwise_hypothesis_testing} we define pairwise
-Granger-causality hypothesis tests, in subsection
-\ref{sec:model_order_selection} a model order selection criteria, in
-subsection \ref{sec:efficient_model_estimation} an efficient
-estimation algorithm, in subsection \ref{sec:error_rate_control} the
-method for choosing an hypothesis testing threshold, and finally in
-subsection \ref{sec:finite_pwgc} the unified finite sample algorithm.
-
-\subsection{Pairwise Hypothesis Testing}
-\label{sec:pairwise_hypothesis_testing}
-In performing pairwise checks for Granger-causality $x_j \pwgc x_i$ we
-follow the simple scheme of estimating the following two linear models:
-
-\begin{align}
-  H_0:&\ \widehat{x}_i^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau),\\
-  H_1:&\ \widehat{x}_i^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau) + \sum_{\tau = 1}^pb_{ij}(\tau)x_j(t - \tau).
-\end{align}
-
-We formulate the statistic
-
-\begin{equation}
-  \label{eqn:gc_statistics}
-  F_{ij}(p) = \frac{T}{p}\Big(\frac{\xi_i(p)}{\xi_{ij}(p)} - 1\Big),
-\end{equation}
-
-where $\xi_i(p)$ is the sample mean square of the
-residuals\footnote{This quantity is often denoted $\widehat{\sigma}$,
-  but we maintain notation from Definition
-  \ref{def:granger_causality}.}  $x_i(t) - \widehat{x}^{(p)}_i(t)$,
-
-\begin{equation*}
-  \xi_i(p) = \frac{1}{T - p}\sum_{t = p + 1}^T (x_i(t) - \widehat{x}_i^{(p)}(t))^2,
-\end{equation*}
-
-and similarly for $\xi_{ij}(p)$.  We test $F_{ij}(p)$ against a
-$\chi^2(p)$ distribution.
-
-If the estimation procedure is consistent, we will have the following
-convergence (in $\P$ or a.s.):
-
-\begin{equation}
-  F_{ij}(p) \rightarrow
-  \left\{
-    \begin{array}{ll}
-      0;\ x_j \npwgc x_i\\
-      \infty;\ x_j \pwgc x_i
-    \end{array}
-  \right. \text{ as } T \rightarrow \infty.  % the '.' after \right is necessary.
-\end{equation}
-
-In our finite sample implementation (see Algorithm
-\ref{alg:finite_pwgc}) we add edges to $\widehat{\gcg}$ in order of
-the decreasing magnitude of $F_{ij}$ instead of proceeding backwards
-through $P_{k - r}$ in Algorithm \ref{alg:pwgr}.  This makes greater
-use of the information provided by the test statistic $F_{ij}$,
-moreover, if $x_i \gc x_j$ and $x_j \gc x_k$, it is expected that
-$F_{kj} > F_{ji}$, thereby providing the same effect as proceeding
-backwards through $P_{k - r}$.
-\subsection{Model Order Selection}
-\label{sec:model_order_selection}
-There are a variety of methods to choose the filter order $p$ (see
-e.g. \cite{lutkepohl2005new}), but we will focus in particular on the
-Bayesian Information Criteria (BIC).  The BIC is substantially more
-conservative than the popular alternative Akaiake Information Criteria
-(the BIC is also asymptotically consistent), and since we are
-searching for \textit{sparse graphs}, we therefore prefer the BIC,
-where we seek to \textit{minimize} over $p$:
-
-\begin{equation}
-  \label{eqn:bic}
-  \begin{aligned}
-    BIC_{\text{univariate}}(p) &= \ln\ \xi_i(p) + p\frac{\ln T}{T},\\
-    BIC_{\text{bivariate}}(p) &= \ln \det \widehat{\Sigma}_{ij}(p) + 4p\frac{\ln T}{T},\\
-  \end{aligned}
-\end{equation}
-
-where $\widehat{\Sigma}_{ij}(p)$ is the $2 \times 2$ residual
-covariance matrix for the $\VAR(p)$ model of $(x_i(t), x_j(t))$.  The
-bivariate errors $\xi_{ij}(p)$ and $\xi_{ji}(p)$ are the diagonal
-entries of $\widehat{\Sigma}_{ij}(p)$.
-
-We carry this out by a simple direct search on each model order
-between $0$ and some prescribed $p_\text{max}$, resulting in a
-collection $p_{ij}$ of model order estimates.  In practice, it is
-sufficient to pick $p_\text{max}$ ad-hoc or via some simple heuristic
-e.g. plotting the sequence $BIC(p)$ over $p$, though it is not
-technically possible to guarantee that the optimal $p$ is less than
-the chosen $p_\text{max}$ (since there can in general be arbitrarily
-long lags from one variable to another).
-
-\subsection{Efficient Model Estimation}
-\label{sec:efficient_model_estimation}
-In practice, the vast majority of computational effort involved in
-implementing our estimation algorithm is spent calculating the error
-estimates $\xi_i(p_i)$ and $\xi_{ij}(p_{ij})$.  This requires fitting a
-total of $n^2p_{\text{max}}$ autoregressive models, where the most
-naive algorithm (e.g. solving a least squares problem for each model)
-for this task will consume $O(n^2p_{\text{max}}^4T)$ time, it is
-possible to carry out this task in a much more modest
-$O(n^2p_{\text{max}}^2 ) + O(n^2p_{\text{max}}T)$ time via the
-autocorrelation method
-\cite{hayes_statistical_digital_signal_processing} which substitutes
-the following autocovariance estimates in the Yule-Walker
-equations:\footnote{The particular indexing and normalization given in
-  equation \ref{eqn:covariance_estimate} is critical to ensure
-  $\widehat{R}$ is positive semidefinite.  The estimate can be viewed
-  as calculating the covariance sequence of a signal multiplied by a
-  rectangular window.}
-
-\begin{equation}
-  \label{eqn:covariance_estimate}
-  \widehat{R}_x(\tau) = \frac{1}{T}\sum_{t = \tau + 1}^T x(t) x(t - \tau)^\T;\ \tau = 0, \ldots, p_{\text{max}},
-\end{equation}
-
-It is imperative that the first index in the summation is $\tau + 1$, as
-opposed perhaps to $p_\text{max}$ and that the normalization is
-$1 / T$, as opposed perhaps to $1 / (T - p_\text{max})$, in order to
-guarantee that $\widehat{R}_x(\tau)$ forms a valid (i.e. positive
-definite) covariance sequence.  This results in some bias, however the
-dramatic computational speedup is worth it for our purposes.
-
-These covariance estimates constitute the $O(n^2p_{\text{max}}T)$
-operation.  Given these particular estimates, the variances $\xi_i(p)$
-for $p = 1, \ldots, p_{\text{max}}$ can be evaluated in
-$O(p_{\text{max}}^2)$ time each by applying the Levinson-Durbin
-recursion to $\widehat{R}_{ii}(\tau)$, which effectively estimates a
-sequence of $AR$ models, producing $\xi_i(p)$ as a side-effect (see
-\cite{hayes_statistical_digital_signal_processing} and
-\cite{levinson_durbin_recursion}).
-
-Similarly, the variance estimates $\widehat{\Sigma}_{ij}(p)$ (which
-include $\xi_{ij}$ and $\xi_{ji}$) can be obtained by estimating
-$\frac{(n + 1)n}{2}$ bivariate AR models, again in
-$O(p_{\text{max}}^2)$ time via Whittle's generalized Levinson-Durbin
-recursion\footnote{We have made use of standalone tailor made
-  implementations of these algorithms, available at
-  \textsf{github.com/RJTK/Levinson-Durbin-Recursion}.}
-\cite{whittle_generalized_levinson_durbin}.
-
-\subsection{Edge Probabilities and Error Rate Controls}
-\label{sec:error_rate_control}
-Denote $F_{ij}$ the Granger-causality statistic of equation
-\ref{eqn:gc_statistics} with model orders chosen by the methods of
-Section \ref{sec:model_order_selection}.  We assume that this
-statistic is asymptotically $\chi^2(p_{ij})$ distributed (the
-disturbances are Gaussian), and denote by $G$ the cumulative
-distribution function thereof.  We will define the matrix
-
-\begin{equation}
-  \label{eqn:edge_inclusion_probability}
-  P_{ij} = G(F_{ij}),
-\end{equation}
-
-to be the matrix of pairwise edge inclusion P-values.  This is
-motivated by the hypothesis test where the hypothesis $H_0$ will be
-rejected (and thence we will conclude that $x_j \pwgc x_i$) if
-$P_{ij} > 1 - \delta$.
-
-The value $\delta$ can be chosen by a variety of methods, in our case
-we apply the Benjamini Hochberg criteria \cite{benjamini_hochberg}
-\cite{all_of_statistics} to control the false discovery rate of
-pairwise edges to a level $\alpha$ (where we generally take
-$\alpha = 0.05$).
-
-\subsection{Finite Sample Recovery Algorithm}
-\label{sec:finite_pwgc}
-
-After the graph topology $\widehat{\gcg}$ has been estimated via
-Algorithm \ref{alg:finite_pwgc}, we refit the entire model with the
-specified sparsity pattern directly via ordinary least squares.
-
-We note that producing graph estimates which are not strongly causal
-can potentially be achieved by performing sequential estimates
-$\widehat{x}_1(t), \widehat{x}_2(t), \ldots$ estimating a strongly causal
-graph with the residuals of the previous model as input, and then
-refitting on the combined sparsity pattern.  We experiment with this
-heuristic in our example application of Section \ref{sec:application},
-but reserve theoretical analysis for future work.
-
-\begin{algorithm}
-    \SetKwInOut{Input}{input}
-    \SetKwInOut{Output}{output}
-    \SetKwInOut{Initialize}{initialize}
-    \DontPrintSemicolon
-
-    \BlankLine
-    \caption{Finite Sample Pairwise Graph Recovery (PWGC)}
-    \label{alg:finite_pwgc}
-
-    \Input{Estimates of pairwise Granger-causality statistics $F_{ij}$
-      (eqn \ref{eqn:gc_statistics}).
-Matrix of edge probabilities $P_{ij}$ (eqn \ref{eqn:edge_inclusion_probability}).
-Hypothesis testing threshold $\delta$ chosen via the Benjamini-Hochberg criterion (Section \ref{sec:error_rate_control})}
-    \Output{A strongly causal graph $\widehat{\gcg}$}
-    \Initialize{$S = [n]$  \texttt{\# unprocessed nodes}\\
-      $E = \emptyset$  \texttt{\# edges of }$\widehat{\gcg}$\\
-      $k = 1$ \texttt{\# a counter used only for notation}}
-
-    \BlankLine
-
-    $W_\delta \leftarrow \{(i, j)\ |\ P_{ji} > 1 - \delta, F_{ji} > F_{ij}\}$  \texttt{\# candidate edges}\\
-    $\mathcal{I}_0 \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# total node incident probability}\\
-    $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) < \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Nodes with fewest incident edges}\\
-    \If{$P_0 = \emptyset$}{
-      $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) \le \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Ensure non-empty}
-    }
-    \BlankLine
-
-    \While{$S \ne \emptyset$}{
-      $S \leftarrow S \setminus P_{k - 1}$ \texttt{\# remove processed nodes}\\
-      % $\mathcal{I}_k \leftarrow \big(\sum_{j \in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# intra-}$S$ \texttt{incident strength}\\
-      $\mathcal{I}_k \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$\\
-      $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) < \ceil{\text{min}(\mathcal{I}_k)}\}$\\
-      \If{$P_k = \emptyset$}{
-        $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) \le \ceil{\text{min}(\mathcal{I}_k)}\}$
-      }
-      \;
-      \texttt{\# add strongest edges, maintaining strong causality}\\
-      $U_k \leftarrow \bigcup_{r = 1}^k P_{k - r}$ \texttt{\# Include all forward edges}\\
-      \For{$(i, j) \in \mathsf{sort}\Big(\{(i, j) \in U_k \times P_k\ |\ (i, j) \in W_\delta\} \mathsf{\ by\ descending\ } F_{ji}\Big)$} {
-        \If{$\mathsf{is\_strongly\_causal}(E \cup \{(i, j)\})$} {
-          \texttt{\# }$\mathsf{is\_strongly\_causal}$ \texttt{can be implemented by keeping}\\
-          \texttt{\# track of ancestor / descendant relationships}\\
-          $E \leftarrow E \cup \{(i, j)\}$
-        }
-      }
-      $k \leftarrow k + 1$\\
-    }
-    \Return{$([n], E)$}
-\end{algorithm}
-
-% \section{Structure Learning}
-% \label{sec:structure_learning}
-% \subsection{Local Search Heuristics}
-
-% \subsection{Edge-Wise Grouped LASSO}
-% Minimize the sparsity promoting regularized least squares problem and choose the hyper-parameters against the Akaike information criteria.
-
-% \begin{equation}
-%   L(\lambda, p) \defeq \underset{B}{\text{minimize}}\ \frac{1}{T}\sum_{t = 1}^T||x(t) - \sum_{\tau = 1}^pB(\tau)x(t - \tau)||_2^2 + \lambda \sum_{i, j}\big[\alpha||B_{i, j}||_2 + (1 - \alpha)||B_{i, j}||_1\big]
-% \end{equation}
-
-% \begin{equation}
-%   \underset{\lambda, p}{\text{minimize}}\ L(\lambda, p) + \mathsf{AIC}(B^{(\lambda, p)})
-% \end{equation}
-
-% \subsection{Bayesian Posterior Thresholding}
-% Consider the Bayesian model
-
-% \begin{equation}
-%   \begin{aligned}
-%     (x(t)\ |\ B, \sigma_v^2, \{x(t - \tau)\}_{\tau = 1}^p) &\sim \mathcal{N}(\sum_{\tau = 1}^pB(\tau)x(t - \tau), \sigma_v^2)\\
-%     (B_{ij}(\tau)\ |\ G_{ij}(\tau)) &\sim G_{ij}(\tau)\mathcal{N}(0, \sigma_\beta^2) + (1 - G_{ij}(\tau))\delta_0(B_{ij}(\tau))\\
-%     (G_{ij}(\tau)) &\sim \mathsf{BER}(p_{ij}(\tau))\\
-%     \sigma_v^2 &\sim \Gamma(a_v, b_v)\\
-%     \sigma_\beta^2 &\sim \Gamma(a_\beta, b_\beta)
-%   \end{aligned}
-% \end{equation}
-
-% Similar models have been studied by various authors in the context of the linear regression model \hl{[(cite them)]} where it is referred to as \textit{stochastic variable selection}, or referred to as a model for Bayesian variable selection.
-
-% Since we have in mind applications where $n$ is large, sampling posterior probabilities can be prohibitively burdensome, so we can instead fit the mean-field variational approximation $p(B, G, \sigma\ |\ X) \approx q_Gq_Bq_\sigma$ and subsequently apply a thresholding operation to the posterior edge inclusion probabilities under $q_G$.
-
-% \subsection{Pairwise Minimum Error Spanning Trees}
-% Inspired by theorem \ref{scg_recovery}, we propose the following
-% structure learning heuristic.
-
-% %% This algorithm should perform coordinate descent or jointly refit all the trees.
-
-% \begin{enumerate}
-%   \item{Set $k = 0$ and $x^0(t) = x(t)$}
-%   \item{Compute all pairwise causality measures $\Xi \defeq \big[\ln \frac{\xi_i}{\xi_{ij}} \big]_{i, j}$}
-%   \item{Find the maximum spanning arborescence\footnote{An ``arborescence'' is a french word for a tree diagram, and refers to a \textit{directed} tree in graph theory} $\mathcal{T}_k$ of a graph having edge weights $\Xi_{ij}$.}
-%   \item{Fit a vector LTI filter $F_k$ having graph defined by $\mathcal{T}_k$ to $x^{k - 1}(t)$ and set $x^k(t) = x^{k - 1}(t) - \hat{x}^{k - 1}(t)$}
-%   \item{Repeat over $k$ until a stopping criteria is met}
-% \end{enumerate}
-
-% \subsection{Layered Strongly Causal Graphs}
-% It is clearn that Algorithm \ref{alg:pwgr} is a purely theoretical construct, and cannot be applied directly to real data.
-Instead, we convert the algorithm into an heuristic and observe it's performance experimentally.
-We first consider the heuristic version of Algorithm \ref{alg:pwgr} in Algorithm \ref{alg:pwgr_heuristic} and show that is a well defined procedure (note that $\ceil{x}$ denotes the integer ceiling function):
-
-% \begin{proposition}
-%   Algorithm \ref{alg:pwgr_heuristic} terminates in at most $n$ iterations and returns a strongly causal graph $\widehat{\gcg}$.
-% \end{proposition}
-% \begin{proof}
-%   \hl{TODO}
-% \end{proof}
-
-% A complete estimation procedure for a sparse filter $\widehat{\B}(z)$
-% used to estimate $x(t + 1)$ from $\H_{t - 1}$ is then given in Algorithm \ref{alg:Bz_estimate}.
-
-% \begin{proposition}
-%   The error sequence in Algorithm \ref{alg:Bz_estimate} is a zero mean
-%   WSS process $\E\epsilon^{(m)}(t) = 0$ with monotonically decreasing
-%   variance: $\E|\epsilon^{(m + 1)}(t)|^2 \le \E|\epsilon^{(m)}(t)|^2$.
-% \end{proposition}
-% \begin{proof}
-%   \hl{TODO -- this is probably not even true}
-% \end{proof}
-
-% \begin{algorithm}
-%   \SetKwInOut{Input}{input}
-%   \SetKwInOut{Output}{output}
-%   \SetKwInOut{Initialize}{initialize}
-%   \DontPrintSemicolon
-
-%   \BlankLine
-%   \caption{Graph Recovery Heuristic $\mathsf{pw\_scg}(F, P, \alpha)$}
-%   \label{alg:pwgr_heuristic}
-%   \Input{Estimates of pairwise Granger-causaltiy statistics $F_{ij}$ (eqn \ref{eqn:gc_statistics}), Matrix of edge probabilities $P_{ij}$ (eqn \ref{eqn:edge_inclusion_probability}), threshold $\delta$.}
-%   \Output{A strongly causal graph $\widehat{\gcg}$}
-%   \Initialize{$S = [n]$  \texttt{\# unprocessed nodes}\\
-%     $E = \emptyset$  \texttt{\# edges of }$\widehat{\gcg}$\\
-%     $k = 1$ \texttt{\# a counter used only for notation}}
-
-%   \BlankLine
-
-%   $W_\delta \leftarrow \{(i, j)\ |\ P_{ji} > 1 - \delta, F_{ji} > F_{ij}\}$  \texttt{\# candidate edges}\\
-%   $\mathcal{I}_0 \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# total node incident probability}\\
-%   $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) < \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Nodes with fewest incident edges}\\
-%   \If{$P_0 = \emptyset$}{
-%     $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) \le \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Ensure non-empty}
-%   }
-%   \BlankLine
-
-%   \While{$S \ne \emptyset$}{
-%     $S \leftarrow S \setminus P_{k - 1}$ \texttt{\# remove processed nodes}\\
-%     % $\mathcal{I}_k \leftarrow \big(\sum_{j \in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# intra-}$S$ \texttt{incident strength}\\
-%     $\mathcal{I}_k \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$\\
-%     $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) < \ceil{\text{min}(\mathcal{I}_k)}\}$\\
-%     \If{$P_k = \emptyset$}{
-%       $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) \le \ceil{\text{min}(\mathcal{I}_k)}\}$
-%     }
-%     \;
-%     \texttt{\# add strongest edges, maintaining strong causality}\\
-%     $U_k \leftarrow \bigcup_{r = 1}^k P_{k - r}$ \texttt{\# Include all forward edges}\\
-%     \For{$(i, j) \in \mathsf{sort}\Big(\{(i, j) \in U_k \times P_k\ |\ (i, j) \in W_\delta\} \mathsf{\ by\ descending\ } F_{ji}\Big)$} {
-%       \If{$\mathsf{is\_strongly\_causal}(E \cup \{(i, j)\})$} {
-%         $E \leftarrow E \cup \{(i, j)\}$
-%       }
-%     }
-%     $k \leftarrow k + 1$\\
-%   }
-%   \Return{$([n], E)$}
-% \end{algorithm}
-
-% \begin{algorithm}
-%   \SetKwInOut{Input}{input}
-%   \SetKwInOut{Output}{output}
-%   \SetKwInOut{Initialize}{initialize}
-%   \DontPrintSemicolon
-
-%   \BlankLine
-%   \caption{Filter Estimator}
-%   \label{alg:Bz_estimate}
-%   \Input{$T$ samples from a WSS process $x: \{x(t)\}_{t = 1}^T$\\
-%     Maximum iterations $M$\\
-%     Sparsity parameters $\{(\delta_m\, b_m, R_m)\}_{m = 1}^M$}
-%   \Output{Linear one-step-ahead estimator $\widehat{\B}(z)$}
-%   \Initialize{$\epsilon^{(0)}(t) = x(t)$ \texttt{\# initial error process}\\
-%     $m = 1$ \texttt{\# iteration counter}}
-%   \BlankLine
-%   \For{$m = 1, \ldots, M$}{
-%     $F \leftarrow \mathsf{pwgc}(\epsilon^{(m - 1)}(t))$ \texttt{\# estimate pairwise Granger-causality statistics}\\
-%     $\widehat{\gcg}_m \leftarrow \mathsf{pw\_scg}(F, \delta_m, b_m, R_m)$ \texttt{\# graph estimate (Algorithm \ref{alg:pwgr_heuristic})}\\
-%     $\widehat{\B}^{(m)}(z) \leftarrow \mathsf{estimate\_B}(\epsilon^{(m - 1)}(t), \widehat{\gcg}_m)$ \texttt{\# Estimate filter with topology} $\widehat{\gcg}_m$\\
-%     $\epsilon^{(m)}(t) \leftarrow \epsilon^{(m - 1)}(t) - \widehat{\B}^{(m)}(z)\epsilon^{(m - 1)}(t)$ \texttt{\# updated error process}\\
-%   }
-%   \BlankLine
-%   $\widehat{\B}(z) \leftarrow \widehat{\B}^{(M)}(z) \prod_{k = 1}^{M - 1}\big(I - \widehat{\B}^{(M - k)}(z)\big)$ \texttt{\# compute total filter}\\
-%   \Return{$\widehat{\B}(z)$}
-% \end{algorithm}
-
-\section{Empirical Evaluation}
-\label{sec:empirical_evaluation}
-We have implemented our empirical experiments in Python \cite{scipy},
-in particular we leverage the LASSO implementation from
-\texttt{sklearn} \cite{sklearn} and the random graph generators from
-\texttt{networkx} \cite{networkx}.  We run experiments using two
-separate graph topologies having $n$ nodes: a strongly causal graph
-(SCG) and a directed acyclic graph (DAG).  These are generated
-respectively by drawing a random tree and a random Erdos Renyi graph
-(with edge probability $q = \frac{2}{n}$ resulting in approximately
-the same number of edges for the SCG as for the DAG), then creating a
-directed graph by directing edges from lower numbered nodes to higher
-numbered nodes.
-
-We populate each of the edges (including self loops) with random
-linear filters constructed by placing $5$ transfer function poles
-(i.e. $p = 5$) uniformly at random in a disc of radius $3 / 4$ (which
-guarantees stability for acyclic graphs).  The resulting system is
-driven by i.i.d. Gaussian random noise, each component having random
-variance $\sigma_i^2 = 1/2 + r_i$ where $r_i \sim \text{exp}(1/2)$.
-We set $p_{\text{max}} = 15$.  Results and representative graphs are
-collected in Figures \ref{fig:random_graph_topologies},
-\ref{fig:simulation_results_comparison1},
-\ref{fig:simulation_results_comparison2},
-\ref{fig:simulation_results_scaling_and_small_T},
-\ref{fig:simulation_results_dense}.
-
-We compare our results against the adaptive LASSO
-\cite{adaptive_lasso_zou2006}, which outperformed substantially both
-the LASSO and the grouped LASSO.  Motivated by scaling, we split the
-squared error term into separate terms, one for each group of incident
-edges on a node, and estimate the collection of $n$ incident filters
-$\big\{\B_{ij}(z)\big\}_{j = 1}^n$ that minimizes
-$\xi_i^{\text{LASSO}}$ in the following:
-
-\begin{equation}
-  \begin{aligned}
-  \xi_i^{\text{LASSO}}(\lambda) &= \underset{B}{\text{min}}\ \frac{1}{T}\sum_{t = p + 1}^T\big(x_i(t) - \sum_{\tau = 1}^p\sum_{j = 1}^n B_{i, j}(\tau) x(t - \tau)\big)^2 + \lambda \sum_{\tau = 1}^p \sum_{j = 1}^n |B_{ij}(\tau)|\\
-  \xi_i^{\text{LASSO}} &= \underset{\lambda \ge 0}{\text{min}}\ \xi_i^{\text{LASSO}}(\lambda) + \mathsf{BIC}\big(B_i^{\text{LASSO}}(\lambda)\big)\\
-  \end{aligned}
-\end{equation}
-
-where we are choosing $\lambda$, the regularization parameters, via the BIC.
-
-\begin{figure}
-  \centering
-  \caption{Representative Random Graph Topologies}
-  \label{fig:random_graph_topologies}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \caption{Random SCG}
-    \includegraphics[width=\linewidth]{example_scg.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \caption{Random DAG $(q = \frac{2}{n})$}
-    \includegraphics[width=\linewidth]{example_dag.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \caption{Random DAG $(q = \frac{4}{n})$}
-    \includegraphics[width=\linewidth]{example_dag_q.pdf}
-  \end{subfigure}
-\end{figure}
-
-\begin{remark}[Graph Topologies]
-  We depict in Figure \ref{fig:random_graph_topologies} the topologies
-  of random graphs used in our empirical evaluation.  For values of
-  $q$ close to $\frac{2}{n}$, the resulting random graphs tend to have
-  a topology which is, at least qualitatively, close to the SCG.  As
-  the value of $q$ increases, the random graphs deviate farther from
-  the SCG topology, and we therefore expect the LASSO to outperform
-  PWGC for larger values of $q$.  This can be observed in Figure
-  \ref{fig:simulation_results_dense}, at least where performance is
-  measured by the support recovery (i.e. via the MCC).
-\end{remark}
-
-\begin{remark}[MCC as a Support Recovery Measurement]
-  We apply Matthew's Correlation Coefficient (MCC)
-  \cite{matthews1975comparison} as a statistic for measuring support
-  recovery performance.  This statistic synthesizes the confusion
-  matrix into a single score appropriate for unbalanced labels and is
-  calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
-  performance, $0$ being the performance of random guessing, and $-1$
-  being perfectly opposed.
-\end{remark}
-
-\begin{remark}[Error Measurement]
-  We estimate the 1-step ahead prediction error by forming the variance matrix estimate
-
-  \begin{equation*}
-    \widehat{\Sigma}_v \defeq \frac{1}{T_{\text{out}}} \sum_{t = 1}^{T_{\text{out}}} (x(t) - \widehat{x}(t))(x(t) - \widehat{x}(t))^\T
-  \end{equation*}
-
-  on a long stream of out-of-sample data.  We then report the quantity
-
-  \begin{equation*}
-    \frac{\ln \tr \widehat{\Sigma}_v}{\ln \tr \Sigma_v}
-  \end{equation*}
-
-  where $\widehat{\Sigma}_v = \Sigma_v$ is the best possible performance.
-\end{remark}
-
-\begin{figure}
-  \centering
-  \caption{PWGC Compared Against AdaLASSO \cite{adaptive_lasso_zou2006} (SCG)}
-  \label{fig:simulation_results_comparison1}
-  \includegraphics[width=0.95\textwidth]{new_lasso_comparison_scg_pmax15_simulation.pdf}
-
-  {\scriptsize Comparison of PWGC and LASSO for $\VAR(p)$ model
-    estimation.  We make comparisons against both the MCC and the
-    relative log mean-squared prediction error
-    $\frac{\ln\tr \widehat{\Sigma}_v}{\ln\tr \Sigma_v}$.  Results
-    in Figure \ref{fig:simulation_results_comparison1} are for systems
-    guaranteed to satisfy the assumptions required for Theorem
-    \ref{thm:scg_recovery}.}
-\end{figure}
-
-\begin{figure}
-  \caption{PWGC vs adaLASSO (DAG, $q = \frac{2}{n}$)}
-  \label{fig:simulation_results_comparison2}
-  \includegraphics[width=0.95\textwidth]{new_lasso_comparison_dag_pmax15_simulation.pdf}
-
-  {\scriptsize Figure \ref{fig:simulation_results_comparison2}
-    provides results for systems which do not guarantee the
-    assumptions of Theorem \ref{thm:scg_recovery}, though the graph
-    has a similar level of sparsity.}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \caption{PWGC Scaling and Small Sample Performance}
-  \label{fig:simulation_results_scaling_and_small_T}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{Fixed $T$, increasing $n$ (SCG)}
-    \label{fig:simulation_results_scaling}
-    \includegraphics[width=\linewidth]{new_increasing_n_simulation.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{MCC Comparison for $T \le 100$}
-    \label{fig:small_T_comparison}
-    \includegraphics[width=\linewidth]{new_mcc_comparison001.pdf}
-  \end{subfigure}
-
-  {\scriptsize Figure \ref{fig:simulation_results_scaling} measures
-    support recovery performance as the number of nodes $n$ increases,
-    and the edge proportion as well as the number of samples $T$ is
-    held fixed.  Remarkably, the degradation as $n$ increases is
-    limited, it is primarily the graph topology (SCG or non-SCG) as
-    well as the level of sparsity (measured by $q$) which are the
-    determining factors for support recovery performance.
-
-    Figure \ref{fig:small_T_comparison} provides a support recovery
-    comparison for very small values of, $T$ typical for many
-    applications.}
-\end{figure}
-
-\begin{figure}
-  \caption{Fixed $T, n$, increasing edges $q$ (DAG)}
-  \label{fig:simulation_results_dense}
-  \includegraphics[width=\linewidth]{dag_increasing_q_small_T_alasso_simulation.pdf}
-
-  {\scriptsize Figure \ref{fig:simulation_results_dense} provides a
-    comparison between PWGC and AdaLASSO as the density of graph edges
-    (as measured by $q$) increases.  For reference,
-    $\frac{2}{n} = 0.04$ has approximately the same level of sparsity
-    as the SCGs we simulated.  As $q$ increases, the AdaLASSO
-    outperforms PWGC as measured by the MCC.  However, PWGC maintains
-    superior performance for 1-step-ahead prediction.  We speculate
-    that this is a result of fitting the sparsity pattern recovered by
-    PWGC via OLS which directly seeks to optimize this metric, whereas
-    the LASSO is encumbered by the sparsity inducing penalty.}
-\end{figure}
-
-In reference to figure \ref{fig:simulation_results_comparison1} it
-should not be overly surprising that our PWGC algorithm performs
-better than the LASSO for the case of a strongly causal graph, since
-in this case the assumptions which guarantee the correctness of
-Algorithm \ref{alg:finite_pwgc} hold.  However, the performance is
-still markedly superior in the case of a more general DAG.  We would
-conjecture that a DAG having a similar degree of sparsity as an SCG is
-likely to be ``close'' to an SCG.  Figure
-\ref{fig:simulation_results_dense} illustrates the severe (expected)
-degradation in performance as the number of edges increases while
-the number of data samples $T$ remains fixed.  For larger values $q$
-in this plot, the number of edges in the graph is comparable to the
-number of data samples.
-
-We have also paid close attention to the performance of PWGC in the
-very small sample ($T \le 100$) regime (see Figure
-\ref{fig:small_T_comparison}), as this is the regime many applications
-must contend with.
-
-In regards scalability, we have observed that performing the $O(n^2)$
-pairwise Granger-causality calculations consumes the vast majority
-($> 90\%$) of the computation time.  Since this step is trivially
-parallelizable, our algorithm also scales well with multiple cores or
-multiple machines.  Figure \ref{fig:simulation_results_scaling} is a
-demonstration of this scalability, where we are able to estimate
-graphs having over $1500$ nodes (over $2.25 \times 10 ^6$ possible edges)
-using only $T = 500$ data points, granted, an SCG on this many nodes
-is extremely sparse.
-
-\section{Application}
-\label{sec:application}
-In this section we apply our methods to a real set of EEG data
-obtained from the ``EEG Database Data
-Set''\footnote{http://archive.ics.uci.edu/ml/datasets/EEG+Database}
-\cite{zhang1995event} on the UCI machine learning repository
-\cite{uci_mlr}.  This dataset contains 1 second long measurements of
-(64 channel) EEG signals from patients who are given visual stimuli.
-The subjects in the study are labeled as being either ``control'' or
-``alcoholic'', however, we will ignore this label and instead focus on
-\textit{distinguishing between subjects} based on the
-Granger-causality graphs inferred from the subject's trials.  Our
-reasoning is to focus on the underlying question: ``Does our PWGC
-algorithm uncover meaningful Granger-causal connections from EEG
-data?''.  Focusing only on the subject label allows us to answer this
-question without simultaneously grappeling with the physiological
-question of whether alcoholic subjects as a group have discernable
-differences in their EEG readings\footnote{Some exploratory analysis
-  actually suggests that alcoholic Granger-causality graphs are not
-  substantively different}, which is a stronger requirement than
-simply that there are meaningful distinctions between the EEG readings
-of subjects generally.
-
-The dataset consists of
-$\mathcal{D} = \{\big(x^{(i)}(t)\big)_{t = 1}^T, y^{(i)} \}_{i = 1}^N$
-where $T = 256$, $x^{(i)}(t) \in \R^{64}$,
-$y^{(i)} \in [N_{subjects}]$ (with $N_{subjects} = 119$), and
-$N = 10723$.  There are on average 90 trials for each subject,
-ranging between 30 and 119.
-
-We have constructed a simple pipeline for discriminating between
-subjects by first applying an iterative PWGC algorithm (see Algorithm
-\ref{alg:iterated_pwgc}) directly to the EEG data $x^{(i)}$ to obtain
-a Granger-causality graph $\widehat{G}^{(i)}$.  We then feed the
-vectorized adjacency matrix of $\widehat{G}^{(i)}$ through a
-polynomial Kernel multinomial logistic regression model with
-parameters fit by cross validation.
-
-\begin{remark}
-  The iterated application of PWGC is an example of a heuristic by
-  which graph estimates that aren't constrained to being strongly
-  causal can be obtained.  We defer to future work the theoretical
-  analysis of more sophisticated heuristics, or algorithms appropriate
-  for different topological assumptions.
-\end{remark}
-
-\begin{algorithm}
-  \SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{output}
-  \SetKwInOut{Initialize}{initialize}
-  \DontPrintSemicolon
-
-  \BlankLine
-  \caption{Iterated PWGC Heuristic}
-  \label{alg:iterated_pwgc}
-
-  \Input{Data $x(t)$, maximum number of iterations $N$,
-    threshold $r \in (0, 1)$.}
-  \Output{Granger-causality graph estimate $\widehat{G}$ obtaind from
-    iterated application of PWGC Algorithm \ref{alg:finite_pwgc}}
-  \Initialize{$i = 1$, $\widehat{x}_0(t) \leftarrow x(t)$,
-    $\sigma_0 \leftarrow Var\ x(t)$}
-
-  \While{$i \le N$} {
-    Estimate $\VAR$ model via PWGC on $x_{i - 1}(t)$ to obtain $\widehat{x}_{i - 1}(t)$ and $\widehat{G}_{i - 1}$\\
-    $\epsilon_i(t) \leftarrow x_{i - 1}(t) - \widehat{x}_{i - 1}(t)$ \texttt{\# Compute residuals}\\
-    $\sigma_i^2 \leftarrow \text{MSE}\big(\epsilon_i(t)\big)$ \texttt{\# Compute MSE}\\
-    \If{$\sigma_i^2 > r \sigma_{i - 1}^2$}{
-      \texttt{break  \# Insignificant Improvement}
-    }
-    $i \leftarrow i + 1$
-  }
-  $\widehat{G} \leftarrow \bigcup_{i} \widehat{G}_{i}$ \texttt{\# Combine PWGC graph estimates}\\
-  \Return{$\widehat{G}$}
-\end{algorithm}
-
-While it is almost certainly possible to achieve much greater
-classification accuracy on this dataset by constructing a classifier
-to act directly on $x^{(i)}(t)$, our purpose is to demonstrate that
-relevant information is being captured by the graph estimates
-$\widehat{G}^{(i)}$.  That is, that the latent structure
-$\widehat{G}^{(i)}$ may provide scientifically relevant insights, as
-opposed to simply being a feature for classification tasks.
-
-Figures \ref{fig:ar_eeg_example} and \ref{fig:eeg_oos_error}
-illustrate the appropriateness of modelling $x(t)$ with $\VAR$ models,
-as opposed simply to a collection of $n$ unidimensional
-autoregressiive models.  The conclusion being that the
-Granger-causality graph estimates are not purely spurious.
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{$\mathsf{AR}(p)$ Model Example}
-    \label{fig:ar_eeg_example}
-    \includegraphics[width=\linewidth]{eeg_estimate_example.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{OOS Error for a Single Subject}
-    \label{fig:eeg_oos_error}
-    \includegraphics[width=\linewidth]{eeg_oos_demonstration.pdf}
-  \end{subfigure}
-
-  {\scriptsize Figure \ref{fig:ar_eeg_example} provides an example of
-    an autoregressive model fir via the Levinson-Durbin algorithm
-    (choosing the system order via the BIC) on a single EEG channel.
-    Linear autoregressive models appear qualitatively to be adequate
-    for this data.  Figure \ref{fig:eeg_oos_error} demonstrates the
-    improvement of a unified $\VAR$ model over independent
-    $\mathsf{AR}$ models, providing evidence that intra-node edges
-    inferred by PWGC are not simply spurious.
-
-  Out-of-Sample estimates are performed on data that were not used in
-  fitting the models.}
-\end{figure}
-
-We provide the final results in Figure
-\ref{fig:logistic_regression_results} where the classification
-accuracy is quantified by the (multiclass) MCC of $\approx 0.20$ on a held
-out validation set consisting of $20\%$ of the data.  This MCC score
-exceeds by a modest margin the performance of randomly guessing,
-lending strong evidence to the assertion that PWGC successfully
-recovers some meaningful differences between the Granger-causality
-graphs of different subjects, particularly if we recall that there are
-generally fewer examples for each subject than there are subjects
-overall.  Moreover, distinguishing only between two particular
-subjects (as opposed to distinguishing between one subject and $118$
-others) is substantially easier -- an illustration is provided in
-Figure \ref{fig:logistic_regression_2pair} where we embed the data
-into the plane through supervised dimensionality reduction.  In
-general, fitting a simple classifier between two subjects achieves an
-$MCC$ in excess of $0.6$ and up to $0.9$, even when the same
-hyperparameters are carried over for different pairs of subjects.
-
-\begin{figure}
-  \centering
-  \caption{Subject Classification from Granger-causality Graphs}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{Distinguishing Individual Subjects}
-    \label{fig:logistic_regression_2pair}
-    \includegraphics[width=\linewidth]{logistic_regression_separation.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{Complete Multiclass Classifier}
-    \label{fig:logistic_regression_results}
-    \includegraphics[width=\linewidth]{logistic_regression_confusion_matrix.pdf}
-
-    {\scriptsize }
-  \end{subfigure}
-
-  {\scriptsize Figure \ref{fig:logistic_regression_2pair} illustrates
-    discriminating between two particular subjects based on their
-    EEG Granger-causality graphs.  In this case $MCC = 0.75$ on held
-    out data.  Visualization is constructed by supervised
-    dimensionality reduction and is purely illustrative.
-
-    Figure \ref{fig:logistic_regression_results} provides the
-    row-normalized confusion matrix (computed on held out validation
-    data) of a multiclass logistic regression classifier used to
-    classify subjects based on their EEG Granger-causality graphs.  $MCC = 0.20$}
-
-\end{figure}
-
 \section{Conclusion}
@@
 properties of Granger-causality networks can provide substantial
-theoretical insights as well as computational benefits.  In
-particular, the notion of a strongly-causal graph has been exploited
-to establish conditions under which pairwise causality testing alone
-is sufficient for recovering a complete Granger-causality graph.
-Moreover, examples from the literature suggest that such topological
-assumptions may be reasonable in some applications and secondly, even
-when the strong-causality assumption is not met, we have provided
-simulation evidence to suggest that our pairwise testing algorithm
-PWGC can still outperform the LASSO and adaLASSO, both of which are
-commonly employed in applications.  Application to inference of more
-general graphs may also be within reach via an iterative graph
-building process as suggested in Section \ref{sec:application}.
-
-Our application in Section \ref{sec:application} provides strong
-evidence that PWGC is capable of uncovering meaningful features from
-networks of time series data by constructing a classifier which
-accurately discriminates between subjects in an EEG experiment based
-purely on the Granger-causality graph topology inferred by the
-iterated application of PWGC.
+insights into the structure of causality graphs with potential for
+improvements in causality graph estimation when structural assumptions
+are met.  In particular, the notion of a strongly-causal graph has
+been exploited to establish conditions under which pairwise causality
+testing alone is sufficient for recovering a complete
+Granger-causality graph.  Moreover, examples from the literature
+suggest that such topological assumptions may be reasonable in some
+applications.
 
@@
 case of Gaussian data \cite{barnett2009granger}, do our results extend
-to causality networks defined via transfer entropy?  Finally, the work
-of \cite{barnett2015granger} has established the superiority of
-Granger-causality testing in state space models (as opposed to pure
-autoregressions) in many cases.  Combining this work with our PWGC
-algorithm (by modifying the approach described in Section
-\ref{sec:pairwise_hypothesis_testing} to instead utilize state-space
-Granger-causality testing) therefore is likely to enable application
-to very large networks of time series data which are not well
-approximated by finite $\VAR(p)$ models.
+to causality networks defined via transfer entropy?
+
+Due to constraints on space, we report upon details for practical
+finite sample implementation as well as more extensive simulation
+results and an application to EEG data elsewhere \todo{where?}.
 

### 2019-07-14 paper/paper.tex
Trimmed down theory to create a draft for an application paper.

@@
 
-The following is a fundamental building block:
+\begin{lemma}
+  \label{lem:separated_ancestor_uncorrelated}
+  Consider distinct nodes $i, j$ in a Granger-causality graph
+  $\gcg$.  If
+
+  \begin{enumerate}[label=(\alph*)]
+    \item{$j \not\in \anc{i}$ and $i \not\in \anc{j}$}
+    \item{$\anc{i}\cap\anc{j} = \emptyset$}
+  \end{enumerate}
+
+  then $\H_t^{(i)} \perp \H_t^{(j)}$, that is,
+  $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.  Moreover,
+  this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_{t - 1}^i} = 0$.
+\end{lemma}
+
+\begin{remark}
+  The possibility that there exist nodes sharing both $i$ and $j$ as
+  ancestors is not excluded, the point being that the temporal nature
+  of Granger-causality eliminates the problems caused by ``colliders''
+  (in the language of Pearl's causal calculus
+  \cite{pearl2009causality}) in static causal inference.
+\end{remark}
+
+\begin{proof}
+  We show directly that
+  $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.  To this end,
+  fix $s, \tau \ge 0$, then by expanding with equation \eqref{eqn:ancestor_expansion} we have
+
+  \begin{align*}
+    \E x_i(t - s)x_j(t - \tau)
+    &= \E \big(\A_{ii}(z)v_i(t - s)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)\\
+    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\E[\big(\A_{ik}(z)v_k(t - s)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)]\\
+    &+ \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[\big(\A_{ii}(z)v_i(t - s)\big) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
+    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ik}(z)v_k(t - s)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)].
+  \end{align*}
+
+  Keeping in mind that $v(t)$ is an isotropic and uncorrelated
+  sequence we see that each of these above four terms are 0: the
+  first term since $i \ne j$, the second and third since
+  $j \not\in \anc{i}$ and $i \not\in \anc{j}$ and finally the fourth since
+  $\anc{i} \cap \anc{j} = \emptyset$.
+\end{proof}
+
+It is possible for components of $x(t)$ to be correlated at some time
+lags without resulting in pairwise causality:
 
@@
 
-  then $H_t^{(i)} \perp \H_t^{(j)}$, that is,
-  $\forall \tau \in \Z_+\ \E[x_i(t)x_j(t - \tau)] = 0$.  Moreover,
-  this implies that $j \npwgc i$.
+  then $j \npwgc i$.
 \end{lemma}
-
-\begin{remark}
-  Note that $H_t^{(i)} \perp \H_t^{(j)}$ states that the \textit{present}
-  of $x_i(t)$ is uncorrelated with the whole history of $x_j(t)$, but
-  we need not have $\H_t^{(i)} \perp \H_t^{(j)}$ since $i \in \anc{j}$ is
-  not excluded.
-\end{remark}
-
 \begin{proof}
-  We show directly that
-  $\forall \tau \in \Z_+\ \E[x_i(t)x_j(t - \tau)] = 0$.  To this end,
-  fix $\tau \ge 0$, then by expanding with equation \eqref{eqn:parent_expansion} we have
+  By Theorem \ref{thm:granger_causality_equivalences} it suffices to show that
+
+  \begin{equation*}
+    \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^i}} = 0,
+  \end{equation*}
+
+  which by the orthogonality principle is equivalent to
+
+  \begin{equation}
+    \label{eqn:proof_inner0}
+    \inner{x_i(t)}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^i}} = 0.
+  \end{equation}
+
+  Define the disjoint sets
 
   \begin{align*}
-    \E x_i(t)x_j(t - \tau) &= \E \big(\A_{ii}(z)v_i(t)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)\\
-    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\E[\big(\A_{ik}(z)v_k(t)\big)v_j(t - \tau)] + \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
-    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ik}(z)v_k(t)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)].
+    C_0(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, k \ne i \}\\
+    C_1(u) &= \{k \in \pa{u}\ |\ i \in \anc{k} \text{ or } k = i \}.
   \end{align*}
 
-% (i.e.
-%   $\forall \tau \in \Z,\ k \ne \ell \Rightarrow \E v_k(t)v_\ell(t -
-%   \tau) = 0$ and
-%   $\forall \tau \in \Z\setminus\{0\}\ \E v_i(t)v_i(t - \tau) = 0$),
-
-  Keeping in mind that $v(t)$ is an isotropic and uncorrelated
-  sequence we will see that each of these above four terms are 0.  The
-  first term since $i \ne j$.  The second since $j \not\in \anc{i}$.
-  The fourth since $\anc{i} \cap \anc{j} = \emptyset$.  For the third
-  term, it is possible that $i \in \anc{j}$, which requires more care.
-
-  Similarly as before we can see immediately that
+  We can then expand the parents of $j$ using Equation \ref{eqn:parent_expansion} as
 
   \begin{equation*}
-    \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[v_i(t) \big(\A_{jk}(z) v_k(t - \tau)\big)] = \E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)].
+    x_j(t) = v_j(t) + \sum_{k \in C_0(j)} B_{jk}(z) x_k(t) + \sum_{k \in C_1(j)} B_{jk}(z) x_k(t),
   \end{equation*}
 
-  However, $k \ne \ell \Rightarrow A_{k\ell}(0) = 0$ (i.e.
-  $\A(z) = I + A(1)z^{-1} + \cdots$) so $\A_{ji}(z)v_i(t - \tau)$ is
-  strictly earlier in time (for $\tau \ge 0$) than $v_i(t)$ and hence
-  $\E[v_i(t)\big(\A_{ji}(z)v_i(t - \tau)\big)] = 0$.
+  which when substituted into the left hand side of Equation
+  \ref{eqn:proof_inner0} results (by Lemma
+  \ref{lem:separated_ancestor_uncorrelated}) in
+
+  \begin{equation*}
+    \inner{x_i(t)}{\sum_{k \in C_1(j)} B_{jk}(z) x_k(t) - \sum_{k \in C_1(j)} \linE{B_{jk}(z) x_k(t)}{\H_{t - 1}^i}}.
+  \end{equation*}
+
+  We can continue this process recursively (i.e. split each
+  $k \in C_1(j)$ into $C_0(k)$ and $C_1(k)$) which must eventually
+  terminate with $C_1(u) = \{i\}$.  Therefore, there exists some
+  causal filter $\Phi(z)$ such that Equation \ref{eqn:proof_inner0} is
+  equivalent to
+
+  \begin{equation}
+    \label{eqn:proof_inner0}
+    \inner{x_i(t)}{\Phi(z)x_i(t) - \linE{\Phi(z)x_i(t)}{\H_{t - 1}^i}},
+  \end{equation}
+
+  which is $0$ since $\linE{\Phi(z)x_i(t)}{\H_{t - 1}^i} = \Phi(z)x_i(t)$.
 \end{proof}

### 2019-07-14 paper_theory/paper_theory.tex
Trimmed down theory to create a draft for an application paper.

@@
-\documentclass[12pt]{article}
+\documentclass{statsoc}
+% \usepackage{mathtime}
+\usepackage{ccfonts,eulervm}
+% \usepackage[T1]{fontenc}
+% \usepackage[lite]{belleek}
+% \fontfamily{belleek}
+% \usepackage[math]{belleek}
+% \usepackage{concretemath-fonts}
+
+% ---------------------------------------------------------------------------------------------------------------------------------
+% The statsoc style file is broken?
+% I have copied this code straight from stack overflow
+% https://tex.stackexchange.com/questions/351768/problem-with-statsoc-class-and-pdflatex
+\usepackage{geometry}
+
+\geometry{
+  textwidth=33pc,
+  textheight=\dimexpr48\baselineskip+\topskip\relax,
+  marginparsep=11pt,
+  marginparwidth=107pt,
+  footnotesep=6.65pt,
+  headheight=9pt,
+  headsep=9pt,
+  footskip=30pt,
+}
+% ---------------------------------------------------------------------------------------------------------------------------------
+
+% \usepackage{natbib}
+\bibliographystyle{chicago}
 
@@
 %% bibliography stuff -- this needs come before the preamble inclusion
-\usepackage[backend=bibtex,sorting=none]{biblatex}
 \usepackage{enumitem}
@@
 \usepackage{todonotes}
-\bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
-% \usepackage{apxproof}
-
-% \bibliography{\string~global_bib.bib}
 
@@
 
-\input{\string~/Documents/academics/global_academics/latex_preamble}
-% \input{\string~latex_preamble}
+% \usepackage{fullpage}
+\usepackage{framed}
+
+% Figures
+\usepackage{graphicx}
+\usepackage{caption}  % This is not recommended?
+\usepackage{subcaption}
+% \usepackage{wrapfig}
+% \usepackage{svg}
+
+% Math packages, theorem definitions and numbering
+\usepackage{amsmath}
+\usepackage{amssymb}
+% \usepackage{amsthm}
+% \usepackage{mathrsfs} % Fancy scripted font
+% \usepackage{bm}  % Bold math
+
+% Misc packages
+\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
+\usepackage{algorithm2e} %{algorithm} environment
+\usepackage{soul}  % \hl highlighting
+\usepackage{color}
+\usepackage{mathtools}  % For my \ceil function
+
+% Theorems (with italics)
+% \theoremstyle{plain}  % Style definition removes italics
+\newtheorem{theorem}{Theorem}
+\newtheorem{corollary}{Corollary}
+\newtheorem{proposition}{Proposition}
+\newtheorem{lemma}{Lemma}
+
+% \theoremstyle{definition}
+\newtheorem{remark}{Remark}
+\newtheorem{definition}{Definition}
+\newtheorem{example}{Example}
+\newtheorem{assumptions}{Assumptions}
+
+% keywords
+\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}
+
+% General
+\def\defeq{\overset{\Delta}{=}}  % Equal with triangle
+\def\cl{\mathsf{cl\ }}  % Closure
+\newcommand{\sgn}[1]{\mathsf{sgn}(#1)}  % sign function
+
+% Calculus
+\def\d{\mathsf{d}}  % Differential operator
+
+% Functions
+\def\ln{\mathsf{ln\ }}  % Natural logarithm
+\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}  % Ceiling
+
+% Probability
+\def\H{\mathcal{H}}  % Hilbert space
+\def\E{\mathbb{E}}  % Expectation
+\def\Var{\text{Var}}  % Variance
+\def\P{\mathbb{P}}  % Probability Measure
+\def\F{\mathcal{F}}  % A sigma algebra
+\def\sX{\mathcal{X}}  % Another sigma algebra
+\def\KL{\mathbf{D}_{KL}}  % KL divergence
+\def\bF{\mathbf{F}}  % Whole F-meas space
+
+% Standard sets
+\def\Z{\mathbb{Z}}  % Set of integers
+\def\R{\mathbb{R}}  % Set of real numbers
+\def\C{\mathbb{C}}  % Set of complex numbers
+\def\N{\mathbb{N}}  % Set of natural numbers
+\def\ball{\mathbb{B}}  % Open ball
+\def\clball{\overline{\ball}}  % Closed ball
+
+% Linear algebra
+\def\rk{\mathsf{rk }}  % The rank
+\def\tr{\mathsf{tr }}  % The trace
+\def\T{\mathsf{T}}  % Transpose notation
+\def\c{\mathsf{c}}  % complement
+\def\dg{\mathsf{dg }}   %  Diagonal vector of a matrix
+\def\Dg{\mathsf{Dg }}   %  Diagonal matrix from a vector
+\def\ind{\mathbf{1}}  % Ones vector or indicator
+\def\matvec{\textbf{vec}}  % Vector operator
+\def\<{\langle}  % < Inner product
+\def\>{\rangle}  % > Inner product
+\newcommand{\inner}[2]{\langle #1, #2 \rangle}  % Inner product
+\newcommand{\innerT}[2]{#1^\T #2}  % Inner product for finite vectors
 
@@
 
-\title{Graph Topological Aspects of Granger Causal Network Learning }
-\author{R. J. Kinnear \\
-  \small\href{mailto:ryan@kinnear.ca}{ryan@kinnear.ca} \\
-  \small\url{https://github.com/RJTK} \and R. R. Mazumdar \\
-  \small\href{mailto:mazum@uwaterloo.ca}{mazum@uwaterloo.ca} }
+\title{Graph Topological Aspects of Granger Causal Network Learning}
+\author[Author 1 {\it et al.}]{R. J. Kinnear}
+\address{
+  University of Waterloo,
+  Waterloo,
+  Canada.}
+\email{Ryan@Kinnear.ca}
+
+\author{R. R. Mazumdar}
+\address{
+  University of Waterloo,
+  Waterloo,
+  Canada.}
 
 \begin{document}
-\maketitle
+
 \abstract{We study Granger-causality in the context of wide-sense
@@
 % \tableofcontents
-% \clearpage
+\clearpage
 
@@
 tests are of interest since the computation of such pairwise relations
-is substantially easier..
+is substantially easier.
 
@@
 case in general, as different paths through $\gcg$ can lead to
-cancellation (see example \ref{ex:diamond_cancellation}).  In fact, we
-do not even have $j \in \pa{i} \Rightarrow j \pwgc i$ (see example
-\ref{ex:lag_cancellation}).
-
-\begin{example}
-  \label{ex:diamond_cancellation}
-  Firstly, on $n = 4$ nodes, ``diamond'' shapes can lead to cancellation on paths of length 2:
-
-\begin{equation*}
-  x(t) =
-  \left[
-    \begin{array}{cccc}
-      0 & 0 & 0 & 0\\
-      a & 0 & 0 & 0\\
-      -a & 0 & 0 & 0\\
-      0 & 1 & 1 & 0\\
-    \end{array}
-  \right] x(t - 1) + v(t),
-\end{equation*}
-
-with $\E v(t) = 0,\ \E v(t)v(t - \tau)^\T = \delta_\tau I$.
-
-By directly calculating
-
-\begin{align*}
-  x_4(t) &= x_2(t - 1) + x_3(t - 1) + v_4(t)\\
-         &= ax_1(t - 2) + av_2(t - 1) - ax_1(t - 2) -av_3(t - 1) + v_4(t)\\
-         &= a(v_2(t - 1) - v_3(t - 1)) + v_4(t),
-\end{align*}
-
-we see that, since $v(t)$ is isotropic white noise, $1 \npwgc 4$.  The problem here is that there are multiple paths from $x_1$ to $x_4$.
-\end{example}
+cancellation (see Figure \ref{fig:diamond_cancellation}).  In fact, we
+do not even have $j \in \pa{i} \Rightarrow j \pwgc i$ (see Figure
+\ref{fig:lag_cancellation}).
+
+% \begin{example}
+%   \label{ex:diamond_cancellation}
+%   Firstly, on $n = 4$ nodes, ``diamond'' shapes can lead to cancellation on paths of length 2:
+
+% \begin{equation*}
+%   x(t) =
+%   \left[
+%     \begin{array}{cccc}
+%       0 & 0 & 0 & 0\\
+%       a & 0 & 0 & 0\\
+%       -a & 0 & 0 & 0\\
+%       0 & 1 & 1 & 0\\
+%     \end{array}
+%   \right] x(t - 1) + v(t),
+% \end{equation*}
+
+% with $\E v(t) = 0,\ \E v(t)v(t - \tau)^\T = \delta_\tau I$.
+
+% By directly calculating
+
+% \begin{align*}
+%   x_4(t) &= x_2(t - 1) + x_3(t - 1) + v_4(t)\\
+%          &= ax_1(t - 2) + av_2(t - 1) - ax_1(t - 2) -av_3(t - 1) + v_4(t)\\
+%          &= a(v_2(t - 1) - v_3(t - 1)) + v_4(t),
+% \end{align*}
+
+% we see that, since $v(t)$ is isotropic white noise, $1 \npwgc 4$.  The problem here is that there are multiple paths from $x_1$ to $x_4$.
+% \end{example}
 
@@
   \centering
-  \begin{subfigure}[b]{0.45\textwidth}
+  \begin{subfigure}[b]{0.35\textwidth}
     \includegraphics[width=\linewidth]{example1.pdf}
-    \caption{Graph Corresponding to Example \ref{ex:diamond_cancellation}}
+    \caption{Example multiple path cancellation: $j \in \anc{i} \nRightarrow j \pwgc i$}
     \label{fig:diamond_cancellation}
-    \centering{$j \in \anc{i} \nRightarrow j \pwgc i$}
   \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
+  \begin{subfigure}[b]{0.35\textwidth}
     \includegraphics[width=\linewidth]{example2.pdf}
-    \caption{Graph Corresponding to Example \ref{ex:lag_cancellation}}
+    \caption{Example cancellation from time lag: $j \in \pa{i} \nRightarrow j \pwgc i$}
     \label{fig:lag_cancellation}
-    \centering{$j \in \pa{i} \nRightarrow j \pwgc i$}
   \end{subfigure}
@@
 
-\begin{example}
-  \label{ex:lag_cancellation}
-  A second example on $n = 3$ nodes is also worth examining, in this case
-  cancellation is a result of differing time lags.
-
-\begin{equation*}
-  x(t) =
-  \left[
-    \begin{array}{ccc}
-      0 & 0 & 0\\
-      -a & 0 & 0\\
-      0 & 1 & 0\\
-    \end{array}
-  \right] x(t - 1) +
-  \left[
-    \begin{array}{ccc}
-      0 & 0 & 0\\
-      0 & 0 & 0\\
-      a & 0 & 0\\
-    \end{array}
-  \right] x(t - 2) + v(t)
-\end{equation*}
-
-Then
-
-\begin{align*}
-  x_2(t) &= v_2(t) - ax_1(t - 1)\\
-  x_3(t) &= v_3(t) + x_2(t - 1) + ax_1(t - 2)\\
-  \Rightarrow x_3(t) &= v_2(t - 1) + v_3(t),
-\end{align*}
-
-and again $1 \npwgc 3$.
-\end{example}
+% \begin{example}
+%   \label{ex:lag_cancellation}
+%   A second example on $n = 3$ nodes is also worth examining, in this case
+%   cancellation is a result of differing time lags.
+
+% \begin{equation*}
+%   x(t) =
+%   \left[
+%     \begin{array}{ccc}
+%       0 & 0 & 0\\
+%       -a & 0 & 0\\
+%       0 & 1 & 0\\
+%     \end{array}
+%   \right] x(t - 1) +
+%   \left[
+%     \begin{array}{ccc}
+%       0 & 0 & 0\\
+%       0 & 0 & 0\\
+%       a & 0 & 0\\
+%     \end{array}
+%   \right] x(t - 2) + v(t)
+% \end{equation*}
+
+% Then
+
+% \begin{align*}
+%   x_2(t) &= v_2(t) - ax_1(t - 1)\\
+%   x_3(t) &= v_3(t) + x_2(t - 1) + ax_1(t - 2)\\
+%   \Rightarrow x_3(t) &= v_2(t - 1) + v_3(t),
+% \end{align*}
+
+% and again $1 \npwgc 3$.
+% \end{example}
 
@@
 
-\clearpage
-\printbibliography
+% \bibliography{\string~global_bib.bib}
+\bibliography{\string~/Documents/academics/global_academics/global_bib}
 \end{document}

### 2019-07-14 software/comparison_table.py
Working on code for comparison table.

@@ -1,0 +1,89 @@
+"""
+Create a condensed table of data for comparison against AdaLASSO.
+"""
+
+import pandas as pd
+import numpy as np
+import xarray as xr
+
+from itertools import product
+
+from pwgc.var_system import (random_gnp_dag, drive_gcg, get_errors,
+                             get_estimation_errors, random_scg,
+                             get_X, attach_node_prop, gcg_to_var)
+from pwgc.gc_methods import (compute_pairwise_gc, estimate_B, pw_scg,
+                             estimate_graph, full_filter_estimator,
+                             compute_MCC_score, estimate_dense_graph,
+                             alasso_fista_estimate_dense_graph)
+
+alpha = 0.05
+
+
+def main():
+    np.random.seed(0)
+    n_nodes, p_lags, p_max = 25, 5, 10
+    N_iters = 50
+
+    params = {"T": [50, 250, 1000],
+              "q": ["SCG", 2. / n_nodes, 4. / n_nodes, 16. / n_nodes],
+              "metric": ["MCC", "Err", "FDR"],
+              "iter": list(range(N_iters))}
+
+    def create_empty_ds():
+        _adalasso = _create_empty_da()
+        _pwgc = _create_empty_da()
+        return xr.Dataset(coords=params,
+                          data_vars={"adalasso": _adalasso,
+                                     "pwgc": _pwgc})
+
+    def _create_empty_da():
+        return xr.DataArray(data=np.nan * np.empty((len(params["T"]),
+                                                    len(params["q"]),
+                                                    N_iters,
+                                                    len(params["metric"]))),
+                            dims=["T", "q", "iter", "metric"],
+                            coords=params)
+
+    def make_random_graph(q):
+        if q == "SCG":
+            return random_scg(n_nodes, p_lags, pole_rad=0.75)
+        else:
+            return random_gnp_dag(n_nodes, p_lags, pole_rad=0.75,
+                                  edge_prob=q)
+
+    def make_test_data(T, G):
+        sv2_true = 0.5 + np.random.exponential(0.5, size=n_nodes)
+        drive_gcg(G, 500 + min(2 * T, T + 15000), sv2_true, filter_attr="b(z)")
+        X = get_X(G)
+        X = X - np.mean(X, axis=0)[None, :]
+        X = X
+        return X, sv2_true
+
+    results = create_empty_ds()
+
[…]

### 2019-07-14 software/lasso_comparison.py

@@ -26,6 +26,9 @@ from pwgc.gc_methods import (compute_pairwise_gc, estimate_B, pw_scg,
                              compute_MCC_score, estimate_dense_graph,
                              alasso_fista_estimate_dense_graph)
 
+# NOTE: The FISTA version does pretty badly.  I'm not yet entirely sure why,
+# NOTE: but I have a hard time trusting my own implementation.
+
 
 class TrackErrors:
     def __init__(self, N_iters, T_iters, n_nodes):
@@ -216,7 +219,7 @@ def lasso_mcc_comparison(simulation_name, showfig=False, fista=False):
 
 def lasso_comparison(simulation_name, graph_type, showfig=False, fista=False):
     np.random.seed(0)
-    n_nodes, p_lags, p_max = 50, 5, 5
+    n_nodes, p_lags, p_max = 50, 5, 15
     alpha, N_iters = 0.05, 2
 
     T_iters = list(map(int, np.linspace(50, 5000, 200)))

### 2019-07-14 software/comparison_table.py
Minor updates -- just need to calculate FDRs.

@@ -14,15 +14,15 @@ from pwgc.var_system import (random_gnp_dag, drive_gcg, get_errors,
 from pwgc.gc_methods import (compute_pairwise_gc, estimate_B, pw_scg,
                              estimate_graph, full_filter_estimator,
                              compute_MCC_score, estimate_dense_graph,
-                             alasso_fista_estimate_dense_graph)
+                             alasso_fista_estimate_dense_graph, compute_fdr_score)
 
 alpha = 0.05
 
 
 def main():
     np.random.seed(0)
-    n_nodes, p_lags, p_max = 25, 5, 10
-    N_iters = 50
+    n_nodes, p_lags, p_max = 5, 5, 10
+    N_iters = 5
 
     params = {"T": [50, 250, 1000],
               "q": ["SCG", 2. / n_nodes, 4. / n_nodes, 16. / n_nodes],
@@ -33,7 +33,7 @@ def main():
         _adalasso = _create_empty_da()
         _pwgc = _create_empty_da()
         return xr.Dataset(coords=params,
-                          data_vars={"adalasso": _adalasso,
+                          data_vars={"alasso": _adalasso,
                                      "pwgc": _pwgc})
 
     def _create_empty_da():
@@ -43,6 +43,11 @@ def main():
                                                     len(params["metric"]))),
                             dims=["T", "q", "iter", "metric"],
                             coords=params)
+
+    def _create_result_da(mcc, err, fdr):
+        return xr.DataArray(data=np.array([mcc, err, fdr]),
+                            dims=["metric"],
+                            coords={"metric": ["MCC", "Err", "FDR"]})
 
     def make_random_graph(q):
         if q == "SCG":
@@ -60,30 +65,47 @@ def main():
         return X, sv2_true
 
     results = create_empty_ds()
-
     for T, q in product(params["T"], params["q"]):
         G = make_random_graph(q)
-        X, sv2_true = make_test_data(T, q)
+        X, sv2_true = make_test_data(T, G)
 
-        for it in N_iters:
+        for it in range(N_iters):
             G_hat_lasso = adalasso(X, G, T, p_max)
-            G_hat_pwgc = adalasso(X, G, T, p_max)
-            results["alasso"].loc[T, q, it, :] = calculate_error(
-                G_hat_lasso, G, sv2_true)
[…]

### 2019-07-17 paper/paper.tex
I can have a supplementary document with proofs and extra figures.

@@
 moreover, if $x_i \gc x_j$ and $x_j \gc x_k$, it is expected that
-$F_{kj} > F_{ji}$, thereby providing the same effect as proceeding
+$F_{kj} > F_{ki}$, thereby providing the same effect as proceeding
 backwards through $P_{k - r}$.

### 2019-07-17 paper_theory/paper_theory.tex
I can have a supplementary document with proofs and extra figures.

@@
 \usepackage{ccfonts,eulervm}
+\usepackage{tabularx}
+\usepackage{xr}
+\externaldocument{proofs}
 % \usepackage[T1]{fontenc}
@@
 % \usepackage{concretemath-fonts}
-
+% \interfootnotelinepenalty=10000
 % ---------------------------------------------------------------------------------------------------------------------------------
@@
 \usepackage{todonotes}
+\usepackage{multirow}
+\usepackage{booktabs}
 
@@
 % \usepackage{mathrsfs} % Fancy scripted font
-% \usepackage{bm}  % Bold math
+\usepackage{bm}  % Bold math
 
@@
 
-\clearpage
-
-% \tableofcontents
-\clearpage
+% \clearpage
+
+% % \tableofcontents
+% \clearpage
 
@@
 
-\subsection{Contributions}
+% \subsection{Contributions}
 In the context of time series data, sparsity assumptions remain
@@
 from pairwise tests alone (sec \ref{sec:pairwise_algorithm}).  We
-report \hl{elsewhere}\todo{where?  Or can I fit in a brief summary?}
-the results of extensive finite sample simulations as well as an
-application to EEG data, where in addition we summarize appropriate
-methods for hypothesis testing, model order selection, computationally
-efficient estimations, and error rate controls in practice.  Our
-simulation results establish significant potential for improvement
-over existing methods.  Concluding remarks on further open problems
-and extentions are provided in Section \ref{sec:conclusion}.
+report a summary of simulation results in \ref{sec:simulation}, with
+additional results reported in the supplementary material
+\ref{apx:simulation}.  Our simulation results establish that there is
+significant potential for improvement over existing methods, and that
+the graph-topological aspects of time series analysis are relevant for
+both theory and practice.  Concluding remarks on further open problems
+and extentions are provided in Section \ref{sec:conclusion}.  The
+proofs of each proposition and theorem are also relegated to the
+supplementary material, simple corollaries have proofs included in the
+main text.
 
@@
 
-As a consequence of the Wold decomposition theorem \cite{lindquist},
+As a consequence of the Wold decomposition theorem (see e.g. \cite{lindquist}),
 every WSS sequence has the moving average $MA(\infty)$
@@
 
-The notion of conditional orthogonality is used throughout, as well as
-some equivalent representations for Granger-causality.
-
-\begin{lemma}[\cite{lindquist} Proposition 2.4.2]
+The notion of conditional orthogonality is the essence of
+Granger-causality, and enables us to obtain results for a fairly
+general class of WSS processes, rather than simply $\VAR(p)$ models.
+
+\begin{definition}[Conditional Orthogonality]
   \label{lem:conditional_orthogonality_equivalence}
   Consider three closed subspaces of a Hilbert space $\mathcal{A}$,
-  $\mathcal{B}$, $\mathcal{X}$.  The following statements are
-  equivalent
-
-  \todo{Do I even use (a) anymore?  I think I can just put the definition of conditional orthogonality into Theorem 1.}
-
-  \begin{enumerate}
-    \item{$\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$}
-    \item{$\linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}}\ \forall \beta \in \mathcal{B}$.}
-    % \item{$\linE{\beta}{\mathcal{A}} = \linE{\linE{\beta}{\mathcal{X}}}{\mathcal{A}}\ \forall \beta \in \mathcal{B}}$}
-    \end{enumerate}
-
-    Where $\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$ denotes
-    conditional orthogonality:
+  $\mathcal{B}$, $\mathcal{X}$.  We say that $\mathcal{A}$ is
+  conditionally orthogonal to $\mathcal{B}$ given $\mathcal{X}$
+  and write $\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$ if
 
@@
     \end{equation*}
-\end{lemma}
+
+  An equivalent condition is that (see \cite{lindquist} Proposition 2.4.2)
+  \begin{equation*}
+    \linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}}\ \forall \beta \in \mathcal{B}
+  \end{equation*}
+\end{definition}
 
@@
 
-\begin{proof}
-  % Geweke uses the log of the ratio of the determinants of the residual variances
-  % 
-  % The equivalence $(1) \iff (3)$ is essentially a restatement of the
-  % definition, but is in line with the seminal work of Geweke
-  % \cite{geweke1982measurement}, \cite{geweke1984}.
-
-  $(1) \Rightarrow (2)$ follows as a result of the uniqueness of orthogonal
-  projection (i.e. the best estimate is necessarily the coefficients
-  of the model).  $(2) \Rightarrow (3)$ follows since in computing
-  $(y - \linE{y}{\H_{t - 1}^{-j}})$ for $y \in H_t^i$ it is sufficient
-  to consider $y = x_i(t)$ by linearity, then since
-  $H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$ we have
-  $(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$ since
-  $\B_{ij}(z) = 0$ and $v_i(t) \perp \H_{t - 1}$.  $(3) \iff (4)$ is a result
-  of Lemma \ref{lem:conditional_orthogonality_equivalence}.  And,
-  $(4) \implies (1)$ follows directly from the Definition.
-\end{proof}
+% \begin{proof}
+%   % Geweke uses the log of the ratio of the determinants of the residual variances
+%   % 
+%   % The equivalence $(1) \iff (3)$ is essentially a restatement of the
+%   % definition, but is in line with the seminal work of Geweke
+%   % \cite{geweke1982measurement}, \cite{geweke1984}.
+
+%   $(a) \Rightarrow (b)$ follows as a result of the uniqueness of orthogonal
+%   projection (i.e. the best estimate is necessarily the coefficients
+%   of the model).  $(b) \Rightarrow (c)$ follows since in computing
+%   $(y - \linE{y}{\H_{t - 1}^{-j}})$ for $y \in H_t^i$ it is sufficient
+%   to consider $y = x_i(t)$ by linearity, then since
+%   $H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$ we have
+%   $(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$ since
+%   $\B_{ij}(z) = 0$ and $v_i(t) \perp \H_{t - 1}$.  $(c) \iff (d)$ is a result
+%   of the equivalence in Definition
+%   \ref{lem:conditional_orthogonality_equivalence}.  And,
+%   $(d) \implies (a)$ follows directly from the Definition.
+% \end{proof}
 
@@
 \eqref{eqn:filter_action}).  Thence, the matrix $\B(z)$ is analogous
-to a \textit{weighted adjacency matrix}\footnote{We are using the
+to a \textit{weighted adjacency matrix}\footnote{\footnotesize We are using the
   convention that $\B_{ij}(z)$ is a filter with input $x_j$ and output
@@
 
-An elementary theorem connecting the adjacency matrix with paths will
-allow us to deduce the sparsity pattern of $\A(z)$.  Proof follows
-easily by induction:
-
-\begin{lemma}
-  \label{lem:adj_matrix}
-  Let $S$ be the transposed adjacency matrix of the Granger-causality
-  graph $\gcg$.  Then, $(S^k)_{ij}$ is the number of paths of length
-  $k$ from node $j$ to node $i$.  Evidently, if
-  $\forall k \in \N,\ (S^k)_{ij} = 0$ then $j \not\in \anc{i}$.
-\end{lemma}
-
 From the $\VAR$ representation of $x(t)$ there is clearly a tight
@@
 $\A_{ij}(z) = 0$.
-
-\begin{proof}
-  Equation \eqref{eqn:parent_expansion} is immediate from the
-  $\VAR(\infty)$ representation of \eqref{eqn:ar_representation} and
-  Theorem \ref{thm:granger_causality_equivalences}, we are left to
-  demonstrate \eqref{eqn:ancestor_expansion}.
-
-  From equation (\ref{eqn:ar_representation}), which we are assuming
-  throughout the paper to be invertible, we can write
-
-  \begin{equation*}
-    x(t) = (I - \B(z))^{-1} v(t),
-  \end{equation*}
-
-  where $(I - \B(z))^{-1} = \A(z)$ due to the uniqueness of
-  (\ref{eqn:wold}).  Since $\B(z)$ is stable we have
-
-  \begin{equation}
-    \label{eqn:resolvant_inv}
-    (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k.
-  \end{equation}
-
-  Invoking the Cayley-Hamilton theorem allows writing the infinite sum
-  of \eqref{eqn:resolvant_inv} in terms of \textit{finite} powers of
-  $\B$.
-
-  Let $S$ be a matrix with elements in $\{0, 1\}$ which represents the
-  sparsity pattern of $\B(z)$, from lemma \ref{lem:adj_matrix} $S$ is
-  the transpose of the adjacency matrix for $\gcg$ and hence
-  $(S^k)_{ij}$ is non-zero if and only if $j \in \gpn{k}{i}$, and
-  therefore $\B(z)^k_{ij} = 0$ if $j \not \in \gpn{k}{i}$.  Finally,
-  since $\anc{i} = \bigcup_{k = 1}^n\gpn{k}{i}$ we see that
-  $\A_{ij}(z)$ is zero if $j \not\in \anc{i}$.
-
-  Thence
-
-  \begin{align*}
-    x_i(t) &= [(I - \B(z))^{-1}v(t)]_i = \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
-    &= \A_{ii}(z) v_i(t) + \sum_{\substack{j \in \anc{i} \\ j \ne i}} \A_{ij}(z) v_j(t)
-  \end{align*}
-\end{proof}
 
@@
 
-\begin{lemma}
-  \label{lem:separated_ancestor_uncorrelated}
+The following propositions are essentially lemmas used for the proof
+of the upcomming Proposition \ref{prop:ancestor_properties}, but
+remain relevant for providing intuitive insight into the problems at
+hand.
+
+\begin{proposition}
   Consider distinct nodes $i, j$ in a Granger-causality graph
@@
   this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_{t - 1}^i} = 0$.
-\end{lemma}
+\end{proposition}
 
@@
 
-\begin{proof}
-  We show directly that
-  $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.  To this end,
-  fix $s, \tau \ge 0$, then by expanding with equation \eqref{eqn:ancestor_expansion} we have
-
-  \begin{align*}
-    \E x_i(t - s)x_j(t - \tau)
-    &= \E \big(\A_{ii}(z)v_i(t - s)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)\\
-    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\E[\big(\A_{ik}(z)v_k(t - s)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)]\\
-    &+ \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[\big(\A_{ii}(z)v_i(t - s)\big) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
-    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ik}(z)v_k(t - s)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)].
-  \end{align*}
-
-  Keeping in mind that $v(t)$ is an isotropic and uncorrelated
-  sequence we see that each of these above four terms are 0: the
-  first term since $i \ne j$, the second and third since
-  $j \not\in \anc{i}$ and $i \not\in \anc{j}$ and finally the fourth since
-  $\anc{i} \cap \anc{j} = \emptyset$.
-\end{proof}
-
-It is possible for components of $x(t)$ to be correlated at some time
-lags without resulting in pairwise causality:
-
-\begin{lemma}
-  \label{lem:ancestor_uncorrelated}
+\begin{remark}
+  It is possible for components of $x(t)$ to be correlated at some
+  time lags without resulting in pairwise causality.  For instance,
+  the conclusion $j \npwgc i$ of Proposition
+  \ref{prop:separated_ancestor_uncorrelated} will still hold even if
+  $i \in \anc{j}$, since $j$ cannot provide any information about $i$
+  that is not available from observing $i$ itself.  However, it is
+  important to note that this is the case only if we have direct
+  access to $x(t)$.  That is, if we instead observed a noisy version
+  of $x$ (i.e. $\tilde{x}(t) = x(t) + n(t)$) the observation of
+  $\tilde{x}_j$ would enable conditionally improved estimates of
+  $\tilde{x}_i$, without there necessarily being any causal flow in
+  the underlying system $x(t)$.
+\end{remark}
+
+\begin{proposition}
   Consider distinct nodes $i, j$ in a Granger-causality graph $\gcg$.
@@
   then $j \npwgc i$.
-\end{lemma}
-\begin{proof}
-  By Theorem \ref{thm:granger_causality_equivalences} it suffices to show that
-
-  \begin{equation*}
-    \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^i}} = 0,
-  \end{equation*}
-
-  which by the orthogonality principle is equivalent to
-
-  \begin{equation}
-    \label{eqn:proof_inner0}
-    \inner{x_i(t)}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^i}} = 0.
-  \end{equation}
-
-  Define the disjoint sets
-
-  \begin{align*}
-    C_0(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, k \ne i \}\\
-    C_1(u) &= \{k \in \pa{u}\ |\ i \in \anc{k} \text{ or } k = i \}.
-  \end{align*}
-
-  We can then expand the parents of $j$ using Equation \ref{eqn:parent_expansion} as
-
-  \begin{equation*}
-    x_j(t) = v_j(t) + \sum_{k \in C_0(j)} B_{jk}(z) x_k(t) + \sum_{k \in C_1(j)} B_{jk}(z) x_k(t),
-  \end{equation*}
-
-  which when substituted into the left hand side of Equation
-  \ref{eqn:proof_inner0} results (by Lemma
-  \ref{lem:separated_ancestor_uncorrelated}) in
-
-  \begin{equation*}
-    \inner{x_i(t)}{\sum_{k \in C_1(j)} B_{jk}(z) x_k(t) - \sum_{k \in C_1(j)} \linE{B_{jk}(z) x_k(t)}{\H_{t - 1}^i}}.
-  \end{equation*}
-
-  We can continue this process recursively (i.e. split each
-  $k \in C_1(j)$ into $C_0(k)$ and $C_1(k)$) which must eventually
-  terminate with $C_1(u) = \{i\}$.  Therefore, there exists some
-  causal filter $\Phi(z)$ such that Equation \ref{eqn:proof_inner0} is
-  equivalent to
-
-  \begin{equation}
-    \label{eqn:proof_inner0}
-    \inner{x_i(t)}{\Phi(z)x_i(t) - \linE{\Phi(z)x_i(t)}{\H_{t - 1}^i}},
-  \end{equation}
-
-  which is $0$ since $\linE{\Phi(z)x_i(t)}{\H_{t - 1}^i} = \Phi(z)x_i(t)$.
-\end{proof}
-
-The previous Lemma can be strengthened significantly; notice
+\end{proposition}
+
+The previous result can still be strengthened significantly; notice
 that it is possible to have some $k \in \anc{i} \cap \anc{j}$ where still
@@
 
-\begin{proof}
-  We will prove by way of contradiction.  To this end, suppose that
-  $j$ is a node such that: $(a)$ $j \not \in \anc{i}$ and $(b)$ for
-  every $k \in \anc{i} \cap \anc{j}$ every
-  $k \rightarrow \cdots \rightarrow j$ path contains $i$.
-
-  Firstly, notice that every $u \in \big(\pa{j} \setminus \{i\}\big)$
-  necessarily inherits these same two properties.  This follows since
-  if we also had $u \in \anc{i}$ then $u \in \anc{i} \cap \anc{j}$ so that every
-  $u \rightarrow \cdots \rightarrow j$ path must contain $i$, but
-  $u \in \pa{j}$, so this is not the case since $u \rightarrow j$ is a
-  path that doesn't contain $i$; moreover, if we consider
-  $w \in \anc{i} \cap \anc{u}$ then we also have
-  $w \in \anc{i} \cap \anc{j}$ so every
-  $w \rightarrow \cdots \rightarrow j$ path must contain $i$.  These
-  properties therefore extend inductively to every
-  $u \in \big(\anc{j} \setminus \{i\}\big)$.
-
-  In order to deploy a recursive argument, define the following
-  partition of $\pa{u}$, for some node $u$:
-
-  \begin{align*}
-    C_0(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, \anc{i} \cap \anc{k} = \emptyset, k \ne i\}\\
-    C_1(u) &= \{k \in \pa{u}\ |\ i \in \anc{k} \text{ or } k = i\}\\
-    C_2(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, \anc{i} \cap \anc{k} \ne \emptyset, k \ne i\}.
-  \end{align*}
-
-  We notice that for any $u$ having the properties $(a), (b)$ above, we
-  must have $C_2(u) = \emptyset$ since if $k \in C_2(u)$ then
-  $\exists w \in \anc{i} \cap \anc{k}$ s.t. $i \not \in \anc{k}$ and
-  therefore there must be a path $\gcgpath{w}{k} \rightarrow u$ which
-  does not contain $i$.
-
-  Using this partition, we will expand $x_j(t)$ in terms of it's
-  parents, and recursively expand nodes in $C_1$ until we reach a case
-  where $C_1 = \emptyset$.  For the first step equation
-  \eqref{eqn:parent_expansion} gives us:
-
-  \begin{equation}
-    \label{eqn:xj_partition_expansion}
-    x_j(t) = \A_{jj}(z)\Big(v_j(t) + \sum_{k \in C_0(j)}\B_{ik}(z)x_k(t) + \sum_{k \in C_1(j)}\B_{ik}(z)x_k(t)\Big).
-  \end{equation}
-
-  Using this representation we choose an arbitrary $\Phi(z) x_i(t - 1) \in \H_{t - 1}^{(j)}$ and show that
-
-  \begin{equation}
-    \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}{\Phi(z)x_j(t - 1) - \linE{\Phi(z)x_j(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
-  \end{equation}
-
-  which will imply (by Theorem \ref{thm:granger_causality_equivalences}) that $j \npwgc i$ and for which it is equivalent to show that
-
-  \begin{equation}
-    \label{eqn:sufficient_inner_prod}
-    \inner{x_i(t)}{\Phi(z)x_j(t - 1) - \linE{\Phi(z)x_j(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
-  \end{equation}
-
-  by the orthogonality principle since $\linE{x_i(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$.
-Substituting (\ref{eqn:xj_partition_expansion}) into (\ref{eqn:sufficient_inner_prod}) and starting with the first term we have
-
-  \begin{align*}
-    &\inner{x_i(t)}{\Phi(z)\A_{jj}(z)v_j(t - 1) - \linE{\Phi(z)\A_{jj}(z)v_j(t - 1)}{\H_{t - 1}^{(i)}}}\\
-    \overset{(\alpha)}{=}\ &\inner{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}{\Phi(z)\A_{jj}(z)v_j(t - 1)}\\
-    \overset{(\beta)}{=}\ &0,
-  \end{align*}
-
-  where $(\alpha)$ follows by expanding $x_i(t)$ with (\ref{eqn:ancestor_expansion}) and $\linE{\Phi(z)\A_{jj}(z)v_j(t - 1)}{\H_{t - 1}^{(i)}} = 0$ because $\forall \tau, s$
-
-  \begin{equation*}
-    \E v_j(t - \tau) x_i(t - s) = \E v_j(t - \tau) \sum_{k \in \anc{i} \cup \{i\}}\A_{ik}(z)v_k(t - s) = 0,
-  \end{equation*}
-
-  since $j \not \in \anc{i}$; $(\beta)$ follows similarly, that is, $j \not \in \anc{i}$.  Secondly we see that $\forall k \in C_0(j)$
-
-  \begin{equation*}
-    \inner{x_i(t)}{\Phi(z)\A_{jj}(z)\B_{ik}(z)x_k(t - 1) - \linE{\Phi(z)\A_{jj}(z)\B_{ik}(z)x_k(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
-  \end{equation*}
-
-  which follows from Lemma \ref{lem:ancestor_uncorrelated}.  Finally,
-  for $k \in C_1(j)$ the case $k = i$ is immediate (since the error in
-  estimating $\Phi(z)\B_{ii}(z)x_i(t)$ given $\H_{t - 1}^{(i)}$ is
-  $0$), so suppose $k \ne i$.  We know from above that $k$ inherits the
-  key properties referred to as $(a)$ and $(b)$ above and therefore we
-  can recursively expand $k$ in the same way as in equation
-  (\ref{eqn:xj_partition_expansion}).  Continuing this recursion for
-  each $k \in C_1(j)$ (where $k \ne i$) must eventually terminate since
-  $i \in \anc{k}$.
-\end{proof}
+\begin{remark}
+  The interpretation of this proposition is that for $j \pwgc i$ then
+  there must either be ``causal flow'' from $j$ to $i$
+  ($j \in \anc{i}$) or there must be a confounder $k$ through which
+  common information is received.
+\end{remark}
 
@@
 
-% \begin{example}
-%   \label{ex:diamond_cancellation}
-%   Firstly, on $n = 4$ nodes, ``diamond'' shapes can lead to cancellation on paths of length 2:
-
-% \begin{equation*}
-%   x(t) =
-%   \left[
-%     \begin{array}{cccc}
-%       0 & 0 & 0 & 0\\
-%       a & 0 & 0 & 0\\
-%       -a & 0 & 0 & 0\\
-%       0 & 1 & 1 & 0\\
-%     \end{array}
-%   \right] x(t - 1) + v(t),
-% \end{equation*}
-
-% with $\E v(t) = 0,\ \E v(t)v(t - \tau)^\T = \delta_\tau I$.
-
-% By directly calculating
-
-% \begin{align*}
-%   x_4(t) &= x_2(t - 1) + x_3(t - 1) + v_4(t)\\
-%          &= ax_1(t - 2) + av_2(t - 1) - ax_1(t - 2) -av_3(t - 1) + v_4(t)\\
-%          &= a(v_2(t - 1) - v_3(t - 1)) + v_4(t),
-% \end{align*}
-
-% we see that, since $v(t)$ is isotropic white noise, $1 \npwgc 4$.  The problem here is that there are multiple paths from $x_1$ to $x_4$.
-% \end{example}
+\begin{example}
+  \label{ex:diamond_cancellation}
+  Firstly, on $n = 4$ nodes, ``diamond'' shapes can lead to cancellation on paths of length 2:
+
+\begin{equation*}
+  x(t) =
+  \left[
+    \begin{array}{cccc}
+      0 & 0 & 0 & 0\\
+      a & 0 & 0 & 0\\
+      -a & 0 & 0 & 0\\
+      0 & 1 & 1 & 0\\
+    \end{array}
+  \right] x(t - 1) + v(t),
+\end{equation*}
+
+with $\E v(t) = 0,\ \E v(t)v(t - \tau)^\T = \delta_\tau I$.
+
+By directly calculating
+
+\begin{align*}
+  x_4(t) &= x_2(t - 1) + x_3(t - 1) + v_4(t)\\
+         &= ax_1(t - 2) + av_2(t - 1) - ax_1(t - 2) -av_3(t - 1) + v_4(t)\\
+         &= a(v_2(t - 1) - v_3(t - 1)) + v_4(t),
+\end{align*}
+
+we see that, since $v(t)$ is isotropic white noise, $1 \npwgc 4$.  The problem here is that there are multiple paths from $x_1$ to $x_4$.
+\end{example}
 
@@
   \centering
+  \caption{Examples illustrating the difficulty of obtaining a converse to Proposition \ref{prop:ancestor_properties}}
   \begin{subfigure}[b]{0.35\textwidth}
     \includegraphics[width=\linewidth]{example1.pdf}
-    \caption{Example multiple path cancellation: $j \in \anc{i} \nRightarrow j \pwgc i$}
+    \caption{Path cancellation: $j \in \anc{i} \nRightarrow j \pwgc i$}
     \label{fig:diamond_cancellation}
@@
     \includegraphics[width=\linewidth]{example2.pdf}
-    \caption{Example cancellation from time lag: $j \in \pa{i} \nRightarrow j \pwgc i$}
+    \caption{Cancellation from time lag: $j \in \pa{i} \nRightarrow j \pwgc i$}
     \label{fig:lag_cancellation}
@@
 
-% \begin{example}
-%   \label{ex:lag_cancellation}
-%   A second example on $n = 3$ nodes is also worth examining, in this case
-%   cancellation is a result of differing time lags.
-
-% \begin{equation*}
-%   x(t) =
-%   \left[
-%     \begin{array}{ccc}
-%       0 & 0 & 0\\
-%       -a & 0 & 0\\
-%       0 & 1 & 0\\
-%     \end{array}
-%   \right] x(t - 1) +
-%   \left[
-%     \begin{array}{ccc}
-%       0 & 0 & 0\\
-%       0 & 0 & 0\\
-%       a & 0 & 0\\
-%     \end{array}
-%   \right] x(t - 2) + v(t)
-% \end{equation*}
-
-% Then
-
-% \begin{align*}
-%   x_2(t) &= v_2(t) - ax_1(t - 1)\\
-%   x_3(t) &= v_3(t) + x_2(t - 1) + ax_1(t - 2)\\
-%   \Rightarrow x_3(t) &= v_2(t - 1) + v_3(t),
-% \end{align*}
-
-% and again $1 \npwgc 3$.
-% \end{example}
+\begin{example}
+  \label{ex:lag_cancellation}
+  A second example on $n = 3$ nodes is also worth examining, in this case
+  cancellation is a result of differing time lags.
+
+\begin{equation*}
+  x(t) =
+  \left[
+    \begin{array}{ccc}
+      0 & 0 & 0\\
+      -a & 0 & 0\\
+      0 & 1 & 0\\
+    \end{array}
+  \right] x(t - 1) +
+  \left[
+    \begin{array}{ccc}
+      0 & 0 & 0\\
+      0 & 0 & 0\\
+      a & 0 & 0\\
+    \end{array}
+  \right] x(t - 2) + v(t)
+\end{equation*}
+
+Then
+
+\begin{align*}
+  x_2(t) &= v_2(t) - ax_1(t - 1)\\
+  x_3(t) &= v_3(t) + x_2(t - 1) + ax_1(t - 2)\\
+  \Rightarrow x_3(t) &= v_2(t - 1) + v_3(t),
+\end{align*}
+
+and again $1 \npwgc 3$.
+\end{example}
 
@@
   We will say that a Granger-causality graph $\gcg$ is
-  \textit{strongly causal} if there is at most 1 directed path between
+  \textit{strongly causal} if there is at most one directed path between
   any two nodes.  Strongly Causal Graphs will be referred to as SCGs.
@@
   regulatory network of \textit{Saccharomyces cerevisiae}.  These
-  networks, which we reproduce\footnote{Figure \ref{fig:gene_network1}
-    is reproduced under the Creative Commons Attribution
-    Non-Commercial License
-    (\url{http://creativecommons.org/licenses/by-nc/2.5}) and Figure
-    \ref{fig:gene_network2} under the Creative Commons Attribution
-    License (\url{https://creativecommons.org/licenses/by/4.0/})} in
-  figure \ref{fig:gene_networks} appear to have at most a small number
-  of edges which violate the strong-causality condition.
+  networks, which we reproduce in figure \ref{fig:gene_networks}
+  appear to have at most a small number of edges which violate the
+  strong-causality condition.
 
@@
     \end{subfigure}
+
+  {\footnotesize Figure \ref{fig:gene_network1}
+    is reproduced under the Creative Commons Attribution
+    Non-Commercial License
+    (http://creativecommons.org/licenses/by-nc/2.5) and Figure
+    \ref{fig:gene_network2} under the Creative Commons Attribution
+    License (https://creativecommons.org/licenses/by/4.0/)}
   \end{figure}
@@
 \end{proposition}
-\begin{proof}
-  Suppose that there is a path from $k$ to $i$ which does not contain
-  $j$.  In this case, there are multiple paths from $k$ to $i$ (one of
-  which \textit{does} go through $j$ since $j \in \anc{i}$) which
-  contradicts the assumption of strong causality.
-\end{proof}
 
@@
 \end{proposition}
-\begin{proof}
-  We will show that for some $\psi \in \H_{t - 1}^{(j)}$ we have
-
-  \begin{equation}
-    \label{eqn:cond_ortho_proof}
-    \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} \ne 0
-  \end{equation}
-
-  and therefore that $H_t^{(i)} \not\perp\ \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(i)}$, which by theorem (\ref{thm:granger_causality_equivalences}) is enough to establish that $j \pwgc i$.
-
-  Firstly, we will establish a representation of $x_i(t)$ that involves $x_j(t)$.
-Denote by $a_{r + 1} \rightarrow a_r \rightarrow \cdots \rightarrow a_1 \rightarrow a_0$ with $a_{r + 1} \defeq j$ and $a_0 \defeq i$ the \textit{unique} $\gcgpath{j}{i}$ path in $\gcg$, we will expand the representation of equation (\ref{eqn:parent_expansion}) backwards along this path:
-
-  % Should this be written as a lemma?
-  \begin{align*}
-    x_i(t) &= v_i(t) + \B_{ii}(z) x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z) x_k(t)\\
-           &= \underbrace{v_{a_0}(t) + \B_{a_0a_0}(z) x_i(t) + \sum_{\substack{k \in \pa{a_0} \\ k \ne a_1}}\B_{a_0 k}(z) x_k(t)}_{\defeq \wtalpha{a_0}{a_1}} + \B_{a_0a_1}(z)x_{a_1}(t)\\
-           &= \wtalpha{a_0}{a_1} + \B_{a_0a_1}(z)\big[\wtalpha{a_1}{a_2} + \B_{a_1a_2}(z)x_{a_2}(t) \big]\\
-           &\overset{(a)}{=} \sum_{\ell = 0}^r \underbrace{\Big(\prod_{m = 0}^{\ell - 1} \B_{a_m a_{m + 1}}(z) \Big)}_{\defeq F_\ell(z)} \wtalpha{a_\ell}{a_{\ell + 1}} + \Big(\prod_{m = 0}^{r}\B_{a_m a_{m + 1}}(z)\Big)x_{a_{r + 1}}(t)\\
-           &= \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}} + F_{r + 1}(z) x_j(t)
-  \end{align*}
-
-  where $(a)$ follows by a routine induction argument and where we define $\prod_{m = 0}^{-1} \bullet \defeq 1$ for notational convenience.
-
-  Using this representation to expand equation (\ref{eqn:cond_ortho_proof}), we obtain the following cumbersome expression:
-
-  \begin{align*}
-    &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}}\\
-    &- \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\linE{\sum_{\ell = 0}^r F_\ell(z)\wtalpha{a_\ell}{a_{\ell + 1}}}{\H_{t - 1}^{(i)}}}\\
-    &+ \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\sum_{\ell = 0}^r F_\ell(z)\wtalpha{a_\ell}{a_{\ell + 1}}}.
-  \end{align*}
-
-  Note that by the orthogonality principle, $\psi - \linE{\psi}{\H_{t - 1}^{(i)}} \perp \H_{t - 1}^{(i)}$, the middle term above is $0$.
-Choosing now the particular value $\psi = F_{r + 1}(z)x_j(t) \in \H_{t - 1}^{(j)}$ we arrive at
-
-  \begin{align*}
-    &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}\\
-    &= \E|F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}|^2\\
-    &+ \inner{F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}}{\sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}}},
-  \end{align*}
-
-  which by the Cauchy-Schwarz inequality is $0$ if and only if
-
-  \begin{equation*}
-    \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}} \overset{\text{a.s.}}{=} \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}} - F_{r + 1}(z)x_j(t),
-  \end{equation*}
-
-  or by rearranging and applying the representation obtained earlier, if and only if
-
-  \begin{equation*}
-    x_i(t) \overset{\text{a.s.}}{=} \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}},
-  \end{equation*}
-
-  but this is impossible since $x_i(t) \not \in \H_{t - 1}^{(i)}$.
-\end{proof}
 
@@
   Granger-causality does not suffer any difficulty with so-called
-  ``colliders'', that is, the topology $i \rightarrow k \leftarrow j$
-  will never result in $i \pwgc j$ or $j \pwgc i$.  This is evidently
-  an advantage of the \textit{temporal} nature of Granger-causality -- there is
-  no backwards causal flow along the edges of $\gcg$.
+  ``colliders'' (in the language of Pearl's causal calculus
+  \cite{pearl2009causality}), that is, the topology
+  $i \rightarrow k \leftarrow j$ will never result in $i \pwgc j$ or
+  $j \pwgc i$.  This is evidently an advantage of the
+  \textit{temporal} nature of Granger-causality -- there is no
+  backwards causal flow along the edges of $\gcg$.
 \end{remark}
@@
   converse to proposition \ref{prop:ancestor_properties} is not
-  possible without additional conditions.  Consider the ``fork'' system on $3$
-  nodes (i.e. $2 \leftarrow 1 \rightarrow 3$) defined by
+  possible without additional conditions.  Consider the ``fork''
+  system on $3$ nodes (i.e. $2 \leftarrow 1 \rightarrow 3$) defined by
 
@@
   \begin{align}
-    \tau_0(\B) &= \text{min}\{\tau \in \Z_+\ |\ b(\tau) \ne 0\},\\
-    \tau_{\infty}(\B) &= \text{sup}\{\tau \in \Z_+\ |\ b(\tau) \ne 0\}.\\
+    \tau_0(\B) &= \text{inf }\{\tau \in \Z_+\ |\ b(\tau) \ne 0\},\\
+    \tau_{\infty}(\B) &= \text{sup }\{\tau \in \Z_+\ |\ b(\tau) \ne 0\}.\\
   \end{align}
@@
   be persistent since these naturally result in an equivalent
-  $\mathsf{MA}(\infty)$ representation, except for the pathological
-  case where $\B(z)$ is Nilpotent.
+  $\mathsf{MA}(\infty)$ representation, except for the pathological case
+  where $\B(z)$ is Nilpotent.  Indeed, in a $\VAR(1)$ model, if the
+  system matrix is diagonalizable and has atleast two distinct
+  eigenvalues, then the system is persistent.  See Example
+  \ref{ex:persistent_system}.
 
@@
   $k$.
-
 \end{remark}
@@
 \begin{example}
+  \label{ex:persistent_system}
   process $x(t)$ generated by the $\VAR(1)$ model\footnote{Recall that
@@
 
-\begin{lemma}
-  \label{lem:time_lag_cancellation}
-  Suppose $v(t)$ is a scalar sequence with unit variance and zero
-  autocorrelation and let $\A(z), \B(z)$ be nonzero and strictly
-  causal (i.e. $\tau_0(\A) \ge 1$) linear filters.  Then,
-
-  \begin{equation}
-    \inner{F(z)\A(z)v(t)}{\B(z)v(t)} = 0\ \forall \text{ strictly causal filters } F(z)
-  \end{equation}
-
-  if and only if $\tau_0(\A) \ge \tau_\infty(\B)$.
-\end{lemma}
-\begin{proof}
-  We have
-
-  \begin{align}
-    \inner{\A(z)v(t)}{\B(z)v(t)} &= \sum_{\tau = 1}^\infty \sum_{s = 1}^\infty a(\tau)b(s)\E[v(t - s)v(t - \tau)]\\
-    &= \sum_{\tau = \text{max}(\tau_0(\A), \tau_0(\B))}^{\text{min}(\tau_\infty(\A), \tau_\infty(\B))} a(\tau) b(\tau)\\
-  \end{align}
-
-  due to the uncorrelatedness assumptions on $v(t)$.  This expression
-  is $0$ if and only if $\tau_0(\A) \ge 1 + \tau_\infty(\B)$ or if
-  $\tau_0(\B) \ge 1 + \tau_\infty(\A)$ or if the coefficients are
-  orthogonal along the common support.
-
-  Specializing this fact to $\inner{F(z)\A(z)v(t)}{\B(z)v(t)} = 0$ we
-  see that the coefficients cannot be orthogonal for every choice of
-  $F$, and that $\text{sup}_F \tau_\infty(F\A) = \infty$, leaving only
-  the possibility that
-
-  \begin{align*}
-    \tau_0(F\A) \ge 1 + \tau_\infty(\B) \forall F &\overset{(a)}{\iff} \tau_0(\A) \ge 1 + \tau_\infty(\B) - \underset{F}{\text{min }} \tau_0(F)\\
-    &\overset{(b)}{\iff} \tau_0(\A) \ge \tau_\infty(\B),
-  \end{align*}
-
-  where $(a)$ follows since $\tau_0(F\A) = \tau_0(F) + \tau_0(\A)$,
-  and $(b)$ since $\text{min}_F\ \tau_0(F) = 1$.
-\end{proof}
-
-\begin{corollary}
-  \label{cor:time_lag_cancellation}
-  For $k \in \anc{i} \cap \anc{j}$ we have
-
-  \begin{align*}
-    \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} &= 0\ \forall \text{ strictly causal } F(z)\\
-    \iff \inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} &= 0\ \forall \text{ strictly causal } F(z)\\
-    \iff \tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})
-  \end{align*}
-\end{corollary}
-\begin{proof}
-  The final equivalence follows immediately from Lemma \ref{lem:time_lag_cancellation}.  For the first equivalence we have
-
-  \begin{align*}
-    \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} &= 0\ \forall \text{ strictly causal } F(z)\\
-    \iff \inner{F(z)A_{jk}(z)v_k(t)}{x_i(t - \tau)} &= 0\ \forall \tau \ge 1, \text{ strictly causal } F(z),
-  \end{align*}
-
-  which can be expanded by equation \eqref{eqn:ancestor_expansion} to
-  obtain (after cancelling all ancestors of $i$ other than $k$)
-
-  \begin{equation*}
-    \inner{F(z)A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t - \tau)} = 0\ \forall \tau \ge 1, \text{ strictly causal } F(z),
-  \end{equation*}
-
-  which by the Lemma is equivalent to $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$ as stated.
-\end{proof}
-
 \begin{proposition}
@@
 \end{proposition}
-\begin{proof}
-  We will show that $j \pwgc i$, the other being symmetric.  First
-  note also that by proposition \ref{prop:sc_graph_common_anc} we
-  cannot have $i \in \anc{j}$ or $j \in \anc{i}$ and therefore every
-  $k \in \anc{i}\cap\anc{j}$ will be a confounder.
-
-  It is sufficient to show that $\exists \psi \in \H_{t - 1}^{(j)}$
-  such that
-
-  \begin{equation*}
-    \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} \ne 0.
-  \end{equation*}
-
-  To this end, let $F(z)$ be an arbitrary but strictly causal linear
-  filter.  We apply equation \eqref{eqn:ancestor_expansion} to
-  $x_i(t)$ and $\psi \defeq F(z)x_j(t)$:
-
-  \begin{align*}
-    &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}\\
-    &\overset{(a)}{=} \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
-    &\overset{(b)}{=} \inner{\sum_{k \in \anc{j}}\big(F(z)\A_{jk}(z)v_k(t) - \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}\big)}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
-    &\overset{(c)}{=} \sum_{k \in \anc{i}\cap\anc{j}}\Big(\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} - \inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{ii}v_i(t)}\\
-    &- \sum_{\ell \in \anc{i}}\inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{i\ell}(z)v_\ell(t)}\Big)
-  \end{align*}
-
-  where in $(a)$ we have removed the $\linE{x_i(t)}{\H_{t - 1}^{(i)}}$
-  term via the orthogonality principle, in $(b)$ there is no
-  $F(z)\A_{jj}(z)v_j(t)$ term since due to $j \not\in \anc{i}$ it is
-  orthogonal to $\H_t^{(i)}$.  Finally, $(c)$ follows by applying
-  orthogonality properties of $v(t)$, as well as the fact that
-  $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$ for
-  $k \not \in \anc{i}$.  Note that
-  $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$
-  and thence there is in general no cancellation in the final term
-  above for $\ell \in \anc{i}$.
-
-  \todo{This is clearly not immediately evident}
-
-  This is $0$ for every $F$ if and only if for all $F$ and
-  $\forall k \in \anc{i} \cap \anc{j}$ we have
-  $$\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$$ and
-  $$\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} = 0,$$ which by
-  Corollary \ref{cor:time_lag_cancellation} occurs if and only if
-  $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$, which is impossible since by persistence
-  $\tau_0(\A_{jk}) < \infty$ and  $\tau_\infty(\A_{ik}) = \infty$.
-\end{proof}
+
+\begin{remark}
+  The importance of this result is that it allows us to eliminate the
+  effects of confounding by 
+\end{remark}
 
@@
   (\ref{alg:pwgr}).
-
-  \begin{algorithm}
+\end{theorem}
+
+\begin{algorithm}[H]
   \SetKwInOut{Input}{input}
@@
 
-    \BlankLine
-    \caption{Pairwise Graph Recovery}
+  % \BlankLine
+  \caption{Pairwise Granger Causality Algorithm}
   \label{alg:pwgr}
   % \TitleOfAlgo{Pairwise Graph Recovery}
-    \Input{Pairwise Granger-causality relations between a persistent
-      process of dimension $n$ whose joint Granger-causality
-      relations are known to form a strongly causal DAG $\gcg$.}
+  % \Input{Pairwise Granger-causality relations between a persistent
+  % process of dimension $n$ whose joint Granger-causality
+  % relations are known to form a strongly causal DAG $\gcg$.}
+  \Input{Pairwise Granger-causality relations}
   \Output{Edges $\gcge = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of
@@
   $W \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$  \texttt{\# candidate edges}\\
-    $P_0 \leftarrow \{i \in S_0\ |\ \forall s \in S_0\ (s, i) \not\in W\}$  \texttt{\# parent-less nodes}\\
+  $P_0 \leftarrow \{i \in S_0\ |\ \forall s \in S_0\ (s, i) \not\in W\}$  \texttt{\# parentless nodes}\\
   \While{$S_{k - 1} \ne \emptyset$}{
     $S_k \leftarrow S_{k - 1} \setminus P_{k - 1}$ \texttt{\# remove nodes with depth }$k - 1$\\
-      $P_k \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W\}$   \texttt{\# candidate children of }$P_{k - 1}$\\
-      \;
+    $P_k \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W\}$   \texttt{\# candidate children}\\
 
@@
       $Q \leftarrow E_{k - 1} \cup \big(\bigcup_{\ell = 0}^{r - 1} D_{k\ell}\big)$ \texttt{\# currently known edges}\\
-        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times P_k\ |\ (i, j) \in W,\ \text{no } \gcgpath{i}{j} \text{ path in } Q\}$
+      $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times P_k\ |\ (i, j) \in W,\ \text{no } i \rightarrow j \text{ path in } Q\}$
       % $E_k \leftarrow E_{k - 1} \cup D_{kr}$ \texttt{\# add edges to }$E_k$ \label{alg:inner_loop_end}\\
@@
     % $W_{k + 1} \leftarrow W_k \setminus \big(\bigcup_{r = 0}^k D_{kr}\big)$  \texttt{\# remove edges from consideration}\\
-      \;
     $k \leftarrow k + 1$
@@
 \end{algorithm}
-\end{theorem}
-
-We will shortly prove the theorem by establishing the correctness of
-algorithm (\ref{alg:pwgr}).  The idea is to iteratively ``peel away
-layers'' of nodes by removing the nodes that have no parents
-remaining, which always exist since the graph is acyclic.  The
-requirement of strong causality ensures that all actual edges of
-$\gcg$ manifest in some way as pairwise relations (by proposition
-\ref{prop:pwgc_anc}), and the requirement of persistence allows
-confounding to be eliminated by removing bidirectional edges.  Without
-persistence, each confounded pair would give rise to $4$ possible
-pairwise topologies consistent with $\gcg$, one for each type of
-pairwise edge (no edge, unidirectional, bidirectional).
+
+The theorem is proven in the supplementary material by establishing
+the correctness correctness of algorithm (\ref{alg:pwgr}).  The idea
+is to iteratively ``peel away layers'' of nodes by removing the nodes
+that have no parents remaining, which always exist since the graph is
+acyclic.  The requirement of strong causality ensures that all actual
+edges of $\gcg$ manifest in some way as pairwise relations (by
+proposition \ref{prop:pwgc_anc}), and the requirement of persistence
+allows confounding to be eliminated by removing bidirectional edges.
+Without persistence, each confounded pair would give rise to $4$
+possible pairwise topologies consistent with $\gcg$, one for each type
+of pairwise edge (no edge, unidirectional, bidirectional).
 
@@
     \caption{Example graph for Algorithm \ref{alg:pwgr}}
-    \footnotesize{Black arrows indicate true parent-child
+    {\footnotesize{Black arrows indicate true parent-child
         relations.  Red dotted arrows indicate pairwise causality (due to
@@
         $1$).  Blue groupings indicate each $P_k$ in Algorithm
-      \ref{alg:pwgr}.}
+        \ref{alg:pwgr}.}}
     \label{fig:example_fig3}
@@
 
-Our proof proceeds in 5 steps stated formally as lemmas.  Firstly, we
-characterize the sets $W$ and $P_k$.  Then we establish a correctness
-result for the inner loop on $r$, a correctness result for the outer
-loop on $k$, and finally that the algorithm terminates in a finite
-number of steps.
-
-\begin{lemma}[$W$ Represents Ancestor Relations]
-  \label{lem:W_subset_E}
-  In Algorithm \ref{alg:pwgr} we have
-  $(i, j) \in W$ if and only if $i \in \anc{j}$.  In particular,
-  $W \subseteq \gcge$.
-\end{lemma}
-\begin{proof}
-  Let $j \in [n]$ and suppose that $i \in \anc{j}$.  Then $i \pwgc j$
-  by Proposition \ref{prop:pwgc_anc}.  Proposition
-  \ref{prop:sc_graph_common_anc} ensures that $(i, j)$ are not
-  confounded and Corollary \ref{cor:parent_corollary} that
-  $j \not\in \anc{i}$ so $j \npwgc i$ and thence by Proposition
-  \ref{prop:ancestor_properties} $(i, j) \in W$.
-
-  Conversely, suppose $(i, j) \in W$.  Then since $j \npwgc i$
-  Proposition \ref{prop:persistence_converse} ensures that $(j, i)$
-  are not confounded and so by Proposition \ref{prop:ancestor_properties}
-  we must have $i \in \anc{j}$.
-\end{proof}
-
-\begin{definition}[Depth]
-  For our present purposes we will define the \textit{depth} $d(j)$ of
-  a node $j$ in $\gcg$ to be the length of the \textit{longest} path
-  from a node in $P_0$ to $j$, where $d(j) = 0$ if $j \in P_0$.  It is
-  apparent that such a path will always exist.  For example, in Figure
-  \ref{fig:example_fig3} we have $d(3) = 1$ and $d(4) = 2$.
-\end{definition}
-
-\begin{lemma}[Depth Characterization of $P_k$]
-  \label{lem:depth_lemma}
-  $i \in P_k \iff d(i) = k$ and $j \in S_k \iff d(j) \ge k$.
-\end{lemma}
-\begin{proof}
-  We proceed by induction, noting that $P_0$ is non-empty since $\gcg$
-  is acyclic and therefore $\gcg$ contains nodes without parents.  The
-  base case $i \in P_0 \iff d(i) = 0$ is by definition, and
-  $j \in S_0 \iff d(j) \ge 0$ is trivial since $S_0 = [n]$.  So
-  suppose that the lemma is true up to $k - 1$.
-
-  ($i \in P_k \implies d(i) = k$): Let $i \in P_k$.  Suppose that
-  $d(i) \ge k + 1$, then $\exists j \in \pa{i}$ such that
-  $j \not\in \cup_{r \ge 1}P_{k - r}$ (otherwise $d(i) \le k$), this
-  implies that $j \in S_k$ with $(j, i) \in W$ (by Lemma
-  \ref{lem:W_subset_E}) which is not possible due to the construction of
-  $P_k$ and therefore $d(i) \le k$.  Moreover,
-  $P_k \subseteq S_k \subseteq S_{k - 1}$ implies that
-  $d(i) \ge k - 1$ by the induction hypothesis, but if $d(i) = k - 1$
-  then $i \in P_{k - 1}$ again by induction which is impossible since
-  $i \in P_k$ and therefore $d(i) = k$.
-
-  ($s \in S_k \implies d(s) \ge k$): Let
-  $s \in S_k \subseteq S_{k - 1}$.  We have by induction that
-  $d(s) \ge k - 1$, but again by induction (this time on $P_{k - 1}$)
-  we have $d(s) \ne k - 1$ since $S_k = S_{k - 1} \setminus P_{k - 1}$
-  and therefore $d(s) \ge k$.
-
-  ($d(i) = k \implies i \in P_k$): Suppose $i \in [n]$ is such that
-  $d(i) = k$.  Then $i \in S_{k - 1}$ by the hypothesis, but also
-  $i \not\in P_{k - 1}$ so then
-  $i \in S_k = S_{k - 1} \setminus P_{k - 1}$ and thus $d(i) \ge k$.
-  Now, recalling the definition of $P_k$
+% \begin{example}
+%   We close this section by noting that the conditions of persistence
+%   and strong causality are only sufficient conditions.  For example,
+%   the complete directed graph with 2 nodes i.e.
+
+%   \begin{equation*}
+%     B(1) = \left[ \begin{array}{cc} 1/2 & 1 \\ 1 & 1/2 \end{array}\right]
+%   \end{equation*}
+
+%   contains a loop but is pairwise recoverable, though not by algorithm
+%   (\ref{alg:pwgr}).  Clearly, this example is somewhat artificial
+%   since when $n = 2$ there is no difference between pairwise
+%   Granger-causality and joint Granger-causality amongst all series --
+%   however, one can add any number of nodes having no parents or
+%   children to a graph containing a length 2 cycle, in which case the
+%   graph clearly remains pairwise recoverable.
+% \end{example}
+
+\section{Simulation}
+\label{sec:simulation}
+We implement an heuristic inspired by Algorithm \ref{alg:pwgr}, the
+details of which can be found in the supplementary material Section
+\ref{sec:finite_pwgc} (see Algorithm \ref{alg:finite_pwgc}.  The
+heuristic is essentially controlling the false discovery rate
+substantially below what it would be with a threshold based pairwise
+scheme.  The methods are easily parallelizable, and can scale to
+graphs with thousands of nodes on a single machine.  By contrast,
+scaling the LASSO to this large of a network (millions of variables)
+is nontrivial and extremely computationally demanding.
+
+We run experiments using two separate graph topologies having $n = 50$
+nodes: a strongly causal graph (SCG) and a directed acyclic graph
+(DAG).  Consult Section \ref{apx:simulation} for details on how data
+is generated from these models.
+
+We compare our results against the adaptive LASSO
+\cite{adaptive_lasso_zou2006}, which outperformed both the LASSO and
+the grouped LASSO by a large margin.  Motivated by scaling, we split the squared error
+term into separate terms, one for each group of incident edges on a
+node, and estimate the collection of $n$ incident filters
+$\big\{\B_{ij}(z)\big\}_{j = 1}^n$ that minimizes
+$\xi_i^{\text{LASSO}}$ in the following:
+
+\begin{equation}
+  \begin{aligned}
+  \xi_i^{\text{LASSO}}(\lambda) &= \underset{B}{\text{min}}\ \frac{1}{T}\sum_{t = p_{\text{max}} + 1}^T\big(x_i(t) - \sum_{\tau = 1}^{p_{\text{max}}}\sum_{j = 1}^n B_{i, j}(\tau) x(t - \tau)\big)^2 + \lambda \sum_{\tau = 1}^{p_{\text{max}}} \sum_{j = 1}^n |B_{ij}(\tau)|\\
+  \xi_i^{\text{LASSO}} &= \underset{\lambda \ge 0}{\text{min}}\ \xi_i^{\text{LASSO}}(\lambda) + \mathsf{BIC}\big(B_i^{\text{LASSO}}(\lambda)\big)\\
+  \end{aligned}
+\end{equation}
+
+where we are choosing $\lambda$, the regularization parameters, via the BIC.
+This is similar to the work of \cite{arnold2007temporal}, except that
+we have replacing the LASSO with the Adaptive LASSO, which provides
+dramatically superior performance.
+
+\begin{figure}
+  \centering
+  \caption{Representative Random Graph Topologies on $n = 50$ Nodes}
+  \label{fig:random_graph_topologies}
+  \begin{subfigure}[b]{0.3\textwidth}
+    \caption{SCG}
+    \includegraphics[width=\linewidth]{example_scg.pdf}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.3\textwidth}
+    \caption{DAG $(q = \frac{2}{n})$}
+    \includegraphics[width=\linewidth]{example_dag.pdf}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.3\textwidth}
+    \caption{DAG $(q = \frac{4}{n})$}
+    \includegraphics[width=\linewidth]{example_dag_q.pdf}
+  \end{subfigure}
+\end{figure}
+
+\begin{remark}[Graph Topologies]
+  We depict in Figure \ref{fig:random_graph_topologies} the topologies
+  of random graphs used in our empirical evaluation.  For values of
+  $q$ close to $\frac{2}{n}$, the resulting random graphs tend to have
+  a topology which is, at least qualitatively, close to the SCG.  As
+  the value of $q$ increases, the random graphs deviate farther from
+  the SCG topology, and we therefore expect the LASSO to outperform
+  PWGC for larger values of $q$.
+\end{remark}
+
+\begin{remark}[MCC as a Support Recovery Measurement]
+  We apply Matthew's Correlation Coefficient (MCC)
+  \cite{matthews1975comparison} as a statistic for measuring support
+  recovery performance.  This statistic synthesizes the confusion
+  matrix into a single score appropriate for unbalanced labels and is
+  calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
+  performance, $0$ being the performance of random guessing, and $-1$
+  being perfectly opposed.
+\end{remark}
+
+\begin{remark}[Error Measurement]
+  We estimate the 1-step ahead prediction error by forming the variance matrix estimate
 
   \begin{equation*}
-    P_k = \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W \},
+    \widehat{\Sigma}_v \defeq \frac{1}{T_{\text{out}}} \sum_{t = 1}^{T_{\text{out}}} (x(t) - \widehat{x}(t))(x(t) - \widehat{x}(t))^\T
   \end{equation*}
 
-  if $s \in S_k$ is such that $(s, i) \in W$ then $s \pwgc i$ and
-  $i \npwgc s$ so that by persistence and Proposition
-  \ref{prop:persistence_converse} there cannot be a confounder of
-  $(s, i)$ (otherwise $i \pwgc s$) so then by Proposition
-  \ref{prop:ancestor_properties} we have $s \in \anc{i}$.  We have
-  shown that $s \in S_k \implies d(s) \ge k$ and so we must have
-  $d(i) > k$, a contradiction, thence $s \not\in \anc{i}$,
-  $s \npwgc i$, $(s, i) \not\in W$ and $i \in P_k$.
-
-  ($d(j) \ge k \implies j \in S_k$): Let $j \in [n]$ such that
-  $d(j) \ge k$, then by induction we have $j \in S_{k - 1}$.  This
-  implies by the construction of $S_k$ that $j \not\in S_k$ only if
-  $j \in P_{k - 1}$, but we have shown that this only occurs when
-  $d(j) = k - 1$, but $d(j) > k - 1$ so $j \in S_k$.
-\end{proof}
-
-\begin{lemma}[Inner Loop]
-  \label{lem:inner_loop_lemma}
-  Fix an integer $k \ge 1$ and suppose that $(i, j) \in E_{k - 1}$ if
-  and only if $(i, j) \in \gcge$ and $d(j) \le k - 1$.  Then, we have
-  $(i, j) \in D_{kr}$ if and only if $(i, j) \in \gcge $, $d(j) = k$,
-  and $d(i) = k - r$.
-\end{lemma}
-\begin{proof}
-  We prove by induction on $r$, keeping in mind the results of Lemmas
-  \ref{lem:W_subset_E} and \ref{lem:depth_lemma}.  For the base case,
-  let $r = 1$ and suppose that $(i, j) \in \gcge$ with $d(j) = k$ and
-  $d(i) = k - 1$.  Then, $(i, j) \in W$ and by our assumptions on
-  $E_{k - 1}$ there is no $\gcgpath{i}{j}$ path in $E_{k - 1}$
-  and therefore $(i, j) \in D_{k1}$.  Conversely, suppose that
-  $(i, j) \in D_{k1}$.  Then, $d(i) = k - 1$ and $d(j) = k$ which, since
-  $(i, j) \in W \implies i \in \anc{j}$ implies that
-  $i \in \pa{j}$ and $(i, j) \in \gcge$.
-
-  Now, fix $r > 1$ and suppose that the result holds up to $r - 1$.
-  Let $(i, j) \in \gcge$ with $d(j) = k$ and $d(i) = k - r$.  Then,
-  $(i, j) \in W$ and by induction and strong causality there cannot
-  already be an $\gcgpath{i}{j}$ path in
-  $E_{k - 1} \cup \big(\bigcup_{\ell = 0}^{r - 1} D_{kr}\big)$,
-  therefore $(i, j) \in D_{kr}$.  Conversely, suppose
-  $(i, j) \in D_{kr}$.  Then we have $d(i) = k - r$, $d(j) = k$, and
-  $i \in \anc{j}$.  Suppose by way of contradiction that
-  $i \not\in \pa{j}$, then there must be some $u \in \pa{j}$ such that
-  $i \in \anc{u}$.  But, this implies that $d(i) < d(u)$ and by
-  induction that $(u, j) \in \bigcup_{\ell = 1}^{r - 1}D_{k\ell}$.
-  Moreover, since $d(u) < k$ (otherwise $d(j) > k$) each edge in
-  the $\gcgpath{i}{u}$ path must already be in $E_{k - 1}$, and so
-  there must be an $\gcgpath{i}{j}$ path in
-  $E_{k - 1}\cup\big(\bigcup_{\ell = 0}^{r - 1}D_{kr}\big)$, which is
-  a contradiction since we assumed $(i, j) \in D_{kr}$.  Therefore
-  $i \in \pa{j}$ and $(i, j) \in \gcge$.
-\end{proof}
-
-\begin{lemma}[Outer Loop]
-  \label{lem:outer_loop_lemma}
-  We have $(i, j) \in E_k$ if and only if $(i, j) \in \gcge$ and
-  $d(j) \le k$.  That is, at iteration $k, E_k$ and $\gcge$ agree on
-  the set of edges whose terminating node is at most $k$ steps away
-  from $P_0$.
-\end{lemma}
-\begin{proof}
-  We will proceed by induction.  The base case $E_0 = \emptyset$ is
-  trivial, so fix some $k \ge 1$, and suppose that the lemma holds for
-  all nodes of depth less than $k$.
-
-  Suppose that
-  $(i, j) \in E_k = E_{k - 1}\cup \big(\bigcup_{r = 1}^k D_{rk}
-  \big)$.  Then clearly there is some $1 \le r \le k$ such that
-  $(i, j) \in D_{kr}$ so that by Lemma \ref{lem:inner_loop_lemma} we
-  have $(i, j) \in \gcge$ and $d(j) = k$.
-
-  Conversely, suppose that $(i, j) \in \gcge$ and $d(j) \le k$.  If
-  $d(j) < k$ then by induction $(i, j) \in E_{k - 1} \subseteq E_k$ so
-  suppose further than $d(j) = k$.  Since $i \in \pa{j}$ we must have
-  $d(i) < k$ (else $d(j) > k$) and again by Lemma
-  \ref{lem:inner_loop_lemma} $(i, j) \in \bigcup_{r = 1}^k D_{kr}$
-  which implies that $(i, j) \in E_k$.  The result follows.
-\end{proof}
-
-\begin{lemma}[Finite Termination]
-  Algorithm \ref{alg:pwgr} terminates and returns $E_{k^\star - 1} = \gcge$
-  for some $k^\star \le n$.
-\end{lemma}
-\begin{proof}
-  If $n = 1$, the algorithm is clearly correct, returning on the first
-  iteration with $E_1 = \emptyset$.  When $n > 1$ Lemma
-  \ref{lem:outer_loop_lemma} ensures that $E_k$ coincides with
-  $\{(i, j) \in \gcge\ |\ d(j) \le k\}$ and since $d(j) \le n - 1$ for
-  any $j \in [n]$ there is some $k^\star \le n$ such that
-  $E_{k^\star - 1} = \gcge$.  We must have $S_{k^\star} = \emptyset$
-  since $j \in S_{k^\star} \iff d(j) \ge k^\star$ (if $d(j) > k - 1$ then
-  $E_{k^\star - 1} \ne \gcge$) and therefore the algorithm terminates.
-\end{proof}
-
-\begin{example}
-  We close this section by noting that the conditions of persistence
-  and strong causality are only sufficient conditions.  For example,
-  the complete directed graph with 2 nodes i.e.
+  on a long stream of out-of-sample data.  We then report the quantity
 
   \begin{equation*}
-    B(1) = \left[ \begin{array}{cc} 1/2 & 1 \\ 1 & 1/2 \end{array}\right]
+    \frac{\ln \tr \widehat{\Sigma}_v}{\ln \tr \Sigma_v}
   \end{equation*}
 
-  contains a loop but is pairwise recoverable, though not by algorithm
-  (\ref{alg:pwgr}).  Clearly, this example is somewhat artificial
-  since when $n = 2$ there is no difference between pairwise
-  Granger-causality and joint Granger-causality amongst all series --
-  however, one can add any number of nodes having no parents or
-  children to a graph containing a length 2 cycle, in which case the
-  graph clearly remains pairwise recoverable.
-\end{example}
+  where $\widehat{\Sigma}_v = \Sigma_v$ is the best possible performance.
+\end{remark}
+
+\subsection{Results}
+\begin{figure}[!h]
+  \centering
+  \caption{Simulation Results: PWGC vs AdaLASSO}
+  \label{tab:simulation_table}
+
+  \begin{tabular}{|ll||ll|ll|ll|}
+    \toprule
+    &\textbf{Algorithm}&alasso&pwgc&alasso&pwgc&alasso&pwgc\\
+    &\textbf{Metric}&LRE&&FDP&&MCC&\\
+    \textbf{T}&\textbf{q}&&&&&&\\
+    \midrule
+    \multirow{4}{*}{\textbf{50}}
+    &\textbf{SCG}&1.83&\textbf{1.57}&0.48&\textbf{0.07}&0.47&\textbf{0.57}\\
+    &\textbf{0.06}&1.91&\textbf{1.69}&0.54&\textbf{0.09}&0.43&\textbf{0.53}\\
+    &\textbf{0.11}&2.89&\textbf{2.58}&0.46&\textbf{0.22}&0.38&0.40\\
+    &\textbf{0.46}&8.49&\textbf{7.72}&0.42&0.42&\textbf{0.16}&0.11\\
+    \midrule
+    \multirow{4}{*}{\textbf{250}}
+    &\textbf{SCG}&1.32&\textbf{1.22}&0.27&\textbf{0.07}&0.71&\textbf{0.80}\\
+    &\textbf{0.06}&1.44&\textbf{1.28}&0.32&\textbf{0.09}&0.66&\textbf{0.76}\\
+    &\textbf{0.11}&2.57&\textbf{2.17}&0.32&\textbf{0.16}&0.53&\textbf{0.57}\\
+    &\textbf{0.46}&8.09&\textbf{7.11}&0.39&0.40&\textbf{0.20}&0.14\\
+    \midrule
+    \multirow{4}{*}{\textbf{1250}}
+    &\textbf{SCG}&1.18&\textbf{1.09}&0.40&\textbf{0.06}&0.69&\textbf{0.89}\\
+    &\textbf{0.06}&1.24&1.17&0.43&\textbf{0.05}&0.65&\textbf{0.86}\\
+    &\textbf{0.11}&2.02&2.10&0.33&\textbf{0.14}&0.62&0.63\\
+    &\textbf{0.46}&7.44&\textbf{7.04}&0.40&\textbf{0.30}&\textbf{0.22}&0.19\\
+    \bottomrule
+  \end{tabular}
+
+  {\footnotesize Results of Monte Carlo simulations comparing PWGC and
+    AdaLASSO $(n = 35, p = 5, p_{\text{max}} = 10)$ for small
+    samples and when the SCG assumption doesn't hold.  The superior
+    result is bolded when the difference is statistically
+    significant, as measured by \texttt{scipy.stats.ttest\_rel}.
+    100 iterations are run for each
+    set of parameters.\\
+
+    LRE: Log-Relative-Error, i.e. the log sum of squared errors at
+    each node relative to the strength of the driving noise
+    $\frac{\ln \tr \widehat{\Sigma}_v}{\ln \tr \Sigma_v}$.  FDP: False
+    Discovery Proportion.  MCC: Matthew's Correlation Coefficient.\\
+
+    Values of $q$ (edge probability) range between
+    $2 / n, 4 / n, 16 / n$ where $2 / n$ has the property that the
+    random graphs have on average the same number of edges as the
+    SCG.}
+\end{figure}
+
+Our simulation results are summarized in \ref{tab:simulation_table},
+with additional figures provided in the supplementary material Section
+\ref{apx:simulation}.  It is clear that the superior performance of
+PWGC in comparison to AdaLASSO is as a result of limiting the false
+discovery rate.  It is unsurprising that PWGC exhibits superior
+performance when the graph is an SCG, but even in the case of more
+general DAGs, the PWGC heuristic is still able to more reliably
+uncover the graph structure for small values of $q$.  As $q$
+increases, there are simply not enough edges allowed by the SGC
+topology for it to be possible to accurately recover $\gcg$.
+
+Interestingly, we can observe that the AdaLASSO appears to perorm
+marginally better on strongly causal graphs than directed acyclic
+graphs with an equivalent number of edges ($q \approx 2 / n$ is chosen for
+this purpose).  This provides supporting evidence for one of the main
+assertions of this work: that the topological structure of $\gcg$ is
+an important distinguishing feature of time series networks in
+comparison to classical multivariate regression where it is only the
+sparsity rate which is considered.
 
@@
 insights into the structure of causality graphs with potential for
-improvements in causality graph estimation when structural assumptions
-are met.  In particular, the notion of a strongly-causal graph has
-been exploited to establish conditions under which pairwise causality
-testing alone is sufficient for recovering a complete
+providing improvements to causality graph estimation when structural
+assumptions are met.  In particular, the notion of a strongly-causal
+graph has been exploited to establish conditions under which pairwise
+causality testing alone is sufficient for recovering a complete
 Granger-causality graph.  Moreover, examples from the literature
 suggest that such topological assumptions may be reasonable in some
-applications.
+applications.  And secondly, even when the strong-causality assumption
+is not met, we have provided simulation evidence to suggest that our
+pairwise testing algorithm PWGC can still outperform the LASSO and
+adaLASSO, both of which are commonly employed in applications.
 
@@
 case of Gaussian data \cite{barnett2009granger}, do our results extend
-to causality networks defined via transfer entropy?
-
-Due to constraints on space, we report upon details for practical
-finite sample implementation as well as more extensive simulation
-results and an application to EEG data elsewhere \todo{where?}.
+to causality networks defined via transfer entropy?  Finally, the work
+of \cite{barnett2015granger} has established the superiority of
+Granger-causality testing in state space models (as opposed to pure
+autoregressions) in many cases.  Combining this work with our PWGC
+algorithm (by modifying the approach described in Section
+\ref{sec:pairwise_hypothesis_testing} to instead utilize state-space
+Granger-causality testing) therefore is likely to enable application
+to very large networks of time series data which are not well
+approximated by finite $\VAR(p)$ models.
 

### 2019-07-17 paper_theory/proofs.tex
I can have a supplementary document with proofs and extra figures.

--- 2019-07-17 paper/paper.tex
@@
-\documentclass[12pt]{article}
-
-% TODO: Are we, in this setting, immune to confounders?  We are removing
-%       all bi-directional edges, so it seems likely that we are?  Can we state
-%       any theorem about this?
-
-%% bibliography stuff -- this needs come before the preamble inclusion
-\usepackage[backend=bibtex,sorting=none]{biblatex}
+\documentclass{statsoc}
+\usepackage{fullpage}
+\usepackage{ccfonts,eulervm}
+\usepackage{tabularx}
+\usepackage{xr}
+\usepackage{chngcntr}
+\usepackage{apptools}
+\AtAppendix{\counterwithin{lemma}{section}}
+\AtAppendix{\counterwithin{definition}{section}}
+\externaldocument{paper_theory}
+
+\usepackage{geometry}
+
+\geometry{
+  textwidth=33pc,
+  textheight=\dimexpr48\baselineskip+\topskip\relax,
+  marginparsep=11pt,
+  marginparwidth=107pt,
+  footnotesep=6.65pt,
+  headheight=9pt,
+  headsep=9pt,
+  footskip=30pt,
+}
+\bibliographystyle{chicago}
 \usepackage{enumitem}
@@
 \usepackage{todonotes}
-\bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
-% \usepackage{apxproof}
-
-% \bibliography{\string~global_bib.bib}
+\usepackage{multirow}
+\usepackage{booktabs}
 
@@
 
-\input{\string~/Documents/academics/global_academics/latex_preamble}
-% \input{\string~latex_preamble}
+% \usepackage{fullpage}
+\usepackage{framed}
+
+% Figures
+\usepackage{graphicx}
+\usepackage{caption}  % This is not recommended?
+\usepackage{subcaption}
+% \usepackage{wrapfig}
+% \usepackage{svg}
+
+% Math packages, theorem definitions and numbering
+\usepackage{amsmath}
+\usepackage{amssymb}
+% \usepackage{amsthm}
+% \usepackage{mathrsfs} % Fancy scripted font
+\usepackage{bm}  % Bold math
+
+% Misc packages
+\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
+\usepackage{algorithm2e} %{algorithm} environment
+\usepackage{soul}  % \hl highlighting
+\usepackage{color}
+\usepackage{mathtools}  % For my \ceil function
+
+% Theorems (with italics)
+% \theoremstyle{plain}  % Style definition removes italics
+\newtheorem{theorem}{Theorem}
+\newtheorem{corollary}{Corollary}
+\newtheorem{proposition}{Proposition}
+\newtheorem{lemma}{Lemma}
+
+% \theoremstyle{definition}
+\newtheorem{remark}{Remark}
+\newtheorem{definition}{Definition}
+\newtheorem{example}{Example}
+\newtheorem{assumptions}{Assumptions}
+
+% keywords
+\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}
+
+% General
+\def\defeq{\overset{\Delta}{=}}  % Equal with triangle
+\def\cl{\mathsf{cl\ }}  % Closure
+\newcommand{\sgn}[1]{\mathsf{sgn}(#1)}  % sign function
+
+% Calculus
+\def\d{\mathsf{d}}  % Differential operator
+
+% Functions
+\def\ln{\mathsf{ln\ }}  % Natural logarithm
+\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}  % Ceiling
+
+% Probability
+\def\H{\mathcal{H}}  % Hilbert space
+\def\E{\mathbb{E}}  % Expectation
+\def\Var{\text{Var}}  % Variance
+\def\P{\mathbb{P}}  % Probability Measure
+\def\F{\mathcal{F}}  % A sigma algebra
+\def\sX{\mathcal{X}}  % Another sigma algebra
+\def\KL{\mathbf{D}_{KL}}  % KL divergence
+\def\bF{\mathbf{F}}  % Whole F-meas space
+
+% Standard sets
+\def\Z{\mathbb{Z}}  % Set of integers
+\def\R{\mathbb{R}}  % Set of real numbers
+\def\C{\mathbb{C}}  % Set of complex numbers
+\def\N{\mathbb{N}}  % Set of natural numbers
+\def\ball{\mathbb{B}}  % Open ball
+\def\clball{\overline{\ball}}  % Closed ball
+
+% Linear algebra
+\def\rk{\mathsf{rk }}  % The rank
+\def\tr{\mathsf{tr }}  % The trace
+\def\T{\mathsf{T}}  % Transpose notation
+\def\c{\mathsf{c}}  % complement
+\def\dg{\mathsf{dg }}   %  Diagonal vector of a matrix
+\def\Dg{\mathsf{Dg }}   %  Diagonal matrix from a vector
+\def\ind{\mathbf{1}}  % Ones vector or indicator
+\def\matvec{\textbf{vec}}  % Vector operator
+\def\<{\langle}  % < Inner product
+\def\>{\rangle}  % > Inner product
+\newcommand{\inner}[2]{\langle #1, #2 \rangle}  % Inner product
+\newcommand{\innerT}[2]{#1^\T #2}  % Inner product for finite vectors
 
@@
 
-\title{Graph Topological Aspects of Granger Causal Network Learning }
-\author{R. J. Kinnear \\
-  \small\href{mailto:ryan@kinnear.ca}{ryan@kinnear.ca} \\
-  \small\url{https://github.com/RJTK} \and R. R. Mazumdar \\
-  \small\href{mailto:mazum@uwaterloo.ca}{mazum@uwaterloo.ca} }
+\title{Graph Topological Aspects of Granger Causal Network Learning\\
+  \large Supplementary Material}
+
+\author[Author 1 {\it et al.}]{R. J. Kinnear}
+\address{
+  University of Waterloo,
+  Waterloo,
+  Canada.}
+\email{Ryan@Kinnear.ca}
+
+\author{R. R. Mazumdar}
+\address{
+  University of Waterloo,
+  Waterloo,
+  Canada.}
 
 \begin{document}
-\maketitle
-\abstract{We study Granger-causality in the context of wide-sense
-  stationary time series, where our focus is on the topological
-  aspects of the underlying causality graph.  We establish sufficient
-  conditions (in particular, we develop the notion of a ``strongly
-  causal'' graph topology) under which the true causality graph can be
-  recovered via pairwise causality testing alone, and provide examples
-  from the gene regulatory network literature suggesting that our
-  concept of a strongly causal graph may be applicable to this field.
-  We implement and detail finite-sample heuristics derived from our
-  theory, and establish through simulation the efficiency gains (both
-  statistical and computational) which can be obtained (in comparison
-  to LASSO algorithms) when structural assumptions are met.  Finally,
-  we provide an example application where we accurately discriminate
-  between subjects in an EEG study based purely on the
-  Granger-causality graphs inferred from data by our algorithms,
-  demonstrating that meaningful features are captured by the
-  Granger-causal graph topology.}
-
-\keywords{causality graph, EEG, gene regulatory networks,
-  Granger-causality, LASSO, network learning, time series, vector
-  autoregression}
-
-\paragraph{Acknowledgement}
-
-We acknowledge the support of the Natural Sciences and Engineering Research Council of Canada (NSERC), [funding reference number 518418-2018].
-Cette recherche a été financée par le Conseil de recherches en sciences naturelles et en génie du Canada (CRSNG), [numéro de référence 518418-2018].
-
-\clearpage
-
-% \tableofcontents
-% \clearpage
-
-\section{Introduction and Review}
-\label{sec:introduction}
-% Consider a collection of stochastic processes producing observations
-% at discrete time intervals.  Are the underlying processes dependent?
-% Can we quantify any of the underlying relationships?  Can the arrow of
-% time help us to distinguish a directionality or flow of dependence
-% among our observed series?  In this paper we contribute to the
-% understanding of the notion of Granger-Causality
-% \cite{granger1969investigating} \cite{Granger1980329} as a tool for
-% answering these questions.
-
-% Heuristically, we state that if an event '$A$' occuring strictly prior
-% in time to an even '$B$' and $A$ provides us with exclusive (that is,
-% not available anywhere else) information about $B$, then $A$ must have
-% had a causal impact on $B$.  This is an entirely model-free notion of
-% cause, and instead leverages the intuition that a cause must precede
-% it's effect.
-
-In this paper we study the notion of Granger-causality
-\cite{granger1969investigating} \cite{Granger1980329} as a means of
-uncovering an underlying causal structure in multivariate time series.
-Though the underlying causality graph cannot be observed directly, we
-will infer it's presence as a latent structure among our observed time
-series data.  This notion is leveraged in a variety of applications
-e.g. in Neuroscience as a means of recovering interactions amongst
-brain regions \cite{bressler2011wiener}, \cite{anna_paper2008},
-\cite{david2008identifying}; in the study of the dependence and
-connectedness of financial institutions \cite{NBERw16223}; gene
-expression networks \cite{Fujita2007},
-\cite{methods_for_inferring_gene_regulatory_networks_from_time_series_expression_data},
-\cite{grouped_graphical_granger_modelling_for_gene_expression_regulatory_networks_discovery},
-\cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty};
-and power system design \cite{Misyrlis2016450}, \cite{yuan2014root}.
-
-Granger-causality can generally be formulated by searching for the
-``best'' graph structure consistent with observed data, which is in
-general an extremely challenging problem (i.e. it may be framed as a
-best subset selection problem, see \cite{bss_mio},
-\cite{hastie_bss_comp}), moreover, the comparison of quality between
-different structures, and hence the notion of ``best'' needs
-qualification.  In applications where we are interested merely in
-minimizing the mean squared error of a linear one-step-ahead
-predictor, then we will be satisfied with an entirely dense graph of
-connections, since each edge can only serve to reduce estimation
-error.  However, since the number of edges scales quadratically in $n$
-(the number of nodes) it becomes imperative to infer a sparse
-causality graph for large systems, both to avoid overfitting observed
-data, as well as to aid the interpretability of the results.
-
-A fairly early approach to the problem in the context of large systems
-is provided by \cite{bach2004learning}, where the authors apply a
-local search heuristic to the Whittle likelihood with an AIC
-penalization.  The local search heuristic where at each iteration an
-edge is either added, removed, or reversed is a common approach to
-combinatorial optimization due to it's simplicity, but is liable to
-get stuck in shallow local minima.
-
-A second and wildly successful heuristic is the LASSO regularizer
-\cite{tibshirani1996regression}, which can be understood as a natural
-convex relaxation to penalizing the count of the non-zero edges.  The
-LASSO enjoys fairly strong theoretical guarantees
-\cite{wainwright2009sharp}, extending largely to the case of
-stationary time series data with a sufficiently fast rate of
-dependence decay \cite{basu2015} \cite{wong2016lasso}
-\cite{autoregressive_process_modelling_via_the_lasso_procedure}, and
-variations on the LASSO have been applied in a number of different
-time series contexts as well as Granger-causality
-\cite{DBLP:journals/corr/HallacPBL17} \cite{haufe2008sparse}
-\cite{bolstad2011causal} \cite{he2013stationary}
-\cite{grouped_graphical_granger_modelling_for_gene_expression_regulatory_networks_discovery}.
-One of the key improvements to the original LASSO algorithm is the
-adaptive (i.e. weighted) ``adaLASSO'' \cite{adaptive_lasso_zou2006},
-for which oracle results (i.e. asymptotic support recovery) are
-established under less restrictive conditions than for the vanilla
-LASSO.  Our experimental comparisons in Section
-\ref{sec:empirical_evaluation} are against the
-adaLASSO. % Success of the LASSO has lead to
-% the famous ``bet on sparsity'' principle ``Use a procedure that does
-% well in sparse problems, since no procedure does well in dense
-% problems.''  \cite{tibshirani2015statistical}.
-
-\subsection{Contributions}
-In the context of time series data, sparsity assumptions remain
-important, but there is significant additional structure that may
-arise as a result of considering the topology of the underlying
-Granger-causality graph, which to our knowledge remains largely
-unexplored.  The focus of this paper is to shed light on some of these
-topological questions, in particular, we study a particularly simple
-notion of causality graph topology which we term ``strongly causal''
-and show that stationary times series whose underlying causality graph
-has this structure satisfy natural intuitive notions of ``information
-flow'' through the graph.  Moreover, we show that such graphs are
-perfectly recoverable with only \textit{pairwise} Granger-causality
-tests, which would otherwise suffer from serious confounding problems.
-Our finite sample results are based on simulations, where we show
-promising results for these graph topologies where our algorithm
-performs substantially better in our simulation setup than do
-competing LASSO algorithms, even for graphs that do not exactly
-satisfy our strongly-causal topology assumptions.
-
-In the case of gene expression networks, we show examples from the
-literature which suggest our concept of a ``strongly causal graph''
-topology may have application in this field (see Section
-\ref{sec:strongly_causal_graphs}).
-
-The principle contributions of this paper are as follows: firstly, in
-section \ref{sec:theory} we study \textit{pairwise} Granger-causality
-relations, providing novel theorems connecting the structure of the
-causality graph to the pairwise ``causality flow'' in the system, as
-well as an interpretation in terms of the graph topology of the
-sparsity pattern of matrices arising in the Wold decomposition,
-generalizing in some sense the notion of ``feedback-free'' processes
-studied by \cite{caines1975feedback} in close connection with
-Granger-causality.  We establish sufficient conditions (sections
-\ref{sec:strongly_causal_graphs}, \ref{sec:persistent_systems}) under
-which a fully conditional Granger-causality graph can be recovered
-from pairwise tests alone (sec \ref{sec:pairwise_algorithm}).
-Secondly, we propose in section \ref{sec:structure_learning} a graph
-search heuristic which implements our theoretical results to finite
-data samples, specifying and summarizing appropriate methods for
-hypothesis testing (section \ref{sec:pairwise_hypothesis_testing}),
-model order selection (section \ref{sec:model_order_selection}),
-computationally efficient estimation (section
-\ref{sec:efficient_model_estimation}), and error rate controls
-(section \ref{sec:error_rate_control}).  Our heuristics are compared
-against the adaLASSO algorithm in Section
-\ref{sec:empirical_evaluation}.  We stress the scalability of our
-algorithm which is capable of comfortably handling hundreds or
-thousands of nodes on a single machine, as opposed to standard LASSO
-algorithms which do not take advantage of the special structure
-associated with stationary time series data.  In section
-\ref{sec:application} we develop an example application where a
-classifier is constructed to accurately discriminate between subjects
-in an EEG study based only on the Granger-causality graph topology
-inferred by our algorithms.  Concluding remarks on further open
-problems and extentions are provided in Section \ref{sec:conclusion}.
-
-\section{Theory}
-\label{sec:theory}
-\subsection{Formal Setting}
-Consider the space $L_2(\Omega)$, the usual Hilbert space of finite
-variance random variables over a probability space
-$(\Omega, \mathcal{F}, \mathbb{P})$ having inner product
-$\inner{x}{y} = \E[xy]$.  We will work with a discrete time and
-wide-sense stationary (WSS) $n$-dimensional vector valued process
-$x(t)$ (with $t \in \Z$) where the $n$ elements take values in $L_2$.  We
-suppose that $x(t)$ has zero mean, $\E x(t) = 0$, and has absolutely
-summable matrix valued covariance sequence
-$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ with an absolutely
-continuous spectral density $S(\omega)$:
-
-\begin{equation}
-  \label{eqn:fourier_pair}
-  \begin{aligned}
-    R(\tau) &= \frac{1}{2\pi}\int_{-\pi}^\pi S(\omega) e^{j\tau\omega}\d \omega,\\
-    S(\omega) &= \sum_{\tau=-\infty}^\infty R(\tau)e^{-j\tau\omega}.
-  \end{aligned}
-\end{equation}
-
-% and that the spectra is bounded uniformly away from 0:
-
-% \begin{equation*}
-%   \int_{-\pi}^\pi \ln\det S(\omega) \d\omega > -\infty.
-% \end{equation*}
-
-We will also work frequently with the spaces spanned by the values of
-such a process
-
-\begin{equation}
-  \label{eq:hilbert_space_defn}
-  \begin{aligned}
-    % \H^x &= \cl \{\sum_{\tau = -T}^T a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, T \in \N\} \subseteq L_2(\Omega),\\
-    \H_t^x &= \cl \{\sum_{\tau = 0}^p a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, p \in \N\} \subseteq L_2(\Omega)\\
-    H_t^x &= \{a x(t)\ |\ a \in \R\} \subseteq L_2(\Omega),
-  \end{aligned}
-\end{equation}
-
-where the closure is naturally in mean-square.  We will often omit the
-superscript $x$ which should be clear from context.  Evidently these
-spaces are separable, and as closed subspaces of a Hilbert space they
-are themselves Hilbert.  We will denote the spaces generated in
-analogous ways by particular components of $x$ as e.g.
-$\H_t^{(i, j)}$, $\H_t^{i}$ or by all but particular components as
-$\H_t^{-j}$.
-
-As a consequence of the Wold decomposition theorem \cite{lindquist},
-every WSS sequence has the moving average $MA(\infty)$
-representation
-
-\begin{equation}
-\label{eqn:wold}
-  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) v(t - \tau),
-\end{equation}
-
-where $c(t)$ is a purely deterministic sequence\footnote{the purely
-  deterministic sequence $c(t)$ is one which lies in the remote past
-  $\bigcap_{\tau=1}^\infty \H_{t - \tau}^x$ of the process.  For such
-  processes a single sample $c(t_0)$ is enough to determine $c(t)$ for
-  every $t$.  For example,
-  $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi,
-  \pi]$}, $v(t)$ is an uncorrelated sequence and $A(0) = I$.  We will
-assume that $c(t) = 0$, which in practice is to say that the process
-has been detrended.  We additionally require that this representation
-can be inverted to yield the $\VAR(\infty)$ form
-
-\begin{equation}
-  \label{eqn:ar_representation}
-  x(t) = \sum_{\tau = 1}^\infty B(\tau) x(t - \tau) + v(t).
-\end{equation}
-
-The equations (\ref{eqn:wold}), (\ref{eqn:ar_representation}) can be
-represented as $x(t) = \A(z)v(t) = \B(z)x(t) + v(t)$ via the action
-(convolution) of the operators (LTI filters)
-$\A(z) \defeq \sum_{\tau = 0}^\infty A(\tau)z^{-\tau}$ and
-$\B(z) \defeq \sum_{\tau = 1}^\infty B(\tau)z^{-\tau}$ where the
-operator $z^{-1}$ is the back shift operator acting on
-$\ell_2^n(\Omega, \mathcal{F}, \mathbb{P})$, that is:
-
-\begin{equation}
-  \label{eqn:filter_action}
-  \B_{ij}(z)x_j(t) \defeq \sum_{\tau = 1}^\infty B_{ij}(\tau)x_j(t - \tau).
-\end{equation}
-
-Finally, since $||z^{-1}|| = 1$ we have the inversion formula
-
-\begin{equation}
-  \label{eqn:lsi_inversion}
-  \A(z) = (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k.
-\end{equation}
-
-The aforementioned assumptions are quite weak.  The strongest
-assumption we require is finally that $\Sigma_v$ is a diagonal matrix,
-which is referred to as a lack of instantaneous feedback in $x(t)$.
-We formally state our setup to conclude this section:
-
-\begin{definition}[Basic Setup]
-  \label{def:basic_setup}
-  The process $x(t)$ is an $n$ dimensional wide sense stationary
-  process having invertible $\VAR(\infty)$ representation
-  \eqref{eqn:ar_representation} where $v(t)$ is sequentially
-  uncorrelated and has a diagonal covariance matrix.  The $MA(\infty)$
-  representation of equation \eqref{eqn:wold} has $c(t) = 0$ and
-  $A(0) = I$.
-\end{definition}
-
-\subsection{Granger Causality}
-
-\begin{definition}[Granger Causality]
-  \label{def:granger_causality}
-  For the WSS series $x(t)$ satisfying the assumptions of Definition
-  \ref{def:basic_setup} we will say that component $x_j$
-  \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x$)
-  and write $x_j \gc x_i$ if given Hilbert spaces $\H_{t - 1}$,
-  $\H^{-j}_{t - 1}$
-
-\begin{equation}
-  \linEerr{x_i(t)}{\H_{t - 1}} < \linEerr{x_i(t)}{\H^{-j}_{t - 1}},
-\end{equation}
-
-where $\xi[x \ |\ \H] = \E (x - \linE{x}{\H})^2$ is the mean squared
-estimation error and $\linE{x}{\H} = \text{proj}_{\H}(x)$ denotes the
-(unique) projection onto the Hilbert space $\H$.
-\end{definition}
-
-This notion captures the idea that the process $x_j$ provides
-information about $x_i$ that is not available from elsewhere.  The
-caveat ``with respect to $x$'' is important in that GC relations can
-change when components are added to or removed from our collection $x$
-of observations, e.g. new GC relations can arise if we remove the
-observations of a common cause, and existing GC relations can
-disappear if we observe a new mediating series.
-
-The notion is closely related to the information theoretic measure of
-transfer entropy, indeed, if the distribution of $v(t)$ is known to be
-Gaussian then they are equivalent \cite{barnett2009granger}.
-
-We require two technical facts we refer to in the sequel, the notion
-of conditional orthogonality is used throughout.
-
-\begin{lemma}[\cite{lindquist} Proposition 2.4.2]
-  \label{lem:conditional_orthogonality_equivalence}
-  Consider three closed subspaces of a Hilbert space $\mathcal{A}$,
-  $\mathcal{B}$, $\mathcal{X}$.  The following statements are
-  equivalent
-
-  \begin{enumerate}
-    \item{$\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$}
-    \item{$\linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}}\ \forall \beta \in \mathcal{B}$.}
-    % \item{$\linE{\beta}{\mathcal{A}} = \linE{\linE{\beta}{\mathcal{X}}}{\mathcal{A}}\ \forall \beta \in \mathcal{B}}$}
-    \end{enumerate}
-
-    Where $\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$ denotes
-    conditional orthogonality:
-
-    \begin{equation*}
-      \inner{a - \linE{a}{\mathcal{X}}}{b - \linE{b}{\mathcal{X}}} = 0\ \forall a \in \mathcal{A}, b \in \mathcal{B}.
-    \end{equation*}
-\end{lemma}
-% \begin{proof}
-%   Proceding from the definition of conditional orthogonality we have
-
-%   \begin{align*}
-%     &\inner{\alpha - \linE{\alpha}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
-%     &{\iff} \inner{\alpha + x - \linE{\alpha + x}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
-%     &\overset{(a)}{\iff} \inner{\alpha + x}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
-%     &\overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A} \vee \mathcal{X}} = 0\ \forall \beta \in \mathcal{B}\\
-%     &\overset{(c)}{\iff} \linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}},
-%   \end{align*}
-
-%   where $(a)$ follows since $(\beta - \linE{\beta}{\mathcal{X}}) \perp \mathcal{X}$ and $\linE{\alpha + x}{\mathcal{X}} \in \mathcal{X}$, $(b)$ since $z \perp \mathcal{X} \vee \mathcal{A} \iff \linE{z}{\mathcal{X} \vee \mathcal{A}} = 0$, and $(d)$ since $\linE{z}{\H_2} = \linE{\linE{z}{\H_2}}{\H_1}$ for $\H_2 \subseteq \H_1$.
-
-%   % This is property (vi) in Lindquist
-%   % \begin{align*}
-%   %   \langle \alpha - \linE{\alpha}{\mathcal{X}}, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
-%   %   \overset{(a)}{\iff} \langle \alpha, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0\ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
-%   %   \overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A}} &= 0\ \forall \beta \in \mathcal{B}
-%   % \end{align*}
-
-%   % where $(a)$ follows because $\linE{\alpha}{\mathcal{X}} \in \mathcal{X}$ and $\beta - \linE{\beta}{\mathcal{X}} \in \mathcal{X}$, and $(b)$ follows since $\beta \perp \mathcal{A} \iff \linE{\beta}{\mathcal{A}} = 0$.
-% \end{proof}
-\begin{lemma}
-  \label{lem:subspace_sum_projection}
-  Let $x \in \H$ where $\H$ is a separable Hilbert space having inner product
-  $\inner{\cdot}{\cdot}$ and let $\{\H_n\}_{n = 0}^N$ be a
-  collection of subspaces of $\H$ such that $x \perp \H_0$.  Then
-
-  \begin{equation*}
-    \linE{x}{\bigvee_{n = 0}^N\H_n} = \linE{x}{\bigvee_{n = 1}^N\H_n},
-  \end{equation*}
-
-  where
-  $\H_1 \vee \H_2 = \cl \{\alpha + \beta\ |\ \alpha \in \H_1, \beta
-  \in \H_2 \}$ is the closed sum of subspaces.
-\end{lemma}
-\begin{proof}
-  Apply the Gram-Schmidt process and directly calculate the projections.
-\end{proof}
-
-\begin{theorem}[Granger Causality Equivalences]
-  \label{thm:granger_causality_equivalences}
-  Let $x(t)$ be a WSS process with absolutely summable covariance
-  sequence, and spectral density uniformly bounded above and below.
-  Denote by $\xi_{ij} \defeq \linEerr{x_i(t)}{\H^{-j}}$ and
-  $\xi_i \defeq \linEerr{x_i(t)}{\H_t}$ for distinct $i, j$.  Then,
-  the following are equivalent:
+
+\appendix
+
+\section{Overview}
+We restate our main results and provide detailed proofs.  Simple
+Corollaries have their proofs in the main text, and are occasionally
+referenced here.  The main Theorem is proven in Section
+\ref{apx:proof_main_theorem}, and all of the building blocks are
+established in Section \ref{apx:ancillary_results}.
+
+We detail the methods used for our simulations and finite sample
+implementation in Section \ref{sec:structure_learning} and provide
+additional simulation results in Section \ref{apx:simulation}.
+
+Code will be made available at
+\url{https://github.com/RJTK/granger_causality}, as well as
+accompanying this supplementary material.
+
+\section{Proofs}
+\subsection{Ancillary Results}
+\label{apx:ancillary_results}
+\begin{theorem}[Granger Causality Equivalences \ref{thm:granger_causality_equivalences}]
+  The following are equivalent:
 
@@
   \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$ i.e. $\B_{ij}(z) = 0$}
-    \item{$F_{ij} \defeq \big(\frac{\xi_i}{\xi_{ij}} - 1\big) = 0$}
-    \item{$\H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j} \iff H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
+  \item{$H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
   \item{$\linE{x_i(t)}{\H_{t - 1}^{-j}} = \linE{x_i(t)}{\H_{t - 1}}$}
-    % \item{\hl{Wold $A(\tau)$ condition}}
-    % \item{Tie in the 0s of $S(\omega)^{-1}$, though this isn't immediately directed.}
   \end{enumerate}
@@
 
-  $(1) \iff (3)$ is a tautology, and $(3) \iff (5)$ is immediate from the
-  definitions i.e. the projections are equal if and only if the errors
-  are equal.  We have $(4) \iff (5)$ as a result of Lemma
-  \ref{lem:conditional_orthogonality_equivalence}
-
-% , the additional
-%   equivalence in $(4)$ is apparent since $\H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$.
-
-  $(1) \Rightarrow (2)$: The projection for $x_i(t)$ onto $\H_{t - 1}$ is
-  (by definition) given by a form similar to equation
-  \ref{eqn:ar_representation}:
-
-\[
-  \E|x_i(t) - \linE{x_i(t)}{\H_{t - 1}}|^2 = \E|v_i(t) + \sum_{\tau = 1}^\infty \sum_{k = 1}^n (B_{i, k}(\tau) - \hat{B}_{i, k}(\tau))x_k(t - \tau)|^2.
-\]
-
-Since $v(t)$ in equation \ref{eqn:ar_representation} is temporally
-uncorrelated, it follows that the optimal projection is given by the
-model coefficients themselves.  This holds similarly for the
-projection onto $\H_{t - 1}^{-j}$.  Then by $(1)$ we have
-
-\[
-  \E|x_i(t) - \sum_{\tau = 1}^\infty \sum_{k = 1}^n B_{i, k}(\tau)x_k(t - \tau)|^2 = \E|x_i(t) - \sum_{\tau = 1}^\infty\sum_{k \ne j}B_{i, k}(\tau)x_k(t - \tau)|^2
-\]
-
-By the uniqueness of the projection we must have
-$\forall \tau\ B_{i, j}(\tau) = 0$.
-
-$(2) \Rightarrow (4)$: In computing $(y - \linE{y}{\H_{t - 1}^{-j}})$ for
-$y \in H_t^i$ it is sufficient to consider $y = x_i(t)$ by linearity, then since
+  $(a) \Rightarrow (b)$ follows as a result of the uniqueness of orthogonal
+  projection (i.e. the best estimate is necessarily the coefficients
+  of the model).  $(b) \Rightarrow (c)$ follows since in computing
+  $(y - \linE{y}{\H_{t - 1}^{-j}})$ for $y \in H_t^i$ it is sufficient
+  to consider $y = x_i(t)$ by linearity, then since
   $H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$ we have
-$(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$ since $\B_{ij}(z) = 0$.  $(4)$ then follows
-since $v_i(t) \perp \H_{t - 1}$ and
-$\forall z \in \H_{t - 1}^j\ (z - \linE{z}{\H_{t - 1}^{-j}}) \in \H_{t
-  - 1}$
-\end{proof}
-
-% The following propositions justify various modifications of $x(t)$
-% applied in practice to massage $x(t)$ into a form amenable to more
-% standard tools.
-
-% \begin{theorem}[General Invariance]
-%   \hl{Under what conditions is this actually true?}
-
-%   Let $\zeta(t)$ be a stationary discrete time stochastic process.  Let
-%   $F, G$ be (possibly nonlinear and time varying) invertible filtering
-%   operations and $f(t)$, $g(t)$ be perfectly predictable and such that
-%   $x_j(t) \defeq F(\zeta_j - f)(t), x_i(t) \defeq G(\zeta_i - g)(t)$ are W.S.S.  Then,
-
-% \begin{equation}
-%     x_j \gc x_i \iff \zeta_j \te \zeta_i
-%   \end{equation}
-% \end{theorem}
-
-% \begin{proposition}[Invariance Under Invertible and Deterministic Modifications]
-%   Let $x(t)$ be a WSS process with absolutely summable covariance
-%   sequence, and spectral density uniformly bounded above and below.  Let
-%   $F(z), G$ represent univariate and invertible
-%   linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
-%   predictable processes.  Then,
-
-%   \begin{equation}
-%     x_j \gc x_i \iff F(x_j - f)(t) \gc G(x_i - g)(t)
-%   \end{equation}
-
-% \begin{proof}
-%   \hl{TODO}
-% \end{proof}
-% \end{proposition}
-
-\subsection{Granger Causality Graphs}
-We first need to establish some graph theoretic notation and
-terminology, collected formally in definitions for the reader's
-convenient reference.
-
-\begin{definition}[Graph Theory Review]
-  A \textit{graph} $\gcg = (V, \gcge)$ is simply a
-  tuple of sets respectively called \textit{nodes} and \textit{edges}.
-  Throughout this paper, we have in all cases
-  $V = [n] \defeq \{1, 2, \ldots, n\}$.  We will also focus solely on
-  \textit{directed} graphs, where the edges
-  $\gcge \subseteq V \times V$ are \textit{ordered} pairs.
-
-  A (directed) \textit{path} (of length $r$) from node $i$ to node
-  $j$, denoted $\gcgpath{i}{j}$, is a sequence
-  $a_0, a_1, \ldots, a_{r - 1}, a_r$ with $a_0 = i$ and $a_r = j$ such
-  that $\forall\ 0 \le k \le r\ (a_k, a_{k + 1}) \in \gcge$, and where
-  $(a_k, a_{k - 1})$ are \textit{distinct} for $0 \le k < r$.
-
-  A \textit{cycle} is a path of length $2$ or more between a node and
-  itself.  An edge between a node and itself $(i, i) \in \gcge$ (which
-  is not a cycle) is referred to as a \textit{loop}.
-
-  A graph $\gcg$ is a \textit{directed acyclic graph} (DAG) if it is a
-  directed graph and does not contain any cycles.
-\end{definition}
-
-\begin{definition}[Parents, Grandparents, Ancestors]
-  A node $j$ is a \textit{parent} of node $i$ if $(j, i) \in \gcge$.
-  The set of all $i$'s parents will be denoted $\pa{i}$, and we
-  explicitly exclude loops as a special case, that is,
-  $i \not\in \pa{i}$ even if $(i, i) \in \gcge$.
-
-  The set of level $\ell$ \textit{grandparents} of node $i$, denoted
-  $\gpn{\ell}{i}$, is the set such that $j \in \gpn{\ell}{i}$ if and
-  only if there is a \textit{directed path} of length $\ell$ in $\gcg$
-  from $j$ to $i$.  Clearly, $\pa{i} = \gpn{1}{i}$.
-
-  Finally, the set of \textit{level $\ell$ ancestors} of $i$:
-  $\ancn{\ell}{i} = \bigcup_{\lambda \le \ell}\gpn{\lambda}{i}$ is the
-  set such that $j \in \ancn{\ell}{i}$ if and only if there is a
-  directed path of length $\ell$ \textit{or less} in $\gcg$ from $j$
-  to $i$.  The set of \textit{all ancestors} of $i$
-  (i.e. $\ancn{n}{i}$) is denoted simply $\anc{i}$.
-
-  Recall that we do not allow a node to be it's own parent, however
-  unless $\gcg$ is a DAG, a node can be it's own ancestor.  We will
-  ocassionally need to explicitly exclude $i$ from $\anc{i}$, in which
-  case we will write $\anc{i}\setminus \{i\}$.
-\end{definition}
-
-Our principle object of study will be a graph determined by
-Granger-causality relations as follows.
-
-\begin{definition}[Causality graph]
-  We define the Granger-causality graph $\gcg = ([n], \gcge)$ to be the directed
-  graph formed on $n$ vertices where an edge $(j, i) \in \gcge$ if and
-  only if $x_j$ Granger-causes $x_i$ (with respect to $x$).  That is,
-  $$(j, i) \in \gcge \iff j \in \pa{i} \iff x_j \gc x_i.$$
-\end{definition}
-
-The edges of the Granger-causality graph $\gcg$ can be given a general
-notion of ``weight'' by associating an edge $(j, i)$ with the
-\textit{strictly causal} LTI filter $\B_{ij}(z)$ (see eqn
-\eqref{eqn:filter_action}).  Thence, the matrix $\B(z)$ is analogous
-to a \textit{weighted adjacency matrix}\footnote{We are using the
-  convention that $\B_{ij}(z)$ is a filter with input $x_j$ and output
-  $x_i$ so as to write the action of the system as $\B(z)x(t)$ with
-  $x(t)$ as a column vector.  This competes with the usual convention
-  for adjacency matrices where $A_{ij} = 1$ if there is an edge
-  $(i, j)$.  In our case, the sparsity pattern of $\B_{ij}$ is the
-  \textit{transposed} conventional adjacency matrix.} for the graph $\gcg$.  And,
-in the same way that the $k^{\text{th}}$ power of an adjacency matrix
-counts the number of paths of length $k$ between nodes,
-$(\B(z)^k)_{ij}$ is a filter isolating the ``action'' of $j$ on $i$ at
-a time lag of $k$ steps, this is exemplified in the inversion formula
-\ref{eqn:lsi_inversion}.
-
-An elementary theorem connecting the adjacency matrix with paths will
-allow us to deduce the sparsity pattern of $\A(z)$.  Proof follows
-easily by induction:
+  $(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$ since
+  $\B_{ij}(z) = 0$ and $v_i(t) \perp \H_{t - 1}$.  $(c) \iff (d)$ is a result
+  of the equivalence in Definition
+  \ref{lem:conditional_orthogonality_equivalence}.  And,
+  $(d) \implies (a)$ follows directly from the Definition.
+\end{proof}
 
@@
   \label{lem:adj_matrix}
-  Let $S$ be the transposed adjacency matrix of the Granger-causality
-  graph $\gcg$.  Then, $(S^k)_{ij}$ is the number of paths of length
-  $k$ from node $j$ to node $i$.  Evidently, if
-  $\forall k \in \N,\ (S^k)_{ij} = 0$ then $j \not\in \anc{i}$.
+  Let $S$ be the transposed adjacency matrix\footnote{\footnotesize We
+    are using the convention that $\B_{ij}(z)$ is a filter with input
+    $x_j$ and output $x_i$ so as to write the action of the system as
+    $\B(z)x(t)$ with $x(t)$ as a column vector.  This competes with
+    the usual convention for adjacency matrices where $A_{ij} = 1$ if
+    there is an edge $(i, j)$.  In our case, the sparsity pattern of
+    $\B_{ij}$ is the \textit{transposed} conventional adjacency
+    matrix.} of the Granger-causality graph $\gcg$.  Then,
+  $(S^k)_{ij}$ is the number of paths of length $k$ from node $j$ to
+  node $i$.  Evidently, if $\forall k \in \N,\ (S^k)_{ij} = 0$ then
+  $j \not\in \anc{i}$.
 \end{lemma}
-
-% What if we take this as the definition of PWGC?  then, in a simply causal graph
-% the "normal" definition becomes a theorem.  Or, if lem:pwgc_anc is true then
-% maybe that is good enough?
-
-From the $\VAR$ representation of $x(t)$ there is clearly a tight
-relationship between each node and it's parent nodes, the relationship
-is quantified through the sparsity pattern of $B(z)$.  Similarly, the
-following proposition is analogous to the definition of feedback free
-processes of \cite{caines1975feedback} and provides an interpretation
-of the sparsity pattern of $A(z)$ (from the MA representation of
-$x(t)$) in terms of the causality graph $\gcg$.
-
-% % When is this guaranteed w.r.t. the full system?
-% Furthermore, if the filter $\B_{ii}$ is invertible, we can write
-
-% \begin{equation*}
-%   \begin{aligned}
-%     (1 - \B_{ii}(z))x_i(t) &= v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\\
-%     \Rightarrow x_i(t) &= (1 - \B_{ii}(z))^{-1}\big[v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\big]\\
-%     &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)
-%   \end{aligned}
-% \end{equation*}
-
-% In general, we will write $\wtB$ for arbitrary filters maintaining the
-% convention that $\wtB_{ii}$ is a filter that modifies the future of
-% $x_i$ from it's past and $\wtB_{ij}$ a filter modifying $x_i$ from the
-% past of $x_j$.  We will use the convention that the filter
-% $\widetilde{\B}$ will represent simply ``a filter'' and is not
-% necessarily the same as the filters serving as edge weights in $\gcg$
-% (indeed, $\wtB$ will often be $0$), and moreover, we will allow for
-% $\widetilde{\B}$ to potentially change from place to place and even
-% from line to line in the same sequence of calculations as it's exact
-% properties or specification are not important.
-
-% We continue to develop the relationship between $x_i$ and it's family
-% tree, where the following proposition is analagous to the definitions
-% of feedback free processes of \cite{caines1975feedback}.
+\begin{proof}
+  This is a well known theorem, proof follows by induction.
+\end{proof}
 
@@
 
-  Thence
+  Therefore
 
   \begin{align*}
-    x_i(t) &= [(I - \B(z))^{-1}v(t)]_i\\
-    &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
+    x_i(t) &= [(I - \B(z))^{-1}v(t)]_i = \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
            &= \A_{ii}(z) v_i(t) + \sum_{\substack{j \in \anc{i} \\ j \ne i}} \A_{ij}(z) v_j(t)
@@
 
-\subsection{Pairwise Granger Causality}
-\label{sec:pwgc}
-Recall that Granger-causality in general must be understood with
-respect to a particular universe of observations.  If $x_j \gc x_i$
-with respect to $x_{-k}$, it may not hold with respect to $x$.  For
-example, $x_k$ may be a common ancestor which when observed, completely
-explains the connection from $x_j$ to $x_i$.  In this section we study
-\textit{pairwise} Granger-causality, and seek to understand when
-knowledge of pairwise relations is sufficient to deduce the true fully
-conditional relations of $\gcg$.
-
-\begin{definition}[Pairwise Granger-causality]
-  We will say that $x_j$ pairwise Granger-causes $x_i$ and write
-  $x_j \pwgc x_i$ if $x_j$ Granger-causes $x_i$ with respect only to
-  $(x_i, x_j)$.
-\end{definition}
-
-This notion is of interest for a variety of reasons.  From a purely
-conceptual standpoint, we will see how the notion can in some sense
-capture the idea of ``flow of information'' in the underlying graph,
-in the sense that if $j \in \anc{i}$ we expect that $j \pwgc i$.  It
-may also be useful for reasoning about the conditions under which
-\textit{unobserved} components of $x(t)$ may or may not interfere with
-inference in the actually observed components.  Finally, motivated
-from a practical standpoint to analyze causation in large systems, we
-will seek to construct practical estimation procedures based purely on
-pairwise causality tests since the computation of such pairwise
-relations is somewhat easier.
-
-\begin{lemma}
-  \label{lem:separated_ancestor_uncorrelated}
+\begin{proposition}
+  \label{prop:separated_ancestor_uncorrelated}
   Consider distinct nodes $i, j$ in a Granger-causality graph
@@
   this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_{t - 1}^i} = 0$.
-\end{lemma}
-
-\begin{remark}
-  The possibility that there exist nodes sharing both $i$ and $j$ as
-  ancestors is not excluded, the point being that the temporal nature
-  of Granger-causality eliminates the problems caused by ``colliders''
-  (in the language of Pearl's causal calculus
-  \cite{pearl2009causality}) in static causal inference.
-\end{remark}
+\end{proposition}
 
@@
 
-It is possible for components of $x(t)$ to be correlated at some time
-lags without resulting in pairwise causality:
-
-\begin{lemma}
-  \label{lem:ancestor_uncorrelated}
+\begin{proposition}
+  \label{prop:ancestor_uncorrelated}
   Consider distinct nodes $i, j$ in a Granger-causality graph $\gcg$.
@@
   then $j \npwgc i$.
-\end{lemma}
+\end{proposition}
 \begin{proof}
@@
   \begin{equation*}
-    \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^i}} = 0,
+    \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^i}} = 0.
   \end{equation*}
@@
   which when substituted into the left hand side of Equation
-  \ref{eqn:proof_inner0} results (by Lemma
-  \ref{lem:separated_ancestor_uncorrelated}) in
+  \ref{eqn:proof_inner0} results (by Proposition
+  \ref{prop:separated_ancestor_uncorrelated}) in
 
@@
 \end{proof}
-
-The previous proposition can be strengthened significantly; notice
-that it is possible to have some $k \in \anc{i} \cap \anc{j}$ where still
-$j \npwgc i$, an example is furnished by the three node graph
-$k \rightarrow i \rightarrow j$ where clearly
-$k \in \anc{i}\cap\anc{j}$ but $j \npwgc i$.  We must introduce the concept
-of a \textit{confounding} variable, which effectively eliminates the
-possibility presented in this example.
-
-\begin{definition}[Confounder]
-  A node $k$ will be referred to as a \textit{confounder} of nodes
-  $i, j$ (neither of which are equal to $k$) if
-  $k \in \anc{i} \cap \anc{j}$ and there exists a path
-  $\gcgpath{k}{i}$ not containing $j$, and a path $\gcgpath{k}{j}$
-  not containing $i$.
-
-  A simple example is furnished by the ``fork'' graph
-  $i \leftarrow k \rightarrow j$.
-\end{definition}
 
@@
 \end{proposition}
-
-\begin{remark}
-  The interpretation of this proposition is that for $j \pwgc i$ then
-  there must either be ``causal flow'' from $j$ to $i$
-  ($j \in \anc{i}$) or there must be a confounder $k$ through which
-  common information is received.
-\end{remark}
 
@@
 
-  which follows from Lemma \ref{lem:ancestor_uncorrelated}.  Finally,
+  which follows from Proposition \ref{prop:ancestor_uncorrelated}.  Finally,
   for $k \in C_1(j)$ the case $k = i$ is immediate (since the error in
@@
 
-An interesting corollary is the following:
-
-\begin{corollary}
-  If the graph $\gcg$ is a DAG then $j \pwgc i, i \pwgc j \implies \exists k \in \anc{i} \cap \anc{j}$ confounding $(i, j)$.
-\end{corollary}
-
-It seems reasonable to expect a converse of proposition
-\ref{prop:ancestor_properties} to hold, i.e.
-$j \in \anc{i} \Rightarrow j \pwgc i$.  Unfortunately, this is not the
-case in general, as different paths through $\gcg$ can lead to
-cancellation (see example \ref{ex:diamond_cancellation}).  In fact, we
-do not even have $j \in \pa{i} \Rightarrow j \pwgc i$ (see example
-\ref{ex:lag_cancellation}).
-
-\begin{example}
-  \label{ex:diamond_cancellation}
-  Firstly, on $n = 4$ nodes, ``diamond'' shapes can lead to cancellation on paths of length 2:
-
-\begin{equation*}
-  x(t) =
-  \left[
-    \begin{array}{cccc}
-      0 & 0 & 0 & 0\\
-      a & 0 & 0 & 0\\
-      -a & 0 & 0 & 0\\
-      0 & 1 & 1 & 0\\
-    \end{array}
-  \right] x(t - 1) + v(t),
-\end{equation*}
-
-with $\E v(t) = 0,\ \E v(t)v(t - \tau)^\T = \delta_\tau I$.
-
-By directly calculating
-
-\begin{align*}
-  x_4(t) &= x_2(t - 1) + x_3(t - 1) + v_4(t)\\
-         &= ax_1(t - 2) + av_2(t - 1) - ax_1(t - 2) -av_3(t - 1) + v_4(t)\\
-         &= a(v_2(t - 1) - v_3(t - 1)) + v_4(t),
-\end{align*}
-
-we see that, since $v(t)$ is isotropic white noise, $1 \npwgc 4$.  The problem here is that there are multiple paths from $x_1$ to $x_4$.
-\end{example}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[b]{0.45\textwidth}
-    \includegraphics[width=\linewidth]{example1.pdf}
-    \caption{Graph Corresponding to Example \ref{ex:diamond_cancellation}}
-    \label{fig:diamond_cancellation}
-    \centering{$j \in \anc{i} \nRightarrow j \pwgc i$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \includegraphics[width=\linewidth]{example2.pdf}
-    \caption{Graph Corresponding to Example \ref{ex:lag_cancellation}}
-    \label{fig:lag_cancellation}
-    \centering{$j \in \pa{i} \nRightarrow j \pwgc i$}
-  \end{subfigure}
-\end{figure}
-
-\begin{example}
-  \label{ex:lag_cancellation}
-  A second example on $n = 3$ nodes is also worth examining, in this case
-  cancellation is a result of differing time lags.
-
-\begin{equation*}
-  x(t) =
-  \left[
-    \begin{array}{ccc}
-      0 & 0 & 0\\
-      -a & 0 & 0\\
-      0 & 1 & 0\\
-    \end{array}
-  \right] x(t - 1) +
-  \left[
-    \begin{array}{ccc}
-      0 & 0 & 0\\
-      0 & 0 & 0\\
-      a & 0 & 0\\
-    \end{array}
-  \right] x(t - 2) + v(t)
-\end{equation*}
-
-Then
-
-\begin{align*}
-  x_2(t) &= v_2(t) - ax_1(t - 1)\\
-  x_3(t) &= v_3(t) + x_2(t - 1) + ax_1(t - 2)\\
-  \Rightarrow x_3(t) &= v_2(t - 1) + v_3(t),
-\end{align*}
-
-and again $1 \npwgc 3$.
-\end{example}
-
-\subsection{Strongly Causal Graphs}
-\label{sec:strongly_causal_graphs}
-% The examples at the conclusion of the previous section seem rather
-% pathological since we have simply constructed a case where different
-% paths in the graph cancel exactly.  It may be possible to rid
-% ourselves of these pathologies for instance by defining a notion of
-% ``robust'' pairwise causality where ``$j \pwgc _R i$'' if $j \pwgc i$
-% almost surely for a ``small'' random perturbation of the system
-% matrix.  However, we do not believe such a modification of our theory
-% can have useful practical implications since the difference between
-% paths that ``nearly cancel'' rather than cancel exactly would in
-% practice require inordinate amounts of data to resolve.
-
-In this section and the next we will seek to understand when converse
-statements of Proposition \ref{prop:ancestor_properties} \textit{do}
-hold.  One possibility is to restrict the coefficients of the system
-matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead,
-we think it more meaningful to focus on the defining feature of
-time series networks, that is, the topology of $\gcg$.
-
-\begin{definition}[Strongly Causal]
-  \label{def:strongly_causal}
-  We will say that a Granger-causality graph $\gcg$ is
-  \textit{strongly causal} if there is at most 1 directed path between
-  any two nodes.  Strongly Causal Graphs will be referred to as SCGs.
-\end{definition}
-
-Examples of strongly causal graphs include directed trees (or
-forests), DAGs where each node has at most one parent, and figure
-\ref{fig:example_fig3} of this paper.  A complete bipartite graph with
-$2n$ nodes is also strongly causal, demonstrating that the number of
-edges of such a graph can still scale quadratically with the number of
-nodes.  It is evident that the strong causal property is inherited by
-subgraphs.
-
-\begin{example}
-  Though examples of SCGs are easy to construct in theory, should
-  practitioners expect SCGs to arise in application?  While a positive
-  answer to this question is not \textit{necessary} for the concept to
-  be useful, it is certainly sufficient.  Though the answer is likely
-  to depend upon the particular application area, examples appear to
-  be available in biology, in particular, the authors of
-  \cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty}
-  cite an example of the so called ``transcription regulatory network
-  of \textit{E.coli}'', and
-  \cite{learning_genome_scale_regulatory_networks} study a much larger
-  regulatory network of \textit{Saccharomyces cerevisiae}.  These
-  networks, which we reproduce\footnote{Figure \ref{fig:gene_network1}
-    is reproduced under the Creative Commons Attribution
-    Non-Commercial License
-    (\url{http://creativecommons.org/licenses/by-nc/2.5}) and Figure
-    \ref{fig:gene_network2} under the Creative Commons Attribution
-    License (\url{https://creativecommons.org/licenses/by/4.0/})} in
-  figure \ref{fig:gene_networks} appear to have at most a small number
-  of edges which violate the strong-causality condition.
-
-  \begin{figure}[h]
-    \centering
-    \caption{Transcription Regulatory Networks}
-    \label{fig:gene_networks}
-    \begin{subfigure}[b]{0.45\textwidth}
-      \caption{\textit{E.Coli} Network of
-        \cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty}}
-      \label{fig:gene_network1}
-      \includegraphics[width=\linewidth, height=\linewidth]{ecoli_regulatory_network.png}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.45\textwidth}
-      \caption{\textit{Saccharomyces cerevisiae} Network of
-        \cite{learning_genome_scale_regulatory_networks}}
-      \label{fig:gene_network2}
-      \includegraphics[width=\linewidth, height=\linewidth]{huge_gene_network.png}
-    \end{subfigure}
-  \end{figure}
-\end{example}
-
-% \begin{lemma}
-%   \label{lem:still_strongly_causal}
-%   If $\gcg$ is a strongly causal graph, any subgraph formed by eliminating nodes
-%   as well as all of the in and out edges thereof is still strongly causal.
-% \end{lemma}
-% \begin{proof}
-%   If $\gcg$ has at most one path between any two nodes, there can only
-%   be fewer paths after removing nodes from $\gcg$.
-% \end{proof}
-
-For later use, and to get a feel for the topological implications of
-strong causality, we explore a number of properties of such graphs
-before moving into the main result of this section.  The following
-important property essentially strengthens proposition
-\ref{prop:ancestor_properties} for the case of strongly causal graphs.
-
 \begin{proposition}
@@
 \end{proof}
-
-% Is pwgc transitive?
-
-\begin{corollary}
-  \label{cor:parent_corollary}
-  If $\gcg$ is a strongly causal DAG then $i \pwgc j$ and $j \in \anc{i}$ are
-  \textit{alternatives}, that is $i \pwgc j \Rightarrow j \notin \anc{i}$.
-\end{corollary}
-\begin{proof}
-  Suppose that $i \pwgc j$ and $j \in \anc{i}$.  Then since $\gcg$ is
-  acyclic $i \not\in \anc{j}$, and by proposition
-  \ref{prop:ancestor_properties} there is some
-  $k \in \anc{i}\cap\anc{j}$ which is a confounder.  However, by
-  proposition \ref{prop:sc_graph_common_anc} $k$ cannot be a
-  confounder, a contradiction.
-\end{proof}
-% This is a direct proof
-  % \begin{proof}
-%   Suppose that $\gcg$ is a strongly causal DAG and that we have both
-%   $i \pwgc j$ and $j \in \anc{i}$, which implies that
-%   $i \not \in \anc{j}$ since $\gcg$ is a DAG.  We will establish the
-%   contradiction $i \in \anc{j}$.
-
-%   Since $j \in \anc{i}$ there is a path $\gcgpath{j}{i}$.
-%   Moreover, since $i \pwgc j$ by proposition \ref{prop:pwgc_anc} there
-%   must be a confounding ancestor $k \in \anc{i} \cap \anc{j}$, where the
-%   node $k$ has a path to $j,\ \gcgpath{k}{j}$, as well as a path
-%   to $i,\ \gcgpath{k}{i}$.
-
-%   \hl{Double check that the ancestor properties proposition implies
-%     that BOTH paths do not contain the other node.}
-
-%   Now, since the graph is strongly causal, the only possible
-%   $\gcgpath{k}{i}$ path is the one obtained by concatenating the
-%   $\gcgpath{k}{j}$ path with the $\gcgpath{j}{i}$ path.
-%   However, this is a contradiction since $k$ is a confounder and
-%   should have a $\gcgpath{k}{i}$ path which does not contain $i$.
-% \end{proof}
-
-\begin{corollary}
-  \label{cor:bidirectional_edge}
-  If $\gcg$ is a strongly causal DAG such that $i \pwgc j$ and
-  $j \pwgc i$, then $i \not\in \anc{j}$ and $j \not\in \anc{i}$.  In
-  particular, a pairwise bidirectional edge indicates the absence of
-  any edge in $\gcg$.
-\end{corollary}
-\begin{proof}
-  This follows directly from applying proposition
-  \ref{cor:parent_corollary} to $i \pwgc j$ and $j \pwgc i$.
-% This is a direct proof
-% \begin{proof}
-  % By way of contradiction, suppose that $i \in \pa{j}$.  We will
-  % consider the two possibilities allowed by proposition
-  % \ref{prop:ancestor_properties} for $j \pwgc i$.  Firstly
-  % $j \in \anc{i}$ is impossible since $\gcg$ is assumed to be acyclic.
-  % Secondly, if there is some confounding node
-  % $u \in \anc{i} \cap \anc{j}$ with a path
-  % $u \rightarrow \cdots \rightarrow j$ which does not contain $i$ we
-  % have a contradiction since there must now be multiple
-  % $u \rightarrow j$ paths: the aforementioned, and a path
-  % $u \rightarrow \cdots \rightarrow i \rightarrow \cdots \rightarrow
-  % j$ which \textit{does} contain $i$.  We conclude that $i \in \pa{j}$
-  % is impossible, and symmetrically that $j \in \pa{i}$ is as well.
-% \end{proof}
-\end{proof}
-
-% \begin{proposition}
-%   \label{prop:independent_confounders}
-%   If $\gcg$ is strongly causal we have distinct
-%   $k_1, k_2 \in \anc{i} \cap \anc{j}$ are confounders of $i, j$, then
-%   $k_1 \not\in \anc{k_2}$, $k_2 \not\in \anc{k_1}$ and
-%   $\not\exists \ell \in \anc{k_1} \cap \anc{k_2}$ confounding
-%   $k_1, k_2$.
-% \end{proposition}
-% \begin{proof}
-%   \hl{TODO}
-% \end{proof}
-
-% \begin{corollary}
-%   If $\gcg$ is strongly causal, then distinct confounders are
-%   uncorrelated.
-% \end{corollary}
-% \begin{proof}
-%   This follows directly by combining proposition
-%   \ref{prop:independent_confounders} with proposition \ref{prop:ancestor_properties}.  \hl{I need to update prop} \ref{prop:ancestor_properties} \hl{to note the strong non-correlation property.}
-% \end{proof}
-
-% \begin{lemma}
-%   If $\gcg$ is a strongly causal DAG and $j \in \anc{i}$ then 
-% \end{lemma}
-
-In light of proposition \ref{prop:sc_graph_common_anc}, the following
-provides a partial converse to proposition
-\ref{prop:ancestor_properties}, and supports the intuition of ``causal
-flow'' through paths in $\gcg$.
 
@@
 
-We immediately obtain the corollary, which we remind the reader is,
-surprisingly, not true in a general graph.
-
-\begin{corollary}
-  \label{cor:gc_implies_pwgc}
-  If $\gcg$ is a strongly causal DAG then $j \gc i \Rightarrow j \pwgc i$.
-\end{corollary}
-
-% \begin{remark}
-%   It can be seen from the conclusion of the proof that simple facts
-%   about Granger causality are indeed reliant on our assumptions about
-%   the nature of $x(t)$ laid out in the section \ref{sec:theory}, in
-%   particular, the innovations process $v(t)$ must be full rank.
-
-%   Moreover, that strong conditions need to be placed on the topology
-%   of $\gcg$ in order for proposition \ref{prop:pwgc_anc} to hold can
-%   be seen through the earlier examples.
-% \end{remark}
-
-% This holds in any DAG
-% \begin{lemma}
-%   If $\gcg$ is strongly causal, then any $\\VAR(p)$ system on $\gcg$ is stable if and only if
-%   $B_{ii}(z)$ is stable for every $i = 1, \ldots, n$.
-% \end{lemma}
-% \begin{proof}
-%   \hl{This is certainly true, but I don't think I need to use it anywhere.}
-% \end{proof}
-
-\begin{remark}
-  As we have seen and as is true in much of statistics, confounding
-  nodes pose challenges for Granger-causality.  However, as opposed to
-  Pearl's causal calculus \cite{pearl2000art}, pairwise
-  Granger-causality does not suffer any difficulty with so-called
-  ``colliders'', that is, the topology $i \rightarrow k \leftarrow j$
-  will never result in $i \pwgc j$ or $j \pwgc i$.  This is evidently
-  an advantage of the \textit{temporal} nature of Granger-causality -- there is
-  no backwards causal flow along the edges of $\gcg$.
-\end{remark}
-
-\begin{example}
-  As a final remark of this subsection we note that a complete
-  converse to proposition \ref{prop:ancestor_properties} is not
-  possible without additional conditions.  Consider the ``fork'' system on $3$
-  nodes (i.e. $2 \leftarrow 1 \rightarrow 3$) defined by
-
-  \begin{equation*}
-    x(t) =
-    \left[
-      \begin{array}{cccc}
-        0 & 0 & 0\\
-        a & 0 & 0\\
-        a & 0 & 0\\
-      \end{array}
-    \right] x(t - 1) + v(t).
-  \end{equation*}
-
-  In this case, node $1$ is a confounder for nodes $2$ and $3$, but
-  $x_3(t) = v_3(t) - v_2(t) + x_2(t)$ and $2 \npwgc 3$ (even
-  though $x_2(t)$ and $x_3(t)$ are contemporaneously correlated)
-
-  If we were to augment this system by simply adding an autoregressive
-  component (i.e. some ``memory'') to $x_1(t)$ e.g.
-  $x_1(t) = v_1(t) + b x_1(t - 1)$ then we \textit{would} have
-  $2 \pwgc 3$ since then
-  $x_3(t) = v_3(t) + av_1(t - 1) - bv_2(t - 1) + bx_2(t - 1)$.  We
-  develop this idea further in the next section.
-\end{example}
-
-\subsection{Persistent Systems}
-\label{sec:persistent_systems}
-In section \ref{sec:strongly_causal_graphs} we obtained a converse to
-part $(a)$ of proposition \ref{prop:ancestor_properties} via the
-notion of a strongly causal graph topology.  In this section, we
-complete a converse by adding the additional requirement we refer to
-as ``persistence''.
-
-\begin{definition}[Lag Function]
-  Given a causal filter $\B(z) = \sum_{\tau = 0}^\infty b(\tau)z^{-\tau}$
-  define
-
-  \begin{align}
-    \tau_0(\B) &= \text{min}\{\tau \in \Z_+\ |\ b(\tau) \ne 0\},\\
-    \tau_{\infty}(\B) &= \text{sup}\{\tau \in \Z_+\ |\ b(\tau) \ne 0\}.\\
-  \end{align}
-
-  i.e. the ``first'' and ``last'' coefficients of the filter $\B(z)$,
-  where $\tau_\infty(\B) \defeq \infty$ if the filter has an infinite
-  length, and $\tau_0(\B) \defeq \infty$ if $\B(z) = 0$.
-\end{definition}
-
-This interpretation of the following persistence condition is that
-each node stores some ``memory'' of the past.
-
-\begin{definition}[Persistent]
-  We will say that the process $x(t)$ with Granger-causality graph
-  $\gcg$ is \textit{persistent} if for every $i \in [n]$ and every
-  $k \in \anc{i}$ we have $\tau_0(\A_{ik}) < \infty$ and $\tau_\infty(\A_{ik}) = \infty$.
-\end{definition}
-
-\begin{remark}
-  In the context of Granger-causality, ``most'' systems should be
-  persistent.  In particular, $\mathsf{\VAR}(p)$ models are likely to
-  be persistent since these naturally result in an equivalent
-  $\mathsf{MA}(\infty)$ representation, except for the pathological
-  case where $\B(z)$ is Nilpotent.
-
-  Moreover, persistence is not the weakest condition necessary for the
-  results of this section, the condition
-  $\tau_0(\A_{jk}) < \tau_\infty(\A_{ik})$ for each $i, j, k$ such that
-  $k \in \anc{i} \cap \anc{j}$ is enough.  The intuition being that nodes
-  $i$ and $j$ are not receiving temporally disjoint information from
-  $k$.
-
-\end{remark}
-
-\begin{example}
-  process $x(t)$ generated by the $\VAR(1)$ model\footnote{Recall that
-    any $\VAR(p)$ model with $p < \infty$ can be written as a
-    $\VAR(1)$ model, so we lose little generality in considering this
-    case.}  having $\B(z) = Bz^{-1}$, we will examine
-  conditions that guarantee persistence.  Pick any
-  $i \in [n], j \in \anc{i} \setminus \{i\}$, then the stability of
-  $B$ allows us to write
-
-  \begin{equation*}
-    \A(z) = \sum_{k = 0}^\infty B^k z^{-k},
-  \end{equation*}
-
-  whereby we see that $\exists k > 0$ such that $[B^k]_{ij} \ne 0$
-  (since $j \in \anc{i}$).  Then consider
-
-  \begin{equation*}
-    \begin{aligned}
-      e_i^\T B^{rk} e_j &\overset{(a)}{=} \big((P^\T e_i)^\T J^{rk} P^{-1}e_j\big)\\
-      &= \tr [(P^\T e_i)^\T J^{rk} P^{-1}e_j]\\
-      &\overset{(b)}{=} \tr [(J^{rk}) (v u^\T)],
-    \end{aligned}
-  \end{equation*}
-
-  where $(a)$ utilizes the Jordan Normal Form of $B$, and $(b)$
-  denotes $u = P^\T e_i$ and $v = P^{-1}e_j$.  In order for
-  $\tau_\infty(\A_{ij}) < \infty$, there must be some $N > 1$ such
-  that $\forall r \ge N$, the above term is $0$.  This may be the case
-  for instance if $B$ is a nilpotent matrix.
-
-  Let us suppose now that $B$ is diagonalizable (i.e. $J$ is a
-  diagonal matrix) with at least $2$ distinct eigenvalues; in this
-  case $B$ is also \textit{not} nilpotent.  We can then rewrite the
-  above as
-
-  \begin{equation*}
-    f(r) \defeq \tr [(J^{rk}) (v u^\T)] = \sum_{\nu = 1}^n \lambda_\nu ^{rk} v_\nu u_\nu \defeq \sum_{\nu = 1}^n \lambda_\nu^{rk} \beta_\nu
-  \end{equation*}
-
-  where $\lambda_\nu$ denotes the eigenvalues of $B$ and
-  $\beta_\nu = u_\nu v_\nu$.  Note that $f(0) = 0$ since $i \ne j$,
-  $u$ is a row of $P$ and $v$ is a column of $P^{-1}$.  Moreover,
-  $f(1) \ne 0$ by hypothesis.  But, in order for
-  $f(r) = 0\ \forall r \ge N$, it would need to be the case that
-
-  \begin{equation*}
-    \Dg(\bm{\lambda})^r \bm{\lambda} = Vz
-  \end{equation*}
-
-  had a solution in $z$ for every $r \ge N$, where $V$ is an
-  $n \times n - 1$ full-rank matrix whose columns span the nullspace of $\beta$,
-  and $\bm{\lambda} = (\lambda_1, \ldots, \lambda_n)$. That is,
-  iterates of $\Dg(\bm{\lambda})$ applied to $\bm{\lambda}$ would need to remain
-  inside $\beta$'s nullspace.  This would imply that
-
-  \begin{equation*}
-    VV^\dagger \bm{\lambda}^{r + 1} = \bm{\lambda}^{r + 1},
-  \end{equation*}
-
-  i.e. that $\bm{\lambda}^{r + 1}$ is an eigenvector of $VV^\dagger$
-  for an infinite number of integers $r$ (the exponentiation is to be
-  understood as a pointwise operation).  However, since there can only
-  be a finite number of (unit length) eigenvectors, this cannot be the
-  case unless every eigenvalue $(\lambda_1, \ldots, \lambda_n)$ were
-  equal.
-
-  We see from this example that the collection of $\VAR(1)$ systems
-  which are not persistent are pathological, in the sense that their
-  system matrices have zero measure when viewed as a subset of $\R^{n^2}$.
-\end{example}
-
 \begin{lemma}
@@
     &\overset{(a)}{=} \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
-    &\overset{(b)}{=} \inner{\sum_{k \in \anc{j}}\big(F(z)\A_{jk}(z)v_k(t) - \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}\big)}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
-    &\overset{(c)}{=} \sum_{k \in \anc{i}\cap\anc{j}}\Big(\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} - \inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{ii}v_i(t)}\\
+    &\overset{(b)}{=} \inner{\sum_{k \in \anc{j}}\big(F(z)\A_{jk}(z)v_k(t) \\&- \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}\big)}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
+    &\overset{(c)}{=} \sum_{k \in \anc{i}\cap\anc{j}}\Big(\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} \\&- \inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{ii}v_i(t)}\\
     &- \sum_{\ell \in \anc{i}}\inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{i\ell}(z)v_\ell(t)}\Big)
@@
   $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$
-  and thence there is in general no cancellation in the final term
+  and therefore there is in general no cancellation in the final term
   above for $\ell \in \anc{i}$.
@@
 
-\subsection{Recovering $\gcg$ via Pairwise Tests}
-\label{sec:pairwise_algorithm}
-In this section we will show that if the $\gcg$ of a persistent
-process is a strongly causal DAG, then it is possible to recover
-$\gcg$ via pairwise tests alone.  This is the main conclusion of the
-theoretical analysis in this paper.
-
-% This is an interesting fact in and
-% of itself, but also has some implications for applications since
-% pairwise testing is trivially parallelizable.  In section
-% \ref{sec:structure_learning} we will also analyze the use of pairwise
-% testing as a heuristic for general graphs.
+\subsection{The Main Theorem}
+\label{apx:proof_main_theorem}
 
@@
 
-    \BlankLine
-    \caption{Pairwise Graph Recovery}
+  % \BlankLine
+  \caption{Pairwise Granger Causality Algorithm (PWGC)}
   \label{alg:pwgr}
-    % \TitleOfAlgo{Pairwise Graph Recovery}
-    \Input{Pairwise Granger-causality relations between a persistent
-      process of dimension $n$ whose joint Granger-causality
-      relations are known to form a strongly causal DAG $\gcg$.}
+  % \TitleOfAlgo{Pairwise Graph Recovery (PWGC)}
+  % \Input{Pairwise Granger-causality relations between a persistent
+  % process of dimension $n$ whose joint Granger-causality
+  % relations are known to form a strongly causal DAG $\gcg$.}
+  \Input{Pairwise Granger-causality relations}
   \Output{Edges $\gcge = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of
@@
 \end{algorithm}
-\end{theorem}
-
-We will shortly prove the theorem by establishing the correctness of
-algorithm (\ref{alg:pwgr}).  The idea is to iteratively ``peel away
-layers'' of nodes by removing the nodes that have no parents
-remaining, which always exist since the graph is acyclic.  The
-requirement of strong causality ensures that all actual edges of
-$\gcg$ manifest in some way as pairwise relations (by proposition
-\ref{prop:pwgc_anc}), and the requirement of persistence allows
-confounding to be eliminated by removing bidirectional edges.  Without
-persistence, each confounded pair would give rise to $4$ possible
-pairwise topologies consistent with $\gcg$, one for each type of
-pairwise edge (no edge, unidirectional, bidirectional).
-
-\begin{example}
-  The set $W$ collects ancestor relations in $\gcg$ (see Lemma
-  \ref{lem:W_subset_E}).  In reference to figure
-  \ref{fig:example_fig3}, each of the solid black edges, as well as
-  the dotted red edges will be included in $W$, but \textit{not} the
-  bidirectional green dash-dotted edges, which we are able to exclude
-  by appealing to Corollary \ref{cor:bidirectional_edge}.  The
-  groupings $P_0, \ldots, P_3$ are also indicated in figure
-  \ref{fig:example_fig3}.
-
-  The algorithm proceeds first with the parentless nodes $1, 2$ on the
-  initial iteration where the edge $(1, 3)$ is added to $E$.  On the
-  next iteration, the edges $(3, 4), (2, 4), (3, 5)$ are added, and
-  the false edges $(1, 4), (1, 5)$ are excluded due to the paths
-  $1 \rightarrow 3 \rightarrow 4$ and $1 \rightarrow 3 \rightarrow 5$
-  already being present.  Finally, edge $(4, 6)$ is added, and the false
-  $(1, 6), (3, 6), (2, 6)$ edges are similarly excluded due to the
-  ordering of the inner loop.
-
-  \begin{figure}
-    \centering
-    \caption{Example graph for Algorithm \ref{alg:pwgr}}
-    \footnotesize{Black arrows indicate true parent-child
-      relations.  Red dotted arrows indicate pairwise causality (due to
-      non-parent relations), green dash-dotted arrows indicate
-      bidirectional pairwise causality (due to the confounding node
-      $1$).  Blue groupings indicate each $P_k$ in Algorithm
-      \ref{alg:pwgr}.}
-    \label{fig:example_fig3}
-
-    \begin{subfigure}[b]{0.45\textwidth}
-      \includegraphics[width=\linewidth]{example_algorithm.pdf}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.45\textwidth}
-      \includegraphics[width=\linewidth]{example_algorithm2.pdf}
-    \end{subfigure}
-  \end{figure}
-
-  That we need to proceed backwards through $P_{k - r}$ as in the
-  inner loop of on $r$ can also be seen from this example, where if
-  instead we simply added the set
-
-  \begin{equation*}
-    D_k' = \{(i, j) \in \Big(\bigcup_{r = 1}^k P_{k - r}\Big) \times P_k\ |\ i \pwgc j \}
-  \end{equation*}
-
-  to $E_k$ then we would infer the false positive edge
-  $1 \rightarrow 4$.  Moreover, the same example shows that simply
-  using the set
-
-  \begin{equation*}
-    D_k'' = \{(i, j) \in P_{k - 1} \times P_k\ |\ i \pwgc j \}  ,
-  \end{equation*}
-
-  causes the edge $1 \rightarrow 3$ to be missed.
-\end{example}
-
-% \begin{lemma}
-%   \label{lem:inner_loop_lemma}
-%   At any given iteration, $k$, of Algorithm \ref{alg:pwgr}, if we have
-%   $e \in E$ if and only if $e \in \gcge$, then the inner loop (line \ref{alg:inner_loop})
-%   will add $(i, j) \in \big(\bigcup_{r = 0}^kP_{k - r}\big) \times C_k$ to $E$
-%   if $(i, j) \in \gcge$.  i.e. any real edge will be included.
-% \end{lemma}
-% \begin{proof}
-%   Let $j \in C_k$, $i \in \bigcup_{r = 0}^k P_{k - r}$, and assume
-%   $i \in \pa{j}$.
-
-%   First, suppose that there is no $i \rightarrow j$ path in $E$.  Then
-%   since $i \gc j$ we have $i \pwgc j$ and $j \npwgc i$ (i.e.
-%   $(i, j) \in W$) by Proposition \ref{prop:pwgc_anc} and Corollary
-%   \ref{cor:bidirectional_edge} so at least one of
-%   $\{D_{kr} \}_{r = 0}^k$ contains $(i, j)$ and hence $(i, j)$ will be
-%   added to $E$ upon the end of the iteration.
-
-%   Now, suppose instead that there is already an $i \rightarrow j$ path
-%   in $E$.  Given our assumption that $E$ contains edges if and only if
-%   they are edges of $\gcg$, then either we already have $(i, j) \in E$
-%   correctly, or there is more than one $i \rightarrow j$ path in
-%   $\gcg$, which is not possible since $\gcg$ is strongly causal.
-% \end{proof}
-
-% \begin{lemma}
-%   \label{lem:W_subset_E}
-%   $W \subseteq \gcge$
-% \end{lemma}
-% \begin{proof}
-%   By proposition \ref{prop:pwgc_anc} we have
-%   $j \in \pa{i} \Rightarrow j \pwgc i$ and by Corollary \ref{cor:bidirectional_edge}
-%   there are no bidirectional edges in $\gcg$.
-% \end{proof}
-
-% \hl{Must show that Algorithm terminates in $n$ steps.}
-% \begin{lemma}
-%   For any $i \in P_k$ we have $i \in C_{k - 1}$.  \hl{[Can't yet prove
-%     this part] Moreover, for each $i \in \bigcup_{k = 0}^n P_k$, if
-%     $(s, i) \in W$ then $(s, i) \in E$ upon the termination of
-%     Algorithm} \ref{alg:pwgr}.
-% \end{lemma}
-% \begin{proof}
-%   We prove the second part by induction, the first part being
-%   established along the way.  By construction $P_0$ has no parent
-%   nodes in $\gcg$ so the base case is immediate.  Fix now some
-%   $k \ge 1$, $i \in P_k$, and let $s$ be such that $(s, i) \in W$.
-%   For every $s$ such that $(s, i) \in W$ we have $s \not\in S_k$ since
-%   otherwise $i \not\in P_k$.  Therefore
-%   $s \in \bigcup_{r = 0}^{k - 1}P_{k - r}$, which implies
-%   $s \not\in S_k$, and finally that $i \in C_{k - 1}$.  Then by Lemma
-%   \ref{lem:W_subset_E} every edge in $\gcge$ is also present in
-%   $W$, and by Lemma \ref{lem:inner_loop_lemma} \hl{[We can't conclude
-%     here because of the strong conditions required on lemma}
-%     \ref{lem:inner_loop_lemma}]
-% \end{proof}
-
-% \begin{lemma}[Loop Invariant]
-%   \label{lem:loop_invariant}
-%   $\forall k \ge 0,\ e \in E_k \Rightarrow e \in \gcge$.
-%   \hl{Not sure if this is useful / the right approach.}
-% \end{lemma}
-% \begin{proof}
-%   We will proceed by induction.  Firstly, the base case
-%   $E_0 = \emptyset$ is trivial, so fix some $k \ge 1$ and consider an
-%   edge $(i, j) \in \big(\bigcup_{r = 0}^k P_{k - r} \big) \times C_k$
-%   such that $(i, j) \in W$ (if the edge is not in $W$ then it is not
-%   in $\gcge$ by Lemma \ref{lem:W_subset_E}).
-
-%   \hl{WIP}
-% \end{proof}
 
@@
 
-% \begin{proof}
-%   Denote the edges of $\gcg$ by $\gcge$, we endevour to show
-%   that upon the termination of Algorithm \ref{alg:pwgr} we have
-%   $E = \gcge$.  On the first iteration of the algorithm
-%   ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
-
-%   Moving on from any trivial cases, notice that
-%   $\gcge \subseteq W$.  This follows since by Proposition
-%   \ref{prop:pwgc_anc} we have $j \in \pa{i} \Rightarrow j \pwgc i$ and by
-%   Corollary \ref{cor:bidirectional_edge} there are no bidirectional
-%   edges in $\gcg$.
-
-% 1. Dispense with trivialities
-% 2. Show that hat{P}_k is nonempty
-% 3. Show that C_k is nonempty
-% 4. Show that edges in D_k are actual edges in G
-% 5. show that each edge incident on C_k is in D_k
-% 6. Induction
-
-% Given non-empty P and C...  Can we show all P -> C edges get collected in D?
-  % Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$ and let us define
-
-  % \begin{equation*}
-  %   A_k = \{i \in [n]\ |\ \dist{i}{P_0} = k\},
-  % \end{equation*}
-
-  % where
-
-  % \begin{equation*}
-  %   \dist{i}{P} = \text{max}\{r \in [n]\ |\ \exists (j \rightarrow i) \text{ path of length } r \text{ in } \gcg \text{ for some } j \in P\}.
-  % \end{equation*}
-
-  % We will proceed by induction and show that when the counter has reached the value $k \ge 1$, it is guaranteed that all nodes in $\bigcup_{\ell \le k}A_\ell$ (nodes having paths of length up to $k$ between themselves and $P_0$) must have all of their incident edges included in $E$, and that every edge in $E$ is an edge of $\gcg$.
-If this is the case, then it can be immediately seen that the algorithm terminates with the correct result, $\gcg = ([n], E)$, once $k = n$; since there can only be paths up to length $n - 1$ in $\gcg$, and there cannot be any nodes or edges that don't involve paths back to $P_0$.
-That is, $\bigcup_{\ell \le n}A_\ell = [n]$.
-
-  % % firstly dispensing with some trivialities, and then establishing that every edge in $\gcg$ makes an appearance in some $D_{kr}$ as well as that everything in any $D_{kr}$ is indeed an edge in $\gcg$.
-
-  % On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
-To quickly dispense with some trivialities: the set $P_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $P_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
-Now, if $P_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
-
-  % Continuing with the non-trivial case: if $P_0 \ne [n]$ then $C_0$ must be non-empty, otherwise there would be driving nodes not included in $P_0$.
-In this case it is also clear that $A_1 \ne \emptyset$, so let us consider some $j \in A_1$.
-We must have $j \in C_0$, otherwise there would be a path of length $2$ from $P_0$ to $j$.
-Moreover, since $\exists i \in P_0$ s.t.
-$i \gc j$ and therefore $i \pwgc j$ (by proposition \ref{prop:pwgc_anc}) we have $(i, j) \in D_{00}$ and therefore $(i, j) \in E$.
-Therefore $A_1 \subseteq D_{00}$.
-\hl{This is nonsensical because $A_1$ does not contain tuples -- need to show that $A_1 \subseteq C_0$ ?}
-
-  % Now consdider some $(i, j) \in D_{00}$ (the same reasoning as for $C_0$ is sufficient to see that $D_{00}$ is non-empty).
-Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).
-Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in P_0 \Rightarrow u \in S_0 \Rightarrow j \not\in C_0$, a contradiction.
-This implies every tuple in $D_{00}$ is a bona-fide edge of $\gcg$, as well as that $(i, j) \in A_1$.
-So then $A_1 = D_{00}$, [\hl{nonsensical, see earlier.}] completing the base case for induction.
-
-  % Suppose now the algorithm has reached step $k$ and that $E = \bigcup_{\ell \le k - 1}A_\ell$.
-
-% The reasoning here is (I think) correct -- but isn't directly establishing the induction hypothesis.
-%   On the other hand, if $P_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $P_0$.
-Consider now an edge $(i, j) \in D_{00}$ ($D_{00}$ is non-empty for the same reason as $C_0$).
-Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).
-Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in P_0 \Rightarrow u \in S_0 \Rightarrow j \not\in C_0$, a contradiction.
-So every tuple in $D_{00}$ is a bona-fide edge of $\gcg$.
-Suppose now that for some $j \in C_0$ there is an $i$ s.t.
-$(i, j) \in \gcg$ but $(i, j) \not\in D_{00}$.
-Since $(i, j) \in \gcg$ we have $i \pwgc j$ (corollary \ref{cor:gc_implies_pwgc}), but then if $i \not\in P_0$ then $j \notin C_0$ and if $i \in P_0, j \in C_0$ then necessarily $(i, j) \in D_{00}$ since we cannot at this stage have any paths in $E$.
-
-%   and show that on each step, every edge incident upon a node in $C_k$ is contained in $D_k$, that every edge in $D_k$ is an edge in $\gcg$, and finally that $\bigcup_{k = 1}^n C_k = [n] \setminus P_0$ and that this implies $\big([n], \bigcup_{k = 1}^n D_k\big) = \gcg$.
-
-%  On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E_0 = \emptyset$.
-The set $P_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $P_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
-Now, if $P_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
-On the other hand, if $P_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $P_0$.
-Consider now an edge $(i, j) \in D_0$ (again $D_0$ is non-empty for the same reason as $C_0$).
-Since $i$ has no parents, we must have $\anc{i}\cap\anc{j} = \emptyset$ and therefore $i \in \anc{j}$, and due to the construction of $C_0$ we in fact have $i \in \pa{j}$, which establishes that every edge in $D_0$ is also an edge in $\gcg$.
-Finally consider $j \in C_0$, by proposition \ref{prop:pwgc_anc} and since the parents of $C_0$ must be in $P_0$, we see that every edge incident on $j$ must be present in $D_0$.
-
-% We also see that since nodes in $P_0$ have no parents,
-
-% we must have $\anc{i}\cap\anc{j} = \emptyset$, and therefore by proposition \ref{prop:ancestor_properties} $i \in \anc{j}$;  moreover, the construction of $P_1$ implies that in fact $i \in \pa{j}$, so $(i, j)$ is an edge of $\gcg$.
-
-%Finally, by similar reasoning, there cannot be an edge $(i, j) \in D_0$ which is not in $\gcg$, since $\anc{i} \cap \anc{j} = \emptyset$.
-
-%Consider now step $K < n$ of algorithm \ref{alg:pwgr}.
-By lemma \ref{lem:still_strongly_causal} the graph remains strongly causal after having removed nodes $\bigcup_{k = 0}^K P_k$ from $S$, and we assume for induction that $\bigcup_{k = 1}^K D_k$ are all edges of $\gcg$.
-
-% \hl{WIP}
-% \end{proof}
-
-\begin{example}
-  We close this section by noting that the conditions of persistence
-  and strong causality are only sufficient conditions.  For example,
-  the complete directed graph with 2 nodes i.e.
-
-  \begin{equation*}
-    B(1) = \left[ \begin{array}{cc} 1/2 & 1 \\ 1 & 1/2 \end{array}\right]
-  \end{equation*}
-
-  contains a loop but is pairwise recoverable, though not by algorithm
-  (\ref{alg:pwgr}).  Clearly, this example is somewhat artificial
-  since when $n = 2$ there is no difference between pairwise
-  Granger-causality and joint Granger-causality amongst all series --
-  however, one can add any number of nodes having no parents or
-  children to a graph containing a length 2 cycle, in which case the
-  graph clearly remains pairwise recoverable.
-\end{example}
-
-\section{Finite Sample Graph Recovery}
+\section{Finite Sample Implementation}
 \label{sec:structure_learning}
@@
 graph with the residuals of the previous model as input, and then
-refitting on the combined sparsity pattern.  We experiment with this
-heuristic in our example application of Section \ref{sec:application},
-but reserve theoretical analysis for future work.
+refitting on the combined sparsity pattern.  We intend to consider
+this heuristic in future work.
 
@@
 
-% \section{Structure Learning}
-% \label{sec:structure_learning}
-% \subsection{Local Search Heuristics}
-
-% \subsection{Edge-Wise Grouped LASSO}
-% Minimize the sparsity promoting regularized least squares problem and choose the hyper-parameters against the Akaike information criteria.
-
-% \begin{equation}
-%   L(\lambda, p) \defeq \underset{B}{\text{minimize}}\ \frac{1}{T}\sum_{t = 1}^T||x(t) - \sum_{\tau = 1}^pB(\tau)x(t - \tau)||_2^2 + \lambda \sum_{i, j}\big[\alpha||B_{i, j}||_2 + (1 - \alpha)||B_{i, j}||_1\big]
-% \end{equation}
-
-% \begin{equation}
-%   \underset{\lambda, p}{\text{minimize}}\ L(\lambda, p) + \mathsf{AIC}(B^{(\lambda, p)})
-% \end{equation}
-
-% \subsection{Bayesian Posterior Thresholding}
-% Consider the Bayesian model
-
-% \begin{equation}
-%   \begin{aligned}
-%     (x(t)\ |\ B, \sigma_v^2, \{x(t - \tau)\}_{\tau = 1}^p) &\sim \mathcal{N}(\sum_{\tau = 1}^pB(\tau)x(t - \tau), \sigma_v^2)\\
-%     (B_{ij}(\tau)\ |\ G_{ij}(\tau)) &\sim G_{ij}(\tau)\mathcal{N}(0, \sigma_\beta^2) + (1 - G_{ij}(\tau))\delta_0(B_{ij}(\tau))\\
-%     (G_{ij}(\tau)) &\sim \mathsf{BER}(p_{ij}(\tau))\\
-%     \sigma_v^2 &\sim \Gamma(a_v, b_v)\\
-%     \sigma_\beta^2 &\sim \Gamma(a_\beta, b_\beta)
-%   \end{aligned}
-% \end{equation}
-
-% Similar models have been studied by various authors in the context of the linear regression model \hl{[(cite them)]} where it is referred to as \textit{stochastic variable selection}, or referred to as a model for Bayesian variable selection.
-
-% Since we have in mind applications where $n$ is large, sampling posterior probabilities can be prohibitively burdensome, so we can instead fit the mean-field variational approximation $p(B, G, \sigma\ |\ X) \approx q_Gq_Bq_\sigma$ and subsequently apply a thresholding operation to the posterior edge inclusion probabilities under $q_G$.
-
-% \subsection{Pairwise Minimum Error Spanning Trees}
-% Inspired by theorem \ref{scg_recovery}, we propose the following
-% structure learning heuristic.
-
-% %% This algorithm should perform coordinate descent or jointly refit all the trees.
-
-% \begin{enumerate}
-%   \item{Set $k = 0$ and $x^0(t) = x(t)$}
-%   \item{Compute all pairwise causality measures $\Xi \defeq \big[\ln \frac{\xi_i}{\xi_{ij}} \big]_{i, j}$}
-%   \item{Find the maximum spanning arborescence\footnote{An ``arborescence'' is a french word for a tree diagram, and refers to a \textit{directed} tree in graph theory} $\mathcal{T}_k$ of a graph having edge weights $\Xi_{ij}$.}
-%   \item{Fit a vector LTI filter $F_k$ having graph defined by $\mathcal{T}_k$ to $x^{k - 1}(t)$ and set $x^k(t) = x^{k - 1}(t) - \hat{x}^{k - 1}(t)$}
-%   \item{Repeat over $k$ until a stopping criteria is met}
-% \end{enumerate}
-
-% \subsection{Layered Strongly Causal Graphs}
-% It is clearn that Algorithm \ref{alg:pwgr} is a purely theoretical construct, and cannot be applied directly to real data.
-Instead, we convert the algorithm into an heuristic and observe it's performance experimentally.
-We first consider the heuristic version of Algorithm \ref{alg:pwgr} in Algorithm \ref{alg:pwgr_heuristic} and show that is a well defined procedure (note that $\ceil{x}$ denotes the integer ceiling function):
-
-% \begin{proposition}
-%   Algorithm \ref{alg:pwgr_heuristic} terminates in at most $n$ iterations and returns a strongly causal graph $\widehat{\gcg}$.
-% \end{proposition}
-% \begin{proof}
-%   \hl{TODO}
-% \end{proof}
-
-% A complete estimation procedure for a sparse filter $\widehat{\B}(z)$
-% used to estimate $x(t + 1)$ from $\H_{t - 1}$ is then given in Algorithm \ref{alg:Bz_estimate}.
-
-% \begin{proposition}
-%   The error sequence in Algorithm \ref{alg:Bz_estimate} is a zero mean
-%   WSS process $\E\epsilon^{(m)}(t) = 0$ with monotonically decreasing
-%   variance: $\E|\epsilon^{(m + 1)}(t)|^2 \le \E|\epsilon^{(m)}(t)|^2$.
-% \end{proposition}
-% \begin{proof}
-%   \hl{TODO -- this is probably not even true}
-% \end{proof}
-
-% \begin{algorithm}
-%   \SetKwInOut{Input}{input}
-%   \SetKwInOut{Output}{output}
-%   \SetKwInOut{Initialize}{initialize}
-%   \DontPrintSemicolon
-
-%   \BlankLine
-%   \caption{Graph Recovery Heuristic $\mathsf{pw\_scg}(F, P, \alpha)$}
-%   \label{alg:pwgr_heuristic}
-%   \Input{Estimates of pairwise Granger-causaltiy statistics $F_{ij}$ (eqn \ref{eqn:gc_statistics}), Matrix of edge probabilities $P_{ij}$ (eqn \ref{eqn:edge_inclusion_probability}), threshold $\delta$.}
-%   \Output{A strongly causal graph $\widehat{\gcg}$}
-%   \Initialize{$S = [n]$  \texttt{\# unprocessed nodes}\\
-%     $E = \emptyset$  \texttt{\# edges of }$\widehat{\gcg}$\\
-%     $k = 1$ \texttt{\# a counter used only for notation}}
-
-%   \BlankLine
-
-%   $W_\delta \leftarrow \{(i, j)\ |\ P_{ji} > 1 - \delta, F_{ji} > F_{ij}\}$  \texttt{\# candidate edges}\\
-%   $\mathcal{I}_0 \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# total node incident probability}\\
-%   $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) < \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Nodes with fewest incident edges}\\
-%   \If{$P_0 = \emptyset$}{
-%     $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) \le \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Ensure non-empty}
-%   }
-%   \BlankLine
-
-%   \While{$S \ne \emptyset$}{
-%     $S \leftarrow S \setminus P_{k - 1}$ \texttt{\# remove processed nodes}\\
-%     % $\mathcal{I}_k \leftarrow \big(\sum_{j \in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# intra-}$S$ \texttt{incident strength}\\
-%     $\mathcal{I}_k \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$\\
-%     $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) < \ceil{\text{min}(\mathcal{I}_k)}\}$\\
-%     \If{$P_k = \emptyset$}{
-%       $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) \le \ceil{\text{min}(\mathcal{I}_k)}\}$
-%     }
-%     \;
-%     \texttt{\# add strongest edges, maintaining strong causality}\\
-%     $U_k \leftarrow \bigcup_{r = 1}^k P_{k - r}$ \texttt{\# Include all forward edges}\\
-%     \For{$(i, j) \in \mathsf{sort}\Big(\{(i, j) \in U_k \times P_k\ |\ (i, j) \in W_\delta\} \mathsf{\ by\ descending\ } F_{ji}\Big)$} {
-%       \If{$\mathsf{is\_strongly\_causal}(E \cup \{(i, j)\})$} {
-%         $E \leftarrow E \cup \{(i, j)\}$
-%       }
-%     }
-%     $k \leftarrow k + 1$\\
-%   }
-%   \Return{$([n], E)$}
-% \end{algorithm}
-
-% \begin{algorithm}
-%   \SetKwInOut{Input}{input}
-%   \SetKwInOut{Output}{output}
-%   \SetKwInOut{Initialize}{initialize}
-%   \DontPrintSemicolon
-
-%   \BlankLine
-%   \caption{Filter Estimator}
-%   \label{alg:Bz_estimate}
-%   \Input{$T$ samples from a WSS process $x: \{x(t)\}_{t = 1}^T$\\
-%     Maximum iterations $M$\\
-%     Sparsity parameters $\{(\delta_m\, b_m, R_m)\}_{m = 1}^M$}
-%   \Output{Linear one-step-ahead estimator $\widehat{\B}(z)$}
-%   \Initialize{$\epsilon^{(0)}(t) = x(t)$ \texttt{\# initial error process}\\
-%     $m = 1$ \texttt{\# iteration counter}}
-%   \BlankLine
-%   \For{$m = 1, \ldots, M$}{
-%     $F \leftarrow \mathsf{pwgc}(\epsilon^{(m - 1)}(t))$ \texttt{\# estimate pairwise Granger-causality statistics}\\
-%     $\widehat{\gcg}_m \leftarrow \mathsf{pw\_scg}(F, \delta_m, b_m, R_m)$ \texttt{\# graph estimate (Algorithm \ref{alg:pwgr_heuristic})}\\
-%     $\widehat{\B}^{(m)}(z) \leftarrow \mathsf{estimate\_B}(\epsilon^{(m - 1)}(t), \widehat{\gcg}_m)$ \texttt{\# Estimate filter with topology} $\widehat{\gcg}_m$\\
-%     $\epsilon^{(m)}(t) \leftarrow \epsilon^{(m - 1)}(t) - \widehat{\B}^{(m)}(z)\epsilon^{(m - 1)}(t)$ \texttt{\# updated error process}\\
-%   }
-%   \BlankLine
-%   $\widehat{\B}(z) \leftarrow \widehat{\B}^{(M)}(z) \prod_{k = 1}^{M - 1}\big(I - \widehat{\B}^{(M - k)}(z)\big)$ \texttt{\# compute total filter}\\
-%   \Return{$\widehat{\B}(z)$}
-% \end{algorithm}
-
-\section{Empirical Evaluation}
-\label{sec:empirical_evaluation}
+\section{Simulation}
+\label{apx:simulation}
 We have implemented our empirical experiments in Python \cite{scipy},
@@
 \texttt{networkx} \cite{networkx}.  We run experiments using two
-separate graph topologies having $n$ nodes: a strongly causal graph
-(SCG) and a directed acyclic graph (DAG).  These are generated
+separate graph topologies having $n = 50$ nodes. These are generated
 respectively by drawing a random tree and a random Erdos Renyi graph
-(with edge probability $q = \frac{2}{n}$ resulting in approximately
-the same number of edges for the SCG as for the DAG), then creating a
-directed graph by directing edges from lower numbered nodes to higher
-numbered nodes.
+then creating a directed graph by directing edges from lower numbered
+nodes to higher numbered nodes.
 
@@
 driven by i.i.d. Gaussian random noise, each component having random
-variance $\sigma_i^2 = 1/2 + r_i$ where $r_i \sim \text{exp}(1/2)$.
-We set $p_{\text{max}} = 15$.  Results and representative graphs are
-collected in Figures \ref{fig:random_graph_topologies},
+variance $\sigma_i^2 = 1/2 + r_i$ where $r_i \sim \text{exp}(1/2)$.  To ensure
+we are generating data from a stationary system, we first discard
+samples during a long burnin period.
+
+For both PWGC and adaLASSO We set the maximum lag length
+$p_{\text{max}} = 10$.
+
+Results are collected in Figures
 \ref{fig:simulation_results_comparison1},
@@
 \ref{fig:simulation_results_dense}.
-
-We compare our results against the adaptive LASSO
-\cite{adaptive_lasso_zou2006}, which outperformed substantially both
-the LASSO and the grouped LASSO.  Motivated by scaling, we split the
-squared error term into separate terms, one for each group of incident
-edges on a node, and estimate the collection of $n$ incident filters
-$\big\{\B_{ij}(z)\big\}_{j = 1}^n$ that minimizes
-$\xi_i^{\text{LASSO}}$ in the following:
-
-\begin{equation}
-  \begin{aligned}
-  \xi_i^{\text{LASSO}}(\lambda) &= \underset{B}{\text{min}}\ \frac{1}{T}\sum_{t = p + 1}^T\big(x_i(t) - \sum_{\tau = 1}^p\sum_{j = 1}^n B_{i, j}(\tau) x(t - \tau)\big)^2 + \lambda \sum_{\tau = 1}^p \sum_{j = 1}^n |B_{ij}(\tau)|\\
-  \xi_i^{\text{LASSO}} &= \underset{\lambda \ge 0}{\text{min}}\ \xi_i^{\text{LASSO}}(\lambda) + \mathsf{BIC}\big(B_i^{\text{LASSO}}(\lambda)\big)\\
-  \end{aligned}
-\end{equation}
-
-where we are choosing $\lambda$, the regularization parameters, via the BIC.
-
-\begin{figure}
-  \centering
-  \caption{Representative Random Graph Topologies}
-  \label{fig:random_graph_topologies}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \caption{Random SCG}
-    \includegraphics[width=\linewidth]{example_scg.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \caption{Random DAG $(q = \frac{2}{n})$}
-    \includegraphics[width=\linewidth]{example_dag.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \caption{Random DAG $(q = \frac{4}{n})$}
-    \includegraphics[width=\linewidth]{example_dag_q.pdf}
-  \end{subfigure}
-\end{figure}
-
-\begin{remark}[Graph Topologies]
-  We depict in Figure \ref{fig:random_graph_topologies} the topologies
-  of random graphs used in our empirical evaluation.  For values of
-  $q$ close to $\frac{2}{n}$, the resulting random graphs tend to have
-  a topology which is, at least qualitatively, close to the SCG.  As
-  the value of $q$ increases, the random graphs deviate farther from
-  the SCG topology, and we therefore expect the LASSO to outperform
-  PWGC for larger values of $q$.  This can be observed in Figure
-  \ref{fig:simulation_results_dense}, at least where performance is
-  measured by the support recovery (i.e. via the MCC).
-\end{remark}
-
-\begin{remark}[MCC as a Support Recovery Measurement]
-  We apply Matthew's Correlation Coefficient (MCC)
-  \cite{matthews1975comparison} as a statistic for measuring support
-  recovery performance.  This statistic synthesizes the confusion
-  matrix into a single score appropriate for unbalanced labels and is
-  calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
-  performance, $0$ being the performance of random guessing, and $-1$
-  being perfectly opposed.
-\end{remark}
-
-\begin{remark}[Error Measurement]
-  We estimate the 1-step ahead prediction error by forming the variance matrix estimate
-
-  \begin{equation*}
-    \widehat{\Sigma}_v \defeq \frac{1}{T_{\text{out}}} \sum_{t = 1}^{T_{\text{out}}} (x(t) - \widehat{x}(t))(x(t) - \widehat{x}(t))^\T
-  \end{equation*}
-
-  on a long stream of out-of-sample data.  We then report the quantity
-
-  \begin{equation*}
-    \frac{\ln \tr \widehat{\Sigma}_v}{\ln \tr \Sigma_v}
-  \end{equation*}
-
-  where $\widehat{\Sigma}_v = \Sigma_v$ is the best possible performance.
-\end{remark}
 
@@
 
-  {\scriptsize Comparison of PWGC and LASSO for $\VAR(p)$ model
+  {\footnotesize Comparison of PWGC and LASSO for $\VAR(p)$ model
     estimation.  We make comparisons against both the MCC and the
@@
 
-  {\scriptsize Figure \ref{fig:simulation_results_comparison2}
+  {\footnotesize Figure \ref{fig:simulation_results_comparison2}
     provides results for systems which do not guarantee the
@@
 
-  {\scriptsize Figure \ref{fig:simulation_results_scaling} measures
+  {\footnotesize Figure \ref{fig:simulation_results_scaling} measures
     support recovery performance as the number of nodes $n$ increases,
@@
 
-  {\scriptsize Figure \ref{fig:simulation_results_dense} provides a
+  {\footnotesize Figure \ref{fig:simulation_results_dense} provides a
     comparison between PWGC and AdaLASSO as the density of graph edges
@@
 better than the LASSO for the case of a strongly causal graph, since
-in this case the assumptions which guarantee the correctness of
-Algorithm \ref{alg:finite_pwgc} hold.  However, the performance is
-still markedly superior in the case of a more general DAG.  We would
-conjecture that a DAG having a similar degree of sparsity as an SCG is
-likely to be ``close'' to an SCG.  Figure
-\ref{fig:simulation_results_dense} illustrates the severe (expected)
-degradation in performance as the number of edges increases while
-the number of data samples $T$ remains fixed.  For larger values $q$
-in this plot, the number of edges in the graph is comparable to the
-number of data samples.
+in this case the theory from which our heuristic derives is valid.
+However, the performance is still markedly superior in the case of a
+more general DAG.  We would conjecture that a DAG having a similar
+degree of sparsity as an SCG is ``likely'' to be ``close'' to an SCG,
+in some appropriate sense.
+
+Figure \ref{fig:simulation_results_dense} illustrates the severe
+(expected) degradation in performance as the number of edges increases
+while the number of data samples $T$ remains fixed.  For larger values
+$q$ in this plot, the number of edges in the graph is comparable to
+the number of data samples.
 
@@
 
-\section{Application}
-\label{sec:application}
-In this section we apply our methods to a real set of EEG data
-obtained from the ``EEG Database Data
-Set''\footnote{http://archive.ics.uci.edu/ml/datasets/EEG+Database}
-\cite{zhang1995event} on the UCI machine learning repository
-\cite{uci_mlr}.  This dataset contains 1 second long measurements of
-(64 channel) EEG signals from patients who are given visual stimuli.
-The subjects in the study are labeled as being either ``control'' or
-``alcoholic'', however, we will ignore this label and instead focus on
-\textit{distinguishing between subjects} based on the
-Granger-causality graphs inferred from the subject's trials.  Our
-reasoning is to focus on the underlying question: ``Does our PWGC
-algorithm uncover meaningful Granger-causal connections from EEG
-data?''.  Focusing only on the subject label allows us to answer this
-question without simultaneously grappeling with the physiological
-question of whether alcoholic subjects as a group have discernable
-differences in their EEG readings\footnote{Some exploratory analysis
-  actually suggests that alcoholic Granger-causality graphs are not
-  substantively different}, which is a stronger requirement than
-simply that there are meaningful distinctions between the EEG readings
-of subjects generally.
-
-The dataset consists of
-$\mathcal{D} = \{\big(x^{(i)}(t)\big)_{t = 1}^T, y^{(i)} \}_{i = 1}^N$
-where $T = 256$, $x^{(i)}(t) \in \R^{64}$,
-$y^{(i)} \in [N_{subjects}]$ (with $N_{subjects} = 119$), and
-$N = 10723$.  There are on average 90 trials for each subject,
-ranging between 30 and 119.
-
-We have constructed a simple pipeline for discriminating between
-subjects by first applying an iterative PWGC algorithm (see Algorithm
-\ref{alg:iterated_pwgc}) directly to the EEG data $x^{(i)}$ to obtain
-a Granger-causality graph $\widehat{G}^{(i)}$.  We then feed the
-vectorized adjacency matrix of $\widehat{G}^{(i)}$ through a
-polynomial Kernel multinomial logistic regression model with
-parameters fit by cross validation.
-
-\begin{remark}
-  The iterated application of PWGC is an example of a heuristic by
-  which graph estimates that aren't constrained to being strongly
-  causal can be obtained.  We defer to future work the theoretical
-  analysis of more sophisticated heuristics, or algorithms appropriate
-  for different topological assumptions.
-\end{remark}
-
-\begin{algorithm}
-  \SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{output}
-  \SetKwInOut{Initialize}{initialize}
-  \DontPrintSemicolon
-
-  \BlankLine
-  \caption{Iterated PWGC Heuristic}
-  \label{alg:iterated_pwgc}
-
-  \Input{Data $x(t)$, maximum number of iterations $N$,
-    threshold $r \in (0, 1)$.}
-  \Output{Granger-causality graph estimate $\widehat{G}$ obtaind from
-    iterated application of PWGC Algorithm \ref{alg:finite_pwgc}}
-  \Initialize{$i = 1$, $\widehat{x}_0(t) \leftarrow x(t)$,
-    $\sigma_0 \leftarrow Var\ x(t)$}
-
-  \While{$i \le N$} {
-    Estimate $\VAR$ model via PWGC on $x_{i - 1}(t)$ to obtain $\widehat{x}_{i - 1}(t)$ and $\widehat{G}_{i - 1}$\\
-    $\epsilon_i(t) \leftarrow x_{i - 1}(t) - \widehat{x}_{i - 1}(t)$ \texttt{\# Compute residuals}\\
-    $\sigma_i^2 \leftarrow \text{MSE}\big(\epsilon_i(t)\big)$ \texttt{\# Compute MSE}\\
-    \If{$\sigma_i^2 > r \sigma_{i - 1}^2$}{
-      \texttt{break  \# Insignificant Improvement}
-    }
-    $i \leftarrow i + 1$
-  }
-  $\widehat{G} \leftarrow \bigcup_{i} \widehat{G}_{i}$ \texttt{\# Combine PWGC graph estimates}\\
-  \Return{$\widehat{G}$}
-\end{algorithm}
-
-While it is almost certainly possible to achieve much greater
-classification accuracy on this dataset by constructing a classifier
-to act directly on $x^{(i)}(t)$, our purpose is to demonstrate that
-relevant information is being captured by the graph estimates
-$\widehat{G}^{(i)}$.  That is, that the latent structure
-$\widehat{G}^{(i)}$ may provide scientifically relevant insights, as
-opposed to simply being a feature for classification tasks.
-
-Figures \ref{fig:ar_eeg_example} and \ref{fig:eeg_oos_error}
-illustrate the appropriateness of modelling $x(t)$ with $\VAR$ models,
-as opposed simply to a collection of $n$ unidimensional
-autoregressiive models.  The conclusion being that the
-Granger-causality graph estimates are not purely spurious.
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{$\mathsf{AR}(p)$ Model Example}
-    \label{fig:ar_eeg_example}
-    \includegraphics[width=\linewidth]{eeg_estimate_example.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{OOS Error for a Single Subject}
-    \label{fig:eeg_oos_error}
-    \includegraphics[width=\linewidth]{eeg_oos_demonstration.pdf}
-  \end{subfigure}
-
-  {\scriptsize Figure \ref{fig:ar_eeg_example} provides an example of
-    an autoregressive model fir via the Levinson-Durbin algorithm
-    (choosing the system order via the BIC) on a single EEG channel.
-    Linear autoregressive models appear qualitatively to be adequate
-    for this data.  Figure \ref{fig:eeg_oos_error} demonstrates the
-    improvement of a unified $\VAR$ model over independent
-    $\mathsf{AR}$ models, providing evidence that intra-node edges
-    inferred by PWGC are not simply spurious.
-
-  Out-of-Sample estimates are performed on data that were not used in
-  fitting the models.}
-\end{figure}
-
-We provide the final results in Figure
-\ref{fig:logistic_regression_results} where the classification
-accuracy is quantified by the (multiclass) MCC of $\approx 0.20$ on a held
-out validation set consisting of $20\%$ of the data.  This MCC score
-exceeds by a modest margin the performance of randomly guessing,
-lending strong evidence to the assertion that PWGC successfully
-recovers some meaningful differences between the Granger-causality
-graphs of different subjects, particularly if we recall that there are
-generally fewer examples for each subject than there are subjects
-overall.  Moreover, distinguishing only between two particular
-subjects (as opposed to distinguishing between one subject and $118$
-others) is substantially easier -- an illustration is provided in
-Figure \ref{fig:logistic_regression_2pair} where we embed the data
-into the plane through supervised dimensionality reduction.  In
-general, fitting a simple classifier between two subjects achieves an
-$MCC$ in excess of $0.6$ and up to $0.9$, even when the same
-hyperparameters are carried over for different pairs of subjects.
-
-\begin{figure}
-  \centering
-  \caption{Subject Classification from Granger-causality Graphs}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{Distinguishing Individual Subjects}
-    \label{fig:logistic_regression_2pair}
-    \includegraphics[width=\linewidth]{logistic_regression_separation.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{Complete Multiclass Classifier}
-    \label{fig:logistic_regression_results}
-    \includegraphics[width=\linewidth]{logistic_regression_confusion_matrix.pdf}
-
-    {\scriptsize }
-  \end{subfigure}
-
-  {\scriptsize Figure \ref{fig:logistic_regression_2pair} illustrates
-    discriminating between two particular subjects based on their
-    EEG Granger-causality graphs.  In this case $MCC = 0.75$ on held
-    out data.  Visualization is constructed by supervised
-    dimensionality reduction and is purely illustrative.
-
-    Figure \ref{fig:logistic_regression_results} provides the
-    row-normalized confusion matrix (computed on held out validation
-    data) of a multiclass logistic regression classifier used to
-    classify subjects based on their EEG Granger-causality graphs.  $MCC = 0.20$}
-
-\end{figure}
-
-\section{Conclusion}
-\label{sec:conclusion}
-In this paper we have argued that considering particular topological
-properties of Granger-causality networks can provide substantial
-theoretical insights as well as computational benefits.  In
-particular, the notion of a strongly-causal graph has been exploited
-to establish conditions under which pairwise causality testing alone
-is sufficient for recovering a complete Granger-causality graph.
-Moreover, examples from the literature suggest that such topological
-assumptions may be reasonable in some applications and secondly, even
-when the strong-causality assumption is not met, we have provided
-simulation evidence to suggest that our pairwise testing algorithm
-PWGC can still outperform the LASSO and adaLASSO, both of which are
-commonly employed in applications.  Application to inference of more
-general graphs may also be within reach via an iterative graph
-building process as suggested in Section \ref{sec:application}.
-
-Our application in Section \ref{sec:application} provides strong
-evidence that PWGC is capable of uncovering meaningful features from
-networks of time series data by constructing a classifier which
-accurately discriminates between subjects in an EEG experiment based
-purely on the Granger-causality graph topology inferred by the
-iterated application of PWGC.
-
-We emphasize that the causality graph topology is one of the key
-defining features of time series analysis in comparison to standard
-multivariate regression and therefore advocate for further study of
-how different topological assumptions may impact the recovery of
-causality graphs.  For example, are there provable guarantees on the
-error rate of PWGC when applied to non strongly-causal graphs?
-Computationally, can constraint systems or cunning adaptive weighting
-schemes impose useful prior knowledge about graph topology for the
-LASSO algorithm?  In another direction, it is known that the more
-general notion of transfer entropy reduces to Granger-causality in the
-case of Gaussian data \cite{barnett2009granger}, do our results extend
-to causality networks defined via transfer entropy?  Finally, the work
-of \cite{barnett2015granger} has established the superiority of
-Granger-causality testing in state space models (as opposed to pure
-autoregressions) in many cases.  Combining this work with our PWGC
-algorithm (by modifying the approach described in Section
-\ref{sec:pairwise_hypothesis_testing} to instead utilize state-space
-Granger-causality testing) therefore is likely to enable application
-to very large networks of time series data which are not well
-approximated by finite $\VAR(p)$ models.
-
 \clearpage
-\printbibliography
+\bibliography{\string~/Documents/academics/global_academics/global_bib}
+
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2019-07-17 software/pwgc/gc_methods.py

@@ -88,6 +88,12 @@ def _estimate_b_lasso(G, X, y):
     return w
 
 
+n, p = 1000 + int(1e4), int(1e4)
+b = np.random.normal(size=p) * np.random.binomial(1, 0.2, size=p)
+X = np.random.normal(size=(n, p))
+y = X @ b + 0.1 * np.random.randn(n)
+
+
 def _standardize_X(f):
     def _f(X, *args, **kwargs):
         _X = X - np.mean(X, axis=0)[None, :]
@@ -567,7 +573,7 @@ def _par_fast_compute_pairwise_gc(n, T, R, k_cores=2):
 
 # TODO: P_j should be an n x n array
 # TODO: Tune alpha by Benjamini Hochberg criterion
-def pw_scg(F, P_edge, alpha):
+def pw_scg(F, P_edge, alpha, eliminate_bidirectional=False):
     """
     Graph recovery heuristic
 
@@ -635,6 +641,7 @@ def pw_scg(F, P_edge, alpha):
 
     n = F.shape[0]
     S = set(range(n))
+
     # Larger value means more confident about an edge
     P_edge[P_edge < 1 - alpha] = 0
 
@@ -644,8 +651,15 @@ def pw_scg(F, P_edge, alpha):
 
     # ------------- Initialization -------------
     # Set of candidate edges
-    W_set = {(i, j) for (i, j) in product(S, S)
-             if (F[j, i] >= F[i, j] and P_edge[j, i] > 0)}
+    if eliminate_bidirectional:
+        W_set = {(i, j) for (i, j) in product(S, S)
+                 if (F[j, i] >= F[i, j] and
+                     P_edge[j, i] > 0 and
+                     P_edge[i, j] == 0)}
+    else:
+        W_set = {(i, j) for (i, j) in product(S, S)
+                 if (F[j, i] >= F[i, j] and
+                     P_edge[j, i] > 0)}
 
     # Predecessor edges of a node, i.e. W[i] = {j | (j, i) \in W}
     W_pred = {i: [j for j in S if (j, i) in W_set]
@@ -799,16 +813,20 @@ def get_residual_graph(G_hat):
     return G_hat
 
 
-def estimate_dense_graph(X, max_lag=10,
-                         max_T=None, method="lasso"):
+def estimate_dense_graph(X, max_lag=10, max_T=None, method="lasso",
+                         post_ols=False):
     # TODO: Replace this with ts_lasso
 
     _, n = X.shape
[…]

### 2019-07-18 paper_theory/paper_theory.tex
Better proofs.

@@
 
+% Should I do this?  It makes cross references much easier,
+% but is possibly extra complication.
+% \counterwithin{theorem}{section}
+% \counterwithin{corollary}{section}
+% \counterwithin{proposition}{section}
+% \counterwithin{lemma}{section}
+
+% \counterwithin{remark}{section}
+% \counterwithin{definition}{section}
+% \counterwithin{example}{section}
+% \counterwithin{assumptions}{section}
+
 % keywords
@@
   from the gene regulatory network literature suggesting that our
-  concept of a strongly causal graph may be applicable to this field.}
+  concept of a strongly causal graph may be applicable to this field.
+  We implement and detail finite-sample heuristics derived from our
+  theory, and establish through simulation the efficiency gains (both
+  statistical and computational) which can be obtained (in comparison
+  to LASSO-type algorithms) when structural assumptions are met.}
 
@@
 Cette recherche a été financée par le Conseil de recherches en sciences naturelles et en génie du Canada (CRSNG), [numéro de référence 518418-2018].
-
-% \clearpage
-
-% % \tableofcontents
-% \clearpage
 
@@
 uncovering an underlying causal structure in multivariate time series.
-Though the underlying causality graph cannot be observed directly, we
-will infer it's presence as a latent structure among our observed time
-series data.  This notion is leveraged in a variety of applications
-e.g. in Neuroscience as a means of recovering interactions amongst
-brain regions \cite{bressler2011wiener}, \cite{anna_paper2008},
-\cite{david2008identifying}; in the study of the dependence and
-connectedness of financial institutions \cite{NBERw16223}; gene
-expression networks \cite{Fujita2007},
+Though the underlying causality graph cannot be observed directly,
+it's it's presence is inferred as a latent structure among observed
+time series data.  This notion is leveraged in a variety of
+applications e.g. in Neuroscience as a means of recovering
+interactions amongst brain regions \cite{bressler2011wiener},
+\cite{anna_paper2008}, \cite{david2008identifying}; in the study of
+the dependence and connectedness of financial institutions
+\cite{NBERw16223}; gene expression networks \cite{Fujita2007},
 \cite{methods_for_inferring_gene_regulatory_networks_from_time_series_expression_data},
@@
 general an extremely challenging problem (i.e. it may be framed as a
-best subset selection problem, see \cite{bss_mio},
-\cite{hastie_bss_comp}), moreover, the comparison of quality between
-different structures, and hence the notion of ``best'' needs
+best subset selection problem, see, \cite{hastie_bss_comp} for recent
+improvements in BSS methods), moreover, the comparison of quality
+between different structures, and hence the notion of ``best'' needs
 qualification.  In applications where we are interested merely in
@@
 local search heuristic to the Whittle likelihood with an AIC
-penalization.  The local search heuristic where at each iteration an
-edge is either added, removed, or reversed is a common approach to
+penalization.  The local search heuristic is a common approach to
 combinatorial optimization due to it's simplicity, but is liable to
@@
 established under less restrictive conditions than for the vanilla
-LASSO.  We study a novel perspective on the problem by establishing
-conditions under which simple \textit{pairwise} Granger-causality
-tests are sufficient to recover the entire graph.
-
-%   Prior assumptions
-% on the \textit{topology} of the underlying Granger-causality graph can
-% result in tests more powerful than more generally applicable
-% algorithms, and we provide examples from the literature where
-% simplifying topological assumptions may be met.
-
-% \subsection{Contributions}
+LASSO.
+
+% We study a novel perspective on the problem by establishing conditions
+% under which simple \textit{pairwise} Granger-causality tests are
+% sufficient to recover the entire graph.  Without particular
+% topological assumptions, pairwise testing naturally results in a large
+% number of false positives \cite{tam2013gene_pwgc}.
+
 In the context of time series data, sparsity assumptions remain
@@
 arise as a result of considering the topology of the underlying
-Granger-causality graph, which to our knowledge remains largely
-unexplored.  The focus of this paper is to shed light on some of these
-topological questions, in particular, we study a particularly simple
-notion of causality graph topology which we term ``strongly causal''
-and show that stationary times series whose underlying causality graph
-has this structure satisfy natural intuitive notions of ``information
-flow'' through the graph.  Moreover, we show that such graphs are
-perfectly recoverable with only \textit{pairwise} Granger-causality
-tests, which would otherwise suffer from serious confounding problems.
+Granger-causality graph.  The focus of this paper is to shed light on
+some of these topological questions, in particular, we study a
+particularly simple notion of graph topology which we term
+``strongly causal'' and show that stationary times series whose
+underlying causality graph has this structure satisfy natural
+intuitive notions of ``information flow'' through the graph.
+Moreover, we show that such graphs are perfectly recoverable with only
+\textit{pairwise} Granger-causality tests, which would otherwise
+suffer from serious confounding problems (see
+e.g. \cite{tam2013gene_pwgc} for earlier work on pairwise testing).
 Aside from being an interesting theoretical perspective, prior
@@
 both theory and practice.  Concluding remarks on further open problems
-and extentions are provided in Section \ref{sec:conclusion}.  The
+and extensions are provided in Section \ref{sec:conclusion}.  The
 proofs of each proposition and theorem are also relegated to the
@@
   \begin{aligned}
-    % \H^x &= \cl \{\sum_{\tau = -T}^T a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, T \in \N\} \subseteq L_2(\Omega),\\
     \H_t^x &= \cl \{\sum_{\tau = 0}^p a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, p \in \N\} \subseteq L_2(\Omega)\\
@@
 analogous ways by particular components of $x$ as e.g.
-$\H_t^{(i, j)}$, $\H_t^{i}$ or by all but particular components as
+$\H_t^{(i, j)}$, $\H_t^{i}$ or by all but a particular component as
 $\H_t^{-j}$.
 
-As a consequence of the Wold decomposition theorem (see e.g. \cite{lindquist}),
-every WSS sequence has the moving average $MA(\infty)$
-representation
+As a consequence of the Wold decomposition theorem (see
+e.g. \cite{lindquist}), every WSS sequence has the moving average
+$MA(\infty)$ representation
 
@@
 
-where $c(t)$ is a purely deterministic sequence$v(t)$ is an
+where $c(t)$ is a purely deterministic sequence, $v(t)$ is an
 uncorrelated sequence and $A(0) = I$.  We will assume that $c(t) = 0$.
-We additionally require that this representation can be inverted to
-yield the $\VAR(\infty)$ form
+Given our setup, this representation can be inverted to yield the
+$\VAR(\infty)$ form
 
@@
   \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x$)
-  and write $x_j \gc x_i$ if given Hilbert spaces $\H_{t - 1}$,
-  $\H^{-j}_{t - 1}$
+  and write $x_j \gc x_i$ if
 
@@
 
-where $\xi[x \ |\ \H] = \E (x - \linE{x}{\H})^2$ is the mean squared
-estimation error and $\linE{x}{\H} = \text{proj}_{\H}(x)$ denotes the
-(unique) projection onto the Hilbert space $\H$.
+where $\xi[x \ |\ \H] \defeq \E (x - \linE{x}{\H})^2$ is the mean
+squared estimation error and $\linE{x}{\H} \defeq \text{proj}_{\H}(x)$
+denotes the (unique) projection onto the Hilbert space $\H$.
 \end{definition}
@@
 
-% \begin{proof}
-%   % Geweke uses the log of the ratio of the determinants of the residual variances
-%   % 
-%   % The equivalence $(1) \iff (3)$ is essentially a restatement of the
-%   % definition, but is in line with the seminal work of Geweke
-%   % \cite{geweke1982measurement}, \cite{geweke1984}.
-
-%   $(a) \Rightarrow (b)$ follows as a result of the uniqueness of orthogonal
-%   projection (i.e. the best estimate is necessarily the coefficients
-%   of the model).  $(b) \Rightarrow (c)$ follows since in computing
-%   $(y - \linE{y}{\H_{t - 1}^{-j}})$ for $y \in H_t^i$ it is sufficient
-%   to consider $y = x_i(t)$ by linearity, then since
-%   $H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$ we have
-%   $(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$ since
-%   $\B_{ij}(z) = 0$ and $v_i(t) \perp \H_{t - 1}$.  $(c) \iff (d)$ is a result
-%   of the equivalence in Definition
-%   \ref{lem:conditional_orthogonality_equivalence}.  And,
-%   $(d) \implies (a)$ follows directly from the Definition.
-% \end{proof}
-
 \subsection{Granger Causality Graphs}
@@
   itself.  An edge between a node and itself $(i, i) \in \gcge$ (which
-  is not a cycle) is referred to as a \textit{loop}.
+  we do not consider to be a cycle) is referred to as a \textit{loop}.
 
@@
 
-  Recall that we do not allow a node to be it's own parent, however
-  unless $\gcg$ is a DAG a node \textit{can} be it's own ancestor.  We will
-  ocassionally need to explicitly exclude $i$ from $\anc{i}$, in which
-  case we will write $\anc{i}\setminus \{i\}$.
+  Recall that we do not allow a node to be it's own parent, although
+  unless $\gcg$ is a DAG, a node \textit{can} be it's own ancestor.
+  We will occasionally need to explicitly exclude $i$ from $\anc{i}$,
+  in which case we will write $\anc{i}\setminus \{i\}$.
 \end{definition}
@@
 The following propositions are essentially lemmas used for the proof
-of the upcomming Proposition \ref{prop:ancestor_properties}, but
+of the upcoming Proposition \ref{prop:ancestor_properties}, but
 remain relevant for providing intuitive insight into the problems at
@@
   $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.  Moreover,
-  this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_{t - 1}^i} = 0$.
+  this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_t^i} = 0$.
 \end{proposition}
@@
   $i \in \anc{j}$, since $j$ cannot provide any information about $i$
-  that is not available from observing $i$ itself.  However, it is
-  important to note that this is the case only if we have direct
-  access to $x(t)$.  That is, if we instead observed a noisy version
-  of $x$ (i.e. $\tilde{x}(t) = x(t) + n(t)$) the observation of
-  $\tilde{x}_j$ would enable conditionally improved estimates of
-  $\tilde{x}_i$, without there necessarily being any causal flow in
-  the underlying system $x(t)$.
+  that is not available from observing $i$ itself.
+  % However, it is important to note that this is the case only if we
+  % have direct access to $x(t)$.  That is, if we instead observed a
+  % noisy version of $x$ (i.e. $\tilde{x}(t) = x(t) + n(t)$) the
+  % observation of $\tilde{x}_j$ would enable conditionally improved
+  % estimates of $\tilde{x}_i$, without there necessarily being any
+  % causal flow in the underlying system $x(t)$.
 \end{remark}
@@
   $k \in \anc{i} \cap \anc{j}$ and there exists a path
-  $\gcgpath{k}{i}$ not containing $j$, and a path $\gcgpath{k}{j}$
-  not containing $i$.
-
-  A simple example is furnished by the ``fork'' graph
+  $\gcgpath{k}{i}$ not containing $j$, and a path $\gcgpath{k}{j}$ not
+  containing $i$. A simple example is furnished by the ``fork'' graph
   $i \leftarrow k \rightarrow j$.
@@
 
-\begin{remark}
-  The interpretation of this proposition is that for $j \pwgc i$ then
-  there must either be ``causal flow'' from $j$ to $i$
-  ($j \in \anc{i}$) or there must be a confounder $k$ through which
-  common information is received.
-\end{remark}
-
 An interesting corollary is the following:
@@
 
-It seems reasonable to expect a converse of proposition
+It seems reasonable to expect a converse of Proposition
 \ref{prop:ancestor_properties} to hold, i.e.
@@
 Examples of strongly causal graphs include directed trees (or
-forests), DAGs where each node has at most one parent, and figure
+forests), DAGs where each node has at most one parent, and Figure
 \ref{fig:example_fig3} of this paper.  A complete bipartite graph with
@@
   regulatory network of \textit{Saccharomyces cerevisiae}.  These
-  networks, which we reproduce in figure \ref{fig:gene_networks}
+  networks, which we reproduce in Figure \ref{fig:gene_networks},
   appear to have at most a small number of edges which violate the
@@
 
-  {\footnotesize Figure \ref{fig:gene_network1}
+    {\footnotesize Figure \ref{fig:gene_network1} has only one edge
+      violating the strong-causality assumption, and Figure
+      \ref{fig:gene_network2} appears qualitatively to be nearly
+      strongly causal.  In particular, most of the edges are emanating
+      from network hubs, and many of the other edges are colliders.
+
+      Figure \ref{fig:gene_network1}
       is reproduced under the Creative Commons Attribution
@@
 before moving into the main result of this section.  The following
-important property essentially strengthens proposition
+important property essentially strengthens Proposition
 \ref{prop:ancestor_properties} for the case of strongly causal graphs.
@@
   Suppose that $i \pwgc j$ and $j \in \anc{i}$.  Then since $\gcg$ is
-  acyclic $i \not\in \anc{j}$, and by proposition
+  acyclic $i \not\in \anc{j}$, and by Proposition
   \ref{prop:ancestor_properties} there is some
   $k \in \anc{i}\cap\anc{j}$ which is a confounder.  However, by
-  proposition \ref{prop:sc_graph_common_anc} $k$ cannot be a
+  Proposition \ref{prop:sc_graph_common_anc} $k$ cannot be a
   confounder, a contradiction.
@@
 %   Since $j \in \anc{i}$ there is a path $\gcgpath{j}{i}$.
-%   Moreover, since $i \pwgc j$ by proposition \ref{prop:pwgc_anc} there
+%   Moreover, since $i \pwgc j$ by Proposition \ref{prop:pwgc_anc} there
 %   must be a confounding ancestor $k \in \anc{i} \cap \anc{j}$, where the
@@
 
-%   \hl{Double check that the ancestor properties proposition implies
+%   \hl{Double check that the ancestor properties Proposition implies
 %     that BOTH paths do not contain the other node.}
@@
 \begin{proof}
-  This follows directly from applying proposition
+  This follows directly from applying Corollary
   \ref{cor:parent_corollary} to $i \pwgc j$ and $j \pwgc i$.
@@
 
-In light of proposition \ref{prop:sc_graph_common_anc}, the following
-provides a partial converse to proposition
+In light of Proposition \ref{prop:sc_graph_common_anc}, the following
+provides a partial converse to Proposition
 \ref{prop:ancestor_properties}, and supports the intuition of ``causal
@@
   nodes pose challenges for Granger-causality.  However, as opposed to
-  Pearl's causal calculus \cite{pearl2000art}, pairwise
+  the causal calculus of \cite{pearl2009causality}, pairwise
   Granger-causality does not suffer any difficulty with so-called
-  ``colliders'' (in the language of Pearl's causal calculus
-  \cite{pearl2009causality}), that is, the topology
-  $i \rightarrow k \leftarrow j$ will never result in $i \pwgc j$ or
-  $j \pwgc i$.  This is evidently an advantage of the
-  \textit{temporal} nature of Granger-causality -- there is no
+  ``colliders'', that is, the topology $i \rightarrow k \leftarrow j$ will never result
+  in $i \pwgc j$ or $j \pwgc i$.  This is evidently an advantage of
+  the \textit{temporal} nature of Granger-causality -- there is no
   backwards causal flow along the edges of $\gcg$.
@@
   As a final remark of this subsection we note that a complete
-  converse to proposition \ref{prop:ancestor_properties} is not
+  converse to Proposition \ref{prop:ancestor_properties} is not
   possible without additional conditions.  Consider the ``fork''
@@
 In section \ref{sec:strongly_causal_graphs} we obtained a converse to
-part $(a)$ of proposition \ref{prop:ancestor_properties} via the
+part $(a)$ of Proposition \ref{prop:ancestor_properties} via the
 notion of a strongly causal graph topology.  In this section, we
@@
 
-This interpretation of the following persistence condition is that
+The interpretation of the following persistence condition is that
 each node stores some ``memory'' of the past.
@@
   where $\B(z)$ is Nilpotent.  Indeed, in a $\VAR(1)$ model, if the
-  system matrix is diagonalizable and has atleast two distinct
+  system matrix is diagonalizable and has at least two distinct
   eigenvalues, then the system is persistent.  See Example
@@
   \label{ex:persistent_system}
-  process $x(t)$ generated by the $\VAR(1)$ model\footnote{Recall that
-    any $\VAR(p)$ model with $p < \infty$ can be written as a
-    $\VAR(1)$ model, so we lose little generality in considering this
-    case.}  having $\B(z) = Bz^{-1}$, we will examine
+  Consider a process $x(t)$ generated by the $\VAR(1)$
+  model\footnote{Recall that any $\VAR(p)$ model with $p < \infty$ can be
+    written as a $\VAR(1)$ model, so we lose little generality in
+    considering this case.}  having $\B(z) = Bz^{-1}$, we will examine
   conditions that guarantee persistence.  Pick any
-  $i \in [n], j \in \anc{i} \setminus \{i\}$, then the stability of
+  $i \in [n], j \in \anc{i} \setminus \{i\}$.  Then the stability of
   $B$ allows us to write
@@
     \begin{aligned}
-      e_i^\T B^{rk} e_j &\overset{(a)}{=} \big((P^\T e_i)^\T J^{rk} P^{-1}e_j\big)\\
+      [B^{rk}]_{ij} &= e_i^\T B^{rk} e_j \\
+      &\overset{(a)}{=} \big((P^\T e_i)^\T J^{rk} P^{-1}e_j\big)\\
       &= \tr [(P^\T e_i)^\T J^{rk} P^{-1}e_j]\\
@@
   where $\lambda_\nu$ denotes the eigenvalues of $B$ and
-  $\beta_\nu = u_\nu v_\nu$.  Note that $f(0) = 0$ since $i \ne j$,
+  $\beta_\nu = u_\nu v_\nu$.  Note that $f(0) = 0$ since $i \ne j$ and
   $u$ is a row of $P$ and $v$ is a column of $P^{-1}$.  Moreover,
@@
   for an infinite number of integers $r$ (the exponentiation is to be
-  understood as a pointwise operation).  However, since there can only
+  understood as a point wise operation).  However, since there can only
   be a finite number of (unit length) eigenvectors, this cannot be the
@@
   The importance of this result is that it allows us to eliminate the
-  effects of confounding by 
+  effects of confounding by removing bidirectional edges.
 \end{remark}
@@
 \label{sec:pairwise_algorithm}
-In this section we will show that if the $\gcg$ of a persistent
-process is a strongly causal DAG, then it is possible to recover
-$\gcg$ via pairwise tests alone.  This is the main conclusion of the
-theoretical analysis in this paper.
+We arrive at the main conclusion of the theoretical analysis in this
+paper.
 
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}
   \SetKwInOut{Input}{input}
@@
   \label{alg:pwgr}
-  % \TitleOfAlgo{Pairwise Graph Recovery}
-  % \Input{Pairwise Granger-causality relations between a persistent
-  % process of dimension $n$ whose joint Granger-causality
-  % relations are known to form a strongly causal DAG $\gcg$.}
-  \Input{Pairwise Granger-causality relations}
+  \TitleOfAlgo{Pairwise Graph Recovery}
+  \Input{Pairwise Granger-causality relations between a persistent
+  process of dimension $n$ whose joint Granger-causality
+  relations are known to form a strongly causal DAG $\gcg$.}
+  % \Input{Pairwise Granger-causality relations}
   \Output{Edges $\gcge = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of
@@
   $W \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$  \texttt{\# candidate edges}\\
-  $P_0 \leftarrow \{i \in S_0\ |\ \forall s \in S_0\ (s, i) \not\in W\}$  \texttt{\# parentless nodes}\\
+  $P_0 \leftarrow \{i \in S_0\ |\ \forall s \in S_0\ (s, i) \not\in W\}$  \texttt{\# parent-less nodes}\\
   \While{$S_{k - 1} \ne \emptyset$}{
@@
 The theorem is proven in the supplementary material by establishing
-the correctness correctness of algorithm (\ref{alg:pwgr}).  The idea
-is to iteratively ``peel away layers'' of nodes by removing the nodes
-that have no parents remaining, which always exist since the graph is
-acyclic.  The requirement of strong causality ensures that all actual
-edges of $\gcg$ manifest in some way as pairwise relations (by
-proposition \ref{prop:pwgc_anc}), and the requirement of persistence
-allows confounding to be eliminated by removing bidirectional edges.
-Without persistence, each confounded pair would give rise to $4$
-possible pairwise topologies consistent with $\gcg$, one for each type
-of pairwise edge (no edge, unidirectional, bidirectional).
+the correctness of Algorithm (\ref{alg:pwgr}).  The idea is to
+iteratively ``peel away layers'' of nodes by removing the nodes that
+have no parents remaining.  The requirement of strong causality
+ensures that all actual edges of $\gcg$ manifest in some way as
+pairwise relations (by Proposition \ref{prop:pwgc_anc}), and the
+requirement of persistence allows confounding to be eliminated by
+removing bidirectional edges(by Proposition
+\ref{prop:persistence_converse} and Corollary
+\ref{cor:bidirectional_edge}).  Without persistence, each confounded
+pair would give rise to $4$ possible pairwise topologies consistent
+with $\gcg$, one for each type of pairwise edge (no edge,
+unidirectional, bidirectional).
 
@@
   The set $W$ collects ancestor relations in $\gcg$ (see Lemma
-  \ref{lem:W_subset_E}).  In reference to figure
+  \ref{lem:W_subset_E}).  In reference to Figure
   \ref{fig:example_fig3}, each of the solid black edges, as well as
@@
   bidirectional green dash-dotted edges, which we are able to exclude
-  by appealing to Corollary \ref{cor:bidirectional_edge}.  The
-  groupings $P_0, \ldots, P_3$ are also indicated in figure
-  \ref{fig:example_fig3}.
-
-  The algorithm proceeds first with the parentless nodes $1, 2$ on the
+  as results of confounding.  The groupings $P_0, \ldots, P_3$ are also
+  indicated in Figure \ref{fig:example_fig3}.
+
+  The algorithm proceeds first with the parent-less nodes $1, 2$ on the
   initial iteration where the edge $(1, 3)$ is added to $E$.  On the
@@
   That we need to proceed backwards through $P_{k - r}$ as in the
-  inner loop of on $r$ can also be seen from this example, where if
+  inner loop on $r$ can also be seen from this example, where if
   instead we simply added the set
@@
 \label{sec:simulation}
-We implement an heuristic inspired by Algorithm \ref{alg:pwgr}, the
+We implement an heuristic inspired by Algorithm \ref{alg:pwgr} by
+replacing the population statistics with finite sample tests, the
 details of which can be found in the supplementary material Section
-\ref{sec:finite_pwgc} (see Algorithm \ref{alg:finite_pwgc}.  The
+\ref{sec:finite_pwgc} (see Algorithm \ref{alg:finite_pwgc}).  The
 heuristic is essentially controlling the false discovery rate
@@
     \multirow{4}{*}{\textbf{50}}
-    &\textbf{SCG}&1.83&\textbf{1.57}&0.48&\textbf{0.07}&0.47&\textbf{0.57}\\
-    &\textbf{0.06}&1.91&\textbf{1.69}&0.54&\textbf{0.09}&0.43&\textbf{0.53}\\
-    &\textbf{0.11}&2.89&\textbf{2.58}&0.46&\textbf{0.22}&0.38&0.40\\
-    &\textbf{0.46}&8.49&\textbf{7.72}&0.42&0.42&\textbf{0.16}&0.11\\
+    &\textbf{SCG}&1.71&\textbf{1.55}&0.52&\textbf{0.08}&0.46&\textbf{0.55}\\
+    &\textbf{0.04}&1.97&\textbf{1.77}&0.57&\textbf{0.10}&0.41&\textbf{0.53}\\
+    &\textbf{0.08}&2.95&\textbf{2.72}&0.50&\textbf{0.23}&0.36&\textbf{0.39}\\
+    &\textbf{0.32}&9.02&\textbf{8.17}&\textbf{0.53}&0.56&\textbf{0.14}&0.10\\
     \midrule
     \multirow{4}{*}{\textbf{250}}
-    &\textbf{SCG}&1.32&\textbf{1.22}&0.27&\textbf{0.07}&0.71&\textbf{0.80}\\
-    &\textbf{0.06}&1.44&\textbf{1.28}&0.32&\textbf{0.09}&0.66&\textbf{0.76}\\
-    &\textbf{0.11}&2.57&\textbf{2.17}&0.32&\textbf{0.16}&0.53&\textbf{0.57}\\
-    &\textbf{0.46}&8.09&\textbf{7.11}&0.39&0.40&\textbf{0.20}&0.14\\
+    &\textbf{SCG}&1.30&\textbf{1.18}&0.29&\textbf{0.06}&0.70&\textbf{0.81}\\
+    &\textbf{0.04}&1.40&\textbf{1.31}&0.30&\textbf{0.07}&0.68&\textbf{0.76}\\
+    &\textbf{0.08}&2.49&\textbf{2.21}&0.32&\textbf{0.16}&0.55&\textbf{0.57}\\
+    &\textbf{0.32}&8.67&\textbf{7.62}&0.48&0.46&\textbf{0.18}&0.15\\
     \midrule
     \multirow{4}{*}{\textbf{1250}}
-    &\textbf{SCG}&1.18&\textbf{1.09}&0.40&\textbf{0.06}&0.69&\textbf{0.89}\\
-    &\textbf{0.06}&1.24&1.17&0.43&\textbf{0.05}&0.65&\textbf{0.86}\\
-    &\textbf{0.11}&2.02&2.10&0.33&\textbf{0.14}&0.62&0.63\\
-    &\textbf{0.46}&7.44&\textbf{7.04}&0.40&\textbf{0.30}&\textbf{0.22}&0.19\\
+    &\textbf{SCG}&1.20&\textbf{1.11}&0.41&\textbf{0.07}&0.68&\textbf{0.88}\\
+    &\textbf{0.04}&1.28&\textbf{1.20}&0.46&\textbf{0.07}&0.64&\textbf{0.84}\\
+    &\textbf{0.08}&2.12&2.05&0.36&\textbf{0.14}&0.60&\textbf{0.64}\\
+    &\textbf{0.32}&7.78&\textbf{7.39}&0.49&\textbf{0.37}&\textbf{0.21}&0.18\\
     \bottomrule
@@
   {\footnotesize Results of Monte Carlo simulations comparing PWGC and
-    AdaLASSO $(n = 35, p = 5, p_{\text{max}} = 10)$ for small
+    AdaLASSO $(n = 50, p = 5, p_{\text{max}} = 10)$ for small
     samples and when the SCG assumption doesn't hold.  The superior
@@
 
-Our simulation results are summarized in \ref{tab:simulation_table},
-with additional figures provided in the supplementary material Section
-\ref{apx:simulation}.  It is clear that the superior performance of
-PWGC in comparison to AdaLASSO is as a result of limiting the false
-discovery rate.  It is unsurprising that PWGC exhibits superior
-performance when the graph is an SCG, but even in the case of more
-general DAGs, the PWGC heuristic is still able to more reliably
-uncover the graph structure for small values of $q$.  As $q$
-increases, there are simply not enough edges allowed by the SGC
-topology for it to be possible to accurately recover $\gcg$.
-
-Interestingly, we can observe that the AdaLASSO appears to perorm
+Our simulation results are summarized in Table
+\ref{tab:simulation_table}, with additional figures provided in the
+supplementary material Section \ref{apx:simulation}.  It is clear that
+the superior performance of PWGC in comparison to AdaLASSO is as a
+result of limiting the false discovery rate.  It is unsurprising that
+PWGC exhibits superior performance when the graph is an SCG, but even
+in the case of more general DAGs, the PWGC heuristic is still able to
+more reliably uncover the graph structure for small values of $q$.  We
+would conjecture that for small $q$, random graphs are ``likely'' to
+be ``close'' to SCGs in some appropriate sense.  As $q$ increases,
+there are simply not enough edges allowed by the SGC topology for it
+to be possible to accurately recover $\gcg$.
+
+Interestingly, we can observe that the AdaLASSO appears to perform
 marginally better on strongly causal graphs than directed acyclic
-graphs with an equivalent number of edges ($q \approx 2 / n$ is chosen for
+graphs with an equivalent number of edges ($q \approx 0.04$ is chosen for
 this purpose).  This provides supporting evidence for one of the main
@@
 causality graphs.  For example, are there provable guarantees on the
-error rate of PWGC when applied to non strongly-causal graphs?
-Computationally, can constraint systems or cunning adaptive weighting
-schemes impose useful prior knowledge about graph topology for the
-LASSO algorithm?  In another direction, it is known that the more
-general notion of transfer entropy reduces to Granger-causality in the
-case of Gaussian data \cite{barnett2009granger}, do our results extend
-to causality networks defined via transfer entropy?  Finally, the work
-of \cite{barnett2015granger} has established the superiority of
-Granger-causality testing in state space models (as opposed to pure
-autoregressions) in many cases.  Combining this work with our PWGC
-algorithm (by modifying the approach described in Section
-\ref{sec:pairwise_hypothesis_testing} to instead utilize state-space
-Granger-causality testing) therefore is likely to enable application
-to very large networks of time series data which are not well
-approximated by finite $\VAR(p)$ models.
+error rate of PWGC when applied to non strongly-causal graphs?  Can
+constraint systems or cunning adaptive weighting schemes impose useful
+prior knowledge about graph topology for the LASSO algorithm?
+Finally, the work of \cite{barnett2015granger} has established the
+superiority of Granger-causality testing by state space models (as
+opposed to pure autoregressions) in many cases.  Combining this work
+with our PWGC algorithm (by modifying the approach described in
+Section \ref{sec:pairwise_hypothesis_testing} to instead utilize
+state-space Granger-causality testing) therefore is likely to enable
+application to very large networks of time series data which are not
+well approximated by finite $\VAR(p)$ models.
 

### 2019-07-18 paper_theory/proofs.tex
Better proofs.

@@
 \AtAppendix{\counterwithin{definition}{section}}
+\AtAppendix{\counterwithin{equation}{section}}
 \externaldocument{paper_theory}
@@
 
+References to equations, lemmas, etc., in this document are prefixed
+with their section, whereas prefix-free equation numbers refer to the
+main document.  e.g. ``Equation (1)'' refers to the first equation in
+the main document, and ``Equation (A.1)'' refers to the first equation
+in this document.
+
 Code will be made available at
@@
 \section{Proofs}
-\subsection{Ancillary Results}
+\subsection{Preparatory Results}
 \label{apx:ancillary_results}
@@
   $(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$ since
-  $\B_{ij}(z) = 0$ and $v_i(t) \perp \H_{t - 1}$.  $(c) \iff (d)$ is a result
-  of the equivalence in Definition
-  \ref{lem:conditional_orthogonality_equivalence}.  And,
+  $\B_{ij}(z) = 0$.  The result follows since
+  $v_i(t) \perp \H_{t - 1}$.  $(c) \iff (d)$ is a result of the equivalence
+  in Definition \ref{lem:conditional_orthogonality_equivalence}.  And,
   $(d) \implies (a)$ follows directly from the Definition.
@@
 
-This statement is ultimately about the sparsity pattern in the Wold
-decomposition matrices $A(\tau)$ since
-$x_i(t) = \sum_{\tau = 0}^\infty \sum_{j = 1}^n A_{ij}(\tau)v_j(t -
-\tau)$.  The proposition states that if $j \not \in \anc{i}$ then
-$\A_{ij}(z) = 0$.
-
 \begin{proof}
   Equation \eqref{eqn:parent_expansion} is immediate from the
-  $\VAR(\infty)$ representation of \eqref{eqn:ar_representation} and
-  Theorem \ref{thm:granger_causality_equivalences}, we are left to
+  $\VAR(\infty)$ representation of Equation \eqref{eqn:ar_representation}
+  and Theorem \ref{thm:granger_causality_equivalences}, we are left to
   demonstrate \eqref{eqn:ancestor_expansion}.
@@
 
-  where $(I - \B(z))^{-1} = \A(z)$ due to the uniqueness of
-  (\ref{eqn:wold}).  Since $\B(z)$ is stable we have
+  where necessarily $(I - \B(z))^{-1} = \A(z)$ due to the uniqueness
+  of the Wold decomposition.  Since $\B(z)$ is stable we have
 
@@
   \begin{align*}
-    x_i(t) &= [(I - \B(z))^{-1}v(t)]_i = \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
+    x_i(t) &= [(I - \B(z))^{-1}v(t)]_i\\
+           &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
            &= \A_{ii}(z) v_i(t) + \sum_{\substack{j \in \anc{i} \\ j \ne i}} \A_{ij}(z) v_j(t)
@@
   We show directly that
-  $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.  To this end,
-  fix $s, \tau \ge 0$, then by expanding with equation \eqref{eqn:ancestor_expansion} we have
+  $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.  To this end, fix
+  $s, \tau \ge 0$, then by expanding with Equation
+  \eqref{eqn:ancestor_expansion} we have
 
@@
     &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\E[\big(\A_{ik}(z)v_k(t - s)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)]\\
-    &+ \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[\big(\A_{ii}(z)v_i(t - s)\big) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
+    &+ \sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ii}(z)v_i(t - s)\big) \big(\A_{j\ell}(z) v_\ell(t - \tau)\big)]\\
     &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ik}(z)v_k(t - s)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)].
@@
 
+\begin{lemma}
+  \label{lem:vj_perp}
+  Consider distinct nodes $i, j$ in a Granger-causality graph $\gcg$.
+  If $j \not \in \anc{i}$, then $\H_t^{v_j} \perp \H_{t}^{i}$, and therefore
+  for any causal filter $\Phi(z)$ we have
+
+  \begin{align*}
+    \linE{\Phi(z)v_j(t)}{\H_{t - 1}^i} &= 0\\
+    \inner{x_i(t)}{\Phi(z)v_j(t)} & = 0.
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  Fix $\tau, s \ge 0$, then by expanding with Equation \ref{eqn:ancestor_expansion}
+
+  \begin{align*}
+    \E[x_i(t - \tau)v_j(t - s)] &= \E[\big(\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t) \big) v_j(t - s)]\\
+    &= 0.
+  \end{align*}
+
+  This follows since $i \ne j$ and $j \not \in \anc{i}$ and $v(t)$ is
+  isotrophic and uncorrelated.
+\end{proof}
+
 \begin{proposition}
@@
   \begin{equation*}
-    \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^i}} = 0.
+    \forall \psi \in \H_{t - 1}^j\ \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}{\psi - \linE{\psi}{\H_{t - 1}^i}} = 0.
   \end{equation*}
 
-  which by the orthogonality principle is equivalent to
+  which by the orthogonality principle and by representing
+  $\psi \in \H_{t - 1}^j$ via the action of some strictly causal filter
+  $\Phi(z)$ on $x_j(t)$ is equivalent to
 
@@
     \label{eqn:proof_inner0}
-    \inner{x_i(t)}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^i}} = 0.
-  \end{equation}
-
-  Define the disjoint sets
-
-  \begin{align*}
-    C_0(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, k \ne i \}\\
-    C_1(u) &= \{k \in \pa{u}\ |\ i \in \anc{k} \text{ or } k = i \}.
-  \end{align*}
-
-  We can then expand the parents of $j$ using Equation \ref{eqn:parent_expansion} as
+    \inner{x_i(t)}{\Phi(z)x_j(t) - \linE{\Phi(z)x_j(t)}{\H_{t - 1}^i}} = 0.
+  \end{equation}
+
+  If we expand $x_j(t)$ using Equation \eqref{eqn:ancestor_expansion},
+  the left hand side of \eqref{eqn:proof_inner0} becomes
 
   \begin{equation*}
-    x_j(t) = v_j(t) + \sum_{k \in C_0(j)} B_{jk}(z) x_k(t) + \sum_{k \in C_1(j)} B_{jk}(z) x_k(t),
+    \inner{x_i(t)}{\sum_{k \in \anc{j} \cup \{j\}} \Big(\Phi(z)\A_{jk}(z)v_k(t) - \linE{\Phi(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^i}\Big)}.
   \end{equation*}
 
-  which when substituted into the left hand side of Equation
-  \ref{eqn:proof_inner0} results (by Proposition
-  \ref{prop:separated_ancestor_uncorrelated}) in
-
-  \begin{equation*}
-    \inner{x_i(t)}{\sum_{k \in C_1(j)} B_{jk}(z) x_k(t) - \sum_{k \in C_1(j)} \linE{B_{jk}(z) x_k(t)}{\H_{t - 1}^i}}.
-  \end{equation*}
-
-  We can continue this process recursively (i.e. split each
-  $k \in C_1(j)$ into $C_0(k)$ and $C_1(k)$) which must eventually
-  terminate with $C_1(u) = \{i\}$.  Therefore, there exists some
-  causal filter $\Phi(z)$ such that Equation \ref{eqn:proof_inner0} is
-  equivalent to
-
-  \begin{equation}
-    \label{eqn:proof_inner0}
-    \inner{x_i(t)}{\Phi(z)x_i(t) - \linE{\Phi(z)x_i(t)}{\H_{t - 1}^i}},
-  \end{equation}
-
-  which is $0$ since $\linE{\Phi(z)x_i(t)}{\H_{t - 1}^i} = \Phi(z)x_i(t)$.
-\end{proof}
+  We see that this is $0$ by Lemma \ref{lem:vj_perp} since
+  $j \not \in \anc{i}$, and
+  \[
+    \anc{i} \cap \anc{j} = \emptyset \implies \forall k \in \anc{j}: k \not \in \anc{i}.
+  \]
+\end{proof}
+
+% I believe this is an incorrect proof since it relies on inverting
+% (1 - B_{jj}(z)), which need not be invertible.  For example,
+% B(z) = [[r, -a], [a, 0]]z^{-1} is invertible whenever
+% max(|r - root(r^2 - 4a^2)|, |r + root(r^2 - 4a^2)|) < 2
+% which includes values of r > 1.
+
+% \begin{proof}
+%   By Theorem \ref{thm:granger_causality_equivalences} it suffices to show that
+
+%   \begin{equation*}
+%     \forall \psi \in \H_{t - 1}^j\ \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}{\psi - \linE{\psi}{\H_{t - 1}^i}} = 0.
+%   \end{equation*}
+
+%   which by the orthogonality principle and by representing
+%   $\psi \in \H_{t - 1}^j$ via the action of some strictly causal filter
+%   $\Phi(z)$ on $x_j(t)$ is equivalent to
+
+%   \begin{equation}
+%     \label{eqn:proof_inner0}
+%     \inner{x_i(t)}{\Phi(z)x_j(t) - \linE{\Phi(z)x_j(t)}{\H_{t - 1}^i}} = 0.
+%   \end{equation}
+
+%   Given a node $u$, consider the partition of $\pa{u}$:
+
+%   \begin{align*}
+%     C_0(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, k \ne i \}\\
+%     C_1(u) &= \{k \in \pa{u}\ |\ i \in \anc{k} \text{ or } k = i \}.
+%   \end{align*}
+
+%   We can then expand the parents of $j$ using Equation \ref{eqn:parent_expansion} as
+
+%   \begin{align*}
+%     x_j(t) &= v_j(t) + \B_{jj}(z)x_j(t) + \sum_{k \in C_0(j)} B_{jk}(z) x_k(t) + \sum_{k \in C_1(j)} B_{jk}(z) x_k(t)\\
+%     &= \widetilde{A}_{jj}(z)\Big(v_j(t) + \sum_{k \in C_0(j)} B_{jk}(z) x_k(t) + \sum_{k \in C_1(j)} B_{jk}(z) x_k(t)\Big),
+%   \end{align*}
+
+%   % \todo{Is this true??}
+%   where $(1 - B_{jj}(z))^{-1} = \widetilde{A}_{jj}(z)$, which exists
+%   due to the invertibility of $\A(z)$. When this expression is
+%   substituted into the left hand side of Equation
+%   \ref{eqn:proof_inner0} it results in
+
+%   \begin{equation}
+%     \label{eqn:proof_inner1}
+%     \begin{aligned}
+%       \inner{x_i(t)}{\sum_{k \in C_1(j)} \big(\widetilde{A}_{jj}(z)\Phi(z)B_{jk}(z) x_k(t) &- \linE{\widetilde{A}_{jj}(z)\Phi(z)B_{jk}(z) x_k(t)}{\H_{t - 1}^i}\big)},
+%     \end{aligned}
+%   \end{equation}
+
+%   where (by Proposition \ref{prop:separated_ancestor_uncorrelated}) we
+%   have eliminated each of the terms from $C_0(u)$, and by Lemma
+%   \ref{lem:vj_perp} the terms involving $v_j(t)$.
+
+%   Now, for any $u \in \anc{j}$, we must have $u \not \in \anc{i}$, since
+%   otherwise $u \in \anc{i} \cap \anc{j}$, which is empty by hypothesis.
+%   Moreover, for any $u \in \anc{j}$, we have
+%   $\anc{i} \cap \anc{u} = \emptyset$, since
+%   $w \in \anc{i} \cap \anc{u} \implies w \in \anc{i} \cap \anc{j}$. Therefore,
+%   we can continue the above process recursively splitting each
+%   $k \in C_1(j)$ into $C_0(k)$ and $C_1(k)$) upwards along each path
+%   from $j$ to $i$, terminating whenever we encounter $i$.  This process
+%   is guaranteed to terminate since $i \in \anc{k}$.  This leaves us with
+%   some strictly causal filter $\Psi(z)$ such that Expression
+%   \ref{eqn:proof_inner1} is equivalent to
+
+%   \begin{equation}
+%     \inner{x_i(t)}{\Psi(z)x_i(t) - \linE{\Psi(z)x_i(t)}{\H_{t - 1}^i}},
+%   \end{equation}
+
+%   which is $0$ since $\Psi(z)x_i(t) \in \H_{t - 1}^i$ and
+%   therefore $\linE{\Psi(z)x_i(t)}{\H_{t - 1}^i} = \Psi(z)x_i(t)$.
+% \end{proof}
 
@@
   necessarily inherits these same two properties.  This follows since
-  if we also had $u \in \anc{i}$ then $u \in \anc{i} \cap \anc{j}$ so that every
-  $u \rightarrow \cdots \rightarrow j$ path must contain $i$, but
-  $u \in \pa{j}$, so this is not the case since $u \rightarrow j$ is a
-  path that doesn't contain $i$; moreover, if we consider
+  if we also had $u \in \anc{i}$ then $u \in \anc{i} \cap \anc{j}$ so by our
+  assumption every $u \rightarrow \cdots \rightarrow j$ path must contain
+  $i$, but $u \in \pa{j}$ so $u \rightarrow j$ is a path that doesn't contain
+  $i$, and therefore $u \not\in \anc{i}$; moreover, if we consider
   $w \in \anc{i} \cap \anc{u}$ then we also have
-  $w \in \anc{i} \cap \anc{j}$ so every
-  $w \rightarrow \cdots \rightarrow j$ path must contain $i$.  These
-  properties therefore extend inductively to every
-  $u \in \big(\anc{j} \setminus \{i\}\big)$.
+  $w \in \anc{i} \cap \anc{j}$ so the assumption implies that every
+  $w \rightarrow \cdots \rightarrow j$ path must contain $i$.  These properties therefore
+  extend inductively to every $u \in \big(\anc{j} \setminus \{i\}\big)$.
 
@@
 
-  We notice that for any $u$ having the properties $(a), (b)$ above, we
-  must have $C_2(u) = \emptyset$ since if $k \in C_2(u)$ then
-  $\exists w \in \anc{i} \cap \anc{k}$ s.t. $i \not \in \anc{k}$ and
-  therefore there must be a path $\gcgpath{w}{k} \rightarrow u$ which
-  does not contain $i$.
-
-  Using this partition, we will expand $x_j(t)$ in terms of it's
-  parents, and recursively expand nodes in $C_1$ until we reach a case
-  where $C_1 = \emptyset$.  For the first step equation
+  We notice that for any $u$ having the properties $(a), (b)$ above,
+  we must have $C_2(u) = \emptyset$ since if $k \in C_2(u)$ then
+  $\exists w \in \anc{i} \cap \anc{k}$ (and
+  $w \in \anc{i} \cap \anc{u}$, since $k \in \pa{u}$) such that
+  $i \not \in \anc{k}$ and therefore there must be a path
+  $\gcgpath{w}{k} \rightarrow u$ which does not contain $i$, contradicting
+  property $(b)$.
+
+  Using this partition, and in a similar fashion to the proof of
+  Proposition \ref{prop:ancestor_uncorrelated}, we will expand
+  $x_j(t)$ in terms of it's parents, and recursively expand nodes in
+  $C_1$ until reaching $i$.  For the first step equation
   \eqref{eqn:parent_expansion} gives us:
@@
     \label{eqn:xj_partition_expansion}
-    x_j(t) = \A_{jj}(z)\Big(v_j(t) + \sum_{k \in C_0(j)}\B_{ik}(z)x_k(t) + \sum_{k \in C_1(j)}\B_{ik}(z)x_k(t)\Big).
-  \end{equation}
-
-  Using this representation we choose an arbitrary $\Phi(z) x_i(t - 1) \in \H_{t - 1}^{(j)}$ and show that
-
-  \begin{equation}
-    \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}{\Phi(z)x_j(t - 1) - \linE{\Phi(z)x_j(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
-  \end{equation}
-
-  which will imply (by Theorem \ref{thm:granger_causality_equivalences}) that $j \npwgc i$ and for which it is equivalent to show that
+    x_j(t) = \widetilde{\A}_{jj}(z) \Big(v_j(t) + \sum_{k \in C_0(j)} \B_{ik}(z)x_k(t) + \sum_{k \in C_1(j)}\B_{ik}(z)x_k(t)\Big).
+  \end{equation}
+
+  Using this representation we choose an arbitrary element of
+  $\H_{t - 1}^j$ and represent it as the action of a strictly causal
+  filter $\Phi(z)$, i.e.  $\Phi(z) x_j(t) \in \H_{t - 1}^{(j)}$ and show that
 
@@
     \label{eqn:sufficient_inner_prod}
-    \inner{x_i(t)}{\Phi(z)x_j(t - 1) - \linE{\Phi(z)x_j(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
-  \end{equation}
-
-  by the orthogonality principle since $\linE{x_i(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$.
-Substituting (\ref{eqn:xj_partition_expansion}) into (\ref{eqn:sufficient_inner_prod}) and starting with the first term we have
-
-  \begin{align*}
-    &\inner{x_i(t)}{\Phi(z)\A_{jj}(z)v_j(t - 1) - \linE{\Phi(z)\A_{jj}(z)v_j(t - 1)}{\H_{t - 1}^{(i)}}}\\
-    \overset{(\alpha)}{=}\ &\inner{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}{\Phi(z)\A_{jj}(z)v_j(t - 1)}\\
-    \overset{(\beta)}{=}\ &0,
-  \end{align*}
-
-  where $(\alpha)$ follows by expanding $x_i(t)$ with (\ref{eqn:ancestor_expansion}) and $\linE{\Phi(z)\A_{jj}(z)v_j(t - 1)}{\H_{t - 1}^{(i)}} = 0$ because $\forall \tau, s$
-
-  \begin{equation*}
-    \E v_j(t - \tau) x_i(t - s) = \E v_j(t - \tau) \sum_{k \in \anc{i} \cup \{i\}}\A_{ik}(z)v_k(t - s) = 0,
-  \end{equation*}
-
-  since $j \not \in \anc{i}$; $(\beta)$ follows similarly, that is, $j \not \in \anc{i}$.  Secondly we see that $\forall k \in C_0(j)$
-
-  \begin{equation*}
-    \inner{x_i(t)}{\Phi(z)\A_{jj}(z)\B_{ik}(z)x_k(t - 1) - \linE{\Phi(z)\A_{jj}(z)\B_{ik}(z)x_k(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
-  \end{equation*}
-
-  which follows from Proposition \ref{prop:ancestor_uncorrelated}.  Finally,
-  for $k \in C_1(j)$ the case $k = i$ is immediate (since the error in
-  estimating $\Phi(z)\B_{ii}(z)x_i(t)$ given $\H_{t - 1}^{(i)}$ is
-  $0$), so suppose $k \ne i$.  We know from above that $k$ inherits the
-  key properties referred to as $(a)$ and $(b)$ above and therefore we
-  can recursively expand $k$ in the same way as in equation
-  (\ref{eqn:xj_partition_expansion}).  Continuing this recursion for
-  each $k \in C_1(j)$ (where $k \ne i$) must eventually terminate since
-  $i \in \anc{k}$.
+    \inner{x_i(t)}{\widetilde{\A}_{jj}(z)\Phi(z)x_j(t) - \linE{\widetilde{\A}_{jj}(z)\Phi(z)x_j(t)}{\H_{t - 1}^{(i)}}} = 0,
+  \end{equation}
+
+  which will imply (by Theorem
+  \ref{thm:granger_causality_equivalences}) that $j \npwgc i$.
+
+  Substituting (\ref{eqn:xj_partition_expansion}) into
+  (\ref{eqn:sufficient_inner_prod}) we have cancellation of all terms
+  in which $k \in C_0(j)$ by Proposition \ref{prop:ancestor_properties},
+  and cancellation of the term involving $v_j(t)$ by Lemma \ref{lem:vj_perp}.
+
+  We are therefore left with
+
+  \begin{equation}
+    \label{eqn:refwornv}  % An arbitrary code for reference in just one place
+    \inner{x_i(t)}{\sum_{k \in C_1(j)} \big(\widetilde{\A}_{jj}(z)\Phi(z) \B_{ik}(z)x_k(t) - \linE{\widetilde{\A}_{jj}(z)\Phi(z) \B_{ik}(z)x_k(t)}{\H_{t - 1}^i}\big)}.
+  \end{equation}
+
+  Since each such $k$ inherits the needed properties $(a), (b)$ we can
+  again (as in Proposition \ref{prop:ancestor_properties}) recursively
+  expand this expression until we reach $i$, obtaining a strictly
+  causal $\Psi(z)$ such that Expression \eqref{eqn:refwornv} is equal to
+
+  \begin{equation}
+    \inner{x_i(t)}{\Psi(z)x_i(t) - \linE{\Psi(z)x_i(t)}{\H_{t - 1}^i}} = 0.
+  \end{equation}
 \end{proof}
@@
   $j$.  In this case, there are multiple paths from $k$ to $i$ (one of
-  which \textit{does} go through $j$ since $j \in \anc{i}$) which
+  which \textit{does} go through $j$, since $j \in \anc{i}$) which
   contradicts the assumption of strong causality.
@@
 
-  % Should this be written as a lemma?
   \begin{align*}
@@
 
-  or by rearranging and applying the representation obtained earlier, if and only if
+  or by rearranging and applying the representation for $x_i(t)$
+  obtained earlier, if and only if
 
   \begin{equation*}
-    x_i(t) \overset{\text{a.s.}}{=} \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}},
+    x_i(t) \overset{\text{a.s.}}{=} \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}.
   \end{equation*}
 
-  but this is impossible since $x_i(t) \not \in \H_{t - 1}^{(i)}$.
+  But, this is impossible since $x_i(t) \not \in \H_{t - 1}^{(i)}$.
 \end{proof}
@@
   \label{lem:time_lag_cancellation}
-  Suppose $v(t)$ is a scalar sequence with unit variance and zero
+  Suppose $v(t)$ is a scalar process with unit variance and zero
   autocorrelation and let $\A(z), \B(z)$ be nonzero and strictly
@@
 
-  due to the uncorrelatedness assumptions on $v(t)$.  This expression
-  is $0$ if and only if $\tau_0(\A) \ge 1 + \tau_\infty(\B)$ or if
-  $\tau_0(\B) \ge 1 + \tau_\infty(\A)$ or if the coefficients are
-  orthogonal along the common support.
-
-  Specializing this fact to $\inner{F(z)\A(z)v(t)}{\B(z)v(t)} = 0$ we
+  since $\E[v(t - s)v(t - \tau)] = \delta_{s - \tau}$.  This expression is
+  $0$ if and only if $\tau_0(\A) \ge 1 + \tau_\infty(\B)$ or if
+  $\tau_0(\B) \ge 1 + \tau_\infty(\A)$ or if the coefficients are orthogonal along
+  the common support.
+
+  Specializing this fact to $\inner{F(z)\A(z)v(t)}{\B(z)v(t)}$ we
   see that the coefficients cannot be orthogonal for every choice of
@@
   \begin{align*}
-    \tau_0(F\A) \ge 1 + \tau_\infty(\B) \forall F &\overset{(a)}{\iff} \tau_0(\A) \ge 1 + \tau_\infty(\B) - \underset{F}{\text{min }} \tau_0(F)\\
+     \forall F\ \tau_0(F\A) \ge 1 + \tau_\infty(\B) &\overset{(a)}{\iff} \tau_0(\A) \ge 1 + \tau_\infty(\B) - \underset{F}{\text{min }} \tau_0(F)\\
     &\overset{(b)}{\iff} \tau_0(\A) \ge \tau_\infty(\B),
@@
   We will show that $j \pwgc i$, the other being symmetric.  First
-  note also that by proposition \ref{prop:sc_graph_common_anc} we
-  cannot have $i \in \anc{j}$ or $j \in \anc{i}$ and therefore every
-  $k \in \anc{i}\cap\anc{j}$ will be a confounder.
+  note that by the contrapositive of proposition
+  \ref{prop:sc_graph_common_anc} we cannot have $i \in \anc{j}$ or
+  $j \in \anc{i}$ (else $k$ would not be a confounder) and therefore
+  \textit{every} $u \in \anc{i}\cap\anc{j}$ will be a confounder.
 

### 2019-07-18 paper_theory/proofs.tex
corrected proof of Prop 4.

@@
   $H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$ we have
-  $(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$ since
+  $(x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{-j}}) = v_i(t)$ since
   $\B_{ij}(z) = 0$.  The result follows since
@@
 
+\begin{remark}
+  In order to prove Proposition \ref{prop:ancestor_properties} we
+  require some additional notation, as well as another representation
+  theorem.  The difficulty addressed by the following Definition
+  \ref{def:scc} and Lemma \ref{lem:scc_expansion} is that in the
+  representation of $x_j(t)$ in terms of it's parents (i.e. Equation
+  \eqref{eqn:parent_expansion})
+
+  \[
+    x_i(t) = v_i(t) + \B_{ii}(z)x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t),
+  \]
+
+  the filter $\B_{ii}(z)$ need not be stable.  That is, the inverse
+  filter $(1 - \B_{ii}(z))^{-1}$ need not exist.  An example of this
+  issue is furnished by
+
+  \[
+    \B(z) = \left[
+    \begin{array}{cc}
+      \rho & -a\\
+      a & 0\\
+    \end{array}
+  \right] z^{-1},
+  \]
+
+  for which, depending on the value of $a$, may still be stable even
+  if $|\rho| > 1$.  This implies that it is not always possible to
+  represent $x_i(t)$ in terms of $v_i(t)$ and $x_k(t), k \in \pa{i}$
+  alone, i.e. as
+
+  \[
+    x_i(t) = (1 - \B_{ii}(z))^{-1}\big(v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\big).
+  \]
+
+  The difficulty presented by the non-existence of such a
+  representation may become apparent upon studying the proof of
+  Proposition \ref{prop:ancestor_properties}.
+\end{remark}
+
+\begin{definition}[Strongly Connected Components]
+  \label{def:scc}
+  In a graph $\gcg$, the subset $S \subseteq [n]$ is \textit{strongly
+    connected} if $\forall i, j \in S$, $i \in \anc{j}$ and
+  $j \in \anc{i}$.  We will denote by $S(j)$ (which may be the
+  singleton $\{j\}$) the largest strongly connected component (SCC)
+  containing $j$.  We will denote $x_{S(j)}(t)$ to be the vector of
+  processes
+
+  \[
+    x_{S(j)}(t) = \big(x_s(t)\ |\ s \in S(j)\big),
+  \]
+
+  whose indices are sorted in ascending order.  Similarly, the
+  sub-filter of $\B(z)$ acting on $x_{S(j)}(t)$ will be denoted
+  $\B_{S(j)}(z)$.
+\end{definition}
+
+\begin{lemma}[Expansion in SCCs]
+  \label{lem:scc_expansion}
+  Given some $j \in [n]$, the process $x_{S(j)}(t)$ can be represented
+  by
+% {\substack{k \in \pa{a_0} \\ k \ne a_1}}
+  \begin{equation}
+    \label{eqn:scc_parent_expansion}
+    x_{S(j)}(t) = v_{S(j)}(t) + \B_{S(j)}(z)x_{S(j)}(t) + \sum_{\substack{s \in S(j) \\ k \in \pa{s} \cap S(j)^\c}} B_{sk}(z)x_k(t)e_s^{S(j)},
+  \end{equation}
+
+  where $e_s^{S(j)}$ denotes the length $|S(j)|$ canonical basis
+  vector with a $1$ in the component corresponding to $x_s$ in the
+  vector $x_{S(j)}$, and the summation is a double sum on $s$ and $k$.
+
+  Moreover, the filter $\B(z)$ is stable with $I - \B_{S(j)}(z)$
+  invertible:
+
+  \begin{equation}
+    \label{eqn:scc_inversion}
+    (I - \B_{S(j)}(z))^{-1} = \sum_{k = 0}^\infty \B_{S(j)}(z)^k,
+  \end{equation}
+
+  therefore
+
+  \begin{equation}
+    \label{eqn:scc_parent_expansion_inverted}
+    x_{S(j)}(t) = (I - \B_{S(j)}(z))^{-1} \big(v_{S(j)}(t) + \sum_{\substack{s \in S(j) \\ k \in \pa{s} \cap S(j)^\c}} B_{sk}(z)x_k(t)e_s^{S(j)}\big),
+  \end{equation}
+
+\end{lemma}
+\begin{proof}
+  The representation of Equation \eqref{eqn:scc_parent_expansion}
+  follows directly from the $\VAR$ representation of $x(t)$ (i.e. Equation
+  \eqref{eqn:ar_representation})
+
+  \[
+    x(t) = \B(z)x(t) + v(t).
+  \]
+
+  \hl{TODO: show stability}
+  % For the stability of $\B_{S(j)}(z)$ we must show that
+  % $\forall z: |z^{-1}| \le 1$ we have $$||\B_{S(j)}(z)||_2 < 1.$$
+\end{proof}
+
 \begin{proposition}
@@
   We will prove by way of contradiction.  To this end, suppose that
-  $j$ is a node such that: $(a)$ $j \not \in \anc{i}$ and $(b)$ for
-  every $k \in \anc{i} \cap \anc{j}$ every
-  $k \rightarrow \cdots \rightarrow j$ path contains $i$.
+  $j$ is a node such that:
+
+  \begin{enumerate}[label=(\alph*)]
+    \item{$j \not \in \anc{i}$}
+    \item{every $k \in \anc{i} \cap \anc{j}$ every
+        $k \rightarrow \cdots \rightarrow j$ path contains $i$.}
+  \end{enumerate}
 
@@
   $i \not \in \anc{k}$ and therefore there must be a path
-  $\gcgpath{w}{k} \rightarrow u$ which does not contain $i$, contradicting
-  property $(b)$.
-
-  Using this partition, and in a similar fashion to the proof of
-  Proposition \ref{prop:ancestor_uncorrelated}, we will expand
-  $x_j(t)$ in terms of it's parents, and recursively expand nodes in
-  $C_1$ until reaching $i$.  For the first step equation
-  \eqref{eqn:parent_expansion} gives us:
-
-  \begin{equation}
-    \label{eqn:xj_partition_expansion}
-    x_j(t) = \widetilde{\A}_{jj}(z) \Big(v_j(t) + \sum_{k \in C_0(j)} \B_{ik}(z)x_k(t) + \sum_{k \in C_1(j)}\B_{ik}(z)x_k(t)\Big).
-  \end{equation}
-
-  Using this representation we choose an arbitrary element of
-  $\H_{t - 1}^j$ and represent it as the action of a strictly causal
-  filter $\Phi(z)$, i.e.  $\Phi(z) x_j(t) \in \H_{t - 1}^{(j)}$ and show that
+  $\gcgpath{w}{k} \rightarrow u$ which does not contain $i$,
+  contradicting property $(b)$.  Moreover, for any
+  $u \in \big(\anc{j} \setminus \{i\}\big)$ and $k \in C_0(u)$,
+  Proposition \ref{prop:ancestor_uncorrelated} shows that
+  $H_t^i \perp \H_{t - 1}^j | \H_{t - 1}^i$.
+
+  In order to establish $j \npwgc i$, choose an arbitrary element of
+  $\H_{t - 1}^j$ and represent it via the action of a strictly causal
+  filter $\Phi(z)$, i.e.  $\Phi(z) x_j(t) \in \H_{t - 1}^{(j)}$, by
+  Theorem \ref{thm:granger_causality_equivalences} it suffices to show
+  that
 
@@
     \label{eqn:sufficient_inner_prod}
-    \inner{x_i(t)}{\widetilde{\A}_{jj}(z)\Phi(z)x_j(t) - \linE{\widetilde{\A}_{jj}(z)\Phi(z)x_j(t)}{\H_{t - 1}^{(i)}}} = 0,
+    \inner{x_i(t)}{\Phi(z)x_j(t) - \linE{\Phi(z)x_j(t)}{\H_{t - 1}^{(i)}}} = 0.
   \end{equation}
 
-  which will imply (by Theorem
-  \ref{thm:granger_causality_equivalences}) that $j \npwgc i$.
-
-  Substituting (\ref{eqn:xj_partition_expansion}) into
-  (\ref{eqn:sufficient_inner_prod}) we have cancellation of all terms
-  in which $k \in C_0(j)$ by Proposition \ref{prop:ancestor_properties},
-  and cancellation of the term involving $v_j(t)$ by Lemma \ref{lem:vj_perp}.
-
-  We are therefore left with
-
-  \begin{equation}
-    \label{eqn:refwornv}  % An arbitrary code for reference in just one place
-    \inner{x_i(t)}{\sum_{k \in C_1(j)} \big(\widetilde{\A}_{jj}(z)\Phi(z) \B_{ik}(z)x_k(t) - \linE{\widetilde{\A}_{jj}(z)\Phi(z) \B_{ik}(z)x_k(t)}{\H_{t - 1}^i}\big)}.
-  \end{equation}
-
-  Since each such $k$ inherits the needed properties $(a), (b)$ we can
-  again (as in Proposition \ref{prop:ancestor_properties}) recursively
-  expand this expression until we reach $i$, obtaining a strictly
-  causal $\Psi(z)$ such that Expression \eqref{eqn:refwornv} is equal to
-
-  \begin{equation}
-    \inner{x_i(t)}{\Psi(z)x_i(t) - \linE{\Psi(z)x_i(t)}{\H_{t - 1}^i}} = 0.
-  \end{equation}
+  Denote $e_j \defeq e_j^{S(j)}$, we can write
+  $x_j(t) = e_j^\T x_{S(j)}(t)$, and therefore from Equation
+  \eqref{eqn:scc_parent_expansion_inverted} there exist strictly
+  causal filters $\Gamma_s(z)$ and $\Lambda_{sk}(z)$ (defined for ease
+  of notation) such that
+
+  \[
+    x_j(t) = \sum_{s \in S(j)} \Gamma_s(z)v_s(t) + \sum_{\substack{s \in S(j) \\ k \in \pa{s} \cap S(j)^\c}} \Lambda_{sk}(z)x_k(t).
+  \]
+
+  When we substitute this expression into the left hand side of
+  Equation \eqref{eqn:sufficient_inner_prod}, we may cancel each term
+  involving $v_s$ by Lemma \ref{lem:vj_perp}, and each $k \in C_0(s)$
+  by our earlier argument, leaving us with
+
+  \[
+    \sum_{\substack{s \in S(j) \\ k \in C_1(s) \cap S(j)^\c}}\inner{x_i(t)}{\Phi(z)\Lambda_{sk}(z)x_k(t) - \linE{\Phi(z)\Lambda_{sk}(z)x_k(t)}{\H_{t - 1}^{(i)}}}.
+  \]
+
+  Since each $k \in C_1(s)$ with $k \ne i$ inherits properties $(a)$
+  and $(b)$ above, we can recursively expand each $x_k$ of the above
+  summation until reaching $k = i$ (which is garaunteed to terminate
+  due to the definition of $C_1(u)$) which leaves us with some
+  strictly causal filter $F(z)$ such that the left hand side of
+  Equation \eqref{eqn:sufficient_inner_prod} is equal to
+
+  \[
+    \inner{x_i(t)}{\Phi(z)F(z)x_i(t) - \linE{\Phi(z)F(z)x_i(t)}{\H_{t - 1}^{(i)}}},
+  \]
+
+  and this is $0$ since $\Phi(z)F(z)x_i(t) \in \H_{t - 1}^i$.
+
+  % Using the above partition of $\pa{j}$, we can expand $x_j(t)$ in
+  % terms of it's parents, and recursively expand nodes in $C_1$
+  % (cancelling all nodes in $C_0$, and recalling that
+  % $C_2 = \emptyset$) until we reach $i$.
+
+  % Using this partition, we will expand $x_j(t)$ in terms of it's
+  % parents, and recursively expand nodes in $C_1$ until reaching $i$.
+  % For the first step equation \eqref{eqn:parent_expansion} gives us:
+
+  % \begin{equation}
+  %   \label{eqn:xj_partition_expansion}
+  %   x_j(t) = \widetilde{\A}_{jj}(z) \Big(v_j(t) + \sum_{k \in C_0(j)} \B_{ik}(z)x_k(t) + \sum_{k \in C_1(j)}\B_{ik}(z)x_k(t)\Big).
+  % \end{equation}
+
+  % Substituting (\ref{eqn:xj_partition_expansion}) into
+  % (\ref{eqn:sufficient_inner_prod}) we have cancellation of all terms
+  % in which $k \in C_0(j)$ by Proposition \ref{prop:ancestor_properties},
+  % and cancellation of the term involving $v_j(t)$ by Lemma \ref{lem:vj_perp}.
+
+  % We are therefore left with
+
+  % \begin{equation}
+  %   \label{eqn:refwornv}  % An arbitrary code for reference in just one place
+  %   \inner{x_i(t)}{\sum_{k \in C_1(j)} \big(\widetilde{\A}_{jj}(z)\Phi(z) \B_{ik}(z)x_k(t) - \linE{\widetilde{\A}_{jj}(z)\Phi(z) \B_{ik}(z)x_k(t)}{\H_{t - 1}^i}\big)}.
+  % \end{equation}
+
+  % Since each such $k$ inherits the needed properties $(a), (b)$ we can
+  % again (as in Proposition \ref{prop:ancestor_properties}) recursively
+  % expand this expression until we reach $i$, obtaining a strictly
+  % causal $\Psi(z)$ such that Expression \eqref{eqn:refwornv} is equal to
+
+  % \begin{equation}
+  %   \inner{x_i(t)}{\Psi(z)x_i(t) - \linE{\Psi(z)x_i(t)}{\H_{t - 1}^i}} = 0.
+  % \end{equation}
 \end{proof}

### 2019-07-23 paper_theory/paper_theory.tex
Substantial updates to the proof of the confounding theorem.

@@
 \newtheorem{example}{Example}
-\newtheorem{assumptions}{Assumptions}
+\newtheorem{assumption}{Assumption}
 
@@
 The aforementioned assumptions are quite weak.  The strongest
-assumption we require is finally that $\Sigma_v$ is a diagonal matrix,
-which is referred to as a lack of instantaneous feedback in $x(t)$.
-We formally state our setup as a definition, which is the setup for
-the remainder of the paper:
+assumption we require is finally that $\Sigma_v$ is a diagonal
+positive-definite matrix, which is referred to as a lack of
+instantaneous feedback in $x(t)$.  We formally state our setup as a
+definition, which is the setup for the remainder of the paper:
 
@@
   \eqref{eqn:ar_representation} where $v(t)$ is sequentially
-  uncorrelated and has a diagonal covariance matrix.  The $MA(\infty)$
-  representation of equation \eqref{eqn:wold} has $c(t) = 0$ and
-  $A(0) = I$.
+  uncorrelated and has a diagonal positive-definite covariance matrix.
+  The $MA(\infty)$ representation of equation \eqref{eqn:wold} has
+  $c(t) = 0$ and $A(0) = I$.
 \end{definition}
@@
 part $(a)$ of Proposition \ref{prop:ancestor_properties} via the
-notion of a strongly causal graph topology.  In this section, we
-complete a converse by adding the additional requirement we refer to
-as ``persistence''.
+notion of a strongly causal graph topology (see Proposition
+\ref{prop:pwgc_anc}).  In this section, we study conditions under
+which a converse to part $(b)$ will hold.
 
@@
 \end{definition}
-
-The interpretation of the following persistence condition is that
-each node stores some ``memory'' of the past.
 
@@
   be persistent since these naturally result in an equivalent
-  $\mathsf{MA}(\infty)$ representation, except for the pathological case
-  where $\B(z)$ is Nilpotent.  Indeed, in a $\VAR(1)$ model, if the
-  system matrix is diagonalizable and has at least two distinct
-  eigenvalues, then the system is persistent.  See Example
+  $\mathsf{MA}(\infty)$ representation, see Example
   \ref{ex:persistent_system}.
@@
   Moreover, persistence is not the weakest condition necessary for the
-  results of this section, the condition
-  $\tau_0(\A_{jk}) < \tau_\infty(\A_{ik})$ for each $i, j, k$ such that
-  $k \in \anc{i} \cap \anc{j}$ is enough.  The intuition being that nodes
-  $i$ and $j$ are not receiving temporally disjoint information from
-  $k$.
+  results of this section, the condition that for each $i, j$ there is
+  some $k \in \anc{i}\cap\anc{j}$ such that
+  $\tau_0(\A_{jk}) < \tau_\infty(\A_{ik})$ is enough.  The intuition
+  being that nodes $i$ and $j$ are not receiving temporally disjoint
+  information from $k$.
+
+  The etymology for the persistence condition can be explained by
+  supposing that the two nodes $i, j$ each have a loop (i.e.
+  $\B_{ii}(z) \ne 0, \B_{jj}(z) \ne 0$) then this autoregressive
+  component acts as ``memory'', and so the influence from the
+  confounder $k$ \textit{persists} in $x_i(t)$, and
+  $\tau_\infty(\A_{ik}) = \infty$ for each confounder is expected.
 \end{remark}
@@
   Consider a process $x(t)$ generated by the $\VAR(1)$
-  model\footnote{Recall that any $\VAR(p)$ model with $p < \infty$ can be
-    written as a $\VAR(1)$ model, so we lose little generality in
-    considering this case.}  having $\B(z) = Bz^{-1}$, we will examine
-  conditions that guarantee persistence.  Pick any
-  $i \in [n], j \in \anc{i} \setminus \{i\}$.  Then the stability of
-  $B$ allows us to write
+  model\footnote{Recall that any $\VAR(p)$ model with $p < \infty$ can
+    be written as a $\VAR(1)$ model, so we lose little generality in
+    considering this case.}  having $\B(z) = Bz^{-1}$.  If $B$ is
+  diagonalizable, and has at least $2$ distinct eigenvalues, then
+  $x(t)$ is persistent.
+
+  See the supplementary material for an analysis of this example.
+\end{example}
+
+In order to eliminate the possibility of a particular sort of
+cancellation, an ad-hoc assumption is required.  Strictly speaking,
+the persistence condition is not a necessary or sufficient condition
+for the following, but cases where the following fails to hold, and
+peristence \textit{does} hold, are unavoidable pathologies.
+
+\begin{assumption}[no-cancellation]
+  \label{ass:T_causality}
+  Fix $i, j \in [n]$ and let $H_i(z)$ be the strictly-causal filter
+  such that
 
   \begin{equation*}
-    \A(z) = \sum_{k = 0}^\infty B^k z^{-k},
+    H_i(z)x_i(t) = \linE{x_i(t)}{\H_{t - 1}^i},
   \end{equation*}
 
-  whereby we see that $\exists k > 0$ such that $[B^k]_{ij} \ne 0$
-  (since $j \in \anc{i}$).  Then consider
-
-  \begin{equation*}
-    \begin{aligned}
-      [B^{rk}]_{ij} &= e_i^\T B^{rk} e_j \\
-      &\overset{(a)}{=} \big((P^\T e_i)^\T J^{rk} P^{-1}e_j\big)\\
-      &= \tr [(P^\T e_i)^\T J^{rk} P^{-1}e_j]\\
-      &\overset{(b)}{=} \tr [(J^{rk}) (v u^\T)],
-    \end{aligned}
-  \end{equation*}
-
-  where $(a)$ utilizes the Jordan Normal Form of $B$, and $(b)$
-  denotes $u = P^\T e_i$ and $v = P^{-1}e_j$.  In order for
-  $\tau_\infty(\A_{ij}) < \infty$, there must be some $N > 1$ such
-  that $\forall r \ge N$, the above term is $0$.  This may be the case
-  for instance if $B$ is a nilpotent matrix.
-
-  Let us suppose now that $B$ is diagonalizable (i.e. $J$ is a
-  diagonal matrix) with at least $2$ distinct eigenvalues; in this
-  case $B$ is also \textit{not} nilpotent.  We can then rewrite the
-  above as
-
-  \begin{equation*}
-    f(r) \defeq \tr [(J^{rk}) (v u^\T)] = \sum_{\nu = 1}^n \lambda_\nu ^{rk} v_\nu u_\nu \defeq \sum_{\nu = 1}^n \lambda_\nu^{rk} \beta_\nu
-  \end{equation*}
-
-  where $\lambda_\nu$ denotes the eigenvalues of $B$ and
-  $\beta_\nu = u_\nu v_\nu$.  Note that $f(0) = 0$ since $i \ne j$ and
-  $u$ is a row of $P$ and $v$ is a column of $P^{-1}$.  Moreover,
-  $f(1) \ne 0$ by hypothesis.  But, in order for
-  $f(r) = 0\ \forall r \ge N$, it would need to be the case that
-
-  \begin{equation*}
-    \Dg(\bm{\lambda})^r \bm{\lambda} = Vz
-  \end{equation*}
-
-  had a solution in $z$ for every $r \ge N$, where $V$ is an
-  $n \times n - 1$ full-rank matrix whose columns span the nullspace of $\beta$,
-  and $\bm{\lambda} = (\lambda_1, \ldots, \lambda_n)$. That is,
-  iterates of $\Dg(\bm{\lambda})$ applied to $\bm{\lambda}$ would need to remain
-  inside $\beta$'s nullspace.  This would imply that
-
-  \begin{equation*}
-    VV^\dagger \bm{\lambda}^{r + 1} = \bm{\lambda}^{r + 1},
-  \end{equation*}
-
-  i.e. that $\bm{\lambda}^{r + 1}$ is an eigenvector of $VV^\dagger$
-  for an infinite number of integers $r$ (the exponentiation is to be
-  understood as a point wise operation).  However, since there can only
-  be a finite number of (unit length) eigenvectors, this cannot be the
-  case unless every eigenvalue $(\lambda_1, \ldots, \lambda_n)$ were
-  equal.
-
-  We see from this example that the collection of $\VAR(1)$ systems
-  which are not persistent are pathological, in the sense that their
-  system matrices have zero measure when viewed as a subset of $\R^{n^2}$.
-\end{example}
+  and similarly for $H_j(z)$.  Then define
+
+  \begin{equation}
+    \label{eqn:T_filter}
+    T_{ij}(z) = \sum_{k \in \anc{i} \cap \anc{j}}\sigma_k^2\A_{ik}(z^{-1})(1 - H_i(z^{-1}))(1 - H_j(z))\A_{jk}(z),
+  \end{equation}
+
+  where $\sigma_k^2 = \E v_k(t)^2$.
+
+  We will say that Assumption \ref{ass:T_causality} is satisfied if
+  for every $i, j \in [n]$, $T_{ij}(z)$ is either constant over $z$
+  (i.e. each $z^k$ coefficient for $k \in \Z \setminus \{0\}$ is $0$),
+  or is \textit{neither} causal (i.e. containing only $z^{-k}$ terms,
+  for $k \ge 0$) \textit{or} anti-causal (i.e. containing only $z^k$
+  terms, for $k \ge 0$).  Put succinctly, $T_{ij}(z)$ must be
+  two-sided.
+\end{assumption}
+
+\begin{remark}
+  Under the condition of persistence, the only way for Assumption
+  \ref{ass:T_causality} to fail is through cancellation in the terms
+  defining $T_{ij}(z)$.  For example, the condition is assured if
+  $x(t)$ is persistent, and there is only a single confounder.
+  Unfortunately, some pathological behaviour resulting from
+  confounding nodes seems to be unavoidable without some assumptions
+  about the parameters of the $MA(\infty)$ system defining $x(t)$.
+\end{remark}
 
@@
   \label{prop:persistence_converse}
-  Suppose $\gcg$ is a strongly causal DAG and that $x(t)$ is
-  persistent, then if there exists a $k$ which confounds $(i, j)$ we
-  have $i \pwgc j$ and $j \pwgc i$.
+  Fix $i, j \in [n]$ and suppose $\exists k \in \anc{i} \cap \anc{j}$
+  which confounds $i, j$.  Then, if $T_{ij}(z)$ is not causal we have
+  $j \pwgc i$, and if $T_{ij}(z)$ is not anti-causal we have
+  $i \pwgc j$.  Moreover, if Assumption \ref{ass:T_causality} is
+  satisfied, then $j \pwgc i \iff i \pwgc j$.
+  % Suppose that Assumption \ref{ass:T_causality} is satisfied.  Then if
+  % there exists a $k$ which confounds $(i, j)$ we have
+  % $i \pwgc j \implies j \pwgc i$.  Moreover, if $T_{ij}(z)$ is not
+  % constant, then $i \pwgc j$.
 \end{proposition}
@@
 \begin{remark}
-  The importance of this result is that it allows us to eliminate the
-  effects of confounding by removing bidirectional edges.
+  The importance of this result is that when $i \pwgc j$ is a result
+  of a condounder $k$, then $i \pwgc j \iff j \pwgc i$.  This
+  implies that in a strongly causal graph every bidirectional pairwise
+  causality relation must be the result of a confounder.  Therefore,
+  in a strongly causal graph, pairwise causality analysis is
+  \textit{immune to confounding} (since we can safely remove all
+  bidrectional edges).
 \end{remark}
@@
   \label{thm:scg_recovery}
-  If the Granger-causality graph $\gcg$ for persistent process $x(t)$
-  is a strongly causal DAG then $\gcg$ can be inferred from pairwise
-  causality tests.  The procedure can be carried out, assuming
-  we have an oracle for pairwise causality, via Algorithm
-  (\ref{alg:pwgr}).
+  If the Granger-causality graph $\gcg$ for the process $x(t)$ is a
+  strongly causal DAG and Assumption \ref{ass:T_causality} holds, then
+  $\gcg$ can be inferred from pairwise causality tests.  The procedure
+  can be carried out, assuming we have an oracle for pairwise
+  causality, via Algorithm (\ref{alg:pwgr}).
 \end{theorem}
@@
 pairwise relations (by Proposition \ref{prop:pwgc_anc}), and the
-requirement of persistence allows confounding to be eliminated by
-removing bidirectional edges(by Proposition
-\ref{prop:persistence_converse} and Corollary
-\ref{cor:bidirectional_edge}).  Without persistence, each confounded
-pair would give rise to $4$ possible pairwise topologies consistent
-with $\gcg$, one for each type of pairwise edge (no edge,
-unidirectional, bidirectional).
+no-cancellation condition of Assumption \ref{ass:T_causality} allows
+confounding to be eliminated by removing bidirectional edges (by
+Proposition \ref{prop:persistence_converse} and Corollary
+\ref{cor:bidirectional_edge}).  Without Assumption
+\ref{ass:T_causality}, then each confounded pair would give rise to
+$4$ possible pairwise topologies consistent with $\gcg$, one for each
+type of pairwise edge (no edge, unidirectional, bidirectional).
 

### 2019-07-23 paper_theory/proofs.tex
Substantial updates to the proof of the confounding theorem.

@@
   \label{def:scc}
-  In a graph $\gcg$, the subset $S \subseteq [n]$ is \textit{strongly
-    connected} if $\forall i, j \in S$, $i \in \anc{j}$ and
-  $j \in \anc{i}$.  We will denote by $S(j)$ (which may be the
-  singleton $\{j\}$) the largest strongly connected component (SCC)
-  containing $j$.  We will denote $x_{S(j)}(t)$ to be the vector of
-  processes
+  In a graph $\gcg$, the \textit{ordered} (by the natural ordering on
+  $\N$) subset $S \subseteq [n]$ is \textit{strongly connected} if
+  $\forall i, j \in S$, $i \in \anc{j}$ and $j \in \anc{i}$.  We will
+  denote by $S(j)$ (which may be the singleton $(j)$) the largest
+  strongly connected component (SCC) containing $j$.  We will denote
+  $x_{S(j)}(t)$ to be the vector of processes
 
@@
 
-  whose indices are sorted in ascending order.  Similarly, the
-  sub-filter of $\B(z)$ acting on $x_{S(j)}(t)$ will be denoted
-  $\B_{S(j)}(z)$.
+  whose indices are given the same (natural) ordering as $S(j)$.
+  Similarly, the sub-filter of $\B(z)$ acting on $x_{S(j)}(t)$ will be
+  denoted $\B_{S(j)}(z)$.
 \end{definition}
@@
     \label{eqn:scc_parent_expansion_inverted}
-    x_{S(j)}(t) = (I - \B_{S(j)}(z))^{-1} \big(v_{S(j)}(t) + \sum_{\substack{s \in S(j) \\ k \in \pa{s} \cap S(j)^\c}} B_{sk}(z)x_k(t)e_s^{S(j)}\big),
+    x_{S(j)}(t) = (I - \B_{S(j)}(z))^{-1} \big(v_{S(j)}(t) + \sum_{\substack{s \in S(j) \\ k \in \pa{s} \cap S(j)^\c}} B_{sk}(z)x_k(t)e_s^{S(j)}\big).
   \end{equation}
@@
   \[
-    x(t) = \B(z)x(t) + v(t).
+    x(t) = \B(z)x(t) + v(t),
   \]
 
-  \hl{TODO: show stability}
-  % For the stability of $\B_{S(j)}(z)$ we must show that
-  % $\forall z: |z^{-1}| \le 1$ we have $$||\B_{S(j)}(z)||_2 < 1.$$
+  which, when rearranged appropriately, can be written as
+
+  \[
+    \left[
+      \begin{array}{c}
+        x_{S(j)}(t)\\
+        x_{S(j)^c}(t)
+      \end{array}
+    \right] =
+    \left[
+      \begin{array}{cc}
+        \B_{S(j)}(z)& \B_{S(j), S(j)^\c}(z)\\
+        \B_{S(j)^\c, S(j)}(z)& \B_{S(j)^\c}(z)
+      \end{array}
+    \right]
+        \left[
+      \begin{array}{c}
+        x_{S(j)}(t)\\
+        x_{S(j)^c}(t)
+      \end{array}
+    \right] +
+    \left[
+      \begin{array}{c}
+        v_{S(j)}(t)\\
+        v_{S(j)^c}(t)
+      \end{array}
+    \right].
+  \]
+
+  Theorem \ref{thm:granger_causality_equivalences} is invoked in order
+  to restrict the summation to $k \in \pa{s}$ (since other elements
+  are $0$).
+
+  Now, we can partition $\gcg$ into it's maximal SCCs
+  $S_1, \ldots, S_N$, (one of which is $S(j)$) and then consider the
+  DAG formed on $N$ nodes with edges $I \rightarrow J$ included on the
+  condition that
+  $\exists j \in S_J, i \in S_I \text{ s.t. } i \in \anc{j}$.  By
+  topologically sorting this DAG, we obtain an ordering $\sigma$ of
+  $[n]$ such that $\B_\sigma(z)$ is block upper triangluar, with one
+  of it's diagonal blocks consisting of the (possibly reordered)
+  matrix $\B_{S(j)}(z)$.  So we have
+
+  \begin{align*}
+    \forall\ |z^{-1}| \le 1: \det \B(z) = \prod_{i = 1}^N \det \B_{S_i}(z) &\ne 0\\
+    \implies \forall\ |z^{-1}| \le 1: \det \B_{S(j)}(z) &\ne 0,
+  \end{align*}
+
+  and therefore $\B_{S(j)}(z)$ is stable, invertible, and Equation
+  \eqref{eqn:scc_inversion} holds.
 \end{proof}
@@
     &= \E|F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}|^2\\
-    &+ \inner{F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}}{\sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}}},
+    &+ \inner{F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}}{\sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}}}.
   \end{align*}
 
-  which by the Cauchy-Schwarz inequality is $0$ if and only if
+  Now since $F_{r + 1}(z) \ne 0$ by Theorem
+  \ref{thm:granger_causality_equivalences}, and
+  $F_{r + 1}(z)x_j(t) \not\in \H_{t - 1}^i$, we have by the
+  Cauchy-Schwarz inequality that this expression is equal to $0$ if
+  and only if
 
@@
 \end{proof}
+
+\begin{example}
+  Consider a process $x(t)$ generated by the $\VAR(1)$
+  model\footnote{Recall that any $\VAR(p)$ model with $p < \infty$ can
+    be written as a $\VAR(1)$ model, so we lose little generality in
+    considering this case.}  having $\B(z) = Bz^{-1}$.  If $B$ is
+  diagonalizable, and has at least $2$ distinct eigenvalues, then
+  $x(t)$ is persistent.
+
+  Pick any $i \in [n], j \in \anc{i} \setminus \{i\}$.  Then the
+  stability of $B$ allows us to write
+
+  \begin{equation*}
+    \A(z) = \sum_{k = 0}^\infty B^k z^{-k},
+  \end{equation*}
+
+  whereby we see that $\exists k > 0$ such that $[B^k]_{ij} \ne 0$
+  (since $j \in \anc{i}$).  Then consider
+
+  \begin{equation*}
+    \begin{aligned}
+      [B^{rk}]_{ij} &= e_i^\T B^{rk} e_j \\
+      &\overset{(a)}{=} \big((P^\T e_i)^\T J^{rk} P^{-1}e_j\big)\\
+      &= \tr [(P^\T e_i)^\T J^{rk} P^{-1}e_j]\\
+      &\overset{(b)}{=} \tr [(J^{rk}) (v u^\T)],
+    \end{aligned}
+  \end{equation*}
+
+  where $(a)$ utilizes the Jordan Normal Form of $B$, and $(b)$
+  denotes $u = P^\T e_i$ and $v = P^{-1}e_j$.  In order for
+  $\tau_\infty(\A_{ij}) < \infty$, there must be some $N > 1$ such
+  that $\forall r \ge N$, the above term is $0$.  This may be the case
+  for instance if $B$ is a nilpotent matrix.
+
+  Using the supposition that $B$ is diagonalizable (i.e. $J$ is a
+  diagonal matrix) with at least $2$ distinct eigenvalues (in this
+  case $B$ is \textit{not} nilpotent), we can then rewrite the above
+  as
+
+  \begin{equation*}
+    f(r) \defeq \tr [(J^{rk}) (v u^\T)] = \sum_{\nu = 1}^n \lambda_\nu ^{rk} v_\nu u_\nu \defeq \sum_{\nu = 1}^n \lambda_\nu^{rk} \beta_\nu
+  \end{equation*}
+
+  where $\lambda_\nu$ denotes the eigenvalues of $B$ and
+  $\beta_\nu = u_\nu v_\nu$.  Note that $f(0) = 0$ since $i \ne j$ and
+  $u$ is a row of $P$ and $v$ is a column of $P^{-1}$.  Moreover,
+  $f(1) \ne 0$ by hypothesis.  But, in order for
+  $f(r) = 0\ \forall r \ge N$, it would need to be the case that
+
+  \begin{equation*}
+    \Dg(\bm{\lambda})^r \bm{\lambda} = Vz
+  \end{equation*}
+
+  had a solution in $z$ for every $r \ge N$, where $V$ is an
+  $n \times n - 1$ full-rank matrix whose columns span the nullspace of $\beta$,
+  and $\bm{\lambda} = (\lambda_1, \ldots, \lambda_n)$. That is,
+  iterates of $\Dg(\bm{\lambda})$ applied to $\bm{\lambda}$ would need to remain
+  inside $\beta$'s nullspace.  This would imply that
+
+  \begin{equation*}
+    VV^\dagger \bm{\lambda}^{r + 1} = \bm{\lambda}^{r + 1},
+  \end{equation*}
+
+  i.e. that $\bm{\lambda}^{r + 1}$ is an eigenvector of $VV^\dagger$
+  for an infinite number of integers $r$ (the exponentiation is to be
+  understood as a point wise operation).  However, since there can only
+  be a finite number of (unit length) eigenvectors, this cannot be the
+  case unless every eigenvalue $(\lambda_1, \ldots, \lambda_n)$ were
+  equal.
+
+  We see from this example that the collection of $\VAR(1)$ systems
+  which are not persistent are pathological, in the sense that their
+  system matrices have zero measure when viewed as a subset of $\R^{n^2}$.
+\end{example}
 
@@
   autocorrelation and let $\A(z), \B(z)$ be nonzero and strictly
-  causal (i.e. $\tau_0(\A) \ge 1$) linear filters.  Then,
+  causal (i.e. $1 \le \tau_0(\A) < \infty$,
+  $1 \le \tau_0(\B) < \infty$) linear filters.  Then,
 
@@
     \inner{\A(z)v(t)}{\B(z)v(t)} &= \sum_{\tau = 1}^\infty \sum_{s = 1}^\infty a(\tau)b(s)\E[v(t - s)v(t - \tau)]\\
-    &= \sum_{\tau = \text{max}(\tau_0(\A), \tau_0(\B))}^{\text{min}(\tau_\infty(\A), \tau_\infty(\B))} a(\tau) b(\tau)\\
+    &= \sum_{\tau = \text{max}(\tau_0(\A), \tau_0(\B))}^{\text{min}(\tau_\infty(\A), \tau_\infty(\B))} a(\tau) b(\tau),\\
   \end{align}
@@
   \label{prop:persistence_converse}
-  Suppose $\gcg$ is a strongly causal DAG and that $x(t)$ is
-  persistent, then if there exists a $k$ which confounds $(i, j)$ we
-  have $i \pwgc j$ and $j \pwgc i$.
+  Fix $i, j \in [n]$ and suppose $\exists k \in \anc{i} \cap \anc{j}$
+  which confounds $i, j$.  Then, if $T_{ij}(z)$ is not causal we have
+  $j \pwgc i$, and if $T_{ij}(z)$ is not anti-causal we have
+  $i \pwgc j$.  Moreover, if Assumption \ref{ass:T_causality} is
+  satisfied, then $j \pwgc i \iff i \pwgc j$.
+  % Suppose that Assumption \ref{ass:T_causality} is satisfied.  Then if
+  % there exists a $k$ which confounds $(i, j)$ we have
+  % $i \pwgc j \implies j \pwgc i$.  Moreover, if $T_{ij}(z)$ is not
+  % constant, then $i \pwgc j$.
 \end{proposition}
+
+% QUESTION: Does $\gcg$ need to be a strongly causal DAG for this proposition?
+
+% \begin{proposition}
+%   \label{prop:persistence_converse}
+%   Suppose $\gcg$ is a strongly causal DAG and that $x(t)$ is
+%   persistent, then if there exists a $k$ which confounds $(i, j)$ we
+%   have $i \pwgc j \implies j \pwgc i$.
+% \end{proposition}
 \begin{proof}
-  We will show that $j \pwgc i$, the other being symmetric.  First
-  note that by the contrapositive of proposition
-  \ref{prop:sc_graph_common_anc} we cannot have $i \in \anc{j}$ or
-  $j \in \anc{i}$ (else $k$ would not be a confounder) and therefore
-  \textit{every} $u \in \anc{i}\cap\anc{j}$ will be a confounder.
-
-  It is sufficient to show that $\exists \psi \in \H_{t - 1}^{(j)}$
-  such that
-
-  \begin{equation*}
-    \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} \ne 0.
-  \end{equation*}
-
-  To this end, let $F(z)$ be an arbitrary but strictly causal linear
-  filter.  We apply equation \eqref{eqn:ancestor_expansion} to
-  $x_i(t)$ and $\psi \defeq F(z)x_j(t)$:
+  Recalling Theorem \ref{thm:granger_causality_equivalences}, consider
+  some $\psi \in \H_{t - 1}^j$ and represent it as
+  $\psi(t) = F(z)x_j(t)$ for some strictly causal filter $F(z)$.
+  Then
 
   \begin{align*}
-    &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}\\
-    &\overset{(a)}{=} \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
-    &\overset{(b)}{=} \inner{\sum_{k \in \anc{j}}\big(F(z)\A_{jk}(z)v_k(t) \\&- \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}\big)}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
-    &\overset{(c)}{=} \sum_{k \in \anc{i}\cap\anc{j}}\Big(\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} \\&- \inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{ii}v_i(t)}\\
-    &- \sum_{\ell \in \anc{i}}\inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{i\ell}(z)v_\ell(t)}\Big)
+    &\inner{\psi(t) - \linE{\psi(t)}{\H_{t - 1}^i}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}\\
+    &\overset{(a)}{=} \inner{F(z)x_j(t)}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}\\
+    &\overset{(b)}{=} \inner{F(z)\big(\A_{jj}(z)v_j(t) + \sum_{k \in \anc{j}}\A_{jk}(z)v_k(t)\big)}{(1 - H_i(z))\big(\A_{ii}(z)v_i(t) + \sum_{\ell \in \anc{i}}\A_{i\ell}(z)v_\ell(t)\big)}\\
+    &\overset{(c)}{=} \sum_{k \in \anc{i}\cap\anc{j}}\inner{F(z)\A_{jk}(z)v_k(t)}{(1 - H_i(z))\A_{ik}(z)v_k(t)},
   \end{align*}
 
-  where in $(a)$ we have removed the $\linE{x_i(t)}{\H_{t - 1}^{(i)}}$
-  term via the orthogonality principle, in $(b)$ there is no
-  $F(z)\A_{jj}(z)v_j(t)$ term since due to $j \not\in \anc{i}$ it is
-  orthogonal to $\H_t^{(i)}$.  Finally, $(c)$ follows by applying
-  orthogonality properties of $v(t)$, as well as the fact that
-  $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$ for
-  $k \not \in \anc{i}$.  Note that
-  $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$
-  and therefore there is in general no cancellation in the final term
-  above for $\ell \in \anc{i}$.
-
-  \todo{This is clearly not immediately evident}
-
-  This is $0$ for every $F$ if and only if for all $F$ and
-  $\forall k \in \anc{i} \cap \anc{j}$ we have
-  $$\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$$ and
-  $$\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} = 0,$$ which by
-  Corollary \ref{cor:time_lag_cancellation} occurs if and only if
-  $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$, which is impossible since by persistence
-  $\tau_0(\A_{jk}) < \infty$ and  $\tau_\infty(\A_{ik}) = \infty$.
+  where $(a)$ applies the orthogonality principle, $(b)$ expands with
+  Equation \eqref{eqn:ancestor_expansion} with
+  $H_i(z)x_i(t) = \linE{x_i(t)}{H_{t - 1}^i}$, and $(c)$ follows by
+  performing cancellations of $v_k(t) \perp v_\ell(t)$ since by the
+  contrapositive of Proposition \ref{prop:sc_graph_common_anc} we
+  cannot have $i \in \anc{j}$ or $j \in \anc{i}$.
+
+% , $(c)$ is obtained through cancellations from Lemma \ref{lem:vj_perp}, $(d)$ follows since $\H_i$ is self-adjoint, $(e)$ follows by definition of $\H_i$: $\exists H_i(z)$ such that $$H_i(z)x_i(t) = \H_i \sum_{\ell \in \anc{i}}\A_{i\ell}v_\ell(t),$$ $(f)$ follows by again expanding $x_i(t)$ with Equation \eqref{eqn:ancestor_expansion}, and finally $(g)$ by cancelling orthogonal components of $v(t)$.
+
+  Through symmetric calculation, we can obtain the expression relevant
+  to the determination of $i \pwgc j$ for $\phi \in \H_{t - 1}^i$ represented by the strictly causal filter $G(z): \phi(t) = G(z)x_i(t)$
+  \begin{align*}
+    &\inner{\phi(t) - \linE{\phi(t)}{\H_{t - 1}^j}}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^j}}\\
+    &= \sum_{k \in \anc{i} \cap \anc{j}}\inner{G(z)\A_{ik}(z)v_k(t)}{(1 - H_j(z))\A_{jk}(z)v_k(t)},
+  \end{align*}
+
+  where $H_j(z)x_j(t) = \linE{x_j(t)}{\H_{t - 1}^j}.$
+
+  We have therefore
+
+  \begin{align}
+      &(j \pwgc i): \exists F(z) \text{ s.t. } \sum_{k \in \anc{i} \cap \anc{j}}\inner{F(z)\A_{jk}(z)v_k(t)}{(1 - H_i(z))\A_{ik}(z)v_k(t)} \ne 0,\\
+      &(i \pwgc j): \exists G(z) \text{ s.t. } \sum_{k \in \anc{i} \cap \anc{j}}\inner{G(z)\A_{ik}(z)v_k(t)}{(1 - H_j(z))\A_{jk}(z)v_k(t)} \ne 0.
+  \end{align}
+
+  The persistence condition, by Corollary
+  \ref{cor:time_lag_cancellation}, ensures that for each
+  $k \in \anc{i}\cap\anc{j}$ there is some $F(z)$ and some $G(z)$ such
+  that at least one of the above terms constituting the sum over $k$
+  is non-zero.  It remains to eliminate the possibility of
+  cancellation in the sum.
+
+  The adjoint of a linear filter $C(z)$ is simply $C(z^{-1})$, which
+  recall is strictly anti-causal if $C(z)$ is strictly causal.  Using
+  this, we can write
+
+  \begin{align*}
+    &\sum_{k \in \anc{i} \cap \anc{j}}\inner{F(z)\A_{jk}(z)v_k(t)}{(1 - H_i(z))\A_{ik}(z)v_k(t)}\\
+    = &\sum_{k \in \anc{i} \cap \anc{j}}\inner{\A_{ik}(z^{-1})(1 - H_i(z^{-1}))F(z)\A_{jk}(z)v_k(t)}{v_k(t)}\\
+  \end{align*}
+
+  Moreover, it is sufficient to find some strictly causal $F(z)$ of
+  the form $F(z)(1 - H_j(z))$ (abusing notation) since $1 - H_j(z)$ is
+  causal.  Similarly for $G(z)$, this leads to symmetric expressions
+  for $j \pwgc i$ and $i \pwgc j$ respectively:
+
+  \begin{equation}
+    \label{eqn:T_F}
+    \sum_{k \in \anc{i} \cap \anc{j}}\inner{\A_{ik}(z^{-1})(1 - H_i(z^{-1}))F(z)(1 - H_j(z))\A_{jk}(z)v_k(t)}{v_k(t)},
+  \end{equation}
+  \begin{equation}
+    \label{eqn:T_G}
+    \sum_{k \in \anc{i} \cap \anc{j}}\inner{\A_{ik}(z^{-1})(1 - H_i(z^{-1}))G(z^{-1})(1 - H_j(z))\A_{jk}(z)v_k(t)}{v_k(t)}.
+  \end{equation}
+
+  Recall the filter from Assumption \ref{ass:T_causality}
+
+  \begin{equation}
+    T_{ij}(z) = \sum_{k \in \anc{i} \cap \anc{j}} \sigma_k^2\A_{ik}(z^{-1})(1 - H_i(z^{-1}))(1 - H_j(z))\A_{jk}(z).
+  \end{equation}
+
+  Since each $v_k(t)$ is uncorrelated through time,
+  $\inner{T_{ij}(z)v_k(t)}{v_k(t)} = \sigma_k^2T_{ij}(0)$, and
+  therefore we have $j \pwgc i$ if $T_{ij}(z)$ is \textit{not} causal
+  and $i \pwgc j$ if $T_{ij}(z)$ it \textit{not} anti-causal.
+  Moreover, we have $i \npwgc j$ \textit{and} $j \pwgc i$ if
+  $T_{ij}(z)$ is a constant.  Therefore, under Assumption
+  \ref{ass:T_causality} $j \pwgc i \implies i \pwgc j$.
+
+  This follows since if $T_{ij}(z)$ is not causal then
+  $\exists k > 0$ such that the $z^k$ coefficient of $T_{ij}(z)$ is
+  non-zero, and we can choose strictly causal $F(z) = z^{-k}$ such
+  that \eqref{eqn:T_F} is non-zero and therefore $j \pwgc i$.
+
+  Similarly, if $T_{ij}(z)$ is not anti-causal, then
+  $\exists k > 0$ such that the $z^{-k}$ coefficient of $T_{ij}(z)$ is
+  non-zero, and we can choose strictly causal $G(z)$ so that
+  $G(z^{-1}) = z^k$, and then \refeq{eqn:T_G} is non-zero and therefore
+  $i \pwgc j$.
 \end{proof}
+
+% This proof is not correct -- there can be cancellation
+% \begin{proof}
+%   We will show that $j \pwgc i$, the other being symmetric.  First
+%   note that by the contrapositive of proposition
+%   \ref{prop:sc_graph_common_anc} we cannot have $i \in \anc{j}$ or
+%   $j \in \anc{i}$ (else $k$ would not be a confounder) and therefore
+%   \textit{every} $u \in \anc{i}\cap\anc{j}$ will be a confounder.
+
+%   It is sufficient to show that $\exists \psi \in \H_{t - 1}^{(j)}$
+%   such that
+
+%   \begin{equation*}
+%     \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} \ne 0.
+%   \end{equation*}
+
+%   To this end, let $F(z)$ be an arbitrary but strictly causal linear
+%   filter.  We apply equation \eqref{eqn:ancestor_expansion} to
+%   $x_i(t)$ and $\psi \defeq F(z)x_j(t)$:
+
+%   \begin{align*}
+%     &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}\\
+%     &\overset{(a)}{=} \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
+%     &\overset{(b)}{=} \inner{\sum_{k \in \anc{j}}\big(F(z)\A_{jk}(z)v_k(t) \\&- \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}\big)}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
+%     &\overset{(c)}{=} \sum_{k \in \anc{i}\cap\anc{j}}\Big(\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} \\&- \inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{ii}v_i(t)}\\
+%     &- \sum_{\ell \in \anc{i}}\inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{i\ell}(z)v_\ell(t)}\Big)
+%   \end{align*}
+
+%   where in $(a)$ we have removed the $\linE{x_i(t)}{\H_{t - 1}^{(i)}}$
+%   term via the orthogonality principle, in $(b)$ there is no
+%   $F(z)\A_{jj}(z)v_j(t)$ term since due to $j \not\in \anc{i}$ it is
+%   orthogonal to $\H_t^{(i)}$.  Finally, $(c)$ follows by applying
+%   orthogonality properties of $v(t)$, as well as the fact that
+%   $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$ for
+%   $k \not \in \anc{i}$.  Note that
+%   $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$
+%   and therefore there is in general no cancellation in the final term
+%   above for $\ell \in \anc{i}$.
+
+%   \todo{This is clearly not immediately evident}
+
+%   This is $0$ for every $F$ if and only if for all $F$ and
+%   $\forall k \in \anc{i} \cap \anc{j}$ we have
+%   $$\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$$ and
+%   $$\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} = 0,$$ which by
+%   Corollary \ref{cor:time_lag_cancellation} occurs if and only if
+%   $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$, which is impossible since by persistence
+%   $\tau_0(\A_{jk}) < \infty$ and  $\tau_\infty(\A_{ik}) = \infty$.
+% \end{proof}
 

### 2019-07-24 paper/paper.tex
LASSO paragraph.

--- 2019-07-14 paper/paper.tex
@@
 \cite{grouped_graphical_granger_modelling_for_gene_expression_regulatory_networks_discovery}.
-One of the key improvements to the original LASSO algorithm is the
-adaptive (i.e. weighted) ``adaLASSO'' \cite{adaptive_lasso_zou2006},
-for which oracle results (i.e. asymptotic support recovery) are
-established under less restrictive conditions than for the vanilla
-LASSO.  Our experimental comparisons in Section
+
+The LASSO algorithm is known to suffer from a large proportion of
+false-positives (e.g. see \cite{fdr_lasso}), except when certain
+stringent conditions are met.  One of the key improvements to the
+original LASSO algorithm is the adaptive (i.e. weighted) ``adaLASSO''
+\cite{adaptive_lasso_zou2006}, for which oracle results
+(i.e. asymptotic support recovery) are established under less
+restrictive conditions.  Our experimental comparisons in Section
 \ref{sec:empirical_evaluation} are against the
-adaLASSO. % Success of the LASSO has lead to
-% the famous ``bet on sparsity'' principle ``Use a procedure that does
-% well in sparse problems, since no procedure does well in dense
-% problems.''  \cite{tibshirani2015statistical}.
+adaLASSO.
 

### 2019-07-24 paper_theory/supplementary_material.tex
Some renaming and prepping for submission.

--- 2019-07-23 paper_theory/proofs.tex
@@
 \AtAppendix{\counterwithin{equation}{section}}
-\externaldocument{paper_theory}
+\externaldocument{graph_topology_granger_causality}
 
@@
   $H_i(z)x_i(t) = \linE{x_i(t)}{H_{t - 1}^i}$, and $(c)$ follows by
-  performing cancellations of $v_k(t) \perp v_\ell(t)$ since by the
-  contrapositive of Proposition \ref{prop:sc_graph_common_anc} we
-  cannot have $i \in \anc{j}$ or $j \in \anc{i}$.
+  performing cancellations of $v_k(t) \perp v_\ell(t)$ and noting that
+  by the contrapositive of Proposition \ref{prop:sc_graph_common_anc}
+  we cannot have $i \in \anc{j}$ or $j \in \anc{i}$.
 
@@
     &\sum_{k \in \anc{i} \cap \anc{j}}\inner{F(z)\A_{jk}(z)v_k(t)}{(1 - H_i(z))\A_{ik}(z)v_k(t)}\\
-    = &\sum_{k \in \anc{i} \cap \anc{j}}\inner{\A_{ik}(z^{-1})(1 - H_i(z^{-1}))F(z)\A_{jk}(z)v_k(t)}{v_k(t)}\\
+    = &\sum_{k \in \anc{i} \cap \anc{j}}\inner{\A_{ik}(z^{-1})(1 - H_i(z^{-1}))F(z)\A_{jk}(z)v_k(t)}{v_k(t)}.\\
   \end{align*}
@@
   $T_{ij}(z)$ is a constant.  Therefore, under Assumption
-  \ref{ass:T_causality} $j \pwgc i \implies i \pwgc j$.
+  \ref{ass:T_causality} $j \pwgc i \iff i \pwgc j$.
 
@@
   confounded and Corollary \ref{cor:parent_corollary} that
-  $j \not\in \anc{i}$ so $j \npwgc i$ and thence by Proposition
+  $j \not\in \anc{i}$ so $j \npwgc i$ by Proposition and therefore
   \ref{prop:ancestor_properties} $(i, j) \in W$.
 
-  Conversely, suppose $(i, j) \in W$.  Then since $j \npwgc i$
+  Conversely, suppose $(i, j) \in W$.  Then since $j \npwgc i$,
   Proposition \ref{prop:persistence_converse} ensures that $(j, i)$
@@
 
-\begin{lemma}[Depth Characterization of $P_k$]
+\begin{lemma}[Depth Characterization of $P_k$ and $S_k$]
   \label{lem:depth_lemma}
@@
   $i \not\in P_{k - 1}$ so then
-  $i \in S_k = S_{k - 1} \setminus P_{k - 1}$ and thus $d(i) \ge k$.
-  Now, recalling the definition of $P_k$
+  $i \in S_k = S_{k - 1} \setminus P_{k - 1}$.  Now, recalling the
+  definition of $P_k$
 
@@
   if $s \in S_k$ is such that $(s, i) \in W$ then $s \pwgc i$ and
-  $i \npwgc s$ so that by persistence and Proposition
-  \ref{prop:persistence_converse} there cannot be a confounder of
-  $(s, i)$ (otherwise $i \pwgc s$) so then by Proposition
-  \ref{prop:ancestor_properties} we have $s \in \anc{i}$.  We have
-  shown that $s \in S_k \implies d(s) \ge k$ and so we must have
-  $d(i) > k$, a contradiction, thence $s \not\in \anc{i}$,
-  $s \npwgc i$, $(s, i) \not\in W$ and $i \in P_k$.
+  $i \npwgc s$ so that by Proposition \ref{prop:persistence_converse}
+  there cannot be a confounder of $(s, i)$ (otherwise $i \pwgc s$) so
+  then by Proposition \ref{prop:ancestor_properties} we have
+  $s \in \anc{i}$.  We have shown that $s \in S_k \implies d(s) \ge k$
+  and so we must have $d(i) > k$, a contradiction, therefore there is
+  no such $s \in S_k$ so $i \in P_k$.
 
@@
   let $r = 1$ and suppose that $(i, j) \in \gcge$ with $d(j) = k$ and
-  $d(i) = k - 1$.  Then, $(i, j) \in W$ and by our assumptions on
-  $E_{k - 1}$ there is no $\gcgpath{i}{j}$ path in $E_{k - 1}$
-  and therefore $(i, j) \in D_{k1}$.  Conversely, suppose that
-  $(i, j) \in D_{k1}$.  Then, $d(i) = k - 1$ and $d(j) = k$ which, since
-  $(i, j) \in W \implies i \in \anc{j}$ implies that
-  $i \in \pa{j}$ and $(i, j) \in \gcge$.
+  $d(i) = k - 1$.  Then, by Corollary \ref{cor:gc_implies_pwgc}
+  $(i, j) \in W$ and by our assumptions on $E_{k - 1}$ there is no
+  $\gcgpath{i}{j}$ path in $E_{k - 1}$ and therefore
+  $(i, j) \in D_{k1}$.  Conversely, suppose that $(i, j) \in D_{k1}$.
+  Then, $d(i) = k - 1$ and $d(j) = k$ which, since
+  $(i, j) \in W \implies i \in \anc{j}$ implies that $i \in \pa{j}$
+  and $(i, j) \in \gcge$.
 
@@
   \ref{lem:inner_loop_lemma} $(i, j) \in \bigcup_{r = 1}^k D_{kr}$
-  which implies that $(i, j) \in E_k$.  The result follows.
+  which implies that $(i, j) \in E_k$.
 \end{proof}
@@
 \begin{lemma}[Finite Termination]
-  Algorithm \ref{alg:pwgr} terminates and returns $E_{k^\star - 1} = \gcge$
-  for some $k^\star \le n$.
+  Algorithm \ref{alg:pwgr} terminates and returns the set
+  $E_{k^\star - 1} = \gcge$ for some $k^\star \le n$.
 \end{lemma}
@@
   H_0:&\ \widehat{x}_i^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau),\\
-  H_1:&\ \widehat{x}_i^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau) + \sum_{\tau = 1}^pb_{ij}(\tau)x_j(t - \tau).
+  H_1:&\ \widehat{x}_{i|j}^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau) + \sum_{\tau = 1}^pb_{ij}(\tau)x_j(t - \tau).
 \end{align}
@@
 Section \ref{sec:model_order_selection}.  We assume that this
-statistic is asymptotically $\chi^2(p_{ij})$ distributed (the
+statistic is asymptotically $\chi^2(p_{ij})$ distributed (e.g. the
 disturbances are Gaussian), and denote by $G$ the cumulative
@@
 \clearpage
-\bibliography{\string~/Documents/academics/global_academics/global_bib}
-
+% \bibliography{\string~/Documents/academics/global_academics/global_bib}
+\bibliography{paper_bib}
 \end{document}

### 2019-07-24 software/pwgc/gc_methods.py

@@ -655,7 +655,7 @@ def pw_scg(F, P_edge, alpha, eliminate_bidirectional=False):
         W_set = {(i, j) for (i, j) in product(S, S)
                  if (F[j, i] >= F[i, j] and
                      P_edge[j, i] > 0 and
-                     P_edge[i, j] == 0)}
+                     P_edges[i, j] == 0)}
     else:
         W_set = {(i, j) for (i, j) in product(S, S)
                  if (F[j, i] >= F[i, j] and

### 2019-07-24 software/pwgc/test_gc_methods.py

@@ -68,7 +68,7 @@ def test_pw_scg():
     n_nodes, p_lags, p_max = 35, 5, 10
     alpha = 0.05
 
-    T_max = 10000
+    T_max = 1000
 
     random_graph = lambda: random_scg(
         n_nodes, p_lags, pole_rad=0.75)
@@ -90,7 +90,7 @@ def test_pw_scg():
         P_edges = normalize_gc_score(F, P, T=len(X), F_distr=F_distr)
         P_values = 1 - P_edges[~np.eye(n_nodes, dtype=bool)].ravel()
         t_bh = benjamini_hochberg(P_values, alpha=alpha, independent=False)
-        G_hat = pw_scg(F, P_edges, t_bh, eliminate_bidirectional=bd)
+        G_hat = pw_scg(F, P_edges, t_bh, eliminate_bidirectional=False)
         draw_graph_estimates(G, G_hat)
         return
 
@@ -116,8 +116,8 @@ def test_pw_scg():
     # est_graph(X, fast=True)
     # est_graph(X, bd=False, F_distr=True)
     # est_graph(X, bd=False, F_distr=False)
-    est_graph(X, bd=False)
-    est_graph(X, bd=True)
+    est_graph(X, bd=True, F_distr=False)
+    est_graph(X, bd=False, F_distr=False)
     return
 
 

### 2019-07-26 paper_theory/graph_topology_granger_causality.tex
Some renaming and prepping for submission.
Essentially at the final version -- pending Ravi

--- 2019-07-23 paper_theory/paper_theory.tex
@@
 \usepackage{xr}
-\externaldocument{proofs}
-% \usepackage[T1]{fontenc}
-% \usepackage[lite]{belleek}
-% \fontfamily{belleek}
-% \usepackage[math]{belleek}
-% \usepackage{concretemath-fonts}
-% \interfootnotelinepenalty=10000
+\externaldocument{supplementary_material}
 % ---------------------------------------------------------------------------------------------------------------------------------
@@
 
-% TODO: Are we, in this setting, immune to confounders?  We are removing
-%       all bi-directional edges, so it seems likely that we are?  Can we state
-%       any theorem about this?
-
 %% bibliography stuff -- this needs come before the preamble inclusion
@@
 \def\VAR{\mathsf{VAR}}  % VAR(p) model
+\def\MA{\mathsf{MA}}  % MA(p) model
 \def\B{\mathsf{B}}  % Filter B
@@
 Though the underlying causality graph cannot be observed directly,
-it's it's presence is inferred as a latent structure among observed
+it's presence is inferred as a latent structure among observed
 time series data.  This notion is leveraged in a variety of
@@
 report a summary of simulation results in \ref{sec:simulation}, with
-additional results reported in the supplementary material
+additional results reported in the supplementary material Section
 \ref{apx:simulation}.  Our simulation results establish that there is
@@
   \begin{aligned}
-    \H_t^x &= \cl \{\sum_{\tau = 0}^p a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, p \in \N\} \subseteq L_2(\Omega)\\
-    H_t^x &= \{a x(t)\ |\ a \in \R\} \subseteq L_2(\Omega),
+    \H_t^{(x)} &= \cl \{\sum_{\tau = 0}^p a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, p \in \N\} \subseteq L_2(\Omega)\\
+    H_t^{(x)} &= \{a x(t)\ |\ a \in \R\} \subseteq L_2(\Omega),
   \end{aligned}
@@
 $\H_t^{(i, j)}$, $\H_t^{i}$ or by all but a particular component as
-$\H_t^{-j}$.
+$\H_t^{(-j)}$.
 
@@
 e.g. \cite{lindquist}), every WSS sequence has the moving average
-$MA(\infty)$ representation
+$\MA(\infty)$ representation
 
@@
 
-The equations (\ref{eqn:wold}), (\ref{eqn:ar_representation}) can be
+The Equations \eqref{eqn:wold}, \eqref{eqn:ar_representation} can be
 represented as $x(t) = \A(z)v(t) = \B(z)x(t) + v(t)$ via the action
 (convolution) of the operators (LTI filters)
-$\A(z) \defeq \sum_{\tau = 0}^\infty A(\tau)z^{-\tau}$ and
-$\B(z) \defeq \sum_{\tau = 1}^\infty B(\tau)z^{-\tau}$ where the
+$$\A(z) \defeq \sum_{\tau = 0}^\infty A(\tau)z^{-\tau}$$ and
+$$\B(z) \defeq \sum_{\tau = 1}^\infty B(\tau)z^{-\tau},$$ where the
 operator $z^{-1}$ is the back shift operator acting on
@@
   uncorrelated and has a diagonal positive-definite covariance matrix.
-  The $MA(\infty)$ representation of equation \eqref{eqn:wold} has
+  The $\MA(\infty)$ representation of Equation \eqref{eqn:wold} has
   $c(t) = 0$ and $A(0) = I$.
@@
 \begin{equation}
-  \linEerr{x_i(t)}{\H_{t - 1}} < \linEerr{x_i(t)}{\H^{-j}_{t - 1}},
+  \linEerr{x_i(t)}{\H_{t - 1}} < \linEerr{x_i(t)}{\H^{(-j)}_{t - 1}},
 \end{equation}
@@
     \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$ i.e. $\B_{ij}(z) = 0$}
-    \item{$H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
-    \item{$\linE{x_i(t)}{\H_{t - 1}^{-j}} = \linE{x_i(t)}{\H_{t - 1}}$}
+    \item{$H_t^{i} \perp \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(-j)}$}
+    \item{$\linE{x_i(t)}{\H_{t - 1}^{(-j)}} = \linE{x_i(t)}{\H_{t - 1}}$}
   \end{enumerate}
@@
 notion of ``weight'' by associating an edge $(j, i)$ with the
-\textit{strictly causal} LTI filter $\B_{ij}(z)$ (see eqn
+\textit{strictly causal} LTI filter $\B_{ij}(z)$ (see Equation
 \eqref{eqn:filter_action}).  Thence, the matrix $\B(z)$ is analogous
@@
 a time lag of $k$ steps, this is exemplified in the inversion formula
-\ref{eqn:lsi_inversion}.
+\eqref{eqn:lsi_inversion}.
 
@@
   where $\A(z) = \sum_{\tau = 0}^\infty A(\tau)z^{-\tau}$ is the filter from
-  the Wold decomposition representation of $x(t)$, equation
-  (\ref{eqn:wold}).
+  the Wold decomposition representation of $x(t)$, Equation
+  \eqref{eqn:wold}.
 \end{proposition}
@@
   $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.  Moreover,
-  this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_t^i} = 0$.
+  this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_t^{(i)}} = 0$.
 \end{proposition}
-
-\begin{remark}
-  The possibility that there exist nodes sharing both $i$ and $j$ as
-  ancestors is not excluded, the point being that the temporal nature
-  of Granger-causality eliminates the problems caused by ``colliders''
-  (in the language of Pearl's causal calculus
-  \cite{pearl2009causality}) in static causal inference.
-\end{remark}
 
@@
   that is not available from observing $i$ itself.
-  % However, it is important to note that this is the case only if we
-  % have direct access to $x(t)$.  That is, if we instead observed a
-  % noisy version of $x$ (i.e. $\tilde{x}(t) = x(t) + n(t)$) the
-  % observation of $\tilde{x}_j$ would enable conditionally improved
-  % estimates of $\tilde{x}_i$, without there necessarily being any
-  % causal flow in the underlying system $x(t)$.
 \end{remark}
@@
       \label{fig:gene_network1}
-      \includegraphics[width=\linewidth, height=\linewidth]{ecoli_regulatory_network.png}
+      \includegraphics[width=\linewidth, height=\linewidth]{ecoli_regulatory_network.pdf}
     \end{subfigure}
@@
       \label{fig:gene_network2}
-      \includegraphics[width=\linewidth, height=\linewidth]{huge_gene_network.png}
+      \includegraphics[width=\linewidth, height=\linewidth]{huge_gene_network.pdf}
     \end{subfigure}
@@
 
-\begin{remark}
-  As we have seen and as is true in much of statistics, confounding
-  nodes pose challenges for Granger-causality.  However, as opposed to
-  the causal calculus of \cite{pearl2009causality}, pairwise
-  Granger-causality does not suffer any difficulty with so-called
-  ``colliders'', that is, the topology $i \rightarrow k \leftarrow j$ will never result
-  in $i \pwgc j$ or $j \pwgc i$.  This is evidently an advantage of
-  the \textit{temporal} nature of Granger-causality -- there is no
-  backwards causal flow along the edges of $\gcg$.
-\end{remark}
-
 \begin{example}
@@
     \tau_0(\B) &= \text{inf }\{\tau \in \Z_+\ |\ b(\tau) \ne 0\},\\
-    \tau_{\infty}(\B) &= \text{sup }\{\tau \in \Z_+\ |\ b(\tau) \ne 0\}.\\
+    \tau_{\infty}(\B) &= \text{sup }\{\tau \in \Z_+\ |\ b(\tau) \ne 0\}.
   \end{align}
@@
 for the following, but cases where the following fails to hold, and
-peristence \textit{does} hold, are unavoidable pathologies.
+persistence \textit{does} hold, are unavoidable pathologies.
 
@@
   \begin{equation*}
-    H_i(z)x_i(t) = \linE{x_i(t)}{\H_{t - 1}^i},
+    H_i(z)x_i(t) = \linE{x_i(t)}{\H_{t - 1}^{(i)}},
   \end{equation*}
@@
   confounding nodes seems to be unavoidable without some assumptions
-  about the parameters of the $MA(\infty)$ system defining $x(t)$.
+  about the parameters of the $\MA(\infty)$ system defining $x(t)$.
 \end{remark}
@@
   The importance of this result is that when $i \pwgc j$ is a result
-  of a condounder $k$, then $i \pwgc j \iff j \pwgc i$.  This
+  of a confounder $k$, then $i \pwgc j \iff j \pwgc i$.  This
   implies that in a strongly causal graph every bidirectional pairwise
@@
   \textit{immune to confounding} (since we can safely remove all
-  bidrectional edges).
+  bidirectional edges).
 \end{remark}
@@
 be ``close'' to SCGs in some appropriate sense.  As $q$ increases,
-there are simply not enough edges allowed by the SGC topology for it
+there are simply not enough edges allowed by the SCG topology for it
 to be possible to accurately recover $\gcg$.
@@
 application to very large networks of time series data which are not
-well approximated by finite $\VAR(p)$ models.
-
-% \bibliography{\string~global_bib.bib}
-\bibliography{\string~/Documents/academics/global_academics/global_bib}
+well approximated by finite $\VAR(p)$ models.  Moreover, our
+heuristics are in principle applicable in a model-free context.  As
+long as a primitive for testing the pairwise causation between two
+components of a multivariate time series is available, our methods may
+be useful.
+
+\bibliography{paper_bib}
 \end{document}

### 2019-07-26 paper_theory/supplementary_material.tex
Essentially at the final version -- pending Ravi

@@
 \usepackage{chngcntr}
-\usepackage{apptools}
-\AtAppendix{\counterwithin{lemma}{section}}
-\AtAppendix{\counterwithin{definition}{section}}
-\AtAppendix{\counterwithin{equation}{section}}
+\usepackage{appendix}
+\newcounter{lemma}
+\newcounter{definition}
+\counterwithin{lemma}{section}
+\counterwithin{definition}{section}
+\counterwithin{equation}{section}
 \externaldocument{graph_topology_granger_causality}
@@
   \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$ i.e. $\B_{ij}(z) = 0$}
-  \item{$H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
-  \item{$\linE{x_i(t)}{\H_{t - 1}^{-j}} = \linE{x_i(t)}{\H_{t - 1}}$}
+  \item{$H_t^{(i)} \perp \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(-j)}$}
+  \item{$\linE{x_i(t)}{\H_{t - 1}^{(-j)}} = \linE{x_i(t)}{\H_{t - 1}}$}
   \end{enumerate}
@@
   of the model).  $(b) \Rightarrow (c)$ follows since in computing
-  $(y - \linE{y}{\H_{t - 1}^{-j}})$ for $y \in H_t^i$ it is sufficient
+  $(y - \linE{y}{\H_{t - 1}^{(-j)}})$ for $y \in H_t^{(i)}$ it is sufficient
   to consider $y = x_i(t)$ by linearity, then since
-  $H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$ we have
-  $(x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{-j}}) = v_i(t)$ since
+  $H_{t - 1}^{(i)} \subseteq \H_{t - 1}^{(-j)}$ we have
+  $(x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(-j)}}) = v_i(t)$ since
   $\B_{ij}(z) = 0$.  The result follows since
@@
 
-  From equation (\ref{eqn:ar_representation}), which we are assuming
+  From Equation \eqref{eqn:ar_representation}, which we are assuming
   throughout the paper to be invertible, we can write
@@
   Let $S$ be a matrix with elements in $\{0, 1\}$ which represents the
-  sparsity pattern of $\B(z)$, from lemma \ref{lem:adj_matrix} $S$ is
+  sparsity pattern of $\B(z)$, from Lemma \ref{lem:adj_matrix} $S$ is
   the transpose of the adjacency matrix for $\gcg$ and hence
@@
   $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.  Moreover,
-  this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_t^i} = 0$.
+  this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_t^{(i)}} = 0$.
 \end{proposition}
@@
   Consider distinct nodes $i, j$ in a Granger-causality graph $\gcg$.
-  If $j \not \in \anc{i}$, then $\H_t^{v_j} \perp \H_{t}^{i}$, and therefore
+  If $j \not \in \anc{i}$, then $\H_t^{(v_j)} \perp \H_{t}^{(i)}$, and therefore
   for any causal filter $\Phi(z)$ we have
@@
   \begin{align*}
-    \linE{\Phi(z)v_j(t)}{\H_{t - 1}^i} &= 0\\
+    \linE{\Phi(z)v_j(t)}{\H_{t - 1}^{(i)}} &= 0\\
     \inner{x_i(t)}{\Phi(z)v_j(t)} & = 0.
@@
 \begin{proof}
-  Fix $\tau, s \ge 0$, then by expanding with Equation \ref{eqn:ancestor_expansion}
+  Fix $\tau, s \ge 0$, then by expanding with Equation \eqref{eqn:ancestor_expansion}
 
@@
   \begin{equation*}
-    \forall \psi \in \H_{t - 1}^j\ \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}{\psi - \linE{\psi}{\H_{t - 1}^i}} = 0.
+    \forall \psi \in \H_{t - 1}^{(j)}\ \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}} = 0.
   \end{equation*}
@@
   which by the orthogonality principle and by representing
-  $\psi \in \H_{t - 1}^j$ via the action of some strictly causal filter
+  $\psi \in \H_{t - 1}^{(j)}$ via the action of some strictly causal filter
   $\Phi(z)$ on $x_j(t)$ is equivalent to
@@
     \label{eqn:proof_inner0}
-    \inner{x_i(t)}{\Phi(z)x_j(t) - \linE{\Phi(z)x_j(t)}{\H_{t - 1}^i}} = 0.
+    \inner{x_i(t)}{\Phi(z)x_j(t) - \linE{\Phi(z)x_j(t)}{\H_{t - 1}^{(i)}}} = 0.
   \end{equation}
@@
   \begin{equation*}
-    \inner{x_i(t)}{\sum_{k \in \anc{j} \cup \{j\}} \Big(\Phi(z)\A_{jk}(z)v_k(t) - \linE{\Phi(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^i}\Big)}.
+    \inner{x_i(t)}{\sum_{k \in \anc{j} \cup \{j\}} \Big(\Phi(z)\A_{jk}(z)v_k(t) - \linE{\Phi(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}\Big)}.
   \end{equation*}
@@
   Proposition \ref{prop:ancestor_uncorrelated} shows that
-  $H_t^i \perp \H_{t - 1}^j | \H_{t - 1}^i$.
+  $H_t^{(i)} \perp \H_{t - 1}^{(j)} | \H_{t - 1}^{(i)}$.
 
   In order to establish $j \npwgc i$, choose an arbitrary element of
-  $\H_{t - 1}^j$ and represent it via the action of a strictly causal
+  $\H_{t - 1}^{(j)}$ and represent it via the action of a strictly causal
   filter $\Phi(z)$, i.e.  $\Phi(z) x_j(t) \in \H_{t - 1}^{(j)}$, by
@@
 
-  and this is $0$ since $\Phi(z)F(z)x_i(t) \in \H_{t - 1}^i$.
+  and this is $0$ since $\Phi(z)F(z)x_i(t) \in \H_{t - 1}^{(i)}$.
 
@@
 
-  and therefore that $H_t^{(i)} \not\perp\ \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(i)}$, which by theorem (\ref{thm:granger_causality_equivalences}) is enough to establish that $j \pwgc i$.
+  and therefore that $H_t^{(i)} \not\perp\ \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(i)}$, which by Theorem (\ref{thm:granger_causality_equivalences}) is enough to establish that $j \pwgc i$.
 
   Firstly, we will establish a representation of $x_i(t)$ that involves $x_j(t)$.
-Denote by $a_{r + 1} \rightarrow a_r \rightarrow \cdots \rightarrow a_1 \rightarrow a_0$ with $a_{r + 1} \defeq j$ and $a_0 \defeq i$ the \textit{unique} $\gcgpath{j}{i}$ path in $\gcg$, we will expand the representation of equation (\ref{eqn:parent_expansion}) backwards along this path:
+Denote by $a_{r + 1} \rightarrow a_r \rightarrow \cdots \rightarrow a_1 \rightarrow a_0$ with $a_{r + 1} \defeq j$ and $a_0 \defeq i$ the \textit{unique} $\gcgpath{j}{i}$ path in $\gcg$, we will expand the representation of Equation \eqref{eqn:parent_expansion} backwards along this path:
 
@@
 
-  Using this representation to expand equation (\ref{eqn:cond_ortho_proof}), we obtain the following cumbersome expression:
+  Using this representation to expand Equation \eqref{eqn:cond_ortho_proof}, we obtain the following cumbersome expression:
 
@@
   \ref{thm:granger_causality_equivalences}, and
-  $F_{r + 1}(z)x_j(t) \not\in \H_{t - 1}^i$, we have by the
+  $F_{r + 1}(z)x_j(t) \not\in \H_{t - 1}^{(i)}$, we have by the
   Cauchy-Schwarz inequality that this expression is equal to $0$ if
@@
 
-  which can be expanded by equation \eqref{eqn:ancestor_expansion} to
+  which can be expanded by Equation \eqref{eqn:ancestor_expansion} to
   obtain (after cancelling all ancestors of $i$ other than $k$)
@@
   Recalling Theorem \ref{thm:granger_causality_equivalences}, consider
-  some $\psi \in \H_{t - 1}^j$ and represent it as
+  some $\psi \in \H_{t - 1}^{(j)}$ and represent it as
   $\psi(t) = F(z)x_j(t)$ for some strictly causal filter $F(z)$.
@@
   \begin{align*}
-    &\inner{\psi(t) - \linE{\psi(t)}{\H_{t - 1}^i}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}\\
-    &\overset{(a)}{=} \inner{F(z)x_j(t)}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}\\
+    &\inner{\psi(t) - \linE{\psi(t)}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}\\
+    &\overset{(a)}{=} \inner{F(z)x_j(t)}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}\\
     &\overset{(b)}{=} \inner{F(z)\big(\A_{jj}(z)v_j(t) + \sum_{k \in \anc{j}}\A_{jk}(z)v_k(t)\big)}{(1 - H_i(z))\big(\A_{ii}(z)v_i(t) + \sum_{\ell \in \anc{i}}\A_{i\ell}(z)v_\ell(t)\big)}\\
@@
   Equation \eqref{eqn:ancestor_expansion} with
-  $H_i(z)x_i(t) = \linE{x_i(t)}{H_{t - 1}^i}$, and $(c)$ follows by
+  $H_i(z)x_i(t) = \linE{x_i(t)}{H_{t - 1}^{(i)}}$, and $(c)$ follows by
   performing cancellations of $v_k(t) \perp v_\ell(t)$ and noting that
@@
   Through symmetric calculation, we can obtain the expression relevant
-  to the determination of $i \pwgc j$ for $\phi \in \H_{t - 1}^i$ represented by the strictly causal filter $G(z): \phi(t) = G(z)x_i(t)$
+  to the determination of $i \pwgc j$ for $\phi \in \H_{t - 1}^{(i)}$ represented by the strictly causal filter $G(z): \phi(t) = G(z)x_i(t)$
   \begin{align*}
-    &\inner{\phi(t) - \linE{\phi(t)}{\H_{t - 1}^j}}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^j}}\\
+    &\inner{\phi(t) - \linE{\phi(t)}{\H_{t - 1}^{(j)}}}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^{(j)}}}\\
     &= \sum_{k \in \anc{i} \cap \anc{j}}\inner{G(z)\A_{ik}(z)v_k(t)}{(1 - H_j(z))\A_{jk}(z)v_k(t)},
@@
 
-  where $H_j(z)x_j(t) = \linE{x_j(t)}{\H_{t - 1}^j}.$
+  where $H_j(z)x_j(t) = \linE{x_j(t)}{\H_{t - 1}^{(j)}}.$
 
@@
 equations:\footnote{The particular indexing and normalization given in
-  equation \ref{eqn:covariance_estimate} is critical to ensure
+  Equation \eqref{eqn:covariance_estimate} is critical to ensure
   $\widehat{R}$ is positive semidefinite.  The estimate can be viewed
@@
 \label{sec:error_rate_control}
-Denote $F_{ij}$ the Granger-causality statistic of equation
-\ref{eqn:gc_statistics} with model orders chosen by the methods of
+Denote $F_{ij}$ the Granger-causality statistic of Equation
+\eqref{eqn:gc_statistics} with model orders chosen by the methods of
 Section \ref{sec:model_order_selection}.  We assume that this
@@
     \Input{Estimates of pairwise Granger-causality statistics $F_{ij}$
-      (eqn \ref{eqn:gc_statistics}).
-Matrix of edge probabilities $P_{ij}$ (eqn \ref{eqn:edge_inclusion_probability}).
+      (eqn.
+\ref{eqn:gc_statistics}).
+Matrix of edge probabilities $P_{ij}$ (eqn.
+\ref{eqn:edge_inclusion_probability}).
 Hypothesis testing threshold $\delta$ chosen via the Benjamini-Hochberg criterion (Section \ref{sec:error_rate_control})}
@@
 
-In reference to figure \ref{fig:simulation_results_comparison1} it
+In reference to Figure \ref{fig:simulation_results_comparison1} it
 should not be overly surprising that our PWGC algorithm performs
@@
 \clearpage
-% \bibliography{\string~/Documents/academics/global_academics/global_bib}
 \bibliography{paper_bib}

### 2019-07-26 paper_appl/paper_appl.tex
Developing a substantive application.

--- 2019-07-17 paper/paper.tex
@@
 \usepackage{todonotes}
+\usepackage{booktabs}
+\usepackage{multirow}
 \bibliography{\string~/Documents/academics/global_academics/global_bib.bib}
@@
 
-\title{Graph Topological Aspects of Granger Causal Network Learning }
+\title{A Study in Pairwise Testing Heuristics for Granger-causal Network Estimation}
 \author{R. J. Kinnear \\
@@
   \small\url{https://github.com/RJTK} \and R. R. Mazumdar \\
-  \small\href{mailto:mazum@uwaterloo.ca}{mazum@uwaterloo.ca} }
+  \small\href{mailto:mazum@uwaterloo.ca}{mazum@uwaterloo.ca}}
 
@@
 \maketitle
-\abstract{We study Granger-causality in the context of wide-sense
-  stationary time series, where our focus is on the topological
-  aspects of the underlying causality graph.  We establish sufficient
-  conditions (in particular, we develop the notion of a ``strongly
-  causal'' graph topology) under which the true causality graph can be
-  recovered via pairwise causality testing alone, and provide examples
-  from the gene regulatory network literature suggesting that our
-  concept of a strongly causal graph may be applicable to this field.
-  We implement and detail finite-sample heuristics derived from our
-  theory, and establish through simulation the efficiency gains (both
-  statistical and computational) which can be obtained (in comparison
-  to LASSO algorithms) when structural assumptions are met.  Finally,
-  we provide an example application where we accurately discriminate
-  between subjects in an EEG study based purely on the
-  Granger-causality graphs inferred from data by our algorithms,
-  demonstrating that meaningful features are captured by the
-  Granger-causal graph topology.}
-
-\keywords{causality graph, EEG, gene regulatory networks,
-  Granger-causality, LASSO, network learning, time series, vector
-  autoregression}
+\abstract{Following recent theoretical developments of
+  Granger-causality graph estimation for ``strongly causal'' causality
+  graphs, we develop in detail a computationally and statistically
+  efficient heuristic for Granger-causality estimation which is based
+  purely on pairwise causality testing.  We perform detailed
+  comparisons against state of the art algorithms in the LASSO family
+  through a simulation study, and demonstrate the capacity for scaling
+  to thousands of nodes in a time series network.  We develop an
+  application example for EEG data where we accurately discriminate
+  between subjects based purely on the Granger-causality graphs
+  inferred from data by our algorithms, demonstrating that meaningful
+  features are being captured in the Granger-causalilt graph topology,
+  as well as providing evidence that individuals exhibit meaningfully
+  distinct EEG causality structures, which may be a result of
+  independent scientific interest.}
+
+\keywords{causality graph, EEG, Granger-causality, LASSO, network
+  learning, time series, vector autoregression}
 
 \paragraph{Acknowledgement}
-
 We acknowledge the support of the Natural Sciences and Engineering Research Council of Canada (NSERC), [funding reference number 518418-2018].
@@
 \label{sec:introduction}
-% Consider a collection of stochastic processes producing observations
-% at discrete time intervals.  Are the underlying processes dependent?
-% Can we quantify any of the underlying relationships?  Can the arrow of
-% time help us to distinguish a directionality or flow of dependence
-% among our observed series?  In this paper we contribute to the
-% understanding of the notion of Granger-Causality
-% \cite{granger1969investigating} \cite{Granger1980329} as a tool for
-% answering these questions.
-
-% Heuristically, we state that if an event '$A$' occuring strictly prior
-% in time to an even '$B$' and $A$ provides us with exclusive (that is,
-% not available anywhere else) information about $B$, then $A$ must have
-% had a causal impact on $B$.  This is an entirely model-free notion of
-% cause, and instead leverages the intuition that a cause must precede
-% it's effect.
-
-In this paper we study the notion of Granger-causality
-\cite{granger1969investigating} \cite{Granger1980329} as a means of
-uncovering an underlying causal structure in multivariate time series.
-Though the underlying causality graph cannot be observed directly, we
-will infer it's presence as a latent structure among our observed time
-series data.  This notion is leveraged in a variety of applications
-e.g. in Neuroscience as a means of recovering interactions amongst
-brain regions \cite{bressler2011wiener}, \cite{anna_paper2008},
-\cite{david2008identifying}; in the study of the dependence and
-connectedness of financial institutions \cite{NBERw16223}; gene
-expression networks \cite{Fujita2007},
+Following recent theoretical developments \cite{my_GC_paper}
+establishing that, under certain topological assumptions on the
+underlying causality graph, it is possible to accurately recover the
+true causality graph by way of pairwise causality tests alone.  That
+is, there are algorithms which are immune to the effects of
+confounders when applied to certain causality graph structures.  In
+this paper we provide simulation evidence that, given structural
+assumptions, such algorithms can dramatically outperform state of the
+art LASSO-type estimators.  Moreover, even in cases where structural
+assumptions do not in fact hold, using similar algorithms as pure
+heuristics can still exhibit superior performance.  In addition to the
+potential for gains in statistical power, that our algorithms make use
+only of pairwise causality testing improves the scalability (both in
+terms of the asymptotics, as well as through trivial parallelization)
+of Granger-causality testing up to thousands of nodes.
+
+Granger-causality is a notion is leveraged in a variety of
+applications e.g. in Neuroscience as a means of recovering
+interactions amongst brain regions \cite{bressler2011wiener},
+\cite{anna_paper2008}, \cite{david2008identifying}; in the study of
+the dependence and connectedness of financial institutions
+\cite{NBERw16223}; gene expression networks \cite{Fujita2007},
 \cite{methods_for_inferring_gene_regulatory_networks_from_time_series_expression_data},
@@
 
-\subsection{Contributions}
-In the context of time series data, sparsity assumptions remain
-important, but there is significant additional structure that may
-arise as a result of considering the topology of the underlying
-Granger-causality graph, which to our knowledge remains largely
-unexplored.  The focus of this paper is to shed light on some of these
-topological questions, in particular, we study a particularly simple
-notion of causality graph topology which we term ``strongly causal''
-and show that stationary times series whose underlying causality graph
-has this structure satisfy natural intuitive notions of ``information
-flow'' through the graph.  Moreover, we show that such graphs are
-perfectly recoverable with only \textit{pairwise} Granger-causality
-tests, which would otherwise suffer from serious confounding problems.
-Our finite sample results are based on simulations, where we show
-promising results for these graph topologies where our algorithm
-performs substantially better in our simulation setup than do
-competing LASSO algorithms, even for graphs that do not exactly
-satisfy our strongly-causal topology assumptions.
-
-In the case of gene expression networks, we show examples from the
-literature which suggest our concept of a ``strongly causal graph''
-topology may have application in this field (see Section
-\ref{sec:strongly_causal_graphs}).
-
-The principle contributions of this paper are as follows: firstly, in
-section \ref{sec:theory} we study \textit{pairwise} Granger-causality
-relations, providing novel theorems connecting the structure of the
-causality graph to the pairwise ``causality flow'' in the system, as
-well as an interpretation in terms of the graph topology of the
-sparsity pattern of matrices arising in the Wold decomposition,
-generalizing in some sense the notion of ``feedback-free'' processes
-studied by \cite{caines1975feedback} in close connection with
-Granger-causality.  We establish sufficient conditions (sections
-\ref{sec:strongly_causal_graphs}, \ref{sec:persistent_systems}) under
-which a fully conditional Granger-causality graph can be recovered
-from pairwise tests alone (sec \ref{sec:pairwise_algorithm}).
-Secondly, we propose in section \ref{sec:structure_learning} a graph
-search heuristic which implements our theoretical results to finite
-data samples, specifying and summarizing appropriate methods for
-hypothesis testing (section \ref{sec:pairwise_hypothesis_testing}),
-model order selection (section \ref{sec:model_order_selection}),
-computationally efficient estimation (section
-\ref{sec:efficient_model_estimation}), and error rate controls
-(section \ref{sec:error_rate_control}).  Our heuristics are compared
-against the adaLASSO algorithm in Section
-\ref{sec:empirical_evaluation}.  We stress the scalability of our
-algorithm which is capable of comfortably handling hundreds or
-thousands of nodes on a single machine, as opposed to standard LASSO
-algorithms which do not take advantage of the special structure
-associated with stationary time series data.  In section
-\ref{sec:application} we develop an example application where a
-classifier is constructed to accurately discriminate between subjects
-in an EEG study based only on the Granger-causality graph topology
-inferred by our algorithms.  Concluding remarks on further open
-problems and extentions are provided in Section \ref{sec:conclusion}.
-
-\section{Theory}
+% \subsection{Contributions}
+% In the context of time series data, sparsity assumptions remain
+% important, but there is significant additional structure that may
+% arise as a result of considering the topology of the underlying
+% Granger-causality graph, which to our knowledge remains largely
+% unexplored.  The focus of this paper is to shed light on some of these
+% topological questions, in particular, we study a particularly simple
+% notion of causality graph topology which we term ``strongly causal''
+% and show that stationary times series whose underlying causality graph
+% has this structure satisfy natural intuitive notions of ``information
+% flow'' through the graph.  Moreover, we show that such graphs are
+% perfectly recoverable with only \textit{pairwise} Granger-causality
+% tests, which would otherwise suffer from serious confounding problems.
+% Our finite sample results are based on simulations, where we show
+% promising results for these graph topologies where our algorithm
+% performs substantially better in our simulation setup than do
+% competing LASSO algorithms, even for graphs that do not exactly
+% satisfy our strongly-causal topology assumptions.
+
+% In the case of gene expression networks, we show examples from the
+% literature which suggest our concept of a ``strongly causal graph''
+% topology may have application in this field (see Section
+% \ref{sec:strongly_causal_graphs}).
+
+% The principle contributions of this paper are as follows: firstly, in
+% section \ref{sec:theory} we study \textit{pairwise} Granger-causality
+% relations, providing novel theorems connecting the structure of the
+% causality graph to the pairwise ``causality flow'' in the system, as
+% well as an interpretation in terms of the graph topology of the
+% sparsity pattern of matrices arising in the Wold decomposition,
+% generalizing in some sense the notion of ``feedback-free'' processes
+% studied by \cite{caines1975feedback} in close connection with
+% Granger-causality.  We establish sufficient conditions (sections
+% \ref{sec:strongly_causal_graphs}, \ref{sec:persistent_systems}) under
+% which a fully conditional Granger-causality graph can be recovered
+% from pairwise tests alone (sec \ref{sec:pairwise_algorithm}).
+% Secondly, we propose in section \ref{sec:structure_learning} a graph
+% search heuristic which implements our theoretical results to finite
+% data samples, specifying and summarizing appropriate methods for
+% hypothesis testing (section \ref{sec:pairwise_hypothesis_testing}),
+% model order selection (section \ref{sec:model_order_selection}),
+% computationally efficient estimation (section
+% \ref{sec:efficient_model_estimation}), and error rate controls
+% (section \ref{sec:error_rate_control}).  Our heuristics are compared
+% against the adaLASSO algorithm in Section
+% \ref{sec:empirical_evaluation}.  We stress the scalability of our
+% algorithm which is capable of comfortably handling hundreds or
+% thousands of nodes on a single machine, as opposed to standard LASSO
+% algorithms which do not take advantage of the special structure
+% associated with stationary time series data.  In section
+% \ref{sec:application} we develop an example application where a
+% classifier is constructed to accurately discriminate between subjects
+% in an EEG study based only on the Granger-causality graph topology
+% inferred by our algorithms.  Concluding remarks on further open
+% problems and extentions are provided in Section \ref{sec:conclusion}.
+
+\section{Granger Causality}
 \label{sec:theory}
-\subsection{Formal Setting}
-Consider the space $L_2(\Omega)$, the usual Hilbert space of finite
-variance random variables over a probability space
-$(\Omega, \mathcal{F}, \mathbb{P})$ having inner product
-$\inner{x}{y} = \E[xy]$.  We will work with a discrete time and
-wide-sense stationary (WSS) $n$-dimensional vector valued process
-$x(t)$ (with $t \in \Z$) where the $n$ elements take values in $L_2$.  We
-suppose that $x(t)$ has zero mean, $\E x(t) = 0$, and has absolutely
-summable matrix valued covariance sequence
-$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ with an absolutely
-continuous spectral density $S(\omega)$:
+For the purposes of this paper, we provide an expediant and simplified
+definition of Granger-causality.  For a more detailed exposition see
+e.g. \cite{my_GC_paper}, \cite{granger1969investigating},
+\cite{geweke1982measurement}, \cite{lutkepohl2005new}.  Suppose that $x(t)$ is an
+$n-$dimensional wide-sense stationary time series generated by the
+following (stable) $VAR(\infty)$ model
 
 \begin{equation}
-  \label{eqn:fourier_pair}
-  \begin{aligned}
-    R(\tau) &= \frac{1}{2\pi}\int_{-\pi}^\pi S(\omega) e^{j\tau\omega}\d \omega,\\
-    S(\omega) &= \sum_{\tau=-\infty}^\infty R(\tau)e^{-j\tau\omega}.
-  \end{aligned}
-\end{equation}
-
-% and that the spectra is bounded uniformly away from 0:
-
-% \begin{equation*}
-%   \int_{-\pi}^\pi \ln\det S(\omega) \d\omega > -\infty.
-% \end{equation*}
-
-We will also work frequently with the spaces spanned by the values of
-such a process
-
-\begin{equation}
-  \label{eq:hilbert_space_defn}
-  \begin{aligned}
-    % \H^x &= \cl \{\sum_{\tau = -T}^T a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, T \in \N\} \subseteq L_2(\Omega),\\
-    \H_t^x &= \cl \{\sum_{\tau = 0}^p a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, p \in \N\} \subseteq L_2(\Omega)\\
-    H_t^x &= \{a x(t)\ |\ a \in \R\} \subseteq L_2(\Omega),
-  \end{aligned}
-\end{equation}
-
-where the closure is naturally in mean-square.  We will often omit the
-superscript $x$ which should be clear from context.  Evidently these
-spaces are separable, and as closed subspaces of a Hilbert space they
-are themselves Hilbert.  We will denote the spaces generated in
-analogous ways by particular components of $x$ as e.g.
-$\H_t^{(i, j)}$, $\H_t^{i}$ or by all but particular components as
-$\H_t^{-j}$.
-
-As a consequence of the Wold decomposition theorem \cite{lindquist},
-every WSS sequence has the moving average $MA(\infty)$
-representation
-
-\begin{equation}
-\label{eqn:wold}
-  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) v(t - \tau),
-\end{equation}
-
-where $c(t)$ is a purely deterministic sequence\footnote{the purely
-  deterministic sequence $c(t)$ is one which lies in the remote past
-  $\bigcap_{\tau=1}^\infty \H_{t - \tau}^x$ of the process.  For such
-  processes a single sample $c(t_0)$ is enough to determine $c(t)$ for
-  every $t$.  For example,
-  $c(t) = \text{sin}(2\pi t + \Theta);\; \Theta \sim \mathcal{U}[-\pi,
-  \pi]$}, $v(t)$ is an uncorrelated sequence and $A(0) = I$.  We will
-assume that $c(t) = 0$, which in practice is to say that the process
-has been detrended.  We additionally require that this representation
-can be inverted to yield the $\VAR(\infty)$ form
-
-\begin{equation}
-  \label{eqn:ar_representation}
+  \label{eq:var_model}
   x(t) = \sum_{\tau = 1}^\infty B(\tau) x(t - \tau) + v(t).
@@
 
-The equations (\ref{eqn:wold}), (\ref{eqn:ar_representation}) can be
-represented as $x(t) = \A(z)v(t) = \B(z)x(t) + v(t)$ via the action
-(convolution) of the operators (LTI filters)
-$\A(z) \defeq \sum_{\tau = 0}^\infty A(\tau)z^{-\tau}$ and
-$\B(z) \defeq \sum_{\tau = 1}^\infty B(\tau)z^{-\tau}$ where the
-operator $z^{-1}$ is the back shift operator acting on
-$\ell_2^n(\Omega, \mathcal{F}, \mathbb{P})$, that is:
-
-\begin{equation}
-  \label{eqn:filter_action}
-  \B_{ij}(z)x_j(t) \defeq \sum_{\tau = 1}^\infty B_{ij}(\tau)x_j(t - \tau).
-\end{equation}
-
-Finally, since $||z^{-1}|| = 1$ we have the inversion formula
-
-\begin{equation}
-  \label{eqn:lsi_inversion}
-  \A(z) = (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k.
-\end{equation}
-
-The aforementioned assumptions are quite weak.  The strongest
-assumption we require is finally that $\Sigma_v$ is a diagonal matrix,
-which is referred to as a lack of instantaneous feedback in $x(t)$.
-We formally state our setup to conclude this section:
-
-\begin{definition}[Basic Setup]
-  \label{def:basic_setup}
-  The process $x(t)$ is an $n$ dimensional wide sense stationary
-  process having invertible $\VAR(\infty)$ representation
-  \eqref{eqn:ar_representation} where $v(t)$ is sequentially
-  uncorrelated and has a diagonal covariance matrix.  The $MA(\infty)$
-  representation of equation \eqref{eqn:wold} has $c(t) = 0$ and
-  $A(0) = I$.
+\begin{definition}[Granger-causality]
+  \label{def:granger_causality}
+  We will say that the $j$th component of $x$, namely $x_j(t)$,
+  Granger-causes component $x_i(t)$ (with respect to $x$) and write
+  $j \gc i$ if $\exists \tau \in \Z_{++}$ such that $B_{ij}(\tau) \ne 0$.
 \end{definition}
 
-\subsection{Granger Causality}
-
-\begin{definition}[Granger Causality]
-  \label{def:granger_causality}
-  For the WSS series $x(t)$ satisfying the assumptions of Definition
-  \ref{def:basic_setup} we will say that component $x_j$
-  \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x$)
-  and write $x_j \gc x_i$ if given Hilbert spaces $\H_{t - 1}$,
-  $\H^{-j}_{t - 1}$
-
-\begin{equation}
-  \linEerr{x_i(t)}{\H_{t - 1}} < \linEerr{x_i(t)}{\H^{-j}_{t - 1}},
-\end{equation}
-
-where $\xi[x \ |\ \H] = \E (x - \linE{x}{\H})^2$ is the mean squared
-estimation error and $\linE{x}{\H} = \text{proj}_{\H}(x)$ denotes the
-(unique) projection onto the Hilbert space $\H$.
-\end{definition}
-
-This notion captures the idea that the process $x_j$ provides
-information about $x_i$ that is not available from elsewhere.  The
-caveat ``with respect to $x$'' is important in that GC relations can
-change when components are added to or removed from our collection $x$
-of observations, e.g. new GC relations can arise if we remove the
-observations of a common cause, and existing GC relations can
-disappear if we observe a new mediating series.
-
-The notion is closely related to the information theoretic measure of
-transfer entropy, indeed, if the distribution of $v(t)$ is known to be
-Gaussian then they are equivalent \cite{barnett2009granger}.
-
-We require two technical facts we refer to in the sequel, the notion
-of conditional orthogonality is used throughout.
-
-\begin{lemma}[\cite{lindquist} Proposition 2.4.2]
-  \label{lem:conditional_orthogonality_equivalence}
-  Consider three closed subspaces of a Hilbert space $\mathcal{A}$,
-  $\mathcal{B}$, $\mathcal{X}$.  The following statements are
-  equivalent
-
-  \begin{enumerate}
-    \item{$\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$}
-    \item{$\linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}}\ \forall \beta \in \mathcal{B}$.}
-    % \item{$\linE{\beta}{\mathcal{A}} = \linE{\linE{\beta}{\mathcal{X}}}{\mathcal{A}}\ \forall \beta \in \mathcal{B}}$}
-    \end{enumerate}
-
-    Where $\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$ denotes
-    conditional orthogonality:
-
-    \begin{equation*}
-      \inner{a - \linE{a}{\mathcal{X}}}{b - \linE{b}{\mathcal{X}}} = 0\ \forall a \in \mathcal{A}, b \in \mathcal{B}.
-    \end{equation*}
-\end{lemma}
-% \begin{proof}
-%   Proceding from the definition of conditional orthogonality we have
-
-%   \begin{align*}
-%     &\inner{\alpha - \linE{\alpha}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
-%     &{\iff} \inner{\alpha + x - \linE{\alpha + x}{\mathcal{X}}}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
-%     &\overset{(a)}{\iff} \inner{\alpha + x}{\beta - \linE{\beta}{\mathcal{X}}} = 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}, x \in \mathcal{X}\\
-%     &\overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A} \vee \mathcal{X}} = 0\ \forall \beta \in \mathcal{B}\\
-%     &\overset{(c)}{\iff} \linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}},
-%   \end{align*}
-
-%   where $(a)$ follows since $(\beta - \linE{\beta}{\mathcal{X}}) \perp \mathcal{X}$ and $\linE{\alpha + x}{\mathcal{X}} \in \mathcal{X}$, $(b)$ since $z \perp \mathcal{X} \vee \mathcal{A} \iff \linE{z}{\mathcal{X} \vee \mathcal{A}} = 0$, and $(d)$ since $\linE{z}{\H_2} = \linE{\linE{z}{\H_2}}{\H_1}$ for $\H_2 \subseteq \H_1$.
-
-%   % This is property (vi) in Lindquist
-%   % \begin{align*}
-%   %   \langle \alpha - \linE{\alpha}{\mathcal{X}}, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0 \ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
-%   %   \overset{(a)}{\iff} \langle \alpha, \beta - \linE{\beta}{\mathcal{X}} \rangle &= 0\ \forall \alpha \in \mathcal{A}\ \beta \in \mathcal{B}\\
-%   %   \overset{(b)}{\iff} \linE{\beta - \linE{\beta}{\mathcal{X}}}{\mathcal{A}} &= 0\ \forall \beta \in \mathcal{B}
-%   % \end{align*}
-
-%   % where $(a)$ follows because $\linE{\alpha}{\mathcal{X}} \in \mathcal{X}$ and $\beta - \linE{\beta}{\mathcal{X}} \in \mathcal{X}$, and $(b)$ follows since $\beta \perp \mathcal{A} \iff \linE{\beta}{\mathcal{A}} = 0$.
-% \end{proof}
-\begin{lemma}
-  \label{lem:subspace_sum_projection}
-  Let $x \in \H$ where $\H$ is a separable Hilbert space having inner product
-  $\inner{\cdot}{\cdot}$ and let $\{\H_n\}_{n = 0}^N$ be a
-  collection of subspaces of $\H$ such that $x \perp \H_0$.  Then
-
-  \begin{equation*}
-    \linE{x}{\bigvee_{n = 0}^N\H_n} = \linE{x}{\bigvee_{n = 1}^N\H_n},
-  \end{equation*}
-
-  where
-  $\H_1 \vee \H_2 = \cl \{\alpha + \beta\ |\ \alpha \in \H_1, \beta
-  \in \H_2 \}$ is the closed sum of subspaces.
-\end{lemma}
-\begin{proof}
-  Apply the Gram-Schmidt process and directly calculate the projections.
-\end{proof}
-
-\begin{theorem}[Granger Causality Equivalences]
-  \label{thm:granger_causality_equivalences}
-  Let $x(t)$ be a WSS process with absolutely summable covariance
-  sequence, and spectral density uniformly bounded above and below.
-  Denote by $\xi_{ij} \defeq \linEerr{x_i(t)}{\H^{-j}}$ and
-  $\xi_i \defeq \linEerr{x_i(t)}{\H_t}$ for distinct $i, j$.  Then,
-  the following are equivalent:
-
-  \begin{enumerate}
-    \item{$x_j \ngc x_i$}
-    \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$ i.e. $\B_{ij}(z) = 0$}
-    \item{$F_{ij} \defeq \big(\frac{\xi_i}{\xi_{ij}} - 1\big) = 0$}
-    \item{$\H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j} \iff H_t^{i} \perp \H_{t - 1}^{j}\ |\ \H_{t - 1}^{-j}$}
-    \item{$\linE{x_i(t)}{\H_{t - 1}^{-j}} = \linE{x_i(t)}{\H_{t - 1}}$}
-    % \item{\hl{Wold $A(\tau)$ condition}}
-    % \item{Tie in the 0s of $S(\omega)^{-1}$, though this isn't immediately directed.}
-  \end{enumerate}
-\end{theorem}
-
-\begin{proof}
-  % Geweke uses the log of the ratio of the determinants of the residual variances
-  % 
-  % The equivalence $(1) \iff (3)$ is essentially a restatement of the
-  % definition, but is in line with the seminal work of Geweke
-  % \cite{geweke1982measurement}, \cite{geweke1984}.
-
-  $(1) \iff (3)$ is a tautology, and $(3) \iff (5)$ is immediate from the
-  definitions i.e. the projections are equal if and only if the errors
-  are equal.  We have $(4) \iff (5)$ as a result of Lemma
-  \ref{lem:conditional_orthogonality_equivalence}
-
-% , the additional
-%   equivalence in $(4)$ is apparent since $\H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$.
-
-  $(1) \Rightarrow (2)$: The projection for $x_i(t)$ onto $\H_{t - 1}$ is
-  (by definition) given by a form similar to equation
-  \ref{eqn:ar_representation}:
-
-\[
-  \E|x_i(t) - \linE{x_i(t)}{\H_{t - 1}}|^2 = \E|v_i(t) + \sum_{\tau = 1}^\infty \sum_{k = 1}^n (B_{i, k}(\tau) - \hat{B}_{i, k}(\tau))x_k(t - \tau)|^2.
-\]
-
-Since $v(t)$ in equation \ref{eqn:ar_representation} is temporally
-uncorrelated, it follows that the optimal projection is given by the
-model coefficients themselves.  This holds similarly for the
-projection onto $\H_{t - 1}^{-j}$.  Then by $(1)$ we have
-
-\[
-  \E|x_i(t) - \sum_{\tau = 1}^\infty \sum_{k = 1}^n B_{i, k}(\tau)x_k(t - \tau)|^2 = \E|x_i(t) - \sum_{\tau = 1}^\infty\sum_{k \ne j}B_{i, k}(\tau)x_k(t - \tau)|^2
-\]
-
-By the uniqueness of the projection we must have
-$\forall \tau\ B_{i, j}(\tau) = 0$.
-
-$(2) \Rightarrow (4)$: In computing $(y - \linE{y}{\H_{t - 1}^{-j}})$ for
-$y \in H_t^i$ it is sufficient to consider $y = x_i(t)$ by linearity, then since
-$H_{t - 1}^i \subseteq \H_{t - 1}^{-j}$ we have
-$(x_i(t) - \linE{x_i(t)}{\H_{t = 1}^{-j}}) = v_i(t)$ since $\B_{ij}(z) = 0$.  $(4)$ then follows
-since $v_i(t) \perp \H_{t - 1}$ and
-$\forall z \in \H_{t - 1}^j\ (z - \linE{z}{\H_{t - 1}^{-j}}) \in \H_{t
-  - 1}$
-\end{proof}
-
-% The following propositions justify various modifications of $x(t)$
-% applied in practice to massage $x(t)$ into a form amenable to more
-% standard tools.
-
-% \begin{theorem}[General Invariance]
-%   \hl{Under what conditions is this actually true?}
-
-%   Let $\zeta(t)$ be a stationary discrete time stochastic process.  Let
-%   $F, G$ be (possibly nonlinear and time varying) invertible filtering
-%   operations and $f(t)$, $g(t)$ be perfectly predictable and such that
-%   $x_j(t) \defeq F(\zeta_j - f)(t), x_i(t) \defeq G(\zeta_i - g)(t)$ are W.S.S.  Then,
-
-% \begin{equation}
-%     x_j \gc x_i \iff \zeta_j \te \zeta_i
-%   \end{equation}
-% \end{theorem}
-
-% \begin{proposition}[Invariance Under Invertible and Deterministic Modifications]
-%   Let $x(t)$ be a WSS process with absolutely summable covariance
-%   sequence, and spectral density uniformly bounded above and below.  Let
-%   $F(z), G$ represent univariate and invertible
-%   linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
-%   predictable processes.  Then,
-
-%   \begin{equation}
-%     x_j \gc x_i \iff F(x_j - f)(t) \gc G(x_i - g)(t)
-%   \end{equation}
-
-% \begin{proof}
-%   \hl{TODO}
-% \end{proof}
-% \end{proposition}
-
-\subsection{Granger Causality Graphs}
-We first need to establish some graph theoretic notation and
-terminology, collected formally in definitions for the reader's
-convenient reference.
-
-\begin{definition}[Graph Theory Review]
-  A \textit{graph} $\gcg = (V, \gcge)$ is simply a
-  tuple of sets respectively called \textit{nodes} and \textit{edges}.
-  Throughout this paper, we have in all cases
-  $V = [n] \defeq \{1, 2, \ldots, n\}$.  We will also focus solely on
-  \textit{directed} graphs, where the edges
-  $\gcge \subseteq V \times V$ are \textit{ordered} pairs.
-
-  A (directed) \textit{path} (of length $r$) from node $i$ to node
-  $j$, denoted $\gcgpath{i}{j}$, is a sequence
-  $a_0, a_1, \ldots, a_{r - 1}, a_r$ with $a_0 = i$ and $a_r = j$ such
-  that $\forall\ 0 \le k \le r\ (a_k, a_{k + 1}) \in \gcge$, and where
-  $(a_k, a_{k - 1})$ are \textit{distinct} for $0 \le k < r$.
-
-  A \textit{cycle} is a path of length $2$ or more between a node and
-  itself.  An edge between a node and itself $(i, i) \in \gcge$ (which
-  is not a cycle) is referred to as a \textit{loop}.
-
-  A graph $\gcg$ is a \textit{directed acyclic graph} (DAG) if it is a
-  directed graph and does not contain any cycles.
-\end{definition}
-
-\begin{definition}[Parents, Grandparents, Ancestors]
-  A node $j$ is a \textit{parent} of node $i$ if $(j, i) \in \gcge$.
-  The set of all $i$'s parents will be denoted $\pa{i}$, and we
-  explicitly exclude loops as a special case, that is,
-  $i \not\in \pa{i}$ even if $(i, i) \in \gcge$.
-
-  The set of level $\ell$ \textit{grandparents} of node $i$, denoted
-  $\gpn{\ell}{i}$, is the set such that $j \in \gpn{\ell}{i}$ if and
-  only if there is a \textit{directed path} of length $\ell$ in $\gcg$
-  from $j$ to $i$.  Clearly, $\pa{i} = \gpn{1}{i}$.
-
-  Finally, the set of \textit{level $\ell$ ancestors} of $i$:
-  $\ancn{\ell}{i} = \bigcup_{\lambda \le \ell}\gpn{\lambda}{i}$ is the
-  set such that $j \in \ancn{\ell}{i}$ if and only if there is a
-  directed path of length $\ell$ \textit{or less} in $\gcg$ from $j$
-  to $i$.  The set of \textit{all ancestors} of $i$
-  (i.e. $\ancn{n}{i}$) is denoted simply $\anc{i}$.
-
-  Recall that we do not allow a node to be it's own parent, however
-  unless $\gcg$ is a DAG, a node can be it's own ancestor.  We will
-  ocassionally need to explicitly exclude $i$ from $\anc{i}$, in which
-  case we will write $\anc{i}\setminus \{i\}$.
-\end{definition}
-
-Our principle object of study will be a graph determined by
-Granger-causality relations as follows.
-
-\begin{definition}[Causality graph]
-  We define the Granger-causality graph $\gcg = ([n], \gcge)$ to be the directed
-  graph formed on $n$ vertices where an edge $(j, i) \in \gcge$ if and
-  only if $x_j$ Granger-causes $x_i$ (with respect to $x$).  That is,
+This definition captures the notion that $x_j$ provides information
+about $x_i$ which is not available through any of the other components
+of $x$.  The motivation being that, since an effect must follow it's
+cause, that $x_j$ ``causes'' $x_i$.  This intuition can also be given
+interpretation as the ``flow'' of ``information'' or ``energy''.
+
+\begin{definition}[Granger-causality Graph]
+  We define the Granger-causality graph $\gcg = ([n], \gcge)$ to be
+  the directed graph formed on $n$ vertices where an edge
+  $(j, i) \in \gcge$ if and only if $x_j$ Granger-causes $x_i$ (with
+  respect to $x$).  That is,
   $$(j, i) \in \gcge \iff j \in \pa{i} \iff x_j \gc x_i.$$
@@
 
-The edges of the Granger-causality graph $\gcg$ can be given a general
-notion of ``weight'' by associating an edge $(j, i)$ with the
-\textit{strictly causal} LTI filter $\B_{ij}(z)$ (see eqn
-\eqref{eqn:filter_action}).  Thence, the matrix $\B(z)$ is analogous
-to a \textit{weighted adjacency matrix}\footnote{We are using the
-  convention that $\B_{ij}(z)$ is a filter with input $x_j$ and output
-  $x_i$ so as to write the action of the system as $\B(z)x(t)$ with
-  $x(t)$ as a column vector.  This competes with the usual convention
-  for adjacency matrices where $A_{ij} = 1$ if there is an edge
-  $(i, j)$.  In our case, the sparsity pattern of $\B_{ij}$ is the
-  \textit{transposed} conventional adjacency matrix.} for the graph $\gcg$.  And,
-in the same way that the $k^{\text{th}}$ power of an adjacency matrix
-counts the number of paths of length $k$ between nodes,
-$(\B(z)^k)_{ij}$ is a filter isolating the ``action'' of $j$ on $i$ at
-a time lag of $k$ steps, this is exemplified in the inversion formula
-\ref{eqn:lsi_inversion}.
-
-An elementary theorem connecting the adjacency matrix with paths will
-allow us to deduce the sparsity pattern of $\A(z)$.  Proof follows
-easily by induction:
-
-\begin{lemma}
-  \label{lem:adj_matrix}
-  Let $S$ be the transposed adjacency matrix of the Granger-causality
-  graph $\gcg$.  Then, $(S^k)_{ij}$ is the number of paths of length
-  $k$ from node $j$ to node $i$.  Evidently, if
-  $\forall k \in \N,\ (S^k)_{ij} = 0$ then $j \not\in \anc{i}$.
-\end{lemma}
-
-% What if we take this as the definition of PWGC?  then, in a simply causal graph
-% the "normal" definition becomes a theorem.  Or, if lem:pwgc_anc is true then
-% maybe that is good enough?
-
-From the $\VAR$ representation of $x(t)$ there is clearly a tight
-relationship between each node and it's parent nodes, the relationship
-is quantified through the sparsity pattern of $B(z)$.  Similarly, the
-following proposition is analogous to the definition of feedback free
-processes of \cite{caines1975feedback} and provides an interpretation
-of the sparsity pattern of $A(z)$ (from the MA representation of
-$x(t)$) in terms of the causality graph $\gcg$.
-
-% % When is this guaranteed w.r.t. the full system?
-% Furthermore, if the filter $\B_{ii}$ is invertible, we can write
-
-% \begin{equation*}
-%   \begin{aligned}
-%     (1 - \B_{ii}(z))x_i(t) &= v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\\
-%     \Rightarrow x_i(t) &= (1 - \B_{ii}(z))^{-1}\big[v_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t)\big]\\
-%     &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)
-%   \end{aligned}
-% \end{equation*}
-
-% In general, we will write $\wtB$ for arbitrary filters maintaining the
-% convention that $\wtB_{ii}$ is a filter that modifies the future of
-% $x_i$ from it's past and $\wtB_{ij}$ a filter modifying $x_i$ from the
-% past of $x_j$.  We will use the convention that the filter
-% $\widetilde{\B}$ will represent simply ``a filter'' and is not
-% necessarily the same as the filters serving as edge weights in $\gcg$
-% (indeed, $\wtB$ will often be $0$), and moreover, we will allow for
-% $\widetilde{\B}$ to potentially change from place to place and even
-% from line to line in the same sequence of calculations as it's exact
-% properties or specification are not important.
-
-% We continue to develop the relationship between $x_i$ and it's family
-% tree, where the following proposition is analagous to the definitions
-% of feedback free processes of \cite{caines1975feedback}.
-
-\begin{proposition}[Ancestor Expansion]
-  \label{prop:parent_expanding}
-  The component $x_i(t)$ of $x(t)$ can be represented in terms of it's
-  parents in $\gcg$:
-
-  \begin{equation}
-    \label{eqn:parent_expansion}
-    x_i(t) = v_i(t) + \B_{ii}(z)x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t).
-  \end{equation}
-
-  Moreover, $x_i$ can be expanded in terms of it's ancestor's $v(t)$
-  components only:
-
-  \begin{equation}
-    \label{eqn:ancestor_expansion}
-    x_i(t) = \A_{ii}(z)v_i(t) + \sum_{\substack{k \in \anc{i} \\ k \ne i}}\A_{ik}(z)v_k(t),
-  \end{equation}
-
-  where $\A(z) = \sum_{\tau = 0}^\infty A(\tau)z^{-\tau}$ is the filter from
-  the Wold decomposition representation of $x(t)$, equation
-  (\ref{eqn:wold}).
-\end{proposition}
-
-This statement is ultimately about the sparsity pattern in the Wold
-decomposition matrices $A(\tau)$ since
-$x_i(t) = \sum_{\tau = 0}^\infty \sum_{j = 1}^n A_{ij}(\tau)v_j(t -
-\tau)$.  The proposition states that if $j \not \in \anc{i}$ then
-$\A_{ij}(z) = 0$.
-
-% TODO: ???
-% , however the implication $j \in \anc{i} \implies \A_{ij}(z) \ne 0$ need not hold.
-
-% \begin{proof}
-%   The formula (\ref{eqn:parent_expansion}) was established in the
-%   preceding paragraph and serves as a starting point for the induction
-%   hypothesis:
-
-%   \begin{equation}
-%     \label{eqn:induction}
-%     x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
-%   \end{equation}
-
-%   The base case follows by expanding each $x_k$ in equation \ref{eqn:induction} via equation \ref{eqn:parent_expansion}, but we skip the calculations since they are nearly identical to the following induction step:
-
-%   \begin{align*}
-%     x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)\\
-%            &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell - 1}{i}\setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)\big[\wtB_{kk}v_k(t)\\
-%     &\ \ \ \  + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t)\big]\\
-%     &\overset{(b)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \ancn{\ell}{i} \setminus \{i\}}\wtB_{ik}(z)v_k(t) + \sum_{k \in \gpn{\ell + 1}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t).
-%   \end{align*}
-
-%   Where equality $(a)$ follows by expanding each $x_k$ with equation \ref{eqn:parent_expansion} and $(b)$ requires the inversion of $\big(1 - \sum_{k \in \gpn{\ell}{i}\setminus\{i\}}\wtB_{ik}(z)\wtB_{ki}(z)\big)$ as well as using the fact that $\ancn{\ell}{i} = \ancn{\ell - 1}{i}\cup\gpn{\ell}{i}$.
-We here remind the reader of our earlier established convention that $\wtB$ simply represents the existence of a filter (possibly 0) and can change from place to place.
-
-% The induction terminates after $n$ steps since in a graph with $n$ nodes $\gpn{n + 1}{i} = \emptyset$
-
-% % Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.
-
-% %   \begin{align*}
-% %     x_i(t) &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t)\\
-% %            &= \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)\big[\wtB_{kk}v_{k}(t) + \wtB_{ki}(z)x_i(t) + \sum_{h \in \pa{k}\setminus \{i\}}\wtB_{kh}(z)x_h(t) \big]\\
-% %            &\overset{(a)}{=} \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)v_{k}(t) + \sum_{k \in \gpn{2}{i} \setminus \{i\}}\wtB_{ik}(z)x_k(t)
-% %   \end{align*}
-
-% % Where step $(a)$ requires the inversion of $1 - \sum_{k \in \pa{i}}\wtB_{ik}(z)\wtB_{ki}(z)$.
-
-% \end{proof}
-% \textbf{this is an alternative proof, I think superior to the one above}.
-
-\begin{proof}
-  Equation \eqref{eqn:parent_expansion} is immediate from the
-  $\VAR(\infty)$ representation of \eqref{eqn:ar_representation} and
-  Theorem \ref{thm:granger_causality_equivalences}, we are left to
-  demonstrate \eqref{eqn:ancestor_expansion}.
-
-  From equation (\ref{eqn:ar_representation}), which we are assuming
-  throughout the paper to be invertible, we can write
-
-  \begin{equation*}
-    x(t) = (I - \B(z))^{-1} v(t),
-  \end{equation*}
-
-  where $(I - \B(z))^{-1} = \A(z)$ due to the uniqueness of
-  (\ref{eqn:wold}).  Since $\B(z)$ is stable we have
-
-  \begin{equation}
-    \label{eqn:resolvant_inv}
-    (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k.
-  \end{equation}
-
-  Invoking the Cayley-Hamilton theorem allows writing the infinite sum
-  of \eqref{eqn:resolvant_inv} in terms of \textit{finite} powers of
-  $\B$.
-
-  Let $S$ be a matrix with elements in $\{0, 1\}$ which represents the
-  sparsity pattern of $\B(z)$, from lemma \ref{lem:adj_matrix} $S$ is
-  the transpose of the adjacency matrix for $\gcg$ and hence
-  $(S^k)_{ij}$ is non-zero if and only if $j \in \gpn{k}{i}$, and
-  therefore $\B(z)^k_{ij} = 0$ if $j \not \in \gpn{k}{i}$.  Finally,
-  since $\anc{i} = \bigcup_{k = 1}^n\gpn{k}{i}$ we see that
-  $\A_{ij}(z)$ is zero if $j \not\in \anc{i}$.
-
-  Thence
-
-  \begin{align*}
-    x_i(t) &= [(I - \B(z))^{-1}v(t)]_i\\
-    &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
-    &= \A_{ii}(z) v_i(t) + \sum_{\substack{j \in \anc{i} \\ j \ne i}} \A_{ij}(z) v_j(t)
-  \end{align*}
-\end{proof}
-
-\subsection{Pairwise Granger Causality}
-\label{sec:pwgc}
-Recall that Granger-causality in general must be understood with
+Referring back to the definition (Def. \ref{def:granger_causality}) of
+Granger-causality it is important that the concept be understood with
 respect to a particular universe of observations.  If $x_j \gc x_i$
 with respect to $x_{-k}$, it may not hold with respect to $x$.  For
-example, $x_k$ may be a common ancestor which when observed, completely
-explains the connection from $x_j$ to $x_i$.  In this section we study
-\textit{pairwise} Granger-causality, and seek to understand when
-knowledge of pairwise relations is sufficient to deduce the true fully
-conditional relations of $\gcg$.
+example, $x_k$ may be a common ancestor which when observed,
+completely explains the connection from $x_j$ to $x_i$.  We introduce
+the pairwise definition of Granger-causality as a special case:
 
@@
 
-\begin{lemma}
-  \label{lem:separated_ancestor_uncorrelated}
-  Consider distinct nodes $i, j$ in a Granger-causality graph
-  $\gcg$.  If
-
-  \begin{enumerate}[label=(\alph*)]
-    \item{$j \not\in \anc{i}$ and $i \not\in \anc{j}$}
-    \item{$\anc{i}\cap\anc{j} = \emptyset$}
-  \end{enumerate}
-
-  then $\H_t^{(i)} \perp \H_t^{(j)}$, that is,
-  $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.  Moreover,
-  this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_{t - 1}^i} = 0$.
-\end{lemma}
-
-\begin{remark}
-  The possibility that there exist nodes sharing both $i$ and $j$ as
-  ancestors is not excluded, the point being that the temporal nature
-  of Granger-causality eliminates the problems caused by ``colliders''
-  (in the language of Pearl's causal calculus
-  \cite{pearl2009causality}) in static causal inference.
-\end{remark}
-
-\begin{proof}
-  We show directly that
-  $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.  To this end,
-  fix $s, \tau \ge 0$, then by expanding with equation \eqref{eqn:ancestor_expansion} we have
-
-  \begin{align*}
-    \E x_i(t - s)x_j(t - \tau)
-    &= \E \big(\A_{ii}(z)v_i(t - s)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)\\
-    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\E[\big(\A_{ik}(z)v_k(t - s)\big)\big(\A_{jj}(z)v_j(t - \tau)\big)]\\
-    &+ \sum_{\substack{k \in \anc{j} \\ k \ne j}}\E[\big(\A_{ii}(z)v_i(t - s)\big) \big(\A_{jk}(z) v_k(t - \tau)\big)]\\
-    &+ \sum_{\substack{k \in \anc{i} \\ k \ne i}}\sum_{\substack{\ell \in \anc{j} \\ \ell \ne j}}\E[\big(\A_{ik}(z)v_k(t - s)\big)\big(\A_{j\ell}(z)v_\ell(t - \tau)\big)].
-  \end{align*}
-
-  Keeping in mind that $v(t)$ is an isotropic and uncorrelated
-  sequence we see that each of these above four terms are 0: the
-  first term since $i \ne j$, the second and third since
-  $j \not\in \anc{i}$ and $i \not\in \anc{j}$ and finally the fourth since
-  $\anc{i} \cap \anc{j} = \emptyset$.
-\end{proof}
-
-It is possible for components of $x(t)$ to be correlated at some time
-lags without resulting in pairwise causality:
-
-\begin{lemma}
-  \label{lem:ancestor_uncorrelated}
-  Consider distinct nodes $i, j$ in a Granger-causality graph $\gcg$.
-  If
-
-  \begin{enumerate}[label=(\alph*)]
-    \item{$j \not\in \anc{i}$}
-    \item{$\anc{i}\cap\anc{j} = \emptyset$}
-  \end{enumerate}
-
-  then $j \npwgc i$.
-\end{lemma}
-\begin{proof}
-  By Theorem \ref{thm:granger_causality_equivalences} it suffices to show that
-
-  \begin{equation*}
-    \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^i}} = 0,
-  \end{equation*}
-
-  which by the orthogonality principle is equivalent to
-
-  \begin{equation}
-    \label{eqn:proof_inner0}
-    \inner{x_i(t)}{x_j(t) - \linE{x_j(t)}{\H_{t - 1}^i}} = 0.
-  \end{equation}
-
-  Define the disjoint sets
-
-  \begin{align*}
-    C_0(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, k \ne i \}\\
-    C_1(u) &= \{k \in \pa{u}\ |\ i \in \anc{k} \text{ or } k = i \}.
-  \end{align*}
-
-  We can then expand the parents of $j$ using Equation \ref{eqn:parent_expansion} as
-
-  \begin{equation*}
-    x_j(t) = v_j(t) + \sum_{k \in C_0(j)} B_{jk}(z) x_k(t) + \sum_{k \in C_1(j)} B_{jk}(z) x_k(t),
-  \end{equation*}
-
-  which when substituted into the left hand side of Equation
-  \ref{eqn:proof_inner0} results (by Lemma
-  \ref{lem:separated_ancestor_uncorrelated}) in
-
-  \begin{equation*}
-    \inner{x_i(t)}{\sum_{k \in C_1(j)} B_{jk}(z) x_k(t) - \sum_{k \in C_1(j)} \linE{B_{jk}(z) x_k(t)}{\H_{t - 1}^i}}.
-  \end{equation*}
-
-  We can continue this process recursively (i.e. split each
-  $k \in C_1(j)$ into $C_0(k)$ and $C_1(k)$) which must eventually
-  terminate with $C_1(u) = \{i\}$.  Therefore, there exists some
-  causal filter $\Phi(z)$ such that Equation \ref{eqn:proof_inner0} is
-  equivalent to
-
-  \begin{equation}
-    \label{eqn:proof_inner0}
-    \inner{x_i(t)}{\Phi(z)x_i(t) - \linE{\Phi(z)x_i(t)}{\H_{t - 1}^i}},
-  \end{equation}
-
-  which is $0$ since $\linE{\Phi(z)x_i(t)}{\H_{t - 1}^i} = \Phi(z)x_i(t)$.
-\end{proof}
-
-The previous proposition can be strengthened significantly; notice
-that it is possible to have some $k \in \anc{i} \cap \anc{j}$ where still
-$j \npwgc i$, an example is furnished by the three node graph
-$k \rightarrow i \rightarrow j$ where clearly
-$k \in \anc{i}\cap\anc{j}$ but $j \npwgc i$.  We must introduce the concept
-of a \textit{confounding} variable, which effectively eliminates the
-possibility presented in this example.
-
-\begin{definition}[Confounder]
-  A node $k$ will be referred to as a \textit{confounder} of nodes
-  $i, j$ (neither of which are equal to $k$) if
-  $k \in \anc{i} \cap \anc{j}$ and there exists a path
-  $\gcgpath{k}{i}$ not containing $j$, and a path $\gcgpath{k}{j}$
-  not containing $i$.
-
-  A simple example is furnished by the ``fork'' graph
-  $i \leftarrow k \rightarrow j$.
-\end{definition}
-
-\begin{proposition}
-  \label{prop:ancestor_properties}
-  If in a Granger-causality graph $\gcg$ where $j \pwgc i$ then
-  $j \in \anc{i}$ or $\exists k \in \anc{i} \cap\anc{j}$ which is a
-  confounder of $(i, j)$.
-\end{proposition}
-
-\begin{remark}
-  The interpretation of this proposition is that for $j \pwgc i$ then
-  there must either be ``causal flow'' from $j$ to $i$
-  ($j \in \anc{i}$) or there must be a confounder $k$ through which
-  common information is received.
-\end{remark}
-
-\begin{proof}
-  We will prove by way of contradiction.  To this end, suppose that
-  $j$ is a node such that: $(a)$ $j \not \in \anc{i}$ and $(b)$ for
-  every $k \in \anc{i} \cap \anc{j}$ every
-  $k \rightarrow \cdots \rightarrow j$ path contains $i$.
-
-  Firstly, notice that every $u \in \big(\pa{j} \setminus \{i\}\big)$
-  necessarily inherits these same two properties.  This follows since
-  if we also had $u \in \anc{i}$ then $u \in \anc{i} \cap \anc{j}$ so that every
-  $u \rightarrow \cdots \rightarrow j$ path must contain $i$, but
-  $u \in \pa{j}$, so this is not the case since $u \rightarrow j$ is a
-  path that doesn't contain $i$; moreover, if we consider
-  $w \in \anc{i} \cap \anc{u}$ then we also have
-  $w \in \anc{i} \cap \anc{j}$ so every
-  $w \rightarrow \cdots \rightarrow j$ path must contain $i$.  These
-  properties therefore extend inductively to every
-  $u \in \big(\anc{j} \setminus \{i\}\big)$.
-
-  In order to deploy a recursive argument, define the following
-  partition of $\pa{u}$, for some node $u$:
-
-  \begin{align*}
-    C_0(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, \anc{i} \cap \anc{k} = \emptyset, k \ne i\}\\
-    C_1(u) &= \{k \in \pa{u}\ |\ i \in \anc{k} \text{ or } k = i\}\\
-    C_2(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, \anc{i} \cap \anc{k} \ne \emptyset, k \ne i\}.
-  \end{align*}
-
-  We notice that for any $u$ having the properties $(a), (b)$ above, we
-  must have $C_2(u) = \emptyset$ since if $k \in C_2(u)$ then
-  $\exists w \in \anc{i} \cap \anc{k}$ s.t. $i \not \in \anc{k}$ and
-  therefore there must be a path $\gcgpath{w}{k} \rightarrow u$ which
-  does not contain $i$.
-
-  Using this partition, we will expand $x_j(t)$ in terms of it's
-  parents, and recursively expand nodes in $C_1$ until we reach a case
-  where $C_1 = \emptyset$.  For the first step equation
-  \eqref{eqn:parent_expansion} gives us:
-
-  \begin{equation}
-    \label{eqn:xj_partition_expansion}
-    x_j(t) = \A_{jj}(z)\Big(v_j(t) + \sum_{k \in C_0(j)}\B_{ik}(z)x_k(t) + \sum_{k \in C_1(j)}\B_{ik}(z)x_k(t)\Big).
-  \end{equation}
-
-  Using this representation we choose an arbitrary $\Phi(z) x_i(t - 1) \in \H_{t - 1}^{(j)}$ and show that
-
-  \begin{equation}
-    \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}{\Phi(z)x_j(t - 1) - \linE{\Phi(z)x_j(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
-  \end{equation}
-
-  which will imply (by Theorem \ref{thm:granger_causality_equivalences}) that $j \npwgc i$ and for which it is equivalent to show that
-
-  \begin{equation}
-    \label{eqn:sufficient_inner_prod}
-    \inner{x_i(t)}{\Phi(z)x_j(t - 1) - \linE{\Phi(z)x_j(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
-  \end{equation}
-
-  by the orthogonality principle since $\linE{x_i(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$.
-Substituting (\ref{eqn:xj_partition_expansion}) into (\ref{eqn:sufficient_inner_prod}) and starting with the first term we have
-
-  \begin{align*}
-    &\inner{x_i(t)}{\Phi(z)\A_{jj}(z)v_j(t - 1) - \linE{\Phi(z)\A_{jj}(z)v_j(t - 1)}{\H_{t - 1}^{(i)}}}\\
-    \overset{(\alpha)}{=}\ &\inner{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}{\Phi(z)\A_{jj}(z)v_j(t - 1)}\\
-    \overset{(\beta)}{=}\ &0,
-  \end{align*}
-
-  where $(\alpha)$ follows by expanding $x_i(t)$ with (\ref{eqn:ancestor_expansion}) and $\linE{\Phi(z)\A_{jj}(z)v_j(t - 1)}{\H_{t - 1}^{(i)}} = 0$ because $\forall \tau, s$
-
-  \begin{equation*}
-    \E v_j(t - \tau) x_i(t - s) = \E v_j(t - \tau) \sum_{k \in \anc{i} \cup \{i\}}\A_{ik}(z)v_k(t - s) = 0,
-  \end{equation*}
-
-  since $j \not \in \anc{i}$; $(\beta)$ follows similarly, that is, $j \not \in \anc{i}$.  Secondly we see that $\forall k \in C_0(j)$
-
-  \begin{equation*}
-    \inner{x_i(t)}{\Phi(z)\A_{jj}(z)\B_{ik}(z)x_k(t - 1) - \linE{\Phi(z)\A_{jj}(z)\B_{ik}(z)x_k(t - 1)}{\H_{t - 1}^{(i)}}} = 0,
-  \end{equation*}
-
-  which follows from Lemma \ref{lem:ancestor_uncorrelated}.  Finally,
-  for $k \in C_1(j)$ the case $k = i$ is immediate (since the error in
-  estimating $\Phi(z)\B_{ii}(z)x_i(t)$ given $\H_{t - 1}^{(i)}$ is
-  $0$), so suppose $k \ne i$.  We know from above that $k$ inherits the
-  key properties referred to as $(a)$ and $(b)$ above and therefore we
-  can recursively expand $k$ in the same way as in equation
-  (\ref{eqn:xj_partition_expansion}).  Continuing this recursion for
-  each $k \in C_1(j)$ (where $k \ne i$) must eventually terminate since
-  $i \in \anc{k}$.
-\end{proof}
-
-An interesting corollary is the following:
-
-\begin{corollary}
-  If the graph $\gcg$ is a DAG then $j \pwgc i, i \pwgc j \implies \exists k \in \anc{i} \cap \anc{j}$ confounding $(i, j)$.
-\end{corollary}
-
-It seems reasonable to expect a converse of proposition
-\ref{prop:ancestor_properties} to hold, i.e.
-$j \in \anc{i} \Rightarrow j \pwgc i$.  Unfortunately, this is not the
-case in general, as different paths through $\gcg$ can lead to
-cancellation (see example \ref{ex:diamond_cancellation}).  In fact, we
-do not even have $j \in \pa{i} \Rightarrow j \pwgc i$ (see example
-\ref{ex:lag_cancellation}).
-
-\begin{example}
-  \label{ex:diamond_cancellation}
-  Firstly, on $n = 4$ nodes, ``diamond'' shapes can lead to cancellation on paths of length 2:
-
-\begin{equation*}
-  x(t) =
-  \left[
-    \begin{array}{cccc}
-      0 & 0 & 0 & 0\\
-      a & 0 & 0 & 0\\
-      -a & 0 & 0 & 0\\
-      0 & 1 & 1 & 0\\
-    \end{array}
-  \right] x(t - 1) + v(t),
-\end{equation*}
-
-with $\E v(t) = 0,\ \E v(t)v(t - \tau)^\T = \delta_\tau I$.
-
-By directly calculating
-
-\begin{align*}
-  x_4(t) &= x_2(t - 1) + x_3(t - 1) + v_4(t)\\
-         &= ax_1(t - 2) + av_2(t - 1) - ax_1(t - 2) -av_3(t - 1) + v_4(t)\\
-         &= a(v_2(t - 1) - v_3(t - 1)) + v_4(t),
-\end{align*}
-
-we see that, since $v(t)$ is isotropic white noise, $1 \npwgc 4$.  The problem here is that there are multiple paths from $x_1$ to $x_4$.
-\end{example}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[b]{0.45\textwidth}
-    \includegraphics[width=\linewidth]{example1.pdf}
-    \caption{Graph Corresponding to Example \ref{ex:diamond_cancellation}}
-    \label{fig:diamond_cancellation}
-    \centering{$j \in \anc{i} \nRightarrow j \pwgc i$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \includegraphics[width=\linewidth]{example2.pdf}
-    \caption{Graph Corresponding to Example \ref{ex:lag_cancellation}}
-    \label{fig:lag_cancellation}
-    \centering{$j \in \pa{i} \nRightarrow j \pwgc i$}
-  \end{subfigure}
-\end{figure}
-
-\begin{example}
-  \label{ex:lag_cancellation}
-  A second example on $n = 3$ nodes is also worth examining, in this case
-  cancellation is a result of differing time lags.
-
-\begin{equation*}
-  x(t) =
-  \left[
-    \begin{array}{ccc}
-      0 & 0 & 0\\
-      -a & 0 & 0\\
-      0 & 1 & 0\\
-    \end{array}
-  \right] x(t - 1) +
-  \left[
-    \begin{array}{ccc}
-      0 & 0 & 0\\
-      0 & 0 & 0\\
-      a & 0 & 0\\
-    \end{array}
-  \right] x(t - 2) + v(t)
-\end{equation*}
-
-Then
-
-\begin{align*}
-  x_2(t) &= v_2(t) - ax_1(t - 1)\\
-  x_3(t) &= v_3(t) + x_2(t - 1) + ax_1(t - 2)\\
-  \Rightarrow x_3(t) &= v_2(t - 1) + v_3(t),
-\end{align*}
-
-and again $1 \npwgc 3$.
-\end{example}
-
-\subsection{Strongly Causal Graphs}
+\subsection{Strongly Causal Graphs and Persistent Systems}
 \label{sec:strongly_causal_graphs}
-% The examples at the conclusion of the previous section seem rather
-% pathological since we have simply constructed a case where different
-% paths in the graph cancel exactly.  It may be possible to rid
-% ourselves of these pathologies for instance by defining a notion of
-% ``robust'' pairwise causality where ``$j \pwgc _R i$'' if $j \pwgc i$
-% almost surely for a ``small'' random perturbation of the system
-% matrix.  However, we do not believe such a modification of our theory
-% can have useful practical implications since the difference between
-% paths that ``nearly cancel'' rather than cancel exactly would in
-% practice require inordinate amounts of data to resolve.
-
-In this section and the next we will seek to understand when converse
-statements of Proposition \ref{prop:ancestor_properties} \textit{do}
-hold.  One possibility is to restrict the coefficients of the system
-matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead,
-we think it more meaningful to focus on the defining feature of
-time series networks, that is, the topology of $\gcg$.
-
-\begin{definition}[Strongly Causal]
+In this section we define the notion of a strongly causal graph, and
+the implications for pairwise causality testing.
+
+\begin{definition}[Strongly Causal \cite{my_GC_paper}]
   \label{def:strongly_causal}
   We will say that a Granger-causality graph $\gcg$ is
-  \textit{strongly causal} if there is at most 1 directed path between
-  any two nodes.  Strongly Causal Graphs will be referred to as SCGs.
+  \textit{strongly causal} if it is acyclic and there is at most 1
+  directed path between any two nodes.  Strongly Causal Graphs will be
+  referred to as SCGs.
 \end{definition}
@@
   Examples of strongly causal graphs include directed trees (or
-forests), DAGs where each node has at most one parent, and figure
-\ref{fig:example_fig3} of this paper.  A complete bipartite graph with
-$2n$ nodes is also strongly causal, demonstrating that the number of
-edges of such a graph can still scale quadratically with the number of
-nodes.  It is evident that the strong causal property is inherited by
-subgraphs.
-
-\begin{example}
-  Though examples of SCGs are easy to construct in theory, should
-  practitioners expect SCGs to arise in application?  While a positive
-  answer to this question is not \textit{necessary} for the concept to
-  be useful, it is certainly sufficient.  Though the answer is likely
-  to depend upon the particular application area, examples appear to
-  be available in biology, in particular, the authors of
+  forests), DAGs where each node has at most one parent, and complete
+  bipartite graphs.  Examples of graphs arising in practice that appear
+  to be ``close'' to strongly causal (i.e. few edges violate the condition)
+  appear to be available in biology, in particular, the authors of
   \cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty}
@@
   \end{figure}
-\end{example}
-
-% \begin{lemma}
-%   \label{lem:still_strongly_causal}
-%   If $\gcg$ is a strongly causal graph, any subgraph formed by eliminating nodes
-%   as well as all of the in and out edges thereof is still strongly causal.
-% \end{lemma}
-% \begin{proof}
-%   If $\gcg$ has at most one path between any two nodes, there can only
-%   be fewer paths after removing nodes from $\gcg$.
-% \end{proof}
-
-For later use, and to get a feel for the topological implications of
-strong causality, we explore a number of properties of such graphs
-before moving into the main result of this section.  The following
-important property essentially strengthens proposition
-\ref{prop:ancestor_properties} for the case of strongly causal graphs.
-
-\begin{proposition}
-  \label{prop:sc_graph_common_anc}
-  In a strongly causal graph if $j \in \anc{i}$ then any
-  $k \in \anc{i} \cap \anc{j}$ is not a confounder, that is,
-  the unique path from $k$ to $i$ contains $j$.
-\end{proposition}
-\begin{proof}
-  Suppose that there is a path from $k$ to $i$ which does not contain
-  $j$.  In this case, there are multiple paths from $k$ to $i$ (one of
-  which \textit{does} go through $j$ since $j \in \anc{i}$) which
-  contradicts the assumption of strong causality.
-\end{proof}
-
-% Is pwgc transitive?
-
-\begin{corollary}
-  \label{cor:parent_corollary}
-  If $\gcg$ is a strongly causal DAG then $i \pwgc j$ and $j \in \anc{i}$ are
-  \textit{alternatives}, that is $i \pwgc j \Rightarrow j \notin \anc{i}$.
-\end{corollary}
-\begin{proof}
-  Suppose that $i \pwgc j$ and $j \in \anc{i}$.  Then since $\gcg$ is
-  acyclic $i \not\in \anc{j}$, and by proposition
-  \ref{prop:ancestor_properties} there is some
-  $k \in \anc{i}\cap\anc{j}$ which is a confounder.  However, by
-  proposition \ref{prop:sc_graph_common_anc} $k$ cannot be a
-  confounder, a contradiction.
-\end{proof}
-% This is a direct proof
-  % \begin{proof}
-%   Suppose that $\gcg$ is a strongly causal DAG and that we have both
-%   $i \pwgc j$ and $j \in \anc{i}$, which implies that
-%   $i \not \in \anc{j}$ since $\gcg$ is a DAG.  We will establish the
-%   contradiction $i \in \anc{j}$.
-
-%   Since $j \in \anc{i}$ there is a path $\gcgpath{j}{i}$.
-%   Moreover, since $i \pwgc j$ by proposition \ref{prop:pwgc_anc} there
-%   must be a confounding ancestor $k \in \anc{i} \cap \anc{j}$, where the
-%   node $k$ has a path to $j,\ \gcgpath{k}{j}$, as well as a path
-%   to $i,\ \gcgpath{k}{i}$.
-
-%   \hl{Double check that the ancestor properties proposition implies
-%     that BOTH paths do not contain the other node.}
-
-%   Now, since the graph is strongly causal, the only possible
-%   $\gcgpath{k}{i}$ path is the one obtained by concatenating the
-%   $\gcgpath{k}{j}$ path with the $\gcgpath{j}{i}$ path.
-%   However, this is a contradiction since $k$ is a confounder and
-%   should have a $\gcgpath{k}{i}$ path which does not contain $i$.
-% \end{proof}
-
-\begin{corollary}
-  \label{cor:bidirectional_edge}
-  If $\gcg$ is a strongly causal DAG such that $i \pwgc j$ and
-  $j \pwgc i$, then $i \not\in \anc{j}$ and $j \not\in \anc{i}$.  In
-  particular, a pairwise bidirectional edge indicates the absence of
-  any edge in $\gcg$.
-\end{corollary}
-\begin{proof}
-  This follows directly from applying proposition
-  \ref{cor:parent_corollary} to $i \pwgc j$ and $j \pwgc i$.
-% This is a direct proof
-% \begin{proof}
-  % By way of contradiction, suppose that $i \in \pa{j}$.  We will
-  % consider the two possibilities allowed by proposition
-  % \ref{prop:ancestor_properties} for $j \pwgc i$.  Firstly
-  % $j \in \anc{i}$ is impossible since $\gcg$ is assumed to be acyclic.
-  % Secondly, if there is some confounding node
-  % $u \in \anc{i} \cap \anc{j}$ with a path
-  % $u \rightarrow \cdots \rightarrow j$ which does not contain $i$ we
-  % have a contradiction since there must now be multiple
-  % $u \rightarrow j$ paths: the aforementioned, and a path
-  % $u \rightarrow \cdots \rightarrow i \rightarrow \cdots \rightarrow
-  % j$ which \textit{does} contain $i$.  We conclude that $i \in \pa{j}$
-  % is impossible, and symmetrically that $j \in \pa{i}$ is as well.
-% \end{proof}
-\end{proof}
-
-% \begin{proposition}
-%   \label{prop:independent_confounders}
-%   If $\gcg$ is strongly causal we have distinct
-%   $k_1, k_2 \in \anc{i} \cap \anc{j}$ are confounders of $i, j$, then
-%   $k_1 \not\in \anc{k_2}$, $k_2 \not\in \anc{k_1}$ and
-%   $\not\exists \ell \in \anc{k_1} \cap \anc{k_2}$ confounding
-%   $k_1, k_2$.
-% \end{proposition}
-% \begin{proof}
-%   \hl{TODO}
-% \end{proof}
-
-% \begin{corollary}
-%   If $\gcg$ is strongly causal, then distinct confounders are
-%   uncorrelated.
-% \end{corollary}
-% \begin{proof}
-%   This follows directly by combining proposition
-%   \ref{prop:independent_confounders} with proposition \ref{prop:ancestor_properties}.  \hl{I need to update prop} \ref{prop:ancestor_properties} \hl{to note the strong non-correlation property.}
-% \end{proof}
-
-% \begin{lemma}
-%   If $\gcg$ is a strongly causal DAG and $j \in \anc{i}$ then 
-% \end{lemma}
-
-In light of proposition \ref{prop:sc_graph_common_anc}, the following
-provides a partial converse to proposition
-\ref{prop:ancestor_properties}, and supports the intuition of ``causal
-flow'' through paths in $\gcg$.
-
-\begin{proposition}
-  \label{prop:pwgc_anc}
-  If $\gcg$ is a strongly causal DAG then $j \in \anc{i} \Rightarrow j \pwgc i$.
-\end{proposition}
-\begin{proof}
-  We will show that for some $\psi \in \H_{t - 1}^{(j)}$ we have
-
-  \begin{equation}
-    \label{eqn:cond_ortho_proof}
-    \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} \ne 0
-  \end{equation}
-
-  and therefore that $H_t^{(i)} \not\perp\ \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(i)}$, which by theorem (\ref{thm:granger_causality_equivalences}) is enough to establish that $j \pwgc i$.
-
-  Firstly, we will establish a representation of $x_i(t)$ that involves $x_j(t)$.
-Denote by $a_{r + 1} \rightarrow a_r \rightarrow \cdots \rightarrow a_1 \rightarrow a_0$ with $a_{r + 1} \defeq j$ and $a_0 \defeq i$ the \textit{unique} $\gcgpath{j}{i}$ path in $\gcg$, we will expand the representation of equation (\ref{eqn:parent_expansion}) backwards along this path:
-
-  % Should this be written as a lemma?
-  \begin{align*}
-    x_i(t) &= v_i(t) + \B_{ii}(z) x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z) x_k(t)\\
-           &= \underbrace{v_{a_0}(t) + \B_{a_0a_0}(z) x_i(t) + \sum_{\substack{k \in \pa{a_0} \\ k \ne a_1}}\B_{a_0 k}(z) x_k(t)}_{\defeq \wtalpha{a_0}{a_1}} + \B_{a_0a_1}(z)x_{a_1}(t)\\
-           &= \wtalpha{a_0}{a_1} + \B_{a_0a_1}(z)\big[\wtalpha{a_1}{a_2} + \B_{a_1a_2}(z)x_{a_2}(t) \big]\\
-           &\overset{(a)}{=} \sum_{\ell = 0}^r \underbrace{\Big(\prod_{m = 0}^{\ell - 1} \B_{a_m a_{m + 1}}(z) \Big)}_{\defeq F_\ell(z)} \wtalpha{a_\ell}{a_{\ell + 1}} + \Big(\prod_{m = 0}^{r}\B_{a_m a_{m + 1}}(z)\Big)x_{a_{r + 1}}(t)\\
-           &= \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}} + F_{r + 1}(z) x_j(t)
-  \end{align*}
-
-  where $(a)$ follows by a routine induction argument and where we define $\prod_{m = 0}^{-1} \bullet \defeq 1$ for notational convenience.
-
-  Using this representation to expand equation (\ref{eqn:cond_ortho_proof}), we obtain the following cumbersome expression:
-
-  \begin{align*}
-    &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}}\\
-    &- \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\linE{\sum_{\ell = 0}^r F_\ell(z)\wtalpha{a_\ell}{a_{\ell + 1}}}{\H_{t - 1}^{(i)}}}\\
-    &+ \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\sum_{\ell = 0}^r F_\ell(z)\wtalpha{a_\ell}{a_{\ell + 1}}}.
-  \end{align*}
-
-  Note that by the orthogonality principle, $\psi - \linE{\psi}{\H_{t - 1}^{(i)}} \perp \H_{t - 1}^{(i)}$, the middle term above is $0$.
-Choosing now the particular value $\psi = F_{r + 1}(z)x_j(t) \in \H_{t - 1}^{(j)}$ we arrive at
-
-  \begin{align*}
-    &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}\\
-    &= \E|F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}|^2\\
-    &+ \inner{F_{r + 1}(z)x_j(t) - \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}}}{\sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}}},
-  \end{align*}
-
-  which by the Cauchy-Schwarz inequality is $0$ if and only if
-
-  \begin{equation*}
-    \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}} \overset{\text{a.s.}}{=} \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}} - F_{r + 1}(z)x_j(t),
-  \end{equation*}
-
-  or by rearranging and applying the representation obtained earlier, if and only if
-
-  \begin{equation*}
-    x_i(t) \overset{\text{a.s.}}{=} \linE{F_{r + 1}(z)x_j(t)}{\H_{t - 1}^{(i)}},
-  \end{equation*}
-
-  but this is impossible since $x_i(t) \not \in \H_{t - 1}^{(i)}$.
-\end{proof}
-
-We immediately obtain the corollary, which we remind the reader is,
-surprisingly, not true in a general graph.
-
-\begin{corollary}
-  \label{cor:gc_implies_pwgc}
-  If $\gcg$ is a strongly causal DAG then $j \gc i \Rightarrow j \pwgc i$.
-\end{corollary}
-
-% \begin{remark}
-%   It can be seen from the conclusion of the proof that simple facts
-%   about Granger causality are indeed reliant on our assumptions about
-%   the nature of $x(t)$ laid out in the section \ref{sec:theory}, in
-%   particular, the innovations process $v(t)$ must be full rank.
-
-%   Moreover, that strong conditions need to be placed on the topology
-%   of $\gcg$ in order for proposition \ref{prop:pwgc_anc} to hold can
-%   be seen through the earlier examples.
-% \end{remark}
-
-% This holds in any DAG
-% \begin{lemma}
-%   If $\gcg$ is strongly causal, then any $\\VAR(p)$ system on $\gcg$ is stable if and only if
-%   $B_{ii}(z)$ is stable for every $i = 1, \ldots, n$.
-% \end{lemma}
-% \begin{proof}
-%   \hl{This is certainly true, but I don't think I need to use it anywhere.}
-% \end{proof}
-
-\begin{remark}
-  As we have seen and as is true in much of statistics, confounding
-  nodes pose challenges for Granger-causality.  However, as opposed to
-  Pearl's causal calculus \cite{pearl2000art}, pairwise
-  Granger-causality does not suffer any difficulty with so-called
-  ``colliders'', that is, the topology $i \rightarrow k \leftarrow j$
-  will never result in $i \pwgc j$ or $j \pwgc i$.  This is evidently
-  an advantage of the \textit{temporal} nature of Granger-causality -- there is
-  no backwards causal flow along the edges of $\gcg$.
-\end{remark}
-
-\begin{example}
-  As a final remark of this subsection we note that a complete
-  converse to proposition \ref{prop:ancestor_properties} is not
-  possible without additional conditions.  Consider the ``fork'' system on $3$
-  nodes (i.e. $2 \leftarrow 1 \rightarrow 3$) defined by
-
-  \begin{equation*}
-    x(t) =
-    \left[
-      \begin{array}{cccc}
-        0 & 0 & 0\\
-        a & 0 & 0\\
-        a & 0 & 0\\
-      \end{array}
-    \right] x(t - 1) + v(t).
-  \end{equation*}
-
-  In this case, node $1$ is a confounder for nodes $2$ and $3$, but
-  $x_3(t) = v_3(t) - v_2(t) + x_2(t)$ and $2 \npwgc 3$ (even
-  though $x_2(t)$ and $x_3(t)$ are contemporaneously correlated)
-
-  If we were to augment this system by simply adding an autoregressive
-  component (i.e. some ``memory'') to $x_1(t)$ e.g.
-  $x_1(t) = v_1(t) + b x_1(t - 1)$ then we \textit{would} have
-  $2 \pwgc 3$ since then
-  $x_3(t) = v_3(t) + av_1(t - 1) - bv_2(t - 1) + bx_2(t - 1)$.  We
-  develop this idea further in the next section.
-\end{example}
-
-\subsection{Persistent Systems}
-\label{sec:persistent_systems}
-In section \ref{sec:strongly_causal_graphs} we obtained a converse to
-part $(a)$ of proposition \ref{prop:ancestor_properties} via the
-notion of a strongly causal graph topology.  In this section, we
-complete a converse by adding the additional requirement we refer to
-as ``persistence''.
-
-\begin{definition}[Lag Function]
-  Given a causal filter $\B(z) = \sum_{\tau = 0}^\infty b(\tau)z^{-\tau}$
-  define
-
-  \begin{align}
-    \tau_0(\B) &= \text{min}\{\tau \in \Z_+\ |\ b(\tau) \ne 0\},\\
-    \tau_{\infty}(\B) &= \text{sup}\{\tau \in \Z_+\ |\ b(\tau) \ne 0\}.\\
-  \end{align}
-
-  i.e. the ``first'' and ``last'' coefficients of the filter $\B(z)$,
-  where $\tau_\infty(\B) \defeq \infty$ if the filter has an infinite
-  length, and $\tau_0(\B) \defeq \infty$ if $\B(z) = 0$.
+
+  In order to exclude the possibility of certain types of cancellation
+  in the $\VAR$ system of Equation \eqref{eq:var_model}, we require
+  the system to have a property we have deemed ``persistence'' in
+  \cite{my_GC_paper}, the intuition being that each node in the system
+  maintains some ``memory''.
+
+  \begin{definition}[Persistent]
+    \hl{TODO:}I can provide the diagonalizability condition for
+    $\VAR(p)$ models, but what about when $p = \infty$?
+    \todo{persistence for $\VAR(\infty)$}
   \end{definition}
 
-This interpretation of the following persistence condition is that
-each node stores some ``memory'' of the past.
-
-\begin{definition}[Persistent]
-  We will say that the process $x(t)$ with Granger-causality graph
-  $\gcg$ is \textit{persistent} if for every $i \in [n]$ and every
-  $k \in \anc{i}$ we have $\tau_0(\A_{ik}) < \infty$ and $\tau_\infty(\A_{ik}) = \infty$.
-\end{definition}
-
-\begin{remark}
-  In the context of Granger-causality, ``most'' systems should be
-  persistent.  In particular, $\mathsf{\VAR}(p)$ models are likely to
-  be persistent since these naturally result in an equivalent
-  $\mathsf{MA}(\infty)$ representation, except for the pathological
-  case where $\B(z)$ is Nilpotent.
-
-  Moreover, persistence is not the weakest condition necessary for the
-  results of this section, the condition
-  $\tau_0(\A_{jk}) < \tau_\infty(\A_{ik})$ for each $i, j, k$ such that
-  $k \in \anc{i} \cap \anc{j}$ is enough.  The intuition being that nodes
-  $i$ and $j$ are not receiving temporally disjoint information from
-  $k$.
-
-\end{remark}
-
-\begin{example}
-  process $x(t)$ generated by the $\VAR(1)$ model\footnote{Recall that
-    any $\VAR(p)$ model with $p < \infty$ can be written as a
-    $\VAR(1)$ model, so we lose little generality in considering this
-    case.}  having $\B(z) = Bz^{-1}$, we will examine
-  conditions that guarantee persistence.  Pick any
-  $i \in [n], j \in \anc{i} \setminus \{i\}$, then the stability of
-  $B$ allows us to write
-
-  \begin{equation*}
-    \A(z) = \sum_{k = 0}^\infty B^k z^{-k},
-  \end{equation*}
-
-  whereby we see that $\exists k > 0$ such that $[B^k]_{ij} \ne 0$
-  (since $j \in \anc{i}$).  Then consider
-
-  \begin{equation*}
-    \begin{aligned}
-      e_i^\T B^{rk} e_j &\overset{(a)}{=} \big((P^\T e_i)^\T J^{rk} P^{-1}e_j\big)\\
-      &= \tr [(P^\T e_i)^\T J^{rk} P^{-1}e_j]\\
-      &\overset{(b)}{=} \tr [(J^{rk}) (v u^\T)],
-    \end{aligned}
-  \end{equation*}
-
-  where $(a)$ utilizes the Jordan Normal Form of $B$, and $(b)$
-  denotes $u = P^\T e_i$ and $v = P^{-1}e_j$.  In order for
-  $\tau_\infty(\A_{ij}) < \infty$, there must be some $N > 1$ such
-  that $\forall r \ge N$, the above term is $0$.  This may be the case
-  for instance if $B$ is a nilpotent matrix.
-
-  Let us suppose now that $B$ is diagonalizable (i.e. $J$ is a
-  diagonal matrix) with at least $2$ distinct eigenvalues; in this
-  case $B$ is also \textit{not} nilpotent.  We can then rewrite the
-  above as
-
-  \begin{equation*}
-    f(r) \defeq \tr [(J^{rk}) (v u^\T)] = \sum_{\nu = 1}^n \lambda_\nu ^{rk} v_\nu u_\nu \defeq \sum_{\nu = 1}^n \lambda_\nu^{rk} \beta_\nu
-  \end{equation*}
-
-  where $\lambda_\nu$ denotes the eigenvalues of $B$ and
-  $\beta_\nu = u_\nu v_\nu$.  Note that $f(0) = 0$ since $i \ne j$,
-  $u$ is a row of $P$ and $v$ is a column of $P^{-1}$.  Moreover,
-  $f(1) \ne 0$ by hypothesis.  But, in order for
-  $f(r) = 0\ \forall r \ge N$, it would need to be the case that
-
-  \begin{equation*}
-    \Dg(\bm{\lambda})^r \bm{\lambda} = Vz
-  \end{equation*}
-
-  had a solution in $z$ for every $r \ge N$, where $V$ is an
-  $n \times n - 1$ full-rank matrix whose columns span the nullspace of $\beta$,
-  and $\bm{\lambda} = (\lambda_1, \ldots, \lambda_n)$. That is,
-  iterates of $\Dg(\bm{\lambda})$ applied to $\bm{\lambda}$ would need to remain
-  inside $\beta$'s nullspace.  This would imply that
-
-  \begin{equation*}
-    VV^\dagger \bm{\lambda}^{r + 1} = \bm{\lambda}^{r + 1},
-  \end{equation*}
-
-  i.e. that $\bm{\lambda}^{r + 1}$ is an eigenvector of $VV^\dagger$
-  for an infinite number of integers $r$ (the exponentiation is to be
-  understood as a pointwise operation).  However, since there can only
-  be a finite number of (unit length) eigenvectors, this cannot be the
-  case unless every eigenvalue $(\lambda_1, \ldots, \lambda_n)$ were
-  equal.
-
-  We see from this example that the collection of $\VAR(1)$ systems
-  which are not persistent are pathological, in the sense that their
-  system matrices have zero measure when viewed as a subset of $\R^{n^2}$.
-\end{example}
-
-\begin{lemma}
-  \label{lem:time_lag_cancellation}
-  Suppose $v(t)$ is a scalar sequence with unit variance and zero
-  autocorrelation and let $\A(z), \B(z)$ be nonzero and strictly
-  causal (i.e. $\tau_0(\A) \ge 1$) linear filters.  Then,
-
-  \begin{equation}
-    \inner{F(z)\A(z)v(t)}{\B(z)v(t)} = 0\ \forall \text{ strictly causal filters } F(z)
-  \end{equation}
-
-  if and only if $\tau_0(\A) \ge \tau_\infty(\B)$.
-\end{lemma}
-\begin{proof}
-  We have
-
-  \begin{align}
-    \inner{\A(z)v(t)}{\B(z)v(t)} &= \sum_{\tau = 1}^\infty \sum_{s = 1}^\infty a(\tau)b(s)\E[v(t - s)v(t - \tau)]\\
-    &= \sum_{\tau = \text{max}(\tau_0(\A), \tau_0(\B))}^{\text{min}(\tau_\infty(\A), \tau_\infty(\B))} a(\tau) b(\tau)\\
-  \end{align}
-
-  due to the uncorrelatedness assumptions on $v(t)$.  This expression
-  is $0$ if and only if $\tau_0(\A) \ge 1 + \tau_\infty(\B)$ or if
-  $\tau_0(\B) \ge 1 + \tau_\infty(\A)$ or if the coefficients are
-  orthogonal along the common support.
-
-  Specializing this fact to $\inner{F(z)\A(z)v(t)}{\B(z)v(t)} = 0$ we
-  see that the coefficients cannot be orthogonal for every choice of
-  $F$, and that $\text{sup}_F \tau_\infty(F\A) = \infty$, leaving only
-  the possibility that
-
-  \begin{align*}
-    \tau_0(F\A) \ge 1 + \tau_\infty(\B) \forall F &\overset{(a)}{\iff} \tau_0(\A) \ge 1 + \tau_\infty(\B) - \underset{F}{\text{min }} \tau_0(F)\\
-    &\overset{(b)}{\iff} \tau_0(\A) \ge \tau_\infty(\B),
-  \end{align*}
-
-  where $(a)$ follows since $\tau_0(F\A) = \tau_0(F) + \tau_0(\A)$,
-  and $(b)$ since $\text{min}_F\ \tau_0(F) = 1$.
-\end{proof}
-
-\begin{corollary}
-  \label{cor:time_lag_cancellation}
-  For $k \in \anc{i} \cap \anc{j}$ we have
-
-  \begin{align*}
-    \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} &= 0\ \forall \text{ strictly causal } F(z)\\
-    \iff \inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} &= 0\ \forall \text{ strictly causal } F(z)\\
-    \iff \tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})
-  \end{align*}
-\end{corollary}
-\begin{proof}
-  The final equivalence follows immediately from Lemma \ref{lem:time_lag_cancellation}.  For the first equivalence we have
-
-  \begin{align*}
-    \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} &= 0\ \forall \text{ strictly causal } F(z)\\
-    \iff \inner{F(z)A_{jk}(z)v_k(t)}{x_i(t - \tau)} &= 0\ \forall \tau \ge 1, \text{ strictly causal } F(z),
-  \end{align*}
-
-  which can be expanded by equation \eqref{eqn:ancestor_expansion} to
-  obtain (after cancelling all ancestors of $i$ other than $k$)
-
-  \begin{equation*}
-    \inner{F(z)A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t - \tau)} = 0\ \forall \tau \ge 1, \text{ strictly causal } F(z),
-  \end{equation*}
-
-  which by the Lemma is equivalent to $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$ as stated.
-\end{proof}
-
-\begin{proposition}
-  \label{prop:persistence_converse}
-  Suppose $\gcg$ is a strongly causal DAG and that $x(t)$ is
-  persistent, then if there exists a $k$ which confounds $(i, j)$ we
-  have $i \pwgc j$ and $j \pwgc i$.
-\end{proposition}
-\begin{proof}
-  We will show that $j \pwgc i$, the other being symmetric.  First
-  note also that by proposition \ref{prop:sc_graph_common_anc} we
-  cannot have $i \in \anc{j}$ or $j \in \anc{i}$ and therefore every
-  $k \in \anc{i}\cap\anc{j}$ will be a confounder.
-
-  It is sufficient to show that $\exists \psi \in \H_{t - 1}^{(j)}$
-  such that
-
-  \begin{equation*}
-    \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} \ne 0.
-  \end{equation*}
-
-  To this end, let $F(z)$ be an arbitrary but strictly causal linear
-  filter.  We apply equation \eqref{eqn:ancestor_expansion} to
-  $x_i(t)$ and $\psi \defeq F(z)x_j(t)$:
-
-  \begin{align*}
-    &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}\\
-    &\overset{(a)}{=} \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
-    &\overset{(b)}{=} \inner{\sum_{k \in \anc{j}}\big(F(z)\A_{jk}(z)v_k(t) - \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}\big)}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
-    &\overset{(c)}{=} \sum_{k \in \anc{i}\cap\anc{j}}\Big(\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} - \inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{ii}v_i(t)}\\
-    &- \sum_{\ell \in \anc{i}}\inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{i\ell}(z)v_\ell(t)}\Big)
-  \end{align*}
-
-  where in $(a)$ we have removed the $\linE{x_i(t)}{\H_{t - 1}^{(i)}}$
-  term via the orthogonality principle, in $(b)$ there is no
-  $F(z)\A_{jj}(z)v_j(t)$ term since due to $j \not\in \anc{i}$ it is
-  orthogonal to $\H_t^{(i)}$.  Finally, $(c)$ follows by applying
-  orthogonality properties of $v(t)$, as well as the fact that
-  $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$ for
-  $k \not \in \anc{i}$.  Note that
-  $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$
-  and thence there is in general no cancellation in the final term
-  above for $\ell \in \anc{i}$.
-
-  \todo{This is clearly not immediately evident}
-
-  This is $0$ for every $F$ if and only if for all $F$ and
-  $\forall k \in \anc{i} \cap \anc{j}$ we have
-  $$\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$$ and
-  $$\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} = 0,$$ which by
-  Corollary \ref{cor:time_lag_cancellation} occurs if and only if
-  $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$, which is impossible since by persistence
-  $\tau_0(\A_{jk}) < \infty$ and  $\tau_\infty(\A_{ik}) = \infty$.
-\end{proof}
-
-\subsection{Recovering $\gcg$ via Pairwise Tests}
-\label{sec:pairwise_algorithm}
-In this section we will show that if the $\gcg$ of a persistent
-process is a strongly causal DAG, then it is possible to recover
-$\gcg$ via pairwise tests alone.  This is the main conclusion of the
-theoretical analysis in this paper.
-
-% This is an interesting fact in and
-% of itself, but also has some implications for applications since
-% pairwise testing is trivially parallelizable.  In section
-% \ref{sec:structure_learning} we will also analyze the use of pairwise
-% testing as a heuristic for general graphs.
-
-\begin{theorem}[Pairwise Recovery]
+  The key property of persistent and strongly causal systems is that
+  they are in some sense immune to confounders.  That is, if $i \pwgc j$
+  \textit{and} $j \pwgc i$, then $i \ngc j$ and $j \ngc i$.  That is,
+  bidirectional pairwise causation (which would otherwise be the
+  result of hidden confounders) is not possible.
+
+  \begin{theorem}[Pairwise Recovery \cite{my_GC_paper}]
     \label{thm:scg_recovery}
-  If the Granger-causality graph $\gcg$ for persistent process $x(t)$
-  is a strongly causal DAG then $\gcg$ can be inferred from pairwise
-  causality tests.  The procedure can be carried out, assuming
-  we have an oracle for pairwise causality, via Algorithm
-  (\ref{alg:pwgr}).
-
-  \begin{algorithm}
-    \SetKwInOut{Input}{input}
-    \SetKwInOut{Output}{output}
-    \SetKwInOut{Initialize}{initialize}
-    \DontPrintSemicolon
-
-    \BlankLine
-    \caption{Pairwise Graph Recovery}
-    \label{alg:pwgr}
-    % \TitleOfAlgo{Pairwise Graph Recovery}
-    \Input{Pairwise Granger-causality relations between a persistent
-      process of dimension $n$ whose joint Granger-causality
-      relations are known to form a strongly causal DAG $\gcg$.}
-    \Output{Edges $\gcge = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of
-      the graph $\gcg$.}
-    \Initialize{$S_0 = [n]$  \texttt{\# unprocessed nodes}\\
-      $E_0 = \emptyset$  \texttt{\# edges of }$\gcg$\\
-      % $P_0 = \emptyset$  \texttt{\# layer by layer driving nodes}\\
-      $k = 1$ \texttt{\# a counter used only for notation}}
-    \BlankLine
-    $W \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$  \texttt{\# candidate edges}\\
-    $P_0 \leftarrow \{i \in S_0\ |\ \forall s \in S_0\ (s, i) \not\in W\}$  \texttt{\# parent-less nodes}\\
-    \While{$S_{k - 1} \ne \emptyset$}{
-      $S_k \leftarrow S_{k - 1} \setminus P_{k - 1}$ \texttt{\# remove nodes with depth }$k - 1$\\
-      $P_k \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W\}$   \texttt{\# candidate children of }$P_{k - 1}$\\
-      \;
-
-      $D_{k0} \leftarrow \emptyset$\\
-      \For{$r = 1, \ldots, k$} 
-      {
-        $Q \leftarrow E_{k - 1} \cup \big(\bigcup_{\ell = 0}^{r - 1} D_{k\ell}\big)$ \texttt{\# currently known edges}\\
-        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times P_k\ |\ (i, j) \in W,\ \text{no } \gcgpath{i}{j} \text{ path in } Q\}$
-        % $E_k \leftarrow E_{k - 1} \cup D_{kr}$ \texttt{\# add edges to }$E_k$ \label{alg:inner_loop_end}\\
-        % $W_{k + 1} \leftarrow W_k \setminus D_{kr}$  \texttt{\# remove edges from consideration}\\
-      }
-      $E_k \leftarrow E_{k - 1} \cup \big(\bigcup_{r = 1}^k D_{kr}\big)$ \texttt{\# update } $E_k$ \texttt{ with new edges}\\
-      % \label{alg_line:inner_loop_end}\\
-      % $W_{k + 1} \leftarrow W_k \setminus \big(\bigcup_{r = 0}^k D_{kr}\big)$  \texttt{\# remove edges from consideration}\\
-      \;
-      $k \leftarrow k + 1$
-    }
-    \Return{$E_{k - 1}$}
-  \end{algorithm}
+    If the Granger-causality graph $\gcg$ for persistent process
+    $x(t)$ is a strongly causal DAG, then $\gcg$ can be inferred from
+    pairwise causality tests, given an oracle for pairwise causation.
+
+    \todo{Contain statement on convergence of the heuristic?}
+    % \todo{Make this precisise} Moreover, given $T < \infty$ samples from
+    % $x(t)$ then Algorithm \ref{alg:finite_pwgc} will correctly return
+    % $\gcg$ with probability $1$ as $T \rightarrow \infty$.
   \end{theorem}
-
-We will shortly prove the theorem by establishing the correctness of
-algorithm (\ref{alg:pwgr}).  The idea is to iteratively ``peel away
-layers'' of nodes by removing the nodes that have no parents
-remaining, which always exist since the graph is acyclic.  The
-requirement of strong causality ensures that all actual edges of
-$\gcg$ manifest in some way as pairwise relations (by proposition
-\ref{prop:pwgc_anc}), and the requirement of persistence allows
-confounding to be eliminated by removing bidirectional edges.  Without
-persistence, each confounded pair would give rise to $4$ possible
-pairwise topologies consistent with $\gcg$, one for each type of
-pairwise edge (no edge, unidirectional, bidirectional).
-
-\begin{example}
-  The set $W$ collects ancestor relations in $\gcg$ (see Lemma
-  \ref{lem:W_subset_E}).  In reference to figure
-  \ref{fig:example_fig3}, each of the solid black edges, as well as
-  the dotted red edges will be included in $W$, but \textit{not} the
-  bidirectional green dash-dotted edges, which we are able to exclude
-  by appealing to Corollary \ref{cor:bidirectional_edge}.  The
-  groupings $P_0, \ldots, P_3$ are also indicated in figure
-  \ref{fig:example_fig3}.
-
-  The algorithm proceeds first with the parentless nodes $1, 2$ on the
-  initial iteration where the edge $(1, 3)$ is added to $E$.  On the
-  next iteration, the edges $(3, 4), (2, 4), (3, 5)$ are added, and
-  the false edges $(1, 4), (1, 5)$ are excluded due to the paths
-  $1 \rightarrow 3 \rightarrow 4$ and $1 \rightarrow 3 \rightarrow 5$
-  already being present.  Finally, edge $(4, 6)$ is added, and the false
-  $(1, 6), (3, 6), (2, 6)$ edges are similarly excluded due to the
-  ordering of the inner loop.
-
-  \begin{figure}
-    \centering
-    \caption{Example graph for Algorithm \ref{alg:pwgr}}
-    \footnotesize{Black arrows indicate true parent-child
-      relations.  Red dotted arrows indicate pairwise causality (due to
-      non-parent relations), green dash-dotted arrows indicate
-      bidirectional pairwise causality (due to the confounding node
-      $1$).  Blue groupings indicate each $P_k$ in Algorithm
-      \ref{alg:pwgr}.}
-    \label{fig:example_fig3}
-
-    \begin{subfigure}[b]{0.45\textwidth}
-      \includegraphics[width=\linewidth]{example_algorithm.pdf}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.45\textwidth}
-      \includegraphics[width=\linewidth]{example_algorithm2.pdf}
-    \end{subfigure}
-  \end{figure}
-
-  That we need to proceed backwards through $P_{k - r}$ as in the
-  inner loop of on $r$ can also be seen from this example, where if
-  instead we simply added the set
-
-  \begin{equation*}
-    D_k' = \{(i, j) \in \Big(\bigcup_{r = 1}^k P_{k - r}\Big) \times P_k\ |\ i \pwgc j \}
-  \end{equation*}
-
-  to $E_k$ then we would infer the false positive edge
-  $1 \rightarrow 4$.  Moreover, the same example shows that simply
-  using the set
-
-  \begin{equation*}
-    D_k'' = \{(i, j) \in P_{k - 1} \times P_k\ |\ i \pwgc j \}  ,
-  \end{equation*}
-
-  causes the edge $1 \rightarrow 3$ to be missed.
-\end{example}
-
-% \begin{lemma}
-%   \label{lem:inner_loop_lemma}
-%   At any given iteration, $k$, of Algorithm \ref{alg:pwgr}, if we have
-%   $e \in E$ if and only if $e \in \gcge$, then the inner loop (line \ref{alg:inner_loop})
-%   will add $(i, j) \in \big(\bigcup_{r = 0}^kP_{k - r}\big) \times C_k$ to $E$
-%   if $(i, j) \in \gcge$.  i.e. any real edge will be included.
-% \end{lemma}
-% \begin{proof}
-%   Let $j \in C_k$, $i \in \bigcup_{r = 0}^k P_{k - r}$, and assume
-%   $i \in \pa{j}$.
-
-%   First, suppose that there is no $i \rightarrow j$ path in $E$.  Then
-%   since $i \gc j$ we have $i \pwgc j$ and $j \npwgc i$ (i.e.
-%   $(i, j) \in W$) by Proposition \ref{prop:pwgc_anc} and Corollary
-%   \ref{cor:bidirectional_edge} so at least one of
-%   $\{D_{kr} \}_{r = 0}^k$ contains $(i, j)$ and hence $(i, j)$ will be
-%   added to $E$ upon the end of the iteration.
-
-%   Now, suppose instead that there is already an $i \rightarrow j$ path
-%   in $E$.  Given our assumption that $E$ contains edges if and only if
-%   they are edges of $\gcg$, then either we already have $(i, j) \in E$
-%   correctly, or there is more than one $i \rightarrow j$ path in
-%   $\gcg$, which is not possible since $\gcg$ is strongly causal.
-% \end{proof}
-
-% \begin{lemma}
-%   \label{lem:W_subset_E}
-%   $W \subseteq \gcge$
-% \end{lemma}
-% \begin{proof}
-%   By proposition \ref{prop:pwgc_anc} we have
-%   $j \in \pa{i} \Rightarrow j \pwgc i$ and by Corollary \ref{cor:bidirectional_edge}
-%   there are no bidirectional edges in $\gcg$.
-% \end{proof}
-
-% \hl{Must show that Algorithm terminates in $n$ steps.}
-% \begin{lemma}
-%   For any $i \in P_k$ we have $i \in C_{k - 1}$.  \hl{[Can't yet prove
-%     this part] Moreover, for each $i \in \bigcup_{k = 0}^n P_k$, if
-%     $(s, i) \in W$ then $(s, i) \in E$ upon the termination of
-%     Algorithm} \ref{alg:pwgr}.
-% \end{lemma}
-% \begin{proof}
-%   We prove the second part by induction, the first part being
-%   established along the way.  By construction $P_0$ has no parent
-%   nodes in $\gcg$ so the base case is immediate.  Fix now some
-%   $k \ge 1$, $i \in P_k$, and let $s$ be such that $(s, i) \in W$.
-%   For every $s$ such that $(s, i) \in W$ we have $s \not\in S_k$ since
-%   otherwise $i \not\in P_k$.  Therefore
-%   $s \in \bigcup_{r = 0}^{k - 1}P_{k - r}$, which implies
-%   $s \not\in S_k$, and finally that $i \in C_{k - 1}$.  Then by Lemma
-%   \ref{lem:W_subset_E} every edge in $\gcge$ is also present in
-%   $W$, and by Lemma \ref{lem:inner_loop_lemma} \hl{[We can't conclude
-%     here because of the strong conditions required on lemma}
-%     \ref{lem:inner_loop_lemma}]
-% \end{proof}
-
-% \begin{lemma}[Loop Invariant]
-%   \label{lem:loop_invariant}
-%   $\forall k \ge 0,\ e \in E_k \Rightarrow e \in \gcge$.
-%   \hl{Not sure if this is useful / the right approach.}
-% \end{lemma}
-% \begin{proof}
-%   We will proceed by induction.  Firstly, the base case
-%   $E_0 = \emptyset$ is trivial, so fix some $k \ge 1$ and consider an
-%   edge $(i, j) \in \big(\bigcup_{r = 0}^k P_{k - r} \big) \times C_k$
-%   such that $(i, j) \in W$ (if the edge is not in $W$ then it is not
-%   in $\gcge$ by Lemma \ref{lem:W_subset_E}).
-
-%   \hl{WIP}
-% \end{proof}
-
-Our proof proceeds in 5 steps stated formally as lemmas.  Firstly, we
-characterize the sets $W$ and $P_k$.  Then we establish a correctness
-result for the inner loop on $r$, a correctness result for the outer
-loop on $k$, and finally that the algorithm terminates in a finite
-number of steps.
-
-\begin{lemma}[$W$ Represents Ancestor Relations]
-  \label{lem:W_subset_E}
-  In Algorithm \ref{alg:pwgr} we have
-  $(i, j) \in W$ if and only if $i \in \anc{j}$.  In particular,
-  $W \subseteq \gcge$.
-\end{lemma}
-\begin{proof}
-  Let $j \in [n]$ and suppose that $i \in \anc{j}$.  Then $i \pwgc j$
-  by Proposition \ref{prop:pwgc_anc}.  Proposition
-  \ref{prop:sc_graph_common_anc} ensures that $(i, j)$ are not
-  confounded and Corollary \ref{cor:parent_corollary} that
-  $j \not\in \anc{i}$ so $j \npwgc i$ and thence by Proposition
-  \ref{prop:ancestor_properties} $(i, j) \in W$.
-
-  Conversely, suppose $(i, j) \in W$.  Then since $j \npwgc i$
-  Proposition \ref{prop:persistence_converse} ensures that $(j, i)$
-  are not confounded and so by Proposition \ref{prop:ancestor_properties}
-  we must have $i \in \anc{j}$.
-\end{proof}
-
-\begin{definition}[Depth]
-  For our present purposes we will define the \textit{depth} $d(j)$ of
-  a node $j$ in $\gcg$ to be the length of the \textit{longest} path
-  from a node in $P_0$ to $j$, where $d(j) = 0$ if $j \in P_0$.  It is
-  apparent that such a path will always exist.  For example, in Figure
-  \ref{fig:example_fig3} we have $d(3) = 1$ and $d(4) = 2$.
-\end{definition}
-
-\begin{lemma}[Depth Characterization of $P_k$]
-  \label{lem:depth_lemma}
-  $i \in P_k \iff d(i) = k$ and $j \in S_k \iff d(j) \ge k$.
-\end{lemma}
-\begin{proof}
-  We proceed by induction, noting that $P_0$ is non-empty since $\gcg$
-  is acyclic and therefore $\gcg$ contains nodes without parents.  The
-  base case $i \in P_0 \iff d(i) = 0$ is by definition, and
-  $j \in S_0 \iff d(j) \ge 0$ is trivial since $S_0 = [n]$.  So
-  suppose that the lemma is true up to $k - 1$.
-
-  ($i \in P_k \implies d(i) = k$): Let $i \in P_k$.  Suppose that
-  $d(i) \ge k + 1$, then $\exists j \in \pa{i}$ such that
-  $j \not\in \cup_{r \ge 1}P_{k - r}$ (otherwise $d(i) \le k$), this
-  implies that $j \in S_k$ with $(j, i) \in W$ (by Lemma
-  \ref{lem:W_subset_E}) which is not possible due to the construction of
-  $P_k$ and therefore $d(i) \le k$.  Moreover,
-  $P_k \subseteq S_k \subseteq S_{k - 1}$ implies that
-  $d(i) \ge k - 1$ by the induction hypothesis, but if $d(i) = k - 1$
-  then $i \in P_{k - 1}$ again by induction which is impossible since
-  $i \in P_k$ and therefore $d(i) = k$.
-
-  ($s \in S_k \implies d(s) \ge k$): Let
-  $s \in S_k \subseteq S_{k - 1}$.  We have by induction that
-  $d(s) \ge k - 1$, but again by induction (this time on $P_{k - 1}$)
-  we have $d(s) \ne k - 1$ since $S_k = S_{k - 1} \setminus P_{k - 1}$
-  and therefore $d(s) \ge k$.
-
-  ($d(i) = k \implies i \in P_k$): Suppose $i \in [n]$ is such that
-  $d(i) = k$.  Then $i \in S_{k - 1}$ by the hypothesis, but also
-  $i \not\in P_{k - 1}$ so then
-  $i \in S_k = S_{k - 1} \setminus P_{k - 1}$ and thus $d(i) \ge k$.
-  Now, recalling the definition of $P_k$
-
-  \begin{equation*}
-    P_k = \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W \},
-  \end{equation*}
-
-  if $s \in S_k$ is such that $(s, i) \in W$ then $s \pwgc i$ and
-  $i \npwgc s$ so that by persistence and Proposition
-  \ref{prop:persistence_converse} there cannot be a confounder of
-  $(s, i)$ (otherwise $i \pwgc s$) so then by Proposition
-  \ref{prop:ancestor_properties} we have $s \in \anc{i}$.  We have
-  shown that $s \in S_k \implies d(s) \ge k$ and so we must have
-  $d(i) > k$, a contradiction, thence $s \not\in \anc{i}$,
-  $s \npwgc i$, $(s, i) \not\in W$ and $i \in P_k$.
-
-  ($d(j) \ge k \implies j \in S_k$): Let $j \in [n]$ such that
-  $d(j) \ge k$, then by induction we have $j \in S_{k - 1}$.  This
-  implies by the construction of $S_k$ that $j \not\in S_k$ only if
-  $j \in P_{k - 1}$, but we have shown that this only occurs when
-  $d(j) = k - 1$, but $d(j) > k - 1$ so $j \in S_k$.
-\end{proof}
-
-\begin{lemma}[Inner Loop]
-  \label{lem:inner_loop_lemma}
-  Fix an integer $k \ge 1$ and suppose that $(i, j) \in E_{k - 1}$ if
-  and only if $(i, j) \in \gcge$ and $d(j) \le k - 1$.  Then, we have
-  $(i, j) \in D_{kr}$ if and only if $(i, j) \in \gcge $, $d(j) = k$,
-  and $d(i) = k - r$.
-\end{lemma}
-\begin{proof}
-  We prove by induction on $r$, keeping in mind the results of Lemmas
-  \ref{lem:W_subset_E} and \ref{lem:depth_lemma}.  For the base case,
-  let $r = 1$ and suppose that $(i, j) \in \gcge$ with $d(j) = k$ and
-  $d(i) = k - 1$.  Then, $(i, j) \in W$ and by our assumptions on
-  $E_{k - 1}$ there is no $\gcgpath{i}{j}$ path in $E_{k - 1}$
-  and therefore $(i, j) \in D_{k1}$.  Conversely, suppose that
-  $(i, j) \in D_{k1}$.  Then, $d(i) = k - 1$ and $d(j) = k$ which, since
-  $(i, j) \in W \implies i \in \anc{j}$ implies that
-  $i \in \pa{j}$ and $(i, j) \in \gcge$.
-
-  Now, fix $r > 1$ and suppose that the result holds up to $r - 1$.
-  Let $(i, j) \in \gcge$ with $d(j) = k$ and $d(i) = k - r$.  Then,
-  $(i, j) \in W$ and by induction and strong causality there cannot
-  already be an $\gcgpath{i}{j}$ path in
-  $E_{k - 1} \cup \big(\bigcup_{\ell = 0}^{r - 1} D_{kr}\big)$,
-  therefore $(i, j) \in D_{kr}$.  Conversely, suppose
-  $(i, j) \in D_{kr}$.  Then we have $d(i) = k - r$, $d(j) = k$, and
-  $i \in \anc{j}$.  Suppose by way of contradiction that
-  $i \not\in \pa{j}$, then there must be some $u \in \pa{j}$ such that
-  $i \in \anc{u}$.  But, this implies that $d(i) < d(u)$ and by
-  induction that $(u, j) \in \bigcup_{\ell = 1}^{r - 1}D_{k\ell}$.
-  Moreover, since $d(u) < k$ (otherwise $d(j) > k$) each edge in
-  the $\gcgpath{i}{u}$ path must already be in $E_{k - 1}$, and so
-  there must be an $\gcgpath{i}{j}$ path in
-  $E_{k - 1}\cup\big(\bigcup_{\ell = 0}^{r - 1}D_{kr}\big)$, which is
-  a contradiction since we assumed $(i, j) \in D_{kr}$.  Therefore
-  $i \in \pa{j}$ and $(i, j) \in \gcge$.
-\end{proof}
-
-\begin{lemma}[Outer Loop]
-  \label{lem:outer_loop_lemma}
-  We have $(i, j) \in E_k$ if and only if $(i, j) \in \gcge$ and
-  $d(j) \le k$.  That is, at iteration $k, E_k$ and $\gcge$ agree on
-  the set of edges whose terminating node is at most $k$ steps away
-  from $P_0$.
-\end{lemma}
-\begin{proof}
-  We will proceed by induction.  The base case $E_0 = \emptyset$ is
-  trivial, so fix some $k \ge 1$, and suppose that the lemma holds for
-  all nodes of depth less than $k$.
-
-  Suppose that
-  $(i, j) \in E_k = E_{k - 1}\cup \big(\bigcup_{r = 1}^k D_{rk}
-  \big)$.  Then clearly there is some $1 \le r \le k$ such that
-  $(i, j) \in D_{kr}$ so that by Lemma \ref{lem:inner_loop_lemma} we
-  have $(i, j) \in \gcge$ and $d(j) = k$.
-
-  Conversely, suppose that $(i, j) \in \gcge$ and $d(j) \le k$.  If
-  $d(j) < k$ then by induction $(i, j) \in E_{k - 1} \subseteq E_k$ so
-  suppose further than $d(j) = k$.  Since $i \in \pa{j}$ we must have
-  $d(i) < k$ (else $d(j) > k$) and again by Lemma
-  \ref{lem:inner_loop_lemma} $(i, j) \in \bigcup_{r = 1}^k D_{kr}$
-  which implies that $(i, j) \in E_k$.  The result follows.
-\end{proof}
-
-\begin{lemma}[Finite Termination]
-  Algorithm \ref{alg:pwgr} terminates and returns $E_{k^\star - 1} = \gcge$
-  for some $k^\star \le n$.
-\end{lemma}
-\begin{proof}
-  If $n = 1$, the algorithm is clearly correct, returning on the first
-  iteration with $E_1 = \emptyset$.  When $n > 1$ Lemma
-  \ref{lem:outer_loop_lemma} ensures that $E_k$ coincides with
-  $\{(i, j) \in \gcge\ |\ d(j) \le k\}$ and since $d(j) \le n - 1$ for
-  any $j \in [n]$ there is some $k^\star \le n$ such that
-  $E_{k^\star - 1} = \gcge$.  We must have $S_{k^\star} = \emptyset$
-  since $j \in S_{k^\star} \iff d(j) \ge k^\star$ (if $d(j) > k - 1$ then
-  $E_{k^\star - 1} \ne \gcge$) and therefore the algorithm terminates.
-\end{proof}
-
-% \begin{proof}
-%   Denote the edges of $\gcg$ by $\gcge$, we endevour to show
-%   that upon the termination of Algorithm \ref{alg:pwgr} we have
-%   $E = \gcge$.  On the first iteration of the algorithm
-%   ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
-
-%   Moving on from any trivial cases, notice that
-%   $\gcge \subseteq W$.  This follows since by Proposition
-%   \ref{prop:pwgc_anc} we have $j \in \pa{i} \Rightarrow j \pwgc i$ and by
-%   Corollary \ref{cor:bidirectional_edge} there are no bidirectional
-%   edges in $\gcg$.
-
-% 1. Dispense with trivialities
-% 2. Show that hat{P}_k is nonempty
-% 3. Show that C_k is nonempty
-% 4. Show that edges in D_k are actual edges in G
-% 5. show that each edge incident on C_k is in D_k
-% 6. Induction
-
-% Given non-empty P and C...  Can we show all P -> C edges get collected in D?
-  % Firstly, if $n = 0$ or $n = 1$ the algorithm is trivially correct, returning on the first iteration with $E = \emptyset$.  Suppose then that $n > 1$ and let us define
-
-  % \begin{equation*}
-  %   A_k = \{i \in [n]\ |\ \dist{i}{P_0} = k\},
-  % \end{equation*}
-
-  % where
-
-  % \begin{equation*}
-  %   \dist{i}{P} = \text{max}\{r \in [n]\ |\ \exists (j \rightarrow i) \text{ path of length } r \text{ in } \gcg \text{ for some } j \in P\}.
-  % \end{equation*}
-
-  % We will proceed by induction and show that when the counter has reached the value $k \ge 1$, it is guaranteed that all nodes in $\bigcup_{\ell \le k}A_\ell$ (nodes having paths of length up to $k$ between themselves and $P_0$) must have all of their incident edges included in $E$, and that every edge in $E$ is an edge of $\gcg$.
-If this is the case, then it can be immediately seen that the algorithm terminates with the correct result, $\gcg = ([n], E)$, once $k = n$; since there can only be paths up to length $n - 1$ in $\gcg$, and there cannot be any nodes or edges that don't involve paths back to $P_0$.
-That is, $\bigcup_{\ell \le n}A_\ell = [n]$.
-
-  % % firstly dispensing with some trivialities, and then establishing that every edge in $\gcg$ makes an appearance in some $D_{kr}$ as well as that everything in any $D_{kr}$ is indeed an edge in $\gcg$.
-
-  % On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E = \emptyset$.
-To quickly dispense with some trivialities: the set $P_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $P_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
-Now, if $P_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
-
-  % Continuing with the non-trivial case: if $P_0 \ne [n]$ then $C_0$ must be non-empty, otherwise there would be driving nodes not included in $P_0$.
-In this case it is also clear that $A_1 \ne \emptyset$, so let us consider some $j \in A_1$.
-We must have $j \in C_0$, otherwise there would be a path of length $2$ from $P_0$ to $j$.
-Moreover, since $\exists i \in P_0$ s.t.
-$i \gc j$ and therefore $i \pwgc j$ (by proposition \ref{prop:pwgc_anc}) we have $(i, j) \in D_{00}$ and therefore $(i, j) \in E$.
-Therefore $A_1 \subseteq D_{00}$.
-\hl{This is nonsensical because $A_1$ does not contain tuples -- need to show that $A_1 \subseteq C_0$ ?}
-
-  % Now consdider some $(i, j) \in D_{00}$ (the same reasoning as for $C_0$ is sufficient to see that $D_{00}$ is non-empty).
-Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).
-Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in P_0 \Rightarrow u \in S_0 \Rightarrow j \not\in C_0$, a contradiction.
-This implies every tuple in $D_{00}$ is a bona-fide edge of $\gcg$, as well as that $(i, j) \in A_1$.
-So then $A_1 = D_{00}$, [\hl{nonsensical, see earlier.}] completing the base case for induction.
-
-  % Suppose now the algorithm has reached step $k$ and that $E = \bigcup_{\ell \le k - 1}A_\ell$.
-
-% The reasoning here is (I think) correct -- but isn't directly establishing the induction hypothesis.
-%   On the other hand, if $P_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $P_0$.
-Consider now an edge $(i, j) \in D_{00}$ ($D_{00}$ is non-empty for the same reason as $C_0$).
-Since $i$ has no parents, $\anc{i} \cap \anc{j} = \emptyset$ and therefore $i \in \anc{j}$ (proposition \ref{prop:ancestor_properties}).
-Furthermore, the construction of $C_k$ implies that $i \in \pa{j}$ since otherwise, there would be some $u \in \anc{j}$ with $i \in \anc{u}$, but then by proposition \ref{prop:pwgc_anc} $i \pwgc u$ and $u \pwgc j$ so that $u \not \in P_0 \Rightarrow u \in S_0 \Rightarrow j \not\in C_0$, a contradiction.
-So every tuple in $D_{00}$ is a bona-fide edge of $\gcg$.
-Suppose now that for some $j \in C_0$ there is an $i$ s.t.
-$(i, j) \in \gcg$ but $(i, j) \not\in D_{00}$.
-Since $(i, j) \in \gcg$ we have $i \pwgc j$ (corollary \ref{cor:gc_implies_pwgc}), but then if $i \not\in P_0$ then $j \notin C_0$ and if $i \in P_0, j \in C_0$ then necessarily $(i, j) \in D_{00}$ since we cannot at this stage have any paths in $E$.
-
-%   and show that on each step, every edge incident upon a node in $C_k$ is contained in $D_k$, that every edge in $D_k$ is an edge in $\gcg$, and finally that $\bigcup_{k = 1}^n C_k = [n] \setminus P_0$ and that this implies $\big([n], \bigcup_{k = 1}^n D_k\big) = \gcg$.
-
-%  On the first iteration of the algorithm ($k = 0$) we have $S_0 = [n]$ and $E_0 = \emptyset$.
-The set $P_0$ is non-empty since $\gcg$ is a DAG and therefore there are nodes in $\gcg$ without parents, thence if $P_0 = \emptyset$ proposition \ref{prop:ancestor_properties} would be contradicted.
-Now, if $P_0 = [n]$ then the graph necessarily contains no edges and we will correctly return $E = \emptyset$.
-On the other hand, if $P_0 \ne [n]$ then $C_0$ must also be non-empty, otherwise there would be driving nodes not included in $P_0$.
-Consider now an edge $(i, j) \in D_0$ (again $D_0$ is non-empty for the same reason as $C_0$).
-Since $i$ has no parents, we must have $\anc{i}\cap\anc{j} = \emptyset$ and therefore $i \in \anc{j}$, and due to the construction of $C_0$ we in fact have $i \in \pa{j}$, which establishes that every edge in $D_0$ is also an edge in $\gcg$.
-Finally consider $j \in C_0$, by proposition \ref{prop:pwgc_anc} and since the parents of $C_0$ must be in $P_0$, we see that every edge incident on $j$ must be present in $D_0$.
-
-% We also see that since nodes in $P_0$ have no parents,
-
-% we must have $\anc{i}\cap\anc{j} = \emptyset$, and therefore by proposition \ref{prop:ancestor_properties} $i \in \anc{j}$;  moreover, the construction of $P_1$ implies that in fact $i \in \pa{j}$, so $(i, j)$ is an edge of $\gcg$.
-
-%Finally, by similar reasoning, there cannot be an edge $(i, j) \in D_0$ which is not in $\gcg$, since $\anc{i} \cap \anc{j} = \emptyset$.
-
-%Consider now step $K < n$ of algorithm \ref{alg:pwgr}.
-By lemma \ref{lem:still_strongly_causal} the graph remains strongly causal after having removed nodes $\bigcup_{k = 0}^K P_k$ from $S$, and we assume for induction that $\bigcup_{k = 1}^K D_k$ are all edges of $\gcg$.
-
-% \hl{WIP}
-% \end{proof}
-
-\begin{example}
-  We close this section by noting that the conditions of persistence
-  and strong causality are only sufficient conditions.  For example,
-  the complete directed graph with 2 nodes i.e.
-
-  \begin{equation*}
-    B(1) = \left[ \begin{array}{cc} 1/2 & 1 \\ 1 & 1/2 \end{array}\right]
-  \end{equation*}
-
-  contains a loop but is pairwise recoverable, though not by algorithm
-  (\ref{alg:pwgr}).  Clearly, this example is somewhat artificial
-  since when $n = 2$ there is no difference between pairwise
-  Granger-causality and joint Granger-causality amongst all series --
-  however, one can add any number of nodes having no parents or
-  children to a graph containing a length 2 cycle, in which case the
-  graph clearly remains pairwise recoverable.
-\end{example}
 
@@
 \label{sec:structure_learning}
-In this section we provide a review of our methods for implementing
-Algorithm 1 given a \textit{finite} sample of $T$ data points.  We
-apply the simplest reasonable methods in order to maintain a focus on
-our main contributions (i.e. Algorithm \ref{alg:pwgr}), more
-sophisticated schemes can only serve to improve the results.  Textbook
-reviews of the following concepts are provided e.g. by
-\cite{all_of_statistics}, \cite{murphy_mlp}, and elsewhere.
-
-In subsection \ref{sec:pairwise_hypothesis_testing} we define pairwise
-Granger-causality hypothesis tests, in subsection
-\ref{sec:model_order_selection} a model order selection criteria, in
-subsection \ref{sec:efficient_model_estimation} an efficient
-estimation algorithm, in subsection \ref{sec:error_rate_control} the
-method for choosing an hypothesis testing threshold, and finally in
-subsection \ref{sec:finite_pwgc} the unified finite sample algorithm.
-
-\subsection{Pairwise Hypothesis Testing}
-\label{sec:pairwise_hypothesis_testing}
-In performing pairwise checks for Granger-causality $x_j \pwgc x_i$ we
-follow the simple scheme of estimating the following two linear models:
-
-\begin{align}
-  H_0:&\ \widehat{x}_i^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau),\\
-  H_1:&\ \widehat{x}_i^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau) + \sum_{\tau = 1}^pb_{ij}(\tau)x_j(t - \tau).
-\end{align}
-
-We formulate the statistic
-
-\begin{equation}
-  \label{eqn:gc_statistics}
-  F_{ij}(p) = \frac{T}{p}\Big(\frac{\xi_i(p)}{\xi_{ij}(p)} - 1\Big),
-\end{equation}
-
-where $\xi_i(p)$ is the sample mean square of the
-residuals\footnote{This quantity is often denoted $\widehat{\sigma}$,
-  but we maintain notation from Definition
-  \ref{def:granger_causality}.}  $x_i(t) - \widehat{x}^{(p)}_i(t)$,
-
-\begin{equation*}
-  \xi_i(p) = \frac{1}{T - p}\sum_{t = p + 1}^T (x_i(t) - \widehat{x}_i^{(p)}(t))^2,
-\end{equation*}
-
-and similarly for $\xi_{ij}(p)$.  We test $F_{ij}(p)$ against a
-$\chi^2(p)$ distribution.
-
-If the estimation procedure is consistent, we will have the following
-convergence (in $\P$ or a.s.):
-
-\begin{equation}
-  F_{ij}(p) \rightarrow
-  \left\{
-    \begin{array}{ll}
-      0;\ x_j \npwgc x_i\\
-      \infty;\ x_j \pwgc x_i
-    \end{array}
-  \right. \text{ as } T \rightarrow \infty.  % the '.' after \right is necessary.
-\end{equation}
-
-In our finite sample implementation (see Algorithm
-\ref{alg:finite_pwgc}) we add edges to $\widehat{\gcg}$ in order of
-the decreasing magnitude of $F_{ij}$ instead of proceeding backwards
-through $P_{k - r}$ in Algorithm \ref{alg:pwgr}.  This makes greater
-use of the information provided by the test statistic $F_{ij}$,
-moreover, if $x_i \gc x_j$ and $x_j \gc x_k$, it is expected that
-$F_{kj} > F_{ki}$, thereby providing the same effect as proceeding
-backwards through $P_{k - r}$.
-\subsection{Model Order Selection}
-\label{sec:model_order_selection}
-There are a variety of methods to choose the filter order $p$ (see
-e.g. \cite{lutkepohl2005new}), but we will focus in particular on the
-Bayesian Information Criteria (BIC).  The BIC is substantially more
-conservative than the popular alternative Akaiake Information Criteria
-(the BIC is also asymptotically consistent), and since we are
-searching for \textit{sparse graphs}, we therefore prefer the BIC,
-where we seek to \textit{minimize} over $p$:
-
-\begin{equation}
-  \label{eqn:bic}
-  \begin{aligned}
-    BIC_{\text{univariate}}(p) &= \ln\ \xi_i(p) + p\frac{\ln T}{T},\\
-    BIC_{\text{bivariate}}(p) &= \ln \det \widehat{\Sigma}_{ij}(p) + 4p\frac{\ln T}{T},\\
-  \end{aligned}
-\end{equation}
-
-where $\widehat{\Sigma}_{ij}(p)$ is the $2 \times 2$ residual
-covariance matrix for the $\VAR(p)$ model of $(x_i(t), x_j(t))$.  The
-bivariate errors $\xi_{ij}(p)$ and $\xi_{ji}(p)$ are the diagonal
-entries of $\widehat{\Sigma}_{ij}(p)$.
-
-We carry this out by a simple direct search on each model order
-between $0$ and some prescribed $p_\text{max}$, resulting in a
-collection $p_{ij}$ of model order estimates.  In practice, it is
-sufficient to pick $p_\text{max}$ ad-hoc or via some simple heuristic
-e.g. plotting the sequence $BIC(p)$ over $p$, though it is not
-technically possible to guarantee that the optimal $p$ is less than
-the chosen $p_\text{max}$ (since there can in general be arbitrarily
-long lags from one variable to another).
-
-\subsection{Efficient Model Estimation}
-\label{sec:efficient_model_estimation}
-In practice, the vast majority of computational effort involved in
-implementing our estimation algorithm is spent calculating the error
-estimates $\xi_i(p_i)$ and $\xi_{ij}(p_{ij})$.  This requires fitting a
-total of $n^2p_{\text{max}}$ autoregressive models, where the most
-naive algorithm (e.g. solving a least squares problem for each model)
-for this task will consume $O(n^2p_{\text{max}}^4T)$ time, it is
-possible to carry out this task in a much more modest
-$O(n^2p_{\text{max}}^2 ) + O(n^2p_{\text{max}}T)$ time via the
-autocorrelation method
-\cite{hayes_statistical_digital_signal_processing} which substitutes
-the following autocovariance estimates in the Yule-Walker
-equations:\footnote{The particular indexing and normalization given in
-  equation \ref{eqn:covariance_estimate} is critical to ensure
-  $\widehat{R}$ is positive semidefinite.  The estimate can be viewed
-  as calculating the covariance sequence of a signal multiplied by a
-  rectangular window.}
-
-\begin{equation}
-  \label{eqn:covariance_estimate}
-  \widehat{R}_x(\tau) = \frac{1}{T}\sum_{t = \tau + 1}^T x(t) x(t - \tau)^\T;\ \tau = 0, \ldots, p_{\text{max}},
-\end{equation}
-
-It is imperative that the first index in the summation is $\tau + 1$, as
-opposed perhaps to $p_\text{max}$ and that the normalization is
-$1 / T$, as opposed perhaps to $1 / (T - p_\text{max})$, in order to
-guarantee that $\widehat{R}_x(\tau)$ forms a valid (i.e. positive
-definite) covariance sequence.  This results in some bias, however the
-dramatic computational speedup is worth it for our purposes.
-
-These covariance estimates constitute the $O(n^2p_{\text{max}}T)$
-operation.  Given these particular estimates, the variances $\xi_i(p)$
-for $p = 1, \ldots, p_{\text{max}}$ can be evaluated in
-$O(p_{\text{max}}^2)$ time each by applying the Levinson-Durbin
-recursion to $\widehat{R}_{ii}(\tau)$, which effectively estimates a
-sequence of $AR$ models, producing $\xi_i(p)$ as a side-effect (see
-\cite{hayes_statistical_digital_signal_processing} and
-\cite{levinson_durbin_recursion}).
-
-Similarly, the variance estimates $\widehat{\Sigma}_{ij}(p)$ (which
-include $\xi_{ij}$ and $\xi_{ji}$) can be obtained by estimating
-$\frac{(n + 1)n}{2}$ bivariate AR models, again in
-$O(p_{\text{max}}^2)$ time via Whittle's generalized Levinson-Durbin
-recursion\footnote{We have made use of standalone tailor made
-  implementations of these algorithms, available at
-  \textsf{github.com/RJTK/Levinson-Durbin-Recursion}.}
-\cite{whittle_generalized_levinson_durbin}.
-
-\subsection{Edge Probabilities and Error Rate Controls}
-\label{sec:error_rate_control}
-Denote $F_{ij}$ the Granger-causality statistic of equation
-\ref{eqn:gc_statistics} with model orders chosen by the methods of
-Section \ref{sec:model_order_selection}.  We assume that this
-statistic is asymptotically $\chi^2(p_{ij})$ distributed (the
-disturbances are Gaussian), and denote by $G$ the cumulative
-distribution function thereof.  We will define the matrix
-
-\begin{equation}
-  \label{eqn:edge_inclusion_probability}
-  P_{ij} = G(F_{ij}),
-\end{equation}
-
-to be the matrix of pairwise edge inclusion P-values.  This is
-motivated by the hypothesis test where the hypothesis $H_0$ will be
-rejected (and thence we will conclude that $x_j \pwgc x_i$) if
-$P_{ij} > 1 - \delta$.
-
-The value $\delta$ can be chosen by a variety of methods, in our case
-we apply the Benjamini Hochberg criteria \cite{benjamini_hochberg}
-\cite{all_of_statistics} to control the false discovery rate of
-pairwise edges to a level $\alpha$ (where we generally take
-$\alpha = 0.05$).
-
-\subsection{Finite Sample Recovery Algorithm}
-\label{sec:finite_pwgc}
-
-After the graph topology $\widehat{\gcg}$ has been estimated via
-Algorithm \ref{alg:finite_pwgc}, we refit the entire model with the
-specified sparsity pattern directly via ordinary least squares.
-
-We note that producing graph estimates which are not strongly causal
-can potentially be achieved by performing sequential estimates
-$\widehat{x}_1(t), \widehat{x}_2(t), \ldots$ estimating a strongly causal
-graph with the residuals of the previous model as input, and then
-refitting on the combined sparsity pattern.  We experiment with this
-heuristic in our example application of Section \ref{sec:application},
-but reserve theoretical analysis for future work.
+In \cite{my_GC_paper} we prove Theorem \ref{thm:scg_recovery} by
+providing an algorithm which, given an oracle for pairwise
+Granger-causality, will correctly recover $\gcg$.  In this section,
+we develop in detail appropriate statistical methods for
+implementing this algorithm on a finite data sample.  When applied
+in practice, the algorithm is a heuristic\footnote{Recall that
+  Granger-causality is necessarily based on heuristics when $n$ is
+  modestly large, as the complexity of searching over causality
+  graphs scales as $O(2^{n^2})$}, since we do not know apriori
+whether the causality graph of $x(t)$ is strongly causal (or if
+$x(t)$ is persistent).  However, we will see in Section
+\ref{sec:empirical_evaluation} that Algorithm \ref{alg:finite_pwgc}
+outperforms LASSO-based heuristics, even when $\gcg$ does not
+exactly satisfy the strong causality assumption.
+
+In the appendix Section \ref{sec:pairwise_hypothesis_testing} we describe
+straightforward methods for implementing Algorithm
+\ref{alg:finite_pwgc}.  In particular, we consider it important to be
+able to test pairwise causation in a computationally efficient manner,
+since $O(n^2)$ pairwise tests are needed.  In practice, carrying out
+each pairwise test consumes the vast majority of computational time,
+though since this step is trivially parallelizable, we are able to
+scale the estimation procedure to over $1000$ nodes on a personal
+computer.
 
@@
 
-% \section{Structure Learning}
-% \label{sec:structure_learning}
-% \subsection{Local Search Heuristics}
-
-% \subsection{Edge-Wise Grouped LASSO}
-% Minimize the sparsity promoting regularized least squares problem and choose the hyper-parameters against the Akaike information criteria.
-
-% \begin{equation}
-%   L(\lambda, p) \defeq \underset{B}{\text{minimize}}\ \frac{1}{T}\sum_{t = 1}^T||x(t) - \sum_{\tau = 1}^pB(\tau)x(t - \tau)||_2^2 + \lambda \sum_{i, j}\big[\alpha||B_{i, j}||_2 + (1 - \alpha)||B_{i, j}||_1\big]
-% \end{equation}
-
-% \begin{equation}
-%   \underset{\lambda, p}{\text{minimize}}\ L(\lambda, p) + \mathsf{AIC}(B^{(\lambda, p)})
-% \end{equation}
-
-% \subsection{Bayesian Posterior Thresholding}
-% Consider the Bayesian model
-
-% \begin{equation}
-%   \begin{aligned}
-%     (x(t)\ |\ B, \sigma_v^2, \{x(t - \tau)\}_{\tau = 1}^p) &\sim \mathcal{N}(\sum_{\tau = 1}^pB(\tau)x(t - \tau), \sigma_v^2)\\
-%     (B_{ij}(\tau)\ |\ G_{ij}(\tau)) &\sim G_{ij}(\tau)\mathcal{N}(0, \sigma_\beta^2) + (1 - G_{ij}(\tau))\delta_0(B_{ij}(\tau))\\
-%     (G_{ij}(\tau)) &\sim \mathsf{BER}(p_{ij}(\tau))\\
-%     \sigma_v^2 &\sim \Gamma(a_v, b_v)\\
-%     \sigma_\beta^2 &\sim \Gamma(a_\beta, b_\beta)
-%   \end{aligned}
-% \end{equation}
-
-% Similar models have been studied by various authors in the context of the linear regression model \hl{[(cite them)]} where it is referred to as \textit{stochastic variable selection}, or referred to as a model for Bayesian variable selection.
-
-% Since we have in mind applications where $n$ is large, sampling posterior probabilities can be prohibitively burdensome, so we can instead fit the mean-field variational approximation $p(B, G, \sigma\ |\ X) \approx q_Gq_Bq_\sigma$ and subsequently apply a thresholding operation to the posterior edge inclusion probabilities under $q_G$.
-
-% \subsection{Pairwise Minimum Error Spanning Trees}
-% Inspired by theorem \ref{scg_recovery}, we propose the following
-% structure learning heuristic.
-
-% %% This algorithm should perform coordinate descent or jointly refit all the trees.
-
-% \begin{enumerate}
-%   \item{Set $k = 0$ and $x^0(t) = x(t)$}
-%   \item{Compute all pairwise causality measures $\Xi \defeq \big[\ln \frac{\xi_i}{\xi_{ij}} \big]_{i, j}$}
-%   \item{Find the maximum spanning arborescence\footnote{An ``arborescence'' is a french word for a tree diagram, and refers to a \textit{directed} tree in graph theory} $\mathcal{T}_k$ of a graph having edge weights $\Xi_{ij}$.}
-%   \item{Fit a vector LTI filter $F_k$ having graph defined by $\mathcal{T}_k$ to $x^{k - 1}(t)$ and set $x^k(t) = x^{k - 1}(t) - \hat{x}^{k - 1}(t)$}
-%   \item{Repeat over $k$ until a stopping criteria is met}
-% \end{enumerate}
-
-% \subsection{Layered Strongly Causal Graphs}
-% It is clearn that Algorithm \ref{alg:pwgr} is a purely theoretical construct, and cannot be applied directly to real data.
-Instead, we convert the algorithm into an heuristic and observe it's performance experimentally.
-We first consider the heuristic version of Algorithm \ref{alg:pwgr} in Algorithm \ref{alg:pwgr_heuristic} and show that is a well defined procedure (note that $\ceil{x}$ denotes the integer ceiling function):
-
-% \begin{proposition}
-%   Algorithm \ref{alg:pwgr_heuristic} terminates in at most $n$ iterations and returns a strongly causal graph $\widehat{\gcg}$.
-% \end{proposition}
-% \begin{proof}
-%   \hl{TODO}
-% \end{proof}
-
-% A complete estimation procedure for a sparse filter $\widehat{\B}(z)$
-% used to estimate $x(t + 1)$ from $\H_{t - 1}$ is then given in Algorithm \ref{alg:Bz_estimate}.
-
-% \begin{proposition}
-%   The error sequence in Algorithm \ref{alg:Bz_estimate} is a zero mean
-%   WSS process $\E\epsilon^{(m)}(t) = 0$ with monotonically decreasing
-%   variance: $\E|\epsilon^{(m + 1)}(t)|^2 \le \E|\epsilon^{(m)}(t)|^2$.
-% \end{proposition}
-% \begin{proof}
-%   \hl{TODO -- this is probably not even true}
-% \end{proof}
-
-% \begin{algorithm}
-%   \SetKwInOut{Input}{input}
-%   \SetKwInOut{Output}{output}
-%   \SetKwInOut{Initialize}{initialize}
-%   \DontPrintSemicolon
-
-%   \BlankLine
-%   \caption{Graph Recovery Heuristic $\mathsf{pw\_scg}(F, P, \alpha)$}
-%   \label{alg:pwgr_heuristic}
-%   \Input{Estimates of pairwise Granger-causaltiy statistics $F_{ij}$ (eqn \ref{eqn:gc_statistics}), Matrix of edge probabilities $P_{ij}$ (eqn \ref{eqn:edge_inclusion_probability}), threshold $\delta$.}
-%   \Output{A strongly causal graph $\widehat{\gcg}$}
-%   \Initialize{$S = [n]$  \texttt{\# unprocessed nodes}\\
-%     $E = \emptyset$  \texttt{\# edges of }$\widehat{\gcg}$\\
-%     $k = 1$ \texttt{\# a counter used only for notation}}
-
-%   \BlankLine
-
-%   $W_\delta \leftarrow \{(i, j)\ |\ P_{ji} > 1 - \delta, F_{ji} > F_{ij}\}$  \texttt{\# candidate edges}\\
-%   $\mathcal{I}_0 \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# total node incident probability}\\
-%   $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) < \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Nodes with fewest incident edges}\\
-%   \If{$P_0 = \emptyset$}{
-%     $P_0 \leftarrow \{i \in S\ |\ \mathcal{I}_0(i) \le \ceil{\text{min}(\mathcal{I}_0)}\}$ \texttt{\# Ensure non-empty}
-%   }
-%   \BlankLine
-
-%   \While{$S \ne \emptyset$}{
-%     $S \leftarrow S \setminus P_{k - 1}$ \texttt{\# remove processed nodes}\\
-%     % $\mathcal{I}_k \leftarrow \big(\sum_{j \in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# intra-}$S$ \texttt{incident strength}\\
-%     $\mathcal{I}_k \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$\\
-%     $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) < \ceil{\text{min}(\mathcal{I}_k)}\}$\\
-%     \If{$P_k = \emptyset$}{
-%       $P_k \leftarrow \{i \in S\ |\ \mathcal{I}_k(i) \le \ceil{\text{min}(\mathcal{I}_k)}\}$
-%     }
-%     \;
-%     \texttt{\# add strongest edges, maintaining strong causality}\\
-%     $U_k \leftarrow \bigcup_{r = 1}^k P_{k - r}$ \texttt{\# Include all forward edges}\\
-%     \For{$(i, j) \in \mathsf{sort}\Big(\{(i, j) \in U_k \times P_k\ |\ (i, j) \in W_\delta\} \mathsf{\ by\ descending\ } F_{ji}\Big)$} {
-%       \If{$\mathsf{is\_strongly\_causal}(E \cup \{(i, j)\})$} {
-%         $E \leftarrow E \cup \{(i, j)\}$
-%       }
-%     }
-%     $k \leftarrow k + 1$\\
-%   }
-%   \Return{$([n], E)$}
-% \end{algorithm}
-
-% \begin{algorithm}
-%   \SetKwInOut{Input}{input}
-%   \SetKwInOut{Output}{output}
-%   \SetKwInOut{Initialize}{initialize}
-%   \DontPrintSemicolon
-
-%   \BlankLine
-%   \caption{Filter Estimator}
-%   \label{alg:Bz_estimate}
-%   \Input{$T$ samples from a WSS process $x: \{x(t)\}_{t = 1}^T$\\
-%     Maximum iterations $M$\\
-%     Sparsity parameters $\{(\delta_m\, b_m, R_m)\}_{m = 1}^M$}
-%   \Output{Linear one-step-ahead estimator $\widehat{\B}(z)$}
-%   \Initialize{$\epsilon^{(0)}(t) = x(t)$ \texttt{\# initial error process}\\
-%     $m = 1$ \texttt{\# iteration counter}}
-%   \BlankLine
-%   \For{$m = 1, \ldots, M$}{
-%     $F \leftarrow \mathsf{pwgc}(\epsilon^{(m - 1)}(t))$ \texttt{\# estimate pairwise Granger-causality statistics}\\
-%     $\widehat{\gcg}_m \leftarrow \mathsf{pw\_scg}(F, \delta_m, b_m, R_m)$ \texttt{\# graph estimate (Algorithm \ref{alg:pwgr_heuristic})}\\
-%     $\widehat{\B}^{(m)}(z) \leftarrow \mathsf{estimate\_B}(\epsilon^{(m - 1)}(t), \widehat{\gcg}_m)$ \texttt{\# Estimate filter with topology} $\widehat{\gcg}_m$\\
-%     $\epsilon^{(m)}(t) \leftarrow \epsilon^{(m - 1)}(t) - \widehat{\B}^{(m)}(z)\epsilon^{(m - 1)}(t)$ \texttt{\# updated error process}\\
-%   }
-%   \BlankLine
-%   $\widehat{\B}(z) \leftarrow \widehat{\B}^{(M)}(z) \prod_{k = 1}^{M - 1}\big(I - \widehat{\B}^{(M - k)}(z)\big)$ \texttt{\# compute total filter}\\
-%   \Return{$\widehat{\B}(z)$}
-% \end{algorithm}
-
 \section{Empirical Evaluation}
@@
 is extremely sparse.
+
+\begin{table}
+  \centering
+  \caption{Simulation Results: PWGC vs AdaLASSO}
+  \label{taab:simulation_table}
+
+  \begin{tabular}{|ll||ll|ll|ll|}
+    \toprule
+    &{Algorithm}&alasso&pwgc&alasso&pwgc&alasso&pwgc\\
+    &\textbf{Metric}&Err&Err&FDR&FDR&MCC&MCC\\
+    \textbf{T}&\textbf{q}&&&&&&\\
+    \midrule
+    \multirow{4}{*}{\textbf{50}}
+    &\textbf{SCG}&1.83&\textbf{1.57}&0.48&\textbf{0.07}&0.47&\textbf{0.57}\\
+    &\textbf{0.06}&1.91&\textbf{1.69}&0.54&\textbf{0.09}&0.43&\textbf{0.53}\\
+    &\textbf{0.11}&2.89&\textbf{2.58}&0.46&\textbf{0.22}&0.38&0.40\\
+    &\textbf{0.46}&8.49&\textbf{7.72}&0.42&0.42&\textbf{0.16}&0.11\\
+    \midrule
+    \multirow{4}{*}{\textbf{250}}
+    &\textbf{SCG}&1.32&\textbf{1.22}&0.27&\textbf{0.07}&0.71&\textbf{0.80}\\
+    &\textbf{0.06}&1.44&\textbf{1.28}&0.32&\textbf{0.09}&0.66&\textbf{0.76}\\
+    &\textbf{0.11}&2.57&\textbf{2.17}&0.32&\textbf{0.16}&0.53&\textbf{0.57}\\
+    &\textbf{0.46}&8.09&\textbf{7.11}&0.39&0.40&\textbf{0.20}&0.14\\
+    \midrule
+    \multirow{4}{*}{\textbf{1250}}
+    &\textbf{SCG}&1.18&\textbf{1.09}&0.40&\textbf{0.06}&0.69&\textbf{0.89}\\
+    &\textbf{0.06}&1.24&1.17&0.43&\textbf{0.05}&0.65&\textbf{0.86}\\
+    &\textbf{0.11}&2.02&2.10&0.33&\textbf{0.14}&0.62&0.63\\
+    &\textbf{0.46}&7.44&\textbf{7.04}&0.40&\textbf{0.30}&\textbf{0.22}&0.19\\
+    \bottomrule
+  \end{tabular}
+
+  % \begin{tabular}{|ll||ll|ll|ll|}
+  %   \toprule
+  %   &\textbf{Algorithm}&alasso&pwgc&alasso&pwgc&alasso&pwgc\\
+  %   &\textbf{Metric}&Err&Err&FDR&FDR&MCC&MCC\\
+  %   \textbf{T}&\textbf{q}&&&&&&\\
+  %   \midrule
+  %   \multirow{4}{*}{\textbf{50}}
+  %   &\textbf{SCG}&1.86&\textbf{1.51}&0.49&\textbf{0.16}&0.47&\textbf{0.63}\\
+  %   &\textbf{0.06}&1.92&\textbf{1.67}&0.52&\textbf{0.17}&0.44&\textbf{0.60}\\
+  %   &\textbf{0.11}&3.06&\textbf{2.81}&0.47&\textbf{0.34}&0.37&\textbf{0.41}\\
+  %   &\textbf{0.46}&8.90&\textbf{7.74}&\textbf{0.42}&0.52&\textbf{0.16}&0.10\\
+  %   \cline{1-8}
+  %   \multirow{4}{*}{\textbf{250}}
+  %   &\textbf{SCG}&1.35&\textbf{1.20}&0.28&\textbf{0.21}&0.70&\textbf{0.77}\\
+  %   &\textbf{0.06}&1.40&\textbf{1.27}&0.33&\textbf{0.22}&0.66&\textbf{0.73}\\
+  %   &\textbf{0.11}&2.48&\textbf{2.18}&0.31&0.30&\textbf{0.55}&0.52\\
+  %   &\textbf{0.46}&7.96&\textbf{6.89}&0.38&0.38&\textbf{0.20}&0.15\\
+  %   \cline{1-8}
+  %   \multirow{4}{*}{\textbf{1250}}
+  %   &\textbf{SCG}&1.15&1.12&0.42&\textbf{0.24}&0.67&\textbf{0.79}\\
+  %   &\textbf{0.06}&1.28&1.24&0.45&\textbf{0.24}&0.64&\textbf{0.75}\\
+  %   &\textbf{0.11}&2.11&2.05&0.34&\textbf{0.29}&\textbf{0.60}&0.56\\
+  %   &\textbf{0.46}&7.12&\textbf{6.79}&0.39&\textbf{0.32}&\textbf{0.24}&0.18\\
+  %   \bottomrule
+  % \end{tabular}
+
+  % \begin{tabular}{|ll||ll|ll|ll|}
+  %   \toprule
+  %   &\textbf{Algorithm}&alasso&pwgc&alasso&pwgc&alasso&pwgc\\
+  %   &\textbf{metric}&RLMSE&&FDR&&MCC&\\
+  %   \textbf{T}&\textbf{q}&&&&&&\\
+  %   \midrule
+  %   \multirow{4}{*}{\textbf{50}}
+  %   &\textbf{SCG}&2.38&\textbf{1.56}&0.49&\textbf{0.16}&0.48&\textbf{0.63}\\
+  %   &\textbf{0.06}&2.66&\textbf{1.68}&0.52&\textbf{0.19}&0.45&\textbf{0.60}\\
+  %   &\textbf{0.11}&3.91&\textbf{2.68}&0.46&\textbf{0.32}&0.38&\textbf{0.42}\\
+  %   &\textbf{0.46}&10.24&\textbf{7.94}&\textbf{0.41}&0.49&\textbf{0.17}&0.11\\
+  %   \cline{1-8}
+  %   \multirow{4}{*}{\textbf{250}}
+  %   &\textbf{SCG}&2.14&\textbf{1.21}&0.30&\textbf{0.20}&0.69&\textbf{0.78}\\
+  %   &\textbf{0.06}&2.26&\textbf{1.29}&0.32&\textbf{0.24}&0.66&\textbf{0.73}\\
+  %   &\textbf{0.11}&3.60&\textbf{2.12}&0.31&0.29&0.55&0.53\\
+  %   &\textbf{0.46}&10.17&\textbf{7.27}&0.40&0.41&\textbf{0.19}&0.14\\
+  %   \cline{1-8}
+  %   \multirow{4}{*}{\textbf{1250}}
+  %   &\textbf{SCG}&1.88&\textbf{1.12}&0.39&\textbf{0.22}&0.70&\textbf{0.81}\\
+  %   &\textbf{0.06}&1.88&\textbf{1.21}&0.41&\textbf{0.25}&0.66&\textbf{0.75}\\
+  %   &\textbf{0.11}&3.24&\textbf{2.17}&0.34&\textbf{0.28}&\textbf{0.59}&0.56\\
+  %   &\textbf{0.46}&9.74&\textbf{6.93}&0.40&\textbf{0.35}&\textbf{0.22}&0.17\\
+  %   \bottomrule
+  % \end{tabular}
+
+  {\scriptsize Results of Monte Carlo simulations comparing PWGC and
+    AdaLASSO for small samples and when the SCG assumption doesn't
+    hold.  The superior result is bolded when the difference is
+    statistically significant, as measured by
+    \texttt{scipy.stats.ttest\_rel}.  100 iterations are run for each
+    set of parameters.}
+\end{table}
 
@@
 \label{sec:conclusion}
-In this paper we have argued that considering particular topological
-properties of Granger-causality networks can provide substantial
-theoretical insights as well as computational benefits.  In
-particular, the notion of a strongly-causal graph has been exploited
-to establish conditions under which pairwise causality testing alone
-is sufficient for recovering a complete Granger-causality graph.
-Moreover, examples from the literature suggest that such topological
-assumptions may be reasonable in some applications and secondly, even
-when the strong-causality assumption is not met, we have provided
-simulation evidence to suggest that our pairwise testing algorithm
-PWGC can still outperform the LASSO and adaLASSO, both of which are
-commonly employed in applications.  Application to inference of more
-general graphs may also be within reach via an iterative graph
-building process as suggested in Section \ref{sec:application}.
+In this paper we have applied the theoretical insights developed in
+\cite{my_GC_paper} into a heuristic for estimating Granger-causality
+graphs.  We have leveraged the notion of a strongly-causal graph which
+enables the recover of causality graphs through simple pairwise
+causality testing (Theorem \ref{thm:scg_recovery}), and demonstrated
+that substantial improvements (both statistically and computationally)
+over LASSO-based heuristics can be obtained.
+
+We emphasize that the improvements are a result of topological
+assumptions placed on the causality graph of the system, though
+examples (see Figure \ref{fig:gene_networks}) from the literature
+suggest that many graphs in practice may have special topological
+properties that can be exploited as prior knowledge.  Further research
+on the implications of Network properties (i.e. degree distributions,
+or small-world properties \todo{cite some network theory}) for
+Granger-causality is of interest.  Algorithms exploiting graphical
+assumptions may, for example, take the form of specially developed
+heuristics (e.g. Algorithm \ref{finite_pwgc}), cunning adaptive
+weighting schemes for the LASSO, or formal Bayesian priors.
 
 Our application in Section \ref{sec:application} provides strong
-evidence that PWGC is capable of uncovering meaningful features from
-networks of time series data by constructing a classifier which
-accurately discriminates between subjects in an EEG experiment based
-purely on the Granger-causality graph topology inferred by the
-iterated application of PWGC.
-
-We emphasize that the causality graph topology is one of the key
-defining features of time series analysis in comparison to standard
-multivariate regression and therefore advocate for further study of
-how different topological assumptions may impact the recovery of
-causality graphs.  For example, are there provable guarantees on the
-error rate of PWGC when applied to non strongly-causal graphs?
-Computationally, can constraint systems or cunning adaptive weighting
-schemes impose useful prior knowledge about graph topology for the
-LASSO algorithm?  In another direction, it is known that the more
-general notion of transfer entropy reduces to Granger-causality in the
-case of Gaussian data \cite{barnett2009granger}, do our results extend
-to causality networks defined via transfer entropy?  Finally, the work
-of \cite{barnett2015granger} has established the superiority of
-Granger-causality testing in state space models (as opposed to pure
-autoregressions) in many cases.  Combining this work with our PWGC
-algorithm (by modifying the approach described in Section
-\ref{sec:pairwise_hypothesis_testing} to instead utilize state-space
-Granger-causality testing) therefore is likely to enable application
-to very large networks of time series data which are not well
-approximated by finite $\VAR(p)$ models.
+evidence that PWGC (Algorithm \ref{finite_pwgc}) is capable of
+uncovering meaningful features from networks of time series data by
+constructing a classifier which accurately discriminates between
+subjects in an EEG experiment based purely on the Granger-causality
+graph topology inferred by the iterated application of PWGC.  This is
+a result which may be of independent scientific interest for EEG.
+
+There are a number of potential extensions to this work.  Firstly, it
+is known that the information theoretic notion of transfer entropy
+reduces to Granger-causality when the data is Gaussian
+\cite{barnett2009granger}, can Theorem \ref{thm:scg_recovery} be
+generalized in a useful way to causality networks defined by transfer
+entropy?  As well, the work of \cite{barnett2015granger} has
+established the superiority of Granger-causality testing in state
+space models (as opposed to pure autoregressions) in many cases.
+Since, our pairwise recovery results do not require $x(t)$ to be
+generated by a $\VAR$ model, such results can replace the Hypothesis
+tests of Section \ref{sec:pairwise_hypothesis_testing} to enable
+application to systems which are not well modelled by finite $\VAR$
+models.
+
+\clearpage
+\printbibliography
+\clearpage
+\appendix
+
+\section{Pairwise Causality Graph Estimation}
+This section details efficient pairwise causality testing methodology
+used in constructing the input data $F, P, \delta$ for Algorithm
+\ref{alg:finite_pwgc}.  We employ the simplest reasonable methods in
+order to maintain a focus on the Graph topological aspects.
+
+\subsection{Pairwise Hypothesis Testing}
+\label{sec:pairwise_hypothesis_testing}
+In performing pairwise checks for Granger-causality $x_j \pwgc x_i$ we
+follow the simple scheme of estimating the following two linear models:
+
+\begin{align}
+  H_0:&\ \widehat{x}_i^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau),\\
+  H_1:&\ \widehat{x}_i^{(p)}(t) = \sum_{\tau = 1}^{p} b_{ii}(\tau)x_i(t - \tau) + \sum_{\tau = 1}^pb_{ij}(\tau)x_j(t - \tau).
+\end{align}
+
+We formulate the statistic
+
+\begin{equation}
+  \label{eqn:gc_statistics}
+  F_{ij}(p) = \frac{T}{p}\Big(\frac{\xi_i(p)}{\xi_{ij}(p)} - 1\Big),
+\end{equation}
+
+where $\xi_i(p)$ is the sample mean square of the
+residuals\footnote{This quantity is often denoted $\widehat{\sigma}$,
+  but we maintain notation from Definition
+  \ref{def:granger_causality}.}  $x_i(t) - \widehat{x}^{(p)}_i(t)$,
+
+\begin{equation*}
+  \xi_i(p) = \frac{1}{T - p}\sum_{t = p + 1}^T (x_i(t) - \widehat{x}_i^{(p)}(t))^2,
+\end{equation*}
+
+and similarly for $\xi_{ij}(p)$.  We test $F_{ij}(p)$ against a
+$\chi^2(p)$ distribution.
+
+If the estimation procedure is consistent, we will have the following
+convergence (in $\P$ or a.s.):
+
+\begin{equation}
+  F_{ij}(p) \rightarrow
+  \left\{
+    \begin{array}{ll}
+      0;\ x_j \npwgc x_i\\
+      \infty;\ x_j \pwgc x_i
+    \end{array}
+  \right. \text{ as } T \rightarrow \infty.  % the '.' after \right is necessary.
+\end{equation}
+
+In our finite sample implementation (see Algorithm
+\ref{alg:finite_pwgc}) we add edges to $\widehat{\gcg}$ in order of
+the decreasing magnitude of $F_{ij}$ instead of proceeding backwards
+through $P_{k - r}$ in Algorithm \ref{alg:pwgr}.  This makes greater
+use of the information provided by the test statistic $F_{ij}$,
+moreover, if $x_i \gc x_j$ and $x_j \gc x_k$, it is expected that
+$F_{kj} > F_{ji}$, thereby providing the same effect as proceeding
+backwards through $P_{k - r}$.
+\subsection{Model Order Selection}
+\label{sec:model_order_selection}
+There are a variety of methods to choose the filter order $p$ (see
+e.g. \cite{lutkepohl2005new}), but we will focus in particular on the
+Bayesian Information Criteria (BIC).  The BIC is substantially more
+conservative than the popular alternative Akaiake Information Criteria
+(the BIC is also asymptotically consistent), and since we are
+searching for \textit{sparse graphs}, we therefore prefer the BIC,
+where we seek to \textit{minimize} over $p$:
+
+\begin{equation}
+  \label{eqn:bic}
+  \begin{aligned}
+    BIC_{\text{univariate}}(p) &= \ln\ \xi_i(p) + p\frac{\ln T}{T},\\
+    BIC_{\text{bivariate}}(p) &= \ln \det \widehat{\Sigma}_{ij}(p) + 4p\frac{\ln T}{T},\\
+  \end{aligned}
+\end{equation}
+
+where $\widehat{\Sigma}_{ij}(p)$ is the $2 \times 2$ residual
+covariance matrix for the $\VAR(p)$ model of $(x_i(t), x_j(t))$.  The
+bivariate errors $\xi_{ij}(p)$ and $\xi_{ji}(p)$ are the diagonal
+entries of $\widehat{\Sigma}_{ij}(p)$.
+
+We carry this out by a simple direct search on each model order
+between $0$ and some prescribed $p_\text{max}$, resulting in a
+collection $p_{ij}$ of model order estimates.  In practice, it is
+sufficient to pick $p_\text{max}$ ad-hoc or via some simple heuristic
+e.g. plotting the sequence $BIC(p)$ over $p$, though it is not
+technically possible to guarantee that the optimal $p$ is less than
+the chosen $p_\text{max}$ (since there can in general be arbitrarily
+long lags from one variable to another).
+
+\subsection{Efficient Model Estimation}
+\label{sec:efficient_model_estimation}
+In practice, the vast majority of computational effort involved in
+implementing our estimation algorithm is spent calculating the error
+estimates $\xi_i(p_i)$ and $\xi_{ij}(p_{ij})$.  This requires fitting a
+total of $n^2p_{\text{max}}$ autoregressive models, where the most
+naive algorithm (e.g. solving a least squares problem for each model)
+for this task will consume $O(n^2p_{\text{max}}^4T)$ time, it is
+possible to carry out this task in a much more modest
+$O(n^2p_{\text{max}}^2 ) + O(n^2p_{\text{max}}T)$ time via the
+autocorrelation method
+\cite{hayes_statistical_digital_signal_processing} which substitutes
+the following autocovariance estimates in the Yule-Walker
+equations:\footnote{The particular indexing and normalization given in
+  equation \ref{eqn:covariance_estimate} is critical to ensure
+  $\widehat{R}$ is positive semidefinite.  The estimate can be viewed
+  as calculating the covariance sequence of a signal multiplied by a
+  rectangular window.}
+
+\begin{equation}
+  \label{eqn:covariance_estimate}
+  \widehat{R}_x(\tau) = \frac{1}{T}\sum_{t = \tau + 1}^T x(t) x(t - \tau)^\T;\ \tau = 0, \ldots, p_{\text{max}},
+\end{equation}
+
+It is imperative that the first index in the summation is $\tau + 1$, as
+opposed perhaps to $p_\text{max}$ and that the normalization is
+$1 / T$, as opposed perhaps to $1 / (T - p_\text{max})$, in order to
+guarantee that $\widehat{R}_x(\tau)$ forms a valid (i.e. positive
+definite) covariance sequence.  This results in some bias, however the
+dramatic computational speedup is worth it for our purposes.
+
+These covariance estimates constitute the $O(n^2p_{\text{max}}T)$
+operation.  Given these particular estimates, the variances $\xi_i(p)$
+for $p = 1, \ldots, p_{\text{max}}$ can be evaluated in
+$O(p_{\text{max}}^2)$ time each by applying the Levinson-Durbin
+recursion to $\widehat{R}_{ii}(\tau)$, which effectively estimates a
+sequence of $AR$ models, producing $\xi_i(p)$ as a side-effect (see
+\cite{hayes_statistical_digital_signal_processing} and
+\cite{levinson_durbin_recursion}).
+
+Similarly, the variance estimates $\widehat{\Sigma}_{ij}(p)$ (which
+include $\xi_{ij}$ and $\xi_{ji}$) can be obtained by estimating
+$\frac{(n + 1)n}{2}$ bivariate AR models, again in
+$O(p_{\text{max}}^2)$ time via Whittle's generalized Levinson-Durbin
+recursion\footnote{We have made use of standalone tailor made
+  implementations of these algorithms, available at
+  \textsf{github.com/RJTK/Levinson-Durbin-Recursion}.}
+\cite{whittle_generalized_levinson_durbin}.
+
+\subsection{Edge Probabilities and Error Rate Controls}
+\label{sec:error_rate_control}
+Denote $F_{ij}$ the Granger-causality statistic of equation
+\ref{eqn:gc_statistics} with model orders chosen by the methods of
+Section \ref{sec:model_order_selection}.  We assume that this
+statistic is asymptotically $\chi^2(p_{ij})$ distributed (the
+disturbances are Gaussian), and denote by $G$ the cumulative
+distribution function thereof.  We will define the matrix
+
+\begin{equation}
+  \label{eqn:edge_inclusion_probability}
+  P_{ij} = G(F_{ij}),
+\end{equation}
+
+to be the matrix of pairwise edge inclusion P-values.  This is
+motivated by the hypothesis test where the hypothesis $H_0$ will be
+rejected (and thence we will conclude that $x_j \pwgc x_i$) if
+$P_{ij} > 1 - \delta$.
+
+The value $\delta$ can be chosen by a variety of methods, in our case
+we apply the Benjamini Hochberg criteria \cite{benjamini_hochberg}
+\cite{all_of_statistics} to control the false discovery rate of
+pairwise edges to a level $\alpha$ (where we generally take
+$\alpha = 0.05$).
+
+\subsection{Finite Sample Recovery Algorithm}
+\label{sec:finite_pwgc}
+
+After the graph topology $\widehat{\gcg}$ has been estimated via
+Algorithm \ref{alg:finite_pwgc}, we refit the entire model with the
+specified sparsity pattern directly via ordinary least squares.
+
+We note that producing graph estimates which are not strongly causal
+can potentially be achieved by performing sequential estimates
+$\widehat{x}_1(t), \widehat{x}_2(t), \ldots$ estimating a strongly causal
+graph with the residuals of the previous model as input, and then
+refitting on the combined sparsity pattern.  We experiment with this
+heuristic in our example application of Section \ref{sec:application},
+but reserve theoretical analysis for future work.
 

### 2019-07-26 paper_appl/paper_appl.tex
Methods to compute causality graphs by different methods on the eeg data.

@@
 is, there are algorithms which are immune to the effects of
-confounders when applied to certain causality graph structures.  In
-this paper we provide simulation evidence that, given structural
+confounders when applied to certain causality graph structures.  See
+\cite{tam2013gene_pwgc} for earlier work on pairwise Granger-causality.
+In this paper we provide simulation evidence that, given structural
 assumptions, such algorithms can dramatically outperform state of the
@@
 
-\subsection{Strongly Causal Graphs and Persistent Systems}
+\subsection{Exact Recovery From Pairwise Tests}
 \label{sec:strongly_causal_graphs}
 In this section we define the notion of a strongly causal graph, and
-the implications for pairwise causality testing.
+review the implications for pairwise causality testing.
 
@@
 
-  In order to exclude the possibility of certain types of cancellation
-  in the $\VAR$ system of Equation \eqref{eq:var_model}, we require
-  the system to have a property we have deemed ``persistence'' in
-  \cite{my_GC_paper}, the intuition being that each node in the system
-  maintains some ``memory''.
-
-  \begin{definition}[Persistent]
-    \hl{TODO:}I can provide the diagonalizability condition for
-    $\VAR(p)$ models, but what about when $p = \infty$?
-    \todo{persistence for $\VAR(\infty)$}
-  \end{definition}
-
-  The key property of persistent and strongly causal systems is that
-  they are in some sense immune to confounders.  That is, if $i \pwgc j$
-  \textit{and} $j \pwgc i$, then $i \ngc j$ and $j \ngc i$.  That is,
-  bidirectional pairwise causation (which would otherwise be the
-  result of hidden confounders) is not possible.
+  There is another technical condition (detailed in
+  \cite{my_GC_paper}) required for the next theorem which eliminates
+  the possibility of certain kinds of pathological cancellation when
+  confounding nodes are present.  However, $\VAR$ systems which do not
+  satisfy the technical condition are pathological.
 
@@
     \label{thm:scg_recovery}
-    If the Granger-causality graph $\gcg$ for persistent process
-    $x(t)$ is a strongly causal DAG, then $\gcg$ can be inferred from
+    Suppose that the Granger-causality graph $\gcg$ for the process
+    $x(t)$ is a strongly causal DAG.  Then, except for certain
+    pathological cases \cite{my_GC_paper}, $\gcg$ can be inferred from
     pairwise causality tests, given an oracle for pairwise causation.
-
-    \todo{Contain statement on convergence of the heuristic?}
-    % \todo{Make this precisise} Moreover, given $T < \infty$ samples from
-    % $x(t)$ then Algorithm \ref{alg:finite_pwgc} will correctly return
-    % $\gcg$ with probability $1$ as $T \rightarrow \infty$.
   \end{theorem}
+
+  The proof of this theorem proceeds by proving the correctness of an
+  algorithm for implementing exact recovery through pairwise testing.
+  This algorithm inspires the heuristic\footnote{Recall that
+    Granger-causality is necessarily based on heuristics when $n$ is
+    modestly large, as the complexity of searching over causality
+    graphs scales as $O(2^{n^2})$} of Section
+  \cite{sec:structure_learning}.
 
@@
 \label{sec:structure_learning}
-In \cite{my_GC_paper} we prove Theorem \ref{thm:scg_recovery} by
-providing an algorithm which, given an oracle for pairwise
-Granger-causality, will correctly recover $\gcg$.  In this section,
-we develop in detail appropriate statistical methods for
-implementing this algorithm on a finite data sample.  When applied
-in practice, the algorithm is a heuristic\footnote{Recall that
-  Granger-causality is necessarily based on heuristics when $n$ is
-  modestly large, as the complexity of searching over causality
-  graphs scales as $O(2^{n^2})$}, since we do not know apriori
-whether the causality graph of $x(t)$ is strongly causal (or if
-$x(t)$ is persistent).  However, we will see in Section
-\ref{sec:empirical_evaluation} that Algorithm \ref{alg:finite_pwgc}
-outperforms LASSO-based heuristics, even when $\gcg$ does not
-exactly satisfy the strong causality assumption.
+In this section we detail our proposed pairwise testing heuristics.
+In particular, Algorithm \ref{alg:finite_pwgc} proceeds first by a
+simple thresholding of pairwise relations.  The algorithm next
+proceeds by an iterative procedure which ``peels away layers'' of the
+causality graph eminating from parentless nodes, while maintaining the
+strong causality property.  We present simulation evidence that this
+heuristic can outperform LASSO-based heuristics in \cite{my_GC_paper}.
+% We will see in Section \ref{sec:empirical_evaluation} that Algorithm
+% \ref{alg:finite_pwgc} outperforms LASSO-based heuristics, even when
+% $\gcg$ does not exactly satisfy the strong causality assumption.
 
@@
 
-\section{Empirical Evaluation}
-\label{sec:empirical_evaluation}
-We have implemented our empirical experiments in Python \cite{scipy},
-in particular we leverage the LASSO implementation from
-\texttt{sklearn} \cite{sklearn} and the random graph generators from
-\texttt{networkx} \cite{networkx}.  We run experiments using two
-separate graph topologies having $n$ nodes: a strongly causal graph
-(SCG) and a directed acyclic graph (DAG).  These are generated
-respectively by drawing a random tree and a random Erdos Renyi graph
-(with edge probability $q = \frac{2}{n}$ resulting in approximately
-the same number of edges for the SCG as for the DAG), then creating a
-directed graph by directing edges from lower numbered nodes to higher
-numbered nodes.
-
-We populate each of the edges (including self loops) with random
-linear filters constructed by placing $5$ transfer function poles
-(i.e. $p = 5$) uniformly at random in a disc of radius $3 / 4$ (which
-guarantees stability for acyclic graphs).  The resulting system is
-driven by i.i.d. Gaussian random noise, each component having random
-variance $\sigma_i^2 = 1/2 + r_i$ where $r_i \sim \text{exp}(1/2)$.
-We set $p_{\text{max}} = 15$.  Results and representative graphs are
-collected in Figures \ref{fig:random_graph_topologies},
-\ref{fig:simulation_results_comparison1},
-\ref{fig:simulation_results_comparison2},
-\ref{fig:simulation_results_scaling_and_small_T},
-\ref{fig:simulation_results_dense}.
-
-We compare our results against the adaptive LASSO
-\cite{adaptive_lasso_zou2006}, which outperformed substantially both
-the LASSO and the grouped LASSO.  Motivated by scaling, we split the
-squared error term into separate terms, one for each group of incident
-edges on a node, and estimate the collection of $n$ incident filters
-$\big\{\B_{ij}(z)\big\}_{j = 1}^n$ that minimizes
-$\xi_i^{\text{LASSO}}$ in the following:
-
-\begin{equation}
-  \begin{aligned}
-  \xi_i^{\text{LASSO}}(\lambda) &= \underset{B}{\text{min}}\ \frac{1}{T}\sum_{t = p + 1}^T\big(x_i(t) - \sum_{\tau = 1}^p\sum_{j = 1}^n B_{i, j}(\tau) x(t - \tau)\big)^2 + \lambda \sum_{\tau = 1}^p \sum_{j = 1}^n |B_{ij}(\tau)|\\
-  \xi_i^{\text{LASSO}} &= \underset{\lambda \ge 0}{\text{min}}\ \xi_i^{\text{LASSO}}(\lambda) + \mathsf{BIC}\big(B_i^{\text{LASSO}}(\lambda)\big)\\
-  \end{aligned}
-\end{equation}
-
-where we are choosing $\lambda$, the regularization parameters, via the BIC.
-
-\begin{figure}
-  \centering
-  \caption{Representative Random Graph Topologies}
-  \label{fig:random_graph_topologies}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \caption{Random SCG}
-    \includegraphics[width=\linewidth]{example_scg.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \caption{Random DAG $(q = \frac{2}{n})$}
-    \includegraphics[width=\linewidth]{example_dag.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \caption{Random DAG $(q = \frac{4}{n})$}
-    \includegraphics[width=\linewidth]{example_dag_q.pdf}
-  \end{subfigure}
-\end{figure}
-
-\begin{remark}[Graph Topologies]
-  We depict in Figure \ref{fig:random_graph_topologies} the topologies
-  of random graphs used in our empirical evaluation.  For values of
-  $q$ close to $\frac{2}{n}$, the resulting random graphs tend to have
-  a topology which is, at least qualitatively, close to the SCG.  As
-  the value of $q$ increases, the random graphs deviate farther from
-  the SCG topology, and we therefore expect the LASSO to outperform
-  PWGC for larger values of $q$.  This can be observed in Figure
-  \ref{fig:simulation_results_dense}, at least where performance is
-  measured by the support recovery (i.e. via the MCC).
-\end{remark}
-
-\begin{remark}[MCC as a Support Recovery Measurement]
-  We apply Matthew's Correlation Coefficient (MCC)
-  \cite{matthews1975comparison} as a statistic for measuring support
-  recovery performance.  This statistic synthesizes the confusion
-  matrix into a single score appropriate for unbalanced labels and is
-  calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
-  performance, $0$ being the performance of random guessing, and $-1$
-  being perfectly opposed.
-\end{remark}
-
-\begin{remark}[Error Measurement]
-  We estimate the 1-step ahead prediction error by forming the variance matrix estimate
-
-  \begin{equation*}
-    \widehat{\Sigma}_v \defeq \frac{1}{T_{\text{out}}} \sum_{t = 1}^{T_{\text{out}}} (x(t) - \widehat{x}(t))(x(t) - \widehat{x}(t))^\T
-  \end{equation*}
-
-  on a long stream of out-of-sample data.  We then report the quantity
-
-  \begin{equation*}
-    \frac{\ln \tr \widehat{\Sigma}_v}{\ln \tr \Sigma_v}
-  \end{equation*}
-
-  where $\widehat{\Sigma}_v = \Sigma_v$ is the best possible performance.
-\end{remark}
-
-\begin{figure}
-  \centering
-  \caption{PWGC Compared Against AdaLASSO \cite{adaptive_lasso_zou2006} (SCG)}
-  \label{fig:simulation_results_comparison1}
-  \includegraphics[width=0.95\textwidth]{new_lasso_comparison_scg_pmax15_simulation.pdf}
-
-  {\scriptsize Comparison of PWGC and LASSO for $\VAR(p)$ model
-    estimation.  We make comparisons against both the MCC and the
-    relative log mean-squared prediction error
-    $\frac{\ln\tr \widehat{\Sigma}_v}{\ln\tr \Sigma_v}$.  Results
-    in Figure \ref{fig:simulation_results_comparison1} are for systems
-    guaranteed to satisfy the assumptions required for Theorem
-    \ref{thm:scg_recovery}.}
-\end{figure}
-
-\begin{figure}
-  \caption{PWGC vs adaLASSO (DAG, $q = \frac{2}{n}$)}
-  \label{fig:simulation_results_comparison2}
-  \includegraphics[width=0.95\textwidth]{new_lasso_comparison_dag_pmax15_simulation.pdf}
-
-  {\scriptsize Figure \ref{fig:simulation_results_comparison2}
-    provides results for systems which do not guarantee the
-    assumptions of Theorem \ref{thm:scg_recovery}, though the graph
-    has a similar level of sparsity.}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \caption{PWGC Scaling and Small Sample Performance}
-  \label{fig:simulation_results_scaling_and_small_T}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{Fixed $T$, increasing $n$ (SCG)}
-    \label{fig:simulation_results_scaling}
-    \includegraphics[width=\linewidth]{new_increasing_n_simulation.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \caption{MCC Comparison for $T \le 100$}
-    \label{fig:small_T_comparison}
-    \includegraphics[width=\linewidth]{new_mcc_comparison001.pdf}
-  \end{subfigure}
-
-  {\scriptsize Figure \ref{fig:simulation_results_scaling} measures
-    support recovery performance as the number of nodes $n$ increases,
-    and the edge proportion as well as the number of samples $T$ is
-    held fixed.  Remarkably, the degradation as $n$ increases is
-    limited, it is primarily the graph topology (SCG or non-SCG) as
-    well as the level of sparsity (measured by $q$) which are the
-    determining factors for support recovery performance.
-
-    Figure \ref{fig:small_T_comparison} provides a support recovery
-    comparison for very small values of, $T$ typical for many
-    applications.}
-\end{figure}
-
-\begin{figure}
-  \caption{Fixed $T, n$, increasing edges $q$ (DAG)}
-  \label{fig:simulation_results_dense}
-  \includegraphics[width=\linewidth]{dag_increasing_q_small_T_alasso_simulation.pdf}
-
-  {\scriptsize Figure \ref{fig:simulation_results_dense} provides a
-    comparison between PWGC and AdaLASSO as the density of graph edges
-    (as measured by $q$) increases.  For reference,
-    $\frac{2}{n} = 0.04$ has approximately the same level of sparsity
-    as the SCGs we simulated.  As $q$ increases, the AdaLASSO
-    outperforms PWGC as measured by the MCC.  However, PWGC maintains
-    superior performance for 1-step-ahead prediction.  We speculate
-    that this is a result of fitting the sparsity pattern recovered by
-    PWGC via OLS which directly seeks to optimize this metric, whereas
-    the LASSO is encumbered by the sparsity inducing penalty.}
-\end{figure}
-
-In reference to figure \ref{fig:simulation_results_comparison1} it
-should not be overly surprising that our PWGC algorithm performs
-better than the LASSO for the case of a strongly causal graph, since
-in this case the assumptions which guarantee the correctness of
-Algorithm \ref{alg:finite_pwgc} hold.  However, the performance is
-still markedly superior in the case of a more general DAG.  We would
-conjecture that a DAG having a similar degree of sparsity as an SCG is
-likely to be ``close'' to an SCG.  Figure
-\ref{fig:simulation_results_dense} illustrates the severe (expected)
-degradation in performance as the number of edges increases while
-the number of data samples $T$ remains fixed.  For larger values $q$
-in this plot, the number of edges in the graph is comparable to the
-number of data samples.
-
-We have also paid close attention to the performance of PWGC in the
-very small sample ($T \le 100$) regime (see Figure
-\ref{fig:small_T_comparison}), as this is the regime many applications
-must contend with.
-
-In regards scalability, we have observed that performing the $O(n^2)$
-pairwise Granger-causality calculations consumes the vast majority
-($> 90\%$) of the computation time.  Since this step is trivially
-parallelizable, our algorithm also scales well with multiple cores or
-multiple machines.  Figure \ref{fig:simulation_results_scaling} is a
-demonstration of this scalability, where we are able to estimate
-graphs having over $1500$ nodes (over $2.25 \times 10 ^6$ possible edges)
-using only $T = 500$ data points, granted, an SCG on this many nodes
-is extremely sparse.
-
-\begin{table}
-  \centering
-  \caption{Simulation Results: PWGC vs AdaLASSO}
-  \label{taab:simulation_table}
-
-  \begin{tabular}{|ll||ll|ll|ll|}
-    \toprule
-    &{Algorithm}&alasso&pwgc&alasso&pwgc&alasso&pwgc\\
-    &\textbf{Metric}&Err&Err&FDR&FDR&MCC&MCC\\
-    \textbf{T}&\textbf{q}&&&&&&\\
-    \midrule
-    \multirow{4}{*}{\textbf{50}}
-    &\textbf{SCG}&1.83&\textbf{1.57}&0.48&\textbf{0.07}&0.47&\textbf{0.57}\\
-    &\textbf{0.06}&1.91&\textbf{1.69}&0.54&\textbf{0.09}&0.43&\textbf{0.53}\\
-    &\textbf{0.11}&2.89&\textbf{2.58}&0.46&\textbf{0.22}&0.38&0.40\\
-    &\textbf{0.46}&8.49&\textbf{7.72}&0.42&0.42&\textbf{0.16}&0.11\\
-    \midrule
-    \multirow{4}{*}{\textbf{250}}
-    &\textbf{SCG}&1.32&\textbf{1.22}&0.27&\textbf{0.07}&0.71&\textbf{0.80}\\
-    &\textbf{0.06}&1.44&\textbf{1.28}&0.32&\textbf{0.09}&0.66&\textbf{0.76}\\
-    &\textbf{0.11}&2.57&\textbf{2.17}&0.32&\textbf{0.16}&0.53&\textbf{0.57}\\
-    &\textbf{0.46}&8.09&\textbf{7.11}&0.39&0.40&\textbf{0.20}&0.14\\
-    \midrule
-    \multirow{4}{*}{\textbf{1250}}
-    &\textbf{SCG}&1.18&\textbf{1.09}&0.40&\textbf{0.06}&0.69&\textbf{0.89}\\
-    &\textbf{0.06}&1.24&1.17&0.43&\textbf{0.05}&0.65&\textbf{0.86}\\
-    &\textbf{0.11}&2.02&2.10&0.33&\textbf{0.14}&0.62&0.63\\
-    &\textbf{0.46}&7.44&\textbf{7.04}&0.40&\textbf{0.30}&\textbf{0.22}&0.19\\
-    \bottomrule
-  \end{tabular}
+In order to adapt Algorithm \ref{alg:finite_pwgc} to graphs which may
+not be strongly causal, we consider an iterative application as
+described in Algorithm \ref{alg:iterated_pwgc}.  We apply Algorithm
+\ref{alg:finite_pwgc} first to the data itself, and then subsequently
+to the resulting residuals before performing a final estimate on the
+combined graph.  We evaluate these algorithms, as well as simpler
+pairwise heuristics and LASSO-type heuristics in the next section.  A
+theoretical analysis of Algorithm \ref{alg:iterated_pwgc} is deferred
+to future work.
+
+\begin{algorithm}
+  \SetKwInOut{Input}{input}
+  \SetKwInOut{Output}{output}
+  \SetKwInOut{Initialize}{initialize}
+  \DontPrintSemicolon
+
+  \BlankLine
+  \caption{Iterated PWGC Heuristic}
+  \label{alg:iterated_pwgc}
+
+  \Input{Data $x(t)$, maximum number of iterations $N$,
+    threshold $r \in (0, 1)$.}
+  \Output{Granger-causality graph estimate $\widehat{G}$ obtaind from
+    iterated application of PWGC Algorithm \ref{alg:finite_pwgc}}
+  \Initialize{$i = 1$, $\widehat{x}_0(t) \leftarrow x(t)$,
+    $\sigma_0 \leftarrow Var\ x(t)$}
+
+  \While{$i \le N$} {
+    Estimate $\VAR$ model via PWGC on $x_{i - 1}(t)$ to obtain $\widehat{x}_{i - 1}(t)$ and $\widehat{G}_{i - 1}$\\
+    $\epsilon_i(t) \leftarrow x_{i - 1}(t) - \widehat{x}_{i - 1}(t)$ \texttt{\# Compute residuals}\\
+    $\sigma_i^2 \leftarrow \text{MSE}\big(\epsilon_i(t)\big)$ \texttt{\# Compute MSE}\\
+    \If{$\sigma_i^2 > r \sigma_{i - 1}^2$}{
+      \texttt{break  \# Insignificant Improvement}
+    }
+    $i \leftarrow i + 1$
+  }
+  $\widehat{G} \leftarrow \bigcup_{i} \widehat{G}_{i}$ \texttt{\# Combine PWGC graph estimates}\\
+  \Return{$\widehat{G}$}
+\end{algorithm}
+
+% \section{Empirical Evaluation}
+% \label{sec:empirical_evaluation}
+Our implementations are in Python \cite{scipy}, in particular we
+leverage the LASSO implementation from \texttt{sklearn} \cite{sklearn}
+and the random graph generators from \texttt{networkx}
+\cite{networkx}.  Our paper \cite{my_GC_paper} provides simulation
+evidence that our PWGC heuristic can outperform LASSO-based heuristics
+for simulated data.
+
+% We run experiments using two separate graph topologies having $n$
+% nodes: a strongly causal graph (SCG) and a directed acyclic graph
+% (DAG).  These are generated respectively by drawing a random tree
+% and a random Erdos Renyi graph (with edge probability
+% $q = \frac{2}{n}$ resulting in approximately the same number of
+% edges for the SCG as for the DAG), then creating a directed graph by
+% directing edges from lower numbered nodes to higher numbered nodes.
+
+% We populate each of the edges (including self loops) with random
+% linear filters constructed by placing $5$ transfer function poles
+% (i.e. $p = 5$) uniformly at random in a disc of radius $3 / 4$ (which
+% guarantees stability for acyclic graphs).  The resulting system is
+% driven by i.i.d. Gaussian random noise, each component having random
+% variance $\sigma_i^2 = 1/2 + r_i$ where $r_i \sim \text{exp}(1/2)$.
+% We set $p_{\text{max}} = 15$.  Results and representative graphs are
+% collected in Figures \ref{fig:random_graph_topologies},
+% \ref{fig:simulation_results_comparison1},
+% \ref{fig:simulation_results_comparison2},
+% \ref{fig:simulation_results_scaling_and_small_T},
+% \ref{fig:simulation_results_dense}.
+
+% We compare our results against the adaptive LASSO
+% \cite{adaptive_lasso_zou2006}, which outperformed substantially both
+% the LASSO and the grouped LASSO.  Motivated by scaling, we split the
+% squared error term into separate terms, one for each group of incident
+% edges on a node, and estimate the collection of $n$ incident filters
+% $\big\{\B_{ij}(z)\big\}_{j = 1}^n$ that minimizes
+% $\xi_i^{\text{LASSO}}$ in the following:
+
+% \begin{equation}
+%   \begin{aligned}
+%   \xi_i^{\text{LASSO}}(\lambda) &= \underset{B}{\text{min}}\ \frac{1}{T}\sum_{t = p + 1}^T\big(x_i(t) - \sum_{\tau = 1}^p\sum_{j = 1}^n B_{i, j}(\tau) x(t - \tau)\big)^2 + \lambda \sum_{\tau = 1}^p \sum_{j = 1}^n |B_{ij}(\tau)|\\
+%   \xi_i^{\text{LASSO}} &= \underset{\lambda \ge 0}{\text{min}}\ \xi_i^{\text{LASSO}}(\lambda) + \mathsf{BIC}\big(B_i^{\text{LASSO}}(\lambda)\big)\\
+%   \end{aligned}
+% \end{equation}
+
+% where we are choosing $\lambda$, the regularization parameters, via the BIC.
+
+% \begin{figure}
+%   \centering
+%   \caption{Representative Random Graph Topologies}
+%   \label{fig:random_graph_topologies}
+%   \begin{subfigure}[b]{0.3\textwidth}
+%     \caption{Random SCG}
+%     \includegraphics[width=\linewidth]{example_scg.pdf}
+%   \end{subfigure}
+%   \begin{subfigure}[b]{0.3\textwidth}
+%     \caption{Random DAG $(q = \frac{2}{n})$}
+%     \includegraphics[width=\linewidth]{example_dag.pdf}
+%   \end{subfigure}
+%   \begin{subfigure}[b]{0.3\textwidth}
+%     \caption{Random DAG $(q = \frac{4}{n})$}
+%     \includegraphics[width=\linewidth]{example_dag_q.pdf}
+%   \end{subfigure}
+% \end{figure}
+
+% \begin{remark}[Graph Topologies]
+%   We depict in Figure \ref{fig:random_graph_topologies} the topologies
+%   of random graphs used in our empirical evaluation.  For values of
+%   $q$ close to $\frac{2}{n}$, the resulting random graphs tend to have
+%   a topology which is, at least qualitatively, close to the SCG.  As
+%   the value of $q$ increases, the random graphs deviate farther from
+%   the SCG topology, and we therefore expect the LASSO to outperform
+%   PWGC for larger values of $q$.  This can be observed in Figure
+%   \ref{fig:simulation_results_dense}, at least where performance is
+%   measured by the support recovery (i.e. via the MCC).
+% \end{remark}
+
+% \begin{remark}[MCC as a Support Recovery Measurement]
+%   We apply Matthew's Correlation Coefficient (MCC)
+%   \cite{matthews1975comparison} as a statistic for measuring support
+%   recovery performance.  This statistic synthesizes the confusion
+%   matrix into a single score appropriate for unbalanced labels and is
+%   calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
+%   performance, $0$ being the performance of random guessing, and $-1$
+%   being perfectly opposed.
+% \end{remark}
+
+% \begin{remark}[Error Measurement]
+%   We estimate the 1-step ahead prediction error by forming the variance matrix estimate
+
+%   \begin{equation*}
+%     \widehat{\Sigma}_v \defeq \frac{1}{T_{\text{out}}} \sum_{t = 1}^{T_{\text{out}}} (x(t) - \widehat{x}(t))(x(t) - \widehat{x}(t))^\T
+%   \end{equation*}
+
+%   on a long stream of out-of-sample data.  We then report the quantity
+
+%   \begin{equation*}
+%     \frac{\ln \tr \widehat{\Sigma}_v}{\ln \tr \Sigma_v}
+%   \end{equation*}
+
+%   where $\widehat{\Sigma}_v = \Sigma_v$ is the best possible performance.
+% \end{remark}
+
+% \begin{figure}
+%   \centering
+%   \caption{PWGC Compared Against AdaLASSO \cite{adaptive_lasso_zou2006} (SCG)}
+%   \label{fig:simulation_results_comparison1}
+%   \includegraphics[width=0.95\textwidth]{new_lasso_comparison_scg_pmax15_simulation.pdf}
+
+%   {\scriptsize Comparison of PWGC and LASSO for $\VAR(p)$ model
+%     estimation.  We make comparisons against both the MCC and the
+%     relative log mean-squared prediction error
+%     $\frac{\ln\tr \widehat{\Sigma}_v}{\ln\tr \Sigma_v}$.  Results
+%     in Figure \ref{fig:simulation_results_comparison1} are for systems
+%     guaranteed to satisfy the assumptions required for Theorem
+%     \ref{thm:scg_recovery}.}
+% \end{figure}
+
+% \begin{figure}
+%   \caption{PWGC vs adaLASSO (DAG, $q = \frac{2}{n}$)}
+%   \label{fig:simulation_results_comparison2}
+%   \includegraphics[width=0.95\textwidth]{new_lasso_comparison_dag_pmax15_simulation.pdf}
+
+%   {\scriptsize Figure \ref{fig:simulation_results_comparison2}
+%     provides results for systems which do not guarantee the
+%     assumptions of Theorem \ref{thm:scg_recovery}, though the graph
+%     has a similar level of sparsity.}
+% \end{figure}
+
+% \begin{figure}
+%   \centering
+%   \caption{PWGC Scaling and Small Sample Performance}
+%   \label{fig:simulation_results_scaling_and_small_T}
+%   \begin{subfigure}[b]{0.45\textwidth}
+%     \caption{Fixed $T$, increasing $n$ (SCG)}
+%     \label{fig:simulation_results_scaling}
+%     \includegraphics[width=\linewidth]{new_increasing_n_simulation.pdf}
+%   \end{subfigure}
+%   \begin{subfigure}[b]{0.45\textwidth}
+%     \caption{MCC Comparison for $T \le 100$}
+%     \label{fig:small_T_comparison}
+%     \includegraphics[width=\linewidth]{new_mcc_comparison001.pdf}
+%   \end{subfigure}
+
+%   {\scriptsize Figure \ref{fig:simulation_results_scaling} measures
+%     support recovery performance as the number of nodes $n$ increases,
+%     and the edge proportion as well as the number of samples $T$ is
+%     held fixed.  Remarkably, the degradation as $n$ increases is
+%     limited, it is primarily the graph topology (SCG or non-SCG) as
+%     well as the level of sparsity (measured by $q$) which are the
+%     determining factors for support recovery performance.
+
+%     Figure \ref{fig:small_T_comparison} provides a support recovery
+%     comparison for very small values of, $T$ typical for many
+%     applications.}
+% \end{figure}
+
+% \begin{figure}
+%   \caption{Fixed $T, n$, increasing edges $q$ (DAG)}
+%   \label{fig:simulation_results_dense}
+%   \includegraphics[width=\linewidth]{dag_increasing_q_small_T_alasso_simulation.pdf}
+
+%   {\scriptsize Figure \ref{fig:simulation_results_dense} provides a
+%     comparison between PWGC and AdaLASSO as the density of graph edges
+%     (as measured by $q$) increases.  For reference,
+%     $\frac{2}{n} = 0.04$ has approximately the same level of sparsity
+%     as the SCGs we simulated.  As $q$ increases, the AdaLASSO
+%     outperforms PWGC as measured by the MCC.  However, PWGC maintains
+%     superior performance for 1-step-ahead prediction.  We speculate
+%     that this is a result of fitting the sparsity pattern recovered by
+%     PWGC via OLS which directly seeks to optimize this metric, whereas
+%     the LASSO is encumbered by the sparsity inducing penalty.}
+% \end{figure}
+
+% In reference to figure \ref{fig:simulation_results_comparison1} it
+% should not be overly surprising that our PWGC algorithm performs
+% better than the LASSO for the case of a strongly causal graph, since
+% in this case the assumptions which guarantee the correctness of
+% Algorithm \ref{alg:finite_pwgc} hold.  However, the performance is
+% still markedly superior in the case of a more general DAG.  We would
+% conjecture that a DAG having a similar degree of sparsity as an SCG is
+% likely to be ``close'' to an SCG.  Figure
+% \ref{fig:simulation_results_dense} illustrates the severe (expected)
+% degradation in performance as the number of edges increases while
+% the number of data samples $T$ remains fixed.  For larger values $q$
+% in this plot, the number of edges in the graph is comparable to the
+% number of data samples.
+
+% We have also paid close attention to the performance of PWGC in the
+% very small sample ($T \le 100$) regime (see Figure
+% \ref{fig:small_T_comparison}), as this is the regime many applications
+% must contend with.
+
+% In regards scalability, we have observed that performing the $O(n^2)$
+% pairwise Granger-causality calculations consumes the vast majority
+% ($> 90\%$) of the computation time.  Since this step is trivially
+% parallelizable, our algorithm also scales well with multiple cores or
+% multiple machines.  Figure \ref{fig:simulation_results_scaling} is a
+% demonstration of this scalability, where we are able to estimate
+% graphs having over $1500$ nodes (over $2.25 \times 10 ^6$ possible edges)
+% using only $T = 500$ data points, granted, an SCG on this many nodes
+% is extremely sparse.
+
+% \begin{table}
+%   \centering
+%   \caption{Simulation Results: PWGC vs AdaLASSO}
+%   \label{taab:simulation_table}
+
+%   \begin{tabular}{|ll||ll|ll|ll|}
+%     \toprule
+%     &{Algorithm}&alasso&pwgc&alasso&pwgc&alasso&pwgc\\
+%     &\textbf{Metric}&Err&Err&FDR&FDR&MCC&MCC\\
+%     \textbf{T}&\textbf{q}&&&&&&\\
+%     \midrule
+%     \multirow{4}{*}{\textbf{50}}
+%     &\textbf{SCG}&1.83&\textbf{1.57}&0.48&\textbf{0.07}&0.47&\textbf{0.57}\\
+%     &\textbf{0.06}&1.91&\textbf{1.69}&0.54&\textbf{0.09}&0.43&\textbf{0.53}\\
+%     &\textbf{0.11}&2.89&\textbf{2.58}&0.46&\textbf{0.22}&0.38&0.40\\
+%     &\textbf{0.46}&8.49&\textbf{7.72}&0.42&0.42&\textbf{0.16}&0.11\\
+%     \midrule
+%     \multirow{4}{*}{\textbf{250}}
+%     &\textbf{SCG}&1.32&\textbf{1.22}&0.27&\textbf{0.07}&0.71&\textbf{0.80}\\
+%     &\textbf{0.06}&1.44&\textbf{1.28}&0.32&\textbf{0.09}&0.66&\textbf{0.76}\\
+%     &\textbf{0.11}&2.57&\textbf{2.17}&0.32&\textbf{0.16}&0.53&\textbf{0.57}\\
+%     &\textbf{0.46}&8.09&\textbf{7.11}&0.39&0.40&\textbf{0.20}&0.14\\
+%     \midrule
+%     \multirow{4}{*}{\textbf{1250}}
+%     &\textbf{SCG}&1.18&\textbf{1.09}&0.40&\textbf{0.06}&0.69&\textbf{0.89}\\
+%     &\textbf{0.06}&1.24&1.17&0.43&\textbf{0.05}&0.65&\textbf{0.86}\\
+%     &\textbf{0.11}&2.02&2.10&0.33&\textbf{0.14}&0.62&0.63\\
+%     &\textbf{0.46}&7.44&\textbf{7.04}&0.40&\textbf{0.30}&\textbf{0.22}&0.19\\
+%     \bottomrule
+%   \end{tabular}
 
@@
 
-  {\scriptsize Results of Monte Carlo simulations comparing PWGC and
-    AdaLASSO for small samples and when the SCG assumption doesn't
-    hold.  The superior result is bolded when the difference is
-    statistically significant, as measured by
-    \texttt{scipy.stats.ttest\_rel}.  100 iterations are run for each
-    set of parameters.}
-\end{table}
+%   {\scriptsize Results of Monte Carlo simulations comparing PWGC and
+%     AdaLASSO for small samples and when the SCG assumption doesn't
+%     hold.  The superior result is bolded when the difference is
+%     statistically significant, as measured by
+%     \texttt{scipy.stats.ttest\_rel}.  100 iterations are run for each
+%     set of parameters.}
+% \end{table}
 
@@
 \label{sec:application}
-In this section we apply our methods to a real set of EEG data
-obtained from the ``EEG Database Data
+In this section we apply our methods, as well as a number of
+previously known methods, to a real set of EEG data obtained from the
+``EEG Database Data
 Set''\footnote{http://archive.ics.uci.edu/ml/datasets/EEG+Database}
@@
 (64 channel) EEG signals from patients who are given visual stimuli.
+
+More precisely, our dataset is described by
+$\mathcal{D} = \{\big(x^{(i)}(t)\big)_{t = 1}^T, y^{(i)} \}_{i = 1}^N$
+where $T = 256$, $x^{(i)}(t) \in \R^{64}$,
+$y^{(i)} \in [N_{subjects}]$ (with $N_{subjects} = 119$), and
+$N = 10723$.  There are on average 90 trials for each subject, ranging
+between 30 and 119.
+
+We have constructed a simple pipeline for discriminating between
+subjects by first applying a Granger-causality algorithm directly to
+the EEG data $x^{(i)}$ to obtain a Granger-causality graph
+$\widehat{G}^{(i)}$.  We then feed the vectorized adjacency matrix of
+$\widehat{G}^{(i)}$ through a polynomial Kernel multinomial logistic
+regression model with parameters fit by cross validation.
+
+\subsection{Explanation}
 The subjects in the study are labeled as being either ``control'' or
@@
 
-The dataset consists of
-$\mathcal{D} = \{\big(x^{(i)}(t)\big)_{t = 1}^T, y^{(i)} \}_{i = 1}^N$
-where $T = 256$, $x^{(i)}(t) \in \R^{64}$,
-$y^{(i)} \in [N_{subjects}]$ (with $N_{subjects} = 119$), and
-$N = 10723$.  There are on average 90 trials for each subject,
-ranging between 30 and 119.
-
-We have constructed a simple pipeline for discriminating between
-subjects by first applying an iterative PWGC algorithm (see Algorithm
-\ref{alg:iterated_pwgc}) directly to the EEG data $x^{(i)}$ to obtain
-a Granger-causality graph $\widehat{G}^{(i)}$.  We then feed the
-vectorized adjacency matrix of $\widehat{G}^{(i)}$ through a
-polynomial Kernel multinomial logistic regression model with
-parameters fit by cross validation.
-
-\begin{remark}
-  The iterated application of PWGC is an example of a heuristic by
-  which graph estimates that aren't constrained to being strongly
-  causal can be obtained.  We defer to future work the theoretical
-  analysis of more sophisticated heuristics, or algorithms appropriate
-  for different topological assumptions.
-\end{remark}
-
-\begin{algorithm}
-  \SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{output}
-  \SetKwInOut{Initialize}{initialize}
-  \DontPrintSemicolon
-
-  \BlankLine
-  \caption{Iterated PWGC Heuristic}
-  \label{alg:iterated_pwgc}
-
-  \Input{Data $x(t)$, maximum number of iterations $N$,
-    threshold $r \in (0, 1)$.}
-  \Output{Granger-causality graph estimate $\widehat{G}$ obtaind from
-    iterated application of PWGC Algorithm \ref{alg:finite_pwgc}}
-  \Initialize{$i = 1$, $\widehat{x}_0(t) \leftarrow x(t)$,
-    $\sigma_0 \leftarrow Var\ x(t)$}
-
-  \While{$i \le N$} {
-    Estimate $\VAR$ model via PWGC on $x_{i - 1}(t)$ to obtain $\widehat{x}_{i - 1}(t)$ and $\widehat{G}_{i - 1}$\\
-    $\epsilon_i(t) \leftarrow x_{i - 1}(t) - \widehat{x}_{i - 1}(t)$ \texttt{\# Compute residuals}\\
-    $\sigma_i^2 \leftarrow \text{MSE}\big(\epsilon_i(t)\big)$ \texttt{\# Compute MSE}\\
-    \If{$\sigma_i^2 > r \sigma_{i - 1}^2$}{
-      \texttt{break  \# Insignificant Improvement}
-    }
-    $i \leftarrow i + 1$
-  }
-  $\widehat{G} \leftarrow \bigcup_{i} \widehat{G}_{i}$ \texttt{\# Combine PWGC graph estimates}\\
-  \Return{$\widehat{G}$}
-\end{algorithm}
+We will provide evidence in the affirmative for the question posed.
+Moreover, we will compare the classification accuracy resulting from
+training the same classifier on the causality graphs estimated by our
+PWGC algorithm, as well as graphs estimated by previously studied
+methods.
+
+Since Granger-causality graphs are latent structures, it is difficult
+to access the quality of graph recovery algorithms when applied to
+real-world data, since the ground truth causality graph cannot be
+known.  Our reasoning for comparing classification accuracy is the
+following: if causality graph estimates are provided by two methods,
+say $M_1$ and $M_2$, it stands to reason that if a classifier trained
+on $M_1$ significantly outerpforms a classifier trained on $M_2$, then
+the latent structure inferred from $M_1$ is more effectively capturing
+the nature of the ground truth causality graph.
 
@@
 classification accuracy on this dataset by constructing a classifier
-to act directly on $x^{(i)}(t)$, our purpose is to demonstrate that
-relevant information is being captured by the graph estimates
-$\widehat{G}^{(i)}$.  That is, that the latent structure
-$\widehat{G}^{(i)}$ may provide scientifically relevant insights, as
-opposed to simply being a feature for classification tasks.
-
+to act directly on $x^{(i)}(t)$, we stress that our purpose is to
+demonstrate that relevant information is being captured by the graph
+estimates $\widehat{G}^{(i)}$.  That is, evidence that the latent
+structure $\widehat{G}^{(i)}$ contains scientifically relevant
+insights, as opposed to simply being a feature for classification
+tasks.
+
+\subsection{Preliminary Analysis}
 Figures \ref{fig:ar_eeg_example} and \ref{fig:eeg_oos_error}
@@
 autoregressiive models.  The conclusion being that the
-Granger-causality graph estimates are not purely spurious.
+Granger-causality graph estimates are not purely spurious.  If there
+were no gains in prediction power resulting from the application of a
+$\VAR$ model, as opposed to a collection of unidimensional
+$\mathsf{AR}$ models, it would be difficult to justify the use of
+linear Granger-causality for EEG data at all.
 
@@
 
+\subsection{Results}
 We provide the final results in Figure
@@
     classify subjects based on their EEG Granger-causality graphs.  $MCC = 0.20$}
-
 \end{figure}

### 2019-07-26 software/eeg_application/eeg_analysis.py

@@ -1,17 +1,29 @@
+"""
+This script produces all of the GCG networks for later analysis.
+"""
+
+import pywren as pw
 import numpy as np
 import pandas as pd
 
 import seaborn as sns
 import os
+import sys
+
+from constants import (FIGURE_DIR, channels, ADJ_MATRICES,
+                       ADJ_ESTIMATES, RAW_DATA, PROCESSED_DATA)
 
 from matplotlib import pyplot as plt
 from levinson.levinson import compute_covariance, lev_durb
 from multiprocessing.pool import ThreadPool
 from threading import Lock
 
+sys.path.append("../")
 from pwgc.gc_methods import (pwgc_estimate_graph, get_residual_graph,
                              get_X, combine_graphs, estimate_B,
-                             attach_X, compute_bic, form_Xy)
+                             attach_X, compute_bic, form_Xy,
+                             pw_threshold_estimate_graph,
+                             estimate_dense_graph)
 from pwgc.var_system import remove_zero_filters
 
 from matplotlib import rc as mpl_rc
@@ -23,40 +35,39 @@ mpl_rc("font", **font)
 mpl_rc("text", usetex=True)
 
 
-# Ensure we always have the same channels and that they are in the same order.
-channels = [
-    "AF1", "AF2", "AF7", "AF8", "AFZ", "C1", "C2", "C3", "C4", "C5", "C6",
-    "CP1", "CP2", "CP3", "CP4", "CP5", "CP6", "CPZ", "CZ", "F1", "F2", "F3",
-    "F4", "F5", "F6", "F7", "F8", "FC1", "FC2", "FC3", "FC4", "FC5", "FC6",
-    "FCZ", "FP1", "FP2", "FPZ", "FT7", "FT8", "FZ", "O1", "O2", "OZ", "P1",
-    "P2", "P3", "P4", "P5", "P6", "P7", "P8", "PO1", "PO2", "PO7", "PO8",
-    "POZ", "PZ", "T7", "T8", "TP7", "TP8", "X", "Y", "nd"]
-
-# X, Y are "EOG" (Electro-oculogram) for eyes
-# nd is a reference electrode
-
 thread_lock = Lock()
+pywren_executor = pw.default_executor()
+USE_PYWREN = False
+K_CORES = "default"
+max_T = 200
+max_lag = 5
+
+# methods available: pwgc, onepass_pwgc, simple_pwgc, alasso
 
 
-def compute_all_networks():
[…]

### 2019-07-26 software/eeg_application/eeg_graph_analysis.py

@@ -6,6 +6,8 @@ import os
 import grakel
 import pandas as pd
 
+from constants import ADJ_MATRICES, FIGURE_DIR
+
 from grakel import GraphKernel  # Graph kernels
 from datetime import datetime
 from matplotlib import pyplot as plt
@@ -46,6 +48,14 @@ mpl_rc("font", **font)
 mpl_rc("text", usetex=True)
 
 
+def stratified_split(X, y, test_size=0.2):
+    ss = StratifiedShuffleSplit(n_splits=1, test_size=0.2)
+    train_ix, valid_ix = next(ss.split(X, y))
+    X_valid, y_valid = X[valid_ix], y[valid_ix]
+    X_train, y_train = X[train_ix], y[train_ix]
+    return X_valid, y_valid, X_train, y_train
+
+
 def logistic_regression_clf():
     """
     Compute and quantify the quality of a multinomial logistic regression
@@ -55,10 +65,7 @@ def logistic_regression_clf():
     le = LabelEncoder()
     y = le.fit_transform(y)
 
-    ss = StratifiedShuffleSplit(n_splits=1, test_size=0.2)
-    train_ix, valid_ix = next(ss.split(X, y))
-    X_valid, y_valid = X[valid_ix], y[valid_ix]
-    X_train, y_train = X[train_ix], y[train_ix]
+    X_valid, y_valid, X_train, y_train = stratified_split(X, y, test_size=0.2)
 
     clf = Pipeline(
         steps=[("kern", Nystroem(kernel="poly")),
@@ -67,7 +74,8 @@ def logistic_regression_clf():
                                                class_weight="balanced",
                                                multi_class="multinomial",
                                                tol=1e-3,
-                                               max_iter=500, warm_start=True))])
+                                               max_iter=500, warm_start=True))]
+                                               )
     param_grid = {"kern__gamma": np.linspace(0.2, 1.5, 20),
                   "kern__degree": [3, 4, 5],
                   "kern__coef0": [0.5, 1.0, 1.5],
@@ -91,9 +99,9 @@ def logistic_regression_clf():
               "".format(matthews_corrcoef(y_hat_valid, y_valid)))
     plt.xlabel("Predicted Label")
     plt.ylabel("True Label")
-    plt.savefig("../figures/logistic_regression_confusion_matrix.pdf",
+    plt.savefig(FIGURE_DIR + "logistic_regression_confusion_matrix.pdf",
                 bbox_inches="tight", pad_inches=0)
[…]

### 2019-07-26 software/pwgc/gc_methods.py

@@ -288,7 +288,10 @@ def estimate_B(G, max_lag=10, copy_G=False,
 
         if len(a_i) == 0:
             G.nodes[i]["sv^2_hat"] = np.var(y_raw)
-            G.nodes[i]["r"] = G.nodes[i]["x"][max_T + max_lag:]
+            if max_T is not None:
+                G.nodes[i]["r"] = G.nodes[i]["x"][max_T + max_lag:]
+            else:
+                G.nodes[i]["r"] = G.nodes[i]["x"][max_lag:]
             continue
         else:
             X_raw = np.hstack([G.nodes[j]["x"][:, None] for j in a_i])
@@ -862,12 +865,40 @@ def alasso_fista_estimate_dense_graph(X, max_lag=10, max_T=None, nu=1.25,
     return G
 
 
-def pwgc_estimate_graph(X, max_lags=10, alpha=0.05,
-                        method="lstsqr"):
+def pw_threshold_estimate_graph(X, max_lags=10, alpha=0.05,
+                                method="lstsqr", K_cores="default"):
     T, n = X.shape
 
-    F, P = fast_compute_pairwise_gc(X, max_lag=max_lags,
-                                    k_cores=int(1 + np.log(n)))
+    if K_cores == "default":
+        F, P = fast_compute_pairwise_gc(X, max_lag=max_lags,
+                                        k_cores=int(1 + np.log(n)))
+    else:
+        F, P = fast_compute_pairwise_gc(X, max_lag=max_lags,
+                                        k_cores=K_cores)
+
+    P_edges = normalize_gc_score(F, P, T, F_distr=True)
+    P_values = 1 - P_edges[~np.eye(n, dtype=bool)].ravel()
+    t_bh = benjamini_hochberg(P_values, alpha=alpha, independent=False)
+
+    P_edges[P_edges < 1 - t_bh] = 0
+    P_edges[P_edges > 0] = 1.0
+
+    G_hat = nx.DiGraph(P_edges.T)
+    attach_X(G_hat, X)
+    G_hat = estimate_B(G_hat, max_lags, method=method)
+    return G_hat
+
+
+def pwgc_estimate_graph(X, max_lags=10, alpha=0.05,
+                        method="lstsqr", K_cores="default"):
+    T, n = X.shape
+
+    if K_cores == "default":
+        F, P = fast_compute_pairwise_gc(X, max_lag=max_lags,
+                                        k_cores=int(1 + np.log(n)))
+    else:
[…]

### 2019-08-27 paper_appl/paper_appl.tex
some preliminary analysis on GC appl paper.

@@
 
+\section{Laten Structure Recovery Performance Metric}
+The Granger causal graph structure described in section
+\ref{sec:theory} is entirely latent.  That is, we can never observe
+the true causality graph (assuming such a structure \textit{exists}),
+which makes evaluating different support recovery algorithms
+difficult.  Broadly speaking, most work on the topic evaluates the
+performance of competing causality graph recovery algorithms in one of
+two ways (or both): (1) theoretical proofs or (2) simulations where
+the true graph is known by construction.  In this work our goal is to
+establish as valid a third possibility: the performance of a second
+step classifier.
+
+\subsection{Introduction and Motivation}
+More precisely, we suppose we have a two distinct data generating
+processes (``subjects'') labeled $S^{(1)}$ and $S^{(2)}$.  Each
+subject has their own fixed and unobservable Granger causality graph
+$\gcg^{(i)}$.  And, when queried, the subject returns a finite sample
+of data $\{x^{(i)}(t)\}_{t = 1}^{T}$ generated by the model
+\ref{eq:var_model}.  In addition, we have $K$ algorithms
+$A_1, \ldots, A_K$ which consume such a dataset and return an estimate
+$\widehat{\gcg}_k^{(i)}$ of graph $\gcg^{(i)}$.
+
+Following this process, we suppose that we collect $N^{(i)}$ samples
+from each subject.  Applying each algorithm to the resulting time
+series data results in $NK$ datasets
+
+\[
+  \mathcal{D}_k = \bigcup_{i \in \{0, 1\}} \bigcup_{n = 1}^{N^{(i)}} \{(\widehat{\gcg}_{k, n}^{(i)}, i)\}.
+\]
+
+Finally, we fit classifiers $C_k: \widehat{\gcg} \rightarrow \{0, 1\}$
+which consume an estimated causality graph and attempt to return the
+subject label $i \in \{0, 1\}$ that generated the data from which the
+graph was estimated.  We will conclude that Algorithm $k_1$ is more
+accurate at estimating causality graphs than Algorithm $k_2$ if
+
+\begin{equation}
+  \label{eqn:comparison_metric}
+  \P\big\{C_{k_1}(\widehat{\gcg}_{k_1}^{(i)}) = i\big\} > \P\big\{C_{k_2}(\widehat{\gcg}_{k_2}^{(i)}) = i\big\}.
+\end{equation}
+
+That is, if it is easier to label the subject from causality graphs
+estimated by $A_{k_1}$ than it is to do so via $A_{k_2}$, it provides
+evidence that $\widehat{\gcg}_{k_1}^{(i)}$ is closer to
+$\gcg^{(i)}$ than is $\widehat{\gcg}_{k_2}^{(i)}$.
+
+\begin{remark}
+We will provide mathematical justification for this evaluation
+strategy below, but some remarks are in order.  Firstly, if Equation
+\ref{eqn:comparison_metric} holds, it does not constitute
+\textit{proof} that $A_{k_1}$ is superior to $A_{k_2}$ for the task of
+estimating Granger causality graphs.  Indeed, it would be possible to
+construct an algorithm $A$ which attempts to encode the idiosyncracies
+of $s^{(i)}$ present in the data $x^{(i)}(t)$, making no attempt at
+actually recovering $\gcg^{(i)}$.  We therefore formulate our
+theoretical justification such that we make no reference to the actual
+recovery algorithm.  Secondly, we claim that this evaluation strategy
+is distinct from the above evaluation metrics (1) and (2), but that we
+do not claim it is \textit{better} than simulation or mathematical
+argument.  Merely, we propose it as a third mechanism for comparing
+different recovery algorithms that may provide further insight beyond
+(1) and (2).
+\end{remark}
+
+\subsection{Theoretical Analysis}
+
 \section{Finite Sample Graph Recovery}

### 2019-08-27 paper_theory/graph_topology_granger_causality.tex
some preliminary analysis on GC appl paper.

@@
 \def\te{\overset{\mathcal{T}}{\rightarrow}}  % Transfer entropy arrow
-\def\gcg{\mathcal{G}}  % Granger-causality graph
+\def\gcg{\mathcal{G}}  % Granger causality graph
 \def\gcge{\mathcal{E}}  % Graph edges
@@
 
-\abstract{We study Granger-causality in the context of wide-sense
+\abstract{We study Granger causality in the context of wide-sense
   stationary time series, where our focus is on the topological
@@
 
-\keywords{causality graph, gene regulatory networks,
-  Granger-causality, network learning, time series, vector
-  autoregression}
+\keywords{causality graph, Granger causality, network learning, time
+  series, vector autoregression, LASSO}
 
@@
 \label{sec:introduction}
-In this paper we study the notion of Granger-causality
+In this paper we study the notion of Granger causality
 \cite{granger1969investigating} \cite{Granger1980329} as a means of
@@
 
-Granger-causality can generally be formulated by searching for the
+Granger causality can generally be formulated by searching for the
 ``best'' graph structure consistent with observed data, which is in
@@
 variations on the LASSO have been applied in a number of different
-time series contexts as well as Granger-causality
+time series contexts as well as Granger causality
 \cite{DBLP:journals/corr/HallacPBL17} \cite{haufe2008sparse}
@@
 % We study a novel perspective on the problem by establishing conditions
-% under which simple \textit{pairwise} Granger-causality tests are
+% under which simple \textit{pairwise} Granger causality tests are
 % sufficient to recover the entire graph.  Without particular
@@
 arise as a result of considering the topology of the underlying
-Granger-causality graph.  The focus of this paper is to shed light on
+Granger causality graph.  The focus of this paper is to shed light on
 some of these topological questions, in particular, we study a
@@
 Moreover, we show that such graphs are perfectly recoverable with only
-\textit{pairwise} Granger-causality tests, which would otherwise
+\textit{pairwise} Granger causality tests, which would otherwise
 suffer from serious confounding problems (see
@@
 
+Detailed study of Granger causality for star structured graphs has
+been carried out in \cite{jozsa2018relationship}.  See as well
+\cite{barnett2015granger}, \cite{jozsa2018_phd_thesis} for state space
+formulations.
+
 In the case of gene expression networks, we show examples from the
@@
 
-The principle contributions of this paper are as follows: firstly, in
-section \ref{sec:theory} we study \textit{pairwise} Granger-causality
+The principal contributions of this paper are as follows: firstly, in
+section \ref{sec:theory} we study \textit{pairwise} Granger causality
 relations, providing novel theorems connecting the structure of the
@@
 studied by \cite{caines1975feedback} in close connection with
-Granger-causality.  We establish sufficient conditions (sections
+Granger causality.  We establish sufficient conditions (sections
 \ref{sec:strongly_causal_graphs}, \ref{sec:persistent_systems}) under
-which a fully conditional Granger-causality graph can be recovered
+which a fully conditional Granger causality graph can be recovered
 from pairwise tests alone (sec \ref{sec:pairwise_algorithm}).  We
@@
 
-\section{Graph Topological Aspects of Granger-causality}
+\section{Graph Topological Aspects of Granger causality}
 \label{sec:theory}
@@
 The notion of conditional orthogonality is the essence of
-Granger-causality, and enables us to obtain results for a fairly
+Granger causality, and enables us to obtain results for a fairly
 general class of WSS processes, rather than simply $\VAR(p)$ models.
@@
 
-Our principle object of study will be a graph determined by
-Granger-causality relations as follows.
+Our principal object of study will be a graph determined by
+Granger causality relations as follows.
 
 \begin{definition}[Causality graph]
-  We define the Granger-causality graph $\gcg = ([n], \gcge)$ to be the directed
+  We define the Granger causality graph $\gcg = ([n], \gcge)$ to be the directed
   graph formed on $n$ vertices where an edge $(j, i) \in \gcge$ if and
@@
 
-The edges of the Granger-causality graph $\gcg$ can be given a general
+The edges of the Granger causality graph $\gcg$ can be given a general
 notion of ``weight'' by associating an edge $(j, i)$ with the
@@
 \label{sec:pwgc}
-Recall that Granger-causality in general must be understood with
+Recall that Granger causality in general must be understood with
 respect to a particular universe of observations.  If $x_j \gc x_i$
@@
 explains the connection from $x_j$ to $x_i$.  In this section we study
-\textit{pairwise} Granger-causality, and seek to understand when
+\textit{pairwise} Granger causality, and seek to understand when
 knowledge of pairwise relations is sufficient to deduce the true fully
@@
 
-\begin{definition}[Pairwise Granger-causality]
+\begin{definition}[Pairwise Granger causality]
   We will say that $x_j$ pairwise Granger-causes $x_i$ and write
@@
 \begin{proposition}
-  Consider distinct nodes $i, j$ in a Granger-causality graph
+  Consider distinct nodes $i, j$ in a Granger causality graph
   $\gcg$.  If
@@
 \begin{proposition}
-  Consider distinct nodes $i, j$ in a Granger-causality graph $\gcg$.
+  Consider distinct nodes $i, j$ in a Granger causality graph $\gcg$.
   If
@@
   \label{prop:ancestor_properties}
-  If in a Granger-causality graph $\gcg$ where $j \pwgc i$ then
+  If in a Granger causality graph $\gcg$ where $j \pwgc i$ then
   $j \in \anc{i}$ or $\exists k \in \anc{i} \cap\anc{j}$ which is a
@@
   \label{def:strongly_causal}
-  We will say that a Granger-causality graph $\gcg$ is
+  We will say that a Granger causality graph $\gcg$ is
   \textit{strongly causal} if there is at most one directed path between
@@
 \begin{definition}[Persistent]
-  We will say that the process $x(t)$ with Granger-causality graph
+  We will say that the process $x(t)$ with Granger causality graph
   $\gcg$ is \textit{persistent} if for every $i \in [n]$ and every
@@
 \begin{remark}
-  In the context of Granger-causality, ``most'' systems should be
+  In the context of Granger causality, ``most'' systems should be
   persistent.  In particular, $\mathsf{\VAR}(p)$ models are likely to
@@
   \label{thm:scg_recovery}
-  If the Granger-causality graph $\gcg$ for the process $x(t)$ is a
+  If the Granger causality graph $\gcg$ for the process $x(t)$ is a
   strongly causal DAG and Assumption \ref{ass:T_causality} holds, then
@@
   \TitleOfAlgo{Pairwise Graph Recovery}
-  \Input{Pairwise Granger-causality relations between a persistent
-  process of dimension $n$ whose joint Granger-causality
+  \Input{Pairwise Granger causality relations between a persistent
+  process of dimension $n$ whose joint Granger causality
   relations are known to form a strongly causal DAG $\gcg$.}
-  % \Input{Pairwise Granger-causality relations}
+  % \Input{Pairwise Granger causality relations}
   \Output{Edges $\gcge = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of
@@
 %   since when $n = 2$ there is no difference between pairwise
-%   Granger-causality and joint Granger-causality amongst all series --
+%   Granger causality and joint Granger causality amongst all series --
 %   however, one can add any number of nodes having no parents or
@@
   \cite{matthews1975comparison} as a statistic for measuring support
-  recovery performance.  This statistic synthesizes the confusion
-  matrix into a single score appropriate for unbalanced labels and is
-  calibrated to fall into the range $[-1, 1]$ with $1$ being perfect
-  performance, $0$ being the performance of random guessing, and $-1$
-  being perfectly opposed.
+  recovery performance (see also \cite{chicco2017ten} tip \# 8).  This
+  statistic synthesizes the confusion matrix into a single score
+  appropriate for unbalanced labels and is calibrated to fall into the
+  range $[-1, 1]$ with $1$ being perfect performance, $0$ being the
+  performance of random guessing, and $-1$ being perfectly opposed.
 \end{remark}
@@
 In this paper we have argued that considering particular topological
-properties of Granger-causality networks can provide substantial
+properties of Granger causality networks can provide substantial
 insights into the structure of causality graphs with potential for
@@
 causality testing alone is sufficient for recovering a complete
-Granger-causality graph.  Moreover, examples from the literature
+Granger causality graph.  Moreover, examples from the literature
 suggest that such topological assumptions may be reasonable in some
@@
 Finally, the work of \cite{barnett2015granger} has established the
-superiority of Granger-causality testing by state space models (as
+superiority of Granger causality testing by state space models (as
 opposed to pure autoregressions) in many cases.  Combining this work
@@
 Section \ref{sec:pairwise_hypothesis_testing} to instead utilize
-state-space Granger-causality testing) therefore is likely to enable
+state-space Granger causality testing) therefore is likely to enable
 application to very large networks of time series data which are not

### 2019-08-27 paper_theory/supplementary_material.tex
some preliminary analysis on GC appl paper.

@@
     $\B_{ij}$ is the \textit{transposed} conventional adjacency
-    matrix.} of the Granger-causality graph $\gcg$.  Then,
+    matrix.} of the Granger causality graph $\gcg$.  Then,
   $(S^k)_{ij}$ is the number of paths of length $k$ from node $j$ to
@@
   \label{prop:separated_ancestor_uncorrelated}
-  Consider distinct nodes $i, j$ in a Granger-causality graph
+  Consider distinct nodes $i, j$ in a Granger causality graph
   $\gcg$.  If
@@
   \label{lem:vj_perp}
-  Consider distinct nodes $i, j$ in a Granger-causality graph $\gcg$.
+  Consider distinct nodes $i, j$ in a Granger causality graph $\gcg$.
   If $j \not \in \anc{i}$, then $\H_t^{(v_j)} \perp \H_{t}^{(i)}$, and therefore
@@
   \label{prop:ancestor_uncorrelated}
-  Consider distinct nodes $i, j$ in a Granger-causality graph $\gcg$.
+  Consider distinct nodes $i, j$ in a Granger causality graph $\gcg$.
   If
@@
   % \TitleOfAlgo{Pairwise Graph Recovery (PWGC)}
-  % \Input{Pairwise Granger-causality relations between a persistent
-  % process of dimension $n$ whose joint Granger-causality
+  % \Input{Pairwise Granger causality relations between a persistent
+  % process of dimension $n$ whose joint Granger causality
   % relations are known to form a strongly causal DAG $\gcg$.}
-  \Input{Pairwise Granger-causality relations}
+  \Input{Pairwise Granger causality relations}
   \Output{Edges $\gcge = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of
@@
 In subsection \ref{sec:pairwise_hypothesis_testing} we define pairwise
-Granger-causality hypothesis tests, in subsection
+Granger causality hypothesis tests, in subsection
 \ref{sec:model_order_selection} a model order selection criteria, in
@@
 \label{sec:pairwise_hypothesis_testing}
-In performing pairwise checks for Granger-causality $x_j \pwgc x_i$ we
+In performing pairwise checks for Granger causality $x_j \pwgc x_i$ we
 follow the simple scheme of estimating the following two linear models:
@@
 \label{sec:error_rate_control}
-Denote $F_{ij}$ the Granger-causality statistic of Equation
+Denote $F_{ij}$ the Granger causality statistic of Equation
 \eqref{eqn:gc_statistics} with model orders chosen by the methods of
@@
 
-    \Input{Estimates of pairwise Granger-causality statistics $F_{ij}$
+    \Input{Estimates of pairwise Granger causality statistics $F_{ij}$
       (eqn.
@@
 In regards scalability, we have observed that performing the $O(n^2)$
-pairwise Granger-causality calculations consumes the vast majority
+pairwise Granger causality calculations consumes the vast majority
 ($> 90\%$) of the computation time.  Since this step is trivially

### 2019-08-27 software/eeg_application/eeg_analysis.py

@@ -88,6 +88,7 @@ def network_comparison_example():
 
 
 def pw_estimate_subject_G_adj(subject, method="pwgc"):
+    # With PyWren
     data_folder = get_subject_data_folder(subject)
     save_folder = PROCESSED_DATA + method + "/" + ADJ_MATRICES + subject + "/"
     if not os.path.exists(save_folder):

### 2019-09-28 paper/paper.tex
Working on presentation.

@@
 
-\title{Graph Topological Aspects of Granger Causal Network Learning }
+\title{Graph Topological Aspects of Granger Causal Network Learning}
 \author{R. J. Kinnear \\
@@
 has been detrended.  We additionally require that this representation
-can be inverted to yield the $\VAR(\infty)$ form
+tcan be inverted to yield the $\VAR(\infty)$ form
 

### 2019-09-28 paper_appl/paper_appl.tex
Working on presentation.

--- 2019-07-26 paper_appl/paper_appl.tex
@@
 
+\section{Laten Structure Recovery Performance Metric}
+The Granger causal graph structure described in section
+\ref{sec:theory} is entirely latent.  That is, we can never observe
+the true causality graph (assuming such a structure \textit{exists}),
+which makes evaluating different support recovery algorithms
+difficult.  Broadly speaking, most work on the topic evaluates the
+performance of competing causality graph recovery algorithms in one of
+two ways (or both): (1) theoretical proofs or (2) simulations where
+the true graph is known by construction.  In this work our goal is to
+establish as valid a third possibility: the performance of a second
+step classifier.
+
+\subsection{Introduction and Motivation}
+More precisely, we suppose we have a two distinct data generating
+processes (``subjects'') labeled $S^{(1)}$ and $S^{(2)}$.  Each
+subject has their own fixed, distinct, and unobservable Granger
+causality graph $\gcg^{(i)}$.  And, when queried, the subject
+returns% \footnote{Formally, we could define a distribution $t_0$ on
+  % $\N$ and take $X^{(i)} = \{x^{(i)}(t)\}_{t = t_0}^{t_0 + T}$.  Since
+  % $x(t)$ is stationary, this doesn't effect the statistics of the
+  % sample}
+a finite length $T$ sample of time series data
+$X^{(i)} \defeq \{x^{(i)}(t)\}_{t = 1}^{T}$ from a stationary process
+whose temporal dependence structure corresponds to the causality graph
+structure of $\gcg^{(i)}$.  We suppose that we collect $N$ independent
+samples $\{X^{(i)}_n\}_{n = 1}^N$ from each subject.
+
+In addition to the subjects, we have $K$ algorithms $A_1, \ldots, A_K$
+which consume such a dataset $X_n^{(i)}$ and return an estimate
+$\widehat{\gcg}_{k,n}^{(i)}$ of graph $\gcg^{(i)}$.  That is, if we
+view graph estimates as points in $\{0, 1\}^{n^2}$ we have
+$A_k: \R^{T \times n} \rightarrow \{0, 1\}^{n^2}$.
+
+Applying each algorithm to each sample of time series data allows us
+to construct $K$ datasets $\mathcal{D}_1, \ldots, \mathcal{D}_K$, each
+of which contains $2N$ labeled graph estimates
+$(\widehat{\gcg}_{k, n}^{(i)}, i)$:
+
+\[
+  \mathcal{D}_k = \bigcup_{i \in \{1, 2\}} \bigcup_{n = 1}^N \{(\widehat{\gcg}_{k, n}^{(i)}, i)\}.
+\]
+
+Finally, we fit a classifier
+$C_k: \{0, 1\}^{n^2} \rightarrow \{1, 2\}$ to the dataset
+$\mathcal{D}_k$ which consumes an estimated causality graph and
+attempts to predict the subject label $i \in \{1, 2\}$ that generated
+the data from which the graph was estimated.  If classifier $C_{k_1}$
+has superior out-of-sample test accuracy in comparison to $C_{k_2}$,
+we consider this as evidence that Algorithm $A_{k_1}$ has made more
+accurate estimates of $\gcg^{(i)}$ than has Algorithm $A_{k_2}$.  That
+is, if for new graph estimates
+$\widehat{\gcg}_{k_1}^{(i)}, \widehat{\gcg}_{k_2}^{(i)}$ we have
+
+\begin{equation}
+  \label{eqn:comparison_metric}
+  \P\big\{C_{k_1}(\widehat{\gcg}_{k_1}^{(i)}) = i\big\} > \P\big\{C_{k_2}(\widehat{\gcg}_{k_2}^{(i)}) = i\big\}.
+\end{equation}
+
+\begin{remark}
+We will provide mathematical justification for this evaluation
+strategy below, but some remarks are in order.
+
+Firstly, if Equation \ref{eqn:comparison_metric} holds, it does not
+constitute \textit{proof} that $A_{k_1}$ is superior to $A_{k_2}$ for
+the task of estimating Granger causality graphs.  Indeed, it would be
+possible to construct an algorithm $A$ which attempts to encode the
+idiosyncracies of $s^{(i)}$ present in the data $x^{(i)}(t)$, making
+no attempt at actually recovering $\gcg^{(i)}$.  We therefore
+formulate our theoretical justification such that we make no reference
+to the actual recovery algorithm.
+
+Secondly, we claim that this evaluation strategy is distinct from the
+above evaluation metrics (1) and (2) (i.e. mathematical proof, and
+simulation), but we do not claim it is \textit{better}.  Merely, we
+propose it as a third mechanism for comparing different recovery
+algorithms that may provide further insight and quantify performance
+along another axis.
+\end{remark}
+
+\subsection{Theoretical Analysis}
+
 \section{Finite Sample Graph Recovery}

### 2019-09-28 presentations/mcgillOct2019/pres.tex
Working on presentation.

\documentclass{beamer} %With pauses
% \input{\string~/Documents/academics/global_academics/latex_preamble}
%\documentclass[handout]{beamer} %Without pauses
\usepackage[backend=bibtex]{biblatex}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{cancel}
\usefonttheme[onlymath]{serif}

\setbeamertemplate{bibliography item}{}
\setbeamertemplate{bibliography entry title}{}
\setbeamertemplate{bibliography entry location}{}
\setbeamertemplate{bibliography entry note}{}

\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
}

\bibliography{\string~/Documents/academics/global_academics/global_bib.bib}

\usepackage{wasysym}
\usepackage{tikz}

\usetheme{PaloAlto}

\newtheorem*{defn}{Definition}

\def\E{\mathbb{E}}  % Expectation
\def\gc{\overset{\text{GC}}{\rightarrow}}  % Granger causality arrow
\def\ngc{\overset{\text{GC}}{\nrightarrow}}  % Negated Granger causality arrow
\def\pwgc{\overset{\text{PW}}{\rightarrow}}  % Pairwise Granger causality arrow
\def\npwgc{\overset{\text{PW}}{\nrightarrow}}  % Negated pairwise Granger causality arrow
\def\te{\overset{\mathcal{T}}{\rightarrow}}  % Transfer entropy arrow
\def\gcg{\mathcal{G}}  % Granger-causality graph
\def\gcge{\mathcal{E}}  % Graph edges
\def\VAR{\mathsf{VAR}}  % VAR(p) model
\def\B{\mathsf{B}}  % Filter B
\def\wtB{\widetilde{\B}}  % General filter B
\def\A{\mathsf{A}}  % Filter A
\def\H{\mathcal{H}}  % Hilbert space
\def\R{\mathbb{R}}  % Reals
\def\N{\mathbb{N}}  % Naturals
\def\T{\mathsf{T}}  % Transpose
\def\cl{\mathsf{cl}}  % Closure

\newcommand{\linE}[2]{\hat{\E}[#1\ |\ #2]}  % Linear projection
\newcommand{\linEerr}[2]{\xi[#1\ |\ #2]}  % Error of linear projection
\newcommand{\pa}[1]{pa(#1)}  % Parents of a node
\newcommand{\anc}[1]{\mathcal{A}(#1)}  % Ancestors of a node
\newcommand{\ancn}[2]{\mathcal{A}_{#1}(#2)}  % nth ancestors of a node
\newcommand{\gpn}[2]{gp_{#1}(#2)}  % nth generation grandparents
\newcommand{\wtalpha}[2]{\widetilde{\alpha}(#1, #2)}  % Some notation for lem:pwgc_anc
\newcommand{\dist}[2]{\mathsf{d}(#1, #2)}  % Distance between things
\newcommand{\gcgpath}[2]{#1 \rightarrow \cdots \rightarrow #2}  % A shorter path command

\title{Graph Topological Aspects of Granger Causal Network Learning}
\author{\texorpdfstring{Ryan J. Kinnear\newline\url{Ryan@Kinnear.ca}}{Ryan J. Kinnear}}% \author{Ryan J. Kinnear} \\
  % \small\href{mailto:ryan@kinnear.ca}{ryan@kinnear.ca} \\
  % \small\url{https://github.com/RJTK}\\
  % Under the Supervision of Ravi R. Mazumdar}

\institute[University of Waterloo] % (optional, but mostly needed)
{
  University of Waterloo\\
  Department of Electrical and Computer Engineering
}

\date{McGill University, Oct 4 2019}

\AtBeginSection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

% Let's get started
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

\section{Introduction}

\begin{frame}{Causality}
  \begin{itemize}
    \item{Do billiard collisions cause motion?}\pause
    \item{Does standing in the street cause cars to stop?}\pause
    \item{Does smoking cause cancer?}\pause
    \item{Does government spending cause an increase in GDP?}\pause
    \item{Granger causality addresses the last question.}
  \end{itemize}
\end{frame}

\begin{frame}{Causality in Time Series}
  \begin{itemize}
    \item{Consider $x_t, y_t \overset{\text{i.i.d.}}{\sim} \text{BER}(1/2)$}\pause
    \item{If $z_t = x_{t - 1}$ then since cause preceeds effect, intuitively $x$ causes $z$}\pause
    \item{$z'_t = x_{t - 1} \oplus y_{t - 1}$ is more ambiguous.}\pause
    \item{Issue raised by: \fullcite{transfer_entropy_criticism}} 
  \end{itemize}
\end{frame}

\begin{frame}{Causality in Linear Time Series}
  \begin{itemize}
    \item{Granger causality formalizes ``cause proceeds effect'' for WSS processes}\pause
    \item{Interpretation as a ``Flow of Information'' or ``Flow of Energy'' is also reasonable}\pause
    \item{Given WSS processes, $x_1(t), \ldots, x_n(t)$, how can we understand causation $x_i \rightarrow x_j$?}\pause
    \item{\fullcite{granger1969investigating}}
  \end{itemize}
\end{frame}

\begin{frame}{Basic Setup}
  \begin{itemize}
    \item{Basic setup is an $n-$dimensional WSS process $x(t)$}\pause
    \item{Wold representation $x(t) = \sum_{\tau = 0}^\infty A(\tau)v(t - \tau)$}\pause
    \item{$\text{VAR}(\infty)$ representation $x(t) = \sum_{\tau = 1}^\infty B(\tau)x(t - \tau) + v(t)$}\pause
    \item{Require $\E[v(t)v(t - \tau)] = \delta_{\tau}\Sigma_v$, with $\Sigma_v$ Diagonal}\pause
  \end{itemize}

  \begin{defn}
    Linear Projection onto Hilbert space $\H \subseteq L_2(\Omega)$
    \begin{equation*}
      \linE{x}{\H} = \underset{u \in \H}{\text{argmin }} ||x - u||
    \end{equation*}

    Minimum Mean Squared Error
    \begin{equation*}
      \xi[x \ |\ \H] = \E (x - \linE{x}{\H})^2      
    \end{equation*}
  \end{defn}
\end{frame}

\begin{frame}{Basic Setup}
  More notation

  \begin{align*}
    \H_t &= \cl \{\sum_{\tau = 0}^p a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, p \in \N\} \subseteq L_2(\Omega)\\
    % H_t^x &= \{a x(t)\ |\ a \in \R\} \subseteq L_2(\Omega),
  \end{align*}\pause

  \begin{itemize}
    \item{$\H_t^{i}$ Hilbert space spanned by $x_i(t)$ alone}
    \item{$\H_t^{-i}$ Hilbert space spanned by $x(t)$ but \textit{excluding} $x_i(t)$}
  \end{itemize}
\end{frame}

\begin{frame}{Granger Causality}
  \begin{defn}
    For the WSS series $x(t)$ we say that component $x_j$
    \textit{Granger-Causes} component $x_i$ (with respect to $x$)
    and write $x_j \gc x_i$ if given Hilbert spaces $\H_{t - 1}$,
    $\H^{-j}_{t - 1}$ if

    \begin{equation*}
      \linEerr{x_i(t)}{\H_{t - 1}} < \linEerr{x_i(t)}{\H^{-j}_{t - 1}},
    \end{equation*}
  \end{defn}\pause

  \begin{defn}
    We say that $x_j$ pairwise Granger-causes $x_i$ and write
    $x_j \pwgc x_i$ if $x_j$ Granger-causes $x_i$ with respect only to
    $(x_i, x_j)$.
  \end{defn}
\end{frame}

\begin{frame}{Simple Examples}
  \begin{equation*}
    x_i(t) = b_i x(t - 1) + v_i(t); i = 0, 1, 2
  \end{equation*}

  \begin{figure}
    \includegraphics[width=0.75\linewidth]{./figures/var1_example.pdf}
  \end{figure}
\end{frame}

\begin{frame}{Simple Examples}
  \begin{align*}
    x_0(t) &= b_0 x_0(t - 1) + b_{0, 2} x_2(t - 150) + v_0(t)\\
    x_1(t) &= b_1 x_1(t - 1) + b_{1, 2} x_2(t - 150) + v_1(t)\\
    x_2(t) &= b_2 x_2(t - 1) + v_2(t)\\
  \end{align*}

  % TODO: Fix the whitespace consumed by this figure!

  \centering
  \includegraphics[width=0.75\linewidth]{./figures/var2_example.pdf}
\end{frame}

\section{Causality Graphs}
\begin{frame}{Graph Terminology}
  \begin{itemize}
    \item{Graph $\gcg = (V, \gcge)$ with nodes (or vertices) $V$ and edges $\gcge$}\pause
    \item{Directed paths: $j \rightarrow a_0 \rightarrow \cdots \rightarrow a_r \rightarrow i$ for $(a_{k}, a_{k + 1}) \in \gcge$ distinct nodes}\pause
    \item{Cycles: paths with $i = j$}
    \item{Directed Acyclic Graph (DAG): A directed graph without any cycles}\pause
    \item{Parents: $j \in \pa{i}$ iff $(j, i) \in \gcge$}\pause
    \item{Ancestors: $j \in \anc{i}$ iff $j \rightarrow i$ path in $\gcge$}\pause
  \end{itemize}
\end{frame}

\begin{frame}{Granger Causality Graph
  A Granger causality graph is simply a graph $\gcg = ([n], \gcge)$ with $(j, i) \in \gcge$ iff $j \gc i$.

  % TODO: Include a gcg example here, maybe a few papers
  % \includegraphics[]{}
    % \item{Confounder: A node $k$ confounds $i, j$ if $k \in \anc{i} \cap \anc{j}$ if exists a
    %     path $\gcgpath{k}{i}$ not containing $j$, and a path
    %     $\gcgpath{k}{j}$ not containing $i$.}
\end{frame}

\section{Pairwise Causal Flow}
\begin{frame}
\end{frame}

\section{Pairwise Testing Recovery Methods}
\begin{frame}
\end{frame}

\section{LASSO Recovery Methods}
\begin{frame}
\end{frame}

\end{document}

### 2019-09-28 presentations/mcgillOct2019/ex_code.py

@@ -1,0 +1,81 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+from scipy.linalg import solve_discrete_lyapunov
+
+
+def plot_vars(x, title=""):
+    plt.plot(x[:, 0], linewidth=2, alpha=0.75, label="$x_0$")
+    plt.plot(x[:, 1], linewidth=2, alpha=0.75, label="$x_1$")
+    plt.plot(x[:, 2], linewidth=2, alpha=0.75, label="$x_2$")
+    plt.xlabel("$t$")
+    plt.title(title)
+    plt.legend()
+    return
+
+
+def var1_example():
+    n, T = 3, 1000
+    t = np.linspace(0, 1, T)
+    Sv = np.diag(np.array([2.0, 1.0, 0.5]))
+    B = np.zeros((3, 3))
+
+    B[0, 0] = 0.5
+    B[1, 1] = 0.7
+    B[2, 2] = 0.99
+
+    x = np.zeros((T, n))
+    v = np.random.multivariate_normal(mean=np.zeros(n),
+                                      cov=Sv, size=T)
+
+    Sx = solve_discrete_lyapunov(a=B.T, q=Sv)
+
+    x[0, :] = np.random.multivariate_normal(mean=np.zeros(n),
+                                            cov=Sx)
+    for t in range(1, T):
+        x[t, :] = B @ x[t - 1, :] + v[t]
+
+    plot_vars(x, "Independent Processes")
+    plt.savefig("figures/var1_example.png")
+    plt.savefig("figures/var1_example.pdf")
+    plt.show()
+    return
+
+
+def var2_example():
+    n, T = 3, 1000
+    t = np.linspace(0, 1, T)
+    Sv = np.diag(np.array([2.0, 1.0, 0.5]))
+    B = np.zeros((3, 3))
+
+    B[0, 0] = 0.5
+    B[1, 1] = 0.7
+    B[2, 2] = 0.99
+
+    B_lag = np.zeros_like(B)
+    B_lag[0, 2] = 0.8
+    B_lag[1, 2] = 0.7
+
+    x = np.zeros((T, n))
+    v = np.random.multivariate_normal(mean=np.zeros(n),
+                                      cov=Sv, size=T)
+
+    Sx = solve_discrete_lyapunov(a=B.T, q=Sv)
+    t0 = 150
+    x[0, :] = np.random.multivariate_normal(mean=np.zeros(n),
+                                            cov=Sx)
[…]

### 2019-09-28 software/eeg_application/eeg_graph_analysis.py

@@ -1,3 +1,7 @@
+"""
+This script performs the classifier training.
+"""
+
 import numpy as np
 
 import seaborn as sns
@@ -14,12 +18,13 @@ from matplotlib import pyplot as plt
 from scipy.sparse import coo_matrix as sparse_matrix
 from eeg_analysis import channels
 
+from skopt import BayesSearchCV
 from sklearn.svm import SVC
 from sklearn.model_selection import (StratifiedShuffleSplit, StratifiedKFold,
                                      GridSearchCV, learning_curve,
                                      RandomizedSearchCV)
 # from sklearn.learning_curves import learning_curve
-from sklearn.metrics import matthews_corrcoef
+from sklearn.metrics import matthews_corrcoef, make_scorer
 from sklearn.decomposition import PCA, KernelPCA
 from sklearn.cross_decomposition import PLSRegression
 from sklearn.manifold import Isomap, TSNE
@@ -53,19 +58,56 @@ def stratified_split(X, y, test_size=0.2):
     train_ix, valid_ix = next(ss.split(X, y))
     X_valid, y_valid = X[valid_ix], y[valid_ix]
     X_train, y_train = X[train_ix], y[train_ix]
-    return X_valid, y_valid, X_train, y_train
+    return X_train, y_train, X_valid, y_valid
 
 
-def logistic_regression_clf():
+def logistic_regression_alcoholism_clf(base_data_folder="data/processed/pwgc",
+                                       alg_name=""):
+    X, y, z = load_subject_Xyz(base_data_folder)
+    z = np.array(z == "A")
+
+    X_train, z_train, X_valid, z_valid = stratified_split(X, z, test_size=0.2)
+
+    clf = Pipeline(
+        steps=[("kern", Nystroem(kernel="poly", n_components=60)),
+               ("qt", QuantileTransformer(output_distribution="normal")),
+               ("logistic", LogisticRegression(dual=False, solver="lbfgs",
+                                               class_weight="balanced",
+                                               tol=1e-3, penalty="l2",
+                                               max_iter=500, warm_start=True))]
+                                               )
+    param_grid = {"kern__gamma": (1e-4, 1e2, "log-uniform"),
[…]

### 2019-09-29 presentations/mcgillOct2019/pres.tex
Presentation updates.

@@
 \usepackage{cancel}
+\usepackage{centernot}
 \usefonttheme[onlymath]{serif}
@@
 \newtheorem*{defn}{Definition}
+\newtheorem*{proposition}{Proposition}
 
@@
 \def\T{\mathsf{T}}  % Transpose
+\def\Z{\mathbb{Z}}  % Transpose
 \def\cl{\mathsf{cl}}  % Closure
@@
     \item{Does government spending cause an increase in GDP?}\pause
-    \item{Granger causality addresses the last question.}
+    \item{Granger causality attemps to address the last question's category.}
   \end{itemize}
@@
 
-\begin{frame}{Granger Causality Graph
-  A Granger causality graph is simply a graph $\gcg = ([n], \gcge)$ with $(j, i) \in \gcge$ iff $j \gc i$.
+\begin{frame}{Granger Causality Graph}
+  A Granger causality graph is simply a graph $\gcg = ([n], \gcge)$
+  with $(j, i) \in \gcge$ iff $j \gc i$.  $\VAR(1)$ models are
+  particularly simple:
 
-  % TODO: Include a gcg example here, maybe a few papers
-  % \includegraphics[]{}
+  \begin{columns}
+    \begin{column}{0.25\textwidth}
+      \includegraphics[width=1.5\linewidth]{../../figures/example1.pdf}
+    \end{column}
+
+    \begin{column}{0.75\textwidth}
+      \begin{equation*}
+        x(t) = \left[
+          \begin{array}{cccc}
+            0 & 0 & 0 & 0\\
+            a & 0 & 0 & 0\\
+            -a & 0 & 0 & 0\\
+            0 & 1 & 1 & 0\\
+          \end{array}\right]
+        x(t - 1) + v(t)
+      \end{equation*}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Application Examples}
+  Analysis of connectedness of financial institutions
+
+  TODO: Finance example
+  \begin{columns}
+
+  \end{columns}
+
+  \footfullcite{NBERw16223}
+\end{frame}
+
+\begin{frame}{Application Examples}
+  Analysis of brain functional connectivity
+
+  TODO: Anna examples or something
+\end{frame}
+
+\begin{frame}{Application Examples}
+  TODO: Gene expression networks
+\end{frame}
+
+\section{Recovery Methods}
+\begin{frame}{LASSO Recovery Methods}
+  \begin{itemize}
+    \item{Small systems ($n = 2$ or $n \lessapprox 10$) with adequate data are amenable to least squares and hypothesis testing}\pause
+    \item{Large systems ($n \gtrapprox 20$) or when data is limited are addressed via the LASSO (and variants)}\pause
+    \item{Standard linear regression case:
+        \begin{equation*}
+          \underset{\beta \in \R^n}{\text{minimize }} \frac{1}{2T}||y - X\beta||_2^2 + \lambda ||\beta||_1
+        \end{equation*}
+      }\pause
+    \item{$\VAR(p)$ models:
+        \begin{equation*}
+          \underset{B(\tau) \in \R^{n \times n}}{\text{minimize }} \frac{1}{2T}||x(t) - \sum_{\tau = 1}^p B(\tau)x(t - \tau)||_2^2 + \lambda \sum_{\tau = 1}^p ||B(\tau)||_1
+        \end{equation*}
+      }
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{LASSO Recovery Theorems}
+  TODO: Static Case
+\end{frame}
+
+\begin{frame}{LASSO Recovery Theorems}
+  TODO: VAR(p) case
+\end{frame}
+
+\begin{frame}{Structure of Granger Causality Graphs}
+  \begin{itemize}
+    \item{Topology of Granger causality graph is a special feature of time series}\pause
+    \item{TODO: Star shaped graph with a self-loop versus a full cycle with graph -- same number of edges, but qualitatively totally different.}
+  \end{itemize}
+\end{frame}
 % \item{Confounder: A node $k$ confounds $i, j$ if $k \in \anc{i} \cap \anc{j}$ if exists a
@@
 %     $\gcgpath{k}{j}$ not containing $i$.}
+
+\section{Pairwise Causality and ``Causal Flow''}
+\begin{frame}{Pairwise Testing}
+  \begin{itemize}
+    \item{Statistical tests for the case $n = 2$ are simple and well established}\pause
+    \item{Can we recover all of $\gcg$ by only testing pairwise relations?}\pause
+    \item{Thresholding schemes suffer from confounding}\pause
+    \item{``Naive'' pairwise testing is so bad, few papers exist on the topic, or it is used as a simple baseline}\pause
+    \item{\fullcite{tam2013gene_pwgc}}
+  \end{itemize}
 \end{frame}
 
-\section{Pairwise Causal Flow}
-\begin{frame}
+\begin{frame}{Pairwise Causality}
+  \begin{itemize}
+    \item{Recall $j \pwgc i$ if $x_j$ Granger causes $x_i$ w.r.t $(x_i, x_j)$}\pause
+  \end{itemize}
+
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+      \begin{itemize}
+        \item{Intuition suggests that $j \in \anc{i} \implies j \pwgc i$}\pause
+      \end{itemize}
+
+      \includegraphics[width=0.75\linewidth]{../../figures/example1.pdf}\pause
+    \end{column}
+
+    \begin{column}{0.5\textwidth}
+      \begin{itemize}
+        \item{At least $j \gc i \implies j \pwgc i$}\pause
+      \end{itemize}
+      \includegraphics[width=0.75\linewidth]{../../figures/example2.pdf}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Necessary Conditions for Pairwise Causality}
+  First, when $i, j$ are completely separated:
+
+  \begin{proposition}
+    Consider distinct nodes $i, j$ in a Granger causality graph
+    $\gcg$.  If
+
+    \begin{enumerate}
+      \item{$j \not\in \anc{i}$ and $i \not\in \anc{j}$}\pause
+      \item{$\anc{i}\cap\anc{j} = \emptyset$}\pause
+    \end{enumerate}
+
+    then $\H_t^{(i)} \perp \H_t^{(j)}$: $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.
+    \pause  Moreover, this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_t^{(i)}} = 0$.\pause
+  \end{proposition}
+
+  Note that $i, j$ may still share descendants.
+\end{frame}
+
+\begin{frame}{Necessary Conditions for Pairwise Causality}
+  $\H_t^{(i)} \perp \H_t^{(j)}$ is a much stronger conclusion that $j \npwgc i$:
+  \begin{proposition}
+    Consider distinct nodes $i, j$ in a Granger causality graph $\gcg$.
+    If
+
+    \begin{enumerate}
+      \item{$j \not\in \anc{i}$}\pause
+      \item{$\anc{i}\cap\anc{j} = \emptyset$}\pause
+    \end{enumerate}
+
+    then $j \npwgc i$.
+  \end{proposition}\pause
+
+  \begin{itemize}
+    \item{What about the simple graph $k \rightarrow i \rightarrow j$?}\pause
+    \item{Surely $j \npwgc i$, even though $k \in \anc{i} \cap \anc{j}$?}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Necessary Conditions for Pairwise Causality}
+  \begin{definition}[Confounder]
+    A node $k$ is called a \textit{confounder} of nodes $i, j$ if
+    $k \in \anc{i} \cap \anc{j}$ and there exists a path
+    $\gcgpath{k}{i}$ not containing $j$, and a path $\gcgpath{k}{j}$
+    not containing $i$.  A simple example is furnished by the ``fork''
+    graph $i \leftarrow k \rightarrow j$.
+  \end{definition}\pause
+
+\begin{proposition}
+  \label{prop:ancestor_properties}
+  If in a Granger causality graph $\gcg$ where $j \pwgc i$ then
+  $j \in \anc{i}$ or $\exists k \in \anc{i} \cap\anc{j}$ which is a
+  confounder of $(i, j)$.
+\end{proposition}
+\end{frame}
+
+\begin{frame}{Sufficient Conditions for Pairwise Causality}
+  \begin{itemize}
+    \item{Unfortunately, there is no general converse (i.e. $j \in \pa{i} \centernot\implies j \pwgc i$)}\pause
+    \item{We seek special (graph-)\textit{topological} conditions when there is a converse}\pause
+  \end{itemize}
+
+  \begin{definition}[Strongly Causal]
+    \label{def:strongly_causal}
+    We will say that a Granger causality graph $\gcg$ is
+    \textit{strongly causal} if there is at most one directed path between
+    any two nodes.  Strongly Causal Graphs will be referred to as SCGs.
+  \end{definition}
+\end{frame}
+
+\begin{frame}{Strongly Causal Graph Examples}
+  \begin{columns}
+    \begin{column}{0.33\linewidth}
+      TODO: Simple example\pause
+    \end{column}
+    \begin{column}{0.33\linewidth}
+      \small{Slightly larger example}
+      \includegraphics[width=\linewidth]{../../figures/example_scg.pdf}\pause
+    \end{column}
+    \begin{column}{0.33\linewidth}
+      \small{Similar number of edges, but not strongly causal}
+      \includegraphics[width=\linewidth]{../../figures/example_dag.pdf}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{(Almost) Strongly Causal Graph Examples}
+  \begin{columns}
+    \begin{column}{0.5\linewidth}
+      \small{E. Coli gene network \tiny{(\fullcite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty})}}
+      \includegraphics[width=\linewidth]{../../figures/ecoli_regulatory_network.pdf}\pause
+    \end{column}
+    \begin{column}{0.5\linewidth}
+      \small{Saccharomyces cerevisiae gene network \tiny{(\fullcite{learning_genome_scale_regulatory_networks})}}
+      \includegraphics[width=\linewidth]{../../figures/huge_gene_network.pdf}\pause
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{A Partial Converse}
+  \begin{itemize}
+    \item{In strongly causal (and acyclic) graphs, we have the appealing results:}
+  \end{itemize}
+
+  \begin{proposition}
+    \label{prop:pwgc_anc}
+    If $\gcg$ is a strongly causal DAG then $j \in \anc{i} \Rightarrow j \pwgc i$.
+  \end{proposition}\pause
+
+  \begin{corollary}
+    \label{cor:gc_implies_pwgc}
+    If $\gcg$ is a strongly causal DAG then $j \gc i \Rightarrow j \pwgc i$.
+  \end{corollary}\pause
+
+  \begin{itemize}
+    \item{Unfortunately, effects of confounding more nuanced}\pause
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Persistent Systems}
+  \begin{itemize}
+    \item{If $k \in \anc{i}\cap\anc{j}$ any of the following may occur:}\pause
+    \item{
+        $(i \npwgc j$, $j \npwgc i)$,
+        $(i \pwgc j$, $j \npwgc i)$,
+        $(i \npwgc j$, $j \pwgc i)$,
+        $(i \pwgc j$, $j \pwgc i)$}\pause
+    \item{Intuitively, $i$ may ``forget'' information from $k$ before the same information reaches $i$.}\pause
+  \end{itemize}
+
+  A heuristic ``definition'' (to avoid introducing more cumbersome notation):
+  \begin{definition}[Persistent]
+    We will say that the process $x(t)$ with Granger causality graph
+    $\gcg$ is \textit{persistent} if every node has a self-loop,
+    i.e. $\B_{ii}(z) \ne 0$.
+  \end{definition}
+
 \end{frame}
@@
 
-\section{LASSO Recovery Methods}
-\begin{frame}
-\end{frame}
-
 \end{document}

### 2019-09-29 presentations/mcgillOct2019/pres.tex
More figures for presentation.

@@
 \usepackage{centernot}
+\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
+\usepackage{subcaption}
+
 \usefonttheme[onlymath]{serif}
@@
 
-  \begin{proposition}
+  \begin{proposition}[Partial Sufficient Conditions]
     \label{prop:pwgc_anc}
@@
   \end{definition}
+\end{frame}
 
+\begin{frame}{Persistent Systems}
+  Define $H_i(z)$ to be the filter s.t.:
+
+  \begin{equation*}
+    H_i(z)x_i(t) = \linE{x_i(t)}{\H_{t - 1}^{(i)}},
+  \end{equation*}\pause
+
+  And then with $\sigma_k^2 = \E v_k(t)^2$
+  \begin{equation*}
+    \label{eqn:T_filter}
+    T_{ij}(z) = \sum_{k \in \anc{i} \cap \anc{j}}\sigma_k^2\A_{ik}(z^{-1})(1 - H_i(z^{-1}))(1 - H_j(z))\A_{jk}(z)
+  \end{equation*}\pause
+
+  \begin{proposition}
+    \label{prop:persistence_converse}
+    Fix $i, j \in [n]$ and suppose $\exists k \in \anc{i} \cap \anc{j}$
+    which confounds $i, j$.  Then, if $T_{ij}(z)$ is not causal we have
+    $j \pwgc i$, and if $T_{ij}(z)$ is not anti-causal we have
+    $i \pwgc j$.  Moreover, if $T_{ij}(z)$ is two-sided, then $j \pwgc i \iff i \pwgc j$.
+  \end{proposition}
 \end{frame}
@@
 \section{Pairwise Testing Recovery Methods}
-\begin{frame}
+\begin{frame}{Pairwise Recovery}
+  \begin{itemize}
+    \item{Persistant systems where $T_{ij}(z)$ is not two-sided are highly pathological}\pause
+    \item{However, that the following requires an assumption on
+        $T_{ij}(z)$ is an unfortunate blemish on the theory as it is
+        not a topological requirement}\pause
+  \end{itemize}
+
+    \begin{theorem}[Pairwise Recovery]
+      \label{thm:scg_recovery}
+      If the Granger causality graph $\gcg$ for the process $x(t)$ is
+      a strongly causal DAG and $T_{ij}(z)$ is two-sided for every
+      $i, j$, then $\gcg$ can be recovered from pairwise causality
+      tests alone.
+    \end{theorem}
+
+    \begin{itemize}
+      \item{Proof is constructive in that I directly provide an algorithm to do so.}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Pairwise Recovery}
+  \begin{algorithm}[H]
+    \SetKwInOut{Input}{input}
+    \SetKwInOut{Output}{output}
+    \SetKwInOut{Initialize}{initialize}
+    \DontPrintSemicolon
+
+    \footnotesize
+    % \BlankLine
+    % \caption{Pairwise Granger Causality Algorithm}
+    \label{alg:pwgr}
+    % \TitleOfAlgo{Pairwise Graph Recovery}
+    % \Input{Pairwise Granger causality relations between a persistent
+    % process of dimension $n$ whose joint Granger causality
+    % relations are known to form a strongly causal DAG $\gcg$.}
+    % \Input{Pairwise Granger causality relations}
+    % \Output{Edges $\gcge = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of
+    %   the graph $\gcg$.}
+    \Initialize{$S_0 = [n]$, $E_0 = \emptyset$, $k = 1$}\pause
+    \BlankLine
+    $W \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$\\\pause
+    $P_0 \leftarrow \{i \in S_0\ |\ \forall s \in S_0\ (s, i) \not\in W\}$\\\pause
+    \While{$S_{k - 1} \ne \emptyset$}{
+      $S_k \leftarrow S_{k - 1} \setminus P_{k - 1}$\\
+      $P_k \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W\}$\\\pause
+
+      $D_{k0} \leftarrow \emptyset$\\\pause
+      \For{$r = 1, \ldots, k$} 
+      {
+        $Q \leftarrow E_{k - 1} \cup \big(\bigcup_{\ell = 0}^{r - 1} D_{k\ell}\big)$\\\pause
+        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times P_k\ |\ (i, j) \in W,\ \text{no } i \rightarrow j \text{ path in } Q\}$\\\pause
+      }
+      $E_k \leftarrow E_{k - 1} \cup \big(\bigcup_{r = 1}^k D_{kr}\big)$\\\pause
+      $k \leftarrow k + 1$\\\pause
+    }
+    \Return{$E_{k - 1}$}
+  \end{algorithm}
+\end{frame}
+
+\begin{frame}{Pairwise Recovery}
+  An Example (TODO: Intersperse Pauses)
+
+  \begin{columns}
+    \begin{column}{0.5\linewidth}
+      \begin{figure}
+        \includegraphics[width=\linewidth]{../../figures/example_algorithm.pdf}
+      \end{figure}
+    \end{column}
+
+    \begin{column}{0.5\linewidth}
+      \begin{figure}
+        \includegraphics[width=\linewidth]{../../figures/example_algorithm2.pdf}
+      \end{figure}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Simulation}
+  \begin{itemize}
+    \item{Construct random graphs and compare PWGC against LASSO method}\pause
+    \item{I obtained the best results with Zou's Adaptive LASSO, but known to suffer somewhat from false positives}
+    \item{\fullcite{adaptive_lasso_zou2006}}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Simulation}
+  Representative Random Graph Topologies on $n = 50$ Nodes.  Parameter
+  $q$ controls the density of edges.
+  \begin{columns}
+    \begin{column}{0.33\linewidth}
+      \begin{figure}
+        \caption{SCG}
+        \includegraphics[width=\linewidth]{../../figures/example_scg.pdf}
+      \end{figure}
+    \end{column}\pause
+    \begin{column}{0.33\linewidth}
+      \begin{figure}
+        \caption{DAG $(q = \frac{2}{n})$}
+        \includegraphics[width=\linewidth]{../../figures/example_dag.pdf}
+      \end{figure}
+    \end{column}\pause
+    \begin{column}{0.33\linewidth}
+      \begin{figure}
+        \caption{DAG $(q = \frac{4}{n})$}
+        \includegraphics[width=\linewidth]{../../figures/example_dag_q.pdf}
+      \end{figure}
+    \end{column}\pause
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Simulation}
+  $\VAR(p)$ model with $n = 50, p = 15$ and randomized edge placements.  Left: SCG, Right: DAG ($q = 2/n$)
+  \begin{columns}
+    \begin{column}{0.5\linewidth}
+      \includegraphics[width=0.9\linewidth]{./figures/new_lasso_comparison_scg_pmax15_simulation_mcc_only.png}
+    \end{column}
+
+    \begin{column}{0.5\linewidth}
+      \includegraphics[width=0.9\linewidth]{./figures/new_lasso_comparison_dag_pmax15_simulation_mcc_only.png}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Simulation}
+  \includegraphics[width=\linewidth]{../../figures/new_increasing_n_simulation.pdf}\pause
+\end{frame}
+
+\begin{frame}{Simulation}
+  \includegraphics[width=\linewidth]{../../figures/new_mcc_comparison001.pdf}
 \end{frame}

### 2019-09-29 presentations/student_dynamical_systems_waterloo_2019/Abstract-Template.tex

--- 2019-08-27 paper_theory/graph_topology_granger_causality.tex
@@
-\documentclass{statsoc}
-% \usepackage{mathtime}
-\usepackage{ccfonts,eulervm}
-\usepackage{tabularx}
-\usepackage{xr}
-\externaldocument{supplementary_material}
-% ---------------------------------------------------------------------------------------------------------------------------------
-% The statsoc style file is broken?
-% I have copied this code straight from stack overflow
-% https://tex.stackexchange.com/questions/351768/problem-with-statsoc-class-and-pdflatex
-\usepackage{geometry}
-
-\geometry{
-  textwidth=33pc,
-  textheight=\dimexpr48\baselineskip+\topskip\relax,
-  marginparsep=11pt,
-  marginparwidth=107pt,
-  footnotesep=6.65pt,
-  headheight=9pt,
-  headsep=9pt,
-  footskip=30pt,
-}
-% ---------------------------------------------------------------------------------------------------------------------------------
-
-% \usepackage{natbib}
-\bibliographystyle{chicago}
-
-%% bibliography stuff -- this needs come before the preamble inclusion
-\usepackage{enumitem}
-\usepackage{etex,etoolbox}
-\usepackage{hyperref}
-\usepackage{fullpage}
-\setlength{\marginparwidth}{2cm}
-\usepackage{todonotes}
-\usepackage{multirow}
-\usepackage{booktabs}
-
-\def\gc{\overset{\text{GC}}{\rightarrow}}  % Granger causality arrow
-\def\ngc{\overset{\text{GC}}{\nrightarrow}}  % Negated Granger causality arrow
-\def\pwgc{\overset{\text{PW}}{\rightarrow}}  % Pairwise Granger causality arrow
-\def\npwgc{\overset{\text{PW}}{\nrightarrow}}  % Negated pairwise Granger causality arrow
-\def\te{\overset{\mathcal{T}}{\rightarrow}}  % Transfer entropy arrow
-\def\gcg{\mathcal{G}}  % Granger causality graph
-\def\gcge{\mathcal{E}}  % Graph edges
-\def\VAR{\mathsf{VAR}}  % VAR(p) model
-\def\MA{\mathsf{MA}}  % MA(p) model
-\def\B{\mathsf{B}}  % Filter B
-\def\wtB{\widetilde{\B}}  % General filter B
-\def\A{\mathsf{A}}  % Filter A
-\def\H{\mathcal{H}}  % Hilbert space
-
-\newcommand{\linE}[2]{\hat{\E}[#1\ |\ #2]}  % Linear projection
-\newcommand{\linEerr}[2]{\xi[#1\ |\ #2]}  % Error of linear projection
-\newcommand{\pa}[1]{pa(#1)}  % Parents of a node
-\newcommand{\anc}[1]{\mathcal{A}(#1)}  % Ancestors of a node
-\newcommand{\ancn}[2]{\mathcal{A}_{#1}(#2)}  % nth ancestors of a node
-\newcommand{\gpn}[2]{gp_{#1}(#2)}  % nth generation grandparents
-\newcommand{\wtalpha}[2]{\widetilde{\alpha}(#1, #2)}  % Some notation for lem:pwgc_anc
-\newcommand{\dist}[2]{\mathsf{d}(#1, #2)}  % Distance between things
-\newcommand{\gcgpath}[2]{#1 \rightarrow \cdots \rightarrow #2}  % A shorter path command
-
-% \usepackage{fullpage}
-\usepackage{framed}
-
-% Figures
-\usepackage{graphicx}
-\usepackage{caption}  % This is not recommended?
-\usepackage{subcaption}
-% \usepackage{wrapfig}
-% \usepackage{svg}
-
-% Math packages, theorem definitions and numbering
-\usepackage{amsmath}
-\usepackage{amssymb}
-% \usepackage{amsthm}
-% \usepackage{mathrsfs} % Fancy scripted font
-\usepackage{bm}  % Bold math
-
-% Misc packages
-\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
-\usepackage{algorithm2e} %{algorithm} environment
-\usepackage{soul}  % \hl highlighting
-\usepackage{color}
-\usepackage{mathtools}  % For my \ceil function
-
-% Theorems (with italics)
-% \theoremstyle{plain}  % Style definition removes italics
-\newtheorem{theorem}{Theorem}
-\newtheorem{corollary}{Corollary}
-\newtheorem{proposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-
-% \theoremstyle{definition}
-\newtheorem{remark}{Remark}
-\newtheorem{definition}{Definition}
-\newtheorem{example}{Example}
-\newtheorem{assumption}{Assumption}
-
-% Should I do this?  It makes cross references much easier,
-% but is possibly extra complication.
-% \counterwithin{theorem}{section}
-% \counterwithin{corollary}{section}
-% \counterwithin{proposition}{section}
-% \counterwithin{lemma}{section}
-
-% \counterwithin{remark}{section}
-% \counterwithin{definition}{section}
-% \counterwithin{example}{section}
-% \counterwithin{assumptions}{section}
-
-% keywords
-\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}
-
-% General
-\def\defeq{\overset{\Delta}{=}}  % Equal with triangle
-\def\cl{\mathsf{cl\ }}  % Closure
-\newcommand{\sgn}[1]{\mathsf{sgn}(#1)}  % sign function
-
-% Calculus
-\def\d{\mathsf{d}}  % Differential operator
-
-% Functions
-\def\ln{\mathsf{ln\ }}  % Natural logarithm
-\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}  % Ceiling
-
-% Probability
-\def\H{\mathcal{H}}  % Hilbert space
-\def\E{\mathbb{E}}  % Expectation
-\def\Var{\text{Var}}  % Variance
-\def\P{\mathbb{P}}  % Probability Measure
-\def\F{\mathcal{F}}  % A sigma algebra
-\def\sX{\mathcal{X}}  % Another sigma algebra
-\def\KL{\mathbf{D}_{KL}}  % KL divergence
-\def\bF{\mathbf{F}}  % Whole F-meas space
-
-% Standard sets
-\def\Z{\mathbb{Z}}  % Set of integers
-\def\R{\mathbb{R}}  % Set of real numbers
-\def\C{\mathbb{C}}  % Set of complex numbers
-\def\N{\mathbb{N}}  % Set of natural numbers
-\def\ball{\mathbb{B}}  % Open ball
-\def\clball{\overline{\ball}}  % Closed ball
-
-% Linear algebra
-\def\rk{\mathsf{rk }}  % The rank
-\def\tr{\mathsf{tr }}  % The trace
-\def\T{\mathsf{T}}  % Transpose notation
-\def\c{\mathsf{c}}  % complement
-\def\dg{\mathsf{dg }}   %  Diagonal vector of a matrix
-\def\Dg{\mathsf{Dg }}   %  Diagonal matrix from a vector
-\def\ind{\mathbf{1}}  % Ones vector or indicator
-\def\matvec{\textbf{vec}}  % Vector operator
-\def\<{\langle}  % < Inner product
-\def\>{\rangle}  % > Inner product
-\newcommand{\inner}[2]{\langle #1, #2 \rangle}  % Inner product
-\newcommand{\innerT}[2]{#1^\T #2}  % Inner product for finite vectors
-
-\graphicspath{{../figures/}}
-
-\title{Graph Topological Aspects of Granger Causal Network Learning}
-\author[Author 1 {\it et al.}]{R. J. Kinnear}
-\address{
-  University of Waterloo,
-  Waterloo,
-  Canada.}
-\email{Ryan@Kinnear.ca}
-
-\author{R. R. Mazumdar}
-\address{
-  University of Waterloo,
-  Waterloo,
-  Canada.}
+%    DO NOT CHANGE THE SETTINGS BELOW
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[12pt]{article}
+\usepackage{epsfig}
+\usepackage{graphicx,float}
+\usepackage{url}
+\usepackage[hidelinks,colorlinks=true,bookmarks=true]{hyperref}
+%*******************************************
+\setlength{\textheight}{24cm}		
+\setlength{\textwidth}{17.5cm}		
+\setlength{\oddsidemargin}{-0.5cm}
+\setlength{\topmargin}{-1.25cm}
+%*******************************************
+ \pagestyle{empty}
 
 \begin{document}
+\hspace{-0.70cm}\rule{17.5cm}{0.01 in} \\
+\vspace{-0.4cm}
+\begin{flushleft}
+\Large \textbf{\noindent
+Insert the Title of Your Talk Here}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\\
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{0.5cm}
+\normalsize
+ %%%%%%             NAMES OF AUTHORS         %%%%%%
+ %%%%%% NAME OF SPEAKER SHOULD BE UNDERLINED %%%%%%%%%
+\normalsize{
+ \underline{Ryan J. Kinnear}$^1$, Ravi R. Mazumdar$^1$
+} \\
+\vspace{5mm}
+\textit{\footnotesize
+ %%%%%% AFFILIATION OF AUTHORS %%%%%%
+$^1$ Department of Electrical and Computer Engineering, University of Waterloo, Waterloo, Ontario N2L 3G1, Canada\\
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+}
+\end{flushleft}
 
-\abstract{We study Granger causality in the context of wide-sense
-  stationary time series, where our focus is on the topological
-  aspects of the underlying causality graph.  We establish sufficient
-  conditions (in particular, we develop the notion of a ``strongly
-  causal'' graph topology) under which the true causality graph can be
-  recovered via pairwise causality testing alone, and provide examples
-  from the gene regulatory network literature suggesting that our
-  concept of a strongly causal graph may be applicable to this field.
-  We implement and detail finite-sample heuristics derived from our
-  theory, and establish through simulation the efficiency gains (both
-  statistical and computational) which can be obtained (in comparison
-  to LASSO-type algorithms) when structural assumptions are met.}
+We study Granger causality in the context of wide-sense stationary
+time series, where our focus is on the topological aspects of the
+underlying causality graph.  We establish sufficient conditions (in
+particular, we develop the notion of a ``strongly causal'' graph
+topology) under which the true causality graph can be recovered via
+pairwise causality testing alone, and provide examples from the gene
+regulatory network literature suggesting that our concept of a
+strongly causal graph may be applicable to this field.  We implement
+and detail finite-sample heuristics derived from our theory, and
+establish through simulation the efficiency gains (both statistical
+and computational) which can be obtained (in comparison to LASSO-type
+algorithms) when structural assumptions are met.
 
-\keywords{causality graph, Granger causality, network learning, time
-  series, vector autoregression, LASSO}
-
-\paragraph{Acknowledgement}
-
-We acknowledge the support of the Natural Sciences and Engineering Research Council of Canada (NSERC), [funding reference number 518418-2018].
-Cette recherche a été financée par le Conseil de recherches en sciences naturelles et en génie du Canada (CRSNG), [numéro de référence 518418-2018].
-
-\section{Introduction and Review}
-\label{sec:introduction}
 In this paper we study the notion of Granger causality
@@
 and power system design \cite{Misyrlis2016450}, \cite{yuan2014root}.
-
-Granger causality can generally be formulated by searching for the
-``best'' graph structure consistent with observed data, which is in
-general an extremely challenging problem (i.e. it may be framed as a
-best subset selection problem, see, \cite{hastie_bss_comp} for recent
-improvements in BSS methods), moreover, the comparison of quality
-between different structures, and hence the notion of ``best'' needs
-qualification.  In applications where we are interested merely in
-minimizing the mean squared error of a linear one-step-ahead
-predictor, then we will be satisfied with an entirely dense graph of
-connections, since each edge can only serve to reduce estimation
-error.  However, since the number of edges scales quadratically in $n$
-(the number of nodes) it becomes imperative to infer a sparse
-causality graph for large systems, both to avoid overfitting observed
-data, as well as to aid the interpretability of the results.
-
-A fairly early approach to the problem in the context of large systems
-is provided by \cite{bach2004learning}, where the authors apply a
-local search heuristic to the Whittle likelihood with an AIC
-penalization.  The local search heuristic is a common approach to
-combinatorial optimization due to it's simplicity, but is liable to
-get stuck in shallow local minima.
-
-A second and wildly successful heuristic is the LASSO regularizer
-\cite{tibshirani1996regression}, which can be understood as a natural
-convex relaxation to penalizing the count of the non-zero edges.  The
-LASSO enjoys fairly strong theoretical guarantees
-\cite{wainwright2009sharp}, extending largely to the case of
-stationary time series data with a sufficiently fast rate of
-dependence decay \cite{basu2015} \cite{wong2016lasso}
-\cite{autoregressive_process_modelling_via_the_lasso_procedure}, and
-variations on the LASSO have been applied in a number of different
-time series contexts as well as Granger causality
-\cite{DBLP:journals/corr/HallacPBL17} \cite{haufe2008sparse}
-\cite{bolstad2011causal} \cite{he2013stationary}
-\cite{grouped_graphical_granger_modelling_for_gene_expression_regulatory_networks_discovery}.
-One of the key improvements to the original LASSO algorithm is the
-adaptive (i.e. weighted) ``adaLASSO'' \cite{adaptive_lasso_zou2006},
-for which oracle results (i.e. asymptotic support recovery) are
-established under less restrictive conditions than for the vanilla
-LASSO.
-
-% We study a novel perspective on the problem by establishing conditions
-% under which simple \textit{pairwise} Granger causality tests are
-% sufficient to recover the entire graph.  Without particular
-% topological assumptions, pairwise testing naturally results in a large
-% number of false positives \cite{tam2013gene_pwgc}.
-
-In the context of time series data, sparsity assumptions remain
-important, but there is significant additional structure that may
-arise as a result of considering the topology of the underlying
-Granger causality graph.  The focus of this paper is to shed light on
-some of these topological questions, in particular, we study a
-particularly simple notion of graph topology which we term
-``strongly causal'' and show that stationary times series whose
-underlying causality graph has this structure satisfy natural
-intuitive notions of ``information flow'' through the graph.
-Moreover, we show that such graphs are perfectly recoverable with only
-\textit{pairwise} Granger causality tests, which would otherwise
-suffer from serious confounding problems (see
-e.g. \cite{tam2013gene_pwgc} for earlier work on pairwise testing).
-Aside from being an interesting theoretical perspective, prior
-assumptions about the underlying graph (similarly to sparsity
-assumptions) can greatly improve upon the statistical power of
-causality graph recovery algorithms when the assumptions are met.
-
-Detailed study of Granger causality for star structured graphs has
-been carried out in \cite{jozsa2018relationship}.  See as well
-\cite{barnett2015granger}, \cite{jozsa2018_phd_thesis} for state space
-formulations.
-
-In the case of gene expression networks, we show examples from the
-literature which suggest our concept of a ``strongly causal graph''
-topology may have application in this field (see Section
-\ref{sec:strongly_causal_graphs}).
 
@@
 
-\section{Graph Topological Aspects of Granger causality}
-\label{sec:theory}
-\subsection{Formal Setting}
-Consider the space $L_2(\Omega)$, the usual Hilbert space of finite
-variance random variables over a probability space
-$(\Omega, \mathcal{F}, \mathbb{P})$ having inner product
-$\inner{x}{y} = \E[xy]$.  We will work with a discrete time and
-wide-sense stationary (WSS) $n$-dimensional vector valued process
-$x(t)$ (with $t \in \Z$) where the $n$ elements take values in $L_2$.
-We suppose that $x(t)$ has zero mean, $\E x(t) = 0$, and has
-absolutely summable matrix valued covariance sequence
-$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$, and an absolutely continuous
-spectral density.
+In a general Granger causality graph, there are natural
+\textit{necessary} conditions for a node $j$ to pairwise cause node
+$i$, namely that $j$ is an ancestor to $i$ or that there is a third
+confounding node $k$.  On the other hand, we will see simple examples
+for why \textit{sufficient} conditions are more nuanced: if $j$ causes
+$i$ conditionally on all other nodes in the graph, it need not be the
+case that $j$ causes $i$ when the two are observed in isolation.  We
+examine a special \textit{graph topological} assumption we term
+``strongly causal'' wherein a natural converse \textit{does} hold,
+satisfying an intuitive notion of ``causality flow'' through the
+graph.  We highlight examples from the literature on gene networks
+suggesting that assumptions of strongly causal graph structure may not
+be out of place in some application areas.
 
-We will also work frequently with the spaces spanned by the values of
-such a process
+As a consequence of the causality graph being strongly causal, we show
+that it is possible to recover the full conditional causality graph
+through pairwise testing alone -- that is, it isn't necessary to
+perform any conditional causality tests.
 
-\begin{equation}
-  \label{eq:hilbert_space_defn}
-  \begin{aligned}
-    \H_t^{(x)} &= \cl \{\sum_{\tau = 0}^p a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, p \in \N\} \subseteq L_2(\Omega)\\
-    H_t^{(x)} &= \{a x(t)\ |\ a \in \R\} \subseteq L_2(\Omega),
-  \end{aligned}
-\end{equation}
+% $\bullet$ Abstracts may contain equations: Equations are to be centered horizontally, with equation numbers aligned with the right margin. See equation (\ref{eq1})
+% \begin{equation}
+% e^{i\pi}+1=0.\label{eq1}
+% \end{equation}
 
-where the closure is naturally in mean-square.  We will often omit the
-superscript $x$ which should be clear from context.  Evidently these
-spaces are separable, and as closed subspaces of a Hilbert space they
-are themselves Hilbert.  We will denote the spaces generated in
-analogous ways by particular components of $x$ as e.g.
-$\H_t^{(i, j)}$, $\H_t^{i}$ or by all but a particular component as
-$\H_t^{(-j)}$.
+% $\bullet$ Abstracts must contain at least two  references. See   \cite{ASY,STE}.
 
-As a consequence of the Wold decomposition theorem (see
-e.g. \cite{lindquist}), every WSS sequence has the moving average
-$\MA(\infty)$ representation
+% $\bullet$ Abstracts may contain figures. For example, Figure \ref{fig1}.
 
-\begin{equation}
-  \label{eqn:wold}
-  x(t) = c(t) + \sum_{\tau = 0}^\infty A(\tau) v(t - \tau),
-\end{equation}
-
-where $c(t)$ is a purely deterministic sequence, $v(t)$ is an
-uncorrelated sequence and $A(0) = I$.  We will assume that $c(t) = 0$.
-Given our setup, this representation can be inverted to yield the
-$\VAR(\infty)$ form
-
-\begin{equation}
-  \label{eqn:ar_representation}
-  x(t) = \sum_{\tau = 1}^\infty B(\tau) x(t - \tau) + v(t).
-\end{equation}
-
-The Equations \eqref{eqn:wold}, \eqref{eqn:ar_representation} can be
-represented as $x(t) = \A(z)v(t) = \B(z)x(t) + v(t)$ via the action
-(convolution) of the operators (LTI filters)
-$$\A(z) \defeq \sum_{\tau = 0}^\infty A(\tau)z^{-\tau}$$ and
-$$\B(z) \defeq \sum_{\tau = 1}^\infty B(\tau)z^{-\tau},$$ where the
-operator $z^{-1}$ is the back shift operator acting on
-$\ell_2^n(\Omega, \mathcal{F}, \mathbb{P})$, that is:
-
-\begin{equation}
-  \label{eqn:filter_action}
-  \B_{ij}(z)x_j(t) \defeq \sum_{\tau = 1}^\infty B_{ij}(\tau)x_j(t - \tau).
-\end{equation}
-
-Finally, we have the inversion formula
-
-\begin{equation}
-  \label{eqn:lsi_inversion}
-  \A(z) = (I - \B(z))^{-1} = \sum_{k = 0}^\infty \B(z)^k.
-\end{equation}
-
-The aforementioned assumptions are quite weak.  The strongest
-assumption we require is finally that $\Sigma_v$ is a diagonal
-positive-definite matrix, which is referred to as a lack of
-instantaneous feedback in $x(t)$.  We formally state our setup as a
-definition, which is the setup for the remainder of the paper:
-
-\begin{definition}[Basic Setup]
-  \label{def:basic_setup}
-  The process $x(t)$ is an $n$ dimensional wide sense stationary
-  process having invertible $\VAR(\infty)$ representation
-  \eqref{eqn:ar_representation} where $v(t)$ is sequentially
-  uncorrelated and has a diagonal positive-definite covariance matrix.
-  The $\MA(\infty)$ representation of Equation \eqref{eqn:wold} has
-  $c(t) = 0$ and $A(0) = I$.
-\end{definition}
-
-\subsection{Granger Causality}
-
-\begin{definition}[Granger Causality]
-  \label{def:granger_causality}
-  For the WSS series $x(t)$ satisfying the assumptions of Definition
-  \ref{def:basic_setup} we will say that component $x_j$
-  \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x$)
-  and write $x_j \gc x_i$ if
-
-\begin{equation}
-  \linEerr{x_i(t)}{\H_{t - 1}} < \linEerr{x_i(t)}{\H^{(-j)}_{t - 1}},
-\end{equation}
-
-where $\xi[x \ |\ \H] \defeq \E (x - \linE{x}{\H})^2$ is the mean
-squared estimation error and $\linE{x}{\H} \defeq \text{proj}_{\H}(x)$
-denotes the (unique) projection onto the Hilbert space $\H$.
-\end{definition}
-
-This notion captures the idea that the process $x_j$ provides
-information about $x_i$ that is not available from elsewhere.  The
-caveat ``with respect to $x$'' is important in that GC relations can
-change when components are added to or removed from our collection $x$
-of observations, e.g. new GC relations can arise if we remove the
-observations of a common cause, and existing GC relations can
-disappear if we observe a new mediating series. The notion is closely
-related to the information theoretic measure of transfer entropy,
-indeed, if the distribution of $v(t)$ is known to be Gaussian then
-they are equivalent \cite{barnett2009granger}.
-
-The notion of conditional orthogonality is the essence of
-Granger causality, and enables us to obtain results for a fairly
-general class of WSS processes, rather than simply $\VAR(p)$ models.
-
-\begin{definition}[Conditional Orthogonality]
-  \label{lem:conditional_orthogonality_equivalence}
-  Consider three closed subspaces of a Hilbert space $\mathcal{A}$,
-  $\mathcal{B}$, $\mathcal{X}$.  We say that $\mathcal{A}$ is
-  conditionally orthogonal to $\mathcal{B}$ given $\mathcal{X}$
-  and write $\mathcal{A} \perp \mathcal{B}\ |\ \mathcal{X}$ if
-
-    \begin{equation*}
-      \inner{a - \linE{a}{\mathcal{X}}}{b - \linE{b}{\mathcal{X}}} = 0\ \forall a \in \mathcal{A}, b \in \mathcal{B}.
-    \end{equation*}
-
-  An equivalent condition is that (see \cite{lindquist} Proposition 2.4.2)
-  \begin{equation*}
-    \linE{\beta}{\mathcal{A} \vee \mathcal{X}} = \linE{\beta}{\mathcal{X}}\ \forall \beta \in \mathcal{B}
-  \end{equation*}
-\end{definition}
-
-\begin{theorem}[Granger Causality Equivalences]
-  \label{thm:granger_causality_equivalences}
-  The following are equivalent:
-
-  \begin{enumerate}
-    \item{$x_j \ngc x_i$}
-    \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$ i.e. $\B_{ij}(z) = 0$}
-    \item{$H_t^{i} \perp \H_{t - 1}^{(j)}\ |\ \H_{t - 1}^{(-j)}$}
-    \item{$\linE{x_i(t)}{\H_{t - 1}^{(-j)}} = \linE{x_i(t)}{\H_{t - 1}}$}
-  \end{enumerate}
-\end{theorem}
-
-\subsection{Granger Causality Graphs}
-We establish some graph theoretic notation and terminology, collected
-formally in definitions for the reader's convenient reference.
-
-\begin{definition}[Graph Theory Review]
-  A \textit{graph} $\gcg = (V, \gcge)$ is simply a
-  tuple of sets respectively called \textit{nodes} and \textit{edges}.
-  Throughout this paper, we have in all cases
-  $V = [n] \defeq \{1, 2, \ldots, n\}$.  We will also focus solely on
-  \textit{directed} graphs, where the edges
-  $\gcge \subseteq V \times V$ are \textit{ordered} pairs.
-
-  A (directed) \textit{path} (of length $r$) from node $i$ to node
-  $j$, denoted $\gcgpath{i}{j}$, is a sequence
-  $a_0, a_1, \ldots, a_{r - 1}, a_r$ with $a_0 = i$ and $a_r = j$ such
-  that $\forall\ 0 \le k \le r\ (a_k, a_{k + 1}) \in \gcge$, and where
-  $(a_k, a_{k - 1})$ are \textit{distinct} for $0 \le k < r$.
-
-  A \textit{cycle} is a path of length $2$ or more between a node and
-  itself.  An edge between a node and itself $(i, i) \in \gcge$ (which
-  we do not consider to be a cycle) is referred to as a \textit{loop}.
-
-  A graph $\gcg$ is a \textit{directed acyclic graph} (DAG) if it is a
-  directed graph and does not contain any cycles.
-\end{definition}
-
-\begin{definition}[Parents, Grandparents, Ancestors]
-  A node $j$ is a \textit{parent} of node $i$ if $(j, i) \in \gcge$.
-  The set of all $i$'s parents will be denoted $\pa{i}$, and we
-  explicitly exclude loops as a special case, that is,
-  $i \not\in \pa{i}$ even if $(i, i) \in \gcge$.
-
-  The set of level $\ell$ \textit{grandparents} of node $i$, denoted
-  $\gpn{\ell}{i}$, is the set such that $j \in \gpn{\ell}{i}$ if and
-  only if there is a \textit{directed path} of length $\ell$ in $\gcg$
-  from $j$ to $i$.  Clearly, $\pa{i} = \gpn{1}{i}$.
-
-  Finally, the set of \textit{level $\ell$ ancestors} of $i$:
-  $\ancn{\ell}{i} = \bigcup_{\lambda \le \ell}\gpn{\lambda}{i}$ is the
-  set such that $j \in \ancn{\ell}{i}$ if and only if there is a
-  directed path of length $\ell$ \textit{or less} in $\gcg$ from $j$
-  to $i$.  The set of \textit{all ancestors} of $i$
-  (i.e. $\ancn{n}{i}$) is denoted simply $\anc{i}$.
-
-  Recall that we do not allow a node to be it's own parent, although
-  unless $\gcg$ is a DAG, a node \textit{can} be it's own ancestor.
-  We will occasionally need to explicitly exclude $i$ from $\anc{i}$,
-  in which case we will write $\anc{i}\setminus \{i\}$.
-\end{definition}
-
-Our principal object of study will be a graph determined by
-Granger causality relations as follows.
-
-\begin{definition}[Causality graph]
-  We define the Granger causality graph $\gcg = ([n], \gcge)$ to be the directed
-  graph formed on $n$ vertices where an edge $(j, i) \in \gcge$ if and
-  only if $x_j$ Granger-causes $x_i$ (with respect to $x$).  That is,
-  $$(j, i) \in \gcge \iff j \in \pa{i} \iff x_j \gc x_i.$$
-\end{definition}
-
-The edges of the Granger causality graph $\gcg$ can be given a general
-notion of ``weight'' by associating an edge $(j, i)$ with the
-\textit{strictly causal} LTI filter $\B_{ij}(z)$ (see Equation
-\eqref{eqn:filter_action}).  Thence, the matrix $\B(z)$ is analogous
-to a \textit{weighted adjacency matrix}\footnote{\footnotesize We are using the
-  convention that $\B_{ij}(z)$ is a filter with input $x_j$ and output
-  $x_i$ so as to write the action of the system as $\B(z)x(t)$ with
-  $x(t)$ as a column vector.  This competes with the usual convention
-  for adjacency matrices where $A_{ij} = 1$ if there is an edge
-  $(i, j)$.  In our case, the sparsity pattern of $\B_{ij}$ is the
-  \textit{transposed} conventional adjacency matrix.} for the graph $\gcg$.  And,
-in the same way that the $k^{\text{th}}$ power of an adjacency matrix
-counts the number of paths of length $k$ between nodes,
-$(\B(z)^k)_{ij}$ is a filter isolating the ``action'' of $j$ on $i$ at
-a time lag of $k$ steps, this is exemplified in the inversion formula
-\eqref{eqn:lsi_inversion}.
-
-From the $\VAR$ representation of $x(t)$ there is clearly a tight
-relationship between each node and it's parent nodes, the relationship
-is quantified through the sparsity pattern of $B(z)$.  Similarly, the
-following proposition is analogous to the definition of feedback free
-processes of \cite{caines1975feedback} and provides an interpretation
-of the sparsity pattern of $A(z)$ (from the MA representation of
-$x(t)$) in terms of the causality graph $\gcg$.
-
-\begin{proposition}[Ancestor Expansion]
-  \label{prop:parent_expanding}
-  The component $x_i(t)$ of $x(t)$ can be represented in terms of it's
-  parents in $\gcg$:
-
-  \begin{equation}
-    \label{eqn:parent_expansion}
-    x_i(t) = v_i(t) + \B_{ii}(z)x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t).
-  \end{equation}
-
-  Moreover, $x_i$ can be expanded in terms of it's ancestor's $v(t)$
-  components only:
-
-  \begin{equation}
-    \label{eqn:ancestor_expansion}
-    x_i(t) = \A_{ii}(z)v_i(t) + \sum_{\substack{k \in \anc{i} \\ k \ne i}}\A_{ik}(z)v_k(t),
-  \end{equation}
-
-  where $\A(z) = \sum_{\tau = 0}^\infty A(\tau)z^{-\tau}$ is the filter from
-  the Wold decomposition representation of $x(t)$, Equation
-  \eqref{eqn:wold}.
-\end{proposition}
-
-This statement is ultimately about the sparsity pattern in the Wold
-decomposition matrices $A(\tau)$ since
-$x_i(t) = \sum_{\tau = 0}^\infty \sum_{j = 1}^n A_{ij}(\tau)v_j(t -
-\tau)$.  The proposition states that if $j \not \in \anc{i}$ then
-$\A_{ij}(z) = 0$.
-
-\subsection{Pairwise Granger Causality}
-\label{sec:pwgc}
-Recall that Granger causality in general must be understood with
-respect to a particular universe of observations.  If $x_j \gc x_i$
-with respect to $x_{-k}$, it may not hold with respect to $x$.  For
-example, $x_k$ may be a common ancestor which when observed, completely
-explains the connection from $x_j$ to $x_i$.  In this section we study
-\textit{pairwise} Granger causality, and seek to understand when
-knowledge of pairwise relations is sufficient to deduce the true fully
-conditional relations of $\gcg$.
-
-\begin{definition}[Pairwise Granger causality]
-  We will say that $x_j$ pairwise Granger-causes $x_i$ and write
-  $x_j \pwgc x_i$ if $x_j$ Granger-causes $x_i$ with respect only to
-  $(x_i, x_j)$.
-\end{definition}
-
-This notion is of interest for a variety of reasons.  From a purely
-conceptual standpoint, we will see how the notion can in some sense
-capture the idea of ``flow of information'' in the underlying graph,
-in the sense that if $j \in \anc{i}$ we expect that $j \pwgc i$.  It may
-also be useful for reasoning about the conditions under which
-\textit{unobserved} components of $x(t)$ may or may not interfere with
-inference in the actually observed components.  Finally, motivated
-from a practical standpoint to analyze causation in large systems,
-practical estimation procedures based purely on pairwise causality
-tests are of interest since the computation of such pairwise relations
-is substantially easier.
-
-The following propositions are essentially lemmas used for the proof
-of the upcoming Proposition \ref{prop:ancestor_properties}, but
-remain relevant for providing intuitive insight into the problems at
-hand.
-
-\begin{proposition}
-  Consider distinct nodes $i, j$ in a Granger causality graph
-  $\gcg$.  If
-
-  \begin{enumerate}[label=(\alph*)]
-    \item{$j \not\in \anc{i}$ and $i \not\in \anc{j}$}
-    \item{$\anc{i}\cap\anc{j} = \emptyset$}
-  \end{enumerate}
-
-  then $\H_t^{(i)} \perp \H_t^{(j)}$, that is,
-  $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.  Moreover,
-  this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_t^{(i)}} = 0$.
-\end{proposition}
-
-\begin{remark}
-  It is possible for components of $x(t)$ to be correlated at some
-  time lags without resulting in pairwise causality.  For instance,
-  the conclusion $j \npwgc i$ of Proposition
-  \ref{prop:separated_ancestor_uncorrelated} will still hold even if
-  $i \in \anc{j}$, since $j$ cannot provide any information about $i$
-  that is not available from observing $i$ itself.
-\end{remark}
-
-\begin{proposition}
-  Consider distinct nodes $i, j$ in a Granger causality graph $\gcg$.
-  If
-
-  \begin{enumerate}[label=(\alph*)]
-    \item{$j \not\in \anc{i}$}
-    \item{$\anc{i}\cap\anc{j} = \emptyset$}
-  \end{enumerate}
-
-  then $j \npwgc i$.
-\end{proposition}
-
-The previous result can still be strengthened significantly; notice
-that it is possible to have some $k \in \anc{i} \cap \anc{j}$ where still
-$j \npwgc i$, an example is furnished by the three node graph
-$k \rightarrow i \rightarrow j$ where clearly
-$k \in \anc{i}\cap\anc{j}$ but $j \npwgc i$.  We must introduce the concept
-of a \textit{confounding} variable, which effectively eliminates the
-possibility presented in this example.
-
-\begin{definition}[Confounder]
-  A node $k$ will be referred to as a \textit{confounder} of nodes
-  $i, j$ (neither of which are equal to $k$) if
-  $k \in \anc{i} \cap \anc{j}$ and there exists a path
-  $\gcgpath{k}{i}$ not containing $j$, and a path $\gcgpath{k}{j}$ not
-  containing $i$. A simple example is furnished by the ``fork'' graph
-  $i \leftarrow k \rightarrow j$.
-\end{definition}
-
-\begin{proposition}
-  \label{prop:ancestor_properties}
-  If in a Granger causality graph $\gcg$ where $j \pwgc i$ then
-  $j \in \anc{i}$ or $\exists k \in \anc{i} \cap\anc{j}$ which is a
-  confounder of $(i, j)$.
-\end{proposition}
-
-\begin{remark}
-  The interpretation of this proposition is that for $j \pwgc i$ then
-  there must either be ``causal flow'' from $j$ to $i$
-  ($j \in \anc{i}$) or there must be a confounder $k$ through which
-  common information is received.
-\end{remark}
-
-An interesting corollary is the following:
-
-\begin{corollary}
-  If the graph $\gcg$ is a DAG then $j \pwgc i, i \pwgc j \implies \exists k \in \anc{i} \cap \anc{j}$ confounding $(i, j)$.
-\end{corollary}
-
-It seems reasonable to expect a converse of Proposition
-\ref{prop:ancestor_properties} to hold, i.e.
-$j \in \anc{i} \Rightarrow j \pwgc i$.  Unfortunately, this is not the
-case in general, as different paths through $\gcg$ can lead to
-cancellation (see Figure \ref{fig:diamond_cancellation}).  In fact, we
-do not even have $j \in \pa{i} \Rightarrow j \pwgc i$ (see Figure
-\ref{fig:lag_cancellation}).
-
-\begin{example}
-  \label{ex:diamond_cancellation}
-  Firstly, on $n = 4$ nodes, ``diamond'' shapes can lead to cancellation on paths of length 2:
-
-\begin{equation*}
-  x(t) =
-  \left[
-    \begin{array}{cccc}
-      0 & 0 & 0 & 0\\
-      a & 0 & 0 & 0\\
-      -a & 0 & 0 & 0\\
-      0 & 1 & 1 & 0\\
-    \end{array}
-  \right] x(t - 1) + v(t),
-\end{equation*}
-
-with $\E v(t) = 0,\ \E v(t)v(t - \tau)^\T = \delta_\tau I$.
-
-By directly calculating
-
-\begin{align*}
-  x_4(t) &= x_2(t - 1) + x_3(t - 1) + v_4(t)\\
-         &= ax_1(t - 2) + av_2(t - 1) - ax_1(t - 2) -av_3(t - 1) + v_4(t)\\
-         &= a(v_2(t - 1) - v_3(t - 1)) + v_4(t),
-\end{align*}
-
-we see that, since $v(t)$ is isotropic white noise, $1 \npwgc 4$.  The problem here is that there are multiple paths from $x_1$ to $x_4$.
-\end{example}
-
-\begin{figure}
+\begin{figure}[H]
 \centering
-  \caption{Examples illustrating the difficulty of obtaining a converse to Proposition \ref{prop:ancestor_properties}}
-  \begin{subfigure}[b]{0.35\textwidth}
-    \includegraphics[width=\linewidth]{example1.pdf}
-    \caption{Path cancellation: $j \in \anc{i} \nRightarrow j \pwgc i$}
-    \label{fig:diamond_cancellation}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.35\textwidth}
-    \includegraphics[width=\linewidth]{example2.pdf}
-    \caption{Cancellation from time lag: $j \in \pa{i} \nRightarrow j \pwgc i$}
-    \label{fig:lag_cancellation}
-  \end{subfigure}
+\framebox[6cm]{\rule{0cm}{0.5cm} Abstracts may contain figures}
+%\includegraphics[width =0.48\linewidth]{image} 
+\caption{Figure one caption.}
+\label{fig1}
 \end{figure}
 
-\begin{example}
-  \label{ex:lag_cancellation}
-  A second example on $n = 3$ nodes is also worth examining, in this case
-  cancellation is a result of differing time lags.
+ Please send your source file  and .pdf file to Alyssa Novelia (anovelia@uwaterloo.ca) or Isam Al-Darabsah (ialdarabsah@uwaterloo.ca)
 
-\begin{equation*}
-  x(t) =
-  \left[
-    \begin{array}{ccc}
-      0 & 0 & 0\\
-      -a & 0 & 0\\
-      0 & 1 & 0\\
-    \end{array}
-  \right] x(t - 1) +
-  \left[
-    \begin{array}{ccc}
-      0 & 0 & 0\\
-      0 & 0 & 0\\
-      a & 0 & 0\\
-    \end{array}
-  \right] x(t - 2) + v(t)
-\end{equation*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% BIBLIOGRAPHY
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{thebibliography}{00}
+\addcontentsline{toc}{chapter}{References}
+\setlength{\itemsep}{-1mm}
+\small
+\bibitem{ASY} K.T. Alligood, T.D. Sauer and J. A. Yorke. \textit{Chaos: An Introduction to Dynamical Systems}, New York, Springer-Verlag (1997).
 
-Then
-
-\begin{align*}
-  x_2(t) &= v_2(t) - ax_1(t - 1)\\
-  x_3(t) &= v_3(t) + x_2(t - 1) + ax_1(t - 2)\\
-  \Rightarrow x_3(t) &= v_2(t - 1) + v_3(t),
-\end{align*}
-
-and again $1 \npwgc 3$.
-\end{example}
-
-\subsection{Strongly Causal Graphs}
-\label{sec:strongly_causal_graphs}
-
-In this section and the next we will seek to understand when converse
-statements of Proposition \ref{prop:ancestor_properties} \textit{do}
-hold.  One possibility is to restrict the coefficients of the system
-matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead,
-we think it more meaningful to focus on the defining feature of
-time series networks, that is, the topology of $\gcg$.
-
-\begin{definition}[Strongly Causal]
-  \label{def:strongly_causal}
-  We will say that a Granger causality graph $\gcg$ is
-  \textit{strongly causal} if there is at most one directed path between
-  any two nodes.  Strongly Causal Graphs will be referred to as SCGs.
-\end{definition}
-
-Examples of strongly causal graphs include directed trees (or
-forests), DAGs where each node has at most one parent, and Figure
-\ref{fig:example_fig3} of this paper.  A complete bipartite graph with
-$2n$ nodes is also strongly causal, demonstrating that the number of
-edges of such a graph can still scale quadratically with the number of
-nodes.  It is evident that the strong causal property is inherited by
-subgraphs.
-
-\begin{example}
-  Though examples of SCGs are easy to construct in theory, should
-  practitioners expect SCGs to arise in application?  While a positive
-  answer to this question is not \textit{necessary} for the concept to
-  be useful, it is certainly sufficient.  Though the answer is likely
-  to depend upon the particular application area, examples appear to
-  be available in biology, in particular, the authors of
-  \cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty}
-  cite an example of the so called ``transcription regulatory network
-  of \textit{E.coli}'', and
-  \cite{learning_genome_scale_regulatory_networks} study a much larger
-  regulatory network of \textit{Saccharomyces cerevisiae}.  These
-  networks, which we reproduce in Figure \ref{fig:gene_networks},
-  appear to have at most a small number of edges which violate the
-  strong-causality condition.
-
-  \begin{figure}[h]
-    \centering
-    \caption{Transcription Regulatory Networks}
-    \label{fig:gene_networks}
-    \begin{subfigure}[b]{0.45\textwidth}
-      \caption{\textit{E.Coli} Network of
-        \cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty}}
-      \label{fig:gene_network1}
-      \includegraphics[width=\linewidth, height=\linewidth]{ecoli_regulatory_network.pdf}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.45\textwidth}
-      \caption{\textit{Saccharomyces cerevisiae} Network of
-        \cite{learning_genome_scale_regulatory_networks}}
-      \label{fig:gene_network2}
-      \includegraphics[width=\linewidth, height=\linewidth]{huge_gene_network.pdf}
-    \end{subfigure}
-
-    {\footnotesize Figure \ref{fig:gene_network1} has only one edge
-      violating the strong-causality assumption, and Figure
-      \ref{fig:gene_network2} appears qualitatively to be nearly
-      strongly causal.  In particular, most of the edges are emanating
-      from network hubs, and many of the other edges are colliders.
-
-      Figure \ref{fig:gene_network1}
-      is reproduced under the Creative Commons Attribution
-      Non-Commercial License
-      (http://creativecommons.org/licenses/by-nc/2.5) and Figure
-      \ref{fig:gene_network2} under the Creative Commons Attribution
-      License (https://creativecommons.org/licenses/by/4.0/)}
-  \end{figure}
-\end{example}
-
-For later use, and to get a feel for the topological implications of
-strong causality, we explore a number of properties of such graphs
-before moving into the main result of this section.  The following
-important property essentially strengthens Proposition
-\ref{prop:ancestor_properties} for the case of strongly causal graphs.
-
-\begin{proposition}
-  \label{prop:sc_graph_common_anc}
-  In a strongly causal graph if $j \in \anc{i}$ then any
-  $k \in \anc{i} \cap \anc{j}$ is not a confounder, that is,
-  the unique path from $k$ to $i$ contains $j$.
-\end{proposition}
-
-\begin{corollary}
-  \label{cor:parent_corollary}
-  If $\gcg$ is a strongly causal DAG then $i \pwgc j$ and $j \in \anc{i}$ are
-  \textit{alternatives}, that is $i \pwgc j \Rightarrow j \notin \anc{i}$.
-\end{corollary}
-\begin{proof}
-  Suppose that $i \pwgc j$ and $j \in \anc{i}$.  Then since $\gcg$ is
-  acyclic $i \not\in \anc{j}$, and by Proposition
-  \ref{prop:ancestor_properties} there is some
-  $k \in \anc{i}\cap\anc{j}$ which is a confounder.  However, by
-  Proposition \ref{prop:sc_graph_common_anc} $k$ cannot be a
-  confounder, a contradiction.
-\end{proof}
-% This is a direct proof
-  % \begin{proof}
-%   Suppose that $\gcg$ is a strongly causal DAG and that we have both
-%   $i \pwgc j$ and $j \in \anc{i}$, which implies that
-%   $i \not \in \anc{j}$ since $\gcg$ is a DAG.  We will establish the
-%   contradiction $i \in \anc{j}$.
-
-%   Since $j \in \anc{i}$ there is a path $\gcgpath{j}{i}$.
-%   Moreover, since $i \pwgc j$ by Proposition \ref{prop:pwgc_anc} there
-%   must be a confounding ancestor $k \in \anc{i} \cap \anc{j}$, where the
-%   node $k$ has a path to $j,\ \gcgpath{k}{j}$, as well as a path
-%   to $i,\ \gcgpath{k}{i}$.
-
-%   \hl{Double check that the ancestor properties Proposition implies
-%     that BOTH paths do not contain the other node.}
-
-%   Now, since the graph is strongly causal, the only possible
-%   $\gcgpath{k}{i}$ path is the one obtained by concatenating the
-%   $\gcgpath{k}{j}$ path with the $\gcgpath{j}{i}$ path.
-%   However, this is a contradiction since $k$ is a confounder and
-%   should have a $\gcgpath{k}{i}$ path which does not contain $i$.
-% \end{proof}
-
-\begin{corollary}
-  \label{cor:bidirectional_edge}
-  If $\gcg$ is a strongly causal DAG such that $i \pwgc j$ and
-  $j \pwgc i$, then $i \not\in \anc{j}$ and $j \not\in \anc{i}$.  In
-  particular, a pairwise bidirectional edge indicates the absence of
-  any edge in $\gcg$.
-\end{corollary}
-\begin{proof}
-  This follows directly from applying Corollary
-  \ref{cor:parent_corollary} to $i \pwgc j$ and $j \pwgc i$.
-% This is a direct proof
-% \begin{proof}
-  % By way of contradiction, suppose that $i \in \pa{j}$.  We will
-  % consider the two possibilities allowed by proposition
-  % \ref{prop:ancestor_properties} for $j \pwgc i$.  Firstly
-  % $j \in \anc{i}$ is impossible since $\gcg$ is assumed to be acyclic.
-  % Secondly, if there is some confounding node
-  % $u \in \anc{i} \cap \anc{j}$ with a path
-  % $u \rightarrow \cdots \rightarrow j$ which does not contain $i$ we
-  % have a contradiction since there must now be multiple
-  % $u \rightarrow j$ paths: the aforementioned, and a path
-  % $u \rightarrow \cdots \rightarrow i \rightarrow \cdots \rightarrow
-  % j$ which \textit{does} contain $i$.  We conclude that $i \in \pa{j}$
-  % is impossible, and symmetrically that $j \in \pa{i}$ is as well.
-% \end{proof}
-\end{proof}
-
-In light of Proposition \ref{prop:sc_graph_common_anc}, the following
-provides a partial converse to Proposition
-\ref{prop:ancestor_properties}, and supports the intuition of ``causal
-flow'' through paths in $\gcg$.
-
-\begin{proposition}
-  \label{prop:pwgc_anc}
-  If $\gcg$ is a strongly causal DAG then $j \in \anc{i} \Rightarrow j \pwgc i$.
-\end{proposition}
-
-We immediately obtain the corollary, which we remind the reader is,
-surprisingly, not true in a general graph.
-
-\begin{corollary}
-  \label{cor:gc_implies_pwgc}
-  If $\gcg$ is a strongly causal DAG then $j \gc i \Rightarrow j \pwgc i$.
-\end{corollary}
-
-\begin{example}
-  As a final remark of this subsection we note that a complete
-  converse to Proposition \ref{prop:ancestor_properties} is not
-  possible without additional conditions.  Consider the ``fork''
-  system on $3$ nodes (i.e. $2 \leftarrow 1 \rightarrow 3$) defined by
-
-  \begin{equation*}
-    x(t) =
-    \left[
-      \begin{array}{cccc}
-        0 & 0 & 0\\
-        a & 0 & 0\\
-        a & 0 & 0\\
-      \end{array}
-    \right] x(t - 1) + v(t).
-  \end{equation*}
-
-  In this case, node $1$ is a confounder for nodes $2$ and $3$, but
-  $x_3(t) = v_3(t) - v_2(t) + x_2(t)$ and $2 \npwgc 3$ (even
-  though $x_2(t)$ and $x_3(t)$ are contemporaneously correlated)
-
-  If we were to augment this system by simply adding an autoregressive
-  component (i.e. some ``memory'') to $x_1(t)$ e.g.
-  $x_1(t) = v_1(t) + b x_1(t - 1)$ then we \textit{would} have
-  $2 \pwgc 3$ since then
-  $x_3(t) = v_3(t) + av_1(t - 1) - bv_2(t - 1) + bx_2(t - 1)$.  We
-  develop this idea further in the next section.
-\end{example}
-
-\subsection{Persistent Systems}
-\label{sec:persistent_systems}
-In section \ref{sec:strongly_causal_graphs} we obtained a converse to
-part $(a)$ of Proposition \ref{prop:ancestor_properties} via the
-notion of a strongly causal graph topology (see Proposition
-\ref{prop:pwgc_anc}).  In this section, we study conditions under
-which a converse to part $(b)$ will hold.
-
-\begin{definition}[Lag Function]
-  Given a causal filter $\B(z) = \sum_{\tau = 0}^\infty b(\tau)z^{-\tau}$
-  define
-
-  \begin{align}
-    \tau_0(\B) &= \text{inf }\{\tau \in \Z_+\ |\ b(\tau) \ne 0\},\\
-    \tau_{\infty}(\B) &= \text{sup }\{\tau \in \Z_+\ |\ b(\tau) \ne 0\}.
-  \end{align}
-
-  i.e. the ``first'' and ``last'' coefficients of the filter $\B(z)$,
-  where $\tau_\infty(\B) \defeq \infty$ if the filter has an infinite
-  length, and $\tau_0(\B) \defeq \infty$ if $\B(z) = 0$.
-\end{definition}
-
-\begin{definition}[Persistent]
-  We will say that the process $x(t)$ with Granger causality graph
-  $\gcg$ is \textit{persistent} if for every $i \in [n]$ and every
-  $k \in \anc{i}$ we have $\tau_0(\A_{ik}) < \infty$ and $\tau_\infty(\A_{ik}) = \infty$.
-\end{definition}
-
-\begin{remark}
-  In the context of Granger causality, ``most'' systems should be
-  persistent.  In particular, $\mathsf{\VAR}(p)$ models are likely to
-  be persistent since these naturally result in an equivalent
-  $\mathsf{MA}(\infty)$ representation, see Example
-  \ref{ex:persistent_system}.
-
-  Moreover, persistence is not the weakest condition necessary for the
-  results of this section, the condition that for each $i, j$ there is
-  some $k \in \anc{i}\cap\anc{j}$ such that
-  $\tau_0(\A_{jk}) < \tau_\infty(\A_{ik})$ is enough.  The intuition
-  being that nodes $i$ and $j$ are not receiving temporally disjoint
-  information from $k$.
-
-  The etymology for the persistence condition can be explained by
-  supposing that the two nodes $i, j$ each have a loop (i.e.
-  $\B_{ii}(z) \ne 0, \B_{jj}(z) \ne 0$) then this autoregressive
-  component acts as ``memory'', and so the influence from the
-  confounder $k$ \textit{persists} in $x_i(t)$, and
-  $\tau_\infty(\A_{ik}) = \infty$ for each confounder is expected.
-\end{remark}
-
-\begin{example}
-  \label{ex:persistent_system}
-  Consider a process $x(t)$ generated by the $\VAR(1)$
-  model\footnote{Recall that any $\VAR(p)$ model with $p < \infty$ can
-    be written as a $\VAR(1)$ model, so we lose little generality in
-    considering this case.}  having $\B(z) = Bz^{-1}$.  If $B$ is
-  diagonalizable, and has at least $2$ distinct eigenvalues, then
-  $x(t)$ is persistent.
-
-  See the supplementary material for an analysis of this example.
-\end{example}
-
-In order to eliminate the possibility of a particular sort of
-cancellation, an ad-hoc assumption is required.  Strictly speaking,
-the persistence condition is not a necessary or sufficient condition
-for the following, but cases where the following fails to hold, and
-persistence \textit{does} hold, are unavoidable pathologies.
-
-\begin{assumption}[no-cancellation]
-  \label{ass:T_causality}
-  Fix $i, j \in [n]$ and let $H_i(z)$ be the strictly-causal filter
-  such that
-
-  \begin{equation*}
-    H_i(z)x_i(t) = \linE{x_i(t)}{\H_{t - 1}^{(i)}},
-  \end{equation*}
-
-  and similarly for $H_j(z)$.  Then define
-
-  \begin{equation}
-    \label{eqn:T_filter}
-    T_{ij}(z) = \sum_{k \in \anc{i} \cap \anc{j}}\sigma_k^2\A_{ik}(z^{-1})(1 - H_i(z^{-1}))(1 - H_j(z))\A_{jk}(z),
-  \end{equation}
-
-  where $\sigma_k^2 = \E v_k(t)^2$.
-
-  We will say that Assumption \ref{ass:T_causality} is satisfied if
-  for every $i, j \in [n]$, $T_{ij}(z)$ is either constant over $z$
-  (i.e. each $z^k$ coefficient for $k \in \Z \setminus \{0\}$ is $0$),
-  or is \textit{neither} causal (i.e. containing only $z^{-k}$ terms,
-  for $k \ge 0$) \textit{or} anti-causal (i.e. containing only $z^k$
-  terms, for $k \ge 0$).  Put succinctly, $T_{ij}(z)$ must be
-  two-sided.
-\end{assumption}
-
-\begin{remark}
-  Under the condition of persistence, the only way for Assumption
-  \ref{ass:T_causality} to fail is through cancellation in the terms
-  defining $T_{ij}(z)$.  For example, the condition is assured if
-  $x(t)$ is persistent, and there is only a single confounder.
-  Unfortunately, some pathological behaviour resulting from
-  confounding nodes seems to be unavoidable without some assumptions
-  about the parameters of the $\MA(\infty)$ system defining $x(t)$.
-\end{remark}
-
-\begin{proposition}
-  \label{prop:persistence_converse}
-  Fix $i, j \in [n]$ and suppose $\exists k \in \anc{i} \cap \anc{j}$
-  which confounds $i, j$.  Then, if $T_{ij}(z)$ is not causal we have
-  $j \pwgc i$, and if $T_{ij}(z)$ is not anti-causal we have
-  $i \pwgc j$.  Moreover, if Assumption \ref{ass:T_causality} is
-  satisfied, then $j \pwgc i \iff i \pwgc j$.
-  % Suppose that Assumption \ref{ass:T_causality} is satisfied.  Then if
-  % there exists a $k$ which confounds $(i, j)$ we have
-  % $i \pwgc j \implies j \pwgc i$.  Moreover, if $T_{ij}(z)$ is not
-  % constant, then $i \pwgc j$.
-\end{proposition}
-
-\begin{remark}
-  The importance of this result is that when $i \pwgc j$ is a result
-  of a confounder $k$, then $i \pwgc j \iff j \pwgc i$.  This
-  implies that in a strongly causal graph every bidirectional pairwise
-  causality relation must be the result of a confounder.  Therefore,
-  in a strongly causal graph, pairwise causality analysis is
-  \textit{immune to confounding} (since we can safely remove all
-  bidirectional edges).
-\end{remark}
-
-\subsection{Recovering $\gcg$ via Pairwise Tests}
-\label{sec:pairwise_algorithm}
-We arrive at the main conclusion of the theoretical analysis in this
-paper.
-
-\begin{theorem}[Pairwise Recovery]
-  \label{thm:scg_recovery}
-  If the Granger causality graph $\gcg$ for the process $x(t)$ is a
-  strongly causal DAG and Assumption \ref{ass:T_causality} holds, then
-  $\gcg$ can be inferred from pairwise causality tests.  The procedure
-  can be carried out, assuming we have an oracle for pairwise
-  causality, via Algorithm (\ref{alg:pwgr}).
-\end{theorem}
-
-\begin{algorithm}
-  \SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{output}
-  \SetKwInOut{Initialize}{initialize}
-  \DontPrintSemicolon
-
-  % \BlankLine
-  \caption{Pairwise Granger Causality Algorithm}
-  \label{alg:pwgr}
-  \TitleOfAlgo{Pairwise Graph Recovery}
-  \Input{Pairwise Granger causality relations between a persistent
-  process of dimension $n$ whose joint Granger causality
-  relations are known to form a strongly causal DAG $\gcg$.}
-  % \Input{Pairwise Granger causality relations}
-  \Output{Edges $\gcge = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of
-    the graph $\gcg$.}
-  \Initialize{$S_0 = [n]$  \texttt{\# unprocessed nodes}\\
-    $E_0 = \emptyset$  \texttt{\# edges of }$\gcg$\\
-    % $P_0 = \emptyset$  \texttt{\# layer by layer driving nodes}\\
-    $k = 1$ \texttt{\# a counter used only for notation}}
-  \BlankLine
-  $W \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$  \texttt{\# candidate edges}\\
-  $P_0 \leftarrow \{i \in S_0\ |\ \forall s \in S_0\ (s, i) \not\in W\}$  \texttt{\# parent-less nodes}\\
-  \While{$S_{k - 1} \ne \emptyset$}{
-    $S_k \leftarrow S_{k - 1} \setminus P_{k - 1}$ \texttt{\# remove nodes with depth }$k - 1$\\
-    $P_k \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W\}$   \texttt{\# candidate children}\\
-
-    $D_{k0} \leftarrow \emptyset$\\
-    \For{$r = 1, \ldots, k$} 
-    {
-      $Q \leftarrow E_{k - 1} \cup \big(\bigcup_{\ell = 0}^{r - 1} D_{k\ell}\big)$ \texttt{\# currently known edges}\\
-      $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times P_k\ |\ (i, j) \in W,\ \text{no } i \rightarrow j \text{ path in } Q\}$
-      % $E_k \leftarrow E_{k - 1} \cup D_{kr}$ \texttt{\# add edges to }$E_k$ \label{alg:inner_loop_end}\\
-      % $W_{k + 1} \leftarrow W_k \setminus D_{kr}$  \texttt{\# remove edges from consideration}\\
-    }
-    $E_k \leftarrow E_{k - 1} \cup \big(\bigcup_{r = 1}^k D_{kr}\big)$ \texttt{\# update } $E_k$ \texttt{ with new edges}\\
-    % \label{alg_line:inner_loop_end}\\
-    % $W_{k + 1} \leftarrow W_k \setminus \big(\bigcup_{r = 0}^k D_{kr}\big)$  \texttt{\# remove edges from consideration}\\
-    $k \leftarrow k + 1$
-  }
-  \Return{$E_{k - 1}$}
-\end{algorithm}
-
-The theorem is proven in the supplementary material by establishing
-the correctness of Algorithm (\ref{alg:pwgr}).  The idea is to
-iteratively ``peel away layers'' of nodes by removing the nodes that
-have no parents remaining.  The requirement of strong causality
-ensures that all actual edges of $\gcg$ manifest in some way as
-pairwise relations (by Proposition \ref{prop:pwgc_anc}), and the
-no-cancellation condition of Assumption \ref{ass:T_causality} allows
-confounding to be eliminated by removing bidirectional edges (by
-Proposition \ref{prop:persistence_converse} and Corollary
-\ref{cor:bidirectional_edge}).  Without Assumption
-\ref{ass:T_causality}, then each confounded pair would give rise to
-$4$ possible pairwise topologies consistent with $\gcg$, one for each
-type of pairwise edge (no edge, unidirectional, bidirectional).
-
-\begin{example}
-  The set $W$ collects ancestor relations in $\gcg$ (see Lemma
-  \ref{lem:W_subset_E}).  In reference to Figure
-  \ref{fig:example_fig3}, each of the solid black edges, as well as
-  the dotted red edges will be included in $W$, but \textit{not} the
-  bidirectional green dash-dotted edges, which we are able to exclude
-  as results of confounding.  The groupings $P_0, \ldots, P_3$ are also
-  indicated in Figure \ref{fig:example_fig3}.
-
-  The algorithm proceeds first with the parent-less nodes $1, 2$ on the
-  initial iteration where the edge $(1, 3)$ is added to $E$.  On the
-  next iteration, the edges $(3, 4), (2, 4), (3, 5)$ are added, and
-  the false edges $(1, 4), (1, 5)$ are excluded due to the paths
-  $1 \rightarrow 3 \rightarrow 4$ and $1 \rightarrow 3 \rightarrow 5$
-  already being present.  Finally, edge $(4, 6)$ is added, and the false
-  $(1, 6), (3, 6), (2, 6)$ edges are similarly excluded due to the
-  ordering of the inner loop.
-
-  \begin{figure}
-    \centering
-    \caption{Example graph for Algorithm \ref{alg:pwgr}}
-    {\footnotesize{Black arrows indicate true parent-child
-        relations.  Red dotted arrows indicate pairwise causality (due to
-        non-parent relations), green dash-dotted arrows indicate
-        bidirectional pairwise causality (due to the confounding node
-        $1$).  Blue groupings indicate each $P_k$ in Algorithm
-        \ref{alg:pwgr}.}}
-    \label{fig:example_fig3}
-
-    \begin{subfigure}[b]{0.45\textwidth}
-      \includegraphics[width=\linewidth]{example_algorithm.pdf}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.45\textwidth}
-      \includegraphics[width=\linewidth]{example_algorithm2.pdf}
-    \end{subfigure}
-  \end{figure}
-
-  That we need to proceed backwards through $P_{k - r}$ as in the
-  inner loop on $r$ can also be seen from this example, where if
-  instead we simply added the set
-
-  \begin{equation*}
-    D_k' = \{(i, j) \in \Big(\bigcup_{r = 1}^k P_{k - r}\Big) \times P_k\ |\ i \pwgc j \}
-  \end{equation*}
-
-  to $E_k$ then we would infer the false positive edge
-  $1 \rightarrow 4$.  Moreover, the same example shows that simply
-  using the set
-
-  \begin{equation*}
-    D_k'' = \{(i, j) \in P_{k - 1} \times P_k\ |\ i \pwgc j \}  ,
-  \end{equation*}
-
-  causes the edge $1 \rightarrow 3$ to be missed.
-\end{example}
-
-% \begin{example}
-%   We close this section by noting that the conditions of persistence
-%   and strong causality are only sufficient conditions.  For example,
-%   the complete directed graph with 2 nodes i.e.
-
-%   \begin{equation*}
-%     B(1) = \left[ \begin{array}{cc} 1/2 & 1 \\ 1 & 1/2 \end{array}\right]
-%   \end{equation*}
-
-%   contains a loop but is pairwise recoverable, though not by algorithm
-%   (\ref{alg:pwgr}).  Clearly, this example is somewhat artificial
-%   since when $n = 2$ there is no difference between pairwise
-%   Granger causality and joint Granger causality amongst all series --
-%   however, one can add any number of nodes having no parents or
-%   children to a graph containing a length 2 cycle, in which case the
-%   graph clearly remains pairwise recoverable.
-% \end{example}
-
-\section{Simulation}
-\label{sec:simulation}
-We implement an heuristic inspired by Algorithm \ref{alg:pwgr} by
-replacing the population statistics with finite sample tests, the
-details of which can be found in the supplementary material Section
-\ref{sec:finite_pwgc} (see Algorithm \ref{alg:finite_pwgc}).  The
-heuristic is essentially controlling the false discovery rate
-substantially below what it would be with a threshold based pairwise
-scheme.  The methods are easily parallelizable, and can scale to
-graphs with thousands of nodes on a single machine.  By contrast,
-scaling the LASSO to this large of a network (millions of variables)
-is nontrivial and extremely computationally demanding.
-
-We run experiments using two separate graph topologies having $n = 50$
-nodes: a strongly causal graph (SCG) and a directed acyclic graph
-(DAG).  Consult Section \ref{apx:simulation} for details on how data
-is generated from these models.
-
-We compare our results against the adaptive LASSO
-\cite{adaptive_lasso_zou2006}, which outperformed both the LASSO and
-the grouped LASSO by a large margin.  Motivated by scaling, we split the squared error
-term into separate terms, one for each group of incident edges on a
-node, and estimate the collection of $n$ incident filters
-$\big\{\B_{ij}(z)\big\}_{j = 1}^n$ that minimizes
-$\xi_i^{\text{LASSO}}$ in the following:
-
-\begin{equation}
-  \begin{aligned}
-  \xi_i^{\text{LASSO}}(\lambda) &= \underset{B}{\text{min}}\ \frac{1}{T}\sum_{t = p_{\text{max}} + 1}^T\big(x_i(t) - \sum_{\tau = 1}^{p_{\text{max}}}\sum_{j = 1}^n B_{i, j}(\tau) x(t - \tau)\big)^2 + \lambda \sum_{\tau = 1}^{p_{\text{max}}} \sum_{j = 1}^n |B_{ij}(\tau)|\\
-  \xi_i^{\text{LASSO}} &= \underset{\lambda \ge 0}{\text{min}}\ \xi_i^{\text{LASSO}}(\lambda) + \mathsf{BIC}\big(B_i^{\text{LASSO}}(\lambda)\big)\\
-  \end{aligned}
-\end{equation}
-
-where we are choosing $\lambda$, the regularization parameters, via the BIC.
-This is similar to the work of \cite{arnold2007temporal}, except that
-we have replacing the LASSO with the Adaptive LASSO, which provides
-dramatically superior performance.
-
-\begin{figure}
-  \centering
-  \caption{Representative Random Graph Topologies on $n = 50$ Nodes}
-  \label{fig:random_graph_topologies}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \caption{SCG}
-    \includegraphics[width=\linewidth]{example_scg.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \caption{DAG $(q = \frac{2}{n})$}
-    \includegraphics[width=\linewidth]{example_dag.pdf}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \caption{DAG $(q = \frac{4}{n})$}
-    \includegraphics[width=\linewidth]{example_dag_q.pdf}
-  \end{subfigure}
-\end{figure}
-
-\begin{remark}[Graph Topologies]
-  We depict in Figure \ref{fig:random_graph_topologies} the topologies
-  of random graphs used in our empirical evaluation.  For values of
-  $q$ close to $\frac{2}{n}$, the resulting random graphs tend to have
-  a topology which is, at least qualitatively, close to the SCG.  As
-  the value of $q$ increases, the random graphs deviate farther from
-  the SCG topology, and we therefore expect the LASSO to outperform
-  PWGC for larger values of $q$.
-\end{remark}
-
-\begin{remark}[MCC as a Support Recovery Measurement]
-  We apply Matthew's Correlation Coefficient (MCC)
-  \cite{matthews1975comparison} as a statistic for measuring support
-  recovery performance (see also \cite{chicco2017ten} tip \# 8).  This
-  statistic synthesizes the confusion matrix into a single score
-  appropriate for unbalanced labels and is calibrated to fall into the
-  range $[-1, 1]$ with $1$ being perfect performance, $0$ being the
-  performance of random guessing, and $-1$ being perfectly opposed.
-\end{remark}
-
-\begin{remark}[Error Measurement]
-  We estimate the 1-step ahead prediction error by forming the variance matrix estimate
-
-  \begin{equation*}
-    \widehat{\Sigma}_v \defeq \frac{1}{T_{\text{out}}} \sum_{t = 1}^{T_{\text{out}}} (x(t) - \widehat{x}(t))(x(t) - \widehat{x}(t))^\T
-  \end{equation*}
-
-  on a long stream of out-of-sample data.  We then report the quantity
-
-  \begin{equation*}
-    \frac{\ln \tr \widehat{\Sigma}_v}{\ln \tr \Sigma_v}
-  \end{equation*}
-
-  where $\widehat{\Sigma}_v = \Sigma_v$ is the best possible performance.
-\end{remark}
-
-\subsection{Results}
-\begin{figure}[!h]
-  \centering
-  \caption{Simulation Results: PWGC vs AdaLASSO}
-  \label{tab:simulation_table}
-
-  \begin{tabular}{|ll||ll|ll|ll|}
-    \toprule
-    &\textbf{Algorithm}&alasso&pwgc&alasso&pwgc&alasso&pwgc\\
-    &\textbf{Metric}&LRE&&FDP&&MCC&\\
-    \textbf{T}&\textbf{q}&&&&&&\\
-    \midrule
-    \multirow{4}{*}{\textbf{50}}
-    &\textbf{SCG}&1.71&\textbf{1.55}&0.52&\textbf{0.08}&0.46&\textbf{0.55}\\
-    &\textbf{0.04}&1.97&\textbf{1.77}&0.57&\textbf{0.10}&0.41&\textbf{0.53}\\
-    &\textbf{0.08}&2.95&\textbf{2.72}&0.50&\textbf{0.23}&0.36&\textbf{0.39}\\
-    &\textbf{0.32}&9.02&\textbf{8.17}&\textbf{0.53}&0.56&\textbf{0.14}&0.10\\
-    \midrule
-    \multirow{4}{*}{\textbf{250}}
-    &\textbf{SCG}&1.30&\textbf{1.18}&0.29&\textbf{0.06}&0.70&\textbf{0.81}\\
-    &\textbf{0.04}&1.40&\textbf{1.31}&0.30&\textbf{0.07}&0.68&\textbf{0.76}\\
-    &\textbf{0.08}&2.49&\textbf{2.21}&0.32&\textbf{0.16}&0.55&\textbf{0.57}\\
-    &\textbf{0.32}&8.67&\textbf{7.62}&0.48&0.46&\textbf{0.18}&0.15\\
-    \midrule
-    \multirow{4}{*}{\textbf{1250}}
-    &\textbf{SCG}&1.20&\textbf{1.11}&0.41&\textbf{0.07}&0.68&\textbf{0.88}\\
-    &\textbf{0.04}&1.28&\textbf{1.20}&0.46&\textbf{0.07}&0.64&\textbf{0.84}\\
-    &\textbf{0.08}&2.12&2.05&0.36&\textbf{0.14}&0.60&\textbf{0.64}\\
-    &\textbf{0.32}&7.78&\textbf{7.39}&0.49&\textbf{0.37}&\textbf{0.21}&0.18\\
-    \bottomrule
-  \end{tabular}
-
-  {\footnotesize Results of Monte Carlo simulations comparing PWGC and
-    AdaLASSO $(n = 50, p = 5, p_{\text{max}} = 10)$ for small
-    samples and when the SCG assumption doesn't hold.  The superior
-    result is bolded when the difference is statistically
-    significant, as measured by \texttt{scipy.stats.ttest\_rel}.
-    100 iterations are run for each
-    set of parameters.\\
-
-    LRE: Log-Relative-Error, i.e. the log sum of squared errors at
-    each node relative to the strength of the driving noise
-    $\frac{\ln \tr \widehat{\Sigma}_v}{\ln \tr \Sigma_v}$.  FDP: False
-    Discovery Proportion.  MCC: Matthew's Correlation Coefficient.\\
-
-    Values of $q$ (edge probability) range between
-    $2 / n, 4 / n, 16 / n$ where $2 / n$ has the property that the
-    random graphs have on average the same number of edges as the
-    SCG.}
-\end{figure}
-
-Our simulation results are summarized in Table
-\ref{tab:simulation_table}, with additional figures provided in the
-supplementary material Section \ref{apx:simulation}.  It is clear that
-the superior performance of PWGC in comparison to AdaLASSO is as a
-result of limiting the false discovery rate.  It is unsurprising that
-PWGC exhibits superior performance when the graph is an SCG, but even
-in the case of more general DAGs, the PWGC heuristic is still able to
-more reliably uncover the graph structure for small values of $q$.  We
-would conjecture that for small $q$, random graphs are ``likely'' to
-be ``close'' to SCGs in some appropriate sense.  As $q$ increases,
-there are simply not enough edges allowed by the SCG topology for it
-to be possible to accurately recover $\gcg$.
-
-Interestingly, we can observe that the AdaLASSO appears to perform
-marginally better on strongly causal graphs than directed acyclic
-graphs with an equivalent number of edges ($q \approx 0.04$ is chosen for
-this purpose).  This provides supporting evidence for one of the main
-assertions of this work: that the topological structure of $\gcg$ is
-an important distinguishing feature of time series networks in
-comparison to classical multivariate regression where it is only the
-sparsity rate which is considered.
-
-\section{Conclusion}
-\label{sec:conclusion}
-In this paper we have argued that considering particular topological
-properties of Granger causality networks can provide substantial
-insights into the structure of causality graphs with potential for
-providing improvements to causality graph estimation when structural
-assumptions are met.  In particular, the notion of a strongly-causal
-graph has been exploited to establish conditions under which pairwise
-causality testing alone is sufficient for recovering a complete
-Granger causality graph.  Moreover, examples from the literature
-suggest that such topological assumptions may be reasonable in some
-applications.  And secondly, even when the strong-causality assumption
-is not met, we have provided simulation evidence to suggest that our
-pairwise testing algorithm PWGC can still outperform the LASSO and
-adaLASSO, both of which are commonly employed in applications.
-
-We emphasize that the causality graph topology is one of the key
-defining features of time series analysis in comparison to standard
-multivariate regression and therefore advocate for further study of
-how different topological assumptions may impact the recovery of
-causality graphs.  For example, are there provable guarantees on the
-error rate of PWGC when applied to non strongly-causal graphs?  Can
-constraint systems or cunning adaptive weighting schemes impose useful
-prior knowledge about graph topology for the LASSO algorithm?
-Finally, the work of \cite{barnett2015granger} has established the
-superiority of Granger causality testing by state space models (as
-opposed to pure autoregressions) in many cases.  Combining this work
-with our PWGC algorithm (by modifying the approach described in
-Section \ref{sec:pairwise_hypothesis_testing} to instead utilize
-state-space Granger causality testing) therefore is likely to enable
-application to very large networks of time series data which are not
-well approximated by finite $\VAR(p)$ models.  Moreover, our
-heuristics are in principle applicable in a model-free context.  As
-long as a primitive for testing the pairwise causation between two
-components of a multivariate time series is available, our methods may
-be useful.
-
-\bibliography{paper_bib}
+\bibitem{STE} L. Schimansky-Geier, A.V. Tolstopjatenko and W. Ebelin \textit{Noise induced transitions due to external additive noise}, Phys. Lett. A, \textbf{108}, 7, pp. 329-332 (1985).
+\end{thebibliography}
+\normalsize
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2019-09-30 presentations/student_dynamical_systems_waterloo_2019/Abstract-Template.tex
abstract for dynamic systems workshop

@@
 
+% \usepackage{graphicx}
+% \usepackage{caption}
+\usepackage{subcaption}
+\newtheorem{theorem}{Theorem}
+
+\def\gcg{\mathcal{G}}  % Granger-causality graph
+
 \begin{document}
@@
 \begin{flushleft}
-\Large \textbf{\noindent
-Insert the Title of Your Talk Here}
+  \Large \textbf{\noindent Graph Topological Aspects of Granger
+    Causality in Stationary Linear Dynamical Systems}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We study Granger causality in the context of wide-sense stationary
-time series, where our focus is on the topological aspects of the
-underlying causality graph.  We establish sufficient conditions (in
-particular, we develop the notion of a ``strongly causal'' graph
-topology) under which the true causality graph can be recovered via
-pairwise causality testing alone, and provide examples from the gene
-regulatory network literature suggesting that our concept of a
-strongly causal graph may be applicable to this field.  We implement
-and detail finite-sample heuristics derived from our theory, and
-establish through simulation the efficiency gains (both statistical
-and computational) which can be obtained (in comparison to LASSO-type
-algorithms) when structural assumptions are met.
+The concept of Granger causality, a notion introduced in 1969
+\cite{granger1969investigating} for the analysis of macroeconomic data
+(e.g. ``does lowering the interest rate cause an increase in
+employment?''), attempts to capture the intuition that a cause must
+proceed it's effect, and has spurred both theoretical and applied
+research.  The concept has been applied in diverse settings, for
+example: in finance, researchers have attempted to quantify the
+interdependence of financial institutions \cite{NBERw16223} and in the
+biological sciences, the concept is used to map out the promoting and
+regulating effects of certain genes upon others
+\cite{methods_for_inferring_gene_regulatory_networks_from_time_series_expression_data}.
 
-In this paper we study the notion of Granger causality
-\cite{granger1969investigating} \cite{Granger1980329} as a means of
-uncovering an underlying causal structure in multivariate time series.
-Though the underlying causality graph cannot be observed directly,
-it's presence is inferred as a latent structure among observed
-time series data.  This notion is leveraged in a variety of
-applications e.g. in Neuroscience as a means of recovering
-interactions amongst brain regions \cite{bressler2011wiener},
-\cite{anna_paper2008}, \cite{david2008identifying}; in the study of
-the dependence and connectedness of financial institutions
-\cite{NBERw16223}; gene expression networks \cite{Fujita2007},
-\cite{methods_for_inferring_gene_regulatory_networks_from_time_series_expression_data},
-\cite{grouped_graphical_granger_modelling_for_gene_expression_regulatory_networks_discovery},
-\cite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty};
-and power system design \cite{Misyrlis2016450}, \cite{yuan2014root}.
+A substantial portion of the work on this topic is devoted to
+generalizing the methods of Granger to nonlinear systems.  For
+example, linear Granger causality is a special case of the information
+theoretic notion of transfer entropy \cite{barnett2009granger}.
+However, like many attempts at quantifying causation, transfer entropy
+suffers from some difficulties of interpretation
+\cite{transfer_entropy_criticism}.  Although the criticisms raised by
+\cite{transfer_entropy_criticism} do not arise in the case of linear
+Granger causality, other pathologies are known.  Our work sheds light
+on some of the challenges that remain in the linear case, and in
+particular focuses on topological particularities of the underlying
+network of interactions.
 
-The principal contributions of this paper are as follows: firstly, in
-section \ref{sec:theory} we study \textit{pairwise} Granger causality
-relations, providing novel theorems connecting the structure of the
-causality graph to the pairwise ``causality flow'' in the system, as
-well as an interpretation in terms of the graph topology of the
-sparsity pattern of matrices arising in the Wold decomposition,
-generalizing in some sense the notion of ``feedback-free'' processes
-studied by \cite{caines1975feedback} in close connection with
-Granger causality.  We establish sufficient conditions (sections
-\ref{sec:strongly_causal_graphs}, \ref{sec:persistent_systems}) under
-which a fully conditional Granger causality graph can be recovered
-from pairwise tests alone (sec \ref{sec:pairwise_algorithm}).  We
-report a summary of simulation results in \ref{sec:simulation}, with
-additional results reported in the supplementary material Section
-\ref{apx:simulation}.  Our simulation results establish that there is
-significant potential for improvement over existing methods, and that
-the graph-topological aspects of time series analysis are relevant for
-both theory and practice.  Concluding remarks on further open problems
-and extensions are provided in Section \ref{sec:conclusion}.  The
-proofs of each proposition and theorem are also relegated to the
-supplementary material, simple corollaries have proofs included in the
-main text.
+In general, we are presented with a multitude of jointly wide sense
+stationary time series $x_1(t), \ldots, x_n(t)$ and seek to construct
+a (Granger-)causal graph $\gcg$ where there is an edge
+$(j, i) \in \gcg$ whenever $x_j$ Granger causes $x_i$ after
+conditioning on each $k \not\in \{i, j\}$.  Depending on the
+interpretation or on the application, the graph $\gcg$ captures the
+flows of ``information'' or ``energy'' throughout a network of
+interacting processes.  However, it may be the case that we observe
+$x_j$ and $x_i$ alone -- in such cases, simple examples establish that
+even if $x_j$ Granger causes $x_i$ conditioned on the rest of the
+observations, it need not be that $x_j$ Granger causes $x_i$ pairwise
+(i.e. without conditioning).
 
-In a general Granger causality graph, there are natural
-\textit{necessary} conditions for a node $j$ to pairwise cause node
-$i$, namely that $j$ is an ancestor to $i$ or that there is a third
-confounding node $k$.  On the other hand, we will see simple examples
-for why \textit{sufficient} conditions are more nuanced: if $j$ causes
-$i$ conditionally on all other nodes in the graph, it need not be the
-case that $j$ causes $i$ when the two are observed in isolation.  We
-examine a special \textit{graph topological} assumption we term
-``strongly causal'' wherein a natural converse \textit{does} hold,
-satisfying an intuitive notion of ``causality flow'' through the
-graph.  We highlight examples from the literature on gene networks
-suggesting that assumptions of strongly causal graph structure may not
-be out of place in some application areas.
+Pathologies of pairwise causation clash with our intuition that causal
+relations in $\gcg$ should be transitive, that is, that if there is
+some ``true'' (conditional) causation path
+$j \rightarrow a_0 \rightarrow \cdots \rightarrow a_r \rightarrow i$
+in $\gcg$ then it should hold that $x_j$ pairwise causes $x_i$.  While
+this does not hold in general, our work analyzes a particular type of
+graph topology, which we deem a ``strongly causal'' graph, wherein
+natural intuitions about causal flow \textit{do} hold.  Moreover, we
+prove that (with modest additional technical assumptions) that in such
+a graph the complete conditional causation graph $\gcg$ can be
+recovered by testing only pairwise interactions of nodes.  That is,
+the complete causal structure of $\gcg$ can be recovered through
+simple one way unconditional causality tests.
 
-As a consequence of the causality graph being strongly causal, we show
-that it is possible to recover the full conditional causality graph
-through pairwise testing alone -- that is, it isn't necessary to
-perform any conditional causality tests.
+The flavour of this work may be conveyed by stating one of our main
+conclusions, where the ideas behind $T_{ij}(z)$ are subtle and not
+expanded upon in this brief summary.
 
-% $\bullet$ Abstracts may contain equations: Equations are to be centered horizontally, with equation numbers aligned with the right margin. See equation (\ref{eq1})
-% \begin{equation}
-% e^{i\pi}+1=0.\label{eq1}
-% \end{equation}
+\begin{theorem}[Pairwise Recovery]
+  \label{thm:scg_recovery}
+  If the Granger causality graph $\gcg$ for the process $x(t)$ is
+  a strongly causal DAG and $T_{ij}(z)$ is a two-sided filter for every
+  $i, j \in [n]$, then $\gcg$ can be recovered from pairwise causality
+  tests alone.
+\end{theorem}
 
-% $\bullet$ Abstracts must contain at least two  references. See   \cite{ASY,STE}.
+% \begin{figure}[h]
+%   \centering
+%   \begin{subfigure}[b]{0.45\textwidth}
+%     \includegraphics[width=\linewidth]{../../figures/example_algorithm.pdf}
+%   \end{subfigure}
+%   \begin{subfigure}[b]{0.45\textwidth}
+%     \includegraphics[width=\linewidth]{../../figures/example_algorithm2.pdf}
+%   \end{subfigure}
+% \end{figure}
 
-% $\bullet$ Abstracts may contain figures. For example, Figure \ref{fig1}.
-
-\begin{figure}[H]
+\begin{figure}[h]
   \centering
-\framebox[6cm]{\rule{0cm}{0.5cm} Abstracts may contain figures}
-%\includegraphics[width =0.48\linewidth]{image} 
-\caption{Figure one caption.}
-\label{fig1}
+  \caption{Recovering Large Networks via Pairwise Causality Tests}
+  \label{fig:large_n_figure}
+  \includegraphics[width=0.7\linewidth]{../../figures/new_increasing_n_simulation.pdf}
 \end{figure}
 
- Please send your source file  and .pdf file to Alyssa Novelia (anovelia@uwaterloo.ca) or Isam Al-Darabsah (ialdarabsah@uwaterloo.ca)
+Our concept of a strongly causal graph topology initially arose
+artificially as a technical condition to understand the intuition of
+causality flow in $\gcg$ and to establish Theorem
+\ref{thm:scg_recovery}.  However, examples from the gene regulatory
+network literature suggest that the dynamics of some natural
+interacting systems may have an underlying interaction graph which is
+in fact strongly causal, or at least nearly so.  Simulation evidence
+also suggests that our algorithm for graph recovery via pairwise
+testing can outperform the widely employed LASSO (and variations
+thereon) for systems which satisfy (or nearly satisfy) the strongly
+causal condition; in the process we make the incidental observation
+that, all else equal, the LASSO exhibits superior performance on the
+strongly causal topology than on a general directed acyclic graph,
+suggesting that topological aspects of the underlying causality graph
+may be of interest as an aspect unique to time series data in
+comparison to ordinary regression.  Finally, the results of Figure
+\ref{fig:large_n_figure} show that when topological assumptions are
+met, it is possible to accurately identify extremely large networks
+($n > 1000$, implying more than $10^6$ possible edges), and that the
+underlying graph topology is an important determination of estimation
+accuracy in addition to the simple size and sparsity of the network.
+
+ % Please send your source file  and .pdf file to Alyssa Novelia (anovelia@uwaterloo.ca) or Isam Al-Darabsah (ialdarabsah@uwaterloo.ca)
 
@@
 \small
-\bibitem{ASY} K.T. Alligood, T.D. Sauer and J. A. Yorke. \textit{Chaos: An Introduction to Dynamical Systems}, New York, Springer-Verlag (1997).
+\bibitem{granger1969investigating} Granger,
+  C. W. J. (1969). \textit{Investigating causal relations by
+    econometric models and cross-spectral methods.} Econometrica 37
+  (3), 424–438.
 
-\bibitem{STE} L. Schimansky-Geier, A.V. Tolstopjatenko and W. Ebelin \textit{Noise induced transitions due to external additive noise}, Phys. Lett. A, \textbf{108}, 7, pp. 329-332 (1985).
+\bibitem{NBERw16223} Billio, M., M. Getmansky, A. W. Lo, and
+  L. Pelizzon (2010, July). \textit{Econometric measures of systemic
+    risk in the finance and insurance sectors.}  Working Paper 16223,
+  National Bureau of Economic Research.
+
+\bibitem{methods_for_inferring_gene_regulatory_networks_from_time_series_expression_data}
+  Nguyen, P. (2019). \textit{Methods for Inferring Gene Regulatory
+    Networks from Time Series Expression Data.}
+  Ph. D. thesis, Northwestern University
+
+\bibitem{barnett2009granger} Barnett, L., A. B. Barrett, and
+  A. K. Seth (2009). \textit{Granger causality and transfer entropy
+    are equivalent for gaussian variables.} Physical review letters
+  103(23), 238701.
+
+\bibitem{transfer_entropy_criticism} Ryan G James, Nix Barnett, and
+  James P Crutchfield. \textit{Information flows? A critique of
+    transfer entropies}. In: Physical review letters 116.23 (2016),
+  p. 238701
 \end{thebibliography}

### 2019-10-02 presentations/mcgillOct2019/pres.tex
Presentation almost done -- just fill in VAR(p) LASSO case.

@@
     \item{Does smoking cause cancer?}\pause
-    \item{Does government spending cause an increase in GDP?}\pause
+    \item{Does lowering interest rates cause lowering unemployment?}\pause
     \item{Granger causality attemps to address the last question's category.}
@@
     \item{Wold representation $x(t) = \sum_{\tau = 0}^\infty A(\tau)v(t - \tau)$}\pause
-    \item{$\text{VAR}(\infty)$ representation $x(t) = \sum_{\tau = 1}^\infty B(\tau)x(t - \tau) + v(t)$}\pause
+    \item{$\VAR(\infty)$ representation $x(t) = \sum_{\tau = 1}^\infty B(\tau)x(t - \tau) + v(t)$}\pause
     \item{Require $\E[v(t)v(t - \tau)] = \delta_{\tau}\Sigma_v$, with $\Sigma_v$ Diagonal}\pause
@@
     \textit{Granger-Causes} component $x_i$ (with respect to $x$)
-    and write $x_j \gc x_i$ if given Hilbert spaces $\H_{t - 1}$,
-    $\H^{-j}_{t - 1}$ if
+    and write $x_j \gc x_i$ if
 
@@
     We say that $x_j$ pairwise Granger-causes $x_i$ and write
-    $x_j \pwgc x_i$ if $x_j$ Granger-causes $x_i$ with respect only to
+    $x_j \pwgc x_i$ if $x_j$ Granger-causes $x_i$ with respect to
     $(x_i, x_j)$.
@@
 \begin{frame}{Simple Examples}
-  \begin{align*}
+  \begin{equation*}
+    \begin{aligned}
       x_0(t) &= b_0 x_0(t - 1) + b_{0, 2} x_2(t - 150) + v_0(t)\\
@@
       x_2(t) &= b_2 x_2(t - 1) + v_2(t)\\
-  \end{align*}
-
-  % TODO: Fix the whitespace consumed by this figure!
+    \end{aligned}
+  \end{equation*}
 
@@
     \item{Parents: $j \in \pa{i}$ iff $(j, i) \in \gcge$}\pause
-    \item{Ancestors: $j \in \anc{i}$ iff $j \rightarrow i$ path in $\gcge$}\pause
+    \item{Ancestors: $j \in \anc{i}$ iff $j \rightarrow i$ path in $\gcge$}
   \end{itemize}
@@
 
+\begin{frame}{Granger Causality Graph}
+  Structural results that tie in $\gcg$.  Recall:
+  \begin{itemize}
+    \item{$\mathsf{MA}(\infty)$ representation $x(t) = \sum_{\tau = 0}^\infty A(\tau)v(t - \tau)$}\pause
+    \item{$\mathsf{VAR}(\infty)$ representation $x(t) = \sum_{\tau = 1}^\infty B(\tau)x(t - \tau) + v(t)$}\pause
+  \end{itemize}
+
+  \begin{proposition}
+  $x_i(t)$ can be represented in terms of it's parents in $\gcg$:
+
+  \begin{equation}
+    \label{eqn:parent_expansion}
+    x_i(t) = v_i(t) + \B_{ii}(z)x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z)x_k(t).
+  \end{equation}\pause
+
+  Or, through it's ancestors:
+
+  \begin{equation}
+    \label{eqn:ancestor_expansion}
+    x_i(t) = \A_{ii}(z)v_i(t) + \sum_{\substack{k \in \anc{i} \\ k \ne i}}\A_{ik}(z)v_k(t).
+  \end{equation}
+
+  \end{proposition}
+\end{frame}
+
 \begin{frame}{Application Examples}
-  Analysis of connectedness of financial institutions
+  Analysis of connectedness of financial institutions: 2002 - 2004
 
-  TODO: Finance example
   \begin{columns}
+    \begin{column}{0.75\linewidth}
+      \includegraphics[width=\linewidth]{./figures/finance_connections2002.png}
+    \end{column}
+  \end{columns}
 
+  \footfullcite{NBERw16223}
+\end{frame}
+
+\begin{frame}{Application Examples}
+  Analysis of connectedness of financial institutions: 2006 - 2008
+
+  \begin{columns}
+    \begin{column}{0.75\linewidth}
+      \includegraphics[width=\linewidth]{./figures/finance_connections2006.png}
+    \end{column}
   \end{columns}
@@
 
-  TODO: Anna examples or something
+  \begin{columns}
+    \begin{column}{0.55\linewidth}
+      \includegraphics[width=\linewidth]{./figures/brain_photo.png}
+    \end{column}
+  \end{columns}
+
+  \footfullcite{anna_paper2008}
 \end{frame}
@@
 \begin{frame}{Application Examples}
-  TODO: Gene expression networks
+  Analysis of Gene regulatory networks
+
+  \begin{columns}
+    \begin{column}{0.5\linewidth}
+      \includegraphics[width=\linewidth]{../../figures/ecoli_regulatory_network.pdf}
+    \end{column}
+  \end{columns}
+
+  \footfullcite{discovering_graphical_Granger_causality_using_the_truncating_lasso_penalty}
 \end{frame}
@@
     \item{Large systems ($n \gtrapprox 20$) or when data is limited are addressed via the LASSO (and variants)}\pause
-    \item{Standard linear regression case:
+    \item{Standard linear regression case $y = X\beta^\star + v$:
         \begin{equation*}
@@
       }\pause
-    \item{$\VAR(p)$ models:
+    \item{$\VAR(p)$ models $x(t) = \sum_{\tau = 1}^p B^\star(\tau)x(t - \tau) + v(t)$:
         \begin{equation*}
@@
 
-\begin{frame}{LASSO Recovery Theorems}
-  TODO: Static Case
+\begin{frame}{LASSO Recovery Theorems (Linear Regression)}
+  \begin{itemize}
+    \item{$y = X\beta^\star + v$; $\beta^\star \in \R^n, y \in \R^T$}\pause
+    \item{$\widehat{\beta} \in \underset{\beta \in \R^n}{\text{argmin }} \frac{1}{2T}||y - X\beta|| + \lambda ||\widehat{\beta}||_1$}\pause
+  \end{itemize}
+
+  \begin{theorem}[Wainwright 2006]
+    Suppose $\beta^\star$ has only $s$ non-zero terms:
+    $||\beta^\star||_0 = s$, and $v \sim \mathcal{N}(0, \sigma^2)$.
+    \pause Then, with high probability, (ignoring some important
+    technicalities) there is a constant $C$ s.t. \pause
+
+    \begin{equation*}
+      ||\widehat{\beta} - \beta^\star||_2 \le C \sigma \sqrt{\frac{s \mathsf{log } n}{T}}.
+    \end{equation*}
+  \end{theorem}
+
+  \begin{itemize}
+    \item{Roughly, $||\beta^\star - \widehat{\beta}||_2$ is $O\big(\frac{s}{T}\big)$.}
+  \end{itemize}
 \end{frame}
 
-\begin{frame}{LASSO Recovery Theorems}
-  TODO: VAR(p) case
+\begin{frame}{LASSO Recovery Theorems ($\VAR(p)$ case)}
+  \begin{itemize}
+    \item{$\frac{1}{2T}||x(t) - \sum_{\tau = 1}^p B(\tau)x(t - \tau)||_2^2 + \lambda \sum_{\tau = 1}^p ||B(\tau)||_1$}\pause
+  \end{itemize}
+  TODO
 \end{frame}
@@
 \begin{frame}{Structure of Granger Causality Graphs}
-  \begin{itemize}
-    \item{Topology of Granger causality graph is a special feature of time series}\pause
-    \item{TODO: Star shaped graph with a self-loop versus a full cycle with graph -- same number of edges, but qualitatively totally different.}
-  \end{itemize}
+  \begin{columns}
+    \begin{column}{0.35\linewidth}
+      \includegraphics[width=\linewidth]{../../figures/example_rotation_shape.pdf}\pause
+    \end{column}
+    \begin{column}{0.35\linewidth}
+      \includegraphics[width=\linewidth]{../../figures/example_star_shaped.pdf}\pause
+    \end{column}
+  \end{columns}
+
+  \begin{columns}
+    \begin{column}{0.35\linewidth}
+      \includegraphics[width=\linewidth]{../../figures/example_loop_shaped.pdf}\pause
+    \end{column}
+    \begin{column}{0.35\linewidth}
+      \includegraphics[width=\linewidth]{../../figures/example_diamond_shaped.pdf}
+    \end{column}
+  \end{columns}
 \end{frame}
@@
     \begin{column}{0.33\linewidth}
-      TODO: Simple example\pause
+      \small{Simple Example}
+      \includegraphics[width=\linewidth]{../../figures/example_algorithm_step0.pdf}\pause
     \end{column}
@@
 \begin{frame}{Pairwise Recovery}
-  An Example (TODO: Intersperse Pauses)
-
   \begin{columns}
@@
       \begin{figure}
-        \includegraphics[width=\linewidth]{../../figures/example_algorithm2.pdf}
+        \includegraphics[width=\linewidth]{../../figures/example_algorithm_step0.pdf}
+      \end{figure}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Pairwise Recovery}
+  \begin{columns}
+    \begin{column}{0.5\linewidth}
+      \begin{figure}
+        \includegraphics[width=\linewidth]{../../figures/example_algorithm0_step1.pdf}
+      \end{figure}
+    \end{column}
+
+    \begin{column}{0.5\linewidth}
+      \begin{figure}
+        \includegraphics[width=\linewidth]{../../figures/example_algorithm_step0.pdf}
+      \end{figure}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Pairwise Recovery}
+  \begin{columns}
+    \begin{column}{0.5\linewidth}
+      \begin{figure}
+        \includegraphics[width=\linewidth]{../../figures/example_algorithm0_step1.pdf}
+      \end{figure}
+    \end{column}
+
+    \begin{column}{0.5\linewidth}
+      \begin{figure}
+        \includegraphics[width=\linewidth]{../../figures/example_algorithm_step1.pdf}
+      \end{figure}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Pairwise Recovery}
+  \begin{columns}
+    \begin{column}{0.5\linewidth}
+      \begin{figure}
+        \includegraphics[width=\linewidth]{../../figures/example_algorithm0_step1.pdf}
+      \end{figure}
+    \end{column}
+
+    \begin{column}{0.5\linewidth}
+      \begin{figure}
+        \includegraphics[width=\linewidth]{../../figures/example_algorithm_step2.pdf}
+      \end{figure}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Pairwise Recovery}
+  \begin{columns}
+    \begin{column}{0.5\linewidth}
+      \begin{figure}
+        \includegraphics[width=\linewidth]{../../figures/example_algorithm0_step1.pdf}
+      \end{figure}
+    \end{column}
+
+    \begin{column}{0.5\linewidth}
+      \begin{figure}
+        \includegraphics[width=\linewidth]{../../figures/example_algorithm_step3.pdf}
+      \end{figure}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Pairwise Recovery}
+  \begin{columns}
+    \begin{column}{0.5\linewidth}
+      \begin{figure}
+        \includegraphics[width=\linewidth]{../../figures/example_algorithm0_step1.pdf}
+      \end{figure}
+    \end{column}
+
+    \begin{column}{0.5\linewidth}
+      \begin{figure}
+        \includegraphics[width=\linewidth]{../../figures/example_algorithm_step4.pdf}
       \end{figure}

### 2019-10-13 presentations/student_dynamical_systems_waterloo_2019/pres.tex
Shortened talk for the CQAM workshop

--- 2019-10-02 presentations/mcgillOct2019/pres.tex
@@
 \def\H{\mathcal{H}}  % Hilbert space
+\def\Ht{\mathsf{H}}  % Hermitian Transpose
 \def\R{\mathbb{R}}  % Reals
@@
 \def\cl{\mathsf{cl}}  % Closure
+\def\ii{\mathsf{i}}  % Imaginary unit
 
@@
 
-\date{McGill University, Oct 4 2019}
+\date{Fields-CQAM Student Workshop on Dynamical Systems and Related Fields, University of Waterloo, Oct 2019}
 
@@
 
-\begin{frame}{Causality}
-  \begin{itemize}
-    \item{Do billiard collisions cause motion?}\pause
-    \item{Does standing in the street cause cars to stop?}\pause
-    \item{Does smoking cause cancer?}\pause
-    \item{Does lowering interest rates cause lowering unemployment?}\pause
-    \item{Granger causality attemps to address the last question's category.}
-  \end{itemize}
-\end{frame}
+% \begin{frame}{Causality}
+%   \begin{itemize}
+%     \item{Do billiard collisions cause motion?}\pause
+%     \item{Does standing in the street cause cars to stop?}\pause
+%     \item{Does smoking cause cancer?}\pause
+%     \item{Does lowering interest rates cause lowering unemployment?}\pause
+%     \item{Granger causality attemps to address the last question's category.}
+%   \end{itemize}
+% \end{frame}
 
-\begin{frame}{Causality in Time Series}
-  \begin{itemize}
-    \item{Consider $x_t, y_t \overset{\text{i.i.d.}}{\sim} \text{BER}(1/2)$}\pause
-    \item{If $z_t = x_{t - 1}$ then since cause preceeds effect, intuitively $x$ causes $z$}\pause
-    \item{$z'_t = x_{t - 1} \oplus y_{t - 1}$ is more ambiguous.}\pause
-    \item{Issue raised by: \fullcite{transfer_entropy_criticism}} 
-  \end{itemize}
-\end{frame}
+% \begin{frame}{Causality in Time Series}
+%   \begin{itemize}
+%     \item{Consider $x_t, y_t \overset{\text{i.i.d.}}{\sim} \text{BER}(1/2)$}\pause
+%     \item{If $z_t = x_{t - 1}$ then since cause preceeds effect, intuitively $x$ causes $z$}\pause
+%     \item{$z'_t = x_{t - 1} \oplus y_{t - 1}$ is more ambiguous.}\pause
+%     \item{Issue raised by: \fullcite{transfer_entropy_criticism}}\pause
+%     \item{This difficulty is absent for \textit{linear} systems -- the domain of Granger causality}
+%   \end{itemize}
+% \end{frame}
 
@@
 
-\begin{frame}{Basic Setup}
+\begin{frame}{Granger Causality}
   \begin{itemize}
     \item{Basic setup is an $n-$dimensional WSS process $x(t)$}\pause
-    \item{Wold representation $x(t) = \sum_{\tau = 0}^\infty A(\tau)v(t - \tau)$}\pause
+    \item{$\mathsf{MA}(\infty)$ (Wold) representation $x(t) = \sum_{\tau = 0}^\infty A(\tau)v(t - \tau)$}\pause
     \item{$\VAR(\infty)$ representation $x(t) = \sum_{\tau = 1}^\infty B(\tau)x(t - \tau) + v(t)$}\pause
@@
 
-  \begin{defn}
-    Linear Projection onto Hilbert space $\H \subseteq L_2(\Omega)$
-    \begin{equation*}
-      \linE{x}{\H} = \underset{u \in \H}{\text{argmin }} ||x - u||
-    \end{equation*}
-
-    Minimum Mean Squared Error
-    \begin{equation*}
-      \xi[x \ |\ \H] = \E (x - \linE{x}{\H})^2      
-    \end{equation*}
-  \end{defn}
-\end{frame}
-
-\begin{frame}{Basic Setup}
-  More notation
-
-  \begin{align*}
-    \H_t &= \cl \{\sum_{\tau = 0}^p a_\tau^\T x(t - \tau)\ |\ a_\tau \in \R^n, p \in \N\} \subseteq L_2(\Omega)\\
-    % H_t^x &= \{a x(t)\ |\ a \in \R\} \subseteq L_2(\Omega),
-  \end{align*}\pause
-
-  \begin{itemize}
-    \item{$\H_t^{i}$ Hilbert space spanned by $x_i(t)$ alone}
-    \item{$\H_t^{-i}$ Hilbert space spanned by $x(t)$ but \textit{excluding} $x_i(t)$}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Granger Causality}
   \begin{defn}
@@
     \begin{equation*}
-      \linEerr{x_i(t)}{\H_{t - 1}} < \linEerr{x_i(t)}{\H^{-j}_{t - 1}},
+      \exists \tau > 0 \text{ s.t. } B_{ij}(\tau) \ne 0.
     \end{equation*}
@@
 
-  \begin{defn}
-    We say that $x_j$ pairwise Granger-causes $x_i$ and write
-    $x_j \pwgc x_i$ if $x_j$ Granger-causes $x_i$ with respect to
-    $(x_i, x_j)$.
-  \end{defn}
+  % Need a VAR(2) definition of pairwise causality...
+  % \begin{defn}
+  %   We say that $x_j$ pairwise Granger-causes $x_i$ and write
+  %   $x_j \pwgc x_i$ if $x_j$ Granger-causes $x_i$ with respect to
+  %   $(x_i, x_j)$.
+  % \end{defn}
 \end{frame}
@@
   \begin{figure}
-    \includegraphics[width=0.75\linewidth]{./figures/var1_example.pdf}
+    \includegraphics[width=0.75\linewidth]{../figures/var1_example.pdf}
   \end{figure}
@@
   \centering
-  \includegraphics[width=0.75\linewidth]{./figures/var2_example.pdf}
+  \includegraphics[width=0.75\linewidth]{../figures/var2_example.pdf}
 \end{frame}
@@
 
-\begin{frame}{Application Examples}
-  Analysis of connectedness of financial institutions: 2002 - 2004
+% \begin{frame}{Application Examples}
+%   Analysis of connectedness of financial institutions: 2002 - 2004
 
-  \begin{columns}
-    \begin{column}{0.75\linewidth}
-      \includegraphics[width=\linewidth]{./figures/finance_connections2002.png}
-    \end{column}
-  \end{columns}
+%   \begin{columns}
+%     \begin{column}{0.75\linewidth}
+%       \includegraphics[width=\linewidth]{../figures/finance_connections2002.png}
+%     \end{column}
+%   \end{columns}
 
-  \footfullcite{NBERw16223}
-\end{frame}
+%   \footfullcite{NBERw16223}
+% \end{frame}
 
-\begin{frame}{Application Examples}
-  Analysis of connectedness of financial institutions: 2006 - 2008
+% \begin{frame}{Application Examples}
+%   Analysis of connectedness of financial institutions: 2006 - 2008
 
-  \begin{columns}
-    \begin{column}{0.75\linewidth}
-      \includegraphics[width=\linewidth]{./figures/finance_connections2006.png}
-    \end{column}
-  \end{columns}
+%   \begin{columns}
+%     \begin{column}{0.75\linewidth}
+%       \includegraphics[width=\linewidth]{../figures/finance_connections2006.png}
+%     \end{column}
+%   \end{columns}
 
-  \footfullcite{NBERw16223}
-\end{frame}
+%   \footfullcite{NBERw16223}
+% \end{frame}
 
-\begin{frame}{Application Examples}
-  Analysis of brain functional connectivity
+% \begin{frame}{Application Examples}
+%   Analysis of brain functional connectivity
 
-  \begin{columns}
-    \begin{column}{0.55\linewidth}
-      \includegraphics[width=\linewidth]{./figures/brain_photo.png}
-    \end{column}
-  \end{columns}
+%   \begin{columns}
+%     \begin{column}{0.55\linewidth}
+%       \includegraphics[width=\linewidth]{../figures/brain_photo.png}
+%     \end{column}
+%   \end{columns}
 
-  \footfullcite{anna_paper2008}
-\end{frame}
+%   \footfullcite{anna_paper2008}
+% \end{frame}
 
-\begin{frame}{Application Examples}
+\begin{frame}{Application Example}
   Analysis of Gene regulatory networks
@@
 
-\section{Recovery Methods}
-\begin{frame}{LASSO Recovery Methods}
+\section{Pairwise Causality and ``Causal Flow''}
+\begin{frame}{Pairwise Granger Causality}
   \begin{itemize}
-    \item{Small systems ($n = 2$ or $n \lessapprox 10$) with adequate data are amenable to least squares and hypothesis testing}\pause
-    \item{Large systems ($n \gtrapprox 20$) or when data is limited are addressed via the LASSO (and variants)}\pause
-    \item{Standard linear regression case $y = X\beta^\star + v$:
-        \begin{equation*}
-          \underset{\beta \in \R^n}{\text{minimize }} \frac{1}{2T}||y - X\beta||_2^2 + \lambda ||\beta||_1
-        \end{equation*}
-      }\pause
-    \item{$\VAR(p)$ models $x(t) = \sum_{\tau = 1}^p B^\star(\tau)x(t - \tau) + v(t)$:
-        \begin{equation*}
-          \underset{B(\tau) \in \R^{n \times n}}{\text{minimize }} \frac{1}{2T}||x(t) - \sum_{\tau = 1}^p B(\tau)x(t - \tau)||_2^2 + \lambda \sum_{\tau = 1}^p ||B(\tau)||_1
-        \end{equation*}
-      }
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{LASSO Recovery Theorems (Linear Regression)}
-  \begin{itemize}
-    \item{$y = X\beta^\star + v$; $\beta^\star \in \R^n, y \in \R^T$}\pause
-    \item{$\widehat{\beta} \in \underset{\beta \in \R^n}{\text{argmin }} \frac{1}{2T}||y - X\beta|| + \lambda ||\widehat{\beta}||_1$}\pause
-  \end{itemize}
-
-  \begin{theorem}[Wainwright 2006]
-    Suppose $\beta^\star$ has only $s$ non-zero terms:
-    $||\beta^\star||_0 = s$, and $v \sim \mathcal{N}(0, \sigma^2)$.
-    \pause Then, with high probability, (ignoring some important
-    technicalities) there is a constant $C$ s.t. \pause
-
-    \begin{equation*}
-      ||\widehat{\beta} - \beta^\star||_2 \le C \sigma \sqrt{\frac{s \mathsf{log } n}{T}}.
-    \end{equation*}
-  \end{theorem}
-
-  \begin{itemize}
-    \item{Roughly, $||\beta^\star - \widehat{\beta}||_2$ is $O\big(\frac{s}{T}\big)$.}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{LASSO Recovery Theorems ($\VAR(p)$ case)}
-  \begin{itemize}
-    \item{$\frac{1}{2T}||x(t) - \sum_{\tau = 1}^p B(\tau)x(t - \tau)||_2^2 + \lambda \sum_{\tau = 1}^p ||B(\tau)||_1$}\pause
-  \end{itemize}
-  TODO
-\end{frame}
-
-\begin{frame}{Structure of Granger Causality Graphs}
-  \begin{columns}
-    \begin{column}{0.35\linewidth}
-      \includegraphics[width=\linewidth]{../../figures/example_rotation_shape.pdf}\pause
-    \end{column}
-    \begin{column}{0.35\linewidth}
-      \includegraphics[width=\linewidth]{../../figures/example_star_shaped.pdf}\pause
-    \end{column}
-  \end{columns}
-
-  \begin{columns}
-    \begin{column}{0.35\linewidth}
-      \includegraphics[width=\linewidth]{../../figures/example_loop_shaped.pdf}\pause
-    \end{column}
-    \begin{column}{0.35\linewidth}
-      \includegraphics[width=\linewidth]{../../figures/example_diamond_shaped.pdf}
-    \end{column}
-  \end{columns}
-\end{frame}
-% \item{Confounder: A node $k$ confounds $i, j$ if $k \in \anc{i} \cap \anc{j}$ if exists a
-%     path $\gcgpath{k}{i}$ not containing $j$, and a path
-%     $\gcgpath{k}{j}$ not containing $i$.}
-
-\section{Pairwise Causality and ``Causal Flow''}
-\begin{frame}{Pairwise Testing}
-  \begin{itemize}
+    \item{Suppose we observe only $x_i, x_j$ (components of a latent $x$)}\pause
+    \item{Test \textit{pairwise} causality: $x_i \pwgc x_j$ (e.g. find
+        the best $2-$dimensional model and check for zeros)}\pause
     \item{Statistical tests for the case $n = 2$ are simple and well established}\pause
@@
     \item{Thresholding schemes suffer from confounding}\pause
-    \item{``Naive'' pairwise testing is so bad, few papers exist on the topic, or it is used as a simple baseline}\pause
     \item{\fullcite{tam2013gene_pwgc}}
@@
 \begin{frame}{Necessary Conditions for Pairwise Causality}
-  First, when $i, j$ are completely separated:
+  \begin{itemize}
+    \item{Interested in relationship between $x_j \gc x_i$ and $x_j \pwgc x_i$}\pause
+  \end{itemize}
 
-  \begin{proposition}
-    Consider distinct nodes $i, j$ in a Granger causality graph
-    $\gcg$.  If
-
-    \begin{enumerate}
-      \item{$j \not\in \anc{i}$ and $i \not\in \anc{j}$}\pause
-      \item{$\anc{i}\cap\anc{j} = \emptyset$}\pause
-    \end{enumerate}
-
-    then $\H_t^{(i)} \perp \H_t^{(j)}$: $\forall s, \tau \in \Z_+\ \E[x_i(t - s)x_j(t - \tau)] = 0$.
-    \pause  Moreover, this means that $j \npwgc i$ and $\linE{x_j(t)}{\H_t^{(i)}} = 0$.\pause
-  \end{proposition}
-
-  Note that $i, j$ may still share descendants.
-\end{frame}
-
-\begin{frame}{Necessary Conditions for Pairwise Causality}
-  $\H_t^{(i)} \perp \H_t^{(j)}$ is a much stronger conclusion that $j \npwgc i$:
   \begin{proposition}
@@
   \begin{itemize}
-    \item{Unfortunately, there is no general converse (i.e. $j \in \pa{i} \centernot\implies j \pwgc i$)}\pause
+    \item{Unfortunately, there is no general converse (e.g. recall $j \in \pa{i} \centernot\implies j \pwgc i$)}\pause
     \item{We seek special (graph-)\textit{topological} conditions when there is a converse}\pause
@@
   \begin{columns}
-    \begin{column}{0.33\linewidth}
+    \begin{column}{0.5\linewidth}
       \small{Simple Example}
@@
     \end{column}
-    \begin{column}{0.33\linewidth}
+    \begin{column}{0.5\linewidth}
       \small{Slightly larger example}
-      \includegraphics[width=\linewidth]{../../figures/example_scg.pdf}\pause
-    \end{column}
-    \begin{column}{0.33\linewidth}
-      \small{Similar number of edges, but not strongly causal}
-      \includegraphics[width=\linewidth]{../../figures/example_dag.pdf}
+      \includegraphics[width=\linewidth]{../../figures/example_scg.pdf}
     \end{column}
@@
   \begin{itemize}
-    \item{Unfortunately, effects of confounding more nuanced}\pause
+    \item{Unfortunately, effects of confounding more are nuanced}\pause
   \end{itemize}
@@
 
-\begin{frame}{Persistent Systems}
-  \begin{itemize}
-    \item{If $k \in \anc{i}\cap\anc{j}$ any of the following may occur:}\pause
-    \item{
-        $(i \npwgc j$, $j \npwgc i)$,
-        $(i \pwgc j$, $j \npwgc i)$,
-        $(i \npwgc j$, $j \pwgc i)$,
-        $(i \pwgc j$, $j \pwgc i)$}\pause
-    \item{Intuitively, $i$ may ``forget'' information from $k$ before the same information reaches $i$.}\pause
-  \end{itemize}
+% \begin{frame}{Persistent Systems}
+%   \begin{itemize}
+%     \item{If $k \in \anc{i}\cap\anc{j}$ any of the following may occur:}\pause
+%     \item{
+%         $(i \npwgc j$, $j \npwgc i)$,
+%         $(i \pwgc j$, $j \npwgc i)$,
+%         $(i \npwgc j$, $j \pwgc i)$,
+%         $(i \pwgc j$, $j \pwgc i)$}\pause
+%     \item{Intuitively, $i$ may ``forget'' information from $k$ before the same information reaches $i$.}\pause
+%   \end{itemize}
 
-  A heuristic ``definition'' (to avoid introducing more cumbersome notation):
-  \begin{definition}[Persistent]
-    We will say that the process $x(t)$ with Granger causality graph
-    $\gcg$ is \textit{persistent} if every node has a self-loop,
-    i.e. $\B_{ii}(z) \ne 0$.
-  \end{definition}
-\end{frame}
+%   \begin{definition}[Persistent]
+%     We will say that the process $x(t)$ with Granger causality graph
+%     $\gcg$ is \textit{persistent} if every node has a self-loop,
+%     i.e. $\B_{ii}(z) \ne 0$.
+%   \end{definition}
+% \end{frame}
 
-\begin{frame}{Persistent Systems}
-  Define $H_i(z)$ to be the filter s.t.:
+% \begin{frame}{Persistent Systems}
+%   Define $H_i(z)$ to be the filter s.t.:
 
-  \begin{equation*}
-    H_i(z)x_i(t) = \linE{x_i(t)}{\H_{t - 1}^{(i)}},
-  \end{equation*}\pause
+%   \begin{equation*}
+%     H_i(z)x_i(t) = \linE{x_i(t)}{\H_{t - 1}^{(i)}},
+%   \end{equation*}\pause
 
-  And then with $\sigma_k^2 = \E v_k(t)^2$
-  \begin{equation*}
-    \label{eqn:T_filter}
-    T_{ij}(z) = \sum_{k \in \anc{i} \cap \anc{j}}\sigma_k^2\A_{ik}(z^{-1})(1 - H_i(z^{-1}))(1 - H_j(z))\A_{jk}(z)
-  \end{equation*}\pause
+%   And then with $\sigma_k^2 = \E v_k(t)^2$
+%   \begin{equation*}
+%     \label{eqn:T_filter}
+%     T_{ij}(z) = \sum_{k \in \anc{i} \cap \anc{j}}\sigma_k^2\A_{ik}(z^{-1})(1 - H_i(z^{-1}))(1 - H_j(z))\A_{jk}(z)
+%   \end{equation*}\pause
 
-  \begin{proposition}
-    \label{prop:persistence_converse}
-    Fix $i, j \in [n]$ and suppose $\exists k \in \anc{i} \cap \anc{j}$
-    which confounds $i, j$.  Then, if $T_{ij}(z)$ is not causal we have
-    $j \pwgc i$, and if $T_{ij}(z)$ is not anti-causal we have
-    $i \pwgc j$.  Moreover, if $T_{ij}(z)$ is two-sided, then $j \pwgc i \iff i \pwgc j$.
-  \end{proposition}
-\end{frame}
+%   \begin{proposition}
+%     \label{prop:persistence_converse}
+%     Fix $i, j \in [n]$ and suppose $\exists k \in \anc{i} \cap \anc{j}$
+%     which confounds $i, j$.  Then, if $T_{ij}(z)$ is not causal we have
+%     $j \pwgc i$, and if $T_{ij}(z)$ is not anti-causal we have
+%     $i \pwgc j$.  Moreover, if $T_{ij}(z)$ is two-sided, then $j \pwgc i \iff i \pwgc j$.
+%   \end{proposition}
+% \end{frame}
 
@@
   \begin{itemize}
-    \item{Persistant systems where $T_{ij}(z)$ is not two-sided are highly pathological}\pause
+    \item{Additional weak assumptions on $x$ are necessary to avoid pathological cancellation}\pause
+    % \item{Persistant systems where $T_{ij}(z)$ is not two-sided are highly pathological}\pause
     \item{However, that the following requires an assumption on
@@
 
+\section{Simulations}
 \begin{frame}{Simulation}
@@
     \item{Construct random graphs and compare PWGC against LASSO method}\pause
-    \item{I obtained the best results with Zou's Adaptive LASSO, but known to suffer somewhat from false positives}
+    \item{I obtained the best competing results with Zou's Adaptive
+        LASSO -- but these methods are known to suffer somewhat from
+        false positives}
     \item{\fullcite{adaptive_lasso_zou2006}}
@@
       \end{figure}
-    \end{column}\pause
+    \end{column}
   \end{columns}
@@
     \begin{column}{0.5\linewidth}
-      \includegraphics[width=0.9\linewidth]{./figures/new_lasso_comparison_scg_pmax15_simulation_mcc_only.png}
+      \includegraphics[width=0.9\linewidth]{../figures/new_lasso_comparison_scg_pmax15_simulation_mcc_only.png}
     \end{column}
@@
     \begin{column}{0.5\linewidth}
-      \includegraphics[width=0.9\linewidth]{./figures/new_lasso_comparison_dag_pmax15_simulation_mcc_only.png}
+      \includegraphics[width=0.9\linewidth]{../figures/new_lasso_comparison_dag_pmax15_simulation_mcc_only.png}
     \end{column}
@@
 
+\section{Conclusion}
+\begin{frame}{Conclusion}
+  \begin{itemize}
+    \item{The topology of $\gcg$, all else equal, may effect LASSO recovery}\pause
+    \item{Granger causality suffers from some pathologies with ``causal flow''}\pause
+    \item{$\gcg$ can be recovered from pairwise tests if it is strongly causal}\pause
+    \item{Pairwise recovery algorithm performs surprisingly well, even if $\gcg$ is not strongly causal.}\pause
+    \item{QUESTIONS?}
+  \end{itemize}
+\end{frame}
+
 \end{document}

### 2019-10-13 presentations/student_dynamical_systems_waterloo_2019/rkinnear_abstract.tex
Shortened talk for the CQAM workshop

--- 2019-09-30 presentations/student_dynamical_systems_waterloo_2019/Abstract-Template.tex
@@
 
-% \usepackage{graphicx}
-% \usepackage{caption}
-\usepackage{subcaption}
 \newtheorem{theorem}{Theorem}
@@
 \cite{transfer_entropy_criticism} do not arise in the case of linear
-Granger causality, other pathologies are known.  Our work sheds light
-on some of the challenges that remain in the linear case, and in
-particular focuses on topological particularities of the underlying
-network of interactions.
+Granger causality, other pathologies are known and our work sheds
+light on some of the challenges that remain in the linear case.  In
+particular, we focus on topological particularities of the underlying
+network of interactions, asking: what class of network topologies are
+compatible with our intuitions about causation?
 
-In general, we are presented with a multitude of jointly wide sense
-stationary time series $x_1(t), \ldots, x_n(t)$ and seek to construct
-a (Granger-)causal graph $\gcg$ where there is an edge
-$(j, i) \in \gcg$ whenever $x_j$ Granger causes $x_i$ after
-conditioning on each $k \not\in \{i, j\}$.  Depending on the
-interpretation or on the application, the graph $\gcg$ captures the
-flows of ``information'' or ``energy'' throughout a network of
-interacting processes.  However, it may be the case that we observe
-$x_j$ and $x_i$ alone -- in such cases, simple examples establish that
-even if $x_j$ Granger causes $x_i$ conditioned on the rest of the
-observations, it need not be that $x_j$ Granger causes $x_i$ pairwise
-(i.e. without conditioning).
+In general, we are presented with jointly wide sense stationary
+processes $x_1(t), \ldots, x_n(t)$ and seek to construct a Granger
+causal graph $\gcg$ where there is an edge $(j, i) \in \gcg$ whenever
+$x_j$ Granger causes $x_i$ conditional on each $k \not\in \{i, j\}$.
+Depending on the interpretation or on the application, the graph
+$\gcg$ captures the flows of ``information'' or ``energy'' throughout
+a network of interacting processes.  However, it may be the case that
+we observe $x_j$ and $x_i$ alone -- in such cases, simple examples
+establish that even if $x_j$ Granger causes $x_i$ conditioned on the
+rest of the observations, it need not be that $x_j$ Granger causes
+$x_i$ pairwise (i.e. without conditioning).
 
-Pathologies of pairwise causation clash with our intuition that causal
-relations in $\gcg$ should be transitive, that is, that if there is
+Pathologies of pairwise causation clash with our intuition that, for
+example, causal relations in $\gcg$ should be transitive: if there is
 some ``true'' (conditional) causation path
@@
 
-% \begin{figure}[h]
-%   \centering
-%   \begin{subfigure}[b]{0.45\textwidth}
-%     \includegraphics[width=\linewidth]{../../figures/example_algorithm.pdf}
-%   \end{subfigure}
-%   \begin{subfigure}[b]{0.45\textwidth}
-%     \includegraphics[width=\linewidth]{../../figures/example_algorithm2.pdf}
-%   \end{subfigure}
-% \end{figure}
-
 \begin{figure}[h]
@@
   \label{fig:large_n_figure}
-  \includegraphics[width=0.7\linewidth]{../../figures/new_increasing_n_simulation.pdf}
+  \includegraphics[width=0.6\linewidth]{abstract_figure.pdf}
 \end{figure}
@@
 testing can outperform the widely employed LASSO (and variations
-thereon) for systems which satisfy (or nearly satisfy) the strongly
-causal condition; in the process we make the incidental observation
-that, all else equal, the LASSO exhibits superior performance on the
-strongly causal topology than on a general directed acyclic graph,
-suggesting that topological aspects of the underlying causality graph
-may be of interest as an aspect unique to time series data in
-comparison to ordinary regression.  Finally, the results of Figure
-\ref{fig:large_n_figure} show that when topological assumptions are
-met, it is possible to accurately identify extremely large networks
-($n > 1000$, implying more than $10^6$ possible edges), and that the
-underlying graph topology is an important determination of estimation
-accuracy in addition to the simple size and sparsity of the network.
+thereon) for systems which satisfy (or nearly satisfy) the strong
+causality condition; in the process we make the incidental observation
+that, all else equal, the LASSO exhibits marginally superior
+performance on the strongly causal topology than on a general directed
+acyclic graph, suggesting that topological aspects of the underlying
+causality graph may be of interest as an aspect unique to time series
+data in comparison to ordinary regression.  Finally, the results of
+Figure \ref{fig:large_n_figure} show that when topological assumptions
+are met, it is possible to accurately identify extremely large
+networks ($n > 1000$, implying more than $10^6$ possible edges), and
+that the underlying graph topology is an important determination of
+estimation accuracy in addition to the simple size and sparsity of the
+network.
 

### 2019-11-16 paper_arxiv/supplementary_material.tex
arxiv version.

--- 2019-08-27 paper_theory/supplementary_material.tex
@@
-\documentclass{statsoc}
-\usepackage{fullpage}
-\usepackage{ccfonts,eulervm}
-\usepackage{tabularx}
-\usepackage{xr}
-\usepackage{chngcntr}
-\usepackage{appendix}
-\newcounter{lemma}
-\newcounter{definition}
-\counterwithin{lemma}{section}
-\counterwithin{definition}{section}
-\counterwithin{equation}{section}
-\externaldocument{graph_topology_granger_causality}
-
-\usepackage{geometry}
-
-\geometry{
-  textwidth=33pc,
-  textheight=\dimexpr48\baselineskip+\topskip\relax,
-  marginparsep=11pt,
-  marginparwidth=107pt,
-  footnotesep=6.65pt,
-  headheight=9pt,
-  headsep=9pt,
-  footskip=30pt,
-}
-\bibliographystyle{chicago}
-\usepackage{enumitem}
-\usepackage{etex,etoolbox}
-\usepackage{hyperref}
-\usepackage{fullpage}
-\setlength{\marginparwidth}{2cm}
-\usepackage{todonotes}
-\usepackage{multirow}
-\usepackage{booktabs}
-
-\def\gc{\overset{\text{GC}}{\rightarrow}}  % Granger causality arrow
-\def\ngc{\overset{\text{GC}}{\nrightarrow}}  % Negated Granger causality arrow
-\def\pwgc{\overset{\text{PW}}{\rightarrow}}  % Pairwise Granger causality arrow
-\def\npwgc{\overset{\text{PW}}{\nrightarrow}}  % Negated pairwise Granger causality arrow
-\def\te{\overset{\mathcal{T}}{\rightarrow}}  % Transfer entropy arrow
-\def\gcg{\mathcal{G}}  % Granger causality graph
-\def\gcge{\mathcal{E}}  % Graph edges
-\def\VAR{\mathsf{VAR}}  % VAR(p) model
-\def\MA{\mathsf{MA}}  % MA(p) model
-\def\B{\mathsf{B}}  % Filter B
-\def\wtB{\widetilde{\B}}  % General filter B
-\def\A{\mathsf{A}}  % Filter A
-\def\H{\mathcal{H}}  % Hilbert space
-
-\newcommand{\linE}[2]{\hat{\E}[#1\ |\ #2]}  % Linear projection
-\newcommand{\linEerr}[2]{\xi[#1\ |\ #2]}  % Error of linear projection
-\newcommand{\pa}[1]{pa(#1)}  % Parents of a node
-\newcommand{\anc}[1]{\mathcal{A}(#1)}  % Ancestors of a node
-\newcommand{\ancn}[2]{\mathcal{A}_{#1}(#2)}  % nth ancestors of a node
-\newcommand{\gpn}[2]{gp_{#1}(#2)}  % nth generation grandparents
-\newcommand{\wtalpha}[2]{\widetilde{\alpha}(#1, #2)}  % Some notation for lem:pwgc_anc
-\newcommand{\dist}[2]{\mathsf{d}(#1, #2)}  % Distance between things
-\newcommand{\gcgpath}[2]{#1 \rightarrow \cdots \rightarrow #2}  % A shorter path command
-
-% \usepackage{fullpage}
-\usepackage{framed}
-
-% Figures
-\usepackage{graphicx}
-\usepackage{caption}  % This is not recommended?
-\usepackage{subcaption}
-% \usepackage{wrapfig}
-% \usepackage{svg}
-
-% Math packages, theorem definitions and numbering
-\usepackage{amsmath}
-\usepackage{amssymb}
-% \usepackage{amsthm}
-% \usepackage{mathrsfs} % Fancy scripted font
-\usepackage{bm}  % Bold math
-
-% Misc packages
-\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
-\usepackage{algorithm2e} %{algorithm} environment
-\usepackage{soul}  % \hl highlighting
-\usepackage{color}
-\usepackage{mathtools}  % For my \ceil function
-
-% Theorems (with italics)
-% \theoremstyle{plain}  % Style definition removes italics
-\newtheorem{theorem}{Theorem}
-\newtheorem{corollary}{Corollary}
-\newtheorem{proposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-
-% \theoremstyle{definition}
-\newtheorem{remark}{Remark}
-\newtheorem{definition}{Definition}
-\newtheorem{example}{Example}
-\newtheorem{assumptions}{Assumptions}
-
-% keywords
-\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}
-
-% General
-\def\defeq{\overset{\Delta}{=}}  % Equal with triangle
-\def\cl{\mathsf{cl\ }}  % Closure
-\newcommand{\sgn}[1]{\mathsf{sgn}(#1)}  % sign function
-
-% Calculus
-\def\d{\mathsf{d}}  % Differential operator
-
-% Functions
-\def\ln{\mathsf{ln\ }}  % Natural logarithm
-\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}  % Ceiling
-
-% Probability
-\def\H{\mathcal{H}}  % Hilbert space
-\def\E{\mathbb{E}}  % Expectation
-\def\Var{\text{Var}}  % Variance
-\def\P{\mathbb{P}}  % Probability Measure
-\def\F{\mathcal{F}}  % A sigma algebra
-\def\sX{\mathcal{X}}  % Another sigma algebra
-\def\KL{\mathbf{D}_{KL}}  % KL divergence
-\def\bF{\mathbf{F}}  % Whole F-meas space
-
-% Standard sets
-\def\Z{\mathbb{Z}}  % Set of integers
-\def\R{\mathbb{R}}  % Set of real numbers
-\def\C{\mathbb{C}}  % Set of complex numbers
-\def\N{\mathbb{N}}  % Set of natural numbers
-\def\ball{\mathbb{B}}  % Open ball
-\def\clball{\overline{\ball}}  % Closed ball
-
-% Linear algebra
-\def\rk{\mathsf{rk }}  % The rank
-\def\tr{\mathsf{tr }}  % The trace
-\def\T{\mathsf{T}}  % Transpose notation
-\def\c{\mathsf{c}}  % complement
-\def\dg{\mathsf{dg }}   %  Diagonal vector of a matrix
-\def\Dg{\mathsf{Dg }}   %  Diagonal matrix from a vector
-\def\ind{\mathbf{1}}  % Ones vector or indicator
-\def\matvec{\textbf{vec}}  % Vector operator
-\def\<{\langle}  % < Inner product
-\def\>{\rangle}  % > Inner product
-\newcommand{\inner}[2]{\langle #1, #2 \rangle}  % Inner product
-\newcommand{\innerT}[2]{#1^\T #2}  % Inner product for finite vectors
-
-\graphicspath{{../figures/}}
-
-\title{Graph Topological Aspects of Granger Causal Network Learning\\
-  \large Supplementary Material}
-
-\author[Author 1 {\it et al.}]{R. J. Kinnear}
-\address{
-  University of Waterloo,
-  Waterloo,
-  Canada.}
-\email{Ryan@Kinnear.ca}
-
-\author{R. R. Mazumdar}
-\address{
-  University of Waterloo,
-  Waterloo,
-  Canada.}
-
-\begin{document}
-
-\appendix
+% Appendix with proofs and more figures
 
@@
 \begin{proof}
-  % Geweke uses the log of the ratio of the determinants of the residual variances
-  % 
-  % The equivalence $(1) \iff (3)$ is essentially a restatement of the
-  % definition, but is in line with the seminal work of Geweke
-  % \cite{geweke1982measurement}, \cite{geweke1984}.
-
   $(a) \Rightarrow (b)$ follows as a result of the uniqueness of orthogonal
@@
 
-% I believe this is an incorrect proof since it relies on inverting
-% (1 - B_{jj}(z)), which need not be invertible.  For example,
-% B(z) = [[r, -a], [a, 0]]z^{-1} is invertible whenever
-% max(|r - root(r^2 - 4a^2)|, |r + root(r^2 - 4a^2)|) < 2
-% which includes values of r > 1.
-
-% \begin{proof}
-%   By Theorem \ref{thm:granger_causality_equivalences} it suffices to show that
-
-%   \begin{equation*}
-%     \forall \psi \in \H_{t - 1}^j\ \inner{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^i}}{\psi - \linE{\psi}{\H_{t - 1}^i}} = 0.
-%   \end{equation*}
-
-%   which by the orthogonality principle and by representing
-%   $\psi \in \H_{t - 1}^j$ via the action of some strictly causal filter
-%   $\Phi(z)$ on $x_j(t)$ is equivalent to
-
-%   \begin{equation}
-%     \label{eqn:proof_inner0}
-%     \inner{x_i(t)}{\Phi(z)x_j(t) - \linE{\Phi(z)x_j(t)}{\H_{t - 1}^i}} = 0.
-%   \end{equation}
-
-%   Given a node $u$, consider the partition of $\pa{u}$:
-
-%   \begin{align*}
-%     C_0(u) &= \{k \in \pa{u}\ |\ i \not\in \anc{k}, k \ne i \}\\
-%     C_1(u) &= \{k \in \pa{u}\ |\ i \in \anc{k} \text{ or } k = i \}.
-%   \end{align*}
-
-%   We can then expand the parents of $j$ using Equation \ref{eqn:parent_expansion} as
-
-%   \begin{align*}
-%     x_j(t) &= v_j(t) + \B_{jj}(z)x_j(t) + \sum_{k \in C_0(j)} B_{jk}(z) x_k(t) + \sum_{k \in C_1(j)} B_{jk}(z) x_k(t)\\
-%     &= \widetilde{A}_{jj}(z)\Big(v_j(t) + \sum_{k \in C_0(j)} B_{jk}(z) x_k(t) + \sum_{k \in C_1(j)} B_{jk}(z) x_k(t)\Big),
-%   \end{align*}
-
-%   % \todo{Is this true??}
-%   where $(1 - B_{jj}(z))^{-1} = \widetilde{A}_{jj}(z)$, which exists
-%   due to the invertibility of $\A(z)$. When this expression is
-%   substituted into the left hand side of Equation
-%   \ref{eqn:proof_inner0} it results in
-
-%   \begin{equation}
-%     \label{eqn:proof_inner1}
-%     \begin{aligned}
-%       \inner{x_i(t)}{\sum_{k \in C_1(j)} \big(\widetilde{A}_{jj}(z)\Phi(z)B_{jk}(z) x_k(t) &- \linE{\widetilde{A}_{jj}(z)\Phi(z)B_{jk}(z) x_k(t)}{\H_{t - 1}^i}\big)},
-%     \end{aligned}
-%   \end{equation}
-
-%   where (by Proposition \ref{prop:separated_ancestor_uncorrelated}) we
-%   have eliminated each of the terms from $C_0(u)$, and by Lemma
-%   \ref{lem:vj_perp} the terms involving $v_j(t)$.
-
-%   Now, for any $u \in \anc{j}$, we must have $u \not \in \anc{i}$, since
-%   otherwise $u \in \anc{i} \cap \anc{j}$, which is empty by hypothesis.
-%   Moreover, for any $u \in \anc{j}$, we have
-%   $\anc{i} \cap \anc{u} = \emptyset$, since
-%   $w \in \anc{i} \cap \anc{u} \implies w \in \anc{i} \cap \anc{j}$. Therefore,
-%   we can continue the above process recursively splitting each
-%   $k \in C_1(j)$ into $C_0(k)$ and $C_1(k)$) upwards along each path
-%   from $j$ to $i$, terminating whenever we encounter $i$.  This process
-%   is guaranteed to terminate since $i \in \anc{k}$.  This leaves us with
-%   some strictly causal filter $\Psi(z)$ such that Expression
-%   \ref{eqn:proof_inner1} is equivalent to
-
-%   \begin{equation}
-%     \inner{x_i(t)}{\Psi(z)x_i(t) - \linE{\Psi(z)x_i(t)}{\H_{t - 1}^i}},
-%   \end{equation}
-
-%   which is $0$ since $\Psi(z)x_i(t) \in \H_{t - 1}^i$ and
-%   therefore $\linE{\Psi(z)x_i(t)}{\H_{t - 1}^i} = \Psi(z)x_i(t)$.
-% \end{proof}
-
 \begin{remark}
@@
   by
-% {\substack{k \in \pa{a_0} \\ k \ne a_1}}
+
   \begin{equation}
@@
   and this is $0$ since $\Phi(z)F(z)x_i(t) \in \H_{t - 1}^{(i)}$.
-
-  % Using the above partition of $\pa{j}$, we can expand $x_j(t)$ in
-  % terms of it's parents, and recursively expand nodes in $C_1$
-  % (cancelling all nodes in $C_0$, and recalling that
-  % $C_2 = \emptyset$) until we reach $i$.
-
-  % Using this partition, we will expand $x_j(t)$ in terms of it's
-  % parents, and recursively expand nodes in $C_1$ until reaching $i$.
-  % For the first step equation \eqref{eqn:parent_expansion} gives us:
-
-  % \begin{equation}
-  %   \label{eqn:xj_partition_expansion}
-  %   x_j(t) = \widetilde{\A}_{jj}(z) \Big(v_j(t) + \sum_{k \in C_0(j)} \B_{ik}(z)x_k(t) + \sum_{k \in C_1(j)}\B_{ik}(z)x_k(t)\Big).
-  % \end{equation}
-
-  % Substituting (\ref{eqn:xj_partition_expansion}) into
-  % (\ref{eqn:sufficient_inner_prod}) we have cancellation of all terms
-  % in which $k \in C_0(j)$ by Proposition \ref{prop:ancestor_properties},
-  % and cancellation of the term involving $v_j(t)$ by Lemma \ref{lem:vj_perp}.
-
-  % We are therefore left with
-
-  % \begin{equation}
-  %   \label{eqn:refwornv}  % An arbitrary code for reference in just one place
-  %   \inner{x_i(t)}{\sum_{k \in C_1(j)} \big(\widetilde{\A}_{jj}(z)\Phi(z) \B_{ik}(z)x_k(t) - \linE{\widetilde{\A}_{jj}(z)\Phi(z) \B_{ik}(z)x_k(t)}{\H_{t - 1}^i}\big)}.
-  % \end{equation}
-
-  % Since each such $k$ inherits the needed properties $(a), (b)$ we can
-  % again (as in Proposition \ref{prop:ancestor_properties}) recursively
-  % expand this expression until we reach $i$, obtaining a strictly
-  % causal $\Psi(z)$ such that Expression \eqref{eqn:refwornv} is equal to
-
-  % \begin{equation}
-  %   \inner{x_i(t)}{\Psi(z)x_i(t) - \linE{\Psi(z)x_i(t)}{\H_{t - 1}^i}} = 0.
-  % \end{equation}
 \end{proof}
@@
   satisfied, then $j \pwgc i \iff i \pwgc j$.
-  % Suppose that Assumption \ref{ass:T_causality} is satisfied.  Then if
-  % there exists a $k$ which confounds $(i, j)$ we have
-  % $i \pwgc j \implies j \pwgc i$.  Moreover, if $T_{ij}(z)$ is not
-  % constant, then $i \pwgc j$.
 \end{proposition}
 
-% QUESTION: Does $\gcg$ need to be a strongly causal DAG for this proposition?
-
-% \begin{proposition}
-%   \label{prop:persistence_converse}
-%   Suppose $\gcg$ is a strongly causal DAG and that $x(t)$ is
-%   persistent, then if there exists a $k$ which confounds $(i, j)$ we
-%   have $i \pwgc j \implies j \pwgc i$.
-% \end{proposition}
 \begin{proof}
@@
   we cannot have $i \in \anc{j}$ or $j \in \anc{i}$.
-
-% , $(c)$ is obtained through cancellations from Lemma \ref{lem:vj_perp}, $(d)$ follows since $\H_i$ is self-adjoint, $(e)$ follows by definition of $\H_i$: $\exists H_i(z)$ such that $$H_i(z)x_i(t) = \H_i \sum_{\ell \in \anc{i}}\A_{i\ell}v_\ell(t),$$ $(f)$ follows by again expanding $x_i(t)$ with Equation \eqref{eqn:ancestor_expansion}, and finally $(g)$ by cancelling orthogonal components of $v(t)$.
 
@@
 \end{proof}
-
-% This proof is not correct -- there can be cancellation
-% \begin{proof}
-%   We will show that $j \pwgc i$, the other being symmetric.  First
-%   note that by the contrapositive of proposition
-%   \ref{prop:sc_graph_common_anc} we cannot have $i \in \anc{j}$ or
-%   $j \in \anc{i}$ (else $k$ would not be a confounder) and therefore
-%   \textit{every} $u \in \anc{i}\cap\anc{j}$ will be a confounder.
-
-%   It is sufficient to show that $\exists \psi \in \H_{t - 1}^{(j)}$
-%   such that
-
-%   \begin{equation*}
-%     \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} \ne 0.
-%   \end{equation*}
-
-%   To this end, let $F(z)$ be an arbitrary but strictly causal linear
-%   filter.  We apply equation \eqref{eqn:ancestor_expansion} to
-%   $x_i(t)$ and $\psi \defeq F(z)x_j(t)$:
-
-%   \begin{align*}
-%     &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}\\
-%     &\overset{(a)}{=} \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
-%     &\overset{(b)}{=} \inner{\sum_{k \in \anc{j}}\big(F(z)\A_{jk}(z)v_k(t) \\&- \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}\big)}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
-%     &\overset{(c)}{=} \sum_{k \in \anc{i}\cap\anc{j}}\Big(\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} \\&- \inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{ii}v_i(t)}\\
-%     &- \sum_{\ell \in \anc{i}}\inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{i\ell}(z)v_\ell(t)}\Big)
-%   \end{align*}
-
-%   where in $(a)$ we have removed the $\linE{x_i(t)}{\H_{t - 1}^{(i)}}$
-%   term via the orthogonality principle, in $(b)$ there is no
-%   $F(z)\A_{jj}(z)v_j(t)$ term since due to $j \not\in \anc{i}$ it is
-%   orthogonal to $\H_t^{(i)}$.  Finally, $(c)$ follows by applying
-%   orthogonality properties of $v(t)$, as well as the fact that
-%   $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$ for
-%   $k \not \in \anc{i}$.  Note that
-%   $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$
-%   and therefore there is in general no cancellation in the final term
-%   above for $\ell \in \anc{i}$.
-
-%   \todo{This is clearly not immediately evident}
-
-%   This is $0$ for every $F$ if and only if for all $F$ and
-%   $\forall k \in \anc{i} \cap \anc{j}$ we have
-%   $$\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$$ and
-%   $$\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} = 0,$$ which by
-%   Corollary \ref{cor:time_lag_cancellation} occurs if and only if
-%   $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$, which is impossible since by persistence
-%   $\tau_0(\A_{jk}) < \infty$ and  $\tau_\infty(\A_{ik}) = \infty$.
-% \end{proof}
 
@@
 
-  % \BlankLine
   \caption{Pairwise Granger Causality Algorithm (PWGC)}
   \label{alg:pwgr}
-  % \TitleOfAlgo{Pairwise Graph Recovery (PWGC)}
-  % \Input{Pairwise Granger causality relations between a persistent
-  % process of dimension $n$ whose joint Granger causality
-  % relations are known to form a strongly causal DAG $\gcg$.}
   \Input{Pairwise Granger causality relations}
@@
     $E_0 = \emptyset$  \texttt{\# edges of }$\gcg$\\
-    % $P_0 = \emptyset$  \texttt{\# layer by layer driving nodes}\\
+
     $k = 1$ \texttt{\# a counter used only for notation}}
@@
       $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times P_k\ |\ (i, j) \in W,\ \text{no } i \rightarrow j \text{ path in } Q\}$
-      % $E_k \leftarrow E_{k - 1} \cup D_{kr}$ \texttt{\# add edges to }$E_k$ \label{alg:inner_loop_end}\\
-      % $W_{k + 1} \leftarrow W_k \setminus D_{kr}$  \texttt{\# remove edges from consideration}\\
     }
     $E_k \leftarrow E_{k - 1} \cup \big(\bigcup_{r = 1}^k D_{kr}\big)$ \texttt{\# update } $E_k$ \texttt{ with new edges}\\
-    % \label{alg_line:inner_loop_end}\\
-    % $W_{k + 1} \leftarrow W_k \setminus \big(\bigcup_{r = 0}^k D_{kr}\big)$  \texttt{\# remove edges from consideration}\\
     $k \leftarrow k + 1$
@@
       $S \leftarrow S \setminus P_{k - 1}$ \texttt{\# remove processed nodes}\\
-      % $\mathcal{I}_k \leftarrow \big(\sum_{j \in S: (j, i) \in W_\delta} F_{ij}, \mathsf{\ for\ }i \in S \big)$ \texttt{\# intra-}$S$ \texttt{incident strength}\\
       $\mathcal{I}_k \leftarrow \big(\sum_{j\in S: (j, i) \in W_\delta} P_{ij}, \mathsf{\ for\ }i \in S \big)$\\
@@
 \clearpage
-\bibliography{paper_bib}
-\end{document}

### 2019-11-16 presentations/student_dynamical_systems_waterloo_2019/pres.tex
arxiv version.

@@
 % \end{frame}
+
+\begin{frame}{Qualitative Structure of Granger Causality Graphs}
+  \begin{columns}
+    \begin{column}{0.35\linewidth}
+      \includegraphics[width=\linewidth]{../../figures/example_rotation_shape.pdf}\pause
+    \end{column}
+    \begin{column}{0.35\linewidth}
+      \includegraphics[width=\linewidth]{../../figures/example_star_shaped.pdf}\pause
+    \end{column}
+  \end{columns}
+
+  \begin{columns}
+    \begin{column}{0.35\linewidth}
+      \includegraphics[width=\linewidth]{../../figures/example_loop_shaped.pdf}\pause
+    \end{column}
+    \begin{column}{0.35\linewidth}
+      \includegraphics[width=\linewidth]{../../figures/example_diamond_shaped.pdf}
+    \end{column}
+  \end{columns}
+\end{frame}
 
